### 2017-06-29 manuscript/manuscript.tex
Started writting manuscript with gji style (Geophysical Journal International)

\documentclass[extra]{gji}

\usepackage[utf8]{inputenc}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{todonotes} % to make annotations on margins

\usepackage{url}
\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
    allcolors=blue,
}

\begin{document}

\title[
    Variable Density Tesseroids]
    {Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities
}
\author
    {Santiado R. Soler$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
    $^1$CONICET, Argentina. e-mail: santiago.r.soler@gmail.com\\
    $^2$Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan, Argentina\\
    $^3$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil
    }

\maketitle

\begin{summary}
Summary of this paper 
\end{summary}

\begin{keywords}
tesseroids, variable density, Gravity, forward, python, fatiando
\end{keywords}

\section{Introduction}

\todo{escribir summary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}

We are indebted to the developers and maintainers of the open-source
software without which this work would not have been possible.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{gji}
\bibliography{references}

\end{document}

### 2017-07-05 manuscript/manuscript.tex
Started writing introduction

@@
 
-\todo{escribir summary}
+Variation of density of the lithosphere with depth has been studied for almost a century. 
+\citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
+And, in the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
+
+Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Welford2010}.
+
+The previous forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
+Now days, satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation in regional scales.
+In order to build forward models that reassemble the gravity anomalies for such scales, the curvature of the Earth must be taken into account, thus we must formulate them in spherical coordinates.
+A common approach, specially applied in the last years, is to discretize the Earth in spherical prisms, also known as tesseroids\todo{Insertar figura de tesseroides de leo}.
+
+The calculation of the gravity potential, gradients and tensor components for an arbitrary tesseroid on any point outside of it involves the resolution of elliptical integrals, so they must be approximated by numerical computations.
+\todo{Write Taylor and GLQ}
+
+\citet{Uieda2016} developed a forward gravity model based on tesseroids with constant densities and using a modified version of the adaptative discretization of \citet{Li2011}.
 
@@
 \bibliographystyle{gji}
-\bibliography{references}
+\bibliography{bibtex/references}
 

### 2017-07-07 manuscript/manuscript.tex
Continue writting introduction

--- 2017-06-29 manuscript/manuscript.tex
@@
 }
-\author
+\author[S.R. Soler, L. Uieda and M.E. Gimenez]
     {Santiado R. Soler$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
@@
 
-\todo{escribir summary}
+Lithosphere density variation with depth has been studied for almost a century. 
+\citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
+And, in the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
+Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Welford2010}.
+
+These forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
+Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation in regional scales.
+This raises the importance of building forward models that reassemble the gravity anomalies for such scales.
+The main issue that should be overcome is that the curvature of the Earth must be taken into account, thus the modelling must be done in spherical coordinates.
+%In order to build forward models that reassemble the gravity anomalies for such scales, the curvature of the Earth must be taken into account, thus we must formulate them in spherical coordinates.
+One way to achieve this is to discretize the Earth in spherical prisms known as tesseroids. \todo{Insertar figura de tesseroides de leo}
+Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume elliptic integrals \todo{citar} that are generally approximated by numerical computations.
+The literature offers two approaches: one involves Taylor series expansion \todo{citar} while the other makes use of Gauss-Legendre Quadrature (GLQ) \todo{citar}.
+The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.
+
+In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
+On the other hand, the GLQ allows us to calculate them for any density function.
+
+\citet{Uieda2016} developed a forward gravity model based on tesseroids with constant densities using the GLQ approximation method.
+The GLQ method becomes less accurate when the computation point is closer to the tesseroid \todo{citar} and the smaller GLQ degree. 
+Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it improves the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
+It consist in splitting the tesseroid into smaller ones if a distance-size ratio ($D$) is greater than a predefined value that has been objectively obtained by \citet{Uieda2016} comparing the analytical and the numerically approximated gravity fields generated by a spherical shell.
+The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
+(1) a faster calculation of the distance from the computation point to the tesseroid and 
+(2) the application of a stack based algorithm instead of recursive function calls that  speeds up the algorithm and gives more control over the recursion step.
+This forward gravity model has been originally implemented in command line programs written in C programming language.
+Although it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \todo{citar a fatiando} for geophysical modelling and inversion.
+
+We have enhanced the later implementation with a new forward gravity model that allows us to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
+It's written in Python and Cython languages, obtaining an easy to use library that runs precompiled code for the more time consuming functions.
+
+It also makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
+Furthermore, we have the intention to include this new method into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
 

### 2017-07-07 manuscript/manuscript.tex
Modifications to introduction

@@
 \documentclass[extra]{gji}
+%\documentclass[extra, referee]{gji}
 
@@
     Variable Density Tesseroids]
-    {Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities
+    {Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth
 }
@@
 \begin{keywords}
-tesseroids, variable density, Gravity, forward, python, fatiando
+up to six keywords from the list
 \end{keywords}
@@
 The main issue that should be overcome is that the curvature of the Earth must be taken into account, thus the modelling must be done in spherical coordinates.
-%In order to build forward models that reassemble the gravity anomalies for such scales, the curvature of the Earth must be taken into account, thus we must formulate them in spherical coordinates.
 One way to achieve this is to discretize the Earth in spherical prisms known as tesseroids. \todo{Insertar figura de tesseroides de leo}
@@
 The GLQ method becomes less accurate when the computation point is closer to the tesseroid \todo{citar} and the smaller GLQ degree. 
-Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it improves the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
-It consist in splitting the tesseroid into smaller ones if a distance-size ratio ($D$) is greater than a predefined value that has been objectively obtained by \citet{Uieda2016} comparing the analytical and the numerically approximated gravity fields generated by a spherical shell.
+Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
+It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.
+% that has been objectively obtained by \citet{Uieda2016} comparing the analytical and the numerically approximated gravity fields generated by a spherical shell.
 The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
@@
 This forward gravity model has been originally implemented in command line programs written in C programming language.
-Although it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \todo{citar a fatiando} for geophysical modelling and inversion.
+Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \todo{citar a fatiando} for geophysical modelling and inversion.
 
@@
 Furthermore, we have the intention to include this new method into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
+
+In order to ensure the accuracy of the numerical approximation, we have compared its results with a spherical shell with linear and exponential density functions, similar to the test performed by \cite{Uieda2016}.
+Although in this case, the analytical expressions for both density functions had to be derived.
+
+In the following sections we will describe how the new algorithm works, its theoretical background and obtain the distance-size ratio values needed to get a good accuracy in the computation.
 

### 2017-07-07 manuscript/manuscript.tex
Modifications to introduction

@@
 And, in the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
-Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Welford2010}.
+Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
 
@@
 Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume elliptic integrals \todo{citar} that are generally approximated by numerical computations.
-The literature offers two approaches: one involves Taylor series expansion \todo{citar} while the other makes use of Gauss-Legendre Quadrature (GLQ) \todo{citar}.
+The literature offers two approaches: one involves Taylor series expansion \cite{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \cite{Asgharzadeh2007, Uieda2016, Uieda2017}.
 The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.
@@
 
-\citet{Uieda2016} developed a forward gravity model based on tesseroids with constant densities using the GLQ approximation method.
-The GLQ method becomes less accurate when the computation point is closer to the tesseroid \todo{citar} and the smaller GLQ degree. 
+\citet{Uieda2016} developed a forward gravity model based on tesseroids with homogeneous densities using the GLQ approximation method.
+The GLQ method becomes less accurate when the computation point is closer to the tesseroid \cite{Ku1977} or for smaller GLQ degree. 
 Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
 It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.
-% that has been objectively obtained by \citet{Uieda2016} comparing the analytical and the numerically approximated gravity fields generated by a spherical shell.
+\citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
 The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
 (1) a faster calculation of the distance from the computation point to the tesseroid and 
-(2) the application of a stack based algorithm instead of recursive function calls that  speeds up the algorithm and gives more control over the recursion step.
+(2) the application of a stack based algorithm that  speeds up the computation and gives more control over the recursion step.
+
 This forward gravity model has been originally implemented in command line programs written in C programming language.
-Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \todo{citar a fatiando} for geophysical modelling and inversion.
+Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.
 
@@
 In the following sections we will describe how the new algorithm works, its theoretical background and obtain the distance-size ratio values needed to get a good accuracy in the computation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Theory}
 

### 2017-07-11 manuscript/manuscript.tex
Minor changes in introduction

@@
 This raises the importance of building forward models that reassemble the gravity anomalies for such scales.
-The main issue that should be overcome is that the curvature of the Earth must be taken into account, thus the modelling must be done in spherical coordinates.
-One way to achieve this is to discretize the Earth in spherical prisms known as tesseroids. \todo{Insertar figura de tesseroides de leo}
-Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume elliptic integrals \todo{citar} that are generally approximated by numerical computations.
-The literature offers two approaches: one involves Taylor series expansion \cite{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \cite{Asgharzadeh2007, Uieda2016, Uieda2017}.
+
+The main issue that should be overcome is taking into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
+One way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries \todo{Insertar figura de tesseroides de leo}.
+Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
+The literature offers two approaches: one involves Taylor series expansion \citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
 The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.
@@
 \citet{Uieda2016} developed a forward gravity model based on tesseroids with homogeneous densities using the GLQ approximation method.
-The GLQ method becomes less accurate when the computation point is closer to the tesseroid \cite{Ku1977} or for smaller GLQ degree. 
+The GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ degree. 
 Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.

### 2017-07-11 manuscript/manuscript.tex
Minor corrections on introduction

@@
 \citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
-And, in the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
+In the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
 Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
@@
 The main issue that should be overcome is taking into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
-One way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries \todo{Insertar figura de tesseroides de leo}.
+A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries \todo{Insertar figura de tesseroides de leo}.
 Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
@@
 In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
-On the other hand, the GLQ allows us to calculate them for any density function.
+On the other hand, the GLQ allows us to calculate them for any density function without any other information but the function itself.
 
@@
 
+In order to ensure the accuracy of the numerical approximation, we have compared its results with a spherical shell with linear and exponential density functions, similar to the test performed by \cite{Uieda2016}.
+Although in this case, the analytical expressions for both density functions had to be derived.
+
 It also makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
 Furthermore, we have the intention to include this new method into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
-
-In order to ensure the accuracy of the numerical approximation, we have compared its results with a spherical shell with linear and exponential density functions, similar to the test performed by \cite{Uieda2016}.
-Although in this case, the analytical expressions for both density functions had to be derived.
 

### 2017-07-11 manuscript/manuscript.tex
Minor changes on introduction

@@
 \citet{Uieda2016} developed a forward gravity model based on tesseroids with homogeneous densities using the GLQ approximation method.
-The GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ degree. 
+It has been originally implemented in command line programs written in C programming language.
+Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.
+
+It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ degree. 
 Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
@@
 (2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
-
-This forward gravity model has been originally implemented in command line programs written in C programming language.
-Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.
 
@@
 It also makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
-Furthermore, we have the intention to include this new method into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
+Furthermore, we have the intention to include it into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
 

### 2017-07-11 manuscript/manuscript.tex
Started writing Theory

@@
 
-It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ degree. 
-Although the model developed by \citet{Uieda2016} uses a second GLQ degree, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
+It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ order. 
+Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
 It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.
@@
 
+The spherical prisms known as tesseroids are mass elements defined in a geocentric spherical coordinate system bounded by a pair of parallels, a pair of  meridians, and two concentric spherical surfaces.
+We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented cartesian coordinate system located at $P$.
+In the special case of homogeneous density, the gravity potential, gradient and Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[see also][]{Uieda2016}.
+
+In our case, we will suppose that the tesseroid has an arbitrary variable density in depth, this means that only depends on the radius spherical coordinate.
+Thus, the integrals for the gravity fields are slightly modified.
+
+\begin{equation}
+    V(r,\phi,\lambda) = G
+    \int\limits_{\lambda_1}^{\lambda_2}
+    \int\limits_{\phi_1}^{\phi_2}
+    \int\limits_{r_1}^{r_2}
+    \frac{\rho(r')}{\ell} \kappa \,  dr' d\phi' d\lambda',
+\label{eq:tesspot}
+\end{equation}
+\begin{equation}
+    g_{\alpha}(r,\phi,\lambda) = G
+    \int\limits_{\lambda_1}^{\lambda_2}
+    \int\limits_{\phi_1}^{\phi_2}
+    \int\limits_{r_1}^{r_2}
+    \rho(r') \frac{\Delta_\alpha}{\ell^3}
+    \kappa \, dr' d\phi' d\lambda',
+\label{eq:tessgrav}
+\end{equation}
+\begin{equation}
+    g_{\alpha\beta}(r,\phi,\lambda) = G
+    \int\limits_{\lambda_1}^{\lambda_2}
+    \int\limits_{\phi_1}^{\phi_2}
+    \int\limits_{r_1}^{r_2}
+    \rho(r') I_{\alpha\beta} \, \kappa \, dr' d\phi' d\lambda' ,
+    \label{eq:tesstensor}
+\end{equation}
+\noindent where
+\begin{equation}
+    I_{\alpha\beta} =
+    \left(
+        \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
+        \frac{\delta_{\alpha\beta}}{\ell^3}
+    \right) ,
+    \label{eq:tesstensorkernel}
+\end{equation}
+
+\noindent $\alpha, \beta \in \{x, y, z\}$,
+$\rho(r')$ is the density function that depends on the radius coordinate,
+$\delta_{\alpha\beta}$ is Kronecker's delta,
+$G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant and
+\begin{equation}
+    \Delta_x = r'[\cos\phi\sin\phi' - \sin\phi\cos\phi'
+               \cos(\lambda' - \lambda)],
+\end{equation}
+\begin{equation}
+    \Delta_y = r' \cos \phi' \sin(\lambda' - \lambda),
+\end{equation}
+\begin{equation}
+    \Delta_z = r' \cos \psi - r,
+\end{equation}
+\begin{equation}
+    \kappa = {r'}^2 \cos \phi',
+\end{equation}
+\begin{equation}
+    \ell = \sqrt{{r'}^2 + r^2 - 2 r' r \cos \psi},
+\end{equation}
+\begin{equation}
+    \cos\psi = \sin\phi\sin\phi' + \cos\phi\cos\phi'
+                 \cos(\lambda' - \lambda).
+\end{equation}
+
+According to \citet[p.390]{Hildebrand1987}, we can approximate any integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e.
+by a weighted sum of the integration kernel evaluated on the roots of the $N$th Legendre polynomial.
+
+\begin{equation}
+    \int\limits_{-1}^1 f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i),
+\end{equation}
+
+\noindent where $x_i$ are the roots of the $N$th Legendre polynomial $P_N(x)$ and $W_i$ are the weights of each term.
+
+\begin{equation}
+    W_i = \frac{2}{(1-x_i^2)[P_N^\prime(x_i)]^2},
+\end{equation}
+
+\begin{equation}
+    P_N(x_i) = 0, \quad \forall i = {1,\dots,N}.
+\end{equation}
+
+In case our integration is defined on a different interval, e.g. $[a,b]$, we can scale the nodes and perform a similar approximation.
+
+\begin{equation}
+    \int\limits_a^b f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i^s),
+\end{equation}
+
+\noindent where $x_i^s$ is the scaled $x_i$ root in the $[a,b]$ interval, also called nodes of the quadrature:
+
+\begin{equation}
+    x_i^s = \frac{b-a}{2} x_i + \frac{b+a}{2}.
+\end{equation}
+
+We are going to use only a second order GLQ, so the roots of the $P_2(x)$ Legendre polynomial are $x_i = \pm 0.577350269$ and the weights are $W_i = 1$.
+
+Our intention is to use GLQ to approximate the gravity fields volume integrals from equations \ref{eq:tesspot}-\ref{eq:tesstensor}.
+Applying the GLQ definition for every integral corresponding to each spherical coordinate we obtain the following expression \citep{Asgharzadeh2007, Uieda2016}:
+
+\begin{equation}
+    \iiint\limits_\Omega f(r', \phi', \lambda') d\Omega \approx A 
+    \sum\limits_{i=1}^{N^r}
+    \sum\limits_{j=1}^{N^\phi}
+    \sum\limits_{k=1}^{N^\lambda}
+    W_i^r W_j^\phi W_k^\lambda f(r_i, \phi_j, \lambda_k),
+\end{equation}
+
+\noindent where
+
+\begin{equation}
+    A = 
+    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8}.
+\end{equation}
+
+In our case with variable density, we can write the integral kernel as:
+
+\begin{equation}
+    f(r', \phi', \lambda') = \rho(r') g(r', \phi', \lambda')
+\end{equation}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-07-12 manuscript/manuscript.tex
Continue writing Theory

@@
 \documentclass[extra]{gji}
-%\documentclass[extra, referee]{gji}
+%~ %\documentclass[extra, referee]{gji}
 
@@
 \usepackage{amsmath}
-\usepackage{amssymb}
 \usepackage{graphicx}
@@
 
-\title[
-    Variable Density Tesseroids]
-    {Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth
+\title[Variable Density Tesseroids]{
+    Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth
 }
-\author[S.R. Soler, L. Uieda and M.E. Gimenez]
-    {Santiado R. Soler$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
+\author[S.R. Soler, L. Uieda and M.E. Gimenez]{
+    Santiado R. Soler$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
     $^1$CONICET, Argentina. e-mail: santiago.r.soler@gmail.com\\
@@
 \citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
-In the following years, other density functions have been proposed for different rock types \citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
+In the following years, other density functions have been proposed for different rock types \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
 Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
@@
     \frac{\rho(r')}{\ell} \kappa \,  dr' d\phi' d\lambda',
-\label{eq:tesspot}
+\label{eq:tesseroid-pot}
 \end{equation}
@@
     \kappa \, dr' d\phi' d\lambda',
-\label{eq:tessgrav}
+\label{eq:tesseroid-grav}
 \end{equation}
@@
     \rho(r') I_{\alpha\beta} \, \kappa \, dr' d\phi' d\lambda' ,
-    \label{eq:tesstensor}
+    \label{eq:tesseroid-tensor}
 \end{equation}
@@
     \right) ,
-    \label{eq:tesstensorkernel}
+    \label{eq:tesseroid-tensor-kernel}
 \end{equation}
@@
 
-According to \citet[p.390]{Hildebrand1987}, we can approximate any integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e.
-by a weighted sum of the integration kernel evaluated on the roots of the $N$th Legendre polynomial.
+According to \citet[p.~390]{Hildebrand1987}, we can approximate any integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e.
+by a weighted sum of the integration kernel evaluated on the roots of the $N$th Legendre polynomial:
 
@@
 
-\noindent where $x_i$ are the roots of the $N$th Legendre polynomial $P_N(x)$ and $W_i$ are the weights of each term.
+\noindent where $x_i$ are the roots of the $N$th Legendre polynomial $P_N(x)$ and $W_i$ are the weights of each term:
 
@@
 
-In case our integration is defined on a different interval, e.g. $[a,b]$, we can scale the nodes and perform a similar approximation.
+In case of an integration defined on a arbitrary interval, e.g. $[a,b]$, we can scale the nodes and perform a similar approximation:
 
@@
 
-We are going to use only a second order GLQ, so the roots of the $P_2(x)$ Legendre polynomial are $x_i = \pm 0.577350269$ and the weights are $W_i = 1$.
+For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$, respectivetly.
 
-Our intention is to use GLQ to approximate the gravity fields volume integrals from equations \ref{eq:tesspot}-\ref{eq:tesstensor}.
-Applying the GLQ definition for every integral corresponding to each spherical coordinate we obtain the following expression \citep{Asgharzadeh2007, Uieda2016}:
+Our intention is to use GLQ to approximate the volume integrals from equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor}. 
+When tesseroids have variable densities, we can write the integral kernels as the product of a certain function $g$ and the density function:
 
 \begin{equation}
-    \iiint\limits_\Omega f(r', \phi', \lambda') d\Omega \approx A 
+    f(r', \phi', \lambda') = \rho(r') g(r', \phi', \lambda'),
+\end{equation}
+
+\noindent and then apply the quadrature to every integral corresponding to each spherical coordinate, obtaining a similar expression to the one shown by \citet{Asgharzadeh2007} and \citet{Uieda2016}:
+
+\iftwocol{
+\begin{equation}
+\begin{split}
+    \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') d\Omega \approx& \\
+    A 
     \sum\limits_{i=1}^{N^r}
@@
     \sum\limits_{k=1}^{N^\lambda}
-    W_i^r W_j^\phi W_k^\lambda f(r_i, \phi_j, \lambda_k),
+    & W_i^r W_j^\phi W_k^\lambda \rho(r_i) g(r_i, \phi_j, \lambda_k),
+\end{split}
+\label{eq:glq-var-dens}
 \end{equation}
+}
+{
+\begin{equation}
+    \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') d\Omega \approx
+    A 
+    \sum\limits_{i=1}^{N^r}
+    \sum\limits_{j=1}^{N^\phi}
+    \sum\limits_{k=1}^{N^\lambda}
+    W_i^r W_j^\phi W_k^\lambda \rho(r_i) g(r_i, \phi_j, \lambda_k),
+\label{eq:glq-var-dens}
+\end{equation}
+}
 
@@
 
-In our case with variable density, we can write the integral kernel as:
-
-\begin{equation}
-    f(r', \phi', \lambda') = \rho(r') g(r', \phi', \lambda')
-\end{equation}
+From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates the gravity fields of a tesseroid with variable density as the ones generated by point masses located at the nodes of the Legendre polynomials.
+Furthermore, the information of the density function is summarised as the values it assumes in those nodes.
+Although this may sound discouraging, we intent to prove otherwise and show that this method, along with a well fitted discretization algorithm, approximates the gravity fields with good accuracy and precision.\todo{no se si poner esta ultima oracion}
 

### 2017-07-12 manuscript/manuscript.tex
Started writing Theory -> Analytical Solutions

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Analytical Solutions for Spherical Shell}
+
+In order to test the accuracy of the GLQ approximation, \citet{Uieda2016} compared the computed gravity fields with the ones generated by a spherical shell, the exceptional case with analytical solution \citep{Misuska2006,}.
+
+In case of variable density tesseroids, the comparison must obviously be made with a variable density spherical shell, whose analytical solution will depend on the chosen density function.
+Moreover, the literature does not develop the analytical solution
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-07-12 manuscript/manuscript.tex
Continue writing analytical solutions

--- 2017-07-12 manuscript/manuscript.tex
@@
 \documentclass[extra]{gji}
-%~ %\documentclass[extra, referee]{gji}
+%~ \documentclass[extra, referee]{gji}
 
@@
 \end{keywords}
+
+\todo{aclarar coordenadas esfericas -> longitud latitud angulos}
 
@@
     \ell = \sqrt{{r'}^2 + r^2 - 2 r' r \cos \psi},
+\label{eq:ell}
 \end{equation}
@@
                  \cos(\lambda' - \lambda).
+\label{eq:cospsi}
 \end{equation}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Analytical Solutions for Spherical Shell}
+
+In order to test the accuracy of the GLQ approximation, \citet{Uieda2016} and \citet{Grombein2013} have compared the computed gravity fields with the ones generated by a spherical shell, the exceptional case with analytical solution \citep{Mikuska2006,Grombein2013}.
+
+In case of variable density tesseroids, the comparison must obviously be made with a variable density spherical shell, whose analytical solution will depend on the chosen density function.
+Because the literature has not offered it yet, we needed to develop ourselves.
+
+Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, which density depends on the radius spherical coordinate.
+The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$ located along the $z$ axis at a distance $r$ can be written as follows:
+
+\todo{figura spherical shell con punto Q}
+
+\begin{equation}
+    V_\text{sh}(r) = G 
+    \int\limits_0^{2\pi}
+    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
+    \int\limits_{R_1}^{R_2}
+    \frac{\rho(r')}{\ell} {r'}^2 \cos\phi' \, 
+    dr' d\phi' d\lambda',
+\end{equation}
+
+\noindent where $\ell$ is defined as in equation \ref{eq:ell}.
+The computation point $Q$ is located at latitude $\phi=90^\circ$, so $\ell$ and $\cos\psi$ (defined in equation \ref{eq:cospsi}) get simplified:
+
+\begin{equation}
+    \cos\psi = \sin\phi', \quad
+    \ell = \sqrt{r'^2 + 2 r' r \sin\phi' + r^2}.
+\end{equation}
+
+Due to the rotational symmetry along the $z$ axis, the integration in $\lambda'$ is straightforward:
+
+\begin{equation}
+    V_\text{sh}(r) = 2\pi G 
+    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
+    \int\limits_{R_1}^{R_2}
+    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + 2 r' r \sin\phi' + r^2}}
+    \, dr' d\phi',
+\end{equation}
+
+\noindent and the integration in $\phi'$ can be performed independently of the density function.
+Using \citet{sagemath} we obtain the following expression:
+
+\iftwocol{
+\begin{equation}
+\begin{split}
+    V_\text{sh}(r) = 2\pi G
+    \int\limits_{R_1}^{R_2}
+    \Big[ \sqrt{r^2 + r'^2 + 2rr'} -& \\
+    \sqrt{r^2 + r'^2 - 2rr'}
+    & \Big] \frac{r'\rho(r')}{r} \, dr'.
+\end{split}
+\label{eq:shell-pot-sqrts}
+\end{equation}
+}
+{
+\begin{equation}
+    V_\text{sh}(r) = 2\pi G
+    \int\limits_{R_1}^{R_2}
+    \Big[ \sqrt{r^2 + r'^2 + 2rr'}  -
+    \sqrt{r^2 + r'^2 - 2rr'} 
+    \Big] \frac{r'\rho(r')}{r} \, dr'.
+\label{eq:shell-pot-sqrts}
+\end{equation}
+}
+
+For our purposes we can assume that the computation point $Q$ is outside the outer radii, i.e. $r>R_2$, and simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
+
+\begin{equation}
+    \sqrt{r^2 + r'^2 + 2rr'} = |r + r'| = r + r',
+\end{equation}
+\begin{equation}
+    \sqrt{r^2 + r'^2 - 2rr'} = |r - r'| = r - r',
+\end{equation}
+
+\noindent what leads to the following expression of the potential of the spherical shell:
+
+\begin{equation}
+    V_\text{sh}(r) = \frac{4\pi G}{r}
+    \int\limits_{R_1}^{R_2} {r'}^2 \rho(r') \, dr'.
+\label{eq:shell-pot}
+\end{equation}
+
+The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity field generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axis.
+Due to the rotational symmetry along any axe that passes through the centre of the shell, the potential in equation \ref{eq:shell-pot} reassembles the gravity field on any outside point at distance $r$ from the its centre.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-07-13 manuscript/manuscript.tex
Continue writing analytical solutions

@@
 \end{equation}
+
 \noindent where
+
 \begin{equation}
@@
 $G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant and
+
 \begin{equation}
@@
     \begin{split}
-    \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') d\Omega \approx& \\
+        \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') 
+        d\Omega \approx& \\
         A 
@@
 \end{equation}
-}
-{
+}{
 \begin{equation}
@@
         \int\limits_{R_1}^{R_2}
-    \Big[ \sqrt{r^2 + r'^2 + 2rr'} -& \\
-    \sqrt{r^2 + r'^2 - 2rr'}
-    & \Big] \frac{r'\rho(r')}{r} \, dr'.
+        \Big[ & \sqrt{r^2 + r'^2 + 2rr'} - \\
+        & \sqrt{r^2 + r'^2 - 2rr'}
+        \Big] \frac{r'\rho(r')}{r} \, dr'.
     \end{split}
@@
 
-For our purposes we can assume that the computation point $Q$ is outside the outer radii, i.e. $r>R_2$, and simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
+For our purposes we can assume that the computation point $Q$ is outside the outer radii, i.e. $r>R_2$ ($R_2 \geq r'$), and simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
 
@@
 
+For any density $\rho(r')$, the gradient and Marussi tensor components can be obtained as \citet{Grombein2013} does.
+Only the vertical gradient component ($g_z$) and the diagonal Marussi tensor components ($g_{xx}$, $g_{yy}$, $g_{zz}$) are not zero, and can be easily determined from the potential $V_\text{sh}$:
+
+\begin{equation}
+    g_z = \frac{V_\text{sh}(r)}{r},
+\end{equation}
+\begin{equation}
+    g_{xx} = g_{yy} = -\frac{V_\text{sh}(r)}{r^2}, \quad
+    g_{zz} = \frac{2V_\text{sh}(r)}{r^2}.
+\end{equation}
+
+For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of they variation in $r'$: a linear density and an exponential density.
+
+\subsubsection{Linear Density}
+
+If the density of the spherical shell is a linear function in the radius coordinate, i.e.:
+
+\begin{equation}
+    \rho(r') = ar' + b,
+\end{equation}
+
+\noindent then the gravity potential generated on any computation point at distance $r$ from its centre can be obtained from the following expression:
+
+\begin{equation}
+    V_\text{sh}^\text{lin}(r) = \pi G \left[ 
+    a \frac{R_2^4 - R_1^4}{r} +
+    b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
+    \label{eq:shell-pot-linear}
+\end{equation}
+
+The second term on equation \ref{eq:shell-pot-linear} reassembles the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
+So, this potential can be read as the combination of a spherical shell with the later homogeneous density and the same shell with variable density $\rho(r') = ar'$.
+
+\subsubsection{Exponential Density}
+
+In case of a spherical shell with exponential density in the radius coordinate,
+
+\begin{equation}
+    \rho(r') = Ae^\frac{r'-\Delta h}{b},
+\end{equation}
+
+\noindent we have used \citet{sagemath} to obtain the following expression for the potential:
+
+\iftwocol{
+\begin{equation}
+    \begin{split}
+        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+        Ab \, e^{-\frac{\Delta h}{b}} \Big[ 
+        & (R_2^2 - 2R_2 b + 2b^2)e^\frac{R_2}{b} - \\
+        & (R_1^2 - 2R_1 b + 2b^2)e^\frac{R_1}{b} \Big].
+    \end{split}
+\end{equation}
+}{
+\begin{equation}
+    V_\text{sh}^\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+    Ab \, e^{-\frac{\Delta h}{b}} \left[ 
+    (R_2^2 - 2R_2 b + 2b^2)e^\frac{R_2}{b} - 
+    (R_1^2 - 2R_1 b + 2b^2)e^\frac{R_1}{b} \right].
+\end{equation}
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-07-14 manuscript/manuscript.tex
Edited proposed keywords
Fixed axe (for single axe) and axis (for plural)

@@
 \begin{keywords}
-up to six keywords from the list
+Satellite gravity??, Numerical modelling, Structure of the Earth, Gravity anomalies and Earth structure, Numerical approximations and analysis, spatial analysis??
 \end{keywords}
@@
 
-Due to the rotational symmetry along the $z$ axis, the integration in $\lambda'$ is straightforward:
+Due to the rotational symmetry along the $z$ axe, the integration in $\lambda'$ is straightforward:
 
@@
 
-The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity field generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axis.
+The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity field generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axe.
 Due to the rotational symmetry along any axe that passes through the centre of the shell, the potential in equation \ref{eq:shell-pot} reassembles the gravity field on any outside point at distance $r$ from the its centre.

### 2017-07-14 manuscript/manuscript.tex
Added figure spherical-shell

@@
 Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, which density depends on the radius spherical coordinate.
-The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$ located along the $z$ axis at a distance $r$ can be written as follows:
+The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$ located along the $z$ axe at a distance $r$ can be written as follows:
 
-\todo{figura spherical shell con punto Q}
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/spherical-shell.pdf}
+\caption{
+Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
+The computation point $Q$ is located in the $z$ axe at a distance $r$ from the centre of the shell.
+For practical purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
+}
+\label{fig:spherical-shell}
+\end{figure}
 

### 2017-07-14 manuscript/manuscript.tex
Fixed wrong single column equation
Added figure spherical-shell reference in text

@@
 
-Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, which density depends on the radius spherical coordinate.
+Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, which density depends on the radius spherical coordinate (Figure \ref{fig:spherical-shell}).
 The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$ located along the $z$ axe at a distance $r$ can be written as follows:
@@
 \begin{equation}
-    V_\text{sh}^\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
     Ab \, e^{-\frac{\Delta h}{b}} \left[ 

### 2017-07-17 manuscript/manuscript.tex
Minor correction in analytical solution
Added tesseroid-uieda figure

@@
 The main issue that should be overcome is taking into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
-A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries \todo{Insertar figura de tesseroides de leo}.
+A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries (see Figure \ref{fig:tesseroid-uieda}).
 Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
@@
 The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/tesseroid-uieda.pdf}
+\caption{
+A spherical prism, Tesseroid, in a geocentric spherical coordinate system, with a computation point $P$ and its local north oriented Cartesian coordinate system. Picture made by \citep{Uieda2015}.
+}
+\label{fig:tesseroid-uieda}
+\end{figure}
 
@@
 
-In case of a spherical shell with exponential density in the radius coordinate,
+If, on the other hand, the spherical shell has an exponential density in the radius coordinate,
 

### 2017-07-17 manuscript/manuscript.tex
Minor changes in manuscript

@@
 \caption{
-A spherical prism, Tesseroid, in a geocentric spherical coordinate system, with a computation point $P$ and its local north oriented Cartesian coordinate system. Picture made by \citep{Uieda2015}.
+A spherical prism, also known as  Tesseroid, in a geocentric spherical coordinate system, with a computation point $P$ and its local north oriented Cartesian coordinate system.
+Picture made by \citep{Uieda2015}.
 }

### 2017-07-17 manuscript/manuscript.tex
Started writing implementation

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Implementation}
+
+We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth that is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}
+
+Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows us generate efficient C code through the Cython compiler.
+This produces a precompiled library whose functions and classes can be called from regular Python code, what allowed us to make it compatible with the preexisting open-source library Fatiando a Terra \citep{Uieda2013}.
+
+Its usage is very similar to the code for homogeneous tesseroid already in Fatiando a Terra (version 0.5):
+
+\begin{enumerate}
+\renewcommand{\theenumi}{(\arabic{enumi})}
+\item Define a single tesseroid (or a tesseroid mesh) with the Fatiando a Terra classes.
+\item Add a predefined Python density function to the tesseroid (or each tesseroid in the mesh).
+\item Calculate the desired gravity field (potential, gradient or tensor component) in any set of computation points with our new functions.
+\end{enumerate}
+
+The existing code in Fatiando a Terra (v.~0.5) for homogeneous tesseroid is optimised through the Numba compiler.
+Due to difficulties when passing a Python function as argument of precompiled Numba methods, we have decided to rewrite them in Cython language, keeping the homogeneous tesseroid calculation and adding the new variable density code.
+
+\todo{speed comparission between homogeneous and variable density}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Acknowledgments}

### 2017-07-31 manuscript/manuscript.tex
Minor corrections in equations
Minor modification in manuscript
Modified exponential function in manuscript: now it's a decreasing function.
Minor correction in exponential function expression
Continue writing implementation

@@
     \cos\psi = \sin\phi', \quad
-    \ell = \sqrt{r'^2 + 2 r' r \sin\phi' + r^2}.
+    \ell = \sqrt{r'^2 + 2 r r' \sin\phi' + r^2}.
 \end{equation}
@@
     \int\limits_{R_1}^{R_2}
-    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + 2 r' r \sin\phi' + r^2}}
+    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + 2 r r' \sin\phi' + r^2}}
     \, dr' d\phi',
@@
 \begin{equation}
-    \rho(r') = Ae^\frac{r'-\Delta h}{b},
+    \rho(r') = Ae^{-(r' - \Delta h)/b},
 \end{equation}
 
-\noindent we have used \citet{sagemath} to obtain the following expression for the potential:
+\noindent the potential can be written as follows:
 
@@
         V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-        Ab \, e^{-\frac{\Delta h}{b}} \Big[ 
-        & (R_2^2 - 2R_2 b + 2b^2)e^\frac{R_2}{b} - \\
-        & (R_1^2 - 2R_1 b + 2b^2)e^\frac{R_1}{b} \Big].
+        Ab \, e^\frac{\Delta h}{b}
+        \Big[
+        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+        \Big].
     \end{split}
+\label{eq:shell-pot-exp}
 \end{equation}
@@
     V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-    Ab \, e^{-\frac{\Delta h}{b}} \left[ 
-    (R_2^2 - 2R_2 b + 2b^2)e^\frac{R_2}{b} - 
-    (R_1^2 - 2R_1 b + 2b^2)e^\frac{R_1}{b} \right].
+    Ab \, e^\frac{\Delta h}{b}
+    \Big[
+    & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+    & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+    \Big].
+\label{eq:shell-pot-exp}
 \end{equation}
@@
 
-We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth that is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}
+We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth through the GLQ to numerically approximate the volume integrals.
+It's freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}
 
@@
 
+From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of $N_\lambda N_\phi N_r$ point masses, where $N_i$ are the orders of the quadrature for the integral on each coordinate, with $i \in \{ \lambda, \phi, r \}$.
+
+If we set the GLQ orders to fixed values, it's easy to conclude that the nearer the computation point is to the tesseroid, the accuracy of the approximation is reduced and the point masses effect becomes more noticeable than the tesseroid we want to approximate.
+A more complete analysis of this conclusion can be found in \citet{Ku1977, Li2011, Uieda2016}.
+
+Instead of increasing the GLQ orders, what leads to a more time consuming computation, adaptative discreatization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
+They basically consist in dividing the tesseroid based on a ratio between its size and the distance between its geometric centre and the computation point.
+
+In order to make the computation more efficient, we use only second order GLQ for each integration, and we ensure the accuracy of the method through the application of the modified adaptative discretization algorithm developed by \citet{Uieda2016}.
+
+Given an arbitrary tesseroid, \citet{Uieda2016} redefined its dimensions: $L_\lambda$, $L_\phi$ and $L_r$, where the former ones are the arc-distances measured along its top surface and $L_r$ is the difference between the radii of the top and bottom surfaces.
+We must check that the following inequality is satisfied:
+
+\begin{equation}
+\frac{d}{L_i} \geq D,
+\label{eq:distance-size-ratio}
+\end{equation}
+
+\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point and $D$ is a predefined value called distance-size ratio.
+If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.
+
+In this way, the numerical approximation efficiently improves its accuracy and
+gives the user more control over the computation by allowing to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g. less than 10cm).
+
+On the other hand, the accuracy is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
+In order to solve this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
+
+We have performed similar tests, but now with variable density spherical shell.
+From the infinite density functions that can be proposed, we have chosen two characteristic ones: a linear and an exponential functions.
+The first one represents the simpler variation, while the second one symbolise the fastest varying function.
+The analytical solutions for these two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-08-01 manuscript/manuscript.tex
Fixed typos

@@
 It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ order. 
-Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptative discretization of \citet{Li2011}.
+Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptive discretization of \citet{Li2011}.
 It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.
 \citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
-The modifications made by \citet{Uieda2016} to the original adaptative discretization method \citep{Li2011} consist in:
+The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} consist in:
 (1) a faster calculation of the distance from the computation point to the tesseroid and 
@@
 The spherical prisms known as tesseroids are mass elements defined in a geocentric spherical coordinate system bounded by a pair of parallels, a pair of  meridians, and two concentric spherical surfaces.
-We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented cartesian coordinate system located at $P$.
+We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented Cartesian coordinate system located at $P$.
 In the special case of homogeneous density, the gravity potential, gradient and Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[see also][]{Uieda2016}.
@@
 
-For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$, respectivetly.
+For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$, respectively.
 
@@
 
-\todo{speed comparission between homogeneous and variable density}
+\todo{speed comparison between homogeneous and variable density}
 
@@
 
-Instead of increasing the GLQ orders, what leads to a more time consuming computation, adaptative discreatization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
+Instead of increasing the GLQ orders, what leads to a more time consuming computation, adaptive discretization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
 They basically consist in dividing the tesseroid based on a ratio between its size and the distance between its geometric centre and the computation point.
 
-In order to make the computation more efficient, we use only second order GLQ for each integration, and we ensure the accuracy of the method through the application of the modified adaptative discretization algorithm developed by \citet{Uieda2016}.
+In order to make the computation more efficient, we use only second order GLQ for each integration, and we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
 
@@
 If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.
+A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
 

### 2017-08-01 manuscript/manuscript.tex
Modifications in manuscript

@@
 In this way, the numerical approximation efficiently improves its accuracy and
-gives the user more control over the computation by allowing to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g. less than 10cm).
+gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
+less than 10cm), which does not modifies significantly the desired gravitational field.
 
@@
 From the infinite density functions that can be proposed, we have chosen two characteristic ones: a linear and an exponential functions.
-The first one represents the simpler variation, while the second one symbolise the fastest varying function.
+The first one constitutes the simplest variation, while the second one represents the fastest varying function.
 The analytical solutions for these two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.

### 2017-08-01 manuscript/manuscript.tex
Continue writing manuscript. Added figs: linear density tests

@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/spherical-shell.pdf}
+\includegraphics[width=0.7\linewidth]{figures/spherical-shell.pdf}
 \caption{
@@
 
-In this way, the numerical approximation efficiently improves its accuracy and
-gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
+In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided and therefore both the accuracy of the algorithm and its computation time.
+In this way, the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
 less than 10cm), which does not modifies significantly the desired gravitational field.
 
-On the other hand, the accuracy is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
+On the other hand, the accuracy of the method is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
 In order to solve this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
@@
 The analytical solutions for these two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Determination of distance-size ratio}
+
+In order to test the new variable density tesseroid code and to estimate the error of the approximation we have compared the numerical model with the analytical solutions for a spherical shell with variable density for two special cases: a linear and an exponential density function.
+This error estimations are also useful in the determination of the value for the distance-size ratio $D$, such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
+
+We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the $g_{zz}$ component of the Marussi tensor on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
+More information about this grids can be found in the Table \ref{tab:grids}.
+
+\todo{explicar pq los otros campos no los testeamos}
+
+We computed the three gravity fields for different values of $D$, ranging from 1 to 10 with a step of 0.5 on every point of these four grids.
+Then, we calculated the maximum difference between these results and the value of the analytical solution at the same height.
+
+Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
+That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
+
+Finally, we set the acceptable value of $D$ at which the error of the numerical approximation is lower than 10$^{-1}\%$.
+
+\begin{table}
+\caption{Description of the synthetic grids where the comparison of the numerical model against the analytical solutions for a spherical shell was done.
+This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height.}
+\label{tab:grids}
+\begin{tabular}{lcccc}
+    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
+    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
+    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
+\end{tabular}
+\end{table}
+
+\subsection{Linear Density}
+
+As a simple case, we proposed a linear density function such as the spherical shell has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:
+
+\begin{equation}
+    \rho(r') = 
+    -\frac{3300 - 2670}{35000} r' + 
+    \left( \frac{3300 - 2670}{35000} R + 2670 \right)
+\end{equation}
+
+\noindent where $R$ = 6378.137 km is the mean Earth radius.
+
+The tests were performed on a thin and a thick spherical shell of 1 km and 35 km of thickness, respectively.
+The results for both cases can be seen in Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
+
+We can see that for a linear density, the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 10$^{-1}\%$.
+On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.
+We will keep the more conservative result of $D=8$ in case of computing the $g_{zz}$ tensor component in case of a linear density tesseroid.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Dlinear-thin-differences.pdf}
+\caption{}
+\label{fig:D-linear-thin}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Dlinear-thick-differences.pdf}
+\caption{}
+\label{fig:D-linear-thick}
+\end{figure}
+
+\subsection{Exponential Density}
 

### 2017-08-01 manuscript/manuscript.tex
Continue writing manuscript

@@
 
-We computed the three gravity fields for different values of $D$, ranging from 1 to 10 with a step of 0.5 on every point of these four grids.
+We computed the three gravity fields for different values of $D$, ranging from 0.5 to 10 with a step of 0.5 on every point of these four grids.
 Then, we calculated the maximum difference between these results and the value of the analytical solution at the same height.
@@
 
+To test the model in case of a exponential density we proposed the following function:
+
+\begin{equation}
+\rho(r') = A e^{-(r' - R)/b}
+\end{equation}
+
+\noindent where $A$ = 1 kg/m$^3$ and $b$ = 10$^6$ m.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-08-02 manuscript/manuscript.tex
Continue writing manuscript: added figures for exponential density

@@
 \begin{table}
-\caption{Description of the synthetic grids where the comparison of the numerical model against the analytical solutions for a spherical shell was done.
-This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height.}
+\caption{
+    Description of the synthetic grids where the comparison of the numerical model against the analytical solutions for a spherical shell was done.
+This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height.
+}
 \label{tab:grids}
@@
     \left( \frac{3300 - 2670}{35000} R + 2670 \right)
+    \label{eq:density-linear}
 \end{equation}
@@
 \includegraphics[width=0.9\linewidth]{figures/Dlinear-thin-differences.pdf}
-\caption{}
+\caption{
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the linear density defined on equation \ref{eq:density-linear}.
+The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
 \label{fig:D-linear-thin}
@@
 \includegraphics[width=0.9\linewidth]{figures/Dlinear-thick-differences.pdf}
-\caption{}
+\caption{
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness and the linear density defined on equation \ref{eq:density-linear}.
+The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
 \label{fig:D-linear-thick}
@@
 
-To test the model in case of a exponential density we proposed the following function:
+To test the model in case of an exponential density we proposed the following function:
 
@@
     \rho(r') = A e^{-(r' - R)/b}
+\label{eq:density-exp}
 \end{equation}
@@
 \noindent where $A$ = 1 kg/m$^3$ and $b$ = 10$^6$ m.
+We have also performed this test with a thin and a thick shell of 1 km and 35 km, respectively.
+The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thin-differences.pdf}
+\caption{
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the exponential density defined on equation \ref{eq:density-exp}.
+The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
+\label{fig:D-exp-thin}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thick-differences.pdf}
+\caption{
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness and the exponential density defined on equation \ref{eq:density-exp}.
+The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
+\label{fig:D-exp-thick}
+\end{figure}
 

### 2017-08-04 manuscript/manuscript.tex
Continue writing manuscript

@@
 
-\todo{explicar pq los otros campos no los testeamos}
-
 We computed the three gravity fields for different values of $D$, ranging from 0.5 to 10 with a step of 0.5 on every point of these four grids.
@@
 
+The tests were only performed on these three magnitudes because the other components of the gravity gradient and Marussi tensor are equal to zero due to the rotational symmetries of the spherical shell.
+
 \begin{table}
@@
 This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height.
+Every grid consists in a set of 10$\times$10 points.
 }
@@
 \noindent where $A$ = 1 kg/m$^3$ and $b$ = 10$^6$ m.
-We have also performed this test with a thin and a thick shell of 1 km and 35 km, respectively.
-The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+We have also performed this test with a thin and a thick shell of 1 km and 35 km, respectively, and the results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+
+In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
+
+Surprisingly both the linear and the exponential density tesseroid models needs the same values of distance-size ratio $D$ in order to assure a good accuracy.
+Because the two functions are significantly different we can expect that these values of $D$ are good for a wide variety of density functions.
 

### 2017-08-09 manuscript/manuscript.tex
Small changes in manuscript

@@
 A spherical prism, also known as  Tesseroid, in a geocentric spherical coordinate system, with a computation point $P$ and its local north oriented Cartesian coordinate system.
-Picture made by \citep{Uieda2015}.
+Picture made by \citet{Uieda2015}.
 }
@@
 \end{figure}
-
-In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
-On the other hand, the GLQ allows us to calculate them for any density function without any other information but the function itself.
 
@@
 
+In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
+On the other hand, the GLQ allows us to calculate them for any density function without any other information but the function itself.
+
 It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ order. 
 Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptive discretization of \citet{Li2011}.
-It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio ($D$) is greater than a predefined value that controls the precision of the computation.
+It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio is greater than a predefined value ($D$) that controls the accuracy of the computation. 
 \citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
@@
 
-In our case, we will suppose that the tesseroid has an arbitrary variable density in depth, this means that only depends on the radius spherical coordinate.
-Thus, the integrals for the gravity fields are slightly modified.
+In our case, we will suppose that the tesseroid has an arbitrary variable density in depth, i.e. only depends on the radius spherical coordinate.
+Thus, the integrals for the gravity fields are slightly modified:
 
@@
 From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates the gravity fields of a tesseroid with variable density as the ones generated by point masses located at the nodes of the Legendre polynomials.
-Furthermore, the information of the density function is summarised as the values it assumes in those nodes.
+Furthermore, the information of the density function is summarised as the values it assumes on those nodes.
 Although this may sound discouraging, we intent to prove otherwise and show that this method, along with a well fitted discretization algorithm, approximates the gravity fields with good accuracy and precision.\todo{no se si poner esta ultima oracion}
@@
 In case of variable density tesseroids, the comparison must obviously be made with a variable density spherical shell, whose analytical solution will depend on the chosen density function.
-Because the literature has not offered it yet, we needed to develop ourselves.
+Because the literature has not offered it yet, we needed to develop it ourselves.
 
 Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, which density depends on the radius spherical coordinate (Figure \ref{fig:spherical-shell}).
-The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$ located along the $z$ axe at a distance $r$ can be written as follows:
+The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$, located along the $z$ axe at a distance $r$ from the origin, can be written as follows:
 
@@
 For any density $\rho(r')$, the gradient and Marussi tensor components can be obtained as \citet{Grombein2013} does.
-Only the vertical gradient component ($g_z$) and the diagonal Marussi tensor components ($g_{xx}$, $g_{yy}$, $g_{zz}$) are not zero, and can be easily determined from the potential $V_\text{sh}$:
+Only the vertical gradient component ($g_z$) and the diagonal Marussi tensor components ($g_{xx}$, $g_{yy}$, $g_{zz}$) are non zero, and can be easily determined from the potential $V_\text{sh}$:
 
@@
 
-For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of they variation in $r'$: a linear density and an exponential density.
+For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of their variation in $r'$: a linear density and an exponential density.
 
@@
 
-If the density of the spherical shell is a linear function in the radius coordinate, i.e.:
+If the density of the spherical shell is a linear function on the radius coordinate, i.e.:
 
@@
 
-Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows us generate efficient C code through the Cython compiler.
+Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows us to generate efficient C code through the Cython compiler.
 This produces a precompiled library whose functions and classes can be called from regular Python code, what allowed us to make it compatible with the preexisting open-source library Fatiando a Terra \citep{Uieda2013}.
@@
 
-Instead of increasing the GLQ orders, what leads to a more time consuming computation, adaptive discretization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
+Instead of increasing the GLQ orders to improve the accuracy, what leads to a more time consuming computation, adaptive discretization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
 They basically consist in dividing the tesseroid based on a ratio between its size and the distance between its geometric centre and the computation point.
@@
 In order to make the computation more efficient, we use only second order GLQ for each integration, and we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
-
 Given an arbitrary tesseroid, \citet{Uieda2016} redefined its dimensions: $L_\lambda$, $L_\phi$ and $L_r$, where the former ones are the arc-distances measured along its top surface and $L_r$ is the difference between the radii of the top and bottom surfaces.
@@
 In this way, the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
-less than 10cm), which does not modifies significantly the desired gravitational field.
+less than 10 cm), which does not modifies significantly the desired gravitational field.
 
@@
 In order to test the new variable density tesseroid code and to estimate the error of the approximation we have compared the numerical model with the analytical solutions for a spherical shell with variable density for two special cases: a linear and an exponential density function.
-This error estimations are also useful in the determination of the value for the distance-size ratio $D$, such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
+This error estimations are also useful in the determination of the minimum value for the distance-size ratio $D$ such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
 
@@
 
-We computed the three gravity fields for different values of $D$, ranging from 0.5 to 10 with a step of 0.5 on every point of these four grids.
+We computed the three gravity fields for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, on every point of these four grids.
 Then, we calculated the maximum difference between these results and the value of the analytical solution at the same height.
@@
 \caption{
-    Description of the synthetic grids where the comparison of the numerical model against the analytical solutions for a spherical shell was done.
+    Description of the synthetic grids on which the comparison of the numerical model against the analytical solutions for a spherical shell was done.
 This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height.
@@
 
-As a simple case, we proposed a linear density function such as the spherical shell has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:
+As a simple case, we proposed a linear density function in such a way that the spherical shell has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:
 
@@
 
-We can see that for a linear density, the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 10$^{-1}\%$.
+For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 10$^{-1}\%$.
 On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.

### 2017-08-09 manuscript/manuscript.tex
Started writing discontinuous density in manuscript

@@
     \int\limits_a^b f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i^s),
+\label{eq:glq-scaled}
 \end{equation}
@@
 For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of their variation in $r'$: a linear density and an exponential density.
+\todo{agregar discontinua}
 
@@
 We have performed similar tests, but now with variable density spherical shell.
-From the infinite density functions that can be proposed, we have chosen two characteristic ones: a linear and an exponential functions.
+From the infinite density functions that can be proposed, we have chosen two characteristic ones: a linear and an exponential functions. \todo{agregar discontinua}
 The first one constitutes the simplest variation, while the second one represents the fastest varying function.
@@
 In order to test the new variable density tesseroid code and to estimate the error of the approximation we have compared the numerical model with the analytical solutions for a spherical shell with variable density for two special cases: a linear and an exponential density function.
+\todo{agregar discontinua}
 This error estimations are also useful in the determination of the minimum value for the distance-size ratio $D$ such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
@@
 Surprisingly both the linear and the exponential density tesseroid models needs the same values of distance-size ratio $D$ in order to assure a good accuracy.
-Because the two functions are significantly different we can expect that these values of $D$ are good for a wide variety of density functions.
+Because the two functions are significantly different we can expect that these values of $D$ are good for a wide variety of density functions.\todo{no se si da para poner esto}
 
@@
 
+\subsection{Discontinuous Density}
+
+As a final case, we propose to perform the same test with a discontinuous density function:
+
+\begin{equation}
+    \rho(r') = 
+    \begin{cases}
+        \rho_1 & \text{if} \, R_1 \leq r' < R_c \\
+        \rho_2 & \text{if} \, R_c \leq r' \leq R_2 \\
+    \end{cases}
+\label{eq:density-discontinuous}
+\end{equation}
+
+\noindent where $\rho_1$ and $\rho_2$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radii on which the 
+density changes.
+
+\todo{poner solucion analitica}
+
+First we assume a symmetric density function, i.e. $R_c=(R_1 + R_2)/2$ is the centre radius of the shell, and perform the test for a thin shell of 1km of thickness.
+The results can be seen in Figure \ref{fig:D-discont-sym}.
+Again, we obtain good accuracy for values of $D=$1,2 and 8 for the computation of the potential, the gradient components and the Marussi tensor components, respectively.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-symmetric-differences.pdf}
+\caption{
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the discontinuous density defined on equation \ref{eq:density-discontinuous}, with $R_c$ as the centre radius of the shell.
+The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
+\label{fig:D-discont-sym}
+\end{figure}
+
+Secondly, lets assume the same density function from equation \ref{eq:density-discontinuous}, but now with an asymmetrical distribution such as $R_c = (R_1 + R_2)/3$, and perform the same test.
+We can see its results in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what
+lead us to conclude that the code does not reassembles the gravity fields for this case with good accuracy.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-asymmetric-differences.pdf}
+\caption{
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the discontinuous density defined on equation \ref{eq:density-discontinuous}, with $R_c=(R_1 + R_2)/3$ is not the centre radius of the shell.
+The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
+\label{fig:D-discont-asym}
+\end{figure}
+
+From equation \ref{eq:glq-var-dens} we have seen that the information of the density function is summarised as the values its assumes on the nodes of the GLQ.
+We can attribute the low performance of the code to this simplification.
+
+On the other hand, we can explain the good performance in the symmetrical density case due to the fact that the GLQ nodes are always symmetrical with respect to the centre of the $[a,b]$ integration interval (see equation \ref{eq:glq-scaled}).
+
+In the light of these results we do not recommend using discontinuous densities functions.
+Instead, divide the tesseroid into several ones with specific homogeneous density or densityfunction for each one.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-08-10 manuscript/manuscript.tex
Minor changes in manuscript
Continue writing manuscript: added changes in the power of exponential density

@@
     Description of the synthetic grids on which the comparison of the numerical model against the analytical solutions for a spherical shell was done.
-This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height.
+This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height over the Earth surface.
 Every grid consists in a set of 10$\times$10 points.
@@
 
+Because the constant $b$ controls the variation of the density function we wanted to test how does the accuracy of the method behaves at different values of $b$.
+We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}).
+For each value of $b$ we computed these difference for the same set of values of $D$ we've been using on the previous tests, obtaining one curve for each kind of variation.
+
+The results can be seen in Figure \ref{fig:D-exp-power}.
+Lower values of $b$ implies a more varying function.
+We can see that for extremely varying functions the numerical model does not perform well: for a $b$=100m the difference is almost 100\%.
+
+\todo{poner analisis con respecto al exponente cuando b=1000m}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences.pdf}
+\caption{
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the exponential density defined on equation \ref{eq:density-exp} for different values of $b$.
+Each computation was performed on the big grid described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
+\label{fig:D-exp-power}
+\end{figure}
+
 \subsection{Discontinuous Density}

### 2017-08-11 manuscript/manuscript.tex
Minor changes in manuscript

@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the exponential density defined on equation \ref{eq:density-exp} for different values of $b$.
+    Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the exponential density defined on equation \ref{eq:density-exp} for different values of $b$.
 Each computation was performed on the big grid described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.

### 2017-08-11 manuscript/manuscript.tex
Changes in manuscript

--- 2017-08-10 manuscript/manuscript.tex
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the exponential density defined on equation \ref{eq:density-exp} for different values of $b$.
-Each computation was performed on the big grid described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+    Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for (a) a thin spherical shell of 1km of thickness and (b) a thick sphericall shell of 35km of thickness, both with the exponential density defined on equation \ref{eq:density-exp}.
+Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and $b$.
 If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
@@
 
-As a final case, we propose to perform the same test with a discontinuous density function:
+As a final case, we performed the same test with a discontinuous density function:
 
@@
 
-\noindent where $\rho_1$ and $\rho_2$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radii on which the 
+\noindent where $\rho_1$ and $\rho_2$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radius on which the 
 density changes.
 
-\todo{poner solucion analitica}
+The analytical solution for this kind of density is the linear combination of two homogeneous shells with densities $\rho_1$ and $\rho_2$:
 
-First we assume a symmetric density function, i.e. $R_c=(R_1 + R_2)/2$ is the centre radius of the shell, and perform the test for a thin shell of 1km of thickness.
+\begin{equation}
+    V(r) = \frac{4}{3} \frac{\pi G}{r}
+    \left[ \rho_1 (R_c^3 - R_1^3) +
+    \rho_2 (R_2^3 - R_c^3) \right]
+\end{equation}
+
+Firstly, lets assume a symmetric density function, i.e. $R_c=(R_1 + R_2)/2$ is the centre radius of the shell, and perform the test for a thin shell of 1km of thickness.
 The results can be seen in Figure \ref{fig:D-discont-sym}.
@@
 
-Secondly, lets assume the same density function from equation \ref{eq:density-discontinuous}, but now with an asymmetrical distribution such as $R_c = (R_1 + R_2)/3$, and perform the same test.
+Secondly, lets assume the same density function from equation \ref{eq:density-discontinuous}, but now with an asymmetrical distribution: $R_c = (R_1 + R_2)/3$, and perform the same test.
 We can see its results in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what

### 2017-08-11 manuscript/manuscript.tex
Continue writing manuscript

@@
 Surprisingly both the linear and the exponential density tesseroid models needs the same values of distance-size ratio $D$ in order to assure a good accuracy.
-Because the two functions are significantly different we can expect that these values of $D$ are good for a wide variety of density functions.\todo{no se si da para poner esto}
 
@@
 
-The results can be seen in Figure \ref{fig:D-exp-power}.
-Lower values of $b$ implies a more varying function.
-We can see that for extremely varying functions the numerical model does not perform well: for a $b$=100m the difference is almost 100\%.
+This tests were performed using both the thin and the thick spherical shell, and the results can be seen in Figure \ref{fig:D-exp-power}.
+Taking into account that lower values of $b$ implies a more varying function, we can see that for extremely varying functions the numerical model does not perform well: for $b$=100m the difference is almost 100\% in the thin shell case, and the same happens with $b$=10km in the thick shell case.
 
-\todo{poner analisis con respecto al exponente cuando b=1000m}
+On both cases, the model performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, in other words, when the absolute value of the exponent in equation \ref{eq:density-exp} is lower than 1.
+Although this sets a limitation to the model, the maximum acceptable variation is very far from real density distributions inside the Earth.
+For example, if we take into account a tesseroid with 35km of thickness, a density of 2670kg/m$^3$ on its top surface and $b=35$km, then the density at its bottom will be approximately of 7000kg/m$^3$, very far from the values we can find at such depths.
 
@@
 We can see its results in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what
-lead us to conclude that the code does not reassembles the gravity fields for this case with good accuracy.
+leads us to conclude that the code does not reassembles the gravity fields for this case with good accuracy.
 
@@
 In the light of these results we do not recommend using discontinuous densities functions.
-Instead, divide the tesseroid into several ones with specific homogeneous density or densityfunction for each one.
+Instead, divide the tesseroid into several ones with specific homogeneous density or density function for each one.
 

### 2017-08-14 manuscript/manuscript.tex
Corrections to manuscript

@@
 \end{keywords}
-
-\todo{aclarar coordenadas esfericas -> longitud latitud angulos}
 
@@
 
-For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of their variation in $r'$: a linear density and an exponential density.
-\todo{agregar discontinua}
+For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of their variation in $r'$: a linear, an exponential and a discontinuous density function.
 
@@
 From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of $N_\lambda N_\phi N_r$ point masses, where $N_i$ are the orders of the quadrature for the integral on each coordinate, with $i \in \{ \lambda, \phi, r \}$.
-
 If we set the GLQ orders to fixed values, it's easy to conclude that the nearer the computation point is to the tesseroid, the accuracy of the approximation is reduced and the point masses effect becomes more noticeable than the tesseroid we want to approximate.
@@
 Instead of increasing the GLQ orders to improve the accuracy, what leads to a more time consuming computation, adaptive discretization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
-They basically consist in dividing the tesseroid based on a ratio between its size and the distance between its geometric centre and the computation point.
+They essentially consist in dividing the tesseroid based on a ratio between its size and the distance between its geometric centre and the computation point.
 
@@
 
-\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point and $D$ is a predefined value called distance-size ratio.
+\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point, and $D$ is a predefined value called distance-size ratio.
 If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.
@@
 In this way, the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
-less than 10 cm), which does not modifies significantly the desired gravitational field.
+less than 10 cm), in which case it does not modifies significantly the desired gravitational field.
 
@@
 We have performed similar tests, but now with variable density spherical shell.
-From the infinite density functions that can be proposed, we have chosen two characteristic ones: a linear and an exponential functions. \todo{agregar discontinua}
+From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
 The first one constitutes the simplest variation, while the second one represents the fastest varying function.
-The analytical solutions for these two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.
+The later one tries to emulate a contact between two different media (e.g.
+crust and mantle interface).
+The analytical solutions for the first two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.
 
@@
 
-In order to test the new variable density tesseroid code and to estimate the error of the approximation we have compared the numerical model with the analytical solutions for a spherical shell with variable density for two special cases: a linear and an exponential density function.
-\todo{agregar discontinua}
+In order to test the new variable density tesseroid code and estimate the error of the approximation we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
 This error estimations are also useful in the determination of the minimum value for the distance-size ratio $D$ such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the linear density defined on equation \ref{eq:density-linear}.
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness and the linear density defined on equation \ref{eq:density-linear}.
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
@@
 
-\noindent where $A$ = 1 kg/m$^3$ and $b$ = 10$^6$ m.
+\noindent where $A$ = 1kg/m$^3$ and $b$ = 1000km.
 We have also performed this test with a thin and a thick shell of 1 km and 35 km, respectively, and the results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the exponential density defined on equation \ref{eq:density-exp}.
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness and the exponential density defined on equation \ref{eq:density-exp}.
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
@@
 \caption{
-    Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for (a) a thin spherical shell of 1km of thickness and (b) a thick sphericall shell of 35km of thickness, both with the exponential density defined on equation \ref{eq:density-exp}.
-Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and $b$.
+    Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for (a) a thin spherical shell of 1km of thickness and (b) a thick sphericall shell of 35km of thickness, both with the exponential density defined on equation \ref{eq:density-exp} with different values of $b$.
+Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
+Each curve corresponds to a value of $b$.
 If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
@@
 The results can be seen in Figure \ref{fig:D-discont-sym}.
-Again, we obtain good accuracy for values of $D=$1,2 and 8 for the computation of the potential, the gradient components and the Marussi tensor components, respectively.
+Again, we obtain good accuracy for values of $D$ equal to 1, 2 and 8 for the computation of the potential, the gradient components and the Marussi tensor components, respectively.
 
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the discontinuous density defined on equation \ref{eq:density-discontinuous}, with $R_c$ as the centre radius of the shell.
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c$ as the centre radius of the shell.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness and the discontinuous density defined on equation \ref{eq:density-discontinuous}, with $R_c=(R_1 + R_2)/3$ is not the centre radius of the shell.
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c=(R_1 + R_2)/3$.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.

### 2017-08-14 manuscript/manuscript.tex
Changes in manuscript

@@
 
+\todo{corregir esto y explicar mejor}
+
 \begin{equation}
@@
 
-For practical purposes we are going to obtain the gravity potential for two very different functions in the perspective of their variation in $r'$: a linear, an exponential and a discontinuous density function.
+For practical purposes we are going to obtain the gravity potential for three different functions in the perspective of their variation in $r'$: a linear, an exponential and a discontinuous density function.
 
@@
 
-Finally, we set the acceptable value of $D$ at which the error of the numerical approximation is lower than 10$^{-1}\%$.
+Finally, we set the acceptable value of $D$ at which the error of the numerical approximation is lower than 0.1\%.
 
 The tests were only performed on these three magnitudes because the other components of the gravity gradient and Marussi tensor are equal to zero due to the rotational symmetries of the spherical shell.
+\todo{corregir esto, los elementos diagonales del tensor fueron testeados, pero no lo mostramos pq son muy similares al gzz}
 
@@
 
-For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 10$^{-1}\%$.
+For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
 On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.
@@
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
 }
@@
 Each curve corresponds to a value of $b$.
-If the difference is less than 10$^{-1}\%$, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
 }

### 2017-08-15 manuscript/manuscript.tex
Modifications to manuscript: added gxx and gyy notice

@@
 
-We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the $g_{zz}$ component of the Marussi tensor on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
+We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
 More information about this grids can be found in the Table \ref{tab:grids}.
@@
 
-The tests were only performed on these three magnitudes because the other components of the gravity gradient and Marussi tensor are equal to zero due to the rotational symmetries of the spherical shell.
-\todo{corregir esto, los elementos diagonales del tensor fueron testeados, pero no lo mostramos pq son muy similares al gzz}
+The tests were only performed on these magnitudes because the other components of the gravity gradient and Marussi tensor are equal to zero due to the rotational symmetries of the spherical shell.
 
@@
 The tests were performed on a thin and a thick spherical shell of 1 km and 35 km of thickness, respectively.
-The results for both cases can be seen in Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
+In Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick} we can see the results for the potential, the vertical component of the gradient and the $g_{zz}$ component of the Marussi tensor.
+Because the other two diagonal components of the tensor show similar results as the ones obtained for the $g_{zz}$, we exclude them from the figure.
+Although they can be found in the repository. \todo{chequear}
 
@@
 On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.
-We will keep the more conservative result of $D=8$ in case of computing the $g_{zz}$ tensor component in case of a linear density tesseroid.
+We will keep the more conservative result of $D=8$ in case of computing the Marussi tensor components in case of a linear density tesseroid.
 
@@
 We have also performed this test with a thin and a thick shell of 1 km and 35 km, respectively, and the results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them. Although, they can be found in the repository.
 
@@
 We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}).
-For each value of $b$ we computed these difference for the same set of values of $D$ we've been using on the previous tests, obtaining one curve for each kind of variation.
+For each value of $b$ we calculated these difference for the same set of values of $D$ we've been using on the previous tests, obtaining one curve for each kind of variation.
 

### 2017-08-15 manuscript/manuscript.tex
Started writing conclusions

@@
 
+\section{Conclusions}
+
+We have developed a Python library that allows us to perform forward gravity computations using tesseroids with variable density in depth.
+The code numerically approximates the integral that defines the gravity potential, its gradient and the Marussi tensor through the Gauss-Legendre Quadrature (GLQ).
+It essentially consists in approximating the gravity fields of the tesseroids with the effect of point masses located on the scaled nodes of the Legendre polynomials.
+We state that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involves Taylor series expansion.
+
+Although this new code uses only second order GLQ on each integration, we've enhanced its accuracy through a modified adaptive discretization method.
+It divides the tesseroid by half if the ratio of the distance to the computation point and its size is lower than a predefined distance-size ratio $D$.
+We can summarise that $D$ controls the accuracy of the method: a higher value of $D$ generates more divisions of the tesseroids, thou more point masses, what leads to a more precise approximation.
+
+On the other hand, there's no direct relation between the value of $D$ and the error of the computation, that's why we had to determine the minimum value of $D$ for each gravity magnitude that produces an acceptable accuracy.
+These distance-size ratio determinations have been made through the comparison of the numerical approximation with the analytical solutions for the exceptional case of a spherical shell.
+From the wide range of possible density variations, we've chosen a (1) linear, (2) an exponential and (3) a discontinuous density functions.
+
+The tests over the first two functions have shown that in order to achieve an accuracy of 0.1\%, a value of $D$ equal to 1, 2 and 8 must be used in the calculation of the gravity potential, its gradient components and the components of the Marussi tensor, respectively.
+We have also found a limitation on the variation of the exponential density function: the absolute value of the exponent must not be greater than 1.
+However, those extremely variations cannot be found inside the Earth, so this limitation does not determines a true restriction for real life models.
+
+On the other hand, the tests over the discontinuous function were not satisfactory.
+In such cases we recommend to divide the model in several tesseroids, each one with a specific homogeneous density or a continuous density function, instead of using one variable density tesseroid.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Acknowledgments}

### 2017-08-16 manuscript/manuscript.tex
Added speed comparison in manuscript

@@
 
-\todo{speed comparison between homogeneous and variable density}
+We have computed the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous and variable density at different heights and recorded their computation times (see Figure \ref{fig:speed-comparison}).
+We can see that the homogeneous density code is faster than the new variable density algorithm: up two twice faster.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
+\caption{Computation times for a single tesseroid model with homogeneous and variable density at different heights.}
+\label{fig:speed-comparison}
+\end{figure}
 

### 2017-08-17 manuscript/manuscript.tex
Changes in manuscript and speed-comparison figure

--- 2017-08-15 manuscript/manuscript.tex
@@
 
-Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows us to generate efficient C code through the Cython compiler.
+Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows to generate efficient C code through the Cython compiler.
 This produces a precompiled library whose functions and classes can be called from regular Python code, what allowed us to make it compatible with the preexisting open-source library Fatiando a Terra \citep{Uieda2013}.
@@
 Due to difficulties when passing a Python function as argument of precompiled Numba methods, we have decided to rewrite them in Cython language, keeping the homogeneous tesseroid calculation and adding the new variable density code.
-
-\todo{speed comparison between homogeneous and variable density}
 
@@
 
+We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous and variable density at different heights (see Figure \ref{fig:speed-comparison}).
+Firstly, we can see that both codes take more time to compute the gravity field for lower heights due to the increased divisions of the tesseroid in such cases.
+Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm, however the later is fast enough to perform the calculations of a real model in a few minutes.
+\todo{real, esta bien la expresion?}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
+\caption{Computation times for a single tesseroid model with homogeneous and variable density at different heights.}
+\label{fig:speed-comparison}
+\end{figure}
+
 On the other hand, the accuracy of the method is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
-In order to solve this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
+In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
 

### 2017-08-17 manuscript/manuscript.tex
Minor corrections in manuscript

@@
 It essentially consists in approximating the gravity fields of the tesseroids with the effect of point masses located on the scaled nodes of the Legendre polynomials.
-We state that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involves Taylor series expansion.
+We state that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involve Taylor series expansion.
 
-Although this new code uses only second order GLQ on each integration, we've enhanced its accuracy through a modified adaptive discretization method.
+Instead of enhancing the accuracy of the method by increasing the order of the GLQ on each integration, we make use of the modified adaptive discretization algorithm.
 It divides the tesseroid by half if the ratio of the distance to the computation point and its size is lower than a predefined distance-size ratio $D$.
-We can summarise that $D$ controls the accuracy of the method: a higher value of $D$ generates more divisions of the tesseroids, thou more point masses, what leads to a more precise approximation.
+We can summarise that $D$ controls the accuracy of the method: a higher value of $D$ generates more divisions of the tesseroids, thus more point masses, what leads to a more precise approximation.
 

### 2017-08-18 manuscript/manuscript.tex
Minor changes in manuscript

@@
 Firstly, we can see that both codes take more time to compute the gravity field for lower heights due to the increased divisions of the tesseroid in such cases.
-Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm, however the later is fast enough to perform the calculations of a real model in a few minutes.
+Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights, however the computation times gets more similar values at greater heights.
+In conclusion, the new variable dnesity code is fast enough to perform the calculations of a real model in a few minutes.
 \todo{real, esta bien la expresion?}
@@
 We computed the three gravity fields for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, on every point of these four grids.
-Then, we calculated the maximum difference between these results and the value of the analytical solution at the same height.
+Then we calculated the maximum difference between these results and the value of the analytical solution at the same height.
 
@@
 
-Finally, we set the acceptable value of $D$ at which the error of the numerical approximation is lower than 0.1\%.
+Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
 
@@
 \begin{equation}
-    \rho(r') = 
-    -\frac{3300 - 2670}{35000} r' + 
-    \left( \frac{3300 - 2670}{35000} R + 2670 \right)
+    \rho(r') = ar' + c,
+    %~ -\frac{3300 - 2670}{35000} r' + 
+    %~ \left( \frac{3300 - 2670}{35000} R + 2670 \right)
     \label{eq:density-linear}
+\end{equation}
+\noindent with 
+\begin{equation}
+    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{35000\text{m}},
+\end{equation}
+\begin{equation}
+    c = \frac{3300\text{kg/m$^3$} - 
+        2670\text{kg/m$^3$}}{35000\text{m}} R + 
+        2670\text{kg/m$^3$},
 \end{equation}

### 2017-08-18 manuscript/manuscript.tex
Minor changes in manuscript

@@
 The tests were performed on a thin and a thick spherical shell of 1 km and 35 km of thickness, respectively.
-In Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick} we can see the results for the potential, the vertical component of the gradient and the $g_{zz}$ component of the Marussi tensor.
-Because the other two diagonal components of the tensor show similar results as the ones obtained for the $g_{zz}$, we exclude them from the figure.
-Although they can be found in the repository. \todo{chequear}
+In Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick} we can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor.
+Because the other two diagonal components of the tensor show similar results as the ones obtained for the $g_{zz}$, we exclude them from the figure, although they can be found in the repository.
+\todo{chequear}
 
 For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
-On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D=8$ while the later only a $D=3.5$.
-We will keep the more conservative result of $D=8$ in case of computing the Marussi tensor components in case of a linear density tesseroid.
+On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D$ equal to 8 while the later only a $D$ of 3.5.
+We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.
 
@@
 We have also performed this test with a thin and a thick shell of 1 km and 35 km, respectively, and the results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
-Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them. Although, they can be found in the repository.
+Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them, although, they can be found in the repository.
 

### 2017-08-23 manuscript/manuscript.tex
Minor corrections manuscript

@@
 
-Lithosphere density variation with depth has been studied for almost a century. 
+Lithosphere's density variation with depth has been studied for almost a century. 
 \citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
-In the following years, other density functions have been proposed for different rock types \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
+In the following years other density functions have been proposed for different rock types, such as linear, quadratic and exponential \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
 Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.

### 2017-08-23 manuscript/manuscript.tex
Changes in manuscript

@@
 Lithosphere's density variation with depth has been studied for almost a century. 
-\citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples.
-In the following years other density functions have been proposed for different rock types, such as linear, quadratic and exponential \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
+On the first years \citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples, while in the following decades other density functions have been proposed for different rock types \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
 Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
@@
 These forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
-Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation in regional scales.
-This raises the importance of building forward models that reassemble the gravity anomalies for such scales.
+Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation on regional scales.
+This raises the importance of building forward models that reproduce the gravity anomalies for such scales.
 
@@
 A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries (see Figure \ref{fig:tesseroid-uieda}).
-Given an arbitrary tesseroid, calculating the gravity fields on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
+Calculating the gravity fields generated by an arbitrary tesseroid on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
 The literature offers two approaches: one involves Taylor series expansion \citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
@@
 \citet{Uieda2016} developed a forward gravity model based on tesseroids with homogeneous densities using the GLQ approximation method.
-It has been originally implemented in command line programs written in C programming language.
-Although, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.
+It has been originally implemented through command line programs written in C programming language.
+However, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.
 
@@
 
-It's well known that the GLQ method becomes less accurate when the computation point is closer to the tesseroid \citep{Ku1977} or for smaller GLQ order. 
+\citet{Ku1977} documented that the GLQ becomes less accurate when the computation point is closer to the tesseroid or for smaller GLQ order. 
 Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptive discretization of \citet{Li2011}.
-It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio is greater than a predefined value ($D$) that controls the accuracy of the computation. 
+It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio is lower than a predefined value ($D$) that controls the accuracy of the computation. 
 \citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
-The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} consist in:
-(1) a faster calculation of the distance from the computation point to the tesseroid and 
-(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
 
 We have enhanced the later implementation with a new forward gravity model that allows us to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
-It's written in Python and Cython languages, obtaining an easy to use library that runs precompiled code for the more time consuming functions.
+It's based on the GLQ approximation, makes use of the modified distance-size discretization algorithm and is written in Python and Cython languages, obtaining an easy to use library that runs precompiled code for the more time consuming functions.
 
-In order to ensure the accuracy of the numerical approximation, we have compared its results with a spherical shell with linear and exponential density functions, similar to the test performed by \cite{Uieda2016}.
-Although in this case, the analytical expressions for both density functions had to be derived.
+In order to ensure the accuracy of the numerical approximation, we have conducted similar tests as the one performed by \citet{Uieda2016} by comparing the results of the numerical model with the analytical solutions for a spherical shell with linear, exponential and discontinuous density.
+Furthermore we had to derive the analytical expressions for these density functions.
 
-It also makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
+This new forward gravity model makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
 Furthermore, we have the intention to include it into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
@@
 
+The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} can be summed up in:
+(1) a faster calculation of the distance from the computation point to the tesseroid and 
+(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
+
 In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided and therefore both the accuracy of the algorithm and its computation time.

### 2017-08-23 manuscript/manuscript.tex
Changes in manuscript

@@
 We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented Cartesian coordinate system located at $P$.
-In the special case of homogeneous density, the gravity potential, gradient and Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[see also][]{Uieda2016}.
+In the special case of homogeneous density, the gravity potential, its gradient and the Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[see also][]{Uieda2016}.
 
-In our case, we will suppose that the tesseroid has an arbitrary variable density in depth, i.e. only depends on the radius spherical coordinate.
+In our case, we will assume that the tesseroid has an arbitrary variable density in depth, i.e. only depends on the radius spherical coordinate.
 Thus, the integrals for the gravity fields are slightly modified:
@@
 
-For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$, respectively.
+For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$.
 
@@
 
-In order to test the accuracy of the GLQ approximation, \citet{Uieda2016} and \citet{Grombein2013} have compared the computed gravity fields with the ones generated by a spherical shell, the exceptional case with analytical solution \citep{Mikuska2006,Grombein2013}.
+In order to test the accuracy of the GLQ approximation, \citet{Uieda2016} and \citet{Grombein2013} have compared the numerical model with the analytical solution of the gravity fields generated by a spherical shell with homogeneous density, the exceptional case with analytical solution \citep{Mikuska2006,Grombein2013}.
 
 In case of variable density tesseroids, the comparison must obviously be made with a variable density spherical shell, whose analytical solution will depend on the chosen density function.
-Because the literature has not offered it yet, we needed to develop it ourselves.
+Because the literature has not offered it yet, we needed to develop it ourselves. \todo{ no se si ponerlo asi}
 
-Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, which density depends on the radius spherical coordinate (Figure \ref{fig:spherical-shell}).
+Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, whose density depends on the radius spherical coordinate (Figure \ref{fig:spherical-shell}).
 The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$, located along the $z$ axe at a distance $r$ from the origin, can be written as follows:
@@
 
-\noindent where $\ell$ is defined as in equation \ref{eq:ell}.
+\noindent where $\ell$ is defined in equation \ref{eq:ell}.
 The computation point $Q$ is located at latitude $\phi=90^\circ$, so $\ell$ and $\cos\psi$ (defined in equation \ref{eq:cospsi}) get simplified:

### 2017-08-24 manuscript/manuscript.tex
Minor changes in manuscript

@@
 
-In order to test the accuracy of the GLQ approximation, \citet{Uieda2016} and \citet{Grombein2013} have compared the numerical model with the analytical solution of the gravity fields generated by a spherical shell with homogeneous density, the exceptional case with analytical solution \citep{Mikuska2006,Grombein2013}.
+The integrals that define the gravity fields in equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have analytical solutions for the exceptional case of a spherical shell with homogeneous density \citep{Mikuska2006,Grombein2013}.
+\citet{Uieda2016} and \citet{Grombein2013} compared their numerical models (tesseroids with homogeneous density) with these analytical solutions in order to test the accuracy of their approximations.
 
 In case of variable density tesseroids, the comparison must obviously be made with a variable density spherical shell, whose analytical solution will depend on the chosen density function.
-Because the literature has not offered it yet, we needed to develop it ourselves. \todo{ no se si ponerlo asi}
+Because the literature has not offered it yet, we had to develop it ourselves. \todo{ no se si ponerlo asi}
 

### 2017-08-24 code/tesseroid_density/tesseroid.py
Changed default value of D for gradient components to 2 in tesseroid.py

@@ -102,7 +102,7 @@ due.cite(Doi("10.1190/geo2015-0204.1"),
          description='Gravity forward modeling with tesseroids',
          path="fatiando.gravmag.tesseroid")
 RATIO_V = 1
-RATIO_G = 1.6
+RATIO_G = 2
 RATIO_GG = 8
 STACK_SIZE = 100
 

### 2017-08-25 manuscript/manuscript.tex
Fixed mistake with a plus sign in \ell
Added disable option in new line of todonotes

@@
 \usepackage{todonotes} % to make annotations on margins
+%~ \usepackage[disable]{todonotes} % to disable annotations on margins
 
@@
 The integrals that define the gravity fields in equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have analytical solutions for the exceptional case of a spherical shell with homogeneous density \citep{Mikuska2006,Grombein2013}.
-\citet{Uieda2016} and \citet{Grombein2013} compared their numerical models (tesseroids with homogeneous density) with these analytical solutions in order to test the accuracy of their approximations.
+In order to test the accuracy of their approximations, \citet{Uieda2016} and \citet{Grombein2013} compared their numerical models (tesseroids with homogeneous density) with these analytical solutions.
 
@@
     \cos\psi = \sin\phi', \quad
-    \ell = \sqrt{r'^2 + 2 r r' \sin\phi' + r^2}.
+    \ell = \sqrt{r'^2 - 2 r r' \sin\phi' + r^2}.
 \end{equation}
@@
     \int\limits_{R_1}^{R_2}
-    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + 2 r r' \sin\phi' + r^2}}
+    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 - 2 r r' \sin\phi' + r^2}}
     \, dr' d\phi',

### 2017-08-25 manuscript/manuscript.tex
Changes in manuscript

@@
 The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity field generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axe.
-Due to the rotational symmetry along any axe that passes through the centre of the shell, the potential in equation \ref{eq:shell-pot} reassembles the gravity field on any outside point at distance $r$ from the its centre.
+Due to the rotational symmetry along any axe that passes through the centre of the shell, the integral in equation \ref{eq:shell-pot} reproduces the gravity potential field on any outside point at distance $r$ from the its centre.
 
-For any density $\rho(r')$, the gradient and Marussi tensor components can be obtained as \citet{Grombein2013} does.
-Only the vertical gradient component ($g_z$) and the diagonal Marussi tensor components ($g_{xx}$, $g_{yy}$, $g_{zz}$) are non zero, and can be easily determined from the potential $V_\text{sh}$:
-
-\todo{corregir esto y explicar mejor}
+The gradient and the Marussi tensor derived from potentials that depends solely on $r$ has only a few non zero components: the vertical component of the gradient ($g_z$) and the diagonal components of the tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
+They can be easily obtained as \citet{Grombein2013} does, even for any density function $\rho(r')$ of our choice:
 
@@
 
-For practical purposes we are going to obtain the gravity potential for three different functions in the perspective of their variation in $r'$: a linear, an exponential and a discontinuous density function.
+For practical purposes we are going to obtain the gravity potential for three different functions from the perspective of their variation in $r'$: a linear and an exponential density function.
 
@@
 
-The second term on equation \ref{eq:shell-pot-linear} reassembles the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
-So, this potential can be read as the combination of a spherical shell with the later homogeneous density and the same shell with variable density $\rho(r') = ar'$.
+The second term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
+So it can be read as the combination of a spherical shell with the later homogeneous density and the same shell with variable density $\rho(r') = ar'$.
 
@@
 Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows to generate efficient C code through the Cython compiler.
-This produces a precompiled library whose functions and classes can be called from regular Python code, what allowed us to make it compatible with the preexisting open-source library Fatiando a Terra \citep{Uieda2013}.
+This produces a precompiled library whose functions can be called from regular Python code, what allowed us to make it compatible with the preexisting open-source library Fatiando a Terra \citep{Uieda2013}.
 
@@
 Instead of increasing the GLQ orders to improve the accuracy, what leads to a more time consuming computation, adaptive discretization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
-They essentially consist in dividing the tesseroid based on a ratio between its size and the distance between its geometric centre and the computation point.
+They essentially consist in dividing the tesseroid based on a ratio between the distance from its geometric centre to the computation point and its dimensions.
 
-In order to make the computation more efficient, we use only second order GLQ for each integration, and we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
+In order to make the computation more efficient, we use only second order GLQ for each integration while we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
 Given an arbitrary tesseroid, \citet{Uieda2016} redefined its dimensions: $L_\lambda$, $L_\phi$ and $L_r$, where the former ones are the arc-distances measured along its top surface and $L_r$ is the difference between the radii of the top and bottom surfaces.

### 2017-08-25 manuscript/manuscript.tex
Modifications in manuscript

@@
 In order to make the computation more efficient, we use only second order GLQ for each integration while we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
-Given an arbitrary tesseroid, \citet{Uieda2016} redefined its dimensions: $L_\lambda$, $L_\phi$ and $L_r$, where the former ones are the arc-distances measured along its top surface and $L_r$ is the difference between the radii of the top and bottom surfaces.
-We must check that the following inequality is satisfied:
+The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} can be summed up in:
+(1) a faster calculation of the distance from the computation point to the tesseroid and 
+(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
+
+More specifically, \citet{Uieda2016} redefined the dimensions of an arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$.
+The former ones are the arc-distances measured along its top surface while $L_r$ is the difference between the radii of the top and bottom surfaces.
+
+The adaptive discretization algorithm starts by checking that the following inequality is satisfied:
 
@@
 A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
+On the contrary, if the inequality is satisfied, the algorithm computes the desired gravity field using the GLQ approximation.
 
-The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} can be summed up in:
-(1) a faster calculation of the distance from the computation point to the tesseroid and 
-(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
-
-In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided and therefore both the accuracy of the algorithm and its computation time.
-In this way, the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
+In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided, and therefore regulates both the accuracy of the algorithm and its computation time.
+In this way the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
 less than 10 cm), in which case it does not modifies significantly the desired gravitational field.
 
-We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous and variable density at different heights (see Figure \ref{fig:speed-comparison}).
-Firstly, we can see that both codes take more time to compute the gravity field for lower heights due to the increased divisions of the tesseroid in such cases.
-Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights, however the computation times gets more similar values at greater heights.
-In conclusion, the new variable dnesity code is fast enough to perform the calculations of a real model in a few minutes.
-\todo{real, esta bien la expresion?}
+On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy is a matter of study.
 
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
-\caption{Computation times for a single tesseroid model with homogeneous and variable density at different heights.}
-\label{fig:speed-comparison}
-\end{figure}
-
-On the other hand, the accuracy of the method is exclusively controlled by the value assigned to the distance-size ratio $D$, which cannot be easily related to the error of the approximation.
 In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
-
 We have performed similar tests, but now with variable density spherical shell.
@@
 crust and mantle interface).
-The analytical solutions for the first two types of functions were previously obtained and shown in equations \ref{eq:shell-pot-linear} and \ref{eq:shell-pot-exp}.
 
@@
 
+\section{Speed Comparison}
+
+We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous and variable density at different heights (see Figure \ref{fig:speed-comparison}).
+Firstly, we can see that both codes take more time to compute the gravity field for lower heights due to the increased divisions of the tesseroid in such cases.
+Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights, however the computation times gets more similar values at greater heights.
+In conclusion, the new variable dnesity code is fast enough to perform the calculations of a real model in a few minutes.
+\todo{real, esta bien la expresion?}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
+\caption{Computation times for a single tesseroid model with homogeneous and variable density at different heights.}
+\label{fig:speed-comparison}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conclusions}

### 2017-08-25 manuscript/manuscript.tex
Modifications on manuscript

@@
 From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
-The first one constitutes the simplest variation, while the second one represents the fastest varying function.
-The later one tries to emulate a contact between two different media (e.g.
-crust and mantle interface).
 
@@
 
-In order to test the new variable density tesseroid code and estimate the error of the approximation we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
+In order to test the new variable density tesseroid code and estimate the error of the approximation, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
+The first one constitutes the simplest variation, while the second one represents the fastest varying function.
+The later one tries to emulate a contact between two different media (e.g.
+crust and mantle interface).
+
 This error estimations are also useful in the determination of the minimum value for the distance-size ratio $D$ such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
@@
 
-We computed the three gravity fields for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, on every point of these four grids.
+We computed the three gravity fields  on every point of these four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
 Then we calculated the maximum difference between these results and the value of the analytical solution at the same height.
@@
 Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
-
-The tests were only performed on these magnitudes because the other components of the gravity gradient and Marussi tensor are equal to zero due to the rotational symmetries of the spherical shell.
 
@@
 
-As a simple case, we proposed a linear density function in such a way that the spherical shell has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:
+In first place, we proposed a spherical shell with linear density in such way that it has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:
 
@@
     \rho(r') = ar' + c,
-    %~ -\frac{3300 - 2670}{35000} r' + 
-    %~ \left( \frac{3300 - 2670}{35000} R + 2670 \right)
     \label{eq:density-linear}
@@
 
-The tests were performed on a thin and a thick spherical shell of 1 km and 35 km of thickness, respectively.
+Then we computed the difference between the numerical model with its corresponding analytical solution (equation \ref{eq:density-linear}) for a thick and a thin spherical shell of 1km and 35km of thickness, respectively.
+
 In Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick} we can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor.
@@
 \noindent where $A$ = 1kg/m$^3$ and $b$ = 1000km.
-We have also performed this test with a thin and a thick shell of 1 km and 35 km, respectively, and the results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, whose results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
 Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them, although, they can be found in the repository.
@@
 
-Because the constant $b$ controls the variation of the density function we wanted to test how does the accuracy of the method behaves at different values of $b$.
-We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}).
+Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$.
+We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the analytical solution.
 For each value of $b$ we calculated these difference for the same set of values of $D$ we've been using on the previous tests, obtaining one curve for each kind of variation.
 
-This tests were performed using both the thin and the thick spherical shell, and the results can be seen in Figure \ref{fig:D-exp-power}.
+This tests were performed using both the thin and the thick spherical shell, whose results can be seen in Figure \ref{fig:D-exp-power}.
 Taking into account that lower values of $b$ implies a more varying function, we can see that for extremely varying functions the numerical model does not perform well: for $b$=100m the difference is almost 100\% in the thin shell case, and the same happens with $b$=10km in the thick shell case.
@@
 We can see its results in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what
-leads us to conclude that the code does not reassembles the gravity fields for this case with good accuracy.
+leads us to conclude that the code does not recreates the gravity fields for this case with good accuracy.
 
@@
 Firstly, we can see that both codes take more time to compute the gravity field for lower heights due to the increased divisions of the tesseroid in such cases.
-Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights, however the computation times gets more similar values at greater heights.
-In conclusion, the new variable dnesity code is fast enough to perform the calculations of a real model in a few minutes.
+Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights.
+However, the computation times gets more similar values at greater heights.
+In conclusion, the new variable density code is fast enough to perform the calculations of a real model in a few minutes.
 \todo{real, esta bien la expresion?}

### 2017-08-25 manuscript/manuscript.tex
Changes in manuscript

@@
 It essentially consists in approximating the gravity fields of the tesseroids with the effect of point masses located on the scaled nodes of the Legendre polynomials.
-We state that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involve Taylor series expansion.
+We have stated that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involve Taylor series expansion.
 
@@
 
-On the other hand, there's no direct relation between the value of $D$ and the error of the computation, that's why we had to determine the minimum value of $D$ for each gravity magnitude that produces an acceptable accuracy.
+On the other hand, there's no direct relation between the value of $D$ and the error of the computation, that's why we had to determine the minimum value of $D$ that produces an acceptable accuracy for each gravity magnitude.
 These distance-size ratio determinations have been made through the comparison of the numerical approximation with the analytical solutions for the exceptional case of a spherical shell.
-From the wide range of possible density variations, we've chosen a (1) linear, (2) an exponential and (3) a discontinuous density functions.
+From the wide range of possible density variations, we've chosen a linear, an exponential and a discontinuous density functions.
 
@@
 On the other hand, the tests over the discontinuous function were not satisfactory.
-In such cases we recommend to divide the model in several tesseroids, each one with a specific homogeneous density or a continuous density function, instead of using one variable density tesseroid.
+In such cases we recommend to divide the model in several tesseroids, each one with a specific homogeneous density or a continuous density function, instead of using a single variable density tesseroid.
 

### 2017-09-14 code/tesseroid_relief.py
Started writing sedimentary basin forward calculation

@@ -1,0 +1,174 @@
+# -*- coding: utf-8 -*-
+"""
+Implements the proposed Moho inversion method.
+
+Defines the classes:
+
+* `TesseroidRelief`: Class to represent a relief undulating around a reference
+  level and discretized into tesseroids.
+
+"""
+from __future__ import division
+from fatiando import gridder
+from fatiando.mesher import Tesseroid
+import numpy as np
+import copy
+
+
+class TesseroidRelief(object):
+    """
+    Implements a tesseroid model of an interface undulating around a reference
+    level.
+
+    This class behaves like a sequence of 'Tesseroid' objects, so you can pass
+    it along to any function that iterates over tesseroids (like the forward
+    modeling function of 'fatiando.gravmag.tesseroid').
+
+    Parameters:
+
+    * area : [s, n, w, e]
+        The south, north, west, and east limits of the mesh in degrees.
+    * shape : (nlat, nlon)
+        The number of tesseroids in the latitude and longitude directions. Will
+        discretize *area* into this number of tesseroids.
+    * relief : 1d-array
+        The height-coordinates of the relief
+    * reference : float
+        The height-coordinate of the reference level.
+    * props : dict
+        Dictionary with the physical properties of the mesh.
+
+    """
+
+    def __init__(self, area, shape, relief, reference, props=None):
+        assert shape[0]*shape[1] == relief.size
+        assert len(area) == 4
+        assert area[0] < area[1] and area[2] < area[3]
+        self.area = area
+        self.shape = shape
+        s, n, w, e = area
+        nlat, nlon = shape
+        self.lons = np.linspace(w, e, nlon, endpoint=False)
+        self.lats = np.linspace(s, n, nlat, endpoint=False)
+        self.lon, self.lat = np.meshgrid(self.lons, self.lats)
+        self.spacing = self.lats[1] - self.lats[0], self.lons[1] - self.lons[0]
+        self._relief = relief
+        self._reference = reference*np.ones_like(relief)
+        self.set_top_bottom()
+        if props is None:
+            self.props = {}
+        else:
+            self.props = props
+        self._i = 0
+
+    @property
+    def clons(self):
+        dlon = self.spacing[1]
+        return self.lons + dlon/2
+
+    @property
+    def clats(self):
+        dlat = self.spacing[0]
+        return self.lats + dlat/2
+
+    @property
+    def clon(self):
+        dlon = self.spacing[1]
+        return self.lon + dlon/2
+
+    @property
+    def clat(self):
+        dlat = self.spacing[0]
[…]

### 2017-09-15 manuscript/manuscript.tex
Temporaly disabled todonotes
Modified speed comparison: absolute times replaced by computation times ratio

@@
 \usepackage{graphicx}
-\usepackage{todonotes} % to make annotations on margins
-%~ \usepackage[disable]{todonotes} % to disable annotations on margins
+%~ \usepackage{todonotes} % to make annotations on margins
+\usepackage[disable]{todonotes} % to disable annotations on margins
 
@@
 
-We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous and variable density at different heights (see Figure \ref{fig:speed-comparison}).
-Firstly, we can see that both codes take more time to compute the gravity field for lower heights due to the increased divisions of the tesseroid in such cases.
-Secondly, we can see that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights.
-However, the computation times gets more similar values at greater heights.
-In conclusion, the new variable density code is fast enough to perform the calculations of a real model in a few minutes.
+We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous ($\Delta t_\text{homogeneous}$) and variable ($\Delta
+t_\text{variable}$) density at different heights. Then we plotted the ratio between these computation times (see Figure \ref{fig:speed-comparison}):
+
+\begin{equation}
+    \text{Computation Times Ratio} =
+        \frac{\Delta t_\text{variable}}{\Delta t_\text{homogeneous}}
+    \label{eq:computation-times-ratio}
+\end{equation}
+
+From these results We can conclude that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights.
+However, the computation times ratio gets lower at greater heights, due to the reduction of the divisions made by the adaptive discretization algorithm.
+In summary, the new variable density code is fast enough to perform the calculations of a real model in a few minutes.
 \todo{real, esta bien la expresion?}
@@
 \includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
-\caption{Computation times for a single tesseroid model with homogeneous and variable density at different heights.}
+\caption{Computation times ratio between a single tesseroid model with variable and with homogeneous density at different heights, defined in equation \ref{eq:computation-times-ratio}.}
 \label{fig:speed-comparison}

### 2017-09-19 manuscript/manuscript.tex
Modification in manuscript: added new exponential variation with b figures

@@
 Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$.
-We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the analytical solution.
-For each value of $b$ we calculated these difference for the same set of values of $D$ we've been using on the previous tests, obtaining one curve for each kind of variation.
+In order to do this we proposed a set of density functions, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900 kg/m$^{3}$ at the bottom of the spherical shell.
+We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the respective analytical solution.
+These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.
 
-This tests were performed using both the thin and the thick spherical shell, whose results can be seen in Figure \ref{fig:D-exp-power}.
+This tests were performed using both the thin and the thick spherical shell.
+On Figures \ref{fig:D-exp-power-thin}a and \ref{fig:D-exp-power-thick}a we can see the density functions used to perform these tests for each spherical shell model, respectively, and their results can be seen in \ref{fig:D-exp-power-thin}b and \ref{fig:D-exp-power-thick}b.
 Taking into account that lower values of $b$ implies a more varying function, we can see that for extremely varying functions the numerical model does not perform well: for $b$=100m the difference is almost 100\% in the thin shell case, and the same happens with $b$=10km in the thick shell case.
 
-On both cases, the model performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, in other words, when the absolute value of the exponent in equation \ref{eq:density-exp} is lower than 1.
+On both cases, the model performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, in other words, when the absolute value of the exponent in equation \ref{eq:density-exp} is lower than one.
 Although this sets a limitation to the model, the maximum acceptable variation is very far from real density distributions inside the Earth.
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences.pdf}
+\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences-thin.pdf}
 \caption{
-    Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for (a) a thin spherical shell of 1km of thickness and (b) a thick sphericall shell of 35km of thickness, both with the exponential density defined on equation \ref{eq:density-exp} with different values of $b$.
+    Results of the accuracy tests for different exponential density functions in case of a thin spherical shell (1km of thickness).
+    (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
+    (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
 Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
-Each curve corresponds to a value of $b$.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
 }
-\label{fig:D-exp-power}
+\label{fig:D-exp-power-thin}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences-thick.pdf}
+\caption{
+    Results of the accuracy tests for different exponential density functions in case of a thick spherical shell (35km of thickness).
+    (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
+    (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
+Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
+If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+}
+\label{fig:D-exp-power-thick}
 \end{figure}

### 2017-09-26 manuscript/figures/to_pdf.sh
Reorganized figures: all in manuscript/figures in pdf

@@ -1,5 +1,0 @@
-#!/bin/bash
-
-for file in ./*.svg; do
-    inkscape --export-pdf ${file%.svg}.pdf ./$file
-done

### 2017-09-26 manuscript/manuscript.tex
Fixed typo
Started writing "Example with real data"

@@
 }
-\author[S.R. Soler, L. Uieda and M.E. Gimenez]{
-    Santiado R. Soler$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
+\author[S.R. Soler, A. Pesce, L. Uieda and M.E. Gimenez]{
+    Santiado R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
     $^1$CONICET, Argentina. e-mail: santiago.r.soler@gmail.com\\
@@
 
-From these results We can conclude that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights.
+From these results we can conclude that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights.
 However, the computation times ratio gets lower at greater heights, due to the reduction of the divisions made by the adaptive discretization algorithm.
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Example with real data}
+
+Besides all the tests we have carried out on the new code, we intend to show a forward modelling example with real data.
+In order to do that we have chosen the Neuquen Basin: a sedimentary basin located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude (see Figure \ref{fig:neuquen-basin}a), that includes continental and marine siliciclastics, carbonates and evaporites accumulated over the Jurassic and the Cretaceous constituting a stratigraphic record up to 5000m of depth \citep{Howell2005}.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
+\caption{(a) Topography of Neuquen Basin (in km),
+         (b) Thickness of the sedimentary basin (in meters),
+         (c)-(h) Computed gravity fields: potential $V$ (in J/kg), gradient components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components $g_{xy}$ and $g_{zz}$ (in Eötvös), calculated at 50km of height over the ellipsoid.}
+\label{fig:neuquen-basin}
+\end{figure*}
+
+The thickness of the sediment pack was obtained from \citet{Heine2007}, which can be seen in Figure \ref{fig:neuquen-basin}.
 

### 2017-09-27 manuscript/manuscript.tex
Continue writing real data example

--- 2017-09-19 manuscript/manuscript.tex
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Example with real data}
+
+Besides all the tests we have carried out on the new code, we intend to show a forward modelling example with real data.
+In order to do that we have chosen the Neuqu\'en Basin: a sedimentary basin located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude (see Figure \ref{fig:neuquen-basin}a), that includes continental and marine siliciclastics, carbonates and evaporites accumulated over the Jurassic and the Cretaceous constituting a stratigraphic record up to 5000m of depth \citep{Howell2005}.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
+\caption{(a) Topography of Neuqu\'en Basin (in km),
+         (b) Thickness of the sedimentary basin (in meters),
+         (c)-(h) Computed gravity fields: potential $V$ (in J/kg), gradient components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components $g_{xy}$ and $g_{zz}$ (in Eötvös), calculated at 50km of height over the ellipsoid.}
+\label{fig:neuquen-basin}
+\end{figure*}
+
+The thickness of the sediment pack was obtained from \citet{Heine2007}.
+We managed to produce a regular grid with a resolution of 0.05$^\circ$ on both longitude and latitude directions with one value of thickness on each node (see Figure \ref{fig:neuquen-basin}).
+
+In order to compute the gravity fields generated by this basin we created a mesh of tesseroids, each one located on one of the nodes of the grid, with dimensions of 0.05$^\circ$ $\times$ 0.05$^\circ$ and depth equal to the thickness of the basin on the corresponding node.
+
+We must also define a density function for the entire model.
+\citet{Sigismondi2012} proposed a minimum and a maximum density for the Neuqu\'en basin of 275kg/m$^3$ and 412kg/m$^3$, respectively.
+So we have chosen, for this specific example, a linear density that assumes the minimum value on the surface and the maximum at 5858m, i.e. the bottom of the basin.
+
+We have finally computed the gravity potential $V$, the gradient components $g_x$, $g_y$ and $g_z$, and the Marussi tensor components $g_{xy}$ and $g_{zz}$ on a computation grid of 159$\times$163 nodes with a 0.05$^\circ$ spacing on both longitude and latitude directions at a 50km height over the reference ellipsoid.
+The resulting gravity fields can be seen in Figures \ref{fig:neuquen-basin}c-h.
+We have excluded the other components of the Marussi tensor from the Figure, although they have also been computed and can be found in the repository.
 
@@
 
+Finally, we have shown a real data example: we have computed the gravity fields generated by the Neuqu\'en sedimentary basin through a tesseroid model with linear increasing density with depth on a regular grid at 50km height.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-09-27 jovenes-investigadores/jovenes-investigadores.tex
Started writing breef work for jovenes investigadores

--- 2017-09-19 manuscript/manuscript.tex
@@
-\documentclass[extra]{gji}
-%~ \documentclass[extra, referee]{gji}
-
+\documentclass[a4paper,10pt]{article}
 \usepackage[utf8]{inputenc}
-\usepackage{timet}
+\usepackage[spanish]{babel}
+\usepackage[hmargin=2.5cm, vmargin=2.5cm]{geometry}
 \usepackage{amsmath}
 \usepackage{graphicx}
-%~ \usepackage{todonotes} % to make annotations on margins
-\usepackage[disable]{todonotes} % to disable annotations on margins
-
+\usepackage{natbib}
 \usepackage{url}
@@
 \hypersetup{
-    allcolors=blue,
+    allcolors=black,
 }
+\usepackage{mathptmx} %Times font
+
+\title{
+    \textbf{
+    TESSEROIDES CON DENSIDADES VARIABLES:
+    CÁLCULO DE CAMPOS GRAVITATORIOS EN COORDENADAS ESFÉRICAS CON
+    DENSIDADES VARIABLES EN PROFUNDIDAD
+    }
+}
+\author{
+    Santiago R. Soler$^{1,2}$ \vspace{0.5em} \\ 
+    \textit{$^1$ Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan} \\
+    \textit{$^2$ Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET)} \vspace{0.4em} \\
+    email: santiago.r.soler@gmail.com
+}
+\date{}
 
@@
 
-\title[Variable Density Tesseroids]{
-    Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth
-}
-\author[S.R. Soler, L. Uieda and M.E. Gimenez]{
-    Santiado R. Soler$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
-    $^1$CONICET, Argentina. e-mail: santiago.r.soler@gmail.com\\
-    $^2$Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan, Argentina\\
-    $^3$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil
-    }
-
 \maketitle
 
-\begin{summary}
-Summary of this paper 
-\end{summary}
+\vspace{-2.5em}
+\begin{center}
+\textbf{Palabras Clave:} poner algo
+\end{center}
+\vspace{0.5em}
 
-\begin{keywords}
-Satellite gravity??, Numerical modelling, Structure of the Earth, Gravity anomalies and Earth structure, Numerical approximations and analysis, spatial analysis??
-\end{keywords}
+\begin{abstract}
+El cálculo directo de campos gravitatorios para grandes extensiones requiere tomar en cuenta la curvatura de la Tierra, lo cual suele resolverse a través del uso de prismas esféricos llamados Tesseroides.
+Además, estructuras como cuencas sedimentarias presentan variaciones de densidad con la profundidad.
 
-\section{Introduction}
+\end{abstract}
 
-Lithosphere's density variation with depth has been studied for almost a century. 
-On the first years \citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples, while in the following decades other density functions have been proposed for different rock types \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
-Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
+\section{Introducción}
+
+Las variaciones de la densidad en la corteza con la profundidad han sido estudiadas por casi un siglo: \citet{Athy1930} obtuvo variaciones exponenciales, mientras que otras funciones fueron propuestas en los siguientes años por \citet{Maxant1980, Rao1986, Rao1993, Rao1994}.
+También se ha realizado modelado directo o modelos de inversión gravimétricos con densidades variables, principalmente aplicados en cuencas y fosas \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
+
+Estos modelos directos fueron desarrollados para cuerpos bidimensionales o tridimensionales en coordenadas cartesianas, los cuales se ajustan adecuadamente para aplicaciones de escalas locales.
+Sin embargo, la disponibilidad de datos gravimétricos satelitales permite llevar a cabo estudios e interpretaciones a escalas regionales, en las cuales es necesario tomar en cuenta la curvatura de la Tierra a la hora de realizar modelados directos.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/tesseroid-uieda.pdf}
-\caption{
-A spherical prism, also known as  Tesseroid, in a geocentric spherical coordinate system, with a computation point $P$ and its local north oriented Cartesian coordinate system.
-Picture made by \citet{Uieda2015}.
-}
-\label{fig:tesseroid-uieda}
-\end{figure}
-
-\citet{Uieda2016} developed a forward gravity model based on tesseroids with homogeneous densities using the GLQ approximation method.
-It has been originally implemented through command line programs written in C programming language.
-However, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.
-
-In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
-On the other hand, the GLQ allows us to calculate them for any density function without any other information but the function itself.
-
-\citet{Ku1977} documented that the GLQ becomes less accurate when the computation point is closer to the tesseroid or for smaller GLQ order. 
-Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptive discretization of \citet{Li2011}.
-It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio is lower than a predefined value ($D$) that controls the accuracy of the computation. 
-\citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
-
-We have enhanced the later implementation with a new forward gravity model that allows us to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
-It's based on the GLQ approximation, makes use of the modified distance-size discretization algorithm and is written in Python and Cython languages, obtaining an easy to use library that runs precompiled code for the more time consuming functions.
-
-In order to ensure the accuracy of the numerical approximation, we have conducted similar tests as the one performed by \citet{Uieda2016} by comparing the results of the numerical model with the analytical solutions for a spherical shell with linear, exponential and discontinuous density.
-Furthermore we had to derive the analytical expressions for these density functions.
-
-This new forward gravity model makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
-Furthermore, we have the intention to include it into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
-
-In the following sections we will describe how the new algorithm works, its theoretical background and obtain the distance-size ratio values needed to get a good accuracy in the computation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Theory}
-
-The spherical prisms known as tesseroids are mass elements defined in a geocentric spherical coordinate system bounded by a pair of parallels, a pair of  meridians, and two concentric spherical surfaces.
-We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented Cartesian coordinate system located at $P$.
-In the special case of homogeneous density, the gravity potential, its gradient and the Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[see also][]{Uieda2016}.
-
-In our case, we will assume that the tesseroid has an arbitrary variable density in depth, i.e. only depends on the radius spherical coordinate.
-Thus, the integrals for the gravity fields are slightly modified:
-
-\begin{equation}
-    V(r,\phi,\lambda) = G
-    \int\limits_{\lambda_1}^{\lambda_2}
-    \int\limits_{\phi_1}^{\phi_2}
-    \int\limits_{r_1}^{r_2}
-    \frac{\rho(r')}{\ell} \kappa \,  dr' d\phi' d\lambda',
-\label{eq:tesseroid-pot}
-\end{equation}
-\begin{equation}
-    g_{\alpha}(r,\phi,\lambda) = G
-    \int\limits_{\lambda_1}^{\lambda_2}
-    \int\limits_{\phi_1}^{\phi_2}
-    \int\limits_{r_1}^{r_2}
-    \rho(r') \frac{\Delta_\alpha}{\ell^3}
-    \kappa \, dr' d\phi' d\lambda',
-\label{eq:tesseroid-grav}
-\end{equation}
-\begin{equation}
-    g_{\alpha\beta}(r,\phi,\lambda) = G
-    \int\limits_{\lambda_1}^{\lambda_2}
-    \int\limits_{\phi_1}^{\phi_2}
-    \int\limits_{r_1}^{r_2}
-    \rho(r') I_{\alpha\beta} \, \kappa \, dr' d\phi' d\lambda' ,
-    \label{eq:tesseroid-tensor}
-\end{equation}
-
-\noindent where
-
-\begin{equation}
-    I_{\alpha\beta} =
-    \left(
-        \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
-        \frac{\delta_{\alpha\beta}}{\ell^3}
-    \right) ,
-    \label{eq:tesseroid-tensor-kernel}
-\end{equation}
-
-\noindent $\alpha, \beta \in \{x, y, z\}$,
-$\rho(r')$ is the density function that depends on the radius coordinate,
-$\delta_{\alpha\beta}$ is Kronecker's delta,
-$G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant and
-
-\begin{equation}
-    \Delta_x = r'[\cos\phi\sin\phi' - \sin\phi\cos\phi'
-               \cos(\lambda' - \lambda)],
-\end{equation}
-\begin{equation}
-    \Delta_y = r' \cos \phi' \sin(\lambda' - \lambda),
-\end{equation}
-\begin{equation}
-    \Delta_z = r' \cos \psi - r,
-\end{equation}
-\begin{equation}
-    \kappa = {r'}^2 \cos \phi',
-\end{equation}
-\begin{equation}
-    \ell = \sqrt{{r'}^2 + r^2 - 2 r' r \cos \psi},
-\label{eq:ell}
-\end{equation}
-\begin{equation}
-    \cos\psi = \sin\phi\sin\phi' + \cos\phi\cos\phi'
-                 \cos(\lambda' - \lambda).
-\label{eq:cospsi}
-\end{equation}
-
-According to \citet[p.~390]{Hildebrand1987}, we can approximate any integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e.
-by a weighted sum of the integration kernel evaluated on the roots of the $N$th Legendre polynomial:
-
-\begin{equation}
-    \int\limits_{-1}^1 f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i),
-\end{equation}
-
-\noindent where $x_i$ are the roots of the $N$th Legendre polynomial $P_N(x)$ and $W_i$ are the weights of each term:
-
-\begin{equation}
-    W_i = \frac{2}{(1-x_i^2)[P_N^\prime(x_i)]^2},
-\end{equation}
-
-\begin{equation}
-    P_N(x_i) = 0, \quad \forall i = {1,\dots,N}.
-\end{equation}
-
-In case of an integration defined on a arbitrary interval, e.g. $[a,b]$, we can scale the nodes and perform a similar approximation:
-
-\begin{equation}
-    \int\limits_a^b f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i^s),
-\label{eq:glq-scaled}
-\end{equation}
-
-\noindent where $x_i^s$ is the scaled $x_i$ root in the $[a,b]$ interval, also called nodes of the quadrature:
-
-\begin{equation}
-    x_i^s = \frac{b-a}{2} x_i + \frac{b+a}{2}.
-\end{equation}
-
-For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$.
-
-Our intention is to use GLQ to approximate the volume integrals from equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor}. 
-When tesseroids have variable densities, we can write the integral kernels as the product of a certain function $g$ and the density function:
-
-\begin{equation}
-    f(r', \phi', \lambda') = \rho(r') g(r', \phi', \lambda'),
-\end{equation}
-
-\noindent and then apply the quadrature to every integral corresponding to each spherical coordinate, obtaining a similar expression to the one shown by \citet{Asgharzadeh2007} and \citet{Uieda2016}:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') 
-        d\Omega \approx& \\
-        A 
-        \sum\limits_{i=1}^{N^r}
-        \sum\limits_{j=1}^{N^\phi}
-        \sum\limits_{k=1}^{N^\lambda}
-        & W_i^r W_j^\phi W_k^\lambda \rho(r_i) g(r_i, \phi_j, \lambda_k),
-    \end{split}
-\label{eq:glq-var-dens}
-\end{equation}
-}{
-\begin{equation}
-    \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') d\Omega \approx
-    A 
-    \sum\limits_{i=1}^{N^r}
-    \sum\limits_{j=1}^{N^\phi}
-    \sum\limits_{k=1}^{N^\lambda}
-    W_i^r W_j^\phi W_k^\lambda \rho(r_i) g(r_i, \phi_j, \lambda_k),
-\label{eq:glq-var-dens}
-\end{equation}
-}
-
-\noindent where
-
-\begin{equation}
-    A = 
-    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8}.
-\end{equation}
-
-From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates the gravity fields of a tesseroid with variable density as the ones generated by point masses located at the nodes of the Legendre polynomials.
-Furthermore, the information of the density function is summarised as the values it assumes on those nodes.
-Although this may sound discouraging, we intent to prove otherwise and show that this method, along with a well fitted discretization algorithm, approximates the gravity fields with good accuracy and precision.\todo{no se si poner esta ultima oracion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Analytical Solutions for Spherical Shell}
-
-The integrals that define the gravity fields in equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have analytical solutions for the exceptional case of a spherical shell with homogeneous density \citep{Mikuska2006,Grombein2013}.
-In order to test the accuracy of their approximations, \citet{Uieda2016} and \citet{Grombein2013} compared their numerical models (tesseroids with homogeneous density) with these analytical solutions.
-
-In case of variable density tesseroids, the comparison must obviously be made with a variable density spherical shell, whose analytical solution will depend on the chosen density function.
-Because the literature has not offered it yet, we had to develop it ourselves. \todo{ no se si ponerlo asi}
-
-Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, whose density depends on the radius spherical coordinate (Figure \ref{fig:spherical-shell}).
-The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$, located along the $z$ axe at a distance $r$ from the origin, can be written as follows:
-
-\begin{figure}
-\centering
-\includegraphics[width=0.7\linewidth]{figures/spherical-shell.pdf}
-\caption{
-Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
-The computation point $Q$ is located in the $z$ axe at a distance $r$ from the centre of the shell.
-For practical purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
-}
-\label{fig:spherical-shell}
-\end{figure}
-
-\begin{equation}
-    V_\text{sh}(r) = G 
-    \int\limits_0^{2\pi}
-    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
-    \int\limits_{R_1}^{R_2}
-    \frac{\rho(r')}{\ell} {r'}^2 \cos\phi' \, 
-    dr' d\phi' d\lambda',
-\end{equation}
-
-\noindent where $\ell$ is defined in equation \ref{eq:ell}.
-The computation point $Q$ is located at latitude $\phi=90^\circ$, so $\ell$ and $\cos\psi$ (defined in equation \ref{eq:cospsi}) get simplified:
-
-\begin{equation}
-    \cos\psi = \sin\phi', \quad
-    \ell = \sqrt{r'^2 - 2 r r' \sin\phi' + r^2}.
-\end{equation}
-
-Due to the rotational symmetry along the $z$ axe, the integration in $\lambda'$ is straightforward:
-
-\begin{equation}
-    V_\text{sh}(r) = 2\pi G 
-    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
-    \int\limits_{R_1}^{R_2}
-    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 - 2 r r' \sin\phi' + r^2}}
-    \, dr' d\phi',
-\end{equation}
-
-\noindent and the integration in $\phi'$ can be performed independently of the density function.
-Using \citet{sagemath} we obtain the following expression:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}(r) = 2\pi G
-        \int\limits_{R_1}^{R_2}
-        \Big[ & \sqrt{r^2 + r'^2 + 2rr'} - \\
-        & \sqrt{r^2 + r'^2 - 2rr'}
-        \Big] \frac{r'\rho(r')}{r} \, dr'.
-    \end{split}
-\label{eq:shell-pot-sqrts}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}(r) = 2\pi G
-    \int\limits_{R_1}^{R_2}
-    \Big[ \sqrt{r^2 + r'^2 + 2rr'}  -
-    \sqrt{r^2 + r'^2 - 2rr'} 
-    \Big] \frac{r'\rho(r')}{r} \, dr'.
-\label{eq:shell-pot-sqrts}
-\end{equation}
-}
-
-For our purposes we can assume that the computation point $Q$ is outside the outer radii, i.e. $r>R_2$ ($R_2 \geq r'$), and simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
-
-\begin{equation}
-    \sqrt{r^2 + r'^2 + 2rr'} = |r + r'| = r + r',
-\end{equation}
-\begin{equation}
-    \sqrt{r^2 + r'^2 - 2rr'} = |r - r'| = r - r',
-\end{equation}
-
-\noindent what leads to the following expression of the potential of the spherical shell:
-
-\begin{equation}
-    V_\text{sh}(r) = \frac{4\pi G}{r}
-    \int\limits_{R_1}^{R_2} {r'}^2 \rho(r') \, dr'.
-\label{eq:shell-pot}
-\end{equation}
-
-The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity field generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axe.
-Due to the rotational symmetry along any axe that passes through the centre of the shell, the integral in equation \ref{eq:shell-pot} reproduces the gravity potential field on any outside point at distance $r$ from the its centre.
-
-The gradient and the Marussi tensor derived from potentials that depends solely on $r$ has only a few non zero components: the vertical component of the gradient ($g_z$) and the diagonal components of the tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
-They can be easily obtained as \citet{Grombein2013} does, even for any density function $\rho(r')$ of our choice:
-
-\begin{equation}
-    g_z = \frac{V_\text{sh}(r)}{r},
-\end{equation}
-\begin{equation}
-    g_{xx} = g_{yy} = -\frac{V_\text{sh}(r)}{r^2}, \quad
-    g_{zz} = \frac{2V_\text{sh}(r)}{r^2}.
-\end{equation}
-
-For practical purposes we are going to obtain the gravity potential for three different functions from the perspective of their variation in $r'$: a linear and an exponential density function.
-
-\subsubsection{Linear Density}
-
-If the density of the spherical shell is a linear function on the radius coordinate, i.e.:
-
-\begin{equation}
-    \rho(r') = ar' + b,
-\end{equation}
-
-\noindent then the gravity potential generated on any computation point at distance $r$ from its centre can be obtained from the following expression:
-
-\begin{equation}
-    V_\text{sh}^\text{lin}(r) = \pi G \left[ 
-    a \frac{R_2^4 - R_1^4}{r} +
-    b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
-    \label{eq:shell-pot-linear}
-\end{equation}
-
-The second term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
-So it can be read as the combination of a spherical shell with the later homogeneous density and the same shell with variable density $\rho(r') = ar'$.
-
-\subsubsection{Exponential Density}
-
-If, on the other hand, the spherical shell has an exponential density in the radius coordinate,
-
-\begin{equation}
-    \rho(r') = Ae^{-(r' - \Delta h)/b},
-\end{equation}
-
-\noindent the potential can be written as follows:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-        Ab \, e^\frac{\Delta h}{b}
-        \Big[
-        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-        \Big].
-    \end{split}
-\label{eq:shell-pot-exp}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-    Ab \, e^\frac{\Delta h}{b}
-    \Big[
-    & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-    & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    \Big].
-\label{eq:shell-pot-exp}
-\end{equation}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Implementation}
-
-We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth through the GLQ to numerically approximate the volume integrals.
-It's freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}
-
-Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows to generate efficient C code through the Cython compiler.
-This produces a precompiled library whose functions can be called from regular Python code, what allowed us to make it compatible with the preexisting open-source library Fatiando a Terra \citep{Uieda2013}.
-
-Its usage is very similar to the code for homogeneous tesseroid already in Fatiando a Terra (version 0.5):
-
-\begin{enumerate}
-\renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Define a single tesseroid (or a tesseroid mesh) with the Fatiando a Terra classes.
-    \item Add a predefined Python density function to the tesseroid (or each tesseroid in the mesh).
-    \item Calculate the desired gravity field (potential, gradient or tensor component) in any set of computation points with our new functions.
-\end{enumerate}
-
-The existing code in Fatiando a Terra (v.~0.5) for homogeneous tesseroid is optimised through the Numba compiler.
-Due to difficulties when passing a Python function as argument of precompiled Numba methods, we have decided to rewrite them in Cython language, keeping the homogeneous tesseroid calculation and adding the new variable density code.
-
-From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of $N_\lambda N_\phi N_r$ point masses, where $N_i$ are the orders of the quadrature for the integral on each coordinate, with $i \in \{ \lambda, \phi, r \}$.
-If we set the GLQ orders to fixed values, it's easy to conclude that the nearer the computation point is to the tesseroid, the accuracy of the approximation is reduced and the point masses effect becomes more noticeable than the tesseroid we want to approximate.
-A more complete analysis of this conclusion can be found in \citet{Ku1977, Li2011, Uieda2016}.
-
-Instead of increasing the GLQ orders to improve the accuracy, what leads to a more time consuming computation, adaptive discretization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
-They essentially consist in dividing the tesseroid based on a ratio between the distance from its geometric centre to the computation point and its dimensions.
-
-In order to make the computation more efficient, we use only second order GLQ for each integration while we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
-The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} can be summed up in:
-(1) a faster calculation of the distance from the computation point to the tesseroid and 
-(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
-
-More specifically, \citet{Uieda2016} redefined the dimensions of an arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$.
-The former ones are the arc-distances measured along its top surface while $L_r$ is the difference between the radii of the top and bottom surfaces.
-
-The adaptive discretization algorithm starts by checking that the following inequality is satisfied:
-
-\begin{equation}
-    \frac{d}{L_i} \geq D,
-\label{eq:distance-size-ratio}
-\end{equation}
-
-\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point, and $D$ is a predefined value called distance-size ratio.
-If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.
-A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
-On the contrary, if the inequality is satisfied, the algorithm computes the desired gravity field using the GLQ approximation.
-
-In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided, and therefore regulates both the accuracy of the algorithm and its computation time.
-In this way the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
-less than 10 cm), in which case it does not modifies significantly the desired gravitational field.
-
-On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy is a matter of study.
-
-In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
-We have performed similar tests, but now with variable density spherical shell.
-From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Determination of distance-size ratio}
-
-In order to test the new variable density tesseroid code and estimate the error of the approximation, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
-The first one constitutes the simplest variation, while the second one represents the fastest varying function.
-The later one tries to emulate a contact between two different media (e.g.
-crust and mantle interface).
-
-This error estimations are also useful in the determination of the minimum value for the distance-size ratio $D$ such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
-
-We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
-More information about this grids can be found in the Table \ref{tab:grids}.
-
-We computed the three gravity fields  on every point of these four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
-Then we calculated the maximum difference between these results and the value of the analytical solution at the same height.
-
-Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
-That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
-
-Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
-
-\begin{table}
-\caption{
-    Description of the synthetic grids on which the comparison of the numerical model against the analytical solutions for a spherical shell was done.
-This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height over the Earth surface.
-Every grid consists in a set of 10$\times$10 points.
-}
-\label{tab:grids}
-\begin{tabular}{lcccc}
-    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
-    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
-    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
-\end{tabular}
-\end{table}
-
-\subsection{Linear Density}
-
-In first place, we proposed a spherical shell with linear density in such way that it has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:
-
-\begin{equation}
-    \rho(r') = ar' + c,
-    \label{eq:density-linear}
-\end{equation}
-\noindent with 
-\begin{equation}
-    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{35000\text{m}},
-\end{equation}
-\begin{equation}
-    c = \frac{3300\text{kg/m$^3$} - 
-        2670\text{kg/m$^3$}}{35000\text{m}} R + 
-        2670\text{kg/m$^3$},
-\end{equation}
-
-\noindent where $R$ = 6378.137 km is the mean Earth radius.
-
-Then we computed the difference between the numerical model with its corresponding analytical solution (equation \ref{eq:density-linear}) for a thick and a thin spherical shell of 1km and 35km of thickness, respectively.
-
-In Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick} we can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor.
-Because the other two diagonal components of the tensor show similar results as the ones obtained for the $g_{zz}$, we exclude them from the figure, although they can be found in the repository.
-\todo{chequear}
-
-For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
-On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D$ equal to 8 while the later only a $D$ of 3.5.
-We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Dlinear-thin-differences.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-linear-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Dlinear-thick-differences.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-linear-thick}
-\end{figure}
-
-\subsection{Exponential Density}
-
-To test the model in case of an exponential density we proposed the following function:
-
-\begin{equation}
-    \rho(r') = A e^{-(r' - R)/b}
-\label{eq:density-exp}
-\end{equation}
-
-\noindent where $A$ = 1kg/m$^3$ and $b$ = 1000km.
-We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, whose results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
-Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them, although, they can be found in the repository.
-
-In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
-
-Surprisingly both the linear and the exponential density tesseroid models needs the same values of distance-size ratio $D$ in order to assure a good accuracy.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thin-differences.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-exp-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thick-differences.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-exp-thick}
-\end{figure}
-
-Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$.
-In order to do this we proposed a set of density functions, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900 kg/m$^{3}$ at the bottom of the spherical shell.
-We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the respective analytical solution.
-These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.
-
-This tests were performed using both the thin and the thick spherical shell.
-On Figures \ref{fig:D-exp-power-thin}a and \ref{fig:D-exp-power-thick}a we can see the density functions used to perform these tests for each spherical shell model, respectively, and their results can be seen in \ref{fig:D-exp-power-thin}b and \ref{fig:D-exp-power-thick}b.
-Taking into account that lower values of $b$ implies a more varying function, we can see that for extremely varying functions the numerical model does not perform well: for $b$=100m the difference is almost 100\% in the thin shell case, and the same happens with $b$=10km in the thick shell case.
-
-On both cases, the model performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, in other words, when the absolute value of the exponent in equation \ref{eq:density-exp} is lower than one.
-Although this sets a limitation to the model, the maximum acceptable variation is very far from real density distributions inside the Earth.
-For example, if we take into account a tesseroid with 35km of thickness, a density of 2670kg/m$^3$ on its top surface and $b=35$km, then the density at its bottom will be approximately of 7000kg/m$^3$, very far from the values we can find at such depths.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences-thin.pdf}
-\caption{
-    Results of the accuracy tests for different exponential density functions in case of a thin spherical shell (1km of thickness).
-    (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
-    (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
-Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-exp-power-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences-thick.pdf}
-\caption{
-    Results of the accuracy tests for different exponential density functions in case of a thick spherical shell (35km of thickness).
-    (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
-    (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
-Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-exp-power-thick}
-\end{figure}
-
-\subsection{Discontinuous Density}
-
-As a final case, we performed the same test with a discontinuous density function:
-
-\begin{equation}
-    \rho(r') = 
-    \begin{cases}
-        \rho_1 & \text{if} \, R_1 \leq r' < R_c \\
-        \rho_2 & \text{if} \, R_c \leq r' \leq R_2 \\
-    \end{cases}
-\label{eq:density-discontinuous}
-\end{equation}
-
-\noindent where $\rho_1$ and $\rho_2$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radius on which the 
-density changes.
-
-The analytical solution for this kind of density is the linear combination of two homogeneous shells with densities $\rho_1$ and $\rho_2$:
-
-\begin{equation}
-    V(r) = \frac{4}{3} \frac{\pi G}{r}
-    \left[ \rho_1 (R_c^3 - R_1^3) +
-    \rho_2 (R_2^3 - R_c^3) \right]
-\end{equation}
-
-Firstly, lets assume a symmetric density function, i.e. $R_c=(R_1 + R_2)/2$ is the centre radius of the shell, and perform the test for a thin shell of 1km of thickness.
-The results can be seen in Figure \ref{fig:D-discont-sym}.
-Again, we obtain good accuracy for values of $D$ equal to 1, 2 and 8 for the computation of the potential, the gradient components and the Marussi tensor components, respectively.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-symmetric-differences.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c$ as the centre radius of the shell.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-discont-sym}
-\end{figure}
-
-Secondly, lets assume the same density function from equation \ref{eq:density-discontinuous}, but now with an asymmetrical distribution: $R_c = (R_1 + R_2)/3$, and perform the same test.
-We can see its results in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what
-leads us to conclude that the code does not recreates the gravity fields for this case with good accuracy.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-asymmetric-differences.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c=(R_1 + R_2)/3$.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-discont-asym}
-\end{figure}
-
-From equation \ref{eq:glq-var-dens} we have seen that the information of the density function is summarised as the values its assumes on the nodes of the GLQ.
-We can attribute the low performance of the code to this simplification.
-
-On the other hand, we can explain the good performance in the symmetrical density case due to the fact that the GLQ nodes are always symmetrical with respect to the centre of the $[a,b]$ integration interval (see equation \ref{eq:glq-scaled}).
-
-In the light of these results we do not recommend using discontinuous densities functions.
-Instead, divide the tesseroid into several ones with specific homogeneous density or density function for each one.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Speed Comparison}
-
-We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous ($\Delta t_\text{homogeneous}$) and variable ($\Delta
-t_\text{variable}$) density at different heights. Then we plotted the ratio between these computation times (see Figure \ref{fig:speed-comparison}):
-
-\begin{equation}
-    \text{Computation Times Ratio} =
-        \frac{\Delta t_\text{variable}}{\Delta t_\text{homogeneous}}
-    \label{eq:computation-times-ratio}
-\end{equation}
-
-From these results We can conclude that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights.
-However, the computation times ratio gets lower at greater heights, due to the reduction of the divisions made by the adaptive discretization algorithm.
-In summary, the new variable density code is fast enough to perform the calculations of a real model in a few minutes.
-\todo{real, esta bien la expresion?}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
-\caption{Computation times ratio between a single tesseroid model with variable and with homogeneous density at different heights, defined in equation \ref{eq:computation-times-ratio}.}
-\label{fig:speed-comparison}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusions}
-
-We have developed a Python library that allows us to perform forward gravity computations using tesseroids with variable density in depth.
-The code numerically approximates the integral that defines the gravity potential, its gradient and the Marussi tensor through the Gauss-Legendre Quadrature (GLQ).
-It essentially consists in approximating the gravity fields of the tesseroids with the effect of point masses located on the scaled nodes of the Legendre polynomials.
-We have stated that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involve Taylor series expansion.
-
-Instead of enhancing the accuracy of the method by increasing the order of the GLQ on each integration, we make use of the modified adaptive discretization algorithm.
-It divides the tesseroid by half if the ratio of the distance to the computation point and its size is lower than a predefined distance-size ratio $D$.
-We can summarise that $D$ controls the accuracy of the method: a higher value of $D$ generates more divisions of the tesseroids, thus more point masses, what leads to a more precise approximation.
-
-On the other hand, there's no direct relation between the value of $D$ and the error of the computation, that's why we had to determine the minimum value of $D$ that produces an acceptable accuracy for each gravity magnitude.
-These distance-size ratio determinations have been made through the comparison of the numerical approximation with the analytical solutions for the exceptional case of a spherical shell.
-From the wide range of possible density variations, we've chosen a linear, an exponential and a discontinuous density functions.
-
-The tests over the first two functions have shown that in order to achieve an accuracy of 0.1\%, a value of $D$ equal to 1, 2 and 8 must be used in the calculation of the gravity potential, its gradient components and the components of the Marussi tensor, respectively.
-We have also found a limitation on the variation of the exponential density function: the absolute value of the exponent must not be greater than 1.
-However, those extremely variations cannot be found inside the Earth, so this limitation does not determines a true restriction for real life models.
-
-On the other hand, the tests over the discontinuous function were not satisfactory.
-In such cases we recommend to divide the model in several tesseroids, each one with a specific homogeneous density or a continuous density function, instead of using a single variable density tesseroid.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Acknowledgments}
-
-We are indebted to the developers and maintainers of the open-source
-software without which this work would not have been possible.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliographystyle{gji}
-\bibliography{bibtex/references}
+\bibliographystyle{apalike-es}
+\bibliography{../manuscript/bibtex/references}
 

### 2017-09-27 jovenes-investigadores/jovenes-investigadores.tex
Continue writing jovenes-investigadores.tex

@@
 
-Las variaciones de la densidad en la corteza con la profundidad han sido estudiadas por casi un siglo: \citet{Athy1930} obtuvo variaciones exponenciales, mientras que otras funciones fueron propuestas en los siguientes años por \citet{Maxant1980, Rao1986, Rao1993, Rao1994}.
-También se ha realizado modelado directo o modelos de inversión gravimétricos con densidades variables, principalmente aplicados en cuencas y fosas \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
+Las variaciones de la densidad en la corteza con la profundidad han sido estudiadas por casi un siglo: se propusieron diferentes tipos de variaciones con la profundidad \citep{Athy1930, Maxant1980, Rao1986, Rao1993, Rao1994} e incluso se han tenido en cuenta en modelos directos y de inversión gravimétrica, principalmente aplicados en cuencas y fosas \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
 
@@
 
-These forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
-Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation on regional scales.
-This raises the importance of building forward models that reproduce the gravity anomalies for such scales.
+\begin{figure}
+\centering
+\includegraphics[width=0.4\linewidth]{../manuscript/figures/tesseroid-uieda.pdf}
+\caption{
+Prisma esférico, también conocido como Tesseroid, en un sistema geocéntrico de coordenadas esféricas, junto a un punto de cómputo $P$ y su respectivo sistema de coordenadas Cartesiano local orientado hacia el Norte.
+Imágen producida por \citet{Uieda2015}.
+}
+\label{fig:tesseroid-uieda}
+\end{figure}
 
-The main issue that should be overcome is taking into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
-A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries (see Figure \ref{fig:tesseroid-uieda}).
-Calculating the gravity fields generated by an arbitrary tesseroid on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
-The literature offers two approaches: one involves Taylor series expansion \citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
-The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.
+Un método muy común para hacerlo es discretizar la Tierra en prismas esféricos conocidos como Tesseroids (ver Figura \ref{fig:tesseroid-uieda}).
+El cómputo de campos gravitatorios generados por un tesseroid arbitrario en un punto externo al mismo involucra la resolución de integrales volumétricas.
+En caso de que el tesseroid posea una densidad variable, las ecuaciones obtenidas por \citet{Grombein2013} \citep[ver también][]{Uieda2016} para el cómputo de los campos gravitatorios se ven ligeramente modificadas.
+Por ejemplo, el potencial generado por un tesseroid arbitrario con densidad $\rho(r')$ sobre un punto de cómputo $P(r, \phi, \lambda)$ se puede obtener como:
+
+\begin{equation}
+    V(r,\phi,\lambda) = G 
+    \int\limits_{\lambda_1}^{\lambda_2}
+    \int\limits_{\phi_1}^{\phi_2}
+    \int\limits_{r_1}^{r_2}
+    \frac{\rho(r')}{\sqrt{{r'}^2 + r^2 - 2 r' r \cos \psi}} \, 
+    r'^2 \cos \phi  dr' d\phi' d\lambda',
+\label{eq:tesseroid-pot}
+\end{equation}
+\noindent donde
+\begin{equation}
+    \cos\psi = \sin\phi\sin\phi' + \cos\phi\cos\phi'
+                 \cos(\lambda' - \lambda).
+\label{eq:cospsi}
+\end{equation}
+
+Este tipos de integrales son generalmente aproximadas por cálculos numéricos.
+Para el caso de densidades variables hemos concluido que uno de los métodos que mejor se adecúa para hacerlo es el que involucra la utilización de la Cuadratura de Gauss Legendre (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
+El mismo consiste en aproximar las integrales por una suma ponderada de los efectos de masas puntuales localizadas en los ceros (previamente rescaldados) de los polinomios de Legendre.
+
+\begin{equation}
+    \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') d\Omega \approx
+    A 
+    \sum\limits_{i=1}^{N^r}
+    \sum\limits_{j=1}^{N^\phi}
+    \sum\limits_{k=1}^{N^\lambda}
+    W_i^r W_j^\phi W_k^\lambda \rho(r_i) g(r_i, \phi_j, \lambda_k),
+\label{eq:glq-var-dens}
+\end{equation}
+\noindent donde
+\begin{equation}
+    g(r', \phi', \lambda') =
+    \frac{r'^2 \cos \phi}{\sqrt{{r'}^2 + r^2 - 2 r' r \cos \psi}},
+    \quad
+    W_i = \frac{2}{(1-x_i^2)[P_N^\prime(x_i)]^2},
+    \quad
+    A = \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8},
+\end{equation}
+
+\noindent $P_N$ es el polinomio de Legendre de orden $N$, $x_i$ sus ceros y $N^i$ con $i \in \{ r, \phi, \lambda\}$ los órdenes de las aproximaciones GLQ.
+
+\citet{Uieda2016} desarrolló un modelo directo que permite calcular los campos gravitatorios generados por tesseroides con densidad homogenea utilzando las GLQ, implementado en Python e incluido en la librería de código abierto Fatiando a Terra \citep{Uieda2013} para el modelado e inversión geofísica.
+
+\citet{Ku1977} advirtió que la aproximación GLQ se vuelve menos precisa a medida que el punto de cómputo se acerca al tesseroide o bien para órdenes menores de GLQ.
+En vez de aumentar los órdenes de la aproximación, \citet{Uieda2016} lo fija a segundo orden y aumenta la precisión a través de una versión modificada del algoritmo de discretización adaptativa de \citet{Li2011}.
+Este consiste en dividir al tesseroide en otros más pequeños en caso de que el cociente entre $d$ y $L_i$ es menor a un valor predefinido $D$ llamado relación distancia-tamaño, donde $d$ es la distancia entre el punto de cómputo y el centro del tesseroide, mientras $L_i$ son las dimensiones del mismo ($i \in \{ r, \phi, \lambda\}$).
+
+\begin{equation}
+    \frac{d}{L_i} \geq D,
+\label{eq:distance-size-ratio}
+\end{equation}
+
+De esta manera, el valor asignado a $D$ controla la precisión de la aproximación.
+Sin embargo, no existe una relación directa entre $D$ y el error de la misma.
+Es por ello que \citet{Uieda2016} obtuvo valores estándar de $D$ comparando los campos gravitatorios generados por un cascarón esférico con densidad constante, un caso especial de las integrales volumétricas que posee solución analítica \citep{LaFehr1991, Mikuska2006, Grombein2013}.
+
+En este trabajo hemos implementado un modelo directo como el de \citet{Uieda2016} que además permite calcular los campos gravitatorios generados por tesseroides con densidades variables arbitrarias en profundidad.
+Al igual que el anterior, hace uso de la aproximación GLQ, de la versión modificada del algoritmo de discretización adaptativa y su implementación está escrita en lenguajes Python y Cython.
+
+Con el objetivo de garantizar la precisión de la aproximación numérica, hemos llevado a cabo pruebas similares a las que realizó \citet{Uieda2016} comparando los resultados del modelo numérico con las soluciones analíticas de un cascarón esférico, en nuestro caso con densidades lineales, exponenciales y discontinuas.
+Para lo cual fue necesario obtener las soluciones analíticas correspondientes a cada función de densidad.
 

### 2017-09-28 jovenes-investigadores/jovenes-investigadores.tex
Continue writting jovenes-investigadores

@@
 \usepackage{graphicx}
+\usepackage{subcaption} % enables subfigure
 \usepackage{natbib}
@@
 
+\renewcommand{\tablename}{Tabla}
+
 \maketitle
@@
 \begin{figure}
+    \begin{subfigure}[t]{0.5\textwidth}
     \centering
-\includegraphics[width=0.4\linewidth]{../manuscript/figures/tesseroid-uieda.pdf}
+    \includegraphics[height=15em]{../manuscript/figures/tesseroid-uieda.pdf}
     \caption{
@@
     \label{fig:tesseroid-uieda}
+    \end{subfigure}
+    \quad
+    \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \includegraphics[height=15em]{../manuscript/figures/spherical-shell.pdf}
+    \caption{
+        Cascarón esférico con radios interior y exteior $R_1$ y $R_2$, respectivamente.
+El punto de cómputo $Q$ está localizado en el eje $z$ a una distancia $r$ al centro del cascarón.
+Vamos a suponer que $Q$ se encuentra por fuera del radio exterior $R_2$, es decir $r > R_2$.
+    }
+    \label{fig:spherical-shell}
+    \end{subfigure}
+    \caption{}
 \end{figure}
@@
 \citet{Ku1977} advirtió que la aproximación GLQ se vuelve menos precisa a medida que el punto de cómputo se acerca al tesseroide o bien para órdenes menores de GLQ.
-En vez de aumentar los órdenes de la aproximación, \citet{Uieda2016} lo fija a segundo orden y aumenta la precisión a través de una versión modificada del algoritmo de discretización adaptativa de \citet{Li2011}.
+En vez de aumentar los órdenes de la aproximación, \citet{Uieda2016} lo fijan a segundo orden y aumentan la precisión a través de una versión modificada del algoritmo de discretización adaptativa de \citet{Li2011}.
 Este consiste en dividir al tesseroide en otros más pequeños en caso de que el cociente entre $d$ y $L_i$ es menor a un valor predefinido $D$ llamado relación distancia-tamaño, donde $d$ es la distancia entre el punto de cómputo y el centro del tesseroide, mientras $L_i$ son las dimensiones del mismo ($i \in \{ r, \phi, \lambda\}$).
 
-\begin{equation}
-    \frac{d}{L_i} \geq D,
-\label{eq:distance-size-ratio}
-\end{equation}
-
-De esta manera, el valor asignado a $D$ controla la precisión de la aproximación.
-Sin embargo, no existe una relación directa entre $D$ y el error de la misma.
+De esta manera, el valor asignado a $D$ controla la precisión de la aproximación: para mayores valores de $D$, el error de la aproximación se reduce.
+Sin embargo, no existe una relación directa entre $D$ y dicho error.
 Es por ello que \citet{Uieda2016} obtuvo valores estándar de $D$ comparando los campos gravitatorios generados por un cascarón esférico con densidad constante, un caso especial de las integrales volumétricas que posee solución analítica \citep{LaFehr1991, Mikuska2006, Grombein2013}.
@@
 
-Con el objetivo de garantizar la precisión de la aproximación numérica, hemos llevado a cabo pruebas similares a las que realizó \citet{Uieda2016} comparando los resultados del modelo numérico con las soluciones analíticas de un cascarón esférico, en nuestro caso con densidades lineales, exponenciales y discontinuas.
+Para garantizar la precisión de la aproximación numérica, hemos llevado a cabo pruebas similares a las que realizaron \citet{Uieda2016} comparando los resultados del modelo numérico con las soluciones analíticas de un cascarón esférico, en nuestro caso con densidades lineales, exponenciales y discontinuas.
 Para lo cual fue necesario obtener las soluciones analíticas correspondientes a cada función de densidad.
+
+\section{Determinación de Relación Distancia-Tamaño}
+
+Con el objetivo de evaluar el nuevo modelo directo con tesseroides de densidad variable y estimar el error de la aproximación hemos comparado el modelo numérico con las soluciones analíticas de un cascarón esférico para tres casos especiales: una densidad linear, una exponencial y una discontinua.
+
+La estimación del error nos es útil para la determinación del mínimo valor de relación distancia-tamaño $D$ de manera tal que la diferencia entre el modelo numérico y las soluciones analíticas estén debajo de un umbral aceptable.
+
+El cascarón esférico ha sido aproximado a través de una malla de tesseroides de $30^\circ \times 30^\circ$ y hemos calculado el potencial gravitatorio, la componente vertical del gradiente $g_z$ y la componente diagonal del tensor de Marussi $g_{zz}$ que la malla genera sobre cuatro grillas: (1) una localizada en el polo, (2) otra en el Ecuador, (3) una localizada a altura satelital y (4) una gran grilla de $30^\circ \times 30^\circ$.
+En la Tabla \ref{tab:grids} podemos obtener más información acerca de estas grillas.
+
+El cómputo de estos
+
+We computed the three gravity fields  on every point of these four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
+Then we calculated the maximum difference between these results and the value of the analytical solution at the same height.
+
+Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
+That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
+
+Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
+
+\begin{table}
+\centering
+\caption{
+    Descripciones de las grillas sintéticas en las cuales la comparación entre el modelo numérico y las soluciones analíticas se lleva a cabo.
+Esta colección de grilla intenta incluir todo posible escenario en el cual el modelo numérico pueda presentar diferentes comportamientos, dependiendo del tamaño de la grilla, su ubicación o su altura sobre la superficie de la Tierra.
+Cada grilla consiste en un conjunto de 10$\times$10 puntos.
+}
+\label{tab:grids}
+\begin{tabular}{lcccc}
+    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
+    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
+    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
+\end{tabular}
+\end{table}
+
+\begin{figure}
+    \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \includegraphics[height=30em]{../manuscript/figures/Dlinear-thin-differences.pdf}
+    \caption{}
+    \label{fig:D-linear-thin}
+    \end{subfigure}
+    \quad
+    \begin{subfigure}[t]{0.5\textwidth}
+    \centering
+    \includegraphics[height=30em]{../manuscript/figures/Dexp-shifted-thin-differences.pdf}
+    \caption{}
+    \label{fig:D-exp-thin}
+    \end{subfigure}
+    \caption{
+        Diferencias entre los campos gravitatorios generados por el modelo numérico y la solución analítica para un cascarón esférico de 1km de grosor con (a) densidad lineal definida en la ecuación \ref{eq:density-linear} y (b) densidad exponencial definida en la ecuación \ref{eq:density-exp}.
+Los computos fueron realizados en las cuatro grillas descriptas en la Tabla \ref{tab:grids} y para diferentes valores de la relación distancia-tamaño $D$.
+Si la diferencia es menor a 0.1\% consideraremos que el modelo ha alcanzado una precisión aceptable para el valor correspondiente de $D$.
+    }
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Ejemplo con datos reales}
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{../manuscript/figures/neuquen-basin.pdf}
+\caption{(a) Topography of Neuqu\'en Basin (in km),
+         (b) Thickness of the sedimentary basin (in meters),
+         (c)-(h) Computed gravity fields: potential $V$ (in J/kg), gradient components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components $g_{xy}$ and $g_{zz}$ (in Eötvös), calculated at 50km of height over the ellipsoid.}
+\label{fig:neuquen-basin}
+\end{figure}
 

### 2017-09-28 jovenes-investigadores/jovenes-investigadores.tex
Continue writing jovenes-investigadores

@@
 
-El cómputo de estos
-
-We computed the three gravity fields  on every point of these four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
-Then we calculated the maximum difference between these results and the value of the analytical solution at the same height.
-
-Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
-That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
-
-Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
+Además, el cómputo de estos campos gravitatorios en cada punto de estas grillas fue realizado para diferentes valores de $D$: desde 0.5 a 10 con un paso de 0.5.
+Y luego calculamos la máxima diferencia entre estos resultados y el valor de la solución analítica a la misma altura.
+Finalmente obtenemos como valor aceptable de $D$ si el error de la correspondiente aproximación numérica es menor a 0.1\%.
 
@@
 \begin{tabular}{lcccc}
-    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
+    Grilla & Tamaño & Extensión Lat. & Extensión Lon. & Altura \\ \hline
     Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
@@
 \end{table}
+
+La determinación de $D$ fue realizada para una función lineal (ecuación \ref{eq:density-linear}) y una exponencial (ecuación \ref{eq:density-exp}), cuyos resultados puede observarse en las Figuras \ref{fig:D-linear-thin} y \ref{fig:D-exp-thin}, respectivamente.
+
+\begin{equation}
+    \rho(r') = ar' + c,
+    \label{eq:density-linear}
+\end{equation}
+
+\begin{equation}
+    \rho(r') = A e^{-(r' - R)/b}
+\label{eq:density-exp}
+\end{equation}
 
@@
 
+A partir de las mismas podemos concluir que para ambos casos es necesario un $D$ igual a 1, 2 y 8 para el cálculo del potencial gravitatorio, de las componentes del gradiente ($g_i$) y las componentes del tensor de Marussi ($g_ij$), $i \in \{ x, y, z\}$, respectivamente.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Ejemplo con datos reales}
+
+Con el objetivo de exponer el comportamiento del nuevo modelo directo hemos decidido aplicarlo sobre datos reales, para lo cual hemos escogido la cuenca Neuquina: una cuenca sedimentaria que se encuentra al Este de los Andes, entre los 32$^\circ$S y los 40$^\circ$S de latitud (ver Figura \ref{fig:neuquen-basin}a) e incluye siliciclastos, carbonatos y evaporitas continentales y marinos acumulados a lo largo del Jurasico y el Cretaceo, constituyendo un registro estratigráfico hasta los 5000m de profundidad \citep{Howell2005}.
 
@@
 
+El espesor del paquete sedimentario fue obtenido de \citet{Heine2007}, a partir del cual construimos una grilla regular con una resolución de 0.05$^\circ$ en ambas direcciones latitudinal y longitudinal con un valor de espesor por cada nodo de la grilla (ver Figura \ref{fig:neuquen-basin}b).
+
+Con el objetivo de calcular los campos gravitatorios generados por esta cuenca, hemos creado una malla de tesseroides, cada uno localizado sobre uno de los nodos de la grilla, con dimensiones de 0.05$^\circ$ $\times$ 0.05$^\circ$ y profundidad igual espesor de la cuenca en el nodo correspondiente.
+
+Además, debimos definir una función densidad para el modelo completo.
+\citet{Sigismondi2012} propuso una densidad mínima y un máxima para la cuenca Nequina de 2755kg/m$^3$ y 412kg/m$^3$, respectivamente.
+Para este ejemplo específico hemos escogido una densidad lineal que asume el mínimo en la superficie y el máximo a los 5858m, es decir en la profundidad máxima de la cuenca.
+
+Finalmente hemos realizado el cómputo del potencial gravitatorio, todas las componentes del gradiente ($g_x$, $g_y$, $g_z$) y las componentes $g_{xy}$ y $g_{zz}$ del tensor de Marussi sobre una grilla de cómputo de 159$\times$163 nodos con una separación de 0.05$^\circ$ en ambas direcciones latitudinal y longitudinal, ubicada a 50km de altura sobre el ellipsoide de referencia.
+Los campos gravitatorios resultantes pueden observarse en las Figuras \ref{fig:neuquen-basin}c-h.
+
+We have finally computed the gravity potential $V$, the gradient components $g_x$, $g_y$ and $g_z$, and the Marussi tensor components $g_{xy}$ and $g_{zz}$ on a computation grid of 159$\times$163 nodes with a 0.05$^\circ$ spacing on both longitude and latitude directions at a 50km height over the reference ellipsoid.
+The resulting gravity fields can be seen in Figures \ref{fig:neuquen-basin}c-h.
+We have excluded the other components of the Marussi tensor from the Figure, although they have also been computed and can be found in the repository.
+
 \bibliographystyle{apalike-es}

### 2017-09-28 manuscript/manuscript.tex
Fixed typo
Fixed bad reference to Figure

@@
 
-We computed the three gravity fields  on every point of these four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
+We computed these gravity fields  on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
 Then we calculated the maximum difference between these results and the value of the analytical solution at the same height.
@@
 The thickness of the sediment pack was obtained from \citet{Heine2007}.
-We managed to produce a regular grid with a resolution of 0.05$^\circ$ on both longitude and latitude directions with one value of thickness on each node (see Figure \ref{fig:neuquen-basin}).
+We managed to produce a regular grid with a resolution of 0.05$^\circ$ on both longitude and latitude directions with one value of thickness on each node (see Figure \ref{fig:neuquen-basin}b).
 

### 2017-09-28 jovenes-investigadores/jovenes-investigadores.tex
First final document for jovenes-investigadores

@@
 \usepackage{natbib}
+\setlength{\bibsep}{0.0pt}      % shrink spaces in references
 \usepackage{url}
@@
     \caption{
-        Prisma esférico, también conocido como Tesseroid, en un sistema geocéntrico de coordenadas esféricas, junto a un punto de cómputo $P$ y su respectivo sistema de coordenadas Cartesiano local orientado hacia el Norte.
+        Prisma esférico, también conocido como Tesseroide, en un sistema geocéntrico de coordenadas esféricas, junto a un punto de cómputo $P$ y su respectivo sistema de coordenadas Cartesiano local orientado hacia el Norte.
 Imágen producida por \citet{Uieda2015}.
@@
 
-Un método muy común para hacerlo es discretizar la Tierra en prismas esféricos conocidos como Tesseroids (ver Figura \ref{fig:tesseroid-uieda}).
-El cómputo de campos gravitatorios generados por un tesseroid arbitrario en un punto externo al mismo involucra la resolución de integrales volumétricas.
-En caso de que el tesseroid posea una densidad variable, las ecuaciones obtenidas por \citet{Grombein2013} \citep[ver también][]{Uieda2016} para el cómputo de los campos gravitatorios se ven ligeramente modificadas.
-Por ejemplo, el potencial generado por un tesseroid arbitrario con densidad $\rho(r')$ sobre un punto de cómputo $P(r, \phi, \lambda)$ se puede obtener como:
+Un método muy común para hacerlo es discretizar la Tierra en prismas esféricos conocidos como Tesseroides (ver Figura \ref{fig:tesseroid-uieda}).
+El cómputo de campos gravitatorios generados por un tesseroide arbitrario en un punto externo al mismo involucra la resolución de integrales volumétricas.
+En caso de que el tesseroide posea una densidad variable, las ecuaciones obtenidas por \citet{Grombein2013} \citep[ver también][]{Uieda2016} para el cómputo de los campos gravitatorios se ven ligeramente modificadas.
+Por ejemplo, el potencial generado por un tesseroide arbitrario con densidad $\rho(r')$ sobre un punto de cómputo $P(r, \phi, \lambda)$ se puede obtener como:
 
@@
 
-Para garantizar la precisión de la aproximación numérica, hemos llevado a cabo pruebas similares a las que realizaron \citet{Uieda2016} comparando los resultados del modelo numérico con las soluciones analíticas de un cascarón esférico, en nuestro caso con densidades lineales, exponenciales y discontinuas.
+Para garantizar la precisión de la aproximación numérica, hemos llevado a cabo pruebas similares a las que realizaron \citet{Uieda2016} comparando los resultados del modelo numérico con las soluciones analíticas de un cascarón esférico, en nuestro caso con densidades lineales y exponenciales.
 Para lo cual fue necesario obtener las soluciones analíticas correspondientes a cada función de densidad.
@@
 
-Con el objetivo de evaluar el nuevo modelo directo con tesseroides de densidad variable y estimar el error de la aproximación hemos comparado el modelo numérico con las soluciones analíticas de un cascarón esférico para tres casos especiales: una densidad linear, una exponencial y una discontinua.
+Con el objetivo de evaluar el nuevo modelo directo con tesseroides de densidad variable y estimar el error de la aproximación hemos comparado el modelo numérico con las soluciones analíticas de un cascarón esférico para dos casos especiales: una densidad linear y una exponencial
 
@@
 
-El cascarón esférico ha sido aproximado a través de una malla de tesseroides de $30^\circ \times 30^\circ$ y hemos calculado el potencial gravitatorio, la componente vertical del gradiente $g_z$ y la componente diagonal del tensor de Marussi $g_{zz}$ que la malla genera sobre cuatro grillas: (1) una localizada en el polo, (2) otra en el Ecuador, (3) una localizada a altura satelital y (4) una gran grilla de $30^\circ \times 30^\circ$.
+Hemos aproximado el cascarón esférico a través de una malla de tesseroides de $30^\circ \times 30^\circ$ para luego calcular el potencial gravitatorio, la componente vertical del gradiente $g_z$ y la componente diagonal del tensor de Marussi $g_{zz}$ que la malla genera sobre cuatro grillas: (1) una localizada en el polo, (2) otra en el Ecuador, (3) una localizada a altura satelital y (4) una gran grilla de $30^\circ \times 30^\circ$.
 En la Tabla \ref{tab:grids} podemos obtener más información acerca de estas grillas.
 
-Además, el cómputo de estos campos gravitatorios en cada punto de estas grillas fue realizado para diferentes valores de $D$: desde 0.5 a 10 con un paso de 0.5.
-Y luego calculamos la máxima diferencia entre estos resultados y el valor de la solución analítica a la misma altura.
+Además, el cómputo de estos campos gravitatorios en cada punto de estas grillas fue realizado para diferentes valores de $D$: desde 0.5 a 10 con un paso de 0.5, para luego calcular la máxima diferencia entre estos resultados y el valor de la solución analítica a la misma altura.
 Finalmente obtenemos como valor aceptable de $D$ si el error de la correspondiente aproximación numérica es menor a 0.1\%.
@@
 
-La determinación de $D$ fue realizada para una función lineal (ecuación \ref{eq:density-linear}) y una exponencial (ecuación \ref{eq:density-exp}), cuyos resultados puede observarse en las Figuras \ref{fig:D-linear-thin} y \ref{fig:D-exp-thin}, respectivamente.
-
-\begin{equation}
-    \rho(r') = ar' + c,
-    \label{eq:density-linear}
-\end{equation}
-
-\begin{equation}
-    \rho(r') = A e^{-(r' - R)/b}
-\label{eq:density-exp}
-\end{equation}
-
-\begin{figure}
+\begin{figure}[t!]
     \begin{subfigure}[t]{0.5\textwidth}
@@
 
+La determinación de $D$ fue realizada para una función lineal (ecuación \ref{eq:density-linear}) y una exponencial (ecuación \ref{eq:density-exp}), cuyos resultados puede observarse en las Figuras \ref{fig:D-linear-thin} y \ref{fig:D-exp-thin}, respectivamente.
+
+\begin{equation}
+    \rho(r') = ar' + c,
+    \label{eq:density-linear}
+\end{equation}
+
+\begin{equation}
+    \rho(r') = A e^{-(r' - R)/b}
+\label{eq:density-exp}
+\end{equation}
+
 A partir de las mismas podemos concluir que para ambos casos es necesario un $D$ igual a 1, 2 y 8 para el cálculo del potencial gravitatorio, de las componentes del gradiente ($g_i$) y las componentes del tensor de Marussi ($g_ij$), $i \in \{ x, y, z\}$, respectivamente.
@@
 
-\begin{figure}
+\begin{figure}[b!]
 \centering
-\includegraphics[width=\linewidth]{../manuscript/figures/neuquen-basin.pdf}
+\includegraphics[width=0.975\linewidth]{../manuscript/figures/neuquen-basin.pdf}
 \caption{(a) Topography of Neuqu\'en Basin (in km),
@@
 Finalmente hemos realizado el cómputo del potencial gravitatorio, todas las componentes del gradiente ($g_x$, $g_y$, $g_z$) y las componentes $g_{xy}$ y $g_{zz}$ del tensor de Marussi sobre una grilla de cómputo de 159$\times$163 nodos con una separación de 0.05$^\circ$ en ambas direcciones latitudinal y longitudinal, ubicada a 50km de altura sobre el ellipsoide de referencia.
-Los campos gravitatorios resultantes pueden observarse en las Figuras \ref{fig:neuquen-basin}c-h.
-
-We have finally computed the gravity potential $V$, the gradient components $g_x$, $g_y$ and $g_z$, and the Marussi tensor components $g_{xy}$ and $g_{zz}$ on a computation grid of 159$\times$163 nodes with a 0.05$^\circ$ spacing on both longitude and latitude directions at a 50km height over the reference ellipsoid.
-The resulting gravity fields can be seen in Figures \ref{fig:neuquen-basin}c-h.
-We have excluded the other components of the Marussi tensor from the Figure, although they have also been computed and can be found in the repository.
+Los campos gravitatorios resultantes pueden observarse en las Figuras \ref{fig:neuquen-basin}c-h..
 

### 2017-09-29 jovenes-investigadores/jovenes-investigadores.tex
Primer borrador del trabajo para jovenes-investigadores

@@
 \usepackage{subcaption} % enables subfigure
+\usepackage{multicol}
 \usepackage{natbib}
@@
     Santiago R. Soler$^{1,2}$ \vspace{0.5em} \\ 
-    \textit{$^1$ Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan} \\
-    \textit{$^2$ Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET)} \vspace{0.4em} \\
-    email: santiago.r.soler@gmail.com
+    \normalsize{\textit{$^1$ Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan}} \\
+    \normalsize{\textit{$^2$ Consejo Nacional de Investigaciones Científicas y Técnicas (CONICET)}} \vspace{0.4em} \\
+    \normalsize{email: santiago.r.soler@gmail.com}
 }
@@
 \begin{center}
-\textbf{Palabras Clave:} poner algo
+\textbf{Palabras Clave:} gravimetría, tesseroid, densidad variable
 \end{center}
@@
 El cálculo directo de campos gravitatorios para grandes extensiones requiere tomar en cuenta la curvatura de la Tierra, lo cual suele resolverse a través del uso de prismas esféricos llamados Tesseroides.
-Además, estructuras como cuencas sedimentarias presentan variaciones de densidad con la profundidad.
+Además, muchas estructuras geológicas presentan variaciones de la densidad con la profundidad.
+En este trabajo hemos sido capaces de implementar un modelo directo que reproduce los campos gravitatorios generados por Tesseroides con densidad variable en profundidad, para lo cual hemos utilizado como aproximación la Cuadratura de Gauss-Legendre y un algoritmo de discretización adaptativa con el objetivo de garantizar la precisión de la misma.
+El código fue escrito en Python y Cython y será liberado como Software Libre.
 
@@
     \centering
-    \includegraphics[height=15em]{../manuscript/figures/tesseroid-uieda.pdf}
+    \includegraphics[height=18em]{../manuscript/figures/tesseroid-uieda.pdf}
     \caption{
@@
     \centering
-    \includegraphics[height=15em]{../manuscript/figures/spherical-shell.pdf}
+    \includegraphics[height=18em]{../manuscript/figures/spherical-shell.pdf}
     \caption{
@@
 Para el caso de densidades variables hemos concluido que uno de los métodos que mejor se adecúa para hacerlo es el que involucra la utilización de la Cuadratura de Gauss Legendre (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
-El mismo consiste en aproximar las integrales por una suma ponderada de los efectos de masas puntuales localizadas en los ceros (previamente rescaldados) de los polinomios de Legendre.
+El mismo consiste en aproximar las integrales por una suma ponderada de los efectos de masas puntuales localizadas en los ceros (previamente rescaldados) de los polinomios de Legendre:
 
@@
 
-\citet{Uieda2016} desarrolló un modelo directo que permite calcular los campos gravitatorios generados por tesseroides con densidad homogenea utilzando las GLQ, implementado en Python e incluido en la librería de código abierto Fatiando a Terra \citep{Uieda2013} para el modelado e inversión geofísica.
+\citet{Uieda2016} desarrollaron un modelo directo que permite calcular los campos gravitatorios generados por tesseroides con densidad homogenea utilzando las GLQ, implementado en Python e incluido en la librería de código abierto Fatiando a Terra \citep{Uieda2013} para el modelado e inversión geofísica.
 
@@
 Sin embargo, no existe una relación directa entre $D$ y dicho error.
-Es por ello que \citet{Uieda2016} obtuvo valores estándar de $D$ comparando los campos gravitatorios generados por un cascarón esférico con densidad constante, un caso especial de las integrales volumétricas que posee solución analítica \citep{LaFehr1991, Mikuska2006, Grombein2013}.
+Es por ello que \citet{Uieda2016} obtuvieron valores estándar de $D$ comparando los campos gravitatorios generados por un cascarón esférico con densidad constante, un caso especial de las integrales volumétricas que posee solución analítica \citep{LaFehr1991, Mikuska2006, Grombein2013}.
 
@@
 Además, el cómputo de estos campos gravitatorios en cada punto de estas grillas fue realizado para diferentes valores de $D$: desde 0.5 a 10 con un paso de 0.5, para luego calcular la máxima diferencia entre estos resultados y el valor de la solución analítica a la misma altura.
-Finalmente obtenemos como valor aceptable de $D$ si el error de la correspondiente aproximación numérica es menor a 0.1\%.
+Finalmente obtenemos como valor aceptable de $D$ el que presente un error de la correspondiente aproximación numérica menor a 0.1\%.
 
-\begin{table}
-\centering
-\caption{
-    Descripciones de las grillas sintéticas en las cuales la comparación entre el modelo numérico y las soluciones analíticas se lleva a cabo.
-Esta colección de grilla intenta incluir todo posible escenario en el cual el modelo numérico pueda presentar diferentes comportamientos, dependiendo del tamaño de la grilla, su ubicación o su altura sobre la superficie de la Tierra.
-Cada grilla consiste en un conjunto de 10$\times$10 puntos.
-}
-\label{tab:grids}
-\begin{tabular}{lcccc}
-    Grilla & Tamaño & Extensión Lat. & Extensión Lon. & Altura \\ \hline
-    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
-    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
-\end{tabular}
-\end{table}
-
-\begin{figure}[t!]
+\begin{figure}[b!]
     \begin{subfigure}[t]{0.5\textwidth}
@@
 
-La determinación de $D$ fue realizada para una función lineal (ecuación \ref{eq:density-linear}) y una exponencial (ecuación \ref{eq:density-exp}), cuyos resultados puede observarse en las Figuras \ref{fig:D-linear-thin} y \ref{fig:D-exp-thin}, respectivamente.
+\begin{table}
+\centering
+\caption{
+    Descripciones de las grillas sintéticas en las cuales la comparación entre el modelo numérico y las soluciones analíticas se lleva a cabo.
+Esta colección de grilla intenta incluir todo posible escenario en el cual el modelo numérico pueda presentar diferentes comportamientos, dependiendo del tamaño de la grilla, su ubicación o su altura sobre la superficie de la Tierra.
+Cada grilla consiste en un conjunto de 10$\times$10 puntos.
+}
+\label{tab:grids}
+\begin{tabular}{lcccc}
+    Grilla & Tamaño & Extensión Lat. & Extensión Lon. & Altura \\ \hline
+    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
+    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
+\end{tabular}
+\end{table}
 
+Este proceso de determinación de $D$ fue realizado para una función lineal (ecuación \ref{eq:density-linear}) y una exponencial (ecuación \ref{eq:density-exp}), cuyos resultados pueden observarse en las Figuras \ref{fig:D-linear-thin} y \ref{fig:D-exp-thin}, respectivamente.
+
+\vspace{-2.5em}
+\begin{multicols}{2}
+    \centering
     \begin{equation}
@@
     \end{equation}
-
+    \break
+    \centering
     \begin{equation}
@@
     \end{equation}
+\end{multicols}
 
@@
 
-Con el objetivo de calcular los campos gravitatorios generados por esta cuenca, hemos creado una malla de tesseroides, cada uno localizado sobre uno de los nodos de la grilla, con dimensiones de 0.05$^\circ$ $\times$ 0.05$^\circ$ y profundidad igual espesor de la cuenca en el nodo correspondiente.
+Con el objetivo de calcular los campos gravitatorios generados por esta cuenca, hemos creado una malla de tesseroides, cada uno localizado sobre uno de los nodos de la grilla, con dimensiones de 0.05$^\circ$ $\times$ 0.05$^\circ$ y profundidad igual al  espesor de la cuenca en el nodo correspondiente.
 
-Además, debimos definir una función densidad para el modelo completo.
+Además, tuvimos que definir una función densidad para el modelo completo.
 \citet{Sigismondi2012} propuso una densidad mínima y un máxima para la cuenca Nequina de 2755kg/m$^3$ y 412kg/m$^3$, respectivamente.
@@
 Finalmente hemos realizado el cómputo del potencial gravitatorio, todas las componentes del gradiente ($g_x$, $g_y$, $g_z$) y las componentes $g_{xy}$ y $g_{zz}$ del tensor de Marussi sobre una grilla de cómputo de 159$\times$163 nodos con una separación de 0.05$^\circ$ en ambas direcciones latitudinal y longitudinal, ubicada a 50km de altura sobre el ellipsoide de referencia.
-Los campos gravitatorios resultantes pueden observarse en las Figuras \ref{fig:neuquen-basin}c-h..
+Los campos gravitatorios resultantes pueden observarse en las Figuras \ref{fig:neuquen-basin}c-h.
 

### 2017-10-05 code/tesseroid_density/tesseroid.py
Simplified line computation in _divider_calculation in tesseroid.py

@@ -268,9 +268,8 @@ def _divider_calculation(tesseroid):
     density_fun = tesseroid.props['density']
     density_top, density_bottom = density_fun(top), density_fun(bottom)
     heights = np.linspace(bottom, top, 101)
-    line = (density_top - density_bottom)/(top - bottom) * heights - \
-        (density_top - density_bottom)/(top - bottom) * bottom + \
-        density_bottom
+    line = (density_top - density_bottom)/(top - bottom) * \
+           (heights - bottom) + density_bottom
     densities = np.array([density_fun(height) for height in heights])
     diff = densities - line
     if np.allclose(diff, diff[0]):

### 2017-10-06 manuscript/manuscript.tex
Added density-based discretization algorithm to manuscript

@@
 We have enhanced the later implementation with a new forward gravity model that allows us to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
-It's based on the GLQ approximation, makes use of the modified distance-size discretization algorithm and is written in Python and Cython languages, obtaining an easy to use library that runs precompiled code for the more time consuming functions.
+It is based on the GLQ approximation, makes use of the modified distance-size discretization algorithm and a new density-based discretization that increases the accuracy for high variating density functions.
+It is written in Python and Cython languages, constituting an easy to use library that runs precompiled code for the more time consuming functions.
 
@@
 
+\subsection{Modified Adaptive Discretization Algorithm}
+
 From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of $N_\lambda N_\phi N_r$ point masses, where $N_i$ are the orders of the quadrature for the integral on each coordinate, with $i \in \{ \lambda, \phi, r \}$.
@@
 From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
+
+\subsection{Density-based Discretization Algorithm}
+
+High variations in the density function of an arbitrary tesseroid introduces a new type of error into the numerical approximation: only a few point masses would not reproduce the effect of this kind of variations.
+The modified version of the adaptive discretization algorithm may help in reducing this kind of error, but as it takes into account only the dimensions of the tesseroid and the distance to the computation point, it's not completely well suited to perform this task.
+
+In order to overcome this, we developed a complementary discretization algorithm (it works together with the previous one) that takes into account the variation of the density inside the tesseroid.
+Basically, we will subdivide any tesseroid on the radius coordinate on the depth at which the ``maximum density variation'' takes place, creating two new tesseroids which we use to compute the desired gravity field by applying the previous adaptive discretization algorithm and the GLQ approximation.
+
+First we need to define a simple density that approximates the one of the tesseroid: a linear function that assumes the same values of the original density on the tesseroid's top and bottom:
+
+\begin{equation}
+    \rho_0(r') =
+    \frac{\rho(R_2) - \rho(R_1)}{R_2 - R_1} (r' - R_1) + \rho(R_1).
+\end{equation}
+
+Then we will define the ``maximum density variation'' as the maximum difference between this density $\rho_0(r')$ and the original density $\rho(r')$ and obtain the radius coordinate $R_\text{max}$ at which this maximum takes place:
+
+\begin{equation}
+    \text{max}\{ | \rho(r') - \rho_0(r') | \} = 
+    | \rho(R_\text{max}) - \rho_0(R_\text{max}) |
+\end{equation}
+
+Finally, we subdivide the original tesseroid at $R_\text{max}$ creating two new tesseroids with the same latitude and longitude boundaries.
+We then apply the modified version of the adaptive discretization algorithm and the GLQ to approximate the gravity fields that it generates on the computation points.
+
+This density-based discretization can be applied several times, creating new pair of subdivisions each time and therefore increasing the accuracy of the resulting gravity field.
+Because each tesseroid is subdivided in two every time we apply this discretization, the number of subdivisions increases exponentially in power of two:
+
+\begin{equation}
+    \text{Number of subdivisions} =
+    \sum\limits_{m=0}^M 2^m,
+\end{equation}
+
+\noindent where $M$ is the order of the discretization, i.e. how many times the density-based discretization algorithm is applied on one tesseroid.
 

### 2017-10-06 manuscript/manuscript.tex
Modofications to implementation in manuscript

@@
 
-We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth through the GLQ to numerically approximate the volume integrals.
-It's freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}
+We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth.
+It is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}
 
-Although it's a Python library, the core functions of the computation are written in Cython, a superset of Python language that allows to generate efficient C code through the Cython compiler.
-This produces a precompiled library whose functions can be called from regular Python code, what allowed us to make it compatible with the preexisting open-source library Fatiando a Terra \citep{Uieda2013}.
+This new library is based on the preexisting code developed by \citet{Uieda2016} (and included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that is able to compute only the gravitational effects of tesseroids with homogeneous density.
+\citet{Uieda2016} have optimised the more time consuming functions through the Numba compiler.
+In contrast, we have decided to develop our new library in Cython language due to difficulties when passing a Python function as argument to precompiled Numba methods, keeping the homogeneous tesseroid calculation and adding the new variable density computations.
+The Cython language is a superset of Python language that allows to generate efficient C code through the Cython compiler, producing a library whose functions can be called from regular Python code.
 
-Its usage is very similar to the code for homogeneous tesseroid already in Fatiando a Terra (version 0.5):
+Moreover, this new software is compatible with the preexisting open-source library Fatiando a Terra and its usage is very similar to the one for homogeneous tesseroids:
 
@@
 \renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Define a single tesseroid (or a tesseroid mesh) with the Fatiando a Terra classes.
-    \item Add a predefined Python density function to the tesseroid (or each tesseroid in the mesh).
-    \item Calculate the desired gravity field (potential, gradient or tensor component) in any set of computation points with our new functions.
+    \item Define a single tesseroid (or a set of several tesseroids) with the Fatiando a Terra classes.
+    \item Add a predefined Python density function to the tesseroid (or to each tesseroid in the set).
+    \item Calculate the desired gravity field (potential, gradient or tensor components) on any set of computation points.
 \end{enumerate}
 
-The existing code in Fatiando a Terra (v.~0.5) for homogeneous tesseroid is optimised through the Numba compiler.
-Due to difficulties when passing a Python function as argument of precompiled Numba methods, we have decided to rewrite them in Cython language, keeping the homogeneous tesseroid calculation and adding the new variable density code.
+In a more detailed way, the new library approximates the gravitational effect of each tessseroid through the GLQ, and increases its accuracy with the modified version of the adaptive discretization algorithm, original developed by \citet{Li2011} and improved by \citet{Uieda2016}.
+
+Due to a new type of approximation error resulting from high variations in the density functions we had to develop a new density-based discretization algorithm that also intent to increase the accuracy, working together with the preexisting adaptive discretization algorithm.
 

### 2017-10-06 manuscript/manuscript.tex
Changes to density-based discretization algorithm in manuscript

@@
 High variations in the density function of an arbitrary tesseroid introduces a new type of error into the numerical approximation: only a few point masses would not reproduce the effect of this kind of variations.
-The modified version of the adaptive discretization algorithm may help in reducing this kind of error, but as it takes into account only the dimensions of the tesseroid and the distance to the computation point, it's not completely well suited to perform this task.
+The modified version of the adaptive discretization algorithm may help in reducing this kind of error, but as it takes into account only the dimensions of the tesseroid and the distance to the computation point, it's not well suited to fully perform this task.
 
 In order to overcome this, we developed a complementary discretization algorithm (it works together with the previous one) that takes into account the variation of the density inside the tesseroid.
-Basically, we will subdivide any tesseroid on the radius coordinate on the depth at which the ``maximum density variation'' takes place, creating two new tesseroids which we use to compute the desired gravity field by applying the previous adaptive discretization algorithm and the GLQ approximation.
+Basically, we will subdivide any tesseroid on the radius coordinate on the depth at which the \emph{maximum density variation} takes place, creating two new tesseroids which we use to compute the desired gravity field by applying the previous adaptive discretization algorithm and the GLQ approximation.
 
@@
 
-Then we will define the ``maximum density variation'' as the maximum difference between this density $\rho_0(r')$ and the original density $\rho(r')$ and obtain the radius coordinate $R_\text{max}$ at which this maximum takes place:
+Then we will define the \emph{maximum density variation} as the maximum difference between this density $\rho_0(r')$ and the original density $\rho(r')$ and obtain the radius coordinate $R_\text{max}$ at which this \emph{maximum} takes place:
 
@@
 Finally, we subdivide the original tesseroid at $R_\text{max}$ creating two new tesseroids with the same latitude and longitude boundaries.
-We then apply the modified version of the adaptive discretization algorithm and the GLQ to approximate the gravity fields that it generates on the computation points.
 
-This density-based discretization can be applied several times, creating new pair of subdivisions each time and therefore increasing the accuracy of the resulting gravity field.
+This density-based discretization can be applied several times, creating a new pair of subdivisions on each iteration and therefore increasing the accuracy of the resulting gravity field.
 Because each tesseroid is subdivided in two every time we apply this discretization, the number of subdivisions increases exponentially in power of two:
@@
 \noindent where $M$ is the order of the discretization, i.e. how many times the density-based discretization algorithm is applied on one tesseroid.
+
+It's worth notice that for $M=0$ or for linear densities, the density-based discretization algorithm will not be applied.
+In the first case because no subdivisions will be made, and in the second one due to the fact that the absolute difference of densities is equal to zero on every depth, thus no having a meaningful maximum.
 

### 2017-10-06 manuscript/manuscript.tex
Added algorithm summary in manuscript

@@
 
+\subsection{Algorithm summary}
+
+When computing any gravity field generated by an arbitrary tesseroid on a specific computation point, the new software:
+
+\begin{enumerate}
+    \item Applies the density-based discretization algorithm in case of variable density producing subdivisions of the original tesseroid,
+    \item Applies the modified adaptive discretization algorithm for each subdivision, generating a set of smaller tesseroids,
+    \item Computes the desired gravity field for every small tesseroid through the GLQ approximation (equation \ref{eq:glq-var-dens}).
+\end{enumerate}
+
+This steps are applied for each computation point, adding the contribution of each tesseroid to the resulting gravity field.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-10-17 manuscript/manuscript.tex
Edited determination of D with linear density in manuscript
Modified manuscript with new density-based discretization algorithm

@@
 
-In order to overcome this, we developed a complementary discretization algorithm (it works together with the previous one) that takes into account the variation of the density inside the tesseroid.
-Basically, we will subdivide any tesseroid on the radius coordinate on the depth at which the \emph{maximum density variation} takes place, creating two new tesseroids which we use to compute the desired gravity field by applying the previous adaptive discretization algorithm and the GLQ approximation.
+In order to overcome this, we developed a complementary discretization algorithm that takes into account the variation of the density inside the tesseroid.
+Basically, it subdivides any tesseroid on the radius coordinate on the depth at which the \emph{maximum density variation} takes place.
 
-First we need to define a simple density that approximates the one of the tesseroid: a linear function that assumes the same values of the original density on the tesseroid's top and bottom:
+Specifically, the algorithm consists in the following steps:
+
+\begin{enumerate}
+\renewcommand{\theenumi}{(\arabic{enumi})}
+    \item Normalise the density function to the [0, 1] interval,
+    \item Compute the absolute difference between the normalised density and a linear function that assumes the same values on the top and bottom radii that this normalised density.
+    \item If the maximum of the absolute difference is greater that a predefined delta ratio ($\delta$), then the tesseroid will be subdivided on the radius at which this maximum takes place.
+Otherwise, the tesseroid won't be subdivided.
+\end{enumerate}
+
+First we need to normalise the density function such as it assumes the values of 0 and 1 on the boundary radii ($R_1$, $R_2$), respectively, depending at which radius the original density function is maximum.
+We call this normalised density $\rho_n(r')$ and we can calculate it as follows:
 
 \begin{equation}
-    \rho_0(r') =
-    \frac{\rho(R_2) - \rho(R_1)}{R_2 - R_1} (r' - R_1) + \rho(R_1).
+    \rho_n(r') = \frac{\rho(r') - \rho_1}{\rho_2 - \rho_1},
+\end{equation}
+\noindent where
+\begin{equation}
+    \rho_1 = \text{min}\{ \rho(R_1), \rho(R_2) \}, \quad
+    \rho_2 = \text{max}\{ \rho(R_1), \rho(R_2) \}.
 \end{equation}
 
-Then we will define the \emph{maximum density variation} as the maximum difference between this density $\rho_0(r')$ and the original density $\rho(r')$ and obtain the radius coordinate $R_\text{max}$ at which this \emph{maximum} takes place:
+Secondly, we must define a linear function $\rho_l(r')$ that assumes the same values of the normalised density $\rho_n(r')$ on the boundary radii:
 
 \begin{equation}
-    \text{max}\{ | \rho(r') - \rho_0(r') | \} = 
-    | \rho(R_\text{max}) - \rho_0(R_\text{max}) |
+    \rho_l(r') =
+    \frac{ \rho_n(R_2) - \rho_n(R_1) }{ R_2 - R_1 } (r' - R_1) + \rho_n(R_1),
 \end{equation}
 
-Finally, we subdivide the original tesseroid at $R_\text{max}$ creating two new tesseroids with the same latitude and longitude boundaries.
-
-This density-based discretization can be applied several times, creating a new pair of subdivisions on each iteration and therefore increasing the accuracy of the resulting gravity field.
-Because each tesseroid is subdivided in two every time we apply this discretization, the number of subdivisions increases exponentially in power of two:
+\noindent and compute the absolute difference between this linear function and the normalised density:
 
 \begin{equation}
-    \text{Number of subdivisions} =
-    \sum\limits_{m=0}^M 2^m,
+    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |
 \end{equation}
 
-\noindent where $M$ is the order of the discretization, i.e. how many times the density-based discretization algorithm is applied on one tesseroid.
+Finally, if the maximum of the absolute difference if greater than a predefined delta ratio $\delta$ we will divide the tesseroid on the radius at which this maximum takes place ($R_\text{max}$):
 
-It's worth notice that for $M=0$ or for linear densities, the density-based discretization algorithm will not be applied.
-In the first case because no subdivisions will be made, and in the second one due to the fact that the absolute difference of densities is equal to zero on every depth, thus no having a meaningful maximum.
+\begin{equation}
+    \text{max}\{ \Delta \rho(r') \} > \delta
+    \label{eq:delta-density}
+\end{equation}
+\begin{equation}
+    \text{max}\{ \Delta \rho(r') \} = \Delta \rho(R_\text{max}),
+\end{equation}
+
+\noindent obtaining a pair of smaller tesseroids with the same longitude and latitude boundaries.
+
+In case that the original density function assumes the same value at both boundary radii, the procedure is the same except that the constants $\rho_1$ and $\rho_2$ are redefined as follows:
+
+\begin{equation}
+    \rho_1 = \text{min}\{ \rho(r') \}, \quad
+    \rho_2 = \text{max}\{ \rho(r') \}.
+\end{equation}
+
+If on the contrary the inequation \ref{eq:delta-density} is not held, then the tesseroid will not be subdivided.
+The same happens if the density function is constant or if the tesseroid radius dimension is bellow the threshold of 1mm.
+On the first case due to the fact that the density-based discretization algorithm is not needed, and on the second one because smaller tesseroids will not introduce meaningful increases in the accuracy of the computation.
+
+This algorithm can be subsequently applied to every subdividing tesseroid, until the inequation \ref{eq:delta-density} is not held or the subdivisions are smaller than 1mm.
+This two conditions guarantee that the algorithm is finite and does not fall into an infinite loop. \todo{chequear esta oracion}
+Once this conditions are met, every subdividing tesseroid is subjected to the modified adaptive discretization algorithm and the GLQ in order to compute the gravity field generated by it.
+
+It's easy to see that the delta ratio controls how many times the tesseroids will be subdivided based on the variation of the density function and therefore it controls the accuracy involving the numerical error that the density variation introduces.
+This raise the need of determine a maximum value of $\delta$ that ensures an acceptable accuracy.
+We will perform this determination with the same methodology as the one applied to the $D$ ratio: compare the numerical model with the analytical solutions for a spherical shell.
 
@@
 \begin{enumerate}
+    \renewcommand{\theenumi}{(\arabic{enumi})}
     \item Applies the density-based discretization algorithm in case of variable density producing subdivisions of the original tesseroid,
@@
 We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.
+
+It's worth to remember that the density-based discretization algorithm is not applied for linear density function comparisons, thus the adaptive discretization algorithm is the only mechanism that controls the accuracy of the approximation.
 

### 2017-10-18 manuscript/manuscript.tex
Changes in manuscript: modified determination of D

@@
 
-In order to test the new variable density tesseroid code and estimate the error of the approximation, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
+In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
 The first one constitutes the simplest variation, while the second one represents the fastest varying function.
@@
 crust and mantle interface).
-
-This error estimations are also useful in the determination of the minimum value for the distance-size ratio $D$ such the difference between the numerical model and the analytical solutions are below an acceptable threshold.
 
@@
 We computed these gravity fields  on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
-Then we calculated the maximum difference between these results and the value of the analytical solution at the same height.
+Then we calculated the maximum difference between these results and the value that the analytical solution assumes at the same height.
 
@@
 That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
+
+Moreover, we did not apply the density-based discretization algorithm in these tests in order to isolate the behaviour of the adaptive discretization algorithm.
 
@@
 \subsection{Linear Density}
-
-In first place, we proposed a spherical shell with linear density in such way that it has a density of 2670~kg/m$^3$ on the surface of the Earth, and 3300~kg/m$^3$ at 35 km of depth:
+Firstly,  we proposed a spherical shell with linear density in such way that it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:
 
@@
 \begin{equation}
-    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{35000\text{m}},
+    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R_2 - R_1},
 \end{equation}
@@
     c = \frac{3300\text{kg/m$^3$} - 
-        2670\text{kg/m$^3$}}{35000\text{m}} R + 
+        2670\text{kg/m$^3$}}{R_2 - R_1} R + 
         2670\text{kg/m$^3$},
@@
 \begin{equation}
-    \rho(r') = A e^{-(r' - R)/b}
+    \rho(r') = A \left( e^{-(r' - R)/b} - 1\right) + 2670 \text{kg/m}^3
 \label{eq:density-exp}
 \end{equation}
+\noindent where
+\begin{equation}
+    A =
+    (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
+    \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
+\end{equation}
 
-\noindent where $A$ = 1kg/m$^3$ and $b$ = 1000km.
-We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, whose results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+\noindent and $b$ = 1000km. 
+In this way, the shell has a density of 2670 kg/m$^3$ and 3300 kg/m$^3$ on its outer and inner radii, respectively.
+The analytical solution of the gravity potential is easily obtained as the sum the solution for a pure exponential function and the one for an homogeneous spherical shell.
+
+We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, without applying the density-based discretization algorithm.
+Its results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
 Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them, although, they can be found in the repository.
 
-In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
+In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
 
@@
 
-Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$.
-In order to do this we proposed a set of density functions, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900 kg/m$^{3}$ at the bottom of the spherical shell.
+Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$ when the density-based discretization algorithm is not applied.
+In order to do this we proposed a set of density functions, each one with a different value of $b$.
 We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the respective analytical solution.
@@
 
-This tests were performed using both the thin and the thick spherical shell.
-On Figures \ref{fig:D-exp-power-thin}a and \ref{fig:D-exp-power-thick}a we can see the density functions used to perform these tests for each spherical shell model, respectively, and their results can be seen in \ref{fig:D-exp-power-thin}b and \ref{fig:D-exp-power-thick}b.
-Taking into account that lower values of $b$ implies a more varying function, we can see that for extremely varying functions the numerical model does not perform well: for $b$=100m the difference is almost 100\% in the thin shell case, and the same happens with $b$=10km in the thick shell case.
+On Figure \ref{fig:D-exp-power-thin}a we can see the density functions used to perform these tests and its corresponding results on \ref{fig:D-exp-power-thin}b .
+Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b$=100m the difference is almost 100\%.
 
-On both cases, the model performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, in other words, when the absolute value of the exponent in equation \ref{eq:density-exp} is lower than one.
-Although this sets a limitation to the model, the maximum acceptable variation is very far from real density distributions inside the Earth.
-For example, if we take into account a tesseroid with 35km of thickness, a density of 2670kg/m$^3$ on its top surface and $b=35$km, then the density at its bottom will be approximately of 7000kg/m$^3$, very far from the values we can find at such depths.
+We can conclude that even when the density-based discretization algorithm is not applied, the numerical approximation performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, i.e.
+for low density variations.
+On the other hand, a density-based discretization algorithm is needed in order to guarantee the accuracy of the approximation in case of a high varying density function.
 
@@
 \caption{
-    Results of the accuracy tests for different exponential density functions in case of a thin spherical shell (1km of thickness).
+    Results of the accuracy tests for different exponential density functions in case of a spherical shell of 1km of thickness.
     (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
@@
 \label{fig:D-exp-power-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences-thick.pdf}
-\caption{
-    Results of the accuracy tests for different exponential density functions in case of a thick spherical shell (35km of thickness).
-    (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
-    (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
-Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-exp-power-thick}
 \end{figure}
@@
 
-\noindent where $\rho_1$ and $\rho_2$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radius on which the 
-density changes.
+\noindent where $\rho_1$ and $\rho_2$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radius on which the density change takes place.
 
-The analytical solution for this kind of density is the linear combination of two homogeneous shells with densities $\rho_1$ and $\rho_2$:
+The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_1$ and $\rho_2$:
 
@@
 
-Firstly, lets assume a symmetric density function, i.e. $R_c=(R_1 + R_2)/2$ is the centre radius of the shell, and perform the test for a thin shell of 1km of thickness.
+Firstly, lets assume a symmetric density function, i.e.
+$R_c=(R_1 + R_2)/2$ is the centre radius of the shell, and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
 The results can be seen in Figure \ref{fig:D-discont-sym}.
-Again, we obtain good accuracy for values of $D$ equal to 1, 2 and 8 for the computation of the potential, the gradient components and the Marussi tensor components, respectively.
+Once again, we obtain good accuracy for values of $D$ equal to 1, 2 and 8 for the computation of the potential, the gradient components and the Marussi tensor components, respectively.
 
@@
 
-From equation \ref{eq:glq-var-dens} we have seen that the information of the density function is summarised as the values its assumes on the nodes of the GLQ.
-We can attribute the low performance of the code to this simplification.
-
-On the other hand, we can explain the good performance in the symmetrical density case due to the fact that the GLQ nodes are always symmetrical with respect to the centre of the $[a,b]$ integration interval (see equation \ref{eq:glq-scaled}).
-
-In the light of these results we do not recommend using discontinuous densities functions.
-Instead, divide the tesseroid into several ones with specific homogeneous density or density function for each one.
+As we did with the high varying exponential densities, we can conclude that for arbitrary discontinuous functions a density-based discretization algorithm is also needed in order to guarantee a good performance.
 

### 2017-10-19 manuscript/manuscript.tex
Changes in manuscript

@@
 
-We have enhanced the later implementation with a new forward gravity model that allows us to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
-It is based on the GLQ approximation, makes use of the modified distance-size discretization algorithm and a new density-based discretization that increases the accuracy for high variating density functions.
+In addition, the varying density inside an arbitrary tesseroid introduces a new kind of error into the numerical approximation.
+This encouraged us to develop a new density-based discretization algorithm that divides the tesseroid into smaller ones taking into account how the density function varies.
+The number of subdivisions and thus the accuracy of the approximation will be controlled by a delta ratio $\delta$, a predefined value by the user.
+
+We have enhanced the implementation of \citet{Uieda2016} with a new forward gravity model that allows us to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
+It is based on the GLQ approximation, makes use of the modified distance-size discretization and the new density-based discretization algorithms that guarantee the accuracy even for high varying density functions.
 It is written in Python and Cython languages, constituting an easy to use library that runs precompiled code for the more time consuming functions.
@@
 
-In the following sections we will describe how the new algorithm works, its theoretical background and obtain the distance-size ratio values needed to get a good accuracy in the computation.
+In the following sections we will describe how the new algorithm works, its theoretical background and obtain default distance-size ratio and delta ratio values needed to get an acceptable accuracy in the numerical approximation.
 
@@
 \begin{equation}
-    \int\limits_{-1}^1 f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i),
+    \int\limits_{-1}^1 f(x) dx \approx \sum_{i=1}^N W_i f(x_i),
 \end{equation}
@@
 From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates the gravity fields of a tesseroid with variable density as the ones generated by point masses located at the nodes of the Legendre polynomials.
-Furthermore, the information of the density function is summarised as the values it assumes on those nodes.
-Although this may sound discouraging, we intent to prove otherwise and show that this method, along with a well fitted discretization algorithm, approximates the gravity fields with good accuracy and precision.\todo{no se si poner esta ultima oracion}
+Furthermore, we can also conclude that the information of the density function is summarised as the values it assumes on those nodes.
+Although this may sound discouraging, we intent to prove otherwise and show that this method, along with well fitted discretization algorithms, approximates the gravity fields with good accuracy.\todo{no se si poner esta ultima oracion}
 
@@
 
-\noindent and the integration in $\phi'$ can be performed independently of the density function.
-Using \citet{sagemath} we obtain the following expression:
+\noindent while the integration in $\phi'$ can be performed independently of the density function.
+Using \citet{sagemath} we obtained the following expression:
 
@@
 
-The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity field generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axe.
+The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity potential generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axe.
 Due to the rotational symmetry along any axe that passes through the centre of the shell, the integral in equation \ref{eq:shell-pot} reproduces the gravity potential field on any outside point at distance $r$ from the its centre.
 
-The gradient and the Marussi tensor derived from potentials that depends solely on $r$ has only a few non zero components: the vertical component of the gradient ($g_z$) and the diagonal components of the tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
-They can be easily obtained as \citet{Grombein2013} does, even for any density function $\rho(r')$ of our choice:
+The gradient and the Marussi tensor derived from potentials that depends solely on $r$ have only a few non zero components: the vertical component of the gradient ($g_z$) and the diagonal components of the tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
+They can be easily obtained as \citet{Grombein2013} does, even for any arbitrary density function $\rho(r')$:
 
@@
 
-Due to a new type of approximation error resulting from high variations in the density functions we had to develop a new density-based discretization algorithm that also intent to increase the accuracy, working together with the preexisting adaptive discretization algorithm.
+In addition, we had to develop a new density-based discretization algorithm in order to decrease a new type of error in the numerical approximation caused by high variations in the density functions.
+It is intended to work together with the preexisting modified adaptive discretization algorithm.
 
@@
 
-On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy is a matter of study.
+On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy must be objectively determined beforehand.
+In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
 
-In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
-We have performed similar tests, but now with variable density spherical shell.
+We will perform similar tests, but now with variable density spherical shell.
 From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
@@
 High variations in the density function of an arbitrary tesseroid introduces a new type of error into the numerical approximation: only a few point masses would not reproduce the effect of this kind of variations.
-The modified version of the adaptive discretization algorithm may help in reducing this kind of error, but as it takes into account only the dimensions of the tesseroid and the distance to the computation point, it's not well suited to fully perform this task.
+The modified version of the adaptive discretization algorithm may help in reducing this kind of error, but as it only takes into account the dimensions of the tesseroid and the distance to the computation point, it's not well suited to fully perform this task.
 
@@
     \item If the maximum of the absolute difference is greater that a predefined delta ratio ($\delta$), then the tesseroid will be subdivided on the radius at which this maximum takes place.
-Otherwise, the tesseroid won't be subdivided.
+Otherwise, the tesseroid won't be split.
 \end{enumerate}
 
-First we need to normalise the density function such as it assumes the values of 0 and 1 on the boundary radii ($R_1$, $R_2$), respectively, depending at which radius the original density function is maximum.
-We call this normalised density $\rho_n(r')$ and we can calculate it as follows:
+First we need to normalise the density function such as it assumes the values of 0 and 1 on the boundary radii ($R_1$, $R_2$), depending at which radius the original density function is maximum.
+We call this normalised density $\rho_n(r')$ and we calculate it as follows:
 
@@
 This algorithm can be subsequently applied to every subdividing tesseroid, until the inequation \ref{eq:delta-density} is not held or the subdivisions are smaller than 1mm.
-This two conditions guarantee that the algorithm is finite and does not fall into an infinite loop. \todo{chequear esta oracion}
+This two conditions guarantee that the algorithm is finite, i.e. it does not fall into an infinite loop. \todo{chequear esta oracion}
 Once this conditions are met, every subdividing tesseroid is subjected to the modified adaptive discretization algorithm and the GLQ in order to compute the gravity field generated by it.
 
-It's easy to see that the delta ratio controls how many times the tesseroids will be subdivided based on the variation of the density function and therefore it controls the accuracy involving the numerical error that the density variation introduces.
-This raise the need of determine a maximum value of $\delta$ that ensures an acceptable accuracy.
+It's easy to see that the delta ratio indirectly controls how many times the tesseroids will be subdivided based on the variation of the density function.
+Therefore it controls the accuracy involving the numerical error that the density variation introduces and also the computation time.
+This raises the need to determine a maximum value of $\delta$ that ensures an acceptable accuracy while minimising the computation time.
 We will perform this determination with the same methodology as the one applied to the $D$ ratio: compare the numerical model with the analytical solutions for a spherical shell.
@@
 
-This steps are applied for each computation point, adding the contribution of each tesseroid to the resulting gravity field.
+These steps are applied for each computation point, adding the contribution of each tesseroid to the resulting gravity field.
 
@@
 In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
-The first one constitutes the simplest variation, while the second one represents the fastest varying function.
+The first one constitutes the simplest variation, while the second one represents a higher varying function.
 The later one tries to emulate a contact between two different media (e.g.
@@
 We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
-More information about this grids can be found in the Table \ref{tab:grids}.
+More information about this grids can be found on Table \ref{tab:grids}.
 
@@
 \begin{equation}
-    \rho(r') = A \left( e^{-(r' - R)/b} - 1\right) + 2670 \text{kg/m}^3
+    \rho(r') = A \left( e^{-(r' - R)/b} - 1\right) + 2670 \text{kg/m}^3,
 \label{eq:density-exp}
@@
 
-\noindent and $b$ = 1000km. 
+\noindent and the value of $b$ depends on the spherical shell on which we performed the test. 
 In this way, the shell has a density of 2670 kg/m$^3$ and 3300 kg/m$^3$ on its outer and inner radii, respectively.
@@
 We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, without applying the density-based discretization algorithm.
-Its results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+On both cases, the constant $b$ is equal to the thickness of the shell: 1km for the thin shell and 35km for the thick one.
+The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
 Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them, although, they can be found in the repository.
 
-In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
-
-Surprisingly both the linear and the exponential density tesseroid models needs the same values of distance-size ratio $D$ in order to assure a good accuracy.
+In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
 
@@
 In order to do this we proposed a set of density functions, each one with a different value of $b$.
-We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the respective analytical solution.
+We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the corresponding analytical solution.
 These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.
@@
 On Figure \ref{fig:D-exp-power-thin}a we can see the density functions used to perform these tests and its corresponding results on \ref{fig:D-exp-power-thin}b .
-Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b$=100m the difference is almost 100\%.
+Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b=100$m the difference is almost 100\%.
 
@@
     \begin{cases}
-        \rho_1 & \text{if} \, R_1 \leq r' < R_c \\
-        \rho_2 & \text{if} \, R_c \leq r' \leq R_2 \\
+        \rho_A & \text{if} \, R_1 \leq r' < R_c \\
+        \rho_B & \text{if} \, R_c \leq r' \leq R_2 \\
     \end{cases}
@@
 
-\noindent where $\rho_1$ and $\rho_2$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radius on which the density change takes place.
+\noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radius on which the density change takes place.
 
-The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_1$ and $\rho_2$:
+The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_A$ and $\rho_B$:
 
@@
     V(r) = \frac{4}{3} \frac{\pi G}{r}
-    \left[ \rho_1 (R_c^3 - R_1^3) +
-    \rho_2 (R_2^3 - R_c^3) \right]
+    \left[ \rho_A (R_c^3 - R_1^3) +
+    \rho_B (R_2^3 - R_c^3) \right]
 \end{equation}
@@
 
+\section{Determination of delta ratio}
+
+In the light of the previous results, a density-based discretization algorithm is needed in order to achieve an acceptable accuracy in those cases on which the modified adaptive discretization algorithm is not sufficient to take into account the numerical error due to density variations.
+
+Besides the implementation of this new algorithm, a determination of the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold is also needed in order to guarantee the good performance of the numerical model.
+
+To do this we applied a similar methodology used in the determination of the distance-size ratio $D$: we proposed a set of exponential density functions (as the one in equation \ref{eq:density-exp}) with different values of $b$ and computed the difference between the gravity fields generated by the numerical model with the corresponding analytical solution, each time with a different value of $\delta$ ranging from 0.1 to 1 with a step of 0.1.
+This tests have been performed for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Speed Comparison}
 
-We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous ($\Delta t_\text{homogeneous}$) and variable ($\Delta
-t_\text{variable}$) density at different heights. Then we plotted the ratio between these computation times (see Figure \ref{fig:speed-comparison}):
+We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous ($\Delta t_\text{homogeneous}$) and variable ($\Delta t_\text{variable}$) density at different heights.
+Then we plotted the ratio between these computation times (see Figure \ref{fig:speed-comparison}):
 

### 2017-10-20 manuscript/manuscript.tex
Figures names changed in manuscript

@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Dlinear-thin-differences.pdf}
+\includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Dlinear-thick-differences.pdf}
+\includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thin-differences.pdf}
+\includegraphics[width=0.9\linewidth]{figures/exponential-D-thin.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-shifted-thick-differences.pdf}
+\includegraphics[width=0.9\linewidth]{figures/exponential-D-thick.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Dexp-power-differences-thin.pdf}
+\includegraphics[width=0.9\linewidth]{figures/exponential-b-thin.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-symmetric-differences.pdf}
+\includegraphics[width=0.9\linewidth]{figures/discontinuous-D-sym.pdf}
 \caption{
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/Ddiscontinuous-asymmetric-differences.pdf}
+\includegraphics[width=0.9\linewidth]{figures/discontinuous-D-asym.pdf}
 \caption{

### 2017-10-20 manuscript/manuscript.tex
Edited manuscript

@@
     Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
@@
     Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
@@
     Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
@@
     Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
+The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
@@
 \caption{
-    Results of the accuracy tests for different exponential density functions in case of a spherical shell of 1km of thickness.
+    Results of the distance-size ratio determination for different exponential density functions in case of a spherical shell of 1km of thickness.
     (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
@@
 This tests have been performed for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
+\caption{
+    Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
+    (a) Density functions used to perform the test.
+    (b) Differences between the numerical model and the corresponding analytical solution for several values of $\delta$ and $b$.}
+\label{fig:exponential-delta-thin}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
+\caption{
+    Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
+    (a) Density functions used to perform the test.
+    (b) Differences between the numerical model and the corresponding analytical solution for several values of $\delta$ and $b$.}
+\label{fig:exponential-delta-thick}
+\end{figure}
+
+On Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a we can see the exponential density functions used on each test for a thin and thick spherical shell, respectively.
+While Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b show their results.
+
+In the thin shell case we can conclude that a delta equal to 0.3 guarantee an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
+We will keep the more conservative value of 0.2 as the default value of delta on future computations.
 

### 2017-10-23 manuscript/manuscript.tex
Edited manuscript

@@
 \begin{equation}
-    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |
+    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |.
 \end{equation}
@@
 \begin{equation}
-    \text{max}\{ \Delta \rho(r') \} > \delta
+    \text{max}\{ \Delta \rho(r') \} > \delta,
     \label{eq:delta-density}
@@
 
-\noindent and the value of $b$ depends on the spherical shell on which we performed the test. 
-In this way, the shell has a density of 2670 kg/m$^3$ and 3300 kg/m$^3$ on its outer and inner radii, respectively.
+\noindent and the value of $b$ equal to the thickness of the spherical shell we use to perform the test.
+In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
 The analytical solution of the gravity potential is easily obtained as the sum the solution for a pure exponential function and the one for an homogeneous spherical shell.
@@
 We can see its results in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what
-leads us to conclude that the code does not recreates the gravity fields for this case with good accuracy.
+leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.
 
@@
 
-As we did with the high varying exponential densities, we can conclude that for arbitrary discontinuous functions a density-based discretization algorithm is also needed in order to guarantee a good performance.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Besides the implementation of this new algorithm, a determination of the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold is also needed in order to guarantee the good performance of the numerical model.
+Besides the implementation of this new algorithm, we must determine the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold in order to guarantee the good performance of the numerical model.
 
-To do this we applied a similar methodology used in the determination of the distance-size ratio $D$: we proposed a set of exponential density functions (as the one in equation \ref{eq:density-exp}) with different values of $b$ and computed the difference between the gravity fields generated by the numerical model with the corresponding analytical solution, each time with a different value of $\delta$ ranging from 0.1 to 1 with a step of 0.1.
+To do this we applied a similar methodology used in the determination of the distance-size ratio $D$: we proposed a set of exponential density functions (as the one in equation \ref{eq:density-exp}) with different values of $b$, and computed the difference between the gravity fields generated by the numerical model with the corresponding analytical solution, each time with a different value of $\delta$ ranging from 0.1 to 1 with a step of 0.1.
 This tests have been performed for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
@@
 On Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a we can see the exponential density functions used on each test for a thin and thick spherical shell, respectively.
-While Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b show their results.
+While Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b show their corresponding results.
 
@@
 We will keep the more conservative value of 0.2 as the default value of delta on future computations.
+
+Furthermore, we computed the difference between the numerical model and the analytical solution in case of a discontinuous density (as the one in equation \ref{eq:density-discontinuous} with $R_c = (R_1 + R_2)/3$) using a $\delta = 0.2$.
+The result for the $g_{zz}$ component of the Marussi tensor is bellow the 0.1\% threshold, achieving a good performance for this default value of $\delta$.
 
@@
 It essentially consists in approximating the gravity fields of the tesseroids with the effect of point masses located on the scaled nodes of the Legendre polynomials.
-We have stated that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involve Taylor series expansion.
+We state that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involve Taylor series expansion.
 
-Instead of enhancing the accuracy of the method by increasing the order of the GLQ on each integration, we make use of the modified adaptive discretization algorithm.
-It divides the tesseroid by half if the ratio of the distance to the computation point and its size is lower than a predefined distance-size ratio $D$.
+Instead of enhancing the accuracy of the method by increasing the order of the GLQ on each integration, we make use of the preexisting modified adaptive discretization algorithm and a new density-based discretization algodithm.
+
+The former one divides the tesseroid by half if the ratio of the distance to the computation point and its size is lower than a predefined distance-size ratio $D$.
 We can summarise that $D$ controls the accuracy of the method: a higher value of $D$ generates more divisions of the tesseroids, thus more point masses, what leads to a more precise approximation.
+Nevertheless it is not sufficient to guarantee the accuracy of the method in case of specific density functions as a highly variable exponential or some discontinuous functions.
 
-On the other hand, there's no direct relation between the value of $D$ and the error of the computation, that's why we had to determine the minimum value of $D$ that produces an acceptable accuracy for each gravity magnitude.
-These distance-size ratio determinations have been made through the comparison of the numerical approximation with the analytical solutions for the exceptional case of a spherical shell.
+For this reason we developed a density-based discretization algorithm that divides the tesseroid on the radius at which the \emph{maximum variation} of the density function takes place, only if a inequation that relates this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
+Therefore the number of subdivisions is indirectly controlled by this delta ratio $\delta$, and thus the accuracy of the computation.
+
+Because there is no direct relation between the values of $D$ and $\delta$ and the error of the computation, we had to empirically determine the minimum value of $D$ and the maximum value of $\delta$ that produce an acceptable accuracy for each gravity field.
+
+The distance-size ratio determinations have been made through the comparison of the numerical approximations with the analytical solutions for the a spherical shell with variable density, which had to be analytically obtained.
 From the wide range of possible density variations, we've chosen a linear, an exponential and a discontinuous density functions.
+These comparisons have been made without applying the density-based discretization algorithm in order to isolate the other one.
+The tests over the first two functions have shown that a value of $D$ equal to 1, 2 and 8 must be used in the calculation of the gravity potential, its gradient components and the components of the Marussi tensor, respectively, in order to achieve an accuracy bellow 0.1\%.
+On the other hand we have found a limitation of this algorithm working on its own in case of highly varying exponential functions and in some cases of discontinuous density.
 
-The tests over the first two functions have shown that in order to achieve an accuracy of 0.1\%, a value of $D$ equal to 1, 2 and 8 must be used in the calculation of the gravity potential, its gradient components and the components of the Marussi tensor, respectively.
-We have also found a limitation on the variation of the exponential density function: the absolute value of the exponent must not be greater than 1.
-However, those extremely variations cannot be found inside the Earth, so this limitation does not determines a true restriction for real life models.
-
-On the other hand, the tests over the discontinuous function were not satisfactory.
-In such cases we recommend to divide the model in several tesseroids, each one with a specific homogeneous density or a continuous density function, instead of using a single variable density tesseroid.
+Nevertheless, the density-based discretization algorithm have shown that a value of $\delta = 0.2$ (together with the previously obtained default values of $D$) guarantees a good accuracy on those cases on which the modified adaptive discretization algorithm does not achieve it on its own.
 

### 2017-10-24 manuscript/manuscript.tex
Changes in manuscript

@@
 These forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
-Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allows geologists and geophysics to perform modelling and interpretation on regional scales.
+Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allow geologists and geophysics to perform modelling and interpretation on regional scales.
 This raises the importance of building forward models that reproduce the gravity anomalies for such scales.
@@
 The literature offers two approaches: one involves Taylor series expansion \citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
-The later consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials.
+
+In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
+On the other hand, the GLQ consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials, where the density information is gathered into these weights.
+Thus, this kind of numerical approximation allows to calculate the gravity fields for any density function without any other information but the function itself.
 
@@
 
-In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
-On the other hand, the GLQ allows us to calculate them for any density function without any other information but the function itself.
-
 \citet{Ku1977} documented that the GLQ becomes less accurate when the computation point is closer to the tesseroid or for smaller GLQ order. 
@@
 It is written in Python and Cython languages, constituting an easy to use library that runs precompiled code for the more time consuming functions.
+It also makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
+Furthermore, we have the intention to include it into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
 
@@
 Furthermore we had to derive the analytical expressions for these density functions.
-
-This new forward gravity model makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
-Furthermore, we have the intention to include it into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
 
@@
 We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented Cartesian coordinate system located at $P$.
-In the special case of homogeneous density, the gravity potential, its gradient and the Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[see also][]{Uieda2016}.
+In the special case of homogeneous density, the gravity potential, its gradient and the Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[for same notation as the one we will use, see][]{Uieda2016}.
 
@@
 \noindent $\alpha, \beta \in \{x, y, z\}$,
-$\rho(r')$ is the density function that depends on the radius coordinate,
+$\rho(r')$ is the density function that depends on the radial coordinate,
 $\delta_{\alpha\beta}$ is Kronecker's delta,
@@
 \begin{equation}
-    \ell = \sqrt{{r'}^2 + r^2 - 2 r' r \cos \psi},
+    \ell = \sqrt{{r'}^2 + r^2 - 2 r r' \cos \psi},
 \label{eq:ell}
@@
 
-\noindent where $x_i^s$ is the scaled $x_i$ root in the $[a,b]$ interval, also called nodes of the quadrature:
+\noindent where $x_i^s$ is the scaled $x_i$ root into the $[a,b]$ interval, also called nodes of the quadrature:
 
@@
 
-For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$.
+\noindent For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$.
 
@@
 
-From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates the gravity fields of a tesseroid with variable density as the ones generated by point masses located at the nodes of the Legendre polynomials.
-Furthermore, we can also conclude that the information of the density function is summarised as the values it assumes on those nodes.
+In the same way as for an homogeneous density, we can observe from equation \ref{eq:glq-var-dens} that the GLQ approximates the gravity fields of a tesseroid as the ones generated by point masses located at the nodes of the Legendre polynomials.
+Nevertheless, in the variable density case, the information of the density function is summarised into the weights of the sum by taking into account only the values it assumes on those nodes.
 Although this may sound discouraging, we intent to prove otherwise and show that this method, along with well fitted discretization algorithms, approximates the gravity fields with good accuracy.\todo{no se si poner esta ultima oracion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     \cos\psi = \sin\phi', \quad
-    \ell = \sqrt{r'^2 - 2 r r' \sin\phi' + r^2}.
+    \ell = \sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}.
 \end{equation}
@@
     \int\limits_{R_1}^{R_2}
-    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 - 2 r r' \sin\phi' + r^2}}
+    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}}
     \, dr' d\phi',
@@
 \noindent while the integration in $\phi'$ can be performed independently of the density function.
-Using \citet{sagemath} we obtained the following expression:
+Taking into account the following integral:
+
+\begin{equation}
+    \int \frac{\cos \phi}{\sqrt{a - b \sin \phi}} \, d\phi =
+    -\frac{2}{b} \sqrt{a - b \sin \phi} + C,
+\end{equation}
+
+\noindent the potential can be written as follows:
 
@@
 
-If the density of the spherical shell is a linear function on the radius coordinate, i.e.:
+If the density of the spherical shell is a linear function on the radial coordinate, i.e.:
 
@@
 
-If, on the other hand, the spherical shell has an exponential density in the radius coordinate,
+If, on the other hand, the spherical shell has an exponential density in the radial coordinate,
 
@@
 In order to overcome this, we developed a complementary discretization algorithm that takes into account the variation of the density inside the tesseroid.
-Basically, it subdivides any tesseroid on the radius coordinate on the depth at which the \emph{maximum density variation} takes place.
+Basically, it subdivides any tesseroid on the radial direction on the depth at which the \emph{maximum density variation} takes place.
 

### 2017-10-25 manuscript/manuscript.tex
Modifications on manuscript

@@
 The main issue that should be overcome is taking into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
-A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radius boundaries (see Figure \ref{fig:tesseroid-uieda}).
+A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radial boundaries (see Figure \ref{fig:tesseroid-uieda}).
 Calculating the gravity fields generated by an arbitrary tesseroid on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
@@
 
-In our case, we will assume that the tesseroid has an arbitrary variable density in depth, i.e. only depends on the radius spherical coordinate.
+In our case, we will assume that the tesseroid has an arbitrary variable density in depth, i.e. only depends on the radial spherical coordinate.
 Thus, the integrals for the gravity fields are slightly modified:
@@
 
-In the same way as for an homogeneous density, we can observe from equation \ref{eq:glq-var-dens} that the GLQ approximates the gravity fields of a tesseroid as the ones generated by point masses located at the nodes of the Legendre polynomials.
+In the same way as for an homogeneous density, we can observe from equation \ref{eq:glq-var-dens} that the GLQ approximates the gravity fields of a tesseroid as the ones generated by $N_\lambda N_\phi N_r$ point masses located at the nodes of the Legendre polynomials, where $N_i$ are the orders of the quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$.
 Nevertheless, in the variable density case, the information of the density function is summarised into the weights of the sum by taking into account only the values it assumes on those nodes.
@@
 
-Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, whose density depends on the radius spherical coordinate (Figure \ref{fig:spherical-shell}).
+Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, whose density depends on the radial spherical coordinate (Figure \ref{fig:spherical-shell}).
 The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$, located along the $z$ axe at a distance $r$ from the origin, can be written as follows:
@@
 Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
-The computation point $Q$ is located in the $z$ axe at a distance $r$ from the centre of the shell.
+The computation point $Q$ is located on the $z$ axe at a distance $r$ from the centre of the shell.
 For practical purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
@@
 \noindent while the integration in $\phi'$ can be performed independently of the density function.
-Taking into account the following integral:
-
-\begin{equation}
-    \int \frac{\cos \phi}{\sqrt{a - b \sin \phi}} \, d\phi =
-    -\frac{2}{b} \sqrt{a - b \sin \phi} + C,
-\end{equation}
-
-\noindent the potential can be written as follows:
+Making use of SymPy \citep{sympy2017}, a Python library for symbolic mathematics, we obtained the following expression for the potential:
 
@@
 
-For our purposes we can assume that the computation point $Q$ is outside the outer radii, i.e. $r>R_2$ ($R_2 \geq r'$), and simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
+For our purposes we can assume that the computation point $Q$ is farther from the origin than the outer radius, i.e. $r>R_2$. 
+And taking into account that $r'$ is always lower than $R_2$, we can simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
 
@@
 
-\noindent what leads to the following expression of the potential of the spherical shell:
+\noindent what leads to the following expression of the potential:
 
@@
 
-The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity potential generated by a spherical shell with a variable density in depth on any outside point located in the $z$ axe.
+The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity potential generated by a spherical shell with a variable density in depth on any outside point located on the $z$ axe.
 Due to the rotational symmetry along any axe that passes through the centre of the shell, the integral in equation \ref{eq:shell-pot} reproduces the gravity potential field on any outside point at distance $r$ from the its centre.
@@
 
-For practical purposes we are going to obtain the gravity potential for three different functions from the perspective of their variation in $r'$: a linear and an exponential density function.
+For future purposes we are going to obtain the gravity potential for three different functions from the perspective of their variation in $r'$: a linear and an exponential density function.
 
@@
 
-The second term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
-So it can be read as the combination of a spherical shell with the later homogeneous density and the same shell with variable density $\rho(r') = ar'$.
+The first term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with variable density $\rho(r') = ar'$, while the second one constitutes the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
+So the potential can be read as the one generated by a combination of both shells.
 
@@
 
-\noindent the potential can be written as follows:
+\noindent the potential (calculated through SymPy \citep{sympy2017}) can be written as follows:
 
@@
         V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-        Ab \, e^\frac{\Delta h}{b}
+        Ab 
         \Big[
-        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
+        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
         \Big].
@@
 
-This new library is based on the preexisting code developed by \citet{Uieda2016} (and included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that is able to compute only the gravitational effects of tesseroids with homogeneous density.
-\citet{Uieda2016} have optimised the more time consuming functions through the Numba compiler.
+This new library approximates the gravitational effect of each tesseroid through the GLQ, and guarantees its accuracy with the modified version of the adaptive discretization algorithm, original developed by \citet{Li2011} and improved by \citet{Uieda2016}, and a new density-based discretization algorithm.
+The later one had to be developed in order to decrease a new type of error in the numerical approximation caused by high variations in the density functions.
+
+Moreover, the new forward model is based on the preexisting code developed by \citet{Uieda2016} (and included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that is able to compute the gravitational effects of tesseroids with homogeneous density.
+On that implementation \citet{Uieda2016} have optimised the more time consuming functions through the Numba compiler.
 In contrast, we have decided to develop our new library in Cython language due to difficulties when passing a Python function as argument to precompiled Numba methods, keeping the homogeneous tesseroid calculation and adding the new variable density computations.
@@
 
-Moreover, this new software is compatible with the preexisting open-source library Fatiando a Terra and its usage is very similar to the one for homogeneous tesseroids:
+It is also compatible with the preexisting open-source library Fatiando a Terra and its usage is very similar to the one for homogeneous tesseroids:
 
@@
 
-In a more detailed way, the new library approximates the gravitational effect of each tessseroid through the GLQ, and increases its accuracy with the modified version of the adaptive discretization algorithm, original developed by \citet{Li2011} and improved by \citet{Uieda2016}.
-
-In addition, we had to develop a new density-based discretization algorithm in order to decrease a new type of error in the numerical approximation caused by high variations in the density functions.
-It is intended to work together with the preexisting modified adaptive discretization algorithm.
-
 \subsection{Modified Adaptive Discretization Algorithm}
 
-From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of $N_\lambda N_\phi N_r$ point masses, where $N_i$ are the orders of the quadrature for the integral on each coordinate, with $i \in \{ \lambda, \phi, r \}$.
+From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of point masses located at the nodes of the quadrature.
 If we set the GLQ orders to fixed values, it's easy to conclude that the nearer the computation point is to the tesseroid, the accuracy of the approximation is reduced and the point masses effect becomes more noticeable than the tesseroid we want to approximate.
@@
 If on the contrary the inequation \ref{eq:delta-density} is not held, then the tesseroid will not be subdivided.
-The same happens if the density function is constant or if the tesseroid radius dimension is bellow the threshold of 1mm.
+The same happens if the density function is constant or if the tesseroid radial dimension is bellow the threshold of 1mm.
 On the first case due to the fact that the density-based discretization algorithm is not needed, and on the second one because smaller tesseroids will not introduce meaningful increases in the accuracy of the computation.
@@
 
-\noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the the radius on which the density change takes place.
+\noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the radius on which the density change takes place.
 
@@
 
-In the thin shell case we can conclude that a delta equal to 0.3 guarantee an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
+In the thin shell case we can conclude that a delta equal to 0.3 guarantees an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
 We will keep the more conservative value of 0.2 as the default value of delta on future computations.

### 2017-10-25 manuscript/manuscript.tex
Modified manuscript

@@
 
-In order to make the computation more efficient, we use only second order GLQ for each integration while we ensure the accuracy of the method through the application of the modified adaptive discretization algorithm developed by \citet{Uieda2016}.
-The modifications made by \citet{Uieda2016} to the original adaptive discretization method \citep{Li2011} can be summed up in:
+The main differences between the modified version made by \citet{Uieda2016} and the original one developed by \citet{Li2011} can be summed up in:
 (1) a faster calculation of the distance from the computation point to the tesseroid and 
 (2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
-
 More specifically, \citet{Uieda2016} redefined the dimensions of an arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$.
@@
 
-The adaptive discretization algorithm starts by checking that the following inequality is satisfied:
+While keeping the GLQ fixed on second orders, the adaptive discretization algorithm starts by checking that the following inequality is satisfied:
 
@@
 \noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point, and $D$ is a predefined value called distance-size ratio.
-If the inequality \ref{eq:distance-size-ratio} is not held for any spherical coordinate, then the tesseroid must be divided by half in that direction.
+If the inequality \ref{eq:distance-size-ratio} is not held for at least one spherical coordinate, then the tesseroid must be divided by half in that direction or directions.
 A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
@@
 
-In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided, and therefore regulates both the accuracy of the algorithm and its computation time.
-In this way the modified version of the adaptive discretization algorithm efficiently improves its accuracy and gives the user more control over the computation: it allows to stop the division of the tesseroids when their dimensions are lower than a certain value (e.g.
-less than 10 cm), in which case it does not modifies significantly the desired gravitational field.
+In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided, and therefore efficiently regulates both the accuracy of the algorithm and its computation time.
+Moreover, it limits the computation time by stopping the division of the tesseroids when their dimensions are lower than a certain value (1mm), assuming that such small tesseroids do not modifies significantly the desired gravitational field.
 
@@
 This algorithm can be subsequently applied to every subdividing tesseroid, until the inequation \ref{eq:delta-density} is not held or the subdivisions are smaller than 1mm.
-This two conditions guarantee that the algorithm is finite, i.e. it does not fall into an infinite loop. \todo{chequear esta oracion}
+This two conditions guarantee that the algorithm is finite.
 Once this conditions are met, every subdividing tesseroid is subjected to the modified adaptive discretization algorithm and the GLQ in order to compute the gravity field generated by it.
@@
 It's easy to see that the delta ratio indirectly controls how many times the tesseroids will be subdivided based on the variation of the density function.
-Therefore it controls the accuracy involving the numerical error that the density variation introduces and also the computation time.
+Therefore it determines the accuracy involving the numerical error that the density variation introduces and also the computation time.
 This raises the need to determine a maximum value of $\delta$ that ensures an acceptable accuracy while minimising the computation time.
@@
 In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
-The first one constitutes the simplest variation, while the second one represents a higher varying function.
-The later one tries to emulate a contact between two different media (e.g.
+The first one constitutes the simplest variation, the second one represents a higher varying function, while the later one tries to emulate a contact between two different media (e.g.
 crust and mantle interface).
@@
 
-We computed these gravity fields on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5.
+We computed these gravity fields on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, without applying the density-based discretization algorithm in order to isolate the behaviour of the adaptive discretization one.
 Then we calculated the maximum difference between these results and the value that the analytical solution assumes at the same height.
@@
 That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
-
-Moreover, we did not apply the density-based discretization algorithm in these tests in order to isolate the behaviour of the adaptive discretization algorithm.
 
@@
 \subsection{Linear Density}
-Firstly,  we proposed a spherical shell with linear density in such way that it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:
+Firstly,  we proposed a spherical shell with linear density in such way it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:
 
@@
 
-In Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick} we can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor.
-Because the other two diagonal components of the tensor show similar results as the ones obtained for the $g_{zz}$, we exclude them from the figure, although they can be found in the repository.
+We can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
+We exclude the results corresponding to the other two diagonal components of the tensor because they are similar to the ones obtained for the $g_{zz}$, although they can be found on the repository.
 \todo{chequear}
@@
 We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.
-
-It's worth to remember that the density-based discretization algorithm is not applied for linear density function comparisons, thus the adaptive discretization algorithm is the only mechanism that controls the accuracy of the approximation.
 
@@
 In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
-The analytical solution of the gravity potential is easily obtained as the sum the solution for a pure exponential function and the one for an homogeneous spherical shell.
+The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function and for an homogeneous spherical shell.
 
@@
 Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them, although, they can be found in the repository.
-
-In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
 
@@
 
+In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
+
+Moreover, we can see that the differences for the potential and the $g_z$ component of the gradient keep constant even for high values of $D$.
+Our interpretation is that the modified adaptive discretization algorithm is limited when increasing the accuracy for this density function, supporting the need of a density-based discretization algorithm.
+
 Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$ when the density-based discretization algorithm is not applied.
-In order to do this we proposed a set of density functions, each one with a different value of $b$.
+In order to do this we proposed a thin spherical shell (1km of thickness) with a set of density functions, each one with a different value of $b$.
 We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the corresponding analytical solution.

### 2017-10-26 manuscript/manuscript.tex
Changes in manuscript

@@
 \begin{equation}
-    \rho(r') = A \left( e^{-(r' - R)/b} - 1\right) + 2670 \text{kg/m}^3,
+    \rho(r') = A e^{-(r' - R)/b} + C,
 \label{eq:density-exp}
@@
 \end{equation}
+\begin{equation}
+    C =
+    2670 \text{kg/m}^3 - A,
+\end{equation}
 
-\noindent and the value of $b$ equal to the thickness of the spherical shell we use to perform the test.
+\noindent and the value of $b$ equal to the thickness of the spherical shell we will use to perform the test.
 In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
@@
 
-Firstly, lets assume a symmetric density function, i.e.
-$R_c=(R_1 + R_2)/2$ is the centre radius of the shell, and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
-The results can be seen in Figure \ref{fig:D-discont-sym}.
-Once again, we obtain good accuracy for values of $D$ equal to 1, 2 and 8 for the computation of the potential, the gradient components and the Marussi tensor components, respectively.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/discontinuous-D-sym.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c$ as the centre radius of the shell.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-discont-sym}
-\end{figure}
-
-Secondly, lets assume the same density function from equation \ref{eq:density-discontinuous}, but now with an asymmetrical distribution: $R_c = (R_1 + R_2)/3$, and perform the same test.
-We can see its results in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what
-leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.
+Lets assume that $R_c = (R_1 + R_2)/3$ and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
+Its results can be seen in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.
 
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous} and $R_c=(R_1 + R_2)/3$.
+    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous}.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.

### 2017-10-30 manuscript/manuscript.tex
Changes in manuscript

@@
 
-For this reason we developed a density-based discretization algorithm that divides the tesseroid on the radius at which the \emph{maximum variation} of the density function takes place, only if a inequation that relates this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
+For this reason we developed a density-based discretization algorithm that divides the tesseroid on the radius at which the \emph{maximum variation} of the density function takes place, only if a certain inequation that relates this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
 Therefore the number of subdivisions is indirectly controlled by this delta ratio $\delta$, and thus the accuracy of the computation.
 
-Because there is no direct relation between the values of $D$ and $\delta$ and the error of the computation, we had to empirically determine the minimum value of $D$ and the maximum value of $\delta$ that produce an acceptable accuracy for each gravity field.
+Because there is no direct relation between the values assigned to $D$ and $\delta$ and the error of the computation, we had to empirically determine the minimum value of $D$ and the maximum value of $\delta$ that produce an acceptable accuracy for each gravity field.
 
@@
 From the wide range of possible density variations, we've chosen a linear, an exponential and a discontinuous density functions.
-These comparisons have been made without applying the density-based discretization algorithm in order to isolate the other one.
-The tests over the first two functions have shown that a value of $D$ equal to 1, 2 and 8 must be used in the calculation of the gravity potential, its gradient components and the components of the Marussi tensor, respectively, in order to achieve an accuracy bellow 0.1\%.
-On the other hand we have found a limitation of this algorithm working on its own in case of highly varying exponential functions and in some cases of discontinuous density.
+These comparisons have been made without applying the density-based discretization algorithm in order to isolate the former one.
+The tests over the first two functions have shown that values of $D$ equal to 1, 2 and 8 for the calculation of the gravity potential, its gradient and the Marussi tensor components, respectively, produce an accuracy of 0.1\% or better.
+On the other hand, the algorithm on its own does not guarantee this amount of accuracy in case of highly varying exponential functions or discontinuous densities.
 
-Nevertheless, the density-based discretization algorithm have shown that a value of $\delta = 0.2$ (together with the previously obtained default values of $D$) guarantees a good accuracy on those cases on which the modified adaptive discretization algorithm does not achieve it on its own.
+We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from a low varying to a high varying one.
+Each comparison was made for different values of $\delta$, keeping the default values of $D$ previously obtained.
+This tests have shown that a value of $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while the same results can be obtained performing this tests with the discontinuous density function.
 

### 2017-10-30 manuscript/manuscript.tex
Edited speed comparison in manuscript

@@
 
-We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model with homogeneous ($\Delta t_\text{homogeneous}$) and variable ($\Delta t_\text{variable}$) density at different heights.
-Then we plotted the ratio between these computation times (see Figure \ref{fig:speed-comparison}):
+We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model ($35$km of thickness) with homogeneous, linear and exponential density (with $b=50$km) at different heights.
+Then we calculated the ratio between these computation times:
 
@@
     \text{Computation Times Ratio} =
-        \frac{\Delta t_\text{variable}}{\Delta t_\text{homogeneous}}
+        \frac{\Delta t_\text{variable}}{\Delta t_\text{homogeneous}},
     \label{eq:computation-times-ratio}
@@
 
-From these results we can conclude that the homogeneous density code is up to twice faster than the new variable density algorithm for lower heights.
-However, the computation times ratio gets lower at greater heights, due to the reduction of the divisions made by the adaptive discretization algorithm.
-In summary, the new variable density code is fast enough to perform the calculations of a real model in a few minutes.
-\todo{real, esta bien la expresion?}
+\noindent where $\Delta t_\text{homogeneous}$ and $\Delta t_\text{variable}$ are the computation times corresponding to the homogeneous model and to the variable density ones (linear and exponential).
+
+On Figures \ref{fig:speed-comparison}a and \ref{fig:speed-comparison}b we can see the results of these comparison.
+In the linear density case the density-based discretization algorithm is not applied because the inequation \ref{eq:delta-density} is never satisfied.
+For this reason, the homogeneous density model is up to 5 times faster than the variable density one in this case.
+On the other hand, the density-based discretization algorithm makes the computation slower: up to 20 times slower than the homogeneous density code.
+Nevertheless, this increase in the computation time pays off with a guarantee on its accuracy.
 
@@
 \includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
-\caption{Computation times ratio between a single tesseroid model with variable and with homogeneous density at different heights, defined in equation \ref{eq:computation-times-ratio}.}
+\caption{Computation time ratio between a single tesseroid model with variable and with homogeneous density at different heights (defined in equation \ref{eq:computation-times-ratio}) for (a) a linear and (b) an exponential density.}
 \label{fig:speed-comparison}

### 2017-10-31 manuscript/manuscript.tex
Reorganized manuscript folder. Fixed Makefile to work properly
Changes in manuscript

@@
 The distance-size ratio determinations have been made through the comparison of the numerical approximations with the analytical solutions for the a spherical shell with variable density, which had to be analytically obtained.
+These tests have been made without applying the density-based discretization algorithm in order to isolate the former one.
 From the wide range of possible density variations, we've chosen a linear, an exponential and a discontinuous density functions.
-These comparisons have been made without applying the density-based discretization algorithm in order to isolate the former one.
-The tests over the first two functions have shown that values of $D$ equal to 1, 2 and 8 for the calculation of the gravity potential, its gradient and the Marussi tensor components, respectively, produce an accuracy of 0.1\% or better.
+The comparisons over the first two functions have shown that values of $D$ equal to 1, 2 and 8 for the calculation of the gravity potential, its gradient and the Marussi tensor components, respectively, produce an accuracy of 0.1\% or better.
 On the other hand, the algorithm on its own does not guarantee this amount of accuracy in case of highly varying exponential functions or discontinuous densities.
 
-We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from a low varying to a high varying one.
+We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from low varying to high varying functions.
 Each comparison was made for different values of $\delta$, keeping the default values of $D$ previously obtained.
 This tests have shown that a value of $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while the same results can be obtained performing this tests with the discontinuous density function.
+
+Moreover, we had performed speed comparisons between tesseroids models with homogeneous and variable densities, showing that a high varying density function requires more computation time than a low varying one, as expected.
 
@@
 \bibliographystyle{gji}
-\bibliography{bibtex/references}
+\bibliography{references}
 

### 2017-11-02 manuscript/manuscript.tex
Fixes in manuscript

@@
 On the first years \citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples, while in the following decades other density functions have been proposed for different rock types \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
-Furthermore, the density variation with depth has been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
+Furthermore, varying densities have been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
 
@@
 
-The main issue that should be overcome is taking into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
+The main issue that should be overcome is to take into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
 A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radial boundaries (see Figure \ref{fig:tesseroid-uieda}).
-Calculating the gravity fields generated by an arbitrary tesseroid on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
+The calculation of the gravity fields generated by an arbitrary tesseroid on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
 The literature offers two approaches: one involves Taylor series expansion \citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
 
-In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need to derive the series expansion terms for each density function.
-On the other hand, the GLQ consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials, where the density information is gathered into these weights.
+In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need the series expansion terms for each density function.
+On the other hand, the GLQ consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials, where the density variation is summarised into the weights.
 Thus, this kind of numerical approximation allows to calculate the gravity fields for any density function without any other information but the function itself.
@@
 
-\citet{Ku1977} documented that the GLQ becomes less accurate when the computation point is closer to the tesseroid or for smaller GLQ order. 
+\citet{Ku1977} proved that the GLQ becomes less accurate when the computation point is closer to the tesseroid or for smaller GLQ order. 
 Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptive discretization of \citet{Li2011}.
 It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio is lower than a predefined value ($D$) that controls the accuracy of the computation. 
-\citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
+\citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the numerical model with the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
 
-In addition, the varying density inside an arbitrary tesseroid introduces a new kind of error into the numerical approximation.
-This encouraged us to develop a new density-based discretization algorithm that divides the tesseroid into smaller ones taking into account how the density function varies.
-The number of subdivisions and thus the accuracy of the approximation will be controlled by a delta ratio $\delta$, a predefined value by the user.
+In addition, a varying density inside an arbitrary tesseroid introduces a new kind of error into the numerical approximation.
+This encouraged us to develop a new density-based discretization algorithm that divides the tesseroid into smaller ones taking into account the variation of the density function.
+The number of subdivisions and thus the accuracy of the approximation will be controlled by a delta ratio $\delta$, predefined by the user.
 
-We have enhanced the implementation of \citet{Uieda2016} with a new forward gravity model that allows us to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
+We have enhanced the implementation of \citet{Uieda2016} with a new forward gravity model that allows to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
 It is based on the GLQ approximation, makes use of the modified distance-size discretization and the new density-based discretization algorithms that guarantee the accuracy even for high varying density functions.
@@
 
-In the same way as for an homogeneous density, we can observe from equation \ref{eq:glq-var-dens} that the GLQ approximates the gravity fields of a tesseroid as the ones generated by $N_\lambda N_\phi N_r$ point masses located at the nodes of the Legendre polynomials, where $N_i$ are the orders of the quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$.
-Nevertheless, in the variable density case, the information of the density function is summarised into the weights of the sum by taking into account only the values it assumes on those nodes.
-Although this may sound discouraging, we intent to prove otherwise and show that this method, along with well fitted discretization algorithms, approximates the gravity fields with good accuracy.\todo{no se si poner esta ultima oracion}
+From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates the gravity fields of a tesseroid as the ones generated by $N_\lambda N_\phi N_r$ point masses located at the nodes of the Legendre polynomials, where $N_i$ are the orders of the quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$, in the same way as the GLQ works in the homogeneous density case.
+Nevertheless, in this new variable density scenario, the information of the density function is summarised into the weights by taking into account only the values it assumes on those nodes.
+Although this may sound discouraging, we intent to prove otherwise and show that this method, along with well suited discretization algorithms, approximates the gravity fields with good accuracy.\todo{no se si poner esta ultima oracion}
 
@@
 The computation point $Q$ is located on the $z$ axe at a distance $r$ from the centre of the shell.
-For practical purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
+For our purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
 }
@@
 The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity potential generated by a spherical shell with a variable density in depth on any outside point located on the $z$ axe.
-Due to the rotational symmetry along any axe that passes through the centre of the shell, the integral in equation \ref{eq:shell-pot} reproduces the gravity potential field on any outside point at distance $r$ from the its centre.
+Due to the rotational symmetry along any axe that passes through the centre of the shell, the expression in equation \ref{eq:shell-pot} reproduces the gravity potential field on any outside point at distance $r$ from the its centre.
 
@@
 
-Moreover, the new forward model is based on the preexisting code developed by \citet{Uieda2016} (and included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that is able to compute the gravitational effects of tesseroids with homogeneous density.
+The new forward model is based on the preexisting code developed by \citet{Uieda2016} (and included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that is able to compute the gravitational effects of tesseroids with homogeneous density.
 On that implementation \citet{Uieda2016} have optimised the more time consuming functions through the Numba compiler.
@@
 
-It is also compatible with the preexisting open-source library Fatiando a Terra and its usage is very similar to the one for homogeneous tesseroids:
+Moreover, it is also compatible with the preexisting open-source library Fatiando a Terra and its usage is very similar to the one for homogeneous tesseroids:
 
@@
 \noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point, and $D$ is a predefined value called distance-size ratio.
-If the inequality \ref{eq:distance-size-ratio} is not held for at least one spherical coordinate, then the tesseroid must be divided by half in that direction or directions.
+If the inequality \ref{eq:distance-size-ratio} is not held for at least one spherical coordinate, then the tesseroid must be divided by half in that direction.
 A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
-On the contrary, if the inequality is satisfied, the algorithm computes the desired gravity field using the GLQ approximation.
+On the contrary, if the inequality is satisfied for every spherical coordinate, the algorithm computes the desired gravity field using the GLQ approximation.
 
@@
 
-On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy must be objectively determined beforehand.
+On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy must be empirically determined beforehand.
 In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
@@
 
-Secondly, we must define a linear function $\rho_l(r')$ that assumes the same values of the normalised density $\rho_n(r')$ on the boundary radii:
+Secondly, we must define a linear function $\rho_l(r')$ that assumes the same values as the normalised density $\rho_n(r')$ on the boundary radii:
 
@@
 
-In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
+In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
 The first one constitutes the simplest variation, the second one represents a higher varying function, while the later one tries to emulate a contact between two different media (e.g.
@@
 In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
-The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function and for an homogeneous spherical shell.
+The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function (like the one in equation \ref{eq:density-exp}) and for an homogeneous spherical shell.
 
@@
 The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
-Once again, the results for $g_{xx}$ and $g_{yy}$ are very similar to the ones obtained for $g_{zz}$, so we exclude them, although, they can be found in the repository.
+Once again, we exclude the ones corresponding to $g_{xx}$ and $g_{yy}$ due to their similarity with the ones obtained for $g_{zz}$, although they can be found in the repository.
 
@@
     Results of the distance-size ratio determination for different exponential density functions in case of a spherical shell of 1km of thickness.
-    (a) Density functions used in the tests, each one with a different value of $b$ and normalised in such way that they reaches a density of 2900kg/m$^{3}$ at the bottom of the shell.
+    (a) Density functions used in the tests, each one with a different value of $b$.
     (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
 Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
+If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy.
 }
@@
 On the other hand, the density-based discretization algorithm makes the computation slower: up to 20 times slower than the homogeneous density code.
-Nevertheless, this increase in the computation time pays off with a guarantee on its accuracy.
+Nevertheless, this increase in the computation time pays off with a guarantee\todo{guarantee is a noun?} on its accuracy.
 
@@
 The resulting gravity fields can be seen in Figures \ref{fig:neuquen-basin}c-h.
-We have excluded the other components of the Marussi tensor from the Figure, although they have also been computed and can be found in the repository.
+We have excluded the other components of the Marussi tensor from the Figure, although they can be found in the repository.
 
@@
 
-The distance-size ratio determinations have been made through the comparison of the numerical approximations with the analytical solutions for the a spherical shell with variable density, which had to be analytically obtained.
+The distance-size ratio determinations have been made through the comparison of the numerical approximations with the analytical solutions for a spherical shell with variable density, which had to be analytically obtained.
 These tests have been made without applying the density-based discretization algorithm in order to isolate the former one.
@@
 
-We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from low varying to high varying functions.
+We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from low to high varying functions.
 Each comparison was made for different values of $\delta$, keeping the default values of $D$ previously obtained.

### 2017-11-02 manuscript/manuscript.tex
Edited Summary and Keywords

@@
 \begin{summary}
-Summary of this paper 
+We present a new open-soruce software that allows to compute the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e.
+a forward gravity model in spherical coordinates for varying densities on the radial direction.
+It is written in Python and Cython language and is compatible with the classes and functions of the library for geophysical modelling Fatiando a Terra.
+It numerically approximates the gravitational fields by applying the Gauss-Legendre Quadrature along with a preexisting adaptive discretization algorithm and a new density-based discretization algorithm that control both the accuracy and the computation time of the forward model.
+The first one is controlled by a predefined distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
+While the second one, regulated by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
+We have also obtained analytical solutions for a spherical shell with variable density and compared them with the results of the numerical model in case of a linear, an exponential and a discontinuous density function.
+These comparisons allowed to obtain default values for the distance-size and the delta ratio in order to guarantee the accuracy of the forward model.
+The resulting optimal values of $D$ for the gravity potential, the components of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
+While a $\delta=0.2$ is needed even for high varying exponential densities and discontinuous functions.
+We have finally tested the new software by modelling the Neuquen sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning a linear density variation to it and computing all the gravity fields it generates.
 \end{summary}
@@
 \begin{keywords}
-Satellite gravity??, Numerical modelling, Structure of the Earth, Gravity anomalies and Earth structure, Numerical approximations and analysis, spatial analysis??
+Numerical modelling, Numerical approximations and analysis, Gravity anomalies and Earth structure, Satellite gravity
 \end{keywords}

### 2017-11-02 manuscript/manuscript.tex
Added potential and gz in delta determination

@@
 While a $\delta=0.2$ is needed even for high varying exponential densities and discontinuous functions.
-We have finally tested the new software by modelling the Neuquen sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning a linear density variation to it and computing all the gravity fields it generates.
+We have finally tested the new software by modelling the Neuqu\'en sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning a linear density variation to it and computing all the gravity fields it generates.
 \end{summary}
@@
 
-To do this we applied a similar methodology used in the determination of the distance-size ratio $D$: we proposed a set of exponential density functions (as the one in equation \ref{eq:density-exp}) with different values of $b$, and computed the difference between the gravity fields generated by the numerical model with the corresponding analytical solution, each time with a different value of $\delta$ ranging from 0.1 to 1 with a step of 0.1.
-This tests have been performed for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
+To do this we applied a similar methodology used in the determination of the distance-size ratio $D$.
+Using different values of $\delta$, ranging from 0.1 to 1 with a step of 0.1, we computed the potential, the $g_z$ component of the gradient and the $g_{zz}$ component of the Marussi tensor generated by a spherical shell with several exponential densities (like the one in equation \ref{eq:density-exp}): each one with a different value of $b$.
+We performed this tests for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
+
+Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a show the exponential density functions used on each test for a thin and thick spherical shell, respectively.
+While on Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b we can see the differences between the resulting $g_{zz}$ with the corresponding analytical solution.
+The results for the other gravity fields were excluded from the figures due to their similarities with the ones for $g_{zz}$.
 
@@
     (a) Density functions used to perform the test.
-    (b) Differences between the numerical model and the corresponding analytical solution for several values of $\delta$ and $b$.}
+    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
 \label{fig:exponential-delta-thin}
@@
     (a) Density functions used to perform the test.
-    (b) Differences between the numerical model and the corresponding analytical solution for several values of $\delta$ and $b$.}
+    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
 \label{fig:exponential-delta-thick}
 \end{figure}
-
-On Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a we can see the exponential density functions used on each test for a thin and thick spherical shell, respectively.
-While Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b show their corresponding results.
 
@@
 This tests have shown that a value of $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while the same results can be obtained performing this tests with the discontinuous density function.
+The default value $\delta$ is the same for every gravity field computation.
 

### 2017-11-02 manuscript/manuscript.tex
Added todonote in real data example

@@
 We must also define a density function for the entire model.
-\citet{Sigismondi2012} proposed a minimum and a maximum density for the Neuqu\'en basin of 275kg/m$^3$ and 412kg/m$^3$, respectively.
+\citet{Sigismondi2012} proposed a minimum and a maximum density for the Neuqu\'en basin of 275kg/m$^3$ and 412kg/m$^3$, respectively.\todo{ver que esos valores son de contraste y tienen q ir negativos.
+poner los valores absolutos.}
 So we have chosen, for this specific example, a linear density that assumes the minimum value on the surface and the maximum at 5858m, i.e. the bottom of the basin.

### 2017-11-06 code/tesseroid_relief.py
Edited manuscript/README.md

@@ -1,5 +1,13 @@
 # -*- coding: utf-8 -*-
 """
+Written by Leonardo Uieda (2016)
+https://github.com/pinga-lab/paper-moho-inversion-tesseroids
+
+Available under a BSD 3-clause license.
+You can freely use and modify the code, without warranty, so long as you
+provide attribution to the authors. See LICENSE.md for the full license text.
+
+
 Implements the proposed Moho inversion method.
 
 Defines the classes:

### 2017-11-06 manuscript/manuscript.tex

--- 2017-11-02 manuscript/manuscript.tex
@@
 \begin{summary}
-We present a new open-soruce software that allows to compute the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e.
+We present a new open-soruce software that computes the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e.
 a forward gravity model in spherical coordinates for varying densities on the radial direction.
@@
 It numerically approximates the gravitational fields by applying the Gauss-Legendre Quadrature along with a preexisting adaptive discretization algorithm and a new density-based discretization algorithm that control both the accuracy and the computation time of the forward model.
-The first one is controlled by a predefined distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
-While the second one, regulated by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
+The first algorithm is regulated by a predefined scalar called distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
+While the second one, controlled by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
 We have also obtained analytical solutions for a spherical shell with variable density and compared them with the results of the numerical model in case of a linear, an exponential and a discontinuous density function.
@@
 The resulting optimal values of $D$ for the gravity potential, the components of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
-While a $\delta=0.2$ is needed even for high varying exponential densities and discontinuous functions.
-We have finally tested the new software by modelling the Neuqu\'en sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning a linear density variation to it and computing all the gravity fields it generates.
+While a $\delta=0.2$ is needed in the computation of every gravity field, guaranteeing the accuracy even for high varying exponential densities and discontinuous functions.
+We have finally tested the new software by modelling the Neuqu\'en sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning an exponential density variation to it and computing all the gravity fields it generates.
 \end{summary}
@@
 \includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
-\caption{(a) Topography of Neuqu\'en Basin (in km),
+\caption{
+         Forward model example with real data: Nequ\'en sedimentary modelled through tesseroids with exponential density.
+         (a) Topography of Neuqu\'en Basin (in km),
          (b) Thickness of the sedimentary basin (in meters),
@@
 We must also define a density function for the entire model.
-\citet{Sigismondi2012} proposed a minimum and a maximum density for the Neuqu\'en basin of 275kg/m$^3$ and 412kg/m$^3$, respectively.
-So we have chosen, for this specific example, a linear density that assumes the minimum value on the surface and the maximum at 5858m, i.e. the bottom of the basin.
+\citet{Sigismondi2012} proposed a minimum and maximum density contrast for the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
+\todo{ver que esos valores son de contraste y tienen q ir negativos.
+poner los valores absolutos.}
+So we have chosen, for this specific example, an exponential density (like the one in equation \ref{eq:density-exp}) that assumes the minimum value on the surface and the maximum at 5858m, i.e.
+the bottom of the basin, with a value of $b$ equal to 3km.
 
@@
 
-Finally, we have shown a real data example: we have computed the gravity fields generated by the Neuqu\'en sedimentary basin through a tesseroid model with linear increasing density with depth on a regular grid at 50km height.
+Finally, we have shown a real data example: we have computed the gravity fields generated by the Neuqu\'en sedimentary basin through a tesseroid model with exponential density with depth on a regular grid at 50km height.
 

### 2017-11-06 jovenes-investigadores/jovenes-investigadores.tex
Fixed typo
Edited jovenes-investigadores.tex for new figures names

@@
 En este trabajo hemos sido capaces de implementar un modelo directo que reproduce los campos gravitatorios generados por Tesseroides con densidad variable en profundidad, para lo cual hemos utilizado como aproximación la Cuadratura de Gauss-Legendre y un algoritmo de discretización adaptativa con el objetivo de garantizar la precisión de la misma.
-El código fue escrito en Python y Cython y será liberado como Software Libre.
+El código fue escrito en Python y Cython y será liberado bajo licencias de Software Libre.
 
@@
     \centering
-    \includegraphics[height=30em]{../manuscript/figures/Dlinear-thin-differences.pdf}
+    \includegraphics[height=30em]{../manuscript/figures/linear-D-thin.pdf}
     \caption{}
@@
     \centering
-    \includegraphics[height=30em]{../manuscript/figures/Dexp-shifted-thin-differences.pdf}
+    \includegraphics[height=30em]{../manuscript/figures/exponential-D-thin.pdf}
     \caption{}
@@
 \bibliographystyle{apalike-es}
-\bibliography{../manuscript/bibtex/references}
+\bibliography{../manuscript/references}
 

### 2018-02-22 manuscript/manuscript.tex

@@
 If we set the GLQ orders to fixed values, it's easy to conclude that the nearer the computation point is to the tesseroid, the accuracy of the approximation is reduced and the point masses effect becomes more noticeable than the tesseroid we want to approximate.
-A more complete analysis of this conclusion can be found in \citet{Ku1977, Li2011, Uieda2016}.
+A complete analysis of this conclusion can be found in \citet{Ku1977, Li2011, Uieda2016}.
 
@@
 If the inequality \ref{eq:distance-size-ratio} is not held for at least one spherical coordinate, then the tesseroid must be divided by half in that direction.
-A more complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
+A complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
 On the contrary, if the inequality is satisfied for every spherical coordinate, the algorithm computes the desired gravity field using the GLQ approximation.

### 2018-02-22 manuscript/manuscript.tex
More theoretical emphasis on the title
This is more of a theoretical paper with a bit of software than the
other way around.

Update my affiliation.

@@
 \title[Variable Density Tesseroids]{
-    Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth
+    Gravitational field calculation in spherical coordinates using variable densities in depth
 }
@@
     $^2$Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan, Argentina\\
-    $^3$Universidade do Estado do Rio de Janeiro, Rio de Janeiro, Brazil
+    $^3$Department of Geology and Geophysics, SOEST, University of Hawai‘i at Manoa, Honolulu, Hawaii, USA
     }
@@
 \begin{summary}
-We present a new open-soruce software that computes the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e.
-a forward gravity model in spherical coordinates for varying densities on the radial direction.
-It is written in Python and Cython language and is compatible with the classes and functions of the library for geophysical modelling Fatiando a Terra.
-It numerically approximates the gravitational fields by applying the Gauss-Legendre Quadrature along with a preexisting adaptive discretization algorithm and a new density-based discretization algorithm that control both the accuracy and the computation time of the forward model.
-The first algorithm is regulated by a predefined scalar called distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
-While the second one, controlled by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
-We have also obtained analytical solutions for a spherical shell with variable density and compared them with the results of the numerical model in case of a linear, an exponential and a discontinuous density function.
-These comparisons allowed to obtain default values for the distance-size and the delta ratio in order to guarantee the accuracy of the forward model.
-The resulting optimal values of $D$ for the gravity potential, the components of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
-While a $\delta=0.2$ is needed in the computation of every gravity field, guaranteeing the accuracy even for high varying exponential densities and discontinuous functions.
-We have finally tested the new software by modelling the Neuqu\'en sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning an exponential density variation to it and computing all the gravity fields it generates.
+We present a new open-soruce software that computes the gravitational fields
+generated by any tesseroid whose density varies arbitrarily in depth, i.e. a
+forward gravity model in spherical coordinates for varying densities on the
+radial direction.
+It is written in Python and Cython language and is compatible with the classes
+and functions of the library for geophysical modelling Fatiando a Terra.
+It numerically approximates the gravitational fields by applying the
+Gauss-Legendre Quadrature along with a preexisting adaptive discretization
+algorithm and a new density-based discretization algorithm that control both
+the accuracy and the computation time of the forward model.
+The first algorithm is regulated by a predefined scalar called distance-size
+ratio ($D$) and reduces the errors due to the distance between the tesseroid
+and the computation point.
+While the second one, controlled by a delta ratio ($\delta$), decreases the
+errors introduced by the variation of the density function.
+We have also obtained analytical solutions for a spherical shell with variable
+density and compared them with the results of the numerical model in case of a
+linear, an exponential and a discontinuous density function.
+These comparisons allowed to obtain default values for the distance-size and
+the delta ratio in order to guarantee the accuracy of the forward model.
+The resulting optimal values of $D$ for the gravity potential, the components
+of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
+While a $\delta=0.2$ is needed in the computation of every gravity field,
+guaranteeing the accuracy even for high varying exponential densities and
+discontinuous functions.
+We have finally tested the new software by modelling the Neuqu\'en sedimentary
+basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S
+latitude, assigning an exponential density variation to it and computing all
+the gravity fields it generates.
 \end{summary}

### 2018-04-04 code/tesseroid_density/tesseroid.py
Defined specific delta values for potential, grandient and tensor componentes

@@ -105,7 +105,9 @@ RATIO_V = 1
 RATIO_G = 2
 RATIO_GG = 8
 STACK_SIZE = 100
-DELTA = 0.2
+DELTA_V = 0.2
+DELTA_G = 0.2
+DELTA_GG = 0.02
 
 
 def _check_input(lon, lat, height, model, ratio, njobs, pool):
@@ -346,7 +348,7 @@ def _split_arrays(arrays, extra_args, nparts):
 
 
 def potential(lon, lat, height, model, dens=None, ratio=RATIO_V,
-              njobs=1, pool=None, delta=DELTA):
+              njobs=1, pool=None, delta=DELTA_V):
     """
     Calculate the gravitational potential due to a tesseroid model.
 
@@ -404,7 +406,7 @@ def potential(lon, lat, height, model, dens=None, ratio=RATIO_V,
 
 
 def gx(lon, lat, height, model, dens=None, ratio=RATIO_G,
-       njobs=1, pool=None, delta=DELTA):
+       njobs=1, pool=None, delta=DELTA_G):
     """
     Calculate the North component of the gravitational attraction.
 
@@ -462,7 +464,7 @@ def gx(lon, lat, height, model, dens=None, ratio=RATIO_G,
 
 
 def gy(lon, lat, height, model, dens=None, ratio=RATIO_G,
-       njobs=1, pool=None, delta=DELTA):
+       njobs=1, pool=None, delta=DELTA_G):
     """
     Calculate the East component of the gravitational attraction.
 
@@ -520,7 +522,7 @@ def gy(lon, lat, height, model, dens=None, ratio=RATIO_G,
 
 
 def gz(lon, lat, height, model, dens=None, ratio=RATIO_G,
-       njobs=1, pool=None, delta=DELTA):
+       njobs=1, pool=None, delta=DELTA_G):
     """
     Calculate the radial component of the gravitational attraction.
 
@@ -583,7 +585,7 @@ def gz(lon, lat, height, model, dens=None, ratio=RATIO_G,
 
 
 def gxx(lon, lat, height, model, dens=None, ratio=RATIO_GG,
-        njobs=1, pool=None, delta=DELTA):
+        njobs=1, pool=None, delta=DELTA_GG):
     """
     Calculate the xx component of the gravity gradient tensor.
 
@@ -641,7 +643,7 @@ def gxx(lon, lat, height, model, dens=None, ratio=RATIO_GG,
 
 
[…]

### 2018-04-04 manuscript/manuscript.tex
Modified manuscript: changes in title and abstract

@@
 \title[Variable Density Tesseroids]{
-    Gravitational field calculation in spherical coordinates using variable densities in depth
+    Gravitational field calculation in spherical coordinates using variable 
+    densities in depth
 }
 \author[S.R. Soler, A. Pesce, L. Uieda and M.E. Gimenez]{
-    Santiado R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and Mario E. Gimenez$^{1,2}$ \\
+    Santiado R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and
+    Mario E. Gimenez$^{1,2}$ \\
     $^1$CONICET, Argentina. e-mail: santiago.r.soler@gmail.com\\
-    $^2$Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan, Argentina\\
-    $^3$Department of Geology and Geophysics, SOEST, University of Hawai‘i at Manoa, Honolulu, Hawaii, USA
+    $^2$Instituto Geofísico Sismológico Volponi, Universidad Nacional de
+    San Juan, Argentina\\
+    $^3$Department of Geology and Geophysics, SOEST, University of Hawai‘i at
+    Manoa, Honolulu, Hawaii, USA
     }
@@
 \begin{summary}
-We present a new open-soruce software that computes the gravitational fields
-generated by any tesseroid whose density varies arbitrarily in depth, i.e. a
-forward gravity model in spherical coordinates for varying densities on the
-radial direction.
-It is written in Python and Cython language and is compatible with the classes
-and functions of the library for geophysical modelling Fatiando a Terra.
-It numerically approximates the gravitational fields by applying the
-Gauss-Legendre Quadrature along with a preexisting adaptive discretization
-algorithm and a new density-based discretization algorithm that control both
-the accuracy and the computation time of the forward model.
+We present a new methodology to compute the gravitational fields generated by 
+any tesseroid whose density varies continuously in depth, i.e. a forward 
+gravity model in spherical coordinates for varying densities on the radial 
+direction.
+It's based on the numerical approximation of the gravitational fields by 
+applying the Gauss-Legendre Quadrature along with a preexisting adaptive 
+discretization algorithm and a new density-based discretization algorithm that 
+control both the accuracy and the computation time of the forward model.
 The first algorithm is regulated by a predefined scalar called distance-size 
@@
 density and compared them with the results of the numerical model in case of a 
-linear, an exponential and a discontinuous density function.
+linear and an exponential density function.
 These comparisons allowed to obtain default values for the distance-size and 
@@
 The resulting optimal values of $D$ for the gravity potential, the components 
-of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
-While a $\delta=0.2$ is needed in the computation of every gravity field,
-guaranteeing the accuracy even for high varying exponential densities and
-discontinuous functions.
+of its gradient and the Marussi tensor components are 1, 2 and 8, 
+respectively.
+While a $\delta=0.2$ is needed in the computation of every gravitational 
+field, guaranteeing the accuracy even for high varying exponential densities.
 We have finally tested the new software by modelling the Neuqu\'en sedimentary 

### 2018-04-04 manuscript/manuscript.tex
Restructured introduction in manuscript

@@
 \begin{keywords}
-Numerical modelling, Numerical approximations and analysis, Gravity anomalies and Earth structure, Satellite gravity
+Numerical modelling, Numerical approximations and analysis, Gravity anomalies 
+and Earth structure, Satellite gravity
 \end{keywords}
@@
 
-Lithosphere's density variation with depth has been studied for almost a century. 
-On the first years \citet{Athy1930} obtained a decreasing exponential relation between both quantities by studying shale samples, while in the following decades other density functions have been proposed for different rock types \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
-Furthermore, varying densities have been taken into account in forward and inversion gravity models, mostly applied to grabens and basins \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, Welford2010}.
+Lithosphere's density variation with depth has been studied for almost a 
+century.
+On the first years \citet{Athy1930} obtained a decreasing exponential relation 
+between both quantities by studying shale samples, while in the following 
+decades other density functions have been proposed for different rock types 
+\citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}. 
+Furthermore, varying densities have been taken into account in forward and 
+inversion gravity models, mostly applied to grabens and basins 
+\citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001, 
+Welford2010}.
 
-These forward gravity models have been developed for two or three dimensional bodies in cartesian coordinates that properly fit local scales applications.
-Nevertheless, the latest satellite missions have provided us gravity measurements with global coverage, which allow geologists and geophysics to perform modelling and interpretation on regional scales.
-This raises the importance of building forward models that reproduce the gravity anomalies for such scales.
+These forward gravity models have been developed for two or three dimensional 
+bodies in cartesian coordinates that properly fit local scales applications. 
+Nevertheless, the latest satellite missions have provided us gravity 
+measurements with global coverage, which allow geologists and geophysics to 
+perform modelling and interpretation on regional scales. This raises the 
+importance of building forward models that reproduce the gravity anomalies for 
+such scales.
 
-The main issue that should be overcome is to take into account the curvature of the Earth, thus the forward model must be defined in spherical coordinates.
-A common way to achieve this is to discretize the Earth in spherical prisms known as tesseroids, which are defined by pairs of latitude, longitude and radial boundaries (see Figure \ref{fig:tesseroid-uieda}).
-The calculation of the gravity fields generated by an arbitrary tesseroid on any external point involves the resolution of volume integrals that are generally approximated by numerical computations.
-The literature offers two approaches: one involves Taylor series expansion \citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
+The main issue that should be overcome is to take into account the curvature 
+of the Earth, thus the forward model must be defined in spherical coordinates. 
+A common way to achieve this is to discretize the Earth in spherical prisms 
+known as tesseroids, which are defined by pairs of latitude, longitude and 
+radial boundaries (see Figure \ref{fig:tesseroid-uieda}).
+The calculation of the gravity fields generated by an arbitrary tesseroid on 
+any external point involves the resolution of volume integrals that are 
+generally approximated by numerical computations.
+The literature offers two approaches: one involves Taylor series expansion 
+\citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre 
+Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
 
-In order to develop a forward model that computes the gravity field generated by any tesseroid with arbitrary variable density, the Taylor series expansion is not well suited, because it would need the series expansion terms for each density function.
-On the other hand, the GLQ consists in approximating the integral by a weighted sum of the effect of point masses located at scaled nodes of the Legendre polynomials, where the density variation is summarised into the weights.
-Thus, this kind of numerical approximation allows to calculate the gravity fields for any density function without any other information but the function itself.
+In order to develop a forward model that computes the gravity field generated 
+by any tesseroid with arbitrary variable density, the Taylor series expansion 
+is not well suited, because it would need the series expansion terms for each 
+density function.
+On the other hand, the GLQ consists in approximating the integral by a 
+weighted sum of the effect of point masses located at scaled nodes of the 
+Legendre polynomials, where the density variation is summarised into the 
+weights.
+Thus, this kind of numerical approximation allows to calculate the gravity 
+fields for any density function without any other information but the function 
+itself.
 
@@
 \caption{
-A spherical prism, also known as  Tesseroid, in a geocentric spherical coordinate system, with a computation point $P$ and its local north oriented Cartesian coordinate system.
-Picture made by \citet{Uieda2015}.
+    A spherical prism, also known as  Tesseroid, in a geocentric spherical 
+    coordinate system, with a computation point $P$ and its local north 
+    oriented Cartesian coordinate system. Picture made by \citet{Uieda2015}.
 }
@@
 
-\citet{Uieda2016} developed a forward gravity model based on tesseroids with homogeneous densities using the GLQ approximation method.
-It has been originally implemented through command line programs written in C programming language.
-However, it has been ultimately ported to Python and included in the open-source library Fatiando a Terra \citep{Uieda2013} for geophysical modelling and inversion.
+\citet{Uieda2016} developed a forward gravity model based on tesseroids with 
+homogeneous densities using the GLQ approximation method.
+It has been originally implemented through command line programs written in C 
+programming language.
+However, it has been ultimately ported to Python and included in the 
+open-source library Fatiando a Terra \citep{Uieda2013} for geophysical 
+modelling and inversion.
+\todo{hace falta aclarar sobre el software de leo?}
 
-\citet{Ku1977} proved that the GLQ becomes less accurate when the computation point is closer to the tesseroid or for smaller GLQ order. 
-Although the model developed by \citet{Uieda2016} uses a second order GLQ, it ensures the accuracy implementing a modified version of the adaptive discretization of \citet{Li2011}.
-It consist in splitting the tesseroid into smaller ones when a certain distance-size ratio is lower than a predefined value ($D$) that controls the accuracy of the computation. 
-\citet{Uieda2016} have also objectively obtained standard values of $D$ for gravity potential, gradients and tensor components comparing the numerical model with the fields generated by a spherical shell, which constitutes a special case of the volume integrals that has analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}\todo{citar mas??}.
+\citet{Ku1977} proved that the GLQ becomes less accurate when the computation 
+point is closer to the tesseroid or for smaller GLQ order.
+Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
+ensures the accuracy implementing a modified version of the adaptive 
+discretization of \citet{Li2011}.
+It consist in splitting the tesseroid into smaller ones when a certain 
+distance-size ratio is lower than a predefined value ($D$) that controls the 
+accuracy of the computation.
+\citet{Uieda2016} have also objectively obtained standard values of $D$ for 
+gravity potential, gradients and tensor components comparing the numerical 
+model with the fields generated by a spherical shell, which constitutes a 
+special case of the volume integrals that has analytical solution 
+\citep{LaFehr1991, Mikuska2006, Grombein2013}
+\todo{citar mas??}.
 
-In addition, a varying density inside an arbitrary tesseroid introduces a new kind of error into the numerical approximation.
-This encouraged us to develop a new density-based discretization algorithm that divides the tesseroid into smaller ones taking into account the variation of the density function.
-The number of subdivisions and thus the accuracy of the approximation will be controlled by a delta ratio $\delta$, predefined by the user.
+In addition, a varying density inside an arbitrary tesseroid introduces a new 
+kind of error into the numerical approximation. This encouraged us to develop 
+a new density-based discretization algorithm that divides the tesseroid into 
+smaller ones taking into account the variation of the density function.
+The number of subdivisions and thus the accuracy of the approximation will be 
+controlled by a delta ratio $\delta$, predefined by the user.
 
-We have enhanced the implementation of \citet{Uieda2016} with a new forward gravity model that allows to compute the gravity fields generated by any tesseroid with an arbitrary density function on any external point.
-It is based on the GLQ approximation, makes use of the modified distance-size discretization and the new density-based discretization algorithms that guarantee the accuracy even for high varying density functions.
-It is written in Python and Cython languages, constituting an easy to use library that runs precompiled code for the more time consuming functions.
-It also makes use of some previously existing classes and functions from Fatiando a Terra, what saved us time and allowed us to focus on the new enhancements.
-Furthermore, we have the intention to include it into a future release of the Fatiando a Terra library in order to make its distribution and maintenance easier.
+We have enhanced the implementation of \citet{Uieda2016} with a new forward 
+gravity model that allows to compute the gravity fields generated by any 
+tesseroid with an arbitrary density function on any external point.
+It is based on the GLQ approximation, makes use of the modified distance-size 
+discretization and the new density-based discretization algorithms that 
+guarantee the accuracy even for high varying density functions.
+It is written in Python and Cython languages, constituting an easy to use 
+library that runs precompiled code for the more time consuming functions.
+It also makes use of some previously existing classes and functions from 
+Fatiando a Terra, what saved us time and allowed us to focus on the new 
+enhancements.
+Furthermore, we have the intention to include it into a future release of the 
+Fatiando a Terra library in order to make its distribution and maintenance 
+easier.
 
-In order to ensure the accuracy of the numerical approximation, we have conducted similar tests as the one performed by \citet{Uieda2016} by comparing the results of the numerical model with the analytical solutions for a spherical shell with linear, exponential and discontinuous density.
-Furthermore we had to derive the analytical expressions for these density functions.
+In order to ensure the accuracy of the numerical approximation, we have 
+conducted similar tests as the one performed by \citet{Uieda2016} by comparing 
+the results of the numerical model with the analytical solutions for a 
+spherical shell with linear, exponential and discontinuous density.
+Furthermore we had to derive the analytical expressions for these density 
+functions.
 
-In the following sections we will describe how the new algorithm works, its theoretical background and obtain default distance-size ratio and delta ratio values needed to get an acceptable accuracy in the numerical approximation.
+In the following sections we will describe how the new algorithm works, its 
+theoretical background and obtain default distance-size ratio and delta ratio 
+values needed to get an acceptable accuracy in the numerical approximation.
 

### 2018-04-04 manuscript/manuscript.tex
Modifications on introduction in manuscript

@@
 \usepackage{graphicx}
-%~ \usepackage{todonotes} % to make annotations on margins
-\usepackage[disable]{todonotes} % to disable annotations on margins
+\usepackage{todonotes} % to make annotations on margins
+%~ \usepackage[disable]{todonotes} % to disable annotations on margins
 
@@
 
-In order to develop a forward model that computes the gravity field generated 
-by any tesseroid with arbitrary variable density, the Taylor series expansion 
-is not well suited, because it would need the series expansion terms for each 
-density function.
+In order to develop a forward model that computes the gravitational 
+field generated by any tesseroid with arbitrary variable density, the 
+Taylor series expansion is not well suited, because it would need the 
+series expansion terms for each density function.
 On the other hand, the GLQ consists in approximating the integral by a 
@@
 weights.
-Thus, this kind of numerical approximation allows to calculate the gravity 
-fields for any density function without any other information but the function 
-itself.
+Thus, this kind of numerical approximation allows to calculate the 
+gravitational fields for any density function without any other 
+information but the function itself.
 
@@
 
-\citet{Uieda2016} developed a forward gravity model based on tesseroids with 
-homogeneous densities using the GLQ approximation method.
-It has been originally implemented through command line programs written in C 
-programming language.
-However, it has been ultimately ported to Python and included in the 
-open-source library Fatiando a Terra \citep{Uieda2013} for geophysical 
-modelling and inversion.
-\todo{hace falta aclarar sobre el software de leo?}
-
-\citet{Ku1977} proved that the GLQ becomes less accurate when the computation 
-point is closer to the tesseroid or for smaller GLQ order.
+\citet{Uieda2016} developed a forward gravitational model based on 
+tesseroids with homogeneous densities using the GLQ approximation 
+method.
+The main issue of the GLQ is that the approximation becomes less 
+accurate when the computation point is closer to the tesseroid of for 
+smaler GLQ order \citep{Ku1977}.
 Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
@@
 accuracy of the computation.
-\citet{Uieda2016} have also objectively obtained standard values of $D$ for 
-gravity potential, gradients and tensor components comparing the numerical 
-model with the fields generated by a spherical shell, which constitutes a 
-special case of the volume integrals that has analytical solution 
-\citep{LaFehr1991, Mikuska2006, Grombein2013}
+\citet{Uieda2016} have also objectively obtained standard values of $D$ 
+for gravitational potential, gradients and tensor components comparing 
+the numerical model with the fields generated by a spherical shell, 
+which constitutes a special case of the volume integrals that has 
+analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013} 
 \todo{citar mas??}.
 
-In addition, a varying density inside an arbitrary tesseroid introduces a new 
-kind of error into the numerical approximation. This encouraged us to develop 
-a new density-based discretization algorithm that divides the tesseroid into 
-smaller ones taking into account the variation of the density function.
-The number of subdivisions and thus the accuracy of the approximation will be 
-controlled by a delta ratio $\delta$, predefined by the user.
+On the other hand, a varying density inside an arbitrary tesseroid 
+introduces a new kind of error into the numerical approximation.
+This encouraged us to develop a new density-based discretization 
+algorithm that divides the tesseroid into smaller ones taking into 
+account the variation of the density function.
+The number of subdivisions and thus the accuracy of the approximation 
+will be controlled by predefined delta ratio $\delta$.
 
-We have enhanced the implementation of \citet{Uieda2016} with a new forward 
-gravity model that allows to compute the gravity fields generated by any 
-tesseroid with an arbitrary density function on any external point.
-It is based on the GLQ approximation, makes use of the modified distance-size 
-discretization and the new density-based discretization algorithms that 
-guarantee the accuracy even for high varying density functions.
-It is written in Python and Cython languages, constituting an easy to use 
-library that runs precompiled code for the more time consuming functions.
-It also makes use of some previously existing classes and functions from 
-Fatiando a Terra, what saved us time and allowed us to focus on the new 
-enhancements.
-Furthermore, we have the intention to include it into a future release of the 
-Fatiando a Terra library in order to make its distribution and maintenance 
-easier.
+In summary, we present a new forward gravitational model that allows to 
+compute the gravitational fields generated by any tesseroid with an 
+arbitrary continuous density function on any external point.
+It is based on the GLQ approximation whose accuracy is controlled by 
+the modified distance-size discretization \citep{Uieda2016} 
+and the new density-based discretization algorithms.
 
-In order to ensure the accuracy of the numerical approximation, we have 
-conducted similar tests as the one performed by \citet{Uieda2016} by comparing 
-the results of the numerical model with the analytical solutions for a 
-spherical shell with linear, exponential and discontinuous density.
-Furthermore we had to derive the analytical expressions for these density 
-functions.
+In order to ensure the accuracy of the numerical approximation we have 
+determined predefined values for the $D$ and $\delta$ parameters by 
+comparing the numerical approximation with the analytical solutions for 
+a spherical shell with linear and exponential density, which had to be 
+derived.
+
+Furthermore, we developed a Python library that implements this forward 
+model.
+I makes use of some previously existing classes and functions 
+from the Python library Fatiando a Terra for geophysical modelling, 
+what saved us time and allowed us to focus on the new model.
+
+Furthermore, we have the intention to include it into a future release 
+of the Fatiando a Terra library in order to make its distribution and 
+maintenance easier.
+\todo{include this?}
 

### 2018-04-04 manuscript/manuscript.tex
Restructured Theory section in manuscript

@@
 
-The spherical prisms known as tesseroids are mass elements defined in a geocentric spherical coordinate system bounded by a pair of parallels, a pair of  meridians, and two concentric spherical surfaces.
-We define an external computation point $P(r, \phi, \lambda)$ where the gravity fields generated by the tesseroid are going to be calculated with respect to the local north oriented Cartesian coordinate system located at $P$.
-In the special case of homogeneous density, the gravity potential, its gradient and the Marussi tensor components can be calculated through the integrals obtained by \citet{Grombein2013} \citep[for same notation as the one we will use, see][]{Uieda2016}.
-
-In our case, we will assume that the tesseroid has an arbitrary variable density in depth, i.e. only depends on the radial spherical coordinate.
+The spherical prisms known as tesseroids are mass elements defined in a 
+geocentric spherical coordinate system bounded by a pair of parallels, 
+a pair of  meridians, and two concentric spherical surfaces.
+We define an external computation point $P(r, \phi, \lambda)$ where the 
+gravity fields generated by the tesseroid are going to be calculated 
+with respect to the local north oriented Cartesian coordinate system 
+located at $P$.
+In the special case of homogeneous density, the gravity potential, its 
+gradient and the Marussi tensor components can be calculated through 
+the integrals obtained by \citet{Grombein2013} \citep[for same notation 
+as the one we will use, see][]{Uieda2016}.
+
+In our case, we will assume that the tesseroid has an arbitrary 
+variable density in depth, i.e. only depends on the radial spherical 
+coordinate.
 Thus, the integrals for the gravity fields are slightly modified:
@@
 
-\noindent $\alpha, \beta \in \{x, y, z\}$,
-$\rho(r')$ is the density function that depends on the radial coordinate,
-$\delta_{\alpha\beta}$ is Kronecker's delta,
-$G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant and
+\noindent $\alpha, \beta \in \{x, y, z\}$, $\rho(r')$ is the density 
+function that depends on the radial coordinate, $\delta_{\alpha\beta}$ 
+is Kronecker's delta, $G = 6.674\times10^{-11}\, 
+\text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant and
 
@@
 
-According to \citet[p.~390]{Hildebrand1987}, we can approximate any integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e.
-by a weighted sum of the integration kernel evaluated on the roots of the $N$th Legendre polynomial:
+According to \citet[p.~390]{Hildebrand1987}, we can approximate any 
+integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e. by a 
+weighted sum of the integration kernel evaluated on the roots of the 
+$N$th Legendre polynomial:
 
@@
 
-\noindent where $x_i$ are the roots of the $N$th Legendre polynomial $P_N(x)$ and $W_i$ are the weights of each term:
+\noindent where $x_i$ are the roots of the $N$th Legendre polynomial 
+$P_N(x)$ and $W_i$ are the weights of each term:
 
@@
 
-In case of an integration defined on a arbitrary interval, e.g. $[a,b]$, we can scale the nodes and perform a similar approximation:
+In case of an integration defined on a arbitrary interval, e.g. 
+$[a,b]$, we can scale the nodes and perform a similar approximation:
 
@@
 
-\noindent where $x_i^s$ is the scaled $x_i$ root into the $[a,b]$ interval, also called nodes of the quadrature:
+\noindent where $x_i^s$ is the scaled $x_i$ root into the $[a,b]$ 
+interval, also called nodes of the quadrature:
 
@@
 
-\noindent For example, if we use a second order GLQ, the roots of the $P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = \pm 0.577350269$ and $W_i = 1$.
-
-Our intention is to use GLQ to approximate the volume integrals from equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor}. 
-When tesseroids have variable densities, we can write the integral kernels as the product of a certain function $g$ and the density function:
+\noindent For example, if we use a second order GLQ, the roots of the 
+$P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = 
+\pm 0.577350269$ and $W_i = 1$.
+
+Our intention is to use GLQ to approximate the volume integrals from 
+equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor}.
+When tesseroids have variable densities, we can write the integral 
+kernels as the product of a certain function $g$ and the density 
+function:
 
@@
 
-\noindent and then apply the quadrature to every integral corresponding to each spherical coordinate, obtaining a similar expression to the one shown by \citet{Asgharzadeh2007} and \citet{Uieda2016}:
+\noindent and then apply the quadrature to every integral corresponding 
+to each spherical coordinate, obtaining a similar expression to the one 
+shown by \citet{Asgharzadeh2007} and \citet{Uieda2016}:
 
@@
 
-From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates the gravity fields of a tesseroid as the ones generated by $N_\lambda N_\phi N_r$ point masses located at the nodes of the Legendre polynomials, where $N_i$ are the orders of the quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$, in the same way as the GLQ works in the homogeneous density case.
-Nevertheless, in this new variable density scenario, the information of the density function is summarised into the weights by taking into account only the values it assumes on those nodes.
-Although this may sound discouraging, we intent to prove otherwise and show that this method, along with well suited discretization algorithms, approximates the gravity fields with good accuracy.\todo{no se si poner esta ultima oracion}
+From equation \ref{eq:glq-var-dens} we can observe that the GLQ 
+approximates the gravity fields of a tesseroid as the ones generated by 
+$N_\lambda N_\phi N_r$ point masses located at the nodes of the 
+Legendre polynomials, where $N_i$ are the orders of the quadrature for 
+each integration, with $i \in \{ \lambda, \phi, r \}$, in the same way 
+as the GLQ works in the homogeneous density case.
+Nevertheless, in this new variable density scenario, the information of 
+the density function is summarised into the weights by taking into 
+account only the values it assumes on those nodes.
+Although this may sound discouraging, we intent to prove otherwise and 
+show that this method, along with well suited discretization 
+algorithms, approximates the gravity fields with good accuracy.\todo{no 
+se si poner esta ultima oracion}
 

### 2018-04-05 manuscript/manuscript.tex
New section in manuscript: Methodology

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Methodology}
+
+From equation \ref{eq:glq-var-dens} we can conclude that the error of 
+the numerical approximation is controlled by the order of the 
+quadrature, i.e. the number of point masses that are going to be used 
+to compute the gravitational field.
+This conclusion could lead to a naive methodology: find the order of 
+the quadrature that produces gravitational fields with an acceptable 
+accuracy.
+Although this simple proposal has a few problems: (a) it won't be 
+efficient and (b) it couldn't be applied to the general case (any
+tesseroid with arbitrary density function on any computational point) 
+without modifying the quadrature order taking into account 
+characteristics of the tesseroid and the computation point (tesseroid 
+size, distance to the computation point, variation of the density 
+function, etc).
+
+That's why we will keep the GLQ order fixed and make use of 
+discretization algorithms.
+Instead of uniformly increase the number of point masses that 
+approximates a certain tesseroid, the discretization algorithms are 
+intended to efficiently subdivide the tesseroid into smaller ones and 
+apply the GLQ to each one of them.
+These algorithms generally involve a discretization inequation that 
+controls if the tesseroid should be subdivided or not and a certain 
+predefined parameter that works as a threshold for the discretization 
+inequation.
+
+On the variable density tesseroid problem we found two main error 
+sources in the GLQ that should be addressed: (a) the relation between 
+the tesseroid size and its distance to the computation point, and (b) 
+the variation of the density function.
+
+The former one is also present in the homogeneous density tesseroid 
+models, such as the one developed by \citet{Uieda2016}, and could be 
+solved by applying a modified version of the adaptive discretization 
+algorithm proposed by \citet{Li2011}.
+Although for the latter we need to introduce a new density-based 
+discretization algorithm that subdivides the tesseroid taking into 
+account the variation of its density function.
+
+\subsection{Modified Adaptive Discretization Algorithm}
+
+From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates 
+the volume integrals as the gravitational effect of point masses 
+located at the nodes of the quadrature. If we set the GLQ orders to 
+fixed values, it's easy to conclude that the nearer the computation 
+point is to the tesseroid, the accuracy of the approximation is reduced 
+and the point masses effect becomes more noticeable than the tesseroid 
+we want to approximate. A complete analysis of this conclusion can be 
+found in \citet{Ku1977, Li2011, Uieda2016}.
+
+Studying the homogeneous density tesseroid problem, \citet{Ku1977} 
+noticed that keeping the GLQ orders fixed, the approximation 
+becomes less accurate when the computation point is closer to the 
+tesseroid, making the point masses effect more noticeable than the 
+tesseroid we want to approximate.
+A complete analysis of this conclusion can be found in \citet{Ku1977, 
+Li2011, Uieda2016}.
+
+Instead of increasing the GLQ orders to improve the accuracy, what 
+leads to a more time consuming computation, \citet{Li2011} proposed an 
+adaptive discretization algorithm which essentially consists in 
+dividing the tesseroid based on a ratio between the distance to the 
+computation point and its dimensions.
+\citet{Uieda2016} proposed a modified version of the algorithm proposed 
+by \citet{Li2011}. These modifications can be summed up in:
+(1) a faster calculation of the distance from the computation point to 
+the tesseroid and 
+(2) the application of a stack based algorithm that speeds up the 
+computation and gives more control over the recursion step.
+More specifically, \citet{Uieda2016} redefined the dimensions of an 
+arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$. The former ones 
+are the arc-distances measured along its top surface while $L_r$ is the 
+difference between the radii of the top and bottom surfaces.
+
+While keeping the GLQ fixed on second orders, the adaptive 
+discretization algorithm starts by checking that the following 
+inequality is satisfied:
+
+\begin{equation}
+    \frac{d}{L_i} \geq D,
+\label{eq:distance-size-ratio}
+\end{equation}
+
+\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance 
+between the geometric centre of the tesseroid and the computation 
+point, and $D$ is a predefined value called distance-size ratio.
+If the inequality \ref{eq:distance-size-ratio} is not held for at least 
+one spherical coordinate, then the tesseroid must be divided by half in 
+that direction.
+On the contrary, if the inequality is satisfied for every spherical 
+coordinate, the algorithm computes the desired gravity field using the 
+GLQ approximation.
+A complete description of the modified adaptive discretization 
+algorithm can be found in \citet{Uieda2016}.
+
+In summary, the distance-size ratio $D$ determines indirectly how many 
+times the tesseroids will be divided, and therefore efficiently 
+regulates both the accuracy of the algorithm and its computation time.
+Moreover, it limits the computation time by stopping the division of 
+the tesseroids when their dimensions are lower than a certain value 
+(1mm), assuming that such small tesseroids do not modifies 
+significantly the desired gravitational field.
+
+On the other hand, the value assigned to the distance-size ratio $D$ 
+cannot be easily related to the error of the approximation, thus the 
+choice of the value of $D$ to get an acceptable accuracy must be 
+empirically determined beforehand.
+In order to overcome this, \citet{Uieda2016} compared the numerical 
+computation with the analytical solution for a spherical shell with 
+homogeneous density.
+
+We will perform similar tests, but now with variable density spherical 
+shell.
+From the infinite density functions that can be proposed, we have 
+chosen three characteristic ones: a linear, an exponential and a 
+discontinuous function.
+\todo{sacaria todo este parrafo y lo introduciria mas adelante}
+
+\subsection{Density-based Discretization Algorithm}
+
+High variations in the density function of an arbitrary tesseroid 
+introduces a new type of error into the numerical approximation: only a 
+few point masses would not reproduce the effect of this kind of 
+variations.
+The modified version of the adaptive discretization algorithm may help 
+in reducing this kind of error, but as it only takes into account the 
+dimensions of the tesseroid and the distance to the computation point, 
+it's not well suited to fully perform this task.
+
+In order to overcome this, we developed a complementary discretization 
+algorithm that takes into account the variation of the density inside 
+the tesseroid. Basically, it subdivides any tesseroid on the radial 
+direction on the depth at which the \emph{maximum density variation} 
+takes place.
+
+In a general way, the algorithm can be summarised in the following steps:
+
+\begin{enumerate}
+\renewcommand{\theenumi}{(\arabic{enumi})}
+    \item Normalise the density function to the [0, 1] interval,
+    \item Compute the absolute difference between the normalised 
+          density and a linear function that assumes the same values on the 
+          top and bottom radii that this normalised density.
+    \item If product of the maximum of the absolute difference and a 
+          tesseroid size ratio is greater that a predefined delta ratio 
+          ($\delta$), then the tesseroid will be subdivided on the 
+          radius at which this maximum takes place. Otherwise, the 
+          tesseroid won't be split.
+    \item If it's discretized, then apply items (2), (3) and (4) for 
+          every subdividing tesseroid.
+\end{enumerate}
+
+Lets assume an arbitrary tesseroid with boundary radii $R_1$ and $R_2$ 
+and a density function $\rho(r')$ which we will call the 
+\emph{original tesseroid}, whose radial dimension can be written as
+$L_r = R_2 - R_1$.
+
+First we need to normalise the density function such as its minimum and 
+maximum will be 0 and 1, respectively.
+We call this normalised density $\rho_n(r')$ and we calculate it as 
+follows:
+
+\begin{equation}
+    \rho_n(r') = \frac{\rho(r') - \rho_1}{\rho_2 - \rho_1},
+\end{equation}
+\noindent where
+\begin{equation}
+    \rho_1 = \text{min}\{ \rho(r') \}, \quad
+    \rho_2 = \text{max}\{ \rho(r') \} \quad
+    \forall \, r' \in [R_1, R_2].
+\end{equation}
+
+\noindent Because this algorithm will be iteratively applied for every 
+subdividing tesseroid, we'd like to note that this normalised density 
+function will be the same for every subdividing tesseroid.
+In case that the density function is constant, both the maximum and 
+minimum densities are equal and the normalised function could not 
+be defined, so the density-based discretization algorithm won't be 
+applied.
+
+Secondly, we must define a linear function $\rho_l(r')$ that assumes 
+the same values as the normalised density $\rho_n(r')$ on the boundary 
+radii $R_1$ and $R_2$:
+
+\begin{equation}
+    \rho_l(r') =
+    \frac{ \rho_n(R_2) - \rho_n(R_1) }{ R_2 - R_1 } (r' - R_1) + \rho_n(R_1),
+    \label{eq:density-reference-line}
+\end{equation}
+
+\noindent and compute the absolute difference between this linear 
+function and the normalised density:
+
+\begin{equation}
+    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |.
+    \label{eq:density-abs-diff}
+\end{equation}
+
+Finally, if the product of the maximum of the absolute difference and a 
+tesseroid size ratio is greater than a predefined delta ratio $\delta$ 
+we will divide the tesseroid on the radius at which this maximum takes 
+place ($R_\text{max}$):
+
+\begin{equation}
+    \text{max}\{ \Delta \rho(r') \} \frac{L_r^\text{sm}}{L_r} > \delta,
+    \label{eq:delta-density}
+\end{equation}
+\begin{equation}
+    \text{max}\{ \Delta \rho(r') \} = \Delta \rho(R_\text{max}),
+\end{equation}
+
+\noindent obtaining a pair of smaller tesseroids with the same 
+longitude and latitude boundaries.
+
+The $L_r^\text{sm}$ is the radial dimension of the tesseroid that is 
+currently analysed. On the first iteration, this tesseroid is the 
+\emph{original} one, so $L_r^\text{sm} = L_r$, but for future 
+iterations this radial dimension will be computed as the difference of 
+the top and bottom radii of the subdividing tesseroid.
+
+This algorithm will be subsequently applied to every subdividing 
+tesseroid.
+We will start from equation \ref{eq:density-reference-line} defining a 
+new linear function $\rho_l(r')$, an absolute difference with equation 
+\ref{eq:density-abs-diff} and evaluate the discretization inequation 
+\ref{eq:delta-density}, taking into account that now $R_1$ and $R_2$ 
+are the radial boundaries of this smaller tesseroid.
+
+The tesseroid size ratio included in the inequation 
+\ref{eq:delta-density} is intended to reduce the number of probable 
+unneeded discretizations and focus on discretizations that will 
+significantly reduce the numerical error.
+A big subdiving tesseroid with small maximum absolute difference would 
+introduce more numerical error than a small subdividing tesseroid with 
+higher maximum absolute difference.
+
+If on the contrary the inequation \ref{eq:delta-density} is not held, 
+then the tesseroid will not be subdivided.
+The same happens if the tesseroid radial dimension is bellow the 
+threshold of 1mm. On the first case due to the fact that no more 
+discretizations are needed, and on the second one because smaller 
+tesseroids will not introduce meaningful increases in the accuracy of 
+the computation.
+This two conditions guarantee that the algorithm is finite.
+Once they are met, every subdividing tesseroid is subjected to the 
+modified adaptive discretization algorithm and the GLQ in order to 
+compute the gravity field generated by it.
+
+It's easy to see that the delta ratio indirectly controls how many 
+times the tesseroids will be subdivided based on the variation of the 
+density function.
+
+Therefore it determines the accuracy involving the numerical error that 
+the density variation introduces and also the computation time.
+This raises the need to determine a maximum value of $\delta$ that 
+ensures an acceptable accuracy while minimising the computation time. 
+We will perform this determination with the same methodology as the one 
+applied to the $D$ ratio: compare the numerical model with the 
+analytical solutions for a spherical shell.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Analytical Solutions for Spherical Shell}
@@
 \end{enumerate}
-
-\subsection{Modified Adaptive Discretization Algorithm}
-
-From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates the volume integrals as the gravitational effect of point masses located at the nodes of the quadrature.
-If we set the GLQ orders to fixed values, it's easy to conclude that the nearer the computation point is to the tesseroid, the accuracy of the approximation is reduced and the point masses effect becomes more noticeable than the tesseroid we want to approximate.
-A complete analysis of this conclusion can be found in \citet{Ku1977, Li2011, Uieda2016}.
-
-Instead of increasing the GLQ orders to improve the accuracy, what leads to a more time consuming computation, adaptive discretization algorithms have been applied in previous works \citep{Li2011, Uieda2016}.
-They essentially consist in dividing the tesseroid based on a ratio between the distance from its geometric centre to the computation point and its dimensions.
-
-The main differences between the modified version made by \citet{Uieda2016} and the original one developed by \citet{Li2011} can be summed up in:
-(1) a faster calculation of the distance from the computation point to the tesseroid and 
-(2) the application of a stack based algorithm that speeds up the computation and gives more control over the recursion step.
-More specifically, \citet{Uieda2016} redefined the dimensions of an arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$.
-The former ones are the arc-distances measured along its top surface while $L_r$ is the difference between the radii of the top and bottom surfaces.
-
-While keeping the GLQ fixed on second orders, the adaptive discretization algorithm starts by checking that the following inequality is satisfied:
-
-\begin{equation}
-    \frac{d}{L_i} \geq D,
-\label{eq:distance-size-ratio}
-\end{equation}
-
-\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance between the geometric centre of the tesseroid and the computation point, and $D$ is a predefined value called distance-size ratio.
-If the inequality \ref{eq:distance-size-ratio} is not held for at least one spherical coordinate, then the tesseroid must be divided by half in that direction.
-A complete description of the modified adaptive discretization algorithm can be found in \citet{Uieda2016}.
-On the contrary, if the inequality is satisfied for every spherical coordinate, the algorithm computes the desired gravity field using the GLQ approximation.
-
-In summary, the distance-size ratio $D$ determines indirectly how many times the tesseroids will be divided, and therefore efficiently regulates both the accuracy of the algorithm and its computation time.
-Moreover, it limits the computation time by stopping the division of the tesseroids when their dimensions are lower than a certain value (1mm), assuming that such small tesseroids do not modifies significantly the desired gravitational field.
-
-On the other hand, the value assigned to the distance-size ratio $D$ cannot be easily related to the error of the approximation, thus the choice of the value of $D$ to get an acceptable accuracy must be empirically determined beforehand.
-In order to overcome this, \citet{Uieda2016} compared the numerical computation with the analytical solution for a spherical shell with homogeneous density.
-
-We will perform similar tests, but now with variable density spherical shell.
-From the infinite density functions that can be proposed, we have chosen three characteristic ones: a linear, an exponential and a discontinuous function.
-
-\subsection{Density-based Discretization Algorithm}
-
-High variations in the density function of an arbitrary tesseroid introduces a new type of error into the numerical approximation: only a few point masses would not reproduce the effect of this kind of variations.
-The modified version of the adaptive discretization algorithm may help in reducing this kind of error, but as it only takes into account the dimensions of the tesseroid and the distance to the computation point, it's not well suited to fully perform this task.
-
-In order to overcome this, we developed a complementary discretization algorithm that takes into account the variation of the density inside the tesseroid.
-Basically, it subdivides any tesseroid on the radial direction on the depth at which the \emph{maximum density variation} takes place.
-
-Specifically, the algorithm consists in the following steps:
-
-\begin{enumerate}
-\renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Normalise the density function to the [0, 1] interval,
-    \item Compute the absolute difference between the normalised density and a linear function that assumes the same values on the top and bottom radii that this normalised density.
-    \item If the maximum of the absolute difference is greater that a predefined delta ratio ($\delta$), then the tesseroid will be subdivided on the radius at which this maximum takes place.
-Otherwise, the tesseroid won't be split.
-\end{enumerate}
-
-First we need to normalise the density function such as it assumes the values of 0 and 1 on the boundary radii ($R_1$, $R_2$), depending at which radius the original density function is maximum.
-We call this normalised density $\rho_n(r')$ and we calculate it as follows:
-
-\begin{equation}
-    \rho_n(r') = \frac{\rho(r') - \rho_1}{\rho_2 - \rho_1},
-\end{equation}
-\noindent where
-\begin{equation}
-    \rho_1 = \text{min}\{ \rho(R_1), \rho(R_2) \}, \quad
-    \rho_2 = \text{max}\{ \rho(R_1), \rho(R_2) \}.
-\end{equation}
-
-Secondly, we must define a linear function $\rho_l(r')$ that assumes the same values as the normalised density $\rho_n(r')$ on the boundary radii:
-
-\begin{equation}
-    \rho_l(r') =
-    \frac{ \rho_n(R_2) - \rho_n(R_1) }{ R_2 - R_1 } (r' - R_1) + \rho_n(R_1),
-\end{equation}
-
-\noindent and compute the absolute difference between this linear function and the normalised density:
-
-\begin{equation}
-    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |.
-\end{equation}
-
-Finally, if the maximum of the absolute difference if greater than a predefined delta ratio $\delta$ we will divide the tesseroid on the radius at which this maximum takes place ($R_\text{max}$):
-
-\begin{equation}
-    \text{max}\{ \Delta \rho(r') \} > \delta,
-    \label{eq:delta-density}
-\end{equation}
-\begin{equation}
-    \text{max}\{ \Delta \rho(r') \} = \Delta \rho(R_\text{max}),
-\end{equation}
-
-\noindent obtaining a pair of smaller tesseroids with the same longitude and latitude boundaries.
-
-In case that the original density function assumes the same value at both boundary radii, the procedure is the same except that the constants $\rho_1$ and $\rho_2$ are redefined as follows:
-
-\begin{equation}
-    \rho_1 = \text{min}\{ \rho(r') \}, \quad
-    \rho_2 = \text{max}\{ \rho(r') \}.
-\end{equation}
-
-If on the contrary the inequation \ref{eq:delta-density} is not held, then the tesseroid will not be subdivided.
-The same happens if the density function is constant or if the tesseroid radial dimension is bellow the threshold of 1mm.
-On the first case due to the fact that the density-based discretization algorithm is not needed, and on the second one because smaller tesseroids will not introduce meaningful increases in the accuracy of the computation.
-
-This algorithm can be subsequently applied to every subdividing tesseroid, until the inequation \ref{eq:delta-density} is not held or the subdivisions are smaller than 1mm.
-This two conditions guarantee that the algorithm is finite.
-Once this conditions are met, every subdividing tesseroid is subjected to the modified adaptive discretization algorithm and the GLQ in order to compute the gravity field generated by it.
-
-It's easy to see that the delta ratio indirectly controls how many times the tesseroids will be subdivided based on the variation of the density function.
-Therefore it determines the accuracy involving the numerical error that the density variation introduces and also the computation time.
-This raises the need to determine a maximum value of $\delta$ that ensures an acceptable accuracy while minimising the computation time.
-We will perform this determination with the same methodology as the one applied to the $D$ ratio: compare the numerical model with the analytical solutions for a spherical shell.
 

### 2018-04-05 manuscript/manuscript.tex
Added Algorithm summary in manuscript

@@
 latitude, assigning an exponential density variation to it and computing all 
-the gravity fields it generates.
+the gravitational fields it generates.
 \end{summary}
@@
 radial boundaries (see Figure \ref{fig:tesseroid-uieda}).
-The calculation of the gravity fields generated by an arbitrary tesseroid on 
-any external point involves the resolution of volume integrals that are 
-generally approximated by numerical computations.
+The calculation of the gravitational fields generated by an arbitrary 
+tesseroid on any external point involves the resolution of volume 
+integrals that are generally approximated by numerical computations.
 The literature offers two approaches: one involves Taylor series expansion 
@@
 We define an external computation point $P(r, \phi, \lambda)$ where the 
-gravity fields generated by the tesseroid are going to be calculated 
+gravitational fields generated by the tesseroid are going to be calculated 
 with respect to the local north oriented Cartesian coordinate system 
@@
 coordinate.
-Thus, the integrals for the gravity fields are slightly modified:
+Thus, the integrals for the gravitational fields are slightly modified:
 
@@
 From equation \ref{eq:glq-var-dens} we can observe that the GLQ 
-approximates the gravity fields of a tesseroid as the ones generated by 
-$N_\lambda N_\phi N_r$ point masses located at the nodes of the 
-Legendre polynomials, where $N_i$ are the orders of the quadrature for 
-each integration, with $i \in \{ \lambda, \phi, r \}$, in the same way 
-as the GLQ works in the homogeneous density case.
+approximates the gravitational fields of a tesseroid as the ones 
+generated by $N_\lambda N_\phi N_r$ point masses located at the nodes 
+of the Legendre polynomials, where $N_i$ are the orders of the 
+quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$, 
+in the same way as the GLQ works in the homogeneous density case.
 Nevertheless, in this new variable density scenario, the information of 
@@
 show that this method, along with well suited discretization 
-algorithms, approximates the gravity fields with good accuracy.\todo{no 
-se si poner esta ultima oracion}
+algorithms, approximates the gravitational fields with good accuracy.
+\todo{no se si poner esta ultima oracion}
 
@@
 On the contrary, if the inequality is satisfied for every spherical 
-coordinate, the algorithm computes the desired gravity field using the 
+coordinate, the algorithm computes the desired gravitational field using the 
 GLQ approximation.
@@
 modified adaptive discretization algorithm and the GLQ in order to 
-compute the gravity field generated by it.
+compute the gravitational field generated by it.
 
@@
 analytical solutions for a spherical shell.
+
+\subsection{Algorithm summary}
+
+When computing any gravitational field generated by an arbitrary tesseroid on 
+a specific computation point, we present the following methodology:
+
+\begin{enumerate}
+    \renewcommand{\theenumi}{(\arabic{enumi})}
+    \item Apply the density-based discretization algorithm in case of 
+          variable density producing subdivisions of the original tesseroid,
+    \item Apply the modified adaptive discretization algorithm for 
+          each subdivision, generating a set of smaller tesseroids,
+    \item Compute the desired gravitational field for every small tesseroid 
+          through the GLQ approximation with fixed quadrature orders
+          (equation \ref{eq:glq-var-dens}).
+\end{enumerate}
+
+In most practical cases, the forward model will consist in a set of 
+tesseroids whose gravitational effect will be computed on a set of 
+computation points.
+These steps can be applied on every tesseroid and for each computation point,
+adding the contribution of each tesseroid to the resulting 
+gravitational field.
 
@@
 \end{enumerate}
-
-\subsection{Algorithm summary}
-
-When computing any gravity field generated by an arbitrary tesseroid on a specific computation point, the new software:
-
-\begin{enumerate}
-    \renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Applies the density-based discretization algorithm in case of variable density producing subdivisions of the original tesseroid,
-    \item Applies the modified adaptive discretization algorithm for each subdivision, generating a set of smaller tesseroids,
-    \item Computes the desired gravity field for every small tesseroid through the GLQ approximation (equation \ref{eq:glq-var-dens}).
-\end{enumerate}
-
-These steps are applied for each computation point, adding the contribution of each tesseroid to the resulting gravity field.
 

### 2018-04-05 manuscript/manuscript.tex
Reorganized Analytical solutions for spherical shell in manuscript

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Determination of distance-size and delta ratios}
+
+In order to determine default values for the distance-size ratio ($D$)
+and the delta ratio ($\delta$) that ensure both numerical accuracy and 
+computation effectiveness we need to compare the numerical 
+approximation with some analytical solutions.
+
+On the homogeneous density tesseroids case, \citet{Uieda2016} and 
+\citet{Grombein2013} compared their numerical model with the analytical 
+solution of a spherical shell.
+We could follow this idea, but for our needs the spherical shell must 
+have the same density function as the numerical model, although the 
+analytical solution in this case must be derived.
+
 \subsection{Analytical Solutions for Spherical Shell}
 
-The integrals that define the gravity fields in equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have analytical solutions for the exceptional case of a spherical shell with homogeneous density \citep{Mikuska2006,Grombein2013}.
-In order to test the accuracy of their approximations, \citet{Uieda2016} and \citet{Grombein2013} compared their numerical models (tesseroids with homogeneous density) with these analytical solutions.
-
-In case of variable density tesseroids, the comparison must obviously be made with a variable density spherical shell, whose analytical solution will depend on the chosen density function.
-Because the literature has not offered it yet, we had to develop it ourselves. \todo{ no se si ponerlo asi}
-
-Lets consider a spherical shell with inner and outer radii $R_1$ and $R_2$, respectivetly, whose density depends on the radial spherical coordinate (Figure \ref{fig:spherical-shell}).
-The gravity potential generated by the shell on an arbitrary external point $Q(0,0,r)$, located along the $z$ axe at a distance $r$ from the origin, can be written as follows:
+The integrals that define the gravity fields in equations 
+\ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have analytical 
+solutions in case of a spherical shell with homogeneous density 
+\citep{Mikuska2006,Grombein2013}.
+But for variable density spherical shell the problem wasn't address in 
+the literature yet.
+
+Lets consider a spherical shell with inner and outer radii $R_1$ and 
+$R_2$, respectivetly, whose density depends on the radial spherical 
+coordinate (Figure \ref{fig:spherical-shell}).
+The gravity potential generated by the shell on an arbitrary external 
+point $Q(0,0,r)$, located along the $z$ axe at a distance $r$ from the 
+origin, can be written as follows:
 
@@
     Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
-The computation point $Q$ is located on the $z$ axe at a distance $r$ from the centre of the shell.
-For our purposes we will assume that $Q$ is outside the outer radii, i.e. $r > R_2$.
+    The computation point $Q$ is located on the $z$ axe at a distance $r$ from
+    the centre of the shell.
+    For our purposes we will assume that $Q$ is outside the outer radii,
+    i.e. $r > R_2$.
     }
@@
 \noindent where $\ell$ is defined in equation \ref{eq:ell}.
-The computation point $Q$ is located at latitude $\phi=90^\circ$, so $\ell$ and $\cos\psi$ (defined in equation \ref{eq:cospsi}) get simplified:
+The computation point $Q$ is located at latitude $\phi=90^\circ$, so 
+$\ell$ and $\cos\psi$ (defined in equation \ref{eq:cospsi}) get 
+simplified:
 
@@
 
-Due to the rotational symmetry along the $z$ axe, the integration in $\lambda'$ is straightforward:
+Due to the rotational symmetry along the $z$ axe, the integration in 
+$\lambda'$ is straightforward:
 
@@
 
-\noindent while the integration in $\phi'$ can be performed independently of the density function.
-Making use of SymPy \citep{sympy2017}, a Python library for symbolic mathematics, we obtained the following expression for the potential:
+\noindent while the integration in $\phi'$ can be performed 
+independently of the density function.
+Making use of SymPy \citep{sympy2017}, a Python library for symbolic 
+mathematics, we obtained the following expression for the potential:
 
@@
 
-For our purposes we can assume that the computation point $Q$ is farther from the origin than the outer radius, i.e. $r>R_2$. 
-And taking into account that $r'$ is always lower than $R_2$, we can simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
+For our purposes we can assume that the computation point $Q$ is 
+farther from the origin than the outer radius, i.e. $r>R_2$. 
+And taking into account that $r'$ is always lower than $R_2$, we can 
+simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
 
@@
 
-The equation \ref{eq:shell-pot} allows to easily obtain the exact gravity potential generated by a spherical shell with a variable density in depth on any outside point located on the $z$ axe.
-Due to the rotational symmetry along any axe that passes through the centre of the shell, the expression in equation \ref{eq:shell-pot} reproduces the gravity potential field on any outside point at distance $r$ from the its centre.
-
-The gradient and the Marussi tensor derived from potentials that depends solely on $r$ have only a few non zero components: the vertical component of the gradient ($g_z$) and the diagonal components of the tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
-They can be easily obtained as \citet{Grombein2013} does, even for any arbitrary density function $\rho(r')$:
+The equation \ref{eq:shell-pot} allows to easily obtain the exact 
+gravity potential generated by a spherical shell with a variable 
+density in depth on any outside point located on the $z$ axe.
+Due to the rotational symmetry along any axe that passes through the 
+centre of the shell, the expression in equation \ref{eq:shell-pot} 
+reproduces the gravity potential field on any outside point at distance 
+$r$ from the its centre.
+
+The gradient and the Marussi tensor derived from potentials that 
+depends solely on $r$ have only a few non zero components: the vertical 
+component of the gradient ($g_z$) and the diagonal components of the 
+tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
+They can be easily obtained as \citet{Grombein2013} does, even for any 
+arbitrary density function $\rho(r')$:
 
@@
 
-For future purposes we are going to obtain the gravity potential for three different functions from the perspective of their variation in $r'$: a linear and an exponential density function.
+For future purposes we are going to obtain the gravity potential for 
+three different functions from the perspective of their variation in 
+$r'$: a linear and an exponential density function.
 

### 2018-04-05 manuscript/manuscript.tex
Reorganized "Linear density" in manuscript

@@
 $r'$: a linear and an exponential density function.
+\todo{lo sacaria}
+
+\subsection{Linear Density}
+
+Lets consider a linear density function like:
+
+\begin{equation}
+    \rho(r') = ar' + b.
+\end{equation}
+
+\noindent The gravitational potential generated by a spherical shell 
+with this density on any external point at distance $r$ from its centre 
+can be obtained from the following expression:
+
+\begin{equation}
+    V_\text{sh}^\text{lin}(r) = \pi G \left[ 
+    a \frac{R_2^4 - R_1^4}{r} +
+    b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
+    \label{eq:shell-pot-linear}
+\end{equation}
+
+\noindent The first term on equation \ref{eq:shell-pot-linear} 
+reproduces the potential generated by a spherical shell with variable 
+density $\rho(r') = ar'$, while the second one constitutes the 
+potential generated by a spherical shell with homogeneous density $\rho 
+= b$ \citep{Mikuska2006,Grombein2013}. So the potential can be read as 
+the one generated by a combination of both shells.
+
+When applying the density-based discretization algorithm on any 
+tesseroid with linear density, the absolute difference defined in 
+equation \ref{eq:density-abs-diff} will always be zero, so the 
+discretization inequation \ref{eq:delta-density} will never be 
+satisfied.
+Therefore, independently of the linear function proposed, the 
+density-based algorithm will never be applied, leaving the modified 
+adaptive discretization algorithm as the only mechanism to control the 
+accuracy of the numerical approximation.
+That's why we only need to determine the minimum value of $D$ needed in 
+order to guarantee an acceptable accuracy while ignoring the value of 
+$\delta$.
+
+In order to compare the numerical model with the analytical solution we 
+must build a spherical shell from a set of tesseroids.
+We propose a mesh of $30^\circ \times 30^\circ$ tesseroids to compute 
+the gravitational potential, the vertical component of the gradient 
+($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, 
+$g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the 
+pole, (2) another one on the Equator, (3) one located at satellite 
+height and (4) a big grid of $30^\circ \times 30^\circ$.
+More information about this grids can be found on Table 
+\ref{tab:grids}.
+
+We have computed these gravitational fields on every point of the four 
+grids for different values of $D$, ranging from 0.5 to 10 with a step 
+of 0.5.
+Then we calculated the maximum difference between these results and the 
+value that the analytical solution assumes at the same height.
+Because the gravity fields generated by a tesseroid are approximated by 
+the effect of point masses, the numerical results may vary between 
+computation points at same height but at different latitude or 
+longitude.
+That's why we choose for comparison the maximum difference between the 
+results on every point of the grid and the analytical solution.
+
+Finally, we will set the acceptable value of $D$ as the minimum value at which the
+corresponding error of the numerical approximation is lower than 0.1\%.
+
+\begin{table}
+\caption{
+    Description of the synthetic grids on which the comparison of the 
+    numerical model against the analytical solutions for a spherical 
+    shell was done.
+    This collection of grids tries to include every possible scenario 
+    on which the numerical model can present a different behaviour 
+    depending on the size of the grid, its location or its height over 
+    the Earth surface. Every grid consists in a set of 10$\times$10 
+    points.
+}
+\label{tab:grids}
+\begin{tabular}{lcccc}
+    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
+    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ &
+        $0^\circ - 1^\circ$ & 2km \\
+    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ &
+        $0^\circ - 1^\circ$ & 2km \\
+    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ &
+        $0^\circ - 1^\circ$ & 260km \\
+    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ &
+        $0^\circ - 30^\circ$ & 2km \\
+\end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     \Big[
-    & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-    & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
     \Big].

### 2018-04-05 manuscript/manuscript.tex
Modified linear density in manuscript

@@
 
-Finally, we will set the acceptable value of $D$ as the minimum value at which the
-corresponding error of the numerical approximation is lower than 0.1\%.
+Finally, we will set the acceptable value of $D$ as the minimum value 
+at which the corresponding error of the numerical approximation is 
+lower than 0.1\%.
 
@@
 
+In order to perform the comparison, we proposed a thin and a thick 
+spherical shells with 1km and 35km of thickness, respectively, whose 
+outer radii is equal to the mean Earth radius.
+On both cases we will suppose that the density assumes a value of 
+2670~kg/m$^3$ on the outer surface and 3300~kg/m$^3$ at the inner 
+radius:
+
+\begin{equation}
+    \rho(r') = ar' + c,
+    \label{eq:density-linear}
+\end{equation}
+\noindent with 
+\begin{equation}
+    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R_2 - R_1},
+\end{equation}
+\begin{equation}
+    c = \frac{3300\text{kg/m$^3$} - 
+        2670\text{kg/m$^3$}}{R_2 - R_1} R + 
+        2670\text{kg/m$^3$},
+\end{equation}
+
+\noindent where $R$ = 6378.137 km is the mean Earth radius.
+
+Then we computed the percentage difference between the numerical model 
+with its corresponding analytical solution (equation 
+\ref{eq:density-linear}) for the thick and the thin spherical shell.
+
+We can see the results for the potential, the vertical component of the 
+gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on 
+Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
+We exclude the results corresponding to the other two diagonal 
+components of the tensor because they are similar to the ones obtained 
+for the $g_{zz}$, although they can be found on the repository. 
+\todo{chequear}
+
+For the linear density case, we can observe that the calculation of the 
+potential and the $g_z$ component of its gradient needs a value of 
+$D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
+On the other hand, the $g_{zz}$ computations show a noticeable 
+difference between the thin and the thick shell: the first one needs a 
+value of $D$ equal to 8 while the later only a $D$ of 3.5.
+We will keep the more conservative result of $D$ equal to 8 in case of 
+computing the Marussi tensor components in case of a linear density 
+tesseroid.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
+\caption{
+    Differences between the gravity fields generated by the numerical 
+    model and the analytical solution for a thin spherical shell of 1km 
+    of thickness with the linear density defined on equation 
+    \ref{eq:density-linear}.
+    The computations were performed on the four grids described in 
+    Table \ref{tab:grids}, for different values of the distance-size 
+    ratio $D$ and without applying the density-based discretization 
+    algorithm.
+    If the difference is less than 0.1\%, we consider that the model 
+    has achieved an acceptable accuracy for the corresponding value of 
+    $D$.
+    }
+\label{fig:D-linear-thin}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
+\caption{
+    Differences between the gravity fields generated by the numerical 
+    model and the analytical solution for a thick spherical shell of 
+    35km of thickness with the linear density defined on equation 
+    \ref{eq:density-linear}.
+    The computations were performed on the four grids described in 
+    Table \ref{tab:grids}, for different values of the distance-size 
+    ratio $D$ and without applying the density-based discretization 
+    algorithm.
+    If the difference is less than 0.1\%, we consider that the model 
+    has achieved an acceptable accuracy for the corresponding value of 
+    $D$.
+    }
+\label{fig:D-linear-thick}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-04-10 manuscript/manuscript.tex
Started writing new exponential density subsection in manuscript

@@
 
+\subsection{Exponential Density}
+
+A more complex density variation than a linear one could be an 
+exponential function like:
+
+\begin{equation}
+    \rho(r') = Ae^{-(r' - \Delta h)/b},
+\end{equation}
+
+\noindent where $A$, $\Delta h$ and $b$ are constants.
+The gravitational potential field generated by a spherical shell with 
+this kind of density function on any external point at a distance $r$ 
+from its centre can be calculated as follows:
+
+\iftwocol{
+\begin{equation}
+    \begin{split}
+        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+        Ab 
+        \Big[
+        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
+        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
+        \Big].
+    \end{split}
+\label{eq:shell-pot-exp}
+\end{equation}
+}{
+\begin{equation}
+    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+    Ab \, e^\frac{\Delta h}{b}
+    \Big[
+    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+    \Big].
+\label{eq:shell-pot-exp}
+\end{equation}
+}
+
+This expression will allow us  to perform comparisons between the 
+anaytical solutions and the numerical approximation, similar as the 
+ones made for the linear density case.
+Although in the exponential density scenario the problem is more 
+complex due to the fact that the density-based discretization must be 
+used together with the modified adaptive discretization algorithm.
+This means that both the distance-size ratio $D$ and the delta ratio 
+$\delta$ must be simultaneously determined.
+
+From every $D$ and $\delta$ combination we want to find the one that 
+produces a numerical error lower than the 0.1\% threshold while keeping 
+an effective computation, i.e. the ($D$, $\delta)$ pair that produces 
+the minimum number of discretizations that ensure an acceptable 
+numerical precision.
+We can propose an heuristic method to perform this task: compute the 
+numerical error for every ($D$, $\delta$) points belonging to a grid on 
+the $D$, $\delta$ space and specify the pair that combines an 
+acceptable numerical error with an efficient computation.
+
+Because the algorithm is intended to work with default $D$ and $\delta$ 
+values that ensures these two conditions for every density function, 
+the default value for $D$ cannot be lower than the one determined for 
+the linear function.
+That's why we can only choose the ($D$, $\delta$) points on which
+$D \geq D_\text{linear}$.
+In this way we can divide the $D$, $\delta$ space in two subregions: 
+one with $D < D_\text{linear}$ and other with $D \ge D_\text{linear}$.
+The former contains points not eligible, while the latter one will 
+contain the one that we want to find.
+
+From the latter subregion we can explore the numerical error of points 
+with $D = D_\text{linear}$ and different values of $\delta$.
+We will call $\delta_p$ as the maximum $\delta$ at which the numerical 
+error is lower than the 0.1\% threshold while keeping
+$D = D_ \text{linear}$.
+Greater $\delta$ values than $\delta_p$ will produce computations 
+with high error, while lower $\delta$ values than $\delta_p$ will 
+produce more discretizations leading to non-efficient computation.
+
+With this definition in mind we can subdivide the subregion in two: one 
+with points with $\delta > \delta_p$ and other with points with $\delta 
+\le \delta_p$.
+
+It worth mention that a greater $D$ or a lower $\delta$ 
+increases the number of discretizations of the tesseroid.
+Although a greater $D$ may subdivide it in the latitude, longitude 
+and/or radial direction; a lower $\delta$ will only discretize the 
+original tesseroid in the radial direction.
+From this analysis we can conclude that a decrease in $\delta$ will 
+produce a more efficient computation than an increase in $D$.
+
+Every point in the subregion with $\delta \le \delta_p$ will have a 
+numerical error lower than the 0.1\% threshold, but they will always 
+produce more discretizations than the ($\D_\text{linear}$, $\delta_p$) 
+point leading to non-efficient computations.
+
+On the other hand, the numerical error of points with $\delta > 
+\delta_p$ can either be lower or greater than 0.1\%.
+Any point in this subregion with acceptable numerical error will 
+produce more discretizations than the ($\D_\text{linear}$, $\delta_p$) 
+point, due to $D$ value is increased.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-04-12 manuscript/manuscript.tex
Fixed wrong math in manuscript

@@
 numerical error lower than the 0.1\% threshold, but they will always 
-produce more discretizations than the ($\D_\text{linear}$, $\delta_p$) 
+produce more discretizations than the ($D_\text{linear}$, $\delta_p$) 
 point leading to non-efficient computations.
@@
 Any point in this subregion with acceptable numerical error will 
-produce more discretizations than the ($\D_\text{linear}$, $\delta_p$) 
+produce more discretizations than the ($D_\text{linear}$, $\delta_p$) 
 point, due to $D$ value is increased.

### 2018-04-12 manuscript/manuscript.tex
Density-based discretization figure changed to only one line
Added density-based discretization figure explanation on manuscript

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\textwidth]
+    {figures/density-based-discretization-algorithm.pdf}
+\caption{
+    Example intended to show how the density-based algorithm works.}
+\label{fig:density-discretization-algorithm}
+\end{figure*}
+
 In a general way, the algorithm can be summarised in the following steps:
@@
 
-Lets assume an arbitrary tesseroid with boundary radii $R_1$ and $R_2$ 
-and a density function $\rho(r')$ which we will call the 
-\emph{original tesseroid}, whose radial dimension can be written as
-$L_r = R_2 - R_1$.
+Lets assume an arbitrary \emph{original} tesseroid with boundary radii 
+$R_1$ and $R_2$ and a density function $\rho(r')$, whose radial 
+dimension can be written as $L_r = R_2 - R_1$.
 
@@
 
-\noindent Because this algorithm will be iteratively applied for every 
+Because the algorithm will be iteratively applied for every 
 subdividing tesseroid, we'd like to note that this normalised density 
@@
 
+On Figure \ref{fig:density-discretization-algorithm}(a) we propose an 
+arbitrary normalised density function as an example of how this 
+algorithm works.
+And we plotted its corresponding linear function $\rho_l(r')$ and a 
+dashed line that represents the maximum absolute difference $\Delta 
+\rho (r')$.
+We will assume that the discretization inequation 
+\ref{eq:delta-density} holds in this case, so we must divide the 
+tesseroid at the height at which the maximum difference occurs.
+The dots on Figure \ref{fig:density-discretization-algorithm} 
+represent the discretization heights, i.e. the radial boundaries of the 
+subdiving tesseroids produced by the algorithm.
+
+On Figure \ref{fig:density-discretization-algorithm}(b) we apply the 
+algorithm on the shallower subdividing tesseroid.
+In the same way, we propose the linear function and compute the maximum 
+absolute difference $\Delta \rho (r')$.
+In this case, the tesseroid size-ratio $L_r^\text{sm}/L_r$ is lower 
+than one and will be taken into account in the discretization 
+inequation, which will be assumed to hold also in this case, leading to 
+the division of this tesseroid.
+
+On Figure \ref{fig:density-discretization-algorithm}(c) we perform the 
+same analysis on the deeper subdiving tesseroid, and assuming that the 
+inequality holds in this case we also subdivide it.
+
+Finally, if we suppose that the value of $\delta$ used in this example 
+does not produce further discretizations, the resulting subdiving 
+tesseroids can be seen on Figure 
+\ref{fig:density-discretization-algorithm}(d).
+
 It's easy to see that the delta ratio indirectly controls how many 
@@
 density function.
-
 Therefore it determines the accuracy involving the numerical error that 

### 2018-04-13 manuscript/manuscript.tex
Small changes in manuscript
Modifications in manuscript: improvements in Exponential Density

@@
 an effective computation, i.e. the ($D$, $\delta)$ pair that produces 
-the minimum number of discretizations that ensure an acceptable 
+the minimum number of discretizations and ensures an acceptable 
 numerical precision.
@@
 one with $D < D_\text{linear}$ and other with $D \ge D_\text{linear}$.
-The former contains points not eligible, while the latter one will 
-contain the one that we want to find.
-
-From the latter subregion we can explore the numerical error of points 
-with $D = D_\text{linear}$ and different values of $\delta$.
+We can ignore points belonging to the former one and focus on the 
+latter.
+
+From the $D \ge D_\text{linear}$ subregion we can explore the numerical 
+error of points with $D = D_\text{linear}$ and different values of 
+$\delta$.
 We will call $\delta_p$ as the maximum $\delta$ at which the numerical 
@@
 increases the number of discretizations of the tesseroid.
-Although a greater $D$ may subdivide it in the latitude, longitude 
+Although a greater $D$ may divide it in the latitude, longitude 
 and/or radial direction; a lower $\delta$ will only discretize the 
 original tesseroid in the radial direction.
-From this analysis we can conclude that a decrease in $\delta$ will 
-produce a more efficient computation than an increase in $D$.
+This means that a decrease in $\delta$ will produce a more efficient 
+computation than an increase in $D$.
+\todo{Should we keep this paragraph? The same is explained bellow.}
 
@@
 
-On the other hand, the numerical error of points with $\delta > 
-\delta_p$ can either be lower or greater than 0.1\%.
-Any point in this subregion with acceptable numerical error will 
-produce more discretizations than the ($D_\text{linear}$, $\delta_p$) 
-point, due to $D$ value is increased.
+On the other hand, we have no a priori information about the numerical 
+error of points with $\delta > \delta_p$.
+Although, if we find a point with acceptable numerical error in this 
+region, we know it will produce more discretizations than the 
+($D_\text{linear}$, $\delta_p$) point.
+This is due to the fact that even an increase in $\delta$ will produce 
+less density-based discretizations, the increase in $D$ will introduce 
+even more subdivisions not only in the radial direction, but 
+also in the latitude and logitude ones.
 

### 2018-04-13 manuscript/manuscript.tex
Changed every todo note to inline todo note

@@
 analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013} 
-\todo{citar mas??}.
+\todo[inline]{citar mas??}.
 
@@
 maintenance easier.
-\todo{include this?}
+\todo[inline]{include this?}
 
@@
 algorithms, approximates the gravitational fields with good accuracy.
-\todo{no se si poner esta ultima oracion}
+\todo[inline]{no se si poner esta ultima oracion}
 
@@
 discontinuous function.
-\todo{sacaria todo este parrafo y lo introduciria mas adelante}
+\todo[inline]{sacaria todo este parrafo y lo introduciria mas adelante}
 
@@
 $r'$: a linear and an exponential density function.
-\todo{lo sacaria}
+\todo[inline]{lo sacaria}
 
@@
 for the $g_{zz}$, although they can be found on the repository. 
-\todo{chequear}
+\todo[inline]{chequear}
 
@@
 computation than an increase in $D$.
-\todo{Should we keep this paragraph? The same is explained bellow.}
+\todo[inline]{Should we keep this paragraph? The same is explained bellow.}
 
@@
 We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth.
-It is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo{agregar repo, vale la pena poner el link?}
+It is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo[inline]{agregar repo, vale la pena poner el link?}
 
@@
 We exclude the results corresponding to the other two diagonal components of the tensor because they are similar to the ones obtained for the $g_{zz}$, although they can be found on the repository.
-\todo{chequear}
+\todo[inline]{chequear}
 
@@
 On the other hand, the density-based discretization algorithm makes the computation slower: up to 20 times slower than the homogeneous density code.
-Nevertheless, this increase in the computation time pays off with a guarantee\todo{guarantee is a noun?} on its accuracy.
+Nevertheless, this increase in the computation time pays off with a guarantee\todo[inline]{guarantee is a noun?} on its accuracy.
 
@@
 \citet{Sigismondi2012} proposed a minimum and maximum density contrast for the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
-\todo{ver que esos valores son de contraste y tienen q ir negativos.
+\todo[inline]{ver que esos valores son de contraste y tienen q ir negativos.
 poner los valores absolutos.}

### 2018-04-13 manuscript/manuscript.tex
Added D-delta-grid-search figure: added svg, pdf and in manuscript

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]
+    {figures/D-delta-grid-search.pdf}
+\caption{
+    $D$, $\delta$ space divided in three regions (A, B and C) based on 
+    the definitions of $D_\text{linear}$ and $\delta_p$ (see text for more 
+    information).
+    Light grey points are the non-eligible ($D$, $\delta$) pairs 
+    belonging to the region A (with $D < D_\text{linear}$).
+    Green points have acceptable numerical error (lower than 0.1\%), 
+    while the red ones represent those whose error is greater than the 
+    threshold.
+    Dark grey points represent those from which we don't have a priori 
+    information about their numerical error.}
+\label{fig:D-delta-grid-search}
+\end{figure}
+
 Because the algorithm is intended to work with default $D$ and $\delta$ 

### 2018-04-13 manuscript/manuscript.tex
Changed textwidth for linewidth in manuscript
Added references to D-delta-grid-search figure in manuscript

@@
 \centering
-\includegraphics[width=\textwidth]
+\includegraphics[width=\linewidth]
     {figures/density-based-discretization-algorithm.pdf}
@@
 one with $D < D_\text{linear}$ and other with $D \ge D_\text{linear}$.
-We can ignore points belonging to the former one and focus on the 
-latter.
+We can ignore points belonging to the former one (region A on Figure 
+\ref{fig:D-delta-grid-search}) and focus on the latter (regions B and C 
+on Figure \ref{fig:D-delta-grid-search}).
 
@@
 With this definition in mind we can subdivide the subregion in two: one 
-with points with $\delta > \delta_p$ and other with points with $\delta 
-\le \delta_p$.
+with points with $\delta \le \delta_p$ (region B on Figure 
+\ref{fig:D-delta-grid-search}) and other with points with $\delta > 
+\delta_p$ (region C on Figure \ref{fig:D-delta-grid-search}).
 
@@
 
-Every point in the subregion with $\delta \le \delta_p$ will have a 
-numerical error lower than the 0.1\% threshold, but they will always 
-produce more discretizations than the ($D_\text{linear}$, $\delta_p$) 
-point leading to non-efficient computations.
+Every point in the subregion with $\delta \le \delta_p$ (region B on 
+Figure \ref{fig:D-delta-grid-search}) will have a numerical error lower 
+than the 0.1\% threshold, but they will always produce more 
+discretizations than the ($D_\text{linear}$, $\delta_p$) point leading 
+to non-efficient computations.
 
 On the other hand, we have no a priori information about the numerical 
-error of points with $\delta > \delta_p$.
+error of points with $\delta > \delta_p$ (region C on 
+Figure \ref{fig:D-delta-grid-search}).
 Although, if we find a point with acceptable numerical error in this 
 region, we know it will produce more discretizations than the 
-($D_\text{linear}$, $\delta_p$) point.
+($D_\text{linear}$, $\delta_p$) point, leading to a non-efficient 
+computation.
 This is due to the fact that even an increase in $\delta$ will produce 
@@
 even more subdivisions not only in the radial direction, but 
-also in the latitude and logitude ones.
+also in the latitude and longitude ones.
+
+In conclusion, the ($D$, $\delta$) point that minimises the number of 
+discretizations while keeping an acceptable numerical error under the 
+0.1\% threshold is the one on which the ratio $D$ is equal to 
+$D_\text{linear}$ and the delta ratio is $\delta_p$.
 

### 2018-04-13 manuscript/manuscript.tex
Fixed typos

@@
 significantly reduce the numerical error.
-A big subdiving tesseroid with small maximum absolute difference would 
+A big subdividing tesseroid with small maximum absolute difference would 
 introduce more numerical error than a small subdividing tesseroid with 
@@
 represent the discretization heights, i.e. the radial boundaries of the 
-subdiving tesseroids produced by the algorithm.
+subdividing tesseroids produced by the algorithm.
 
@@
 On Figure \ref{fig:density-discretization-algorithm}(c) we perform the 
-same analysis on the deeper subdiving tesseroid, and assuming that the 
+same analysis on the deeper subdividing tesseroid, and assuming that the 
 inequality holds in this case we also subdivide it.
@@
 Finally, if we suppose that the value of $\delta$ used in this example 
-does not produce further discretizations, the resulting subdiving 
+does not produce further discretizations, the resulting subdividing 
 tesseroids can be seen on Figure 
@@
 Lets consider a spherical shell with inner and outer radii $R_1$ and 
-$R_2$, respectivetly, whose density depends on the radial spherical 
+$R_2$, respectively, whose density depends on the radial spherical 
 coordinate (Figure \ref{fig:spherical-shell}).
@@
 This expression will allow us  to perform comparisons between the 
-anaytical solutions and the numerical approximation, similar as the 
+analytical solutions and the numerical approximation, similar as the 
 ones made for the linear density case.

### 2018-04-17 manuscript/manuscript.tex
Added single figure for linear density results. Improved density-based discret figure.

@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
+\includegraphics[width=0.9\linewidth]{figures/linear-D.pdf}
 \caption{
     Differences between the gravity fields generated by the numerical 
-    model and the analytical solution for a thin spherical shell of 1km 
-    of thickness with the linear density defined on equation 
+    model and the analytical solution for (a) a thin spherical shell of 
+    1km of thickness and (b) a thick spherical shell of 35km of thickness,
+    both with the linear density defined on equation 
     \ref{eq:density-linear}.
@@
     }
-\label{fig:D-linear-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical 
-    model and the analytical solution for a thick spherical shell of 
-    35km of thickness with the linear density defined on equation 
-    \ref{eq:density-linear}.
-    The computations were performed on the four grids described in 
-    Table \ref{tab:grids}, for different values of the distance-size 
-    ratio $D$ and without applying the density-based discretization 
-    algorithm.
-    If the difference is less than 0.1\%, we consider that the model 
-    has achieved an acceptable accuracy for the corresponding value of 
-    $D$.
-    }
-\label{fig:D-linear-thick}
-\end{figure}
+\label{fig:D-linear}
+\end{figure*}
 
@@
 
+\todo[inline, color=red]{Text bellow is old}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-04-17 manuscript/manuscript.tex
Removed unnecesary equation from manuscript
Modified exponential-delta.pdf figure and added in the manuscript

@@
 \end{equation}
-\begin{equation}
-    \text{max}\{ \Delta \rho(r') \} = \Delta \rho(R_\text{max}),
-\end{equation}
 
@@
     Table \ref{tab:grids}, for different values of the distance-size 
-    ratio $D$ and without applying the density-based discretization 
-    algorithm.
+    ratio $D$.
+    Due to the linearity of the density function the density-based 
+    discretization algorithm is not applied.
     If the difference is less than 0.1\%, we consider that the model 
@@
 
+Taking this conclusion into account we have determined the $\delta$ 
+value needed to ensure an acceptable accuracy through a comparison 
+between the numerical approximation and the analytical solution for a 
+spherical shell with an exponential density.
+In this specific case we proposed an exponential density function  that 
+assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's 
+outer and inner radii, respectively. It can be defined as follows:
+
+\begin{equation}
+    \rho(r') = A e^{-(r' - R)/b} + C,
+\label{eq:density-exp}
+\end{equation}
+\noindent where
+\begin{equation}
+    A =
+    (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
+    \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
+\end{equation}
+\begin{equation}
+    C =
+    2670 \text{kg/m}^3 - A,
+\end{equation}
+
+\noindent where $R$ is the mean Earth radius and $b$ is a constant 
+that determines the variability of the function: a low value of $b$ 
+increases the maximum slope of the density function.
+
+In order to perform this comparison we must compute the analytical 
+solution of the gravitational fields generated by this spherical shell 
+and define a tesseroid model to numerically approximate this fields.
+
+Firstly, the analytical solution can be computed as the sum of the 
+fields generated by a spherical shell with constant density $\rho(r') = 
+C$ and the one determined in equation \ref{eq:shell-pot-exp}.
+
+Secondly, we proposed a thin and a thick spherical shells with 1km and 
+35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
+30^\circ$ tesseroids.
+We have assigned the exponential density function proposed on equation 
+\ref{eq:density-exp} to each spherical shell model, choosing few values 
+of the constant $b$ in order to test the numerical approximation for 
+different density variations.
+
+Finally, we computed the gravitational potential ($V$), the 
+vertical component of its gradient ($g_z$) and the $g_zz$ component of 
+the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
+keeping the distance-size ratio equal to the $D$ values obtained in the 
+density case for its corresponding gravitational field and exploring 
+different values of $\delta$.
+We then computed the percentage difference between this numerical 
+results and the analytical solutions.
+
+On Figures \ref{fig:delta-exponential}(a) and 
+\ref{fig:delta-exponential}(b) we show the density functions used in 
+the comparison and the resulting percentage difference for each 
+gravitational field.
+The former shows the results for the thin spherical shell, while the 
+later shows the ones corresponding to the thick shell.
+From these results we can conclude that for the potential and $g_z$ 
+computation, a $\delta = 0.3$ is needed in order to achieve an 
+acceptable numerical error, while a $\delta = 0.02$ is required for the 
+$g_{zz}$ Marussi tensor component.
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
+\caption{
+    Density functions used in the model comparison and the differences 
+    between the gravitational fields generated by the numerical model 
+    and the analytical solution for (a) a thin spherical shell of 1km 
+    of thickness and (b) a thick spherical shell of 35km of thickness, 
+    both with the exponential density defined on equation 
+    \ref{eq:density-exp}.
+    A few values of the constant $b$ were used in order to test the 
+    numerical model for different density variations.
+    The computations were performed on the big grid described in 
+    Table \ref{tab:grids}, with fixed values of the distance-size ratio 
+    $D$ obtained for in the linear density case ($D=1, 2, 8$ for the 
+    potential, gradient components and tensor components, respectively)
+    and different values of the $\delta$ ratio.
+    If the difference is less than 0.1\%, we consider that the model 
+    has achieved an acceptable accuracy for the corresponding value of 
+    $D$.
+    }
+\label{fig:delta-exponential}
+\end{figure*}
+
 \todo[inline, color=red]{Text bellow is old}

### 2018-04-17 manuscript/manuscript.tex
Commented out old text in manuscript

@@
 
-\subsubsection{Linear Density}
-
-If the density of the spherical shell is a linear function on the radial coordinate, i.e.:
-
-\begin{equation}
-    \rho(r') = ar' + b,
-\end{equation}
-
-\noindent then the gravity potential generated on any computation point at distance $r$ from its centre can be obtained from the following expression:
-
-\begin{equation}
-    V_\text{sh}^\text{lin}(r) = \pi G \left[ 
-    a \frac{R_2^4 - R_1^4}{r} +
-    b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
-    \label{eq:shell-pot-linear}
-\end{equation}
-
-The first term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with variable density $\rho(r') = ar'$, while the second one constitutes the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
-So the potential can be read as the one generated by a combination of both shells.
-
-\subsubsection{Exponential Density}
-
-If, on the other hand, the spherical shell has an exponential density in the radial coordinate,
-
-\begin{equation}
-    \rho(r') = Ae^{-(r' - \Delta h)/b},
-\end{equation}
-
-\noindent the potential (calculated through SymPy \citep{sympy2017}) can be written as follows:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-        Ab 
-        \Big[
-        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
-        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
-        \Big].
-    \end{split}
-\label{eq:shell-pot-exp}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-    Ab \, e^\frac{\Delta h}{b}
-    \Big[
-    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    \Big].
-\label{eq:shell-pot-exp}
-\end{equation}
-}
+%~ \subsubsection{Linear Density}
+
+%~ If the density of the spherical shell is a linear function on the radial coordinate, i.e.:
+
+%~ \begin{equation}
+    %~ \rho(r') = ar' + b,
+%~ \end{equation}
+
+%~ \noindent then the gravity potential generated on any computation point at distance $r$ from its centre can be obtained from the following expression:
+
+%~ \begin{equation}
+    %~ V_\text{sh}^\text{lin}(r) = \pi G \left[ 
+    %~ a \frac{R_2^4 - R_1^4}{r} +
+    %~ b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
+    %~ \label{eq:shell-pot-linear}
+%~ \end{equation}
+
+%~ The first term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with variable density $\rho(r') = ar'$, while the second one constitutes the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
+%~ So the potential can be read as the one generated by a combination of both shells.
+
+%~ \subsubsection{Exponential Density}
+
+%~ If, on the other hand, the spherical shell has an exponential density in the radial coordinate,
+
+%~ \begin{equation}
+    %~ \rho(r') = Ae^{-(r' - \Delta h)/b},
+%~ \end{equation}
+
+%~ \noindent the potential (calculated through SymPy \citep{sympy2017}) can be written as follows:
+
+%~ \iftwocol{
+%~ \begin{equation}
+    %~ \begin{split}
+        %~ V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+        %~ Ab 
+        %~ \Big[
+        %~ & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
+        %~ & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
+        %~ \Big].
+    %~ \end{split}
+%~ \label{eq:shell-pot-exp}
+%~ \end{equation}
+%~ }{
+%~ \begin{equation}
+    %~ V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+    %~ Ab \, e^\frac{\Delta h}{b}
+    %~ \Big[
+    %~ (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+    %~ (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+    %~ \Big].
+%~ \label{eq:shell-pot-exp}
+%~ \end{equation}
+%~ }
 
@@
 
-\section{Determination of distance-size ratio}
-
-In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
-The first one constitutes the simplest variation, the second one represents a higher varying function, while the later one tries to emulate a contact between two different media (e.g.
+%~ \section{Determination of distance-size ratio}
+
+%~ In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
+%~ The first one constitutes the simplest variation, the second one represents a higher varying function, while the later one tries to emulate a contact between two different media (e.g.
 crust and mantle interface).
 
-We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
-More information about this grids can be found on Table \ref{tab:grids}.
-
-We computed these gravity fields on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, without applying the density-based discretization algorithm in order to isolate the behaviour of the adaptive discretization one.
-Then we calculated the maximum difference between these results and the value that the analytical solution assumes at the same height.
-
-Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
-That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
-
-Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
-
-\begin{table}
-\caption{
-    Description of the synthetic grids on which the comparison of the numerical model against the analytical solutions for a spherical shell was done.
+%~ We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
+%~ More information about this grids can be found on Table \ref{tab:grids}.
+
+%~ We computed these gravity fields on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, without applying the density-based discretization algorithm in order to isolate the behaviour of the adaptive discretization one.
+%~ Then we calculated the maximum difference between these results and the value that the analytical solution assumes at the same height.
+
+%~ Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
+%~ That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
+
+%~ Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
+
+%~ \begin{table}
+%~ \caption{
+    %~ Description of the synthetic grids on which the comparison of the numerical model against the analytical solutions for a spherical shell was done.
 This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height over the Earth surface.
 Every grid consists in a set of 10$\times$10 points.
-}
-\label{tab:grids}
-\begin{tabular}{lcccc}
-    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
-    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
-    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
-\end{tabular}
-\end{table}
-
-\subsection{Linear Density}
-Firstly,  we proposed a spherical shell with linear density in such way it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:
-
-\begin{equation}
-    \rho(r') = ar' + c,
-    \label{eq:density-linear}
-\end{equation}
-\noindent with 
-\begin{equation}
-    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R_2 - R_1},
-\end{equation}
-\begin{equation}
-    c = \frac{3300\text{kg/m$^3$} - 
-        2670\text{kg/m$^3$}}{R_2 - R_1} R + 
-        2670\text{kg/m$^3$},
-\end{equation}
-
-\noindent where $R$ = 6378.137 km is the mean Earth radius.
-
-Then we computed the difference between the numerical model with its corresponding analytical solution (equation \ref{eq:density-linear}) for a thick and a thin spherical shell of 1km and 35km of thickness, respectively.
-
-We can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
-We exclude the results corresponding to the other two diagonal components of the tensor because they are similar to the ones obtained for the $g_{zz}$, although they can be found on the repository.
+%~ }
+%~ \label{tab:grids}
+%~ \begin{tabular}{lcccc}
+    %~ Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
+    %~ Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    %~ Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
+    %~ Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
+    %~ Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
+%~ \end{tabular}
+%~ \end{table}
+
+%~ \subsection{Linear Density}
+%~ Firstly,  we proposed a spherical shell with linear density in such way it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:
+
+%~ \begin{equation}
+    %~ \rho(r') = ar' + c,
+    %~ \label{eq:density-linear}
+%~ \end{equation}
+%~ \noindent with 
+%~ \begin{equation}
+    %~ a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R_2 - R_1},
+%~ \end{equation}
+%~ \begin{equation}
+    %~ c = \frac{3300\text{kg/m$^3$} - 
+        %~ 2670\text{kg/m$^3$}}{R_2 - R_1} R + 
+        %~ 2670\text{kg/m$^3$},
+%~ \end{equation}
+
+%~ \noindent where $R$ = 6378.137 km is the mean Earth radius.
+
+%~ Then we computed the difference between the numerical model with its corresponding analytical solution (equation \ref{eq:density-linear}) for a thick and a thin spherical shell of 1km and 35km of thickness, respectively.
+
+%~ We can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
+%~ We exclude the results corresponding to the other two diagonal components of the tensor because they are similar to the ones obtained for the $g_{zz}$, although they can be found on the repository.
 \todo[inline]{chequear}
 
-For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
-On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D$ equal to 8 while the later only a $D$ of 3.5.
-We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}.
+%~ For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
+%~ On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D$ equal to 8 while the later only a $D$ of 3.5.
+%~ We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
+%~ \caption{
+    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}.
 The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-linear-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}.
+%~ }
+%~ \label{fig:D-linear-thin}
+%~ \end{figure}
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
+%~ \caption{
+    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}.
 The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-linear-thick}
-\end{figure}
-
-\subsection{Exponential Density}
-
-To test the model in case of an exponential density we proposed the following function:
-
-\begin{equation}
-    \rho(r') = A e^{-(r' - R)/b} + C,
-\label{eq:density-exp}
-\end{equation}
-\noindent where
-\begin{equation}
-    A =
-    (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
-    \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
-\end{equation}
-\begin{equation}
-    C =
-    2670 \text{kg/m}^3 - A,
-\end{equation}
-
-\noindent and the value of $b$ equal to the thickness of the spherical shell we will use to perform the test.
-In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
-The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function (like the one in equation \ref{eq:density-exp}) and for an homogeneous spherical shell.
-
-We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, without applying the density-based discretization algorithm.
-On both cases, the constant $b$ is equal to the thickness of the shell: 1km for the thin shell and 35km for the thick one.
-The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
-Once again, we exclude the ones corresponding to $g_{xx}$ and $g_{yy}$ due to their similarity with the ones obtained for $g_{zz}$, although they can be found in the repository.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/exponential-D-thin.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
+%~ }
+%~ \label{fig:D-linear-thick}
+%~ \end{figure}
+
+%~ \subsection{Exponential Density}
+
+%~ To test the model in case of an exponential density we proposed the following function:
+
+%~ \begin{equation}
+    %~ \rho(r') = A e^{-(r' - R)/b} + C,
+%~ \label{eq:density-exp}
+%~ \end{equation}
+%~ \noindent where
+%~ \begin{equation}
+    %~ A =
+    %~ (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
+    %~ \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
+%~ \end{equation}
+%~ \begin{equation}
+    %~ C =
+    %~ 2670 \text{kg/m}^3 - A,
+%~ \end{equation}
+
+%~ \noindent and the value of $b$ equal to the thickness of the spherical shell we will use to perform the test.
+%~ In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
+%~ The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function (like the one in equation \ref{eq:density-exp}) and for an homogeneous spherical shell.
+
+%~ We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, without applying the density-based discretization algorithm.
+%~ On both cases, the constant $b$ is equal to the thickness of the shell: 1km for the thin shell and 35km for the thick one.
+%~ The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
+%~ Once again, we exclude the ones corresponding to $g_{xx}$ and $g_{yy}$ due to their similarity with the ones obtained for $g_{zz}$, although they can be found in the repository.
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/exponential-D-thin.pdf}
+%~ \caption{
+    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
 The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-exp-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/exponential-D-thick.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
+%~ }
+%~ \label{fig:D-exp-thin}
+%~ \end{figure}
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/exponential-D-thick.pdf}
+%~ \caption{
+    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
 The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-exp-thick}
-\end{figure}
-
-In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
-
-Moreover, we can see that the differences for the potential and the $g_z$ component of the gradient keep constant even for high values of $D$.
-Our interpretation is that the modified adaptive discretization algorithm is limited when increasing the accuracy for this density function, supporting the need of a density-based discretization algorithm.
-
-Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$ when the density-based discretization algorithm is not applied.
-In order to do this we proposed a thin spherical shell (1km of thickness) with a set of density functions, each one with a different value of $b$.
-We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the corresponding analytical solution.
-These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.
-
-On Figure \ref{fig:D-exp-power-thin}a we can see the density functions used to perform these tests and its corresponding results on \ref{fig:D-exp-power-thin}b .
-Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b=100$m the difference is almost 100\%.
-
-We can conclude that even when the density-based discretization algorithm is not applied, the numerical approximation performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, i.e.
+%~ }
+%~ \label{fig:D-exp-thick}
+%~ \end{figure}
+
+%~ In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
+
+%~ Moreover, we can see that the differences for the potential and the $g_z$ component of the gradient keep constant even for high values of $D$.
+%~ Our interpretation is that the modified adaptive discretization algorithm is limited when increasing the accuracy for this density function, supporting the need of a density-based discretization algorithm.
+
+%~ Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$ when the density-based discretization algorithm is not applied.
+%~ In order to do this we proposed a thin spherical shell (1km of thickness) with a set of density functions, each one with a different value of $b$.
+%~ We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the corresponding analytical solution.
+%~ These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.
+
+%~ On Figure \ref{fig:D-exp-power-thin}a we can see the density functions used to perform these tests and its corresponding results on \ref{fig:D-exp-power-thin}b .
+%~ Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b=100$m the difference is almost 100\%.
+
+%~ We can conclude that even when the density-based discretization algorithm is not applied, the numerical approximation performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, i.e.
 for low density variations.
-On the other hand, a density-based discretization algorithm is needed in order to guarantee the accuracy of the approximation in case of a high varying density function.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/exponential-b-thin.pdf}
-\caption{
-    Results of the distance-size ratio determination for different exponential density functions in case of a spherical shell of 1km of thickness.
-    (a) Density functions used in the tests, each one with a different value of $b$.
-    (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
+%~ On the other hand, a density-based discretization algorithm is needed in order to guarantee the accuracy of the approximation in case of a high varying density function.
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/exponential-b-thin.pdf}
+%~ \caption{
+    %~ Results of the distance-size ratio determination for different exponential density functions in case of a spherical shell of 1km of thickness.
+    %~ (a) Density functions used in the tests, each one with a different value of $b$.
+    %~ (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
 Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy.
-}
-\label{fig:D-exp-power-thin}
-\end{figure}
-
-\subsection{Discontinuous Density}
-
-As a final case, we performed the same test with a discontinuous density function:
-
-\begin{equation}
-    \rho(r') = 
-    \begin{cases}
-        \rho_A & \text{if} \, R_1 \leq r' < R_c \\
-        \rho_B & \text{if} \, R_c \leq r' \leq R_2 \\
-    \end{cases}
-\label{eq:density-discontinuous}
-\end{equation}
-
-\noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the radius on which the density change takes place.
-
-The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_A$ and $\rho_B$:
-
-\begin{equation}
-    V(r) = \frac{4}{3} \frac{\pi G}{r}
-    \left[ \rho_A (R_c^3 - R_1^3) +
-    \rho_B (R_2^3 - R_c^3) \right]
-\end{equation}
-
-Lets assume that $R_c = (R_1 + R_2)/3$ and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
-Its results can be seen in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/discontinuous-D-asym.pdf}
-\caption{
-    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous}.
+%~ }
+%~ \label{fig:D-exp-power-thin}
+%~ \end{figure}
+
+%~ \subsection{Discontinuous Density}
+
+%~ As a final case, we performed the same test with a discontinuous density function:
+
+%~ \begin{equation}
+    %~ \rho(r') = 
+    %~ \begin{cases}
+        %~ \rho_A & \text{if} \, R_1 \leq r' < R_c \\
+        %~ \rho_B & \text{if} \, R_c \leq r' \leq R_2 \\
+    %~ \end{cases}
+%~ \label{eq:density-discontinuous}
+%~ \end{equation}
+
+%~ \noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the radius on which the density change takes place.
+
+%~ The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_A$ and $\rho_B$:
+
+%~ \begin{equation}
+    %~ V(r) = \frac{4}{3} \frac{\pi G}{r}
+    %~ \left[ \rho_A (R_c^3 - R_1^3) +
+    %~ \rho_B (R_2^3 - R_c^3) \right]
+%~ \end{equation}
+
+%~ Lets assume that $R_c = (R_1 + R_2)/3$ and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
+%~ Its results can be seen in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/discontinuous-D-asym.pdf}
+%~ \caption{
+    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous}.
 The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
 If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-}
-\label{fig:D-discont-asym}
-\end{figure}
+%~ }
+%~ \label{fig:D-discont-asym}
+%~ \end{figure}
 
@@
 
-\section{Determination of delta ratio}
-
-In the light of the previous results, a density-based discretization algorithm is needed in order to achieve an acceptable accuracy in those cases on which the modified adaptive discretization algorithm is not sufficient to take into account the numerical error due to density variations.
-
-Besides the implementation of this new algorithm, we must determine the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold in order to guarantee the good performance of the numerical model.
-
-To do this we applied a similar methodology used in the determination of the distance-size ratio $D$.
-Using different values of $\delta$, ranging from 0.1 to 1 with a step of 0.1, we computed the potential, the $g_z$ component of the gradient and the $g_{zz}$ component of the Marussi tensor generated by a spherical shell with several exponential densities (like the one in equation \ref{eq:density-exp}): each one with a different value of $b$.
-We performed this tests for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
-
-Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a show the exponential density functions used on each test for a thin and thick spherical shell, respectively.
-While on Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b we can see the differences between the resulting $g_{zz}$ with the corresponding analytical solution.
-The results for the other gravity fields were excluded from the figures due to their similarities with the ones for $g_{zz}$.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
-\caption{
-    Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
-    (a) Density functions used to perform the test.
-    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
-\label{fig:exponential-delta-thin}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
-\caption{
-    Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
-    (a) Density functions used to perform the test.
-    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
-\label{fig:exponential-delta-thick}
-\end{figure}
-
-In the thin shell case we can conclude that a delta equal to 0.3 guarantees an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
-We will keep the more conservative value of 0.2 as the default value of delta on future computations.
-
-Furthermore, we computed the difference between the numerical model and the analytical solution in case of a discontinuous density (as the one in equation \ref{eq:density-discontinuous} with $R_c = (R_1 + R_2)/3$) using a $\delta = 0.2$.
-The result for the $g_{zz}$ component of the Marussi tensor is bellow the 0.1\% threshold, achieving a good performance for this default value of $\delta$.
+%~ \section{Determination of delta ratio}
+
+%~ In the light of the previous results, a density-based discretization algorithm is needed in order to achieve an acceptable accuracy in those cases on which the modified adaptive discretization algorithm is not sufficient to take into account the numerical error due to density variations.
+
+%~ Besides the implementation of this new algorithm, we must determine the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold in order to guarantee the good performance of the numerical model.
+
+%~ To do this we applied a similar methodology used in the determination of the distance-size ratio $D$.
+%~ Using different values of $\delta$, ranging from 0.1 to 1 with a step of 0.1, we computed the potential, the $g_z$ component of the gradient and the $g_{zz}$ component of the Marussi tensor generated by a spherical shell with several exponential densities (like the one in equation \ref{eq:density-exp}): each one with a different value of $b$.
+%~ We performed this tests for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
+
+%~ Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a show the exponential density functions used on each test for a thin and thick spherical shell, respectively.
+%~ While on Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b we can see the differences between the resulting $g_{zz}$ with the corresponding analytical solution.
+%~ The results for the other gravity fields were excluded from the figures due to their similarities with the ones for $g_{zz}$.
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
+%~ \caption{
+    %~ Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
+    %~ (a) Density functions used to perform the test.
+    %~ (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
+%~ \label{fig:exponential-delta-thin}
+%~ \end{figure}
+
+%~ \begin{figure}
+%~ \centering
+%~ \includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
+%~ \caption{
+    %~ Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
+    %~ (a) Density functions used to perform the test.
+    %~ (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
+%~ \label{fig:exponential-delta-thick}
+%~ \end{figure}
+
+%~ In the thin shell case we can conclude that a delta equal to 0.3 guarantees an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
+%~ We will keep the more conservative value of 0.2 as the default value of delta on future computations.
+
+%~ Furthermore, we computed the difference between the numerical model and the analytical solution in case of a discontinuous density (as the one in equation \ref{eq:density-discontinuous} with $R_c = (R_1 + R_2)/3$) using a $\delta = 0.2$.
+%~ The result for the $g_{zz}$ component of the Marussi tensor is bellow the 0.1\% threshold, achieving a good performance for this default value of $\delta$.
 

### 2018-04-18 manuscript/manuscript.tex
Improved density-based discretization algorithm description in manuscript

@@
 
-High variations in the density function of an arbitrary tesseroid 
-introduces a new type of error into the numerical approximation: only a 
-few point masses would not reproduce the effect of this kind of 
-variations.
+Variations in the density function of an arbitrary tesseroid introduces 
+a new type of error into the numerical approximation: only a few point 
+masses would not reproduce the effect of the density variations.
 The modified version of the adaptive discretization algorithm may help 
@@
 direction on the depth at which the \emph{maximum density variation} 
-takes place.
+takes place, only if a certain discretization inequation holds.
 
@@
 function will be the same for every subdividing tesseroid.
+
 In case that the density function is constant, both the maximum and 
@@
 
-Finally, if the product of the maximum of the absolute difference and a 
-tesseroid size ratio is greater than a predefined delta ratio $\delta$ 
-we will divide the tesseroid on the radius at which this maximum takes 
-place ($R_\text{max}$):
+Finally, if the following inequation holds, the tesseroid will be divided on 
+the radius $R_\text{max}$ at which the maximum absolute difference takes 
+place:
 
@@
 
-\noindent obtaining a pair of smaller tesseroids with the same 
-longitude and latitude boundaries.
-
-The $L_r^\text{sm}$ is the radial dimension of the tesseroid that is 
-currently analysed. On the first iteration, this tesseroid is the 
+\noindent where the delta ratio $\delta$ is a predefined constant value 
+and $L_r^\text{sm}/L_r$ is a tesseroid size ratio, where 
+$L_r^\text{sm}$ and $L_r$ are the radial dimensions of the tesseroid 
+that is currently analysed and of the \emph{original} one, 
+respectively.
+On the first iteration, the former tesseroid corresponds to the 
 \emph{original} one, so $L_r^\text{sm} = L_r$, but for future 
@@
 the top and bottom radii of the subdividing tesseroid.
-
-This algorithm will be subsequently applied to every subdividing 
-tesseroid.
+The tesseroid size ratio is intended to reduce the number of probable 
+unneeded discretizations and focus on discretizations that will 
+significantly reduce the numerical error.
+A big subdividing tesseroid with small maximum absolute difference would 
+introduce more numerical error than a small subdividing tesseroid with 
+higher maximum absolute difference.
+
+Once the \emph{original} tesseroid is divided, we proceed to 
+subsequently apply the same algorithm on every subdividing tesseroid.
 We will start from equation \ref{eq:density-reference-line} defining a 
@@
 are the radial boundaries of this smaller tesseroid.
-
-The tesseroid size ratio included in the inequation 
-\ref{eq:delta-density} is intended to reduce the number of probable 
-unneeded discretizations and focus on discretizations that will 
-significantly reduce the numerical error.
-A big subdividing tesseroid with small maximum absolute difference would 
-introduce more numerical error than a small subdividing tesseroid with 
-higher maximum absolute difference.
+If the inequation holds, the current tesseroid is discretized and the 
+algorithm is then applied to the resulting subdividing tesseroids.
 
@@
 This two conditions guarantee that the algorithm is finite.
-Once they are met, every subdividing tesseroid is subjected to the 
-modified adaptive discretization algorithm and the GLQ in order to 
-compute the gravitational field generated by it.
+Once any of them is met and no more discretizations are carried out, 
+every subdividing tesseroid is subjected to the modified adaptive 
+discretization algorithm and then a second order GLQ must be applied in 
+order to compute the gravitational field generated by it.
 
@@
 algorithm works.
-And we plotted its corresponding linear function $\rho_l(r')$ and a 
-dashed line that represents the maximum absolute difference $\Delta 
-\rho (r')$.
+We plotted its corresponding linear function $\rho_l(r')$ and a 
+dashed line that represents the maximum absolute difference.
 We will assume that the discretization inequation 
 \ref{eq:delta-density} holds in this case, so we must divide the 
-tesseroid at the height at which the maximum difference occurs.
+tesseroid at the depth at which the maximum difference occurs.
 The dots on Figure \ref{fig:density-discretization-algorithm} 
-represent the discretization heights, i.e. the radial boundaries of the 
+represent the discretization points, i.e. the radial boundaries of the 
 subdividing tesseroids produced by the algorithm.
@@
 algorithm on the shallower subdividing tesseroid.
-In the same way, we propose the linear function and compute the maximum 
+In the same way, we plot the linear function and compute the maximum 
 absolute difference $\Delta \rho (r')$.

### 2018-04-18 manuscript/manuscript.tex
Improved linear density text in manuscript

@@
 On the homogeneous density tesseroids case, \citet{Uieda2016} and 
-\citet{Grombein2013} compared their numerical model with the analytical 
+\citet{Grombein2013} compared their numerical models with the analytical 
 solution of a spherical shell.
 We could follow this idea, but for our needs the spherical shell must 
-have the same density function as the numerical model, although the 
-analytical solution in this case must be derived.
+have the same density function as the numerical model.
+In contrast with the homogeneous density case, the analytical solutions 
+for a variable density spherical shell must be obtained.
+
+We will then perform these comparisons between the analytical solutions 
+and the numerical model for specific density functions: a linear and an 
+exponential one.
+And from these results we will finally obtain default values for $D$ 
+and $\delta$ that ensure a numerical error lower than 0.1\%.
 
@@
 The integrals that define the gravity fields in equations 
-\ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have analytical 
-solutions in case of a spherical shell with homogeneous density 
-\citep{Mikuska2006,Grombein2013}.
-But for variable density spherical shell the problem wasn't address in 
-the literature yet.
+\ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have almost trivial 
+analytical solutions in case of a spherical shell with homogeneous 
+density \citep{Mikuska2006,Grombein2013}.
+But for a variable density spherical shell the problem haven't been 
+addressed in the literature yet.
 
@@
 
-Lets consider a linear density function like:
+Lets perform a comparison between our numerical model and the 
+analytical solution for a spherical shell with a linear density 
+function like the following one:
 
@@
 
-\noindent The gravitational potential generated by a spherical shell 
-with this density on any external point at distance $r$ from its centre 
-can be obtained from the following expression:
+Firstly we need to solve the integral on equation \ref{eq:shell-pot} in 
+order to obtain the analytical solution in this case.
+Thus, the gravitational potential generated by a spherical shell with this 
+density on any external point at distance $r$ from its centre can be 
+obtained from the following expression:
 
@@
 potential generated by a spherical shell with homogeneous density $\rho 
-= b$ \citep{Mikuska2006,Grombein2013}. So the potential can be read as 
-the one generated by a combination of both shells.
-
+= b$ \citep{Mikuska2006,Grombein2013}.
+So the potential can be read as the one generated by a combination of 
+both shells.
+
+Secondly we can analyse how the numerical algorithm works for any 
+tesseroid with linear density.
 When applying the density-based discretization algorithm on any 
-tesseroid with linear density, the absolute difference defined in 
+tesseroid with this kind of density, the absolute difference defined in 
 equation \ref{eq:density-abs-diff} will always be zero, so the 
@@
 Then we calculated the maximum difference between these results and the 
-value that the analytical solution assumes at the same height.
+value that the analytical solution assumes at the same height of the 
+computation grid.
 Because the gravity fields generated by a tesseroid are approximated by 
@@
 spherical shells with 1km and 35km of thickness, respectively, whose 
-outer radii is equal to the mean Earth radius.
+outer radii are equal to the mean Earth radius.
 On both cases we will suppose that the density assumes a value of 
@@
 with its corresponding analytical solution (equation 
-\ref{eq:density-linear}) for the thick and the thin spherical shell.
+\ref{eq:shell-pot-linear}) for the thick and the thin spherical shell.
 
@@
 gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on 
-Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
+Figure \ref{fig:D-linear}.
 We exclude the results corresponding to the other two diagonal 
@@
 We will keep the more conservative result of $D$ equal to 8 in case of 
-computing the Marussi tensor components in case of a linear density 
-tesseroid.
+computing the Marussi tensor components.
 

### 2018-04-18 manuscript/manuscript.tex
Improved exponential density in manuscript

@@
 Therefore it determines the accuracy involving the numerical error that 
-the density variation introduces and also the computation time.
+the density variation introduces and also the computation time: a lower 
+value of $\delta$ will produce more discretizations leading to an 
+increase in the accuracy of the approximation but also in the 
+computation time.
 This raises the need to determine a maximum value of $\delta$ that 
@@
 
-A more complex density variation than a linear one could be an 
-exponential function like:
-
-\begin{equation}
-    \rho(r') = Ae^{-(r' - \Delta h)/b},
-\end{equation}
-
-\noindent where $A$, $\Delta h$ and $b$ are constants.
-The gravitational potential field generated by a spherical shell with 
-this kind of density function on any external point at a distance $r$ 
-from its centre can be calculated as follows:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-        Ab 
-        \Big[
-        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
-        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
-        \Big].
-    \end{split}
-\label{eq:shell-pot-exp}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-    Ab \, e^\frac{\Delta h}{b}
-    \Big[
-    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    \Big].
-\label{eq:shell-pot-exp}
-\end{equation}
-}
-
-This expression will allow us  to perform comparisons between the 
-analytical solutions and the numerical approximation, similar as the 
-ones made for the linear density case.
-Although in the exponential density scenario the problem is more 
-complex due to the fact that the density-based discretization must be 
-used together with the modified adaptive discretization algorithm.
+Lets perform a similar comparison between the numerical model and 
+the analytical solution but now for a spherical shell with exponential 
+density.
+On this scenario the problem is more complex due to the fact that the 
+density-based discretization will be certainly applied together with 
+the modified adaptive discretization algorithm.
 This means that both the distance-size ratio $D$ and the delta ratio 
@@
 
-From every $D$ and $\delta$ combination we want to find the one that 
-produces a numerical error lower than the 0.1\% threshold while keeping 
-an effective computation, i.e. the ($D$, $\delta)$ pair that produces 
-the minimum number of discretizations and ensures an acceptable 
-numerical precision.
+From every combination of the $D$ and $\delta$ ratios we want to find 
+the one that produces a numerical error lower than the 0.1\% threshold 
+while keeping an effective computation, i.e. the ($D$, $\delta)$ pair 
+that produces the minimum number of discretizations and ensures an 
+acceptable numerical precision.
 We can propose an heuristic method to perform this task: compute the 
 numerical error for every ($D$, $\delta$) points belonging to a grid on 
-the $D$, $\delta$ space and specify the pair that combines an 
+the $D$, $\delta$ space and choose the pair that combines an 
 acceptable numerical error with an efficient computation.
@@
 the default value for $D$ cannot be lower than the one determined for 
-the linear function.
+the linear function, which will be called $D_\text{linear}$.
 That's why we can only choose the ($D$, $\delta$) points on which
 $D \geq D_\text{linear}$.
-In this way we can divide the $D$, $\delta$ space in two subregions: 
-one with $D < D_\text{linear}$ and other with $D \ge D_\text{linear}$.
-We can ignore points belonging to the former one (region A on Figure 
-\ref{fig:D-delta-grid-search}) and focus on the latter (regions B and C 
-on Figure \ref{fig:D-delta-grid-search}).
+Taking this fact into account we can divide the $D$, $\delta$ space in 
+two subregions: one with $D < D_\text{linear}$ and other with $D \ge 
+D_\text{linear}$. We can ignore points belonging to the former one 
+(region A on Figure \ref{fig:D-delta-grid-search}) and focus on the 
+latter (regions B and C on Figure \ref{fig:D-delta-grid-search}).
 
@@
 $\delta$.
-We will call $\delta_p$ as the maximum $\delta$ at which the numerical 
+We will call $\delta_p$ to the maximum $\delta$ at which the numerical 
 error is lower than the 0.1\% threshold while keeping
 $D = D_ \text{linear}$.
-Greater $\delta$ values than $\delta_p$ will produce computations 
-with high error, while lower $\delta$ values than $\delta_p$ will 
-produce more discretizations leading to non-efficient computation.
-
-With this definition in mind we can subdivide the subregion in two: one 
-with points with $\delta \le \delta_p$ (region B on Figure 
-\ref{fig:D-delta-grid-search}) and other with points with $\delta > 
-\delta_p$ (region C on Figure \ref{fig:D-delta-grid-search}).
-
-It worth mention that a greater $D$ or a lower $\delta$ 
+Greater $\delta$ values than $\delta_p$ will produce computations with 
+high error (red points in Figure \ref{fig:D-delta-grid-search}), while 
+lower $\delta$ values than $\delta_p$ will produce more discretizations 
+leading to non-efficient computation (green dots under the 
+$(D_\text{linear}, \delta_p)$ point on Figure \ref{fig:D-delta-grid-search}).
+
+With the definition of $\delta_p$ in mind we can subdivide the 
+subregion in two: one with points with $\delta \le \delta_p$ (region B 
+on Figure \ref{fig:D-delta-grid-search}) and other with points with 
+$\delta > \delta_p$ (region C on Figure \ref{fig:D-delta-grid-search}).
+
+It worth mention that either a greater $D$ or a lower $\delta$ 
 increases the number of discretizations of the tesseroid.
@@
 
-Taking this conclusion into account we have determined the $\delta$ 
-value needed to ensure an acceptable accuracy through a comparison 
-between the numerical approximation and the analytical solution for a 
-spherical shell with an exponential density.
-In this specific case we proposed an exponential density function  that 
-assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's 
-outer and inner radii, respectively. It can be defined as follows:
+In order to determine the value of $\delta_p$ we should compute the 
+differences between the numerical model and the analytical solution for 
+a spherical shell with an exponential density.
+Thus we must obtain this analytical solution and define a tesseroid 
+model to numerically approximate the gravitational fields.
+
+Lets consider a decreasing exponential density function that assumes 
+the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and 
+inner radii, respectively, that can be defined as follows:
 
@@
 
-\noindent where $R$ is the mean Earth radius and $b$ is a constant 
+\noindent $R$ is the mean Earth radius and $b$ is a constant 
 that determines the variability of the function: a low value of $b$ 
@@
 
-In order to perform this comparison we must compute the analytical 
-solution of the gravitational fields generated by this spherical shell 
-and define a tesseroid model to numerically approximate this fields.
-
-Firstly, the analytical solution can be computed as the sum of the 
-fields generated by a spherical shell with constant density $\rho(r') = 
-C$ and the one determined in equation \ref{eq:shell-pot-exp}.
+Firstly, the analytical solution of the gravitational fields generated 
+by a spherical shell with a density functions as the one in equation 
+\ref{eq:density-exp} can be computed as the sum of the fields generated 
+by a spherical shell with constant density $\rho(r') = C$ and another 
+shell with an exponential density like:
+
+\begin{equation}
+    \rho(r') = Ae^{-(r' - \Delta h)/b},
+\end{equation}
+
+\noindent where $A$, $\Delta h$ and $b$ are constants.
+The gravitational potential field generated by a spherical shell with 
+this density on any external point at a distance $r$ from its centre 
+can be calculated by solving the integral in equation \ref{eq:shell-pot},
+resulting as follows:
+
+\iftwocol{
+\begin{equation}
+    \begin{split}
+        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+        Ab 
+        \Big[
+        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
+        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
+        \Big].
+    \end{split}
+\label{eq:shell-pot-exp}
+\end{equation}
+}{
+\begin{equation}
+    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+    Ab \, e^\frac{\Delta h}{b}
+    \Big[
+    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+    \Big].
+\label{eq:shell-pot-exp}
+\end{equation}
+}
 
@@
 We have assigned the exponential density function proposed on equation 
-\ref{eq:density-exp} to each spherical shell model, choosing few values 
+\ref{eq:density-exp} to each spherical shell model, choosing some values 
 of the constant $b$ in order to test the numerical approximation for 

### 2018-04-24 manuscript/manuscript.tex
Shrinked Theory in manuscript

@@
 
-According to \citet[p.~390]{Hildebrand1987}, we can approximate any 
-integral in the interval $[-1, 1]$ using a $N$th order GLQ, i.e. by a 
-weighted sum of the integration kernel evaluated on the roots of the 
-$N$th Legendre polynomial:
-
-\begin{equation}
-    \int\limits_{-1}^1 f(x) dx \approx \sum_{i=1}^N W_i f(x_i),
-\end{equation}
-
-\noindent where $x_i$ are the roots of the $N$th Legendre polynomial 
-$P_N(x)$ and $W_i$ are the weights of each term:
-
-\begin{equation}
-    W_i = \frac{2}{(1-x_i^2)[P_N^\prime(x_i)]^2},
-\end{equation}
-
-\begin{equation}
-    P_N(x_i) = 0, \quad \forall i = {1,\dots,N}.
-\end{equation}
-
-In case of an integration defined on a arbitrary interval, e.g. 
-$[a,b]$, we can scale the nodes and perform a similar approximation:
-
-\begin{equation}
-    \int\limits_a^b f(x) dx \approx \frac{b-a}{2} \sum_{i=1}^N W_i f(x_i^s),
-\label{eq:glq-scaled}
-\end{equation}
-
-\noindent where $x_i^s$ is the scaled $x_i$ root into the $[a,b]$ 
-interval, also called nodes of the quadrature:
-
-\begin{equation}
-    x_i^s = \frac{b-a}{2} x_i + \frac{b+a}{2}.
-\end{equation}
-
-\noindent For example, if we use a second order GLQ, the roots of the 
-$P_2(x)$ Legendre polynomial and its corresponding weights are $x_i = 
-\pm 0.577350269$ and $W_i = 1$.
-
-Our intention is to use GLQ to approximate the volume integrals from 
-equations \ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor}.
-When tesseroids have variable densities, we can write the integral 
-kernels as the product of a certain function $g$ and the density 
-function:
-
-\begin{equation}
-    f(r', \phi', \lambda') = \rho(r') g(r', \phi', \lambda'),
-\end{equation}
-
-\noindent and then apply the quadrature to every integral corresponding 
-to each spherical coordinate, obtaining a similar expression to the one 
-shown by \citet{Asgharzadeh2007} and \citet{Uieda2016}:
+According to \citet[p.~390]{Hildebrand1987}, we can approximate each 
+integral on equations \ref{eq:tesseroid-pot}, \ref{eq:tesseroid-grav} 
+and \ref{eq:tesseroid-tensor} by applying a $N$th order GLQ, i.e. as 
+a weighted sum of the integration kernel evaluated on the roots of the 
+$N$th Legendre polynomial.
+In the variable density tesseroid case the GLQ is applied in a similar 
+way as \citet{Asgharzadeh2007} and \citet{Uieda2016} do, but in this 
+scenario the density function must be also evaluated on the quadrature 
+nodes:
 
@@
     A = 
-    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8}.
-\end{equation}
+    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8},
+\end{equation}
+
+\noindent $(r_i, \phi_j, \lambda_k)$ are the GLQ nodes and $W_i^r$ are 
+the quadrature weights \citep[see their definitions 
+on][]{Hildebrand1987}.
 

### 2018-04-24 manuscript/manuscript.tex
Splitted sentences in manuscript
Modified Methodology in manuscript

@@
 
-From equation \ref{eq:glq-var-dens} we can conclude that the error of 
-the numerical approximation is controlled by the order of the 
-quadrature, i.e. the number of point masses that are going to be used 
-to compute the gravitational field.
-This conclusion could lead to a naive methodology: find the order of 
-the quadrature that produces gravitational fields with an acceptable 
-accuracy.
-Although this simple proposal has a few problems: (a) it won't be 
-efficient and (b) it couldn't be applied to the general case (any
-tesseroid with arbitrary density function on any computational point) 
-without modifying the quadrature order taking into account 
-characteristics of the tesseroid and the computation point (tesseroid 
-size, distance to the computation point, variation of the density 
-function, etc).
+In a general way, the numerical error of the GLQ is controlled by the 
+order of the quadrature: a higher GLQ order produces a more accurate 
+approximation \citep{Hildebrand1987}.
+This can also be understood in terms of the tesseroid problem: a higher 
+quadrature order increases the number of point masses used to compute 
+the gravitational fields and thus lowers the numerical error.
+
+This reasoning could lead to a naive methodology in order to ensure 
+accurate gravitational field computations: find the order of the 
+quadrature that produces an acceptable numerical error.
+Nevertheless, this simple proposal has a few problems: (a) it won't be 
+efficient \citep{Li2011, Uieda2016} and (b) it couldn't be applied to 
+the general case, i.e. the needed quadrature order probably won't be 
+the same for every tesseroid, computation point or density function.
 
@@
 These algorithms generally involve a discretization inequation that 
-controls if the tesseroid should be subdivided or not and a certain 
+controls if the tesseroid should be subdivided or not, and a certain 
 predefined parameter that works as a threshold for the discretization 
@@
 On the variable density tesseroid problem we found two main error 
-sources in the GLQ that should be addressed: (a) the relation between 
+sources in the GLQ that should be addressed: (a) the ratio between 
 the tesseroid size and its distance to the computation point, and (b) 
@@
 the volume integrals as the gravitational effect of point masses 
-located at the nodes of the quadrature. If we set the GLQ orders to 
-fixed values, it's easy to conclude that the nearer the computation 
-point is to the tesseroid, the accuracy of the approximation is reduced 
-and the point masses effect becomes more noticeable than the tesseroid 
-we want to approximate. A complete analysis of this conclusion can be 
-found in \citet{Ku1977, Li2011, Uieda2016}.
+located at the nodes of the quadrature.
+If we set the GLQ orders to fixed values, it's easy to conclude that 
+the nearer the computation point is to the tesseroid, the accuracy of 
+the approximation is reduced and the point masses effect becomes more 
+noticeable than the tesseroid we want to approximate.
+A complete analysis of this conclusion can be found in \citet{Ku1977, 
+Li2011, Uieda2016}.
 

### 2018-04-24 manuscript/manuscript.tex
Improved Modified Adaptive Discretization Algorithm in manuscript

@@
 
-From equation \ref{eq:glq-var-dens} we saw that the GLQ approximates 
-the volume integrals as the gravitational effect of point masses 
-located at the nodes of the quadrature.
-If we set the GLQ orders to fixed values, it's easy to conclude that 
-the nearer the computation point is to the tesseroid, the accuracy of 
-the approximation is reduced and the point masses effect becomes more 
-noticeable than the tesseroid we want to approximate.
-A complete analysis of this conclusion can be found in \citet{Ku1977, 
-Li2011, Uieda2016}.
-
 Studying the homogeneous density tesseroid problem, \citet{Ku1977} 
@@
 tesseroid we want to approximate.
-A complete analysis of this conclusion can be found in \citet{Ku1977, 
-Li2011, Uieda2016}.
-
-Instead of increasing the GLQ orders to improve the accuracy, what 
-leads to a more time consuming computation, \citet{Li2011} proposed an 
-adaptive discretization algorithm which essentially consists in 
-dividing the tesseroid based on a ratio between the distance to the 
-computation point and its dimensions.
-\citet{Uieda2016} proposed a modified version of the algorithm proposed 
-by \citet{Li2011}. These modifications can be summed up in:
+One way to prevent this could be increasing the GLQ order in case that 
+the computation point is near the tesseroid.
+This would uniformly increase the number of point masses inside the 
+tesseroid volume.
+
+Instead of doing this, \citet{Li2011} proposed an adaptive 
+discretization algorithm which essentially consists in dividing the 
+tesseroid based on a ratio between the distance to the computation 
+point and its dimensions.
+This algorithm produces a more efficient computation because the 
+increase of point masses is produced where are more needed.
+
+\citet{Uieda2016} proposed a modified version of this algorithm, whose
+modifications can be summed up in:
 (1) a faster calculation of the distance from the computation point to 
@@
 homogeneous density.
-
-We will perform similar tests, but now with variable density spherical 
-shell.
-From the infinite density functions that can be proposed, we have 
-chosen three characteristic ones: a linear, an exponential and a 
-discontinuous function.
-\todo[inline]{sacaria todo este parrafo y lo introduciria mas adelante}
 

### 2018-04-24 manuscript/manuscript.tex
Modified density-based discretization algorithm in manuscript

@@
 
-Variations in the density function of an arbitrary tesseroid introduces 
+Variations in the density function of an arbitrary tesseroid introduce 
 a new type of error into the numerical approximation: only a few point 
@@
 algorithm that takes into account the variation of the density inside 
-the tesseroid. Basically, it subdivides any tesseroid on the radial 
-direction on the depth at which the \emph{maximum density variation} 
-takes place, only if a certain discretization inequation holds.
+the tesseroid.
+Basically, it subdivides any tesseroid on the radial direction on the 
+depth at which the \emph{maximum density variation} takes place, only 
+if a certain discretization inequation holds.
 
@@
 On the first iteration, the former tesseroid corresponds to the 
-\emph{original} one, so $L_r^\text{sm} = L_r$, but for future 
-iterations this radial dimension will be computed as the difference of 
-the top and bottom radii of the subdividing tesseroid.
+\emph{original} one, so $L_r^\text{sm}/L_r$ is equal to one, but for 
+future iterations this radial dimension will be computed as the 
+difference of the top and bottom radii of the subdividing tesseroid.
 The tesseroid size ratio is intended to reduce the number of probable 
@@
 The same happens if the tesseroid radial dimension is bellow the 
-threshold of 1mm. On the first case due to the fact that no more 
-discretizations are needed, and on the second one because smaller 
-tesseroids will not introduce meaningful increases in the accuracy of 
-the computation.
+threshold of 1mm.
+On the first case due to the fact that no more discretizations are 
+needed, and on the second one because smaller tesseroids will not 
+introduce meaningful increases in the accuracy of the computation.
 This two conditions guarantee that the algorithm is finite.
@@
 ensures an acceptable accuracy while minimising the computation time.
-We will perform this determination with the same methodology as the one 
-applied to the $D$ ratio: compare the numerical model with the 
-analytical solutions for a spherical shell.
 

### 2018-04-24 manuscript/manuscript.tex
Modified determination of D and delta in manuscript

@@
 
-In order to determine default values for the distance-size ratio ($D$)
-and the delta ratio ($\delta$) that ensure both numerical accuracy and 
-computation effectiveness we need to compare the numerical 
-approximation with some analytical solutions.
-
-On the homogeneous density tesseroids case, \citet{Uieda2016} and 
-\citet{Grombein2013} compared their numerical models with the analytical 
-solution of a spherical shell.
+The discretization algorithms described on the previous section are 
+introduced to improve the accuracy of the numerical approximation of 
+the gravitational fields.
+They involve two predefined ratios that determine how many 
+times each tesseroid will be discretized and thus indirectly control 
+the numerical error of the approximation.
+The modified version of the adaptive discretization algorithm involves  
+the distance-size ratio $D$, while the density-based discretization 
+algorithm makes use of the delta ratio $\delta$.
+These algorithms need default predefined values for $D$ and $\delta$ 
+in order to ensure both acceptable numerical accuracy and computation 
+effectiveness.
+
+On the homogeneous density tesseroids case, \citet{Uieda2016} compared 
+their numerical model with the analytical solution of a spherical 
+shell in order to obtain default values for the distance-size ratio 
+$D$.
 We could follow this idea, but for our needs the spherical shell must 
@@
 exponential one.
-And from these results we will finally obtain default values for $D$ 
+From these results we will finally obtain default values for $D$ 
 and $\delta$ that ensure a numerical error lower than 0.1\%.

### 2018-04-24 manuscript/manuscript.tex
Modified analytical solutions for spherical shell in manuscript
Modified linear density in manuscript

@@
 
-For future purposes we are going to obtain the gravity potential for 
-three different functions from the perspective of their variation in 
-$r'$: a linear and an exponential density function.
-\todo[inline]{lo sacaria}
-
 \subsection{Linear Density}
@@
 
-We have computed these gravitational fields on every point of the four 
+We will then compute these gravitational fields on every point of the four 
 grids for different values of $D$, ranging from 0.5 to 10 with a step 
 of 0.5.
-Then we calculated the maximum difference between these results and the 
+Then we will calculate the maximum difference between these results and the 
 value that the analytical solution assumes at the same height of the 
@@
 longitude.
-That's why we choose for comparison the maximum difference between the 
+That's why we will choose for comparison the maximum difference between the 
 results on every point of the grid and the analytical solution.
@@
 
-Then we computed the percentage difference between the numerical model 
+We have computed the percentage difference between the numerical model 
 with its corresponding analytical solution (equation 
-\ref{eq:shell-pot-linear}) for the thick and the thin spherical shell.
-
+\ref{eq:shell-pot-linear}) for the thick and the thin spherical shell 
+on each of the four grids described in Table \ref{tab:grids}.
 We can see the results for the potential, the vertical component of the 

### 2018-04-25 manuscript/manuscript.tex
Fixed typo

@@
 The main issue of the GLQ is that the approximation becomes less 
-accurate when the computation point is closer to the tesseroid of for 
+accurate when the computation point is closer to the tesseroid or for 
 smaler GLQ order \citep{Ku1977}.

### 2018-05-02 manuscript/manuscript.tex
Fixed typo

@@
 accurate when the computation point is closer to the tesseroid or for 
-smaler GLQ order \citep{Ku1977}.
+smaller GLQ order \citep{Ku1977}.
 Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 

### 2018-05-08 manuscript/manuscript.tex
Added empirical grid search figure in manuscript and more editing

@@
 
+Lets consider a decreasing exponential density function that assumes 
+the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and 
+inner radii, respectively, that can be defined as follows:
+
+\begin{equation}
+    \rho(r') = A e^{-(r' - R)/b} + C,
+\label{eq:density-exp}
+\end{equation}
+\noindent where
+\begin{equation}
+    A =
+    (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
+    \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
+\end{equation}
+\begin{equation}
+    C =
+    2670 \text{kg/m}^3 - A,
+\end{equation}
+
+\noindent $R$ is the mean Earth radius and $b$ is a constant 
+that determines the variability of the function: a low value of $b$ 
+increases the maximum slope of the density function.
+
+The analytical solution of the gravitational fields generated 
+by a spherical shell with a density functions as the one in equation 
+\ref{eq:density-exp} can be computed as the sum of the fields generated 
+by a spherical shell with constant density $\rho(r') = C$ and another 
+shell with an exponential density like:
+
+\begin{equation}
+    \rho(r') = Ae^{-(r' - \Delta h)/b},
+\end{equation}
+
+\noindent where $A$, $\Delta h$ and $b$ are constants.
+The gravitational potential field generated by a spherical shell with 
+the latter density on any external point at a distance $r$ from its centre 
+can be calculated by solving the integral in equation \ref{eq:shell-pot},
+resulting as follows:
+
+\iftwocol{
+\begin{equation}
+    \begin{split}
+        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+        Ab 
+        \Big[
+        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
+        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
+        \Big].
+    \end{split}
+\label{eq:shell-pot-exp}
+\end{equation}
+}{
+\begin{equation}
+    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
+    Ab \, e^\frac{\Delta h}{b}
+    \Big[
+    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+    \Big].
+\label{eq:shell-pot-exp}
+\end{equation}
+}
+
+Then, we proposed a thick spherical shell with a thickness of 35km and 
+modelled by a mesh of $30^\circ \times 30^\circ$ tesseroids.
+We have assigned the density function proposed on equation 
+\ref{eq:density-exp} to our spherical shell model with a $b$ value of 
+1km, i.e. a highly varying density function.
+
+We computed the gravitational potential ($V$), the vertical component 
+of its gradient ($g_z$) and the $g_zz$ component of the Marussi tensor 
+on the big grid (see Table \ref{tab:grids}) by exploring different 
+values of $D$ and $\delta$.
+These results were used to obtain the percentage difference between 
+them and their corresponding analytical solutions.
+
+On Figure \todo[inline]{referencia a figura grid search} we show 
+the resulting percentage difference for each gravitational field.
+The points inside the dotted lines are the ones that present a 
+numerical error lower than the 0.1\% threshold, i.e. they represent the 
+eligible ($D$, $\delta$) pairs.
+The one that produces the more efficient computation must be chosen as 
+the default $D$ and $\delta$ value.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]
+    {figures/grid-search.pdf}
+\caption{
+    }
+\label{fig:grid-search}
+\end{figure}
+
 \begin{figure}
@@
 
-It worth mention that either a greater $D$ or a lower $\delta$ 
-increases the number of discretizations of the tesseroid.
-Although a greater $D$ may divide it in the latitude, longitude 
-and/or radial direction; a lower $\delta$ will only discretize the 
-original tesseroid in the radial direction.
-This means that a decrease in $\delta$ will produce a more efficient 
-computation than an increase in $D$.
-\todo[inline]{Should we keep this paragraph? The same is explained bellow.}
+%~ It worth mention that either a greater $D$ or a lower $\delta$ 
+%~ increases the number of discretizations of the tesseroid.
+%~ Although a greater $D$ may divide it in the latitude, longitude 
+%~ and/or radial direction; a lower $\delta$ will only discretize the 
+%~ original tesseroid in the radial direction.
+%~ This means that a decrease in $\delta$ will produce a more efficient 
+%~ computation than an increase in $D$.
+%~ \todo[inline]{Should we keep this paragraph? The same is explained bellow.}
 
@@
 Figure \ref{fig:D-delta-grid-search}).
-Although, if we find a point with acceptable numerical error in this 
-region, we know it will produce more discretizations than the 
+Nevertheless, the numerical error exploration shown on
+Figure \todo[inline]{ref a figura grid search} doesn't present any 
+point in this region with error lower than the threshold.
+Even if we are analysing a single case, it proves that no point in 
+region C is eligible as default ($D$, $\delta$) pair because it should 
+ensure numerical accuracy for every exponential density scenario.
+
+We can arrive to the same conclusion by analysing the computation 
+efficiency: even if we find a point with acceptable numerical error in 
+this region, we know it will produce more discretizations than the 
 ($D_\text{linear}$, $\delta_p$) point, leading to a non-efficient 
@@
 
-In order to determine the value of $\delta_p$ we should compute the 
-differences between the numerical model and the analytical solution for 
-a spherical shell with an exponential density.
-Thus we must obtain this analytical solution and define a tesseroid 
-model to numerically approximate the gravitational fields.
-
-Lets consider a decreasing exponential density function that assumes 
-the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and 
-inner radii, respectively, that can be defined as follows:
-
-\begin{equation}
-    \rho(r') = A e^{-(r' - R)/b} + C,
-\label{eq:density-exp}
-\end{equation}
-\noindent where
-\begin{equation}
-    A =
-    (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
-    \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
-\end{equation}
-\begin{equation}
-    C =
-    2670 \text{kg/m}^3 - A,
-\end{equation}
-
-\noindent $R$ is the mean Earth radius and $b$ is a constant 
-that determines the variability of the function: a low value of $b$ 
-increases the maximum slope of the density function.
-
-Firstly, the analytical solution of the gravitational fields generated 
-by a spherical shell with a density functions as the one in equation 
-\ref{eq:density-exp} can be computed as the sum of the fields generated 
-by a spherical shell with constant density $\rho(r') = C$ and another 
-shell with an exponential density like:
-
-\begin{equation}
-    \rho(r') = Ae^{-(r' - \Delta h)/b},
-\end{equation}
-
-\noindent where $A$, $\Delta h$ and $b$ are constants.
-The gravitational potential field generated by a spherical shell with 
-this density on any external point at a distance $r$ from its centre 
-can be calculated by solving the integral in equation \ref{eq:shell-pot},
-resulting as follows:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-        Ab 
-        \Big[
-        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
-        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
-        \Big].
-    \end{split}
-\label{eq:shell-pot-exp}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-    Ab \, e^\frac{\Delta h}{b}
-    \Big[
-    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    \Big].
-\label{eq:shell-pot-exp}
-\end{equation}
-}
-
-Secondly, we proposed a thin and a thick spherical shells with 1km and 
-35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
-30^\circ$ tesseroids.
-We have assigned the exponential density function proposed on equation 
-\ref{eq:density-exp} to each spherical shell model, choosing some values 
-of the constant $b$ in order to test the numerical approximation for 
-different density variations.
-
-Finally, we computed the gravitational potential ($V$), the 
-vertical component of its gradient ($g_z$) and the $g_zz$ component of 
-the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
-keeping the distance-size ratio equal to the $D$ values obtained in the 
-density case for its corresponding gravitational field and exploring 
-different values of $\delta$.
-We then computed the percentage difference between this numerical 
-results and the analytical solutions.
+%~ In order to determine the value of $\delta_p$ we should compute the 
+%~ differences between the numerical model and the analytical solution for 
+%~ a spherical shell with an exponential density.
+%~ Thus we must obtain this analytical solution and define a tesseroid 
+%~ model to numerically approximate the gravitational fields.
+
+%~ Secondly, we proposed a thin and a thick spherical shells with 1km and 
+%~ 35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
+%~ 30^\circ$ tesseroids.
+%~ We have assigned the exponential density function proposed on equation 
+%~ \ref{eq:density-exp} to each spherical shell model, choosing some values 
+%~ of the constant $b$ in order to test the numerical approximation for 
+%~ different density variations.
+
+%~ Finally, we computed the gravitational potential ($V$), the 
+%~ vertical component of its gradient ($g_z$) and the $g_zz$ component of 
+%~ the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
+%~ keeping the distance-size ratio equal to the $D$ values obtained in the 
+%~ density case for its corresponding gravitational field and exploring 
+%~ different values of $\delta$.
+%~ We then computed the percentage difference between this numerical 
+%~ results and the analytical solutions.
 

### 2018-05-09 manuscript/manuscript.tex
Changes in manuscript

@@
 \caption{
-    Differences between the gravity fields generated by the numerical 
-    model and the analytical solution for (a) a thin spherical shell of 
-    1km of thickness and (b) a thick spherical shell of 35km of thickness,
-    both with the linear density defined on equation 
+    Differences between the gravity fields generated by the numerical model 
+    and the analytical solution for a thin and a thick spherical shell of 1km 
+    and 35km of thickness, respectively.
+    Both models have the linear density defined on equation 
     \ref{eq:density-linear}.
@@
 
-On Figure \todo[inline]{referencia a figura grid search} we show 
+On Figure \ref{fig:grid-search} we show 
 the resulting percentage difference for each gravitational field.
@@
 Nevertheless, the numerical error exploration shown on
-Figure \todo[inline]{ref a figura grid search} doesn't present any 
+Figure \ref{fig:grid-search} doesn't present any 
 point in this region with error lower than the threshold.
-Even if we are analysing a single case, it proves that no point in 
+Even though we are analysing a single case, it proves that no point in 
 region C is eligible as default ($D$, $\delta$) pair because it should 
@@
 discretizations while keeping an acceptable numerical error under the 
-0.1\% threshold is the one on which the ratio $D$ is equal to 
-$D_\text{linear}$ and the delta ratio is $\delta_p$.
+0.1\% threshold is the one whose ratio $D$ is equal to 
+$D_\text{linear}$ and its delta ratio is $\delta_p$.
+
+In order to determine a default value of $\delta$ that ensures an 
+acceptable accuracy for every exponential density case we have 
+performed comparisons between the numerical model and the analytical 
+solutions, keeping the distance-size ratio $D$ equal to the values 
+obtained for the linear density case while exploring different values of 
+$\delta$.
+
+We have considered the same spherical shell models we have used in the 
+linear density case: a thin one (1km of thickness) and a thick one 
+(35km of thickness) composed by $30^\circ \times 30^\circ$ tesseroids.
+But in this case, we assigned an exponential density function like the 
+one shown in equation \ref{eq:density-exp} to each shell, choosing 
+different values of $b$ in order to explore distinct density 
+variations.
+
+Finally, we have computed the gravitational potential ($V$), the 
+vertical component of its gradient ($g_z$) and the $g_zz$ component of 
+the Marussi tensor on the big grid (see Table \ref{tab:grids}) and 
+calculated the percentage difference between these numerical results and 
+their corresponding analytical solutions.
+Our intention is to set the default $\delta$ as the highest value whose 
+numerical error is bellow the $0.1\%$ threshold.
 
@@
 
-On Figures \ref{fig:delta-exponential}(a) and 
-\ref{fig:delta-exponential}(b) we show the density functions used in 
-the comparison and the resulting percentage difference for each 
+On Figure \ref{fig:delta-exponential} we show the density functions 
+used on each comparison and the resulting percentage difference for each 
 gravitational field.
-The former shows the results for the thin spherical shell, while the 
-later shows the ones corresponding to the thick shell.
-From these results we can conclude that for the potential and $g_z$ 
-computation, a $\delta = 0.3$ is needed in order to achieve an 
-acceptable numerical error, while a $\delta = 0.02$ is required for the 
-$g_{zz}$ Marussi tensor component.
+The first column shows the results for the thin spherical shell, while the 
+second one shows the ones corresponding to the thick shell.
+From these we can conclude that a $\delta = 0.2$ is needed in order to 
+achieve an acceptable numerical error for the computation of the 
+potential and the $g_z$ component.
+Because the same comparison using the other components of the gradient ($g_x$ 
+and $g_y$) could not be carried out, we will extrapolate this result 
+setting a default $\delta = 0.2$ for their computation.
+
+On the other hand, the comparison with the $g_{zz}$ component of the 
+Tensor shows a difference between the thin and the thick spherical 
+shell: the former needs a $\delta = 0.2$, but the latter requires a 
+$\delta = 0.01$.
+In order to guarantee a good accuracy for every case, we will keep the 
+more conservative value of $\delta = 0.02$ for the computation of the 
+Marussi tensor components.
 

### 2018-05-09 manuscript/manuscript.tex
Changes in manuscript

@@
 On the variable density tesseroid problem we found two main error 
-sources in the GLQ that should be addressed: (a) the ratio between 
-the tesseroid size and its distance to the computation point, and (b) 
+sources in the GLQ that should be addressed: (a) the distance between 
+the tesseroid and the computation point, and (b) 
 the variation of the density function.
@@
 
-Because the algorithm will be iteratively applied for every 
+\noindent Because the algorithm will be iteratively applied for every 
 subdividing tesseroid, we'd like to note that this normalised density 
 function will be the same for every subdividing tesseroid.
-
 In case that the density function is constant, both the maximum and 
@@
 \noindent where the delta ratio $\delta$ is a predefined constant value 
-and $L_r^\text{sm}/L_r$ is a tesseroid size ratio, where 
-$L_r^\text{sm}$ and $L_r$ are the radial dimensions of the tesseroid 
-that is currently analysed and of the \emph{original} one, 
+and $L_r^\text{sm}/L_r$ is defined as the tesseroid size ratio.
+Besides, $L_r^\text{sm}$ and $L_r$ are the radial dimensions of the 
+tesseroid that is currently analysed and of the \emph{original} one, 
 respectively.
@@
 A big subdividing tesseroid with small maximum absolute difference would 
-introduce more numerical error than a small subdividing tesseroid with 
+contribute more numerical error than a small subdividing tesseroid with 
 higher maximum absolute difference.
@@
 In contrast with the homogeneous density case, the analytical solutions 
-for a variable density spherical shell must be obtained.
+for a variable density spherical shell are not present in the 
+literature, so they must be obtained.
 
@@
 reproduces the gravity potential field on any outside point at distance 
-$r$ from the its centre.
+$r$ from its centre.
 
@@
 obtained from the following expression:
+\todo[inline]{revisar redaccion}
 
@@
 = b$ \citep{Mikuska2006,Grombein2013}.
-So the potential can be read as the one generated by a combination of 
-both shells.
-
-Secondly we can analyse how the numerical algorithm works for any 
-tesseroid with linear density.
-When applying the density-based discretization algorithm on any 
-tesseroid with this kind of density, the absolute difference defined in 
-equation \ref{eq:density-abs-diff} will always be zero, so the 
-discretization inequation \ref{eq:delta-density} will never be 
-satisfied.
+
+Secondly we can analyse how the density-based discretization algorithm 
+works in case of a linear density function.
+Because the absolute difference defined on equation 
+\ref{eq:density-abs-diff} will always be zero, the discretization 
+inequation \ref{eq:delta-density} will never be satisfied.
 Therefore, independently of the linear function proposed, the 
@@
 accuracy of the numerical approximation.
-That's why we only need to determine the minimum value of $D$ needed in 
-order to guarantee an acceptable accuracy while ignoring the value of 
-$\delta$.
+That's why, for linear density scenarios, we only need to determine 
+the minimum value of $D$ needed in order to guarantee an acceptable 
+accuracy while ignoring the value of $\delta$.
 
@@
 
-We will then compute these gravitational fields on every point of the four 
+We will compute these gravitational fields on every point of the four 
 grids for different values of $D$, ranging from 0.5 to 10 with a step 
@@
 potential and the $g_z$ component of its gradient needs a value of 
-$D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
+$D=1$ and $D=2$, respectively, in order to ensure a numerical error 
+bellow the 0.1\% threshold.
 On the other hand, the $g_{zz}$ computations show a noticeable 
@@
 
-Lets consider a decreasing exponential density function that assumes 
-the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and 
-inner radii, respectively, that can be defined as follows:
+Firstly, lets consider a decreasing exponential density function that 
+assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's 
+outer and inner radii, respectively, that can be defined as follows:
 
@@
 
-Then, we proposed a thick spherical shell with a thickness of 35km and 
-modelled by a mesh of $30^\circ \times 30^\circ$ tesseroids.
+Secondly, we propose a thick spherical shell with a thickness of 35km and 
+modelled by a mesh of $30^\circ \times 30^\circ$ tesseroids in order to 
+perform the exploration of the numerical error in the $D$, $\delta$ 
+space.
 We have assigned the density function proposed on equation 
@@
 
-We computed the gravitational potential ($V$), the vertical component 
-of its gradient ($g_z$) and the $g_zz$ component of the Marussi tensor 
-on the big grid (see Table \ref{tab:grids}) by exploring different 
-values of $D$ and $\delta$.
+Finally, we have computed the gravitational potential ($V$), the vertical 
+component of its gradient ($g_z$) and the $g_{zz}$ component of the 
+Marussi tensor on the big grid (see Table \ref{tab:grids}) by exploring 
+different values of $D$ and $\delta$.
 These results were used to obtain the percentage difference between 
@@
 The points inside the dotted lines are the ones that present a 
-numerical error lower than the 0.1\% threshold, i.e. they represent the 
-eligible ($D$, $\delta$) pairs.
+numerical error lower than the 0.1\% threshold.
 The one that produces the more efficient computation must be chosen as 
@@
 to non-efficient computations.
+It's worth remembering that a lower $\delta$ or a greter $D$ will 
+produce more discretizations, increasing the accuracy but also the 
+computation time.
 
@@
 
-We can arrive to the same conclusion by analysing the computation 
-efficiency: even if we find a point with acceptable numerical error in 
-this region, we know it will produce more discretizations than the 
-($D_\text{linear}$, $\delta_p$) point, leading to a non-efficient 
-computation.
-This is due to the fact that even an increase in $\delta$ will produce 
-less density-based discretizations, the increase in $D$ will introduce 
-even more subdivisions not only in the radial direction, but 
-also in the latitude and longitude ones.
-
 In conclusion, the ($D$, $\delta$) point that minimises the number of 
@@
 
-We have considered the same spherical shell models we have used in the 
-linear density case: a thin one (1km of thickness) and a thick one 
-(35km of thickness) composed by $30^\circ \times 30^\circ$ tesseroids.
+Firstly, we have considered the same spherical shell models we have 
+used in the linear density case: a thin one (1km of thickness) and a 
+thick one (35km of thickness) composed by $30^\circ \times 30^\circ$ 
+tesseroids.
 But in this case, we assigned an exponential density function like the 
 one shown in equation \ref{eq:density-exp} to each shell, choosing 
-different values of $b$ in order to explore distinct density 
+a few values of $b$ in order to explore different density 
 variations.
@@
 Finally, we have computed the gravitational potential ($V$), the 
-vertical component of its gradient ($g_z$) and the $g_zz$ component of 
+vertical component of its gradient ($g_z$) and the $g_{zz}$ component of 
 the Marussi tensor on the big grid (see Table \ref{tab:grids}) and 

### 2018-05-09 manuscript/manuscript.tex
Added caption to grid search figure in manuscript

@@
 \caption{
+    Numerical error exploration in the $D$, $\delta$ space.
+    The percentage difference values were obtained from the comparison 
+    between the analytical solution and the numerical approximation of 
+    the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a 
+    spherical shell with an exponential density as the one defined on 
+    equation \ref{eq:density-exp}.
+    These comparison have been carried out with a spherical shell of 
+    35km of thickness, whose density function has a value of $b$ of 1km 
+    and the gravitational fields were computed on the big grid 
+    (defined in Table \ref{tab:grids}).
+    The points inside the dashed line are the ones that present an 
+    error lower than 0.1\%.
     }

### 2018-05-10 manuscript/manuscript.tex
Minor modification in manuscript
Updated Implementation section in manuscript

@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/linear-D.pdf}
+\includegraphics[width=\linewidth]{figures/linear-D.pdf}
 \caption{
@@
 In order to guarantee a good accuracy for every case, we will keep the 
-more conservative value of $\delta = 0.02$ for the computation of the 
+more conservative value of $\delta = 0.01$ for the computation of the 
 Marussi tensor components.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Implementation}
+
+In order to perform the determinations of the distance-size and the 
+delta ratio, we have developed a Python code that implements the 
+discretization algorithms and the GLQ to finally compute the 
+gravitational fields.
+
+It is based on the pre-existing code developed by \citet{Uieda2016} (and 
+included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that computes 
+the gravitational effects of tesseroids with homogeneous density.
+Our implementation keeps the the homogeneous tesseroid calculation and 
+the modified adaptive discretization algorithm while incorporates the 
+new variable density computations with the density-based discretization 
+algorithm.
+On their implementation, \citet{Uieda2016} have optimised the more time 
+consuming functions through the Numba compiler.
+In contrast, we have decided to develop our new library in Cython 
+language due to difficulties when passing a Python function as argument 
+to precompiled Numba methods.
+
+This new code is freely available under the BSD 3-clause open-source 
+license and makes use of the pre-existing classes and functions of 
+Fatiando a Terra v0.5 in order to define single tesseroid or meshes of 
+them.
+It can be downloaded from ...
+\todo[inline]{include repo url}
+
 \todo[inline, color=red]{Text bellow is old}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%~ \subsubsection{Linear Density}
-
-%~ If the density of the spherical shell is a linear function on the radial coordinate, i.e.:
-
-%~ \begin{equation}
-    %~ \rho(r') = ar' + b,
-%~ \end{equation}
-
-%~ \noindent then the gravity potential generated on any computation point at distance $r$ from its centre can be obtained from the following expression:
-
-%~ \begin{equation}
-    %~ V_\text{sh}^\text{lin}(r) = \pi G \left[ 
-    %~ a \frac{R_2^4 - R_1^4}{r} +
-    %~ b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
-    %~ \label{eq:shell-pot-linear}
-%~ \end{equation}
-
-%~ The first term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with variable density $\rho(r') = ar'$, while the second one constitutes the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
-%~ So the potential can be read as the one generated by a combination of both shells.
-
-%~ \subsubsection{Exponential Density}
-
-%~ If, on the other hand, the spherical shell has an exponential density in the radial coordinate,
-
-%~ \begin{equation}
-    %~ \rho(r') = Ae^{-(r' - \Delta h)/b},
-%~ \end{equation}
-
-%~ \noindent the potential (calculated through SymPy \citep{sympy2017}) can be written as follows:
-
-%~ \iftwocol{
-%~ \begin{equation}
-    %~ \begin{split}
-        %~ V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-        %~ Ab 
-        %~ \Big[
-        %~ & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
-        %~ & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
-        %~ \Big].
-    %~ \end{split}
-%~ \label{eq:shell-pot-exp}
-%~ \end{equation}
-%~ }{
-%~ \begin{equation}
-    %~ V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
-    %~ Ab \, e^\frac{\Delta h}{b}
-    %~ \Big[
-    %~ (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
-    %~ (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    %~ \Big].
-%~ \label{eq:shell-pot-exp}
-%~ \end{equation}
-%~ }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Implementation}
-
-We have developed a Python library that computes the gravity fields generated by any Tesseroid with variable density in depth.
-It is freely available under the BSD 3-clause open-source library and can be downloaded from the following repository: \todo[inline]{agregar repo, vale la pena poner el link?}
-
-This new library approximates the gravitational effect of each tesseroid through the GLQ, and guarantees its accuracy with the modified version of the adaptive discretization algorithm, original developed by \citet{Li2011} and improved by \citet{Uieda2016}, and a new density-based discretization algorithm.
-The later one had to be developed in order to decrease a new type of error in the numerical approximation caused by high variations in the density functions.
-
-The new forward model is based on the preexisting code developed by \citet{Uieda2016} (and included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that is able to compute the gravitational effects of tesseroids with homogeneous density.
-On that implementation \citet{Uieda2016} have optimised the more time consuming functions through the Numba compiler.
-In contrast, we have decided to develop our new library in Cython language due to difficulties when passing a Python function as argument to precompiled Numba methods, keeping the homogeneous tesseroid calculation and adding the new variable density computations.
-The Cython language is a superset of Python language that allows to generate efficient C code through the Cython compiler, producing a library whose functions can be called from regular Python code.
-
-Moreover, it is also compatible with the preexisting open-source library Fatiando a Terra and its usage is very similar to the one for homogeneous tesseroids:
-
-\begin{enumerate}
-\renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Define a single tesseroid (or a set of several tesseroids) with the Fatiando a Terra classes.
-    \item Add a predefined Python density function to the tesseroid (or to each tesseroid in the set).
-    \item Calculate the desired gravity field (potential, gradient or tensor components) on any set of computation points.
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%~ \section{Determination of distance-size ratio}
-
-%~ In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
-%~ The first one constitutes the simplest variation, the second one represents a higher varying function, while the later one tries to emulate a contact between two different media (e.g.
-crust and mantle interface).
-
-%~ We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
-%~ More information about this grids can be found on Table \ref{tab:grids}.
-
-%~ We computed these gravity fields on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, without applying the density-based discretization algorithm in order to isolate the behaviour of the adaptive discretization one.
-%~ Then we calculated the maximum difference between these results and the value that the analytical solution assumes at the same height.
-
-%~ Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
-%~ That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.
-
-%~ Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.
-
-%~ \begin{table}
-%~ \caption{
-    %~ Description of the synthetic grids on which the comparison of the numerical model against the analytical solutions for a spherical shell was done.
-This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height over the Earth surface.
-Every grid consists in a set of 10$\times$10 points.
-%~ }
-%~ \label{tab:grids}
-%~ \begin{tabular}{lcccc}
-    %~ Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
-    %~ Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    %~ Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
-    %~ Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
-    %~ Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
-%~ \end{tabular}
-%~ \end{table}
-
-%~ \subsection{Linear Density}
-%~ Firstly,  we proposed a spherical shell with linear density in such way it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:
-
-%~ \begin{equation}
-    %~ \rho(r') = ar' + c,
-    %~ \label{eq:density-linear}
-%~ \end{equation}
-%~ \noindent with 
-%~ \begin{equation}
-    %~ a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R_2 - R_1},
-%~ \end{equation}
-%~ \begin{equation}
-    %~ c = \frac{3300\text{kg/m$^3$} - 
-        %~ 2670\text{kg/m$^3$}}{R_2 - R_1} R + 
-        %~ 2670\text{kg/m$^3$},
-%~ \end{equation}
-
-%~ \noindent where $R$ = 6378.137 km is the mean Earth radius.
-
-%~ Then we computed the difference between the numerical model with its corresponding analytical solution (equation \ref{eq:density-linear}) for a thick and a thin spherical shell of 1km and 35km of thickness, respectively.
-
-%~ We can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
-%~ We exclude the results corresponding to the other two diagonal components of the tensor because they are similar to the ones obtained for the $g_{zz}$, although they can be found on the repository.
-\todo[inline]{chequear}
-
-%~ For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
-%~ On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D$ equal to 8 while the later only a $D$ of 3.5.
-%~ We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
-%~ \caption{
-    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}.
-The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-%~ }
-%~ \label{fig:D-linear-thin}
-%~ \end{figure}
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
-%~ \caption{
-    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}.
-The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-%~ }
-%~ \label{fig:D-linear-thick}
-%~ \end{figure}
-
-%~ \subsection{Exponential Density}
-
-%~ To test the model in case of an exponential density we proposed the following function:
-
-%~ \begin{equation}
-    %~ \rho(r') = A e^{-(r' - R)/b} + C,
-%~ \label{eq:density-exp}
-%~ \end{equation}
-%~ \noindent where
-%~ \begin{equation}
-    %~ A =
-    %~ (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
-    %~ \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
-%~ \end{equation}
-%~ \begin{equation}
-    %~ C =
-    %~ 2670 \text{kg/m}^3 - A,
-%~ \end{equation}
-
-%~ \noindent and the value of $b$ equal to the thickness of the spherical shell we will use to perform the test.
-%~ In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
-%~ The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function (like the one in equation \ref{eq:density-exp}) and for an homogeneous spherical shell.
-
-%~ We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, without applying the density-based discretization algorithm.
-%~ On both cases, the constant $b$ is equal to the thickness of the shell: 1km for the thin shell and 35km for the thick one.
-%~ The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
-%~ Once again, we exclude the ones corresponding to $g_{xx}$ and $g_{yy}$ due to their similarity with the ones obtained for $g_{zz}$, although they can be found in the repository.
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/exponential-D-thin.pdf}
-%~ \caption{
-    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
-The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-%~ }
-%~ \label{fig:D-exp-thin}
-%~ \end{figure}
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/exponential-D-thick.pdf}
-%~ \caption{
-    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}.
-The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-%~ }
-%~ \label{fig:D-exp-thick}
-%~ \end{figure}
-
-%~ In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.
-
-%~ Moreover, we can see that the differences for the potential and the $g_z$ component of the gradient keep constant even for high values of $D$.
-%~ Our interpretation is that the modified adaptive discretization algorithm is limited when increasing the accuracy for this density function, supporting the need of a density-based discretization algorithm.
-
-%~ Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$ when the density-based discretization algorithm is not applied.
-%~ In order to do this we proposed a thin spherical shell (1km of thickness) with a set of density functions, each one with a different value of $b$.
-%~ We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the corresponding analytical solution.
-%~ These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.
-
-%~ On Figure \ref{fig:D-exp-power-thin}a we can see the density functions used to perform these tests and its corresponding results on \ref{fig:D-exp-power-thin}b .
-%~ Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b=100$m the difference is almost 100\%.
-
-%~ We can conclude that even when the density-based discretization algorithm is not applied, the numerical approximation performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, i.e.
-for low density variations.
-%~ On the other hand, a density-based discretization algorithm is needed in order to guarantee the accuracy of the approximation in case of a high varying density function.
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/exponential-b-thin.pdf}
-%~ \caption{
-    %~ Results of the distance-size ratio determination for different exponential density functions in case of a spherical shell of 1km of thickness.
-    %~ (a) Density functions used in the tests, each one with a different value of $b$.
-    %~ (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function.
-Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy.
-%~ }
-%~ \label{fig:D-exp-power-thin}
-%~ \end{figure}
-
-%~ \subsection{Discontinuous Density}
-
-%~ As a final case, we performed the same test with a discontinuous density function:
-
-%~ \begin{equation}
-    %~ \rho(r') = 
-    %~ \begin{cases}
-        %~ \rho_A & \text{if} \, R_1 \leq r' < R_c \\
-        %~ \rho_B & \text{if} \, R_c \leq r' \leq R_2 \\
-    %~ \end{cases}
-%~ \label{eq:density-discontinuous}
-%~ \end{equation}
-
-%~ \noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the radius on which the density change takes place.
-
-%~ The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_A$ and $\rho_B$:
-
-%~ \begin{equation}
-    %~ V(r) = \frac{4}{3} \frac{\pi G}{r}
-    %~ \left[ \rho_A (R_c^3 - R_1^3) +
-    %~ \rho_B (R_2^3 - R_c^3) \right]
-%~ \end{equation}
-
-%~ Lets assume that $R_c = (R_1 + R_2)/3$ and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
-%~ Its results can be seen in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/discontinuous-D-asym.pdf}
-%~ \caption{
-    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous}.
-The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$.
-If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
-%~ }
-%~ \label{fig:D-discont-asym}
-%~ \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%~ \section{Determination of delta ratio}
-
-%~ In the light of the previous results, a density-based discretization algorithm is needed in order to achieve an acceptable accuracy in those cases on which the modified adaptive discretization algorithm is not sufficient to take into account the numerical error due to density variations.
-
-%~ Besides the implementation of this new algorithm, we must determine the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold in order to guarantee the good performance of the numerical model.
-
-%~ To do this we applied a similar methodology used in the determination of the distance-size ratio $D$.
-%~ Using different values of $\delta$, ranging from 0.1 to 1 with a step of 0.1, we computed the potential, the $g_z$ component of the gradient and the $g_{zz}$ component of the Marussi tensor generated by a spherical shell with several exponential densities (like the one in equation \ref{eq:density-exp}): each one with a different value of $b$.
-%~ We performed this tests for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.
-
-%~ Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a show the exponential density functions used on each test for a thin and thick spherical shell, respectively.
-%~ While on Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b we can see the differences between the resulting $g_{zz}$ with the corresponding analytical solution.
-%~ The results for the other gravity fields were excluded from the figures due to their similarities with the ones for $g_{zz}$.
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
-%~ \caption{
-    %~ Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
-    %~ (a) Density functions used to perform the test.
-    %~ (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
-%~ \label{fig:exponential-delta-thin}
-%~ \end{figure}
-
-%~ \begin{figure}
-%~ \centering
-%~ \includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
-%~ \caption{
-    %~ Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
-    %~ (a) Density functions used to perform the test.
-    %~ (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
-%~ \label{fig:exponential-delta-thick}
-%~ \end{figure}
-
-%~ In the thin shell case we can conclude that a delta equal to 0.3 guarantees an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
-%~ We will keep the more conservative value of 0.2 as the default value of delta on future computations.
-
-%~ Furthermore, we computed the difference between the numerical model and the analytical solution in case of a discontinuous density (as the one in equation \ref{eq:density-discontinuous} with $R_c = (R_1 + R_2)/3$) using a $\delta = 0.2$.
-%~ The result for the $g_{zz}$ component of the Marussi tensor is bellow the 0.1\% threshold, achieving a good performance for this default value of $\delta$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Speed Comparison}
+
+\subsection{Speed Comparison}
 

### 2018-05-10 code/tesseroid_density/tesseroid.py
Changed default delta value for g_zz in tesseroid.py

@@ -107,7 +107,7 @@ RATIO_GG = 8
 STACK_SIZE = 100
 DELTA_V = 0.2
 DELTA_G = 0.2
-DELTA_GG = 0.02
+DELTA_GG = 0.01
 
 
 def _check_input(lon, lat, height, model, ratio, njobs, pool):

### 2018-05-10 manuscript/manuscript.tex
Fixed exponential density test figure caption

@@
 \caption{
-    Density functions used in the model comparison and the differences 
-    between the gravitational fields generated by the numerical model 
-    and the analytical solution for (a) a thin spherical shell of 1km 
-    of thickness and (b) a thick spherical shell of 35km of thickness, 
-    both with the exponential density defined on equation 
+    Density functions used in the $\delta$ ratio determination and the 
+    differences between the gravitational fields generated by the 
+    numerical model and the analytical solution for a thin and a thick 
+    spherical shell of 1km adn 35km of thickness, respectively, both 
+    with the exponential density defined on equation 
     \ref{eq:density-exp}.
@@
     has achieved an acceptable accuracy for the corresponding value of 
-    $D$.
+    $\delta$.
     }

### 2018-05-15 manuscript/manuscript.tex
Edited Example with real data in manuscript

@@
 
-\todo[inline, color=red]{Text bellow is old}
+\todo[inline]{I'm thinking to remove the speed comparison}
 
@@
 
-Besides all the tests we have carried out on the new code, we intend to show a forward modelling example with real data.
-In order to do that we have chosen the Neuqu\'en Basin: a sedimentary basin located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude (see Figure \ref{fig:neuquen-basin}a), that includes continental and marine siliciclastics, carbonates and evaporites accumulated over the Jurassic and the Cretaceous constituting a stratigraphic record up to 5000m of depth \citep{Howell2005}.
+Besides all the tests we have carried out on the new code, we intend to show a 
+forward modelling example with real data.
+
+Besides the determination of the distance-size and the delta ratio, we 
+wanted to show how the algorithm performs under a real data forward model.
+In order to do it we have chosen the Neuqu\'en Basin: a sedimentary basin 
+located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude 
+(see Figure \ref{fig:neuquen-basin}a), that includes continental and marine 
+siliciclastics, carbonates and evaporites accumulated over the Jurassic and 
+the Cretaceous constituting a stratigraphic record up to 5000m of depth 
+\citep{Howell2005}.
 
@@
 \caption{
-         Forward model example with real data: Nequ\'en sedimentary modelled through tesseroids with exponential density.
+    Forward model example with real data: Nequ\'en sedimentary modelled 
+    through tesseroids with exponential density.
     (a) Topography of Neuqu\'en Basin (in km),
-         (b) Thickness of the sedimentary basin (in meters),
-         (c)-(h) Computed gravity fields: potential $V$ (in J/kg), gradient components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components $g_{xy}$ and $g_{zz}$ (in Eötvös), calculated at 50km of height over the ellipsoid.}
+    (b) thickness of the sedimentary basin (in meters),
+    (c) density function used to model the basin (depth in km and density in 
+    kg/m$^3$),
+    (d)-(l) computed gravity fields: potential $V$ (in J/kg), gradient 
+    components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components 
+    (in Eötvös), calculated at 50km of height over the ellipsoid.}
 \label{fig:neuquen-basin}
@@
 The thickness of the sediment pack was obtained from \citet{Heine2007}.
-We managed to produce a regular grid with a resolution of 0.05$^\circ$ on both longitude and latitude directions with one value of thickness on each node (see Figure \ref{fig:neuquen-basin}b).
-
-In order to compute the gravity fields generated by this basin we created a mesh of tesseroids, each one located on one of the nodes of the grid, with dimensions of 0.05$^\circ$ $\times$ 0.05$^\circ$ and depth equal to the thickness of the basin on the corresponding node.
+We managed to produce a regular grid with a resolution of 0.05$^\circ$ on both 
+longitude and latitude directions with one value of thickness on each node 
+(see Figure \ref{fig:neuquen-basin}(b).
+
+In order to compute the gravity fields generated by this basin we created a 
+mesh of tesseroids, each one located on one of the nodes of the grid, with 
+dimensions of 0.05$^\circ$ $\times$ 0.05$^\circ$ and depth equal to the 
+thickness of the basin on the corresponding node.
 
 We must also define a density function for the entire model.
-\citet{Sigismondi2012} proposed a minimum and maximum density contrast for the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
-\todo[inline]{ver que esos valores son de contraste y tienen q ir negativos.
-poner los valores absolutos.}
-So we have chosen, for this specific example, an exponential density (like the one in equation \ref{eq:density-exp}) that assumes the minimum value on the surface and the maximum at 5858m, i.e.
-the bottom of the basin, with a value of $b$ equal to 3km.
-
-We have finally computed the gravity potential $V$, the gradient components $g_x$, $g_y$ and $g_z$, and the Marussi tensor components $g_{xy}$ and $g_{zz}$ on a computation grid of 159$\times$163 nodes with a 0.05$^\circ$ spacing on both longitude and latitude directions at a 50km height over the reference ellipsoid.
-The resulting gravity fields can be seen in Figures \ref{fig:neuquen-basin}c-h.
-We have excluded the other components of the Marussi tensor from the Figure, although they can be found in the repository.
+\citet{Sigismondi2012} proposed a minimum and maximum density contrast for the 
+Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively. 
+So we have chosen, for this specific example, an exponential density (like the 
+one in equation \ref{eq:density-exp}) that assumes the minimum value on the 
+surface and the maximum at 5858m, i.e. the bottom of the basin, with a value 
+of $b$ equal to 2km.
+This density function can be seen on Figure \ref{fig:neuquen-basin}(c).
+
+We have computed the gravity potential $V$, the gradient components $g_x$, 
+$g_y$ and $g_z$, and the Marussi tensor components $g_{xx}$, $g_{xy}$, 
+$g_{xz}$, $g_{yy}$  and $g_{zz}$ on a computation grid of 159$\times$163 nodes 
+with a 0.05$^\circ$ spacing on both longitude and latitude directions at a 
+50km height over the reference ellipsoid.
+The resulting gravity fields can be seen in Figures 
+\ref{fig:neuquen-basin}(d)-(l).
+We have excluded the other components of the Marussi tensor from the Figure, 
+although they can be found in the repository.
+
+\todo[inline, color=red]{Text bellow is old}
 

### 2018-05-16 manuscript/manuscript.tex
Edited conclusions

@@
 
-We have developed a Python library that allows us to perform forward gravity computations using tesseroids with variable density in depth.
-The code numerically approximates the integral that defines the gravity potential, its gradient and the Marussi tensor through the Gauss-Legendre Quadrature (GLQ).
-It essentially consists in approximating the gravity fields of the tesseroids with the effect of point masses located on the scaled nodes of the Legendre polynomials.
-We state that this method it's well suited to approximate the effect of tesseroids with arbitrary densities, in contrast with the ones that involve Taylor series expansion.
-
-Instead of enhancing the accuracy of the method by increasing the order of the GLQ on each integration, we make use of the preexisting modified adaptive discretization algorithm and a new density-based discretization algodithm.
-
-The former one divides the tesseroid by half if the ratio of the distance to the computation point and its size is lower than a predefined distance-size ratio $D$.
-We can summarise that $D$ controls the accuracy of the method: a higher value of $D$ generates more divisions of the tesseroids, thus more point masses, what leads to a more precise approximation.
-Nevertheless it is not sufficient to guarantee the accuracy of the method in case of specific density functions as a highly variable exponential or some discontinuous functions.
-
-For this reason we developed a density-based discretization algorithm that divides the tesseroid on the radius at which the \emph{maximum variation} of the density function takes place, only if a certain inequation that relates this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
-Therefore the number of subdivisions is indirectly controlled by this delta ratio $\delta$, and thus the accuracy of the computation.
-
-Because there is no direct relation between the values assigned to $D$ and $\delta$ and the error of the computation, we had to empirically determine the minimum value of $D$ and the maximum value of $\delta$ that produce an acceptable accuracy for each gravity field.
-
-The distance-size ratio determinations have been made through the comparison of the numerical approximations with the analytical solutions for a spherical shell with variable density, which had to be analytically obtained.
-These tests have been made without applying the density-based discretization algorithm in order to isolate the former one.
-From the wide range of possible density variations, we've chosen a linear, an exponential and a discontinuous density functions.
-The comparisons over the first two functions have shown that values of $D$ equal to 1, 2 and 8 for the calculation of the gravity potential, its gradient and the Marussi tensor components, respectively, produce an accuracy of 0.1\% or better.
-On the other hand, the algorithm on its own does not guarantee this amount of accuracy in case of highly varying exponential functions or discontinuous densities.
-
-We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from low to high varying functions.
-Each comparison was made for different values of $\delta$, keeping the default values of $D$ previously obtained.
-This tests have shown that a value of $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while the same results can be obtained performing this tests with the discontinuous density function.
-The default value $\delta$ is the same for every gravity field computation.
-
-Moreover, we had performed speed comparisons between tesseroids models with homogeneous and variable densities, showing that a high varying density function requires more computation time than a low varying one, as expected.
-
-Finally, we have shown a real data example: we have computed the gravity fields generated by the Neuqu\'en sedimentary basin through a tesseroid model with exponential density with depth on a regular grid at 50km height.
+We have developed a new methodology that computes the gravitational generated 
+by a tesseroid with variable density in depth.
+It numerically solves the integrals that define the gravitational potential, 
+its gradient and the Marussi tensor through the Gauss-Legendre Quadrature 
+(GLQ), which essentially consists in approximating the gravitational fields of 
+the tesseroids with the effect of point masses located on the scaled nodes of 
+the Legendre polynomials.
+
+%~ We state that this method it's well suited to approximate the 
+%~ effect of tesseroids with arbitrary densities, in contrast with the ones that 
+%~ involve Taylor series expansion.
+
+Instead of enhancing the accuracy of the method by increasing the order of the 
+GLQ on each integration, we make use of the pre-existing modified adaptive 
+discretization algorithm and a new density-based discretization algorithm.
+
+The former one divides the tesseroid by half if the ratio of the distance to 
+the computation point and its size is lower than a predefined distance-size 
+ratio $D$.
+It's intended to minimise the error when the computation point is close to the 
+tesseroid.
+We can summarise that the distance-size ratio $D$ controls the accuracy of the 
+method: a higher value of $D$ generates more divisions of the tesseroids, thus 
+more point masses, what leads to a more precise approximation.
+Nevertheless, this algorithm is not sufficient to guarantee the accuracy of 
+the method in case of tesseroid with variable density.
+
+For this reason we have developed a density-based discretization algorithm that 
+divides the tesseroid on the radius at which the \emph{maximum variation} of 
+the density function takes place, only if a certain inequation that relates 
+this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
+Therefore the number of subdivisions is indirectly controlled by this delta 
+ratio $\delta$, and thus the accuracy of the computation.
+
+Because there is no direct relation between the values assigned to $D$ and 
+$\delta$ with the error of the computation, we had to empirically determine 
+the minimum value of $D$ and the maximum value of $\delta$ that produce an 
+acceptable accuracy for each gravitational field, while maintaining an 
+efficient computation.
+In order to determine default values of $D$ and $\delta$ we have chosen two 
+density functions: a linear and an exponential one.
+
+In the linear density case, the densty-based discretization algorithm is not 
+applied due to how it's build, so only a default value of $D$ had to be 
+determined.
+By comparing the analytical solutions of the gravitational fields generated by 
+a spherical shell with the ones produced by the numerical model for different 
+values of $D$, we have obtained default values for the distance-size ratio of 
+1, 2 and 8 for the potential, its gradient and the Marussi tensor components, 
+respectively.
+From these comparisons we can guarantee a precision of 0.1\% or greater on the 
+computations of the gravitational fields generated by a tesseroid with a 
+linear density.
+
+The exponential density case is more complex, due to the fact that both $D$ 
+and $\delta$ must be determined simultaneously.
+We have explored point in the $D$, $\delta$ space and computed the percentage 
+difference between the analytical solutions for a spherical shell and the 
+numerical model.
+We had to take into account that the $D$ values cannot be lower than the ones 
+determined for the linear density case due to the fact that they must 
+guarantee acceptable accuracy for every density function, including the linear 
+one.
+From these comparisons we have concluded that using the same $D$ values 
+obtained in the linear density scenario, a $\delta = 0.2$ is needed for the 
+computation of the potential and its gradient components, while a $\delta = 
+0.01$ must be used for the Marussi tensor components in order to ensure a 
+0.1\% precision.
+
+%~ This distance-size determinations have been carried out by comparing the 
+%~ analytical solutions of the gravitational fields generated by a spherical 
+%~ shell with a linear density with the gravitational fields generated by the 
+%~ numerical model for different values of $D$.
+%~ This comparisons were taken into account to obtain a percentage difference 
+%~ between the gravitational potential, the $g_z$ component of its gradient and 
+%~ the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
+%~ We have set the default value of density-based discretization as the minimum
+
+%~ The distance-size ratio determinations have been made through the comparison 
+%~ of the numerical approximations with the analytical solutions for a spherical 
+%~ shell with variable density, which had to be analytically obtained. These 
+%~ tests have been made without applying the density-based discretization 
+%~ algorithm in order to isolate the former one. From the wide range of possible 
+%~ density variations, we've chosen a linear, an exponential and a discontinuous 
+%~ density functions. The comparisons over the first two functions have shown 
+%~ that values of $D$ equal to 1, 2 and 8 for the calculation of the gravity 
+%~ potential, its gradient and the Marussi tensor components, respectively, 
+%~ produce an accuracy of 0.1\% or better. On the other hand, the algorithm on 
+%~ its own does not guarantee this amount of accuracy in case of highly varying 
+%~ exponential functions or discontinuous densities.
+
+%~ We have performed similar tests for the density-based discretization algorithm 
+%~ comparing the numerical model with the analytical solutions for a spherical 
+%~ shell with various exponential densities: from low to high varying functions. 
+%~ Each comparison was made for different values of $\delta$, keeping the default 
+%~ values of $D$ previously obtained. This tests have shown that a value of 
+%~ $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while 
+%~ the same results can be obtained performing this tests with the discontinuous 
+%~ density function. The default value $\delta$ is the same for every gravity 
+%~ field computation.
+
+%~ Moreover, we had performed speed comparisons between tesseroids models with 
+%~ homogeneous and variable densities, showing that a high varying density 
+%~ function requires more computation time than a low varying one, as expected.
+
+Finally, we have shown a real data example: we have computed the gravity 
+fields generated by the Neuqu\'en sedimentary basin through a tesseroid model 
+with exponential density with depth on a regular grid at 50km height.
 

### 2018-05-16 manuscript/manuscript.tex
Edited abstract

@@
 respectively.
-While a $\delta=0.2$ is needed in the computation of every gravitational 
-field, guaranteeing the accuracy even for high varying exponential densities.
+While a $\delta=0.2$ is needed for the computation of the gravitational 
+potential and its gradient components, and a $\delta=0.01$ must be used for 
+the Marussi tensor components in order to guarantee the accuracy even for high 
+varying exponential densities.
 We have finally tested the new software by modelling the Neuqu\'en sedimentary 

### 2018-05-16 manuscript/manuscript.tex
Removed todo note
Minor modification in abstract

--- 2018-05-16 manuscript/manuscript.tex
@@
 respectively.
-While a $\delta=0.2$ is needed in the computation of every gravitational 
-field, guaranteeing the accuracy even for high varying exponential densities.
+While a $\delta=0.2$ is needed for the computation of the gravitational 
+potential and its gradient components and a $\delta=0.01$ must be used for 
+the Marussi tensor components in order to guarantee the accuracy even for high 
+varying exponential densities.
 We have finally tested the new software by modelling the Neuqu\'en sedimentary 
@@
 
-\todo[inline, color=red]{Text bellow is old}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-05-17 manuscript/manuscript.tex
Modifications in conclusions

@@
 
-We have developed a new methodology that computes the gravitational generated 
-by a tesseroid with variable density in depth.
+We have developed a new methodology that computes the gravitational fields 
+generated by a tesseroid with continuously variable density in depth.
 It numerically solves the integrals that define the gravitational potential, 
-its gradient and the Marussi tensor through the Gauss-Legendre Quadrature 
-(GLQ), which essentially consists in approximating the gravitational fields of 
-the tesseroids with the effect of point masses located on the scaled nodes of 
-the Legendre polynomials.
+its gradient and the Marussi tensor components through the Gauss-Legendre 
+Quadrature (GLQ), which essentially consists in approximating the 
+gravitational fields of the tesseroids with the effect of point masses located 
+on nodes of the Quadrature.
 
@@
 ratio $D$.
-It's intended to minimise the error when the computation point is close to the 
-tesseroid.
-We can summarise that the distance-size ratio $D$ controls the accuracy of the 
-method: a higher value of $D$ generates more divisions of the tesseroids, thus 
-more point masses, what leads to a more precise approximation.
+This algorithm is introduced in order to minimise the error when the 
+computation point is close to the tesseroid, in which case the effect of the 
+point masses becomes more noticeable and the approximation lose precision.
+Following this idea, we can summarise that the distance-size ratio $D$ 
+controls the accuracy of the method: a higher value of $D$ generates more 
+divisions of the tesseroids, thus more point masses, what leads to a more 
+precise approximation.
 Nevertheless, this algorithm is not sufficient to guarantee the accuracy of 
@@
 this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
-Therefore the number of subdivisions is indirectly controlled by this delta 
+The the number of subdivisions is indirectly controlled by this delta 
 ratio $\delta$, and thus the accuracy of the computation.
+This new algorithm is intended to minimise the error due to the inability of 
+the GLQ to produce precise approximations in case of highly varying density: 
+the information about the density function is summed up with the values that 
+it assumes at the node of the quadrature.
+Increasing the subdivisions of the tesseroid where the \emph{maximum 
+variation} of the density takes place becomes an efficient way to reduce this 
+kind of error.
 
@@
 efficient computation.
-In order to determine default values of $D$ and $\delta$ we have chosen two 
-density functions: a linear and an exponential one.
+The default values of $D$ and $\delta$ must produce accurate computations for 
+most continuous and smooth densities, that's why we have chosen two density 
+functions in order to establish them: a linear and an exponential one.
 
@@
 and $\delta$ must be determined simultaneously.
-We have explored point in the $D$, $\delta$ space and computed the percentage 
-difference between the analytical solutions for a spherical shell and the 
-numerical model.
+We have explored regular gridded points in the $D$, $\delta$ space and 
+computed the percentage difference between the analytical solutions for a 
+spherical shell and the numerical model.
 We had to take into account that the $D$ values cannot be lower than the ones 
 determined for the linear density case due to the fact that they must 
-guarantee acceptable accuracy for every density function, including the linear 
-one.
-From these comparisons we have concluded that using the same $D$ values 
-obtained in the linear density scenario, a $\delta = 0.2$ is needed for the 
-computation of the potential and its gradient components, while a $\delta = 
-0.01$ must be used for the Marussi tensor components in order to ensure a 
-0.1\% precision.
+guarantee acceptable accuracy for every continuous density function, including 
+the linear one.
+From these comparisons we have concluded that the more efficient computations 
+take places when the $D$ values obtained in the linear density scenario are 
+used, i.e. the minimum ones.
+This can be understood in terms of which kind of discretization produces more 
+subdivisions: a higher $D$ may produce more discretizations than a lower 
+$\delta$ because the first one divides the tesseroid in the radial, longitude 
+and latitude directions, while the latter only divides it in the radial 
+direction.
+In conclusion, we found out that keeping the $D$ values as the ones obtained 
+for the linear density case, a $\delta = 0.2$ is needed for the computation of 
+the potential and its gradient components, while a $\delta = 0.01$ must be 
+used for the Marussi tensor components in order to ensure a 0.1\% precision.
 
@@
 
-Finally, we have shown a real data example: we have computed the gravity 
-fields generated by the Neuqu\'en sedimentary basin through a tesseroid model 
-with exponential density with depth on a regular grid at 50km height.
+Finally, we have shown how this new method performs under a real data example: 
+we have computed the gravitational fields generated by the Neuqu\'en 
+sedimentary basin through a tesseroid model with an increasing exponential 
+density in depth on a regular grid at 50km height.
 

### 2018-05-17 code/tests/test_variable_density.py
Edited tests/test_variable_density.py: changed str(tess) for str(tess.get_bounds())
The Tesseroid class doesn't handle density function when __print__

@@ -282,7 +282,8 @@ def test_laplace_equation():
         trace = gxx + gyy + gzz
         assert_array_almost_equal(trace, np.zeros_like(lon), 9,
                                   'Failed tesseroid %s. Max diff %.15g'
-                                  % (str(tess), np.abs(trace).max()))
+                                  % (str(tess.get_bounds()),
+                                  np.abs(trace).max()))
 
 
 def test_stack_overflow():

### 2018-05-17 code/tests/test_exponential_density.py
Added test_exponential_density.py

@@ -1,0 +1,121 @@
+from __future__ import division, absolute_import
+import numpy as np
+from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
+from fatiando.mesher import TesseroidMesh
+from fatiando import gridder
+
+from tesseroid_density import tesseroid
+
+
+def shell_exponential_density(height, top, bottom, a, b, c, deltah):
+    r = height + MEAN_EARTH_RADIUS
+    r1 = bottom + MEAN_EARTH_RADIUS
+    r2 = top + MEAN_EARTH_RADIUS
+    constant = 4*np.pi*G*a*b*((r1**2 + 2*r1*b + 2*b**2)*np.exp(-(r1 - deltah)/b) -
+                              (r2**2 + 2*r2*b + 2*b**2)*np.exp(-(r2 - deltah)/b))
+    constant += 4/3*np.pi*G*c*(r2**3 - r1**3)
+    potential = constant/r
+    data = {'potential': potential,
+            'gx': 0,
+            'gy': 0,
+            'gz': SI2MGAL*(potential/r),
+            'gxx': SI2EOTVOS*(-potential/r**2),
+            'gxy': 0,
+            'gxz': 0,
+            'gyy': SI2EOTVOS*(-potential/r**2),
+            'gyz': 0,
+            'gzz': SI2EOTVOS*(2*potential/r**2)}
+    return data
+
+
+def test_thin_shell():
+
+    def density_exponential(height):
+        r = height + MEAN_EARTH_RADIUS
+        return a*np.exp(-(r - deltah)/b) + c
+
+    # Spherical shell model with a Tesseroid Mesh
+    top, bottom = 0, -1000
+    density_1, density_2 = 2670, 3300
+    b = 50.
+    a = (density_2 - density_1)/(np.exp((abs(top - bottom))/b) - 1)
+    c = density_1 - a
+    deltah = MEAN_EARTH_RADIUS
+    model = TesseroidMesh((0, 360, -90, 90, top, bottom), (1, 6, 12))
+    model.addprop("density", [density_exponential for i in range(model.size)])
+
+    # Computation grids
+    shape = (10, 10)
+    grids = {"pole": gridder.regular((89, 90, 0, 1), shape, z=2e3),
+             "equator": gridder.regular((0, 1, 0, 1), shape, z=2e3),
+             "260km": gridder.regular((89, 90, 0, 1), shape, z=260e3),
[…]

### 2018-05-24 manuscript/manuscript.tex
Removed old commented lines in manuscript
Changes in abstract

@@
 direction.
-It's based on the numerical approximation of the gravitational fields by 
-applying the Gauss-Legendre Quadrature along with a preexisting adaptive 
-discretization algorithm and a new density-based discretization algorithm that 
-control both the accuracy and the computation time of the forward model.
-The first algorithm is regulated by a predefined scalar called distance-size 
-ratio ($D$) and reduces the errors due to the distance between the tesseroid 
-and the computation point.
-While the second one, controlled by a delta ratio ($\delta$), decreases the 
-errors introduced by the variation of the density function.
+It numerical approximates these gravitational fields through the 
+Gauss-Legendre Quadrature along with two discretization algorithms that 
+increase its accuracy by subdividing the tesseroid into smaller ones.
+The first one is a preexisting adaptive discretization algorithm that reduces 
+the errors due to the distance between the tesseroid and the computation 
+point, while being regulated by a predefined scalar called distance-size ratio 
+($D)$.
+The second one, controlled by a predefined delta ratio ($\delta$), is a new 
+density-based discretization algorithm that decreases the errors introduced by 
+the variation of the density function.
 We have also obtained analytical solutions for a spherical shell with variable 
@@
 varying exponential densities.
-We have finally tested the new software by modelling the Neuqu\'en sedimentary 
-basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S 
-latitude, assigning an exponential density variation to it and computing all 
-the gravitational fields it generates.
+We have finally tested this new methodology by modelling the Neuqu\'en 
+sedimentary basin, located to the east of the Andes between 32$^\circ$S and 
+40$^\circ$S latitude, assigning an exponential density variation to it and 
+computing all the gravitational fields it generates.
 \end{summary}
@@
 
-%~ It worth mention that either a greater $D$ or a lower $\delta$ 
-%~ increases the number of discretizations of the tesseroid.
-%~ Although a greater $D$ may divide it in the latitude, longitude 
-%~ and/or radial direction; a lower $\delta$ will only discretize the 
-%~ original tesseroid in the radial direction.
-%~ This means that a decrease in $\delta$ will produce a more efficient 
-%~ computation than an increase in $D$.
-%~ \todo[inline]{Should we keep this paragraph? The same is explained bellow.}
-
 Every point in the subregion with $\delta \le \delta_p$ (region B on 
@@
 numerical error is bellow the $0.1\%$ threshold.
-
-%~ In order to determine the value of $\delta_p$ we should compute the 
-%~ differences between the numerical model and the analytical solution for 
-%~ a spherical shell with an exponential density.
-%~ Thus we must obtain this analytical solution and define a tesseroid 
-%~ model to numerically approximate the gravitational fields.
-
-%~ Secondly, we proposed a thin and a thick spherical shells with 1km and 
-%~ 35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
-%~ 30^\circ$ tesseroids.
-%~ We have assigned the exponential density function proposed on equation 
-%~ \ref{eq:density-exp} to each spherical shell model, choosing some values 
-%~ of the constant $b$ in order to test the numerical approximation for 
-%~ different density variations.
-
-%~ Finally, we computed the gravitational potential ($V$), the 
-%~ vertical component of its gradient ($g_z$) and the $g_zz$ component of 
-%~ the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
-%~ keeping the distance-size ratio equal to the $D$ values obtained in the 
-%~ density case for its corresponding gravitational field and exploring 
-%~ different values of $\delta$.
-%~ We then computed the percentage difference between this numerical 
-%~ results and the analytical solutions.
 
@@
 
-%~ We state that this method it's well suited to approximate the 
-%~ effect of tesseroids with arbitrary densities, in contrast with the ones that 
-%~ involve Taylor series expansion.
-
 Instead of enhancing the accuracy of the method by increasing the order of the 
@@
 
-%~ This distance-size determinations have been carried out by comparing the 
-%~ analytical solutions of the gravitational fields generated by a spherical 
-%~ shell with a linear density with the gravitational fields generated by the 
-%~ numerical model for different values of $D$.
-%~ This comparisons were taken into account to obtain a percentage difference 
-%~ between the gravitational potential, the $g_z$ component of its gradient and 
-%~ the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
-%~ We have set the default value of density-based discretization as the minimum
-
-%~ The distance-size ratio determinations have been made through the comparison 
-%~ of the numerical approximations with the analytical solutions for a spherical 
-%~ shell with variable density, which had to be analytically obtained. These 
-%~ tests have been made without applying the density-based discretization 
-%~ algorithm in order to isolate the former one. From the wide range of possible 
-%~ density variations, we've chosen a linear, an exponential and a discontinuous 
-%~ density functions. The comparisons over the first two functions have shown 
-%~ that values of $D$ equal to 1, 2 and 8 for the calculation of the gravity 
-%~ potential, its gradient and the Marussi tensor components, respectively, 
-%~ produce an accuracy of 0.1\% or better. On the other hand, the algorithm on 
-%~ its own does not guarantee this amount of accuracy in case of highly varying 
-%~ exponential functions or discontinuous densities.
-
-%~ We have performed similar tests for the density-based discretization algorithm 
-%~ comparing the numerical model with the analytical solutions for a spherical 
-%~ shell with various exponential densities: from low to high varying functions. 
-%~ Each comparison was made for different values of $\delta$, keeping the default 
-%~ values of $D$ previously obtained. This tests have shown that a value of 
-%~ $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while 
-%~ the same results can be obtained performing this tests with the discontinuous 
-%~ density function. The default value $\delta$ is the same for every gravity 
-%~ field computation.
-
-%~ Moreover, we had performed speed comparisons between tesseroids models with 
-%~ homogeneous and variable densities, showing that a high varying density 
-%~ function requires more computation time than a low varying one, as expected.
-
 Finally, we have shown how this new method performs under a real data example: 

### 2018-05-24 manuscript/manuscript.tex
More changes in abstract

@@
 the errors due to the distance between the tesseroid and the computation 
-point, while being regulated by a predefined scalar called distance-size ratio 
-($D)$.
-The second one, controlled by a predefined delta ratio ($\delta$), is a new 
-density-based discretization algorithm that decreases the errors introduced by 
-the variation of the density function.
+point.
+While the second one  is a new density-based discretization algorithm that 
+decreases the errors introduced by the variation of the density function.
+The amount of discretizations made by each algorithm is indirectly controlled 
+by two predefined scalar values: the distance-size ratio $D$ for the former 
+and the delta ratio $\delta$ for the latter.
 We have also obtained analytical solutions for a spherical shell with variable 
@@
 potential and its gradient components and a $\delta=0.01$ must be used for 
-the Marussi tensor components in order to guarantee the accuracy even for high 
-varying exponential densities.
+the Marussi tensor components.
 We have finally tested this new methodology by modelling the Neuqu\'en 

### 2018-05-24 manuscript/manuscript.tex
Minor changes in Introduction

@@
 Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
-ensures the accuracy implementing a modified version of the adaptive 
+ensures the accuracy through a modified version of the adaptive 
 discretization of \citet{Li2011}.
@@
 which constitutes a special case of the volume integrals that has 
-analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013} 
-\todo[inline]{citar mas??}.
+analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}.
+\todo[inline]{citar mas??}
 
@@
 In the following sections we will describe how the new algorithm works, its 
-theoretical background and obtain default distance-size ratio and delta ratio 
-values needed to get an acceptable accuracy in the numerical approximation.
+theoretical background, obtain default distance-size ratio and delta ratio 
+values needed to get a numerical approximation with acceptable accuracy and 
+finally show a simple application of the methodology by modelling the Nequen 
+basin through a set of tesseroids with exponential density.
 

### 2018-05-24 manuscript/manuscript.tex
Minor changes in Theory

@@
 
-From equation \ref{eq:glq-var-dens} we can observe that the GLQ 
-approximates the gravitational fields of a tesseroid as the ones 
-generated by $N_\lambda N_\phi N_r$ point masses located at the nodes 
-of the Legendre polynomials, where $N_i$ are the orders of the 
-quadrature for each integration, with $i \in \{ \lambda, \phi, r \}$, 
-in the same way as the GLQ works in the homogeneous density case.
+From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates 
+the gravitational fields of a tesseroid as the ones generated by $N_\lambda 
+N_\phi N_r$ point masses located at the nodes of the Legendre polynomials, in 
+the same way as the GLQ works in the homogeneous density case, where $N_i$ ($i 
+\in \{ \lambda, \phi, r \}$) are the orders of the quadrature for each 
+integration.
 Nevertheless, in this new variable density scenario, the information of 

### 2018-05-24 manuscript/manuscript.tex
Minor changes in Methodology

@@
 the variation of the density function.
-
 The former one is also present in the homogeneous density tesseroid 
@@
 $R_1$ and $R_2$ and a density function $\rho(r')$, whose radial 
-dimension can be written as $L_r = R_2 - R_1$.
+dimension is $L_r = R_2 - R_1$.
 
@@
 \begin{equation}
-    \rho_n(r') = \frac{\rho(r') - \rho_1}{\rho_2 - \rho_1},
+    \rho_n(r') =
+    \frac{\rho(r') - \rho_\text{min}}{\rho_\text{max} - \rho_\text{min}},
 \end{equation}
@@
 \begin{equation}
-    \rho_1 = \text{min}\{ \rho(r') \}, \quad
-    \rho_2 = \text{max}\{ \rho(r') \} \quad
+    \rho_\text{min} = \text{min}\{ \rho(r') \}, \quad
+    \rho_\text{max} = \text{max}\{ \rho(r') \} \quad
     \forall \, r' \in [R_1, R_2].
@@
 
-\noindent where the delta ratio $\delta$ is a predefined constant value 
-and $L_r^\text{sm}/L_r$ is defined as the tesseroid size ratio.
-Besides, $L_r^\text{sm}$ and $L_r$ are the radial dimensions of the 
-tesseroid that is currently analysed and of the \emph{original} one, 
-respectively.
+\noindent where the delta ratio $\delta$ is a predefined constant value and 
+$L_r^\text{sm}/L_r$ is defined as the tesseroid size ratio, while 
+$L_r^\text{sm}$ and $L_r$ are the radial dimensions of the tesseroid that is 
+currently analysed and of the \emph{original} one, respectively.
 On the first iteration, the former tesseroid corresponds to the 
@@
 future iterations this radial dimension will be computed as the 
-difference of the top and bottom radii of the subdividing tesseroid.
-The tesseroid size ratio is intended to reduce the number of probable 
-unneeded discretizations and focus on discretizations that will 
-significantly reduce the numerical error.
-A big subdividing tesseroid with small maximum absolute difference would 
-contribute more numerical error than a small subdividing tesseroid with 
+difference of the top and bottom radii of the subdividing tesseroid, making 
+$L_r^\text{sm}/L_r$ lower than one.
+The tesseroid size ratio is intended to reduce the number of unneeded 
+discretizations and focus on the ones that will significantly reduce the 
+numerical error:
+a big subdividing tesseroid with small maximum absolute difference would 
+cause more relative numerical error than a small subdividing tesseroid with 
 higher maximum absolute difference.
 
-Once the \emph{original} tesseroid is divided, we proceed to 
+Once the \emph{original} tesseroid is discretized, we proceed to 
 subsequently apply the same algorithm on every subdividing tesseroid.
@@
 If the inequation holds, the current tesseroid is discretized and the 
-algorithm is then applied to the resulting subdividing tesseroids.
+algorithm is subsequently applied to the resulting subdividing tesseroids.
 
 If on the contrary the inequation \ref{eq:delta-density} is not held, 
-then the tesseroid will not be subdivided.
+then the tesseroid will not be discretized.
 The same happens if the tesseroid radial dimension is bellow the 
@@
 every subdividing tesseroid is subjected to the modified adaptive 
-discretization algorithm and then a second order GLQ must be applied in 
+discretization algorithm and then a second order GLQ will be applied in 
 order to compute the gravitational field generated by it.

### 2018-05-24 manuscript/manuscript.tex
Modifications in Determinations of D and delta

@@
 algorithm makes use of the delta ratio $\delta$.
-These algorithms need default predefined values for $D$ and $\delta$ 
+These algorithms need predefined default values for $D$ and $\delta$ 
 in order to ensure both acceptable numerical accuracy and computation 
@@
 
-\noindent where $\ell$ is defined in equation \ref{eq:ell}.
-The computation point $Q$ is located at latitude $\phi=90^\circ$, so 
-$\ell$ and $\cos\psi$ (defined in equation \ref{eq:cospsi}) get 
-simplified:
+\noindent where $\ell$ is defined in equation \ref{eq:ell}. The computation 
+point $Q$ is located at latitude $\phi=90^\circ$, so $\ell$ and $\cos\psi$ 
+(defined on equations \ref{eq:ell} and \ref{eq:cospsi}) get simplified:
 
@@
 
-Firstly we need to solve the integral on equation \ref{eq:shell-pot} in 
-order to obtain the analytical solution in this case.
-Thus, the gravitational potential generated by a spherical shell with this 
-density on any external point at distance $r$ from its centre can be 
-obtained from the following expression:
-\todo[inline]{revisar redaccion}
+First, by solving the integral on equation \ref{eq:shell-pot}, we obtained the 
+gravitational potential generated by a spherical shell with this density on 
+any external point at distance $r$ from its centre:
 
@@
 
-\noindent The first term on equation \ref{eq:shell-pot-linear} 
-reproduces the potential generated by a spherical shell with variable 
-density $\rho(r') = ar'$, while the second one constitutes the 
-potential generated by a spherical shell with homogeneous density $\rho 
-= b$ \citep{Mikuska2006,Grombein2013}.
+\noindent The first term on this equation reproduces the potential generated 
+by a spherical shell with variable density $\rho(r') = ar'$, while the second 
+one constitutes the potential generated by a spherical shell with homogeneous 
+density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
 
@@
 In order to compare the numerical model with the analytical solution we 
-must build a spherical shell from a set of tesseroids.
-We propose a mesh of $30^\circ \times 30^\circ$ tesseroids to compute 
+must build a spherical shell made of tesseroids.
+We propose a mesh of $30^\circ \times 30^\circ$ tesseroids whose outer and 
+inner radii are the same ones as the spherical shell in order to compute 
 the gravitational potential, the vertical component of the gradient 
@@
     ratio $D$.
-    Due to the linearity of the density function the density-based 
+    Due to the linearity of the density function, the density-based 
     discretization algorithm is not applied.
@@
 $\delta$ must be simultaneously determined.
+
+\subsubsection{$D$, $\delta$ space exploration}
 
@@
 increases the maximum slope of the density function.
-
-The analytical solution of the gravitational fields generated 
-by a spherical shell with a density functions as the one in equation 
-\ref{eq:density-exp} can be computed as the sum of the fields generated 
-by a spherical shell with constant density $\rho(r') = C$ and another 
-shell with an exponential density like:
-
-\begin{equation}
-    \rho(r') = Ae^{-(r' - \Delta h)/b},
-\end{equation}
-
-\noindent where $A$, $\Delta h$ and $b$ are constants.
-The gravitational potential field generated by a spherical shell with 
-the latter density on any external point at a distance $r$ from its centre 
-can be calculated by solving the integral in equation \ref{eq:shell-pot},
-resulting as follows:
+The analytical solution of the gravitational potential generated by a 
+spherical shell with a density functions as the one in equation 
+\ref{eq:density-exp} can be obtained by solving the equation 
+\ref{eq:shell-pot} as follows:
 
@@
         & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
-        \Big].
+        \Big] + \\
+        & \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
     \end{split}
@@
     \Big[
-    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
+    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} -
     (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    \Big].
+    \Big] +
+    \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
 \label{eq:shell-pot-exp}
@@
 numerical error lower than the 0.1\% threshold.
-The one that produces the more efficient computation must be chosen as 
-the default $D$ and $\delta$ value.
+Our intention is to choose the one that produces the most efficient computation
+and set it as the default values for $D$ and $\delta$.
 
@@
 
-Because the algorithm is intended to work with default $D$ and $\delta$ 
+Because the algorithm is thought to work with default $D$ and $\delta$ 
 values that ensures these two conditions for every density function, 
@@
 $D_\text{linear}$ and its delta ratio is $\delta_p$.
+
+\subsubsection{Delta ratio determination}
 

### 2018-05-24 manuscript/manuscript.tex
Minor changes on Implementation
Modifications in Example with real data

@@
 algorithm.
-On their implementation, \citet{Uieda2016} have optimised the more time 
+On their software, \citet{Uieda2016} have optimised the more time 
 consuming functions through the Numba compiler.
@@
 
-Besides all the tests we have carried out on the new code, we intend to show a 
-forward modelling example with real data.
-
 Besides the determination of the distance-size and the delta ratio, we 
-wanted to show how the algorithm performs under a real data forward model.
-In order to do it we have chosen the Neuqu\'en Basin: a sedimentary basin 
+wanted to illustrate how the algorithm performs under a real data model.
+For this task we have chosen the Neuqu\'en Basin: a sedimentary basin 
 located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude 
@@
 longitude and latitude directions with one value of thickness on each node 
-(see Figure \ref{fig:neuquen-basin}(b).
+(see Figure \ref{fig:neuquen-basin}(b)).
 
@@
 We must also define a density function for the entire model. 
-\citet{Sigismondi2012} proposed a minimum and maximum density contrast for the 
-Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
+\citet{Sigismondi2012} measured a minimum and maximum density contrast for 
+the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
 So we have chosen, for this specific example, an exponential density (like the 

### 2018-05-24 manuscript/manuscript.tex
Modifications in Conclusions

@@
 this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
-The the number of subdivisions is indirectly controlled by this delta 
+The number of subdivisions is indirectly controlled by this delta 
 ratio $\delta$, and thus the accuracy of the computation.
@@
 From these comparisons we can guarantee a precision of 0.1\% or greater on the 
-computations of the gravitational fields generated by a tesseroid with a 
+computations of the gravitational fields generated by any tesseroid with a 
 linear density.
@@
 
+In order to implement this new methodology we have developed a Python code 
+that carries out both discretization algorithms and the gravitational field 
+computation through the GLQ. It's freely available under the BSD 3-clause 
+open-source license and can be downloaded from
+\todo[inline]{add repo url}
+
 Finally, we have shown how this new method performs under a real data example: 
@@
 sedimentary basin through a tesseroid model with an increasing exponential 
-density in depth on a regular grid at 50km height.
+density in depth on a regular grid at a computation height of 50km over the 
+reference ellipsoid.
 

### 2018-06-22 manuscript/manuscript.tex

@@
 two subregions: one with $D < D_\text{linear}$ and other with $D \ge 
-D_\text{linear}$. We can ignore points belonging to the former one 
-(region A on Figure \ref{fig:D-delta-grid-search}) and focus on the 
-latter (regions B and C on Figure \ref{fig:D-delta-grid-search}).
+D_\text{linear}$.
+We can ignore points belonging to the former one (region A on Figure 
+\ref{fig:D-delta-grid-search}) and focus on the latter (regions B and C on 
+Figure \ref{fig:D-delta-grid-search}).
 

### 2018-09-19 manuscript/manuscript.tex
Added Wild-Pfeiffer2008 reference in manuscript
Reviewed the text until Adaptive Discretization
Mostly simplified the wording, cut repeated text, and removed some
sections that are well explained in the references. There is enough new
material here that we don't need to go into details of the GLQ, which
has been extensively covered elsewhere.

@@
     $^3$Department of Geology and Geophysics, SOEST, University of Hawai‘i at
-    Manoa, Honolulu, Hawaii, USA
+    M\={a}noa, Honolulu, Hawaii, USA
 }
@@
 
-Lithosphere's density variation with depth has been studied for almost a 
+The lithosphere's density variation with depth has been studied for close to a
 century.
-On the first years \citet{Athy1930} obtained a decreasing exponential relation 
-between both quantities by studying shale samples, while in the following 
-decades other density functions have been proposed for different rock types 
+Over this time period,
+several density-depth relations have been proposed for different rock types
 \citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
-Furthermore, varying densities have been taken into account in forward and 
-inversion gravity models, mostly applied to grabens and basins 
+Furthermore, depth-variable densities have been used in forward and
+inverse gravity modeling, mostly applied to sedimentary basins
 \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001,
 Welford2010}.
-
 These forward gravity models have been developed for two or three dimensional
-bodies in cartesian coordinates that properly fit local scales applications. 
-Nevertheless, the latest satellite missions have provided us gravity 
-measurements with global coverage, which allow geologists and geophysics to 
-perform modelling and interpretation on regional scales. This raises the 
-importance of building forward models that reproduce the gravity anomalies for 
-such scales.
-
-The main issue that should be overcome is to take into account the curvature 
-of the Earth, thus the forward model must be defined in spherical coordinates. 
-A common way to achieve this is to discretize the Earth in spherical prisms 
-known as tesseroids, which are defined by pairs of latitude, longitude and 
-radial boundaries (see Figure \ref{fig:tesseroid-uieda}).
-The calculation of the gravitational fields generated by an arbitrary 
-tesseroid on any external point involves the resolution of volume 
-integrals that are generally approximated by numerical computations.
-The literature offers two approaches: one involves Taylor series expansion 
+bodies in Cartesian coordinates, which limits the applications to local scales.
+The advent of satellite gravimetry has provided gravity field
+measurements with global coverage, enabling modelling and interpretation on regional and
+global scales.
+Hence, designing forward modeling methods that reproduce the gravity anomalies for
+such scales is of high importance.
+
+To take into account the curvature of the Earth, many global forward modeling methods
+are defined in geocentric spherical coordinates.
+A common approach is to discretize the Earth into tesseroids (spherical prisms),
+which are defined by pairs of latitude, longitude, and
+radial boundaries (see Fig.~\ref{fig:tesseroid}).
+The gravitational fields generated by an arbitrary
+tesseroid on any external point are given by volume
+integrals that are generally approximated by numerically.
+The literature offers two main approaches: one involves Taylor series expansion
 \citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre
-Quadrature (GLQ) \citep{Asgharzadeh2007, Uieda2016, Uieda2017}.
-
-In order to develop a forward model that computes the gravitational 
-field generated by any tesseroid with arbitrary variable density, the 
-Taylor series expansion is not well suited, because it would need the 
-series expansion terms for each density function.
-On the other hand, the GLQ consists in approximating the integral by a 
-weighted sum of the effect of point masses located at scaled nodes of the 
-Legendre polynomials, where the density variation is summarised into the 
-weights.
-Thus, this kind of numerical approximation allows to calculate the 
-gravitational fields for any density function without any other 
-information but the function itself.
+Quadrature (GLQ)
+\citep{Asgharzadeh2007, Wild-Pfeiffer2008, Li2011, Uieda2016}.
+The Taylor series expansion is not well suited to develop an algorithm for
+a density varying with depth according to an arbitrary function.
+Different series expansion terms would have to obtained for each density function
+desired.
+Conversely, an arbitrary density function can be included in the GLQ without
+any change to the integration method.
 
@@
 \caption{
-    A spherical prism, also known as  Tesseroid, in a geocentric spherical 
-    coordinate system, with a computation point $P$ and its local north 
-    oriented Cartesian coordinate system. Picture made by \citet{Uieda2015}.
+    A tesseroid (spherical prism) in a geocentric spherical coordinate system, with a
+    computation point $P$ and its local north oriented Cartesian coordinate system.
+    After \citet{Uieda2015}.
 }
-\label{fig:tesseroid-uieda}
+\label{fig:tesseroid}
 \end{figure}
 
-\citet{Uieda2016} developed a forward gravitational model based on 
-tesseroids with homogeneous densities using the GLQ approximation 
-method.
-The main issue of the GLQ is that the approximation becomes less 
-accurate when the computation point is closer to the tesseroid or for 
-smaller GLQ order \citep{Ku1977}.
-Although the model developed by \citet{Uieda2016} uses a second order GLQ, it 
-ensures the accuracy through a modified version of the adaptive 
-discretization of \citet{Li2011}.
-It consist in splitting the tesseroid into smaller ones when a certain 
-distance-size ratio is lower than a predefined value ($D$) that controls the 
-accuracy of the computation.
-\citet{Uieda2016} have also objectively obtained standard values of $D$ 
-for gravitational potential, gradients and tensor components comparing 
-the numerical model with the fields generated by a spherical shell, 
-which constitutes a special case of the volume integrals that has 
-analytical solution \citep{LaFehr1991, Mikuska2006, Grombein2013}.
-\todo[inline]{citar mas??}
-
-On the other hand, a varying density inside an arbitrary tesseroid 
-introduces a new kind of error into the numerical approximation.
-This encouraged us to develop a new density-based discretization 
-algorithm that divides the tesseroid into smaller ones taking into 
-account the variation of the density function.
-The number of subdivisions and thus the accuracy of the approximation 
-will be controlled by predefined delta ratio $\delta$.
-
-In summary, we present a new forward gravitational model that allows to 
-compute the gravitational fields generated by any tesseroid with an 
-arbitrary continuous density function on any external point.
-It is based on the GLQ approximation whose accuracy is controlled by 
-the modified distance-size discretization \citep{Uieda2016} 
-and the new density-based discretization algorithms.
-
+The main challenge of the GLQ integration is loss of accuracy when the computation point
+gets closer to the tesseroid \citep{Ku1977}.
+\citet{Uieda2016} built on the adaptive discretization algorithm of \citet{Li2011}
+to automatically obtain integration results with 0.1\% accuracy.
+The algorithm consist in recursively splitting the tesseroid into smaller ones when a
+threshold is exceeded,
+namely when the normalized distance to the computation point is greater than a
+``distance-size ratio'' ($D$).
+\citet{Uieda2016} have also obtained standard values of $D$
+for the gravitational potential, acceleration, and gradient tensor components
+by comparing the numerical model with the fields generated by a spherical shell.
+
+We present a new algorithm for computing the gravitational fields generated by any
+tesseroid with an arbitrary continuous density function on any external point.
+It is based on the GLQ approximation and the adaptive discretization of
+\citep{Uieda2016}, which we extend to include a new density-based discretization step.
 In order to ensure the accuracy of the numerical approximation we have
-determined predefined values for the $D$ and $\delta$ parameters by 
-comparing the numerical approximation with the analytical solutions for 
-a spherical shell with linear and exponential density, which had to be 
-derived.
-
-Furthermore, we developed a Python library that implements this forward 
-model.
-I makes use of some previously existing classes and functions 
-from the Python library Fatiando a Terra for geophysical modelling, 
-what saved us time and allowed us to focus on the new model.
-
-Furthermore, we have the intention to include it into a future release 
-of the Fatiando a Terra library in order to make its distribution and 
-maintenance easier.
-\todo[inline]{include this?}
-
-In the following sections we will describe how the new algorithm works, its 
-theoretical background, obtain default distance-size ratio and delta ratio 
-values needed to get a numerical approximation with acceptable accuracy and 
-finally show a simple application of the methodology by modelling the Nequen 
-basin through a set of tesseroids with exponential density.
+determined optimal values for the controlling parameters by
+comparing the numerical approximation with analytical solutions for
+spherical shells.
+Finally, we applied the methodology to model the Nequen basin, Argentina, using
+tesseroids with an exponentially increasing density with depth.
 
@@
 
-\section{Theory}
-
-The spherical prisms known as tesseroids are mass elements defined in a 
-geocentric spherical coordinate system bounded by a pair of parallels, 
-a pair of  meridians, and two concentric spherical surfaces.
-We define an external computation point $P(r, \phi, \lambda)$ where the 
-gravitational fields generated by the tesseroid are going to be calculated 
-with respect to the local north oriented Cartesian coordinate system 
-located at $P$.
-In the special case of homogeneous density, the gravity potential, its 
-gradient and the Marussi tensor components can be calculated through 
-the integrals obtained by \citet{Grombein2013} \citep[for same notation 
-as the one we will use, see][]{Uieda2016}.
-
-In our case, we will assume that the tesseroid has an arbitrary 
-variable density in depth, i.e. only depends on the radial spherical 
-coordinate.
-Thus, the integrals for the gravitational fields are slightly modified:
+\section{Methodology}
+
+We define an external computation point $P(r, \phi, \lambda)$ in a geocentric spherical
+coordinate system at a radius $r$, geocentric latitude $\phi$, and longitude $\lambda$
+where the gravitational fields are going to be calculated.
+The first and second derivatives of the gravitational potential are taken with
+respect to the local north-oriented Cartesian coordinate system of $P$
+(Fig.~\ref{fig:tesseroid}).
+\citet{Grombein2013} provide efficient formulations for the volume integrals of the
+gravitational potential and its first and second derivatives of a tesseroid with
+homogeneous density.
+Here, we will assume that the tesseroid has a density varying with $r$ according to an
+arbitrary function $\rho(r)$.
+Thus, the integrals for the gravitational fields are slightly modified to
 
@@
 \end{equation}
+
 \begin{equation}
@@
 \end{equation}
+
+\noindent and
+
 \begin{equation}
@@
 
-\noindent where
+\noindent in which
 
@@
 
-\noindent $\alpha, \beta \in \{x, y, z\}$, $\rho(r')$ is the density 
-function that depends on the radial coordinate, $\delta_{\alpha\beta}$ 
-is Kronecker's delta, $G = 6.674\times10^{-11}\, 
-\text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant and
+\noindent $\alpha, \beta \in \{x, y, z\}$, $\delta_{\alpha\beta}$ is Kronecker's delta,
+$G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant
+and
 
@@
 
-According to \citet[p.~390]{Hildebrand1987}, we can approximate each 
-integral on equations \ref{eq:tesseroid-pot}, \ref{eq:tesseroid-grav} 
-and \ref{eq:tesseroid-tensor} by applying a $N$th order GLQ, i.e. as 
-a weighted sum of the integration kernel evaluated on the roots of the 
-$N$th Legendre polynomial.
-In the variable density tesseroid case the GLQ is applied in a similar 
-way as \citet{Asgharzadeh2007} and \citet{Uieda2016} do, but in this 
-scenario the density function must be also evaluated on the quadrature 
-nodes:
+\subsection{Gauss-Legendre Quadrature integration}
+
+Applying a $N$th order GLQ, we can approximate each integral in equations
+\ref{eq:tesseroid-pot}, \ref{eq:tesseroid-grav} and \ref{eq:tesseroid-tensor} by a
+weighted sum of the integration kernel evaluated on the roots of an $N$th order Legendre
+polynomial \citep[p.~390]{Hildebrand1987}.
+Unlike the homogeneous density case, the radial density function $\rho(r)$ must also be
+included in the integration and evaluated on the Legendre polynomial roots (i.e.,
+quadrature nodes).
 
@@
 
-\noindent $(r_i, \phi_j, \lambda_k)$ are the GLQ nodes and $W_i^r$ are 
-the quadrature weights \citep[see their definitions 
-on][]{Hildebrand1987}.
-
-From equation \ref{eq:glq-var-dens} we can observe that the GLQ approximates 
-the gravitational fields of a tesseroid as the ones generated by $N_\lambda 
-N_\phi N_r$ point masses located at the nodes of the Legendre polynomials, in 
-the same way as the GLQ works in the homogeneous density case, where $N_i$ ($i 
-\in \{ \lambda, \phi, r \}$) are the orders of the quadrature for each 
-integration.
-Nevertheless, in this new variable density scenario, the information of 
-the density function is summarised into the weights by taking into 
-account only the values it assumes on those nodes.
-Although this may sound discouraging, we intent to prove otherwise and 
-show that this method, along with well suited discretization 
-algorithms, approximates the gravitational fields with good accuracy.
-\todo[inline]{no se si poner esta ultima oracion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Methodology}
-
-In a general way, the numerical error of the GLQ is controlled by the 
-order of the quadrature: a higher GLQ order produces a more accurate 
-approximation \citep{Hildebrand1987}.
-This can also be understood in terms of the tesseroid problem: a higher 
-quadrature order increases the number of point masses used to compute 
-the gravitational fields and thus lowers the numerical error.
-
-This reasoning could lead to a naive methodology in order to ensure 
-accurate gravitational field computations: find the order of the 
-quadrature that produces an acceptable numerical error.
-Nevertheless, this simple proposal has a few problems: (a) it won't be 
-efficient \citep{Li2011, Uieda2016} and (b) it couldn't be applied to 
-the general case, i.e. the needed quadrature order probably won't be 
-the same for every tesseroid, computation point or density function.
-
-That's why we will keep the GLQ order fixed and make use of 
-discretization algorithms.
-Instead of uniformly increase the number of point masses that 
-approximates a certain tesseroid, the discretization algorithms are 
-intended to efficiently subdivide the tesseroid into smaller ones and 
-apply the GLQ to each one of them.
-These algorithms generally involve a discretization inequation that 
-controls if the tesseroid should be subdivided or not, and a certain 
-predefined parameter that works as a threshold for the discretization 
-inequation.
-
-On the variable density tesseroid problem we found two main error 
-sources in the GLQ that should be addressed: (a) the distance between 
-the tesseroid and the computation point, and (b) 
-the variation of the density function.
-The former one is also present in the homogeneous density tesseroid 
-models, such as the one developed by \citet{Uieda2016}, and could be 
-solved by applying a modified version of the adaptive discretization 
-algorithm proposed by \citet{Li2011}.
-Although for the latter we need to introduce a new density-based 
-discretization algorithm that subdivides the tesseroid taking into 
-account the variation of its density function.
-
-\subsection{Modified Adaptive Discretization Algorithm}
+\noindent $(r_i, \phi_j, \lambda_k)$ are the GLQ nodes, and $W_i^r$ are
+the quadrature weights.
+
+\subsection{Adaptive Discretization}
 

### 2018-09-20 manuscript/manuscript.tex
Start editing algorithm section

@@
 It is based on the GLQ approximation and the adaptive discretization of
-\citep{Uieda2016}, which we extend to include a new density-based discretization step.
+\citet{Uieda2016}, which we extend to include a new density-based discretization step.
 In order to ensure the accuracy of the numerical approximation we have
@@
 
-\noindent $(r_i, \phi_j, \lambda_k)$ are the GLQ nodes, and $W_i^r$ are
-the quadrature weights.
+\noindent $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$ are the
+quadrature weights.
 
@@
 
-Studying the homogeneous density tesseroid problem, \citet{Ku1977}
-noticed that keeping the GLQ orders fixed, the approximation
+\citet{Ku1977} noticed that the GLQ integration
 becomes less accurate when the computation point is closer to the
-tesseroid, making the point masses effect more noticeable than the
-tesseroid we want to approximate.
-One way to prevent this could be increasing the GLQ order in case that
-the computation point is near the tesseroid.
-This would uniformly increase the number of point masses inside the
+mass element.
+One way to prevent this from happening would be to increase the GLQ order.
+Doing so would uniformly increase the number of point masses inside the
 tesseroid volume.
-
-Instead of doing this, \citet{Li2011} proposed an adaptive
-discretization algorithm which essentially consists in dividing the
+However, an increase in the point mass concentration is only required close to the
+computation point \citep{Uieda2016}.
+Alternatively, \citet{Li2011} proposed an adaptive
+discretization algorithm which keeps the GLQ order fixed and divides the
 tesseroid based on a ratio between the distance to the computation
 point and its dimensions.
-This algorithm produces a more efficient computation because the
-increase of point masses is produced where are more needed.
-
-\citet{Uieda2016} proposed a modified version of this algorithm, whose
-modifications can be summed up in:
-(1) a faster calculation of the distance from the computation point to
-the tesseroid and
-(2) the application of a stack based algorithm that speeds up the
-computation and gives more control over the recursion step.
-More specifically, \citet{Uieda2016} redefined the dimensions of an
-arbitrary tesseroid: $L_\lambda$, $L_\phi$ and $L_r$. The former ones
-are the arc-distances measured along its top surface while $L_r$ is the
-difference between the radii of the top and bottom surfaces.
+This algorithm produces a more efficient computation because an increased concentration
+of point masses is produced where it is needed more.
+\citet{Uieda2016} developed a modified version of this algorithm, which we will use
+here.
+What follows is a summary of the algorithm and the reader is referred to
+\citet{Uieda2016} for a detailed description.
+
+The following algorithm computes the gravitational fields at a given point at $(r, \phi,
+\lambda)$ due to a tesseroid with its geometric center at $(r_t, \phi_t, \lambda_t)$ and
+dimensions $L_r$, $L_\phi$, and $L_\lambda$ given by
+
+\begin{equation}
+    L_\lambda = r_2 \arccos(\sin^2\phi_t +
+        \cos^2\phi_t\cos(\lambda_2 - \lambda_1)),
+    \label{eq:sizelon}
+\end{equation}
+
+\begin{equation}
+    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1),
+\end{equation}
+
+and
+
+\begin{equation}
+    L_r = r_2 - r_1.
+    \label{eq:sizer}
+\end{equation}
+
+\noindent
+\textit{Step 1}: Check that the tesseroid satisfies the following inequality for each
+dimension $L_i$ of the tesseroid:
+
+\begin{equation}
+    \frac{d}{L_i} \geq D,
+    \label{eq:condition}
+\end{equation}
+
+\noindent
+in which $D$ is a positive scalar called the distance-size ratio and $d$ is the distance
+between the computation point and the geometric center of the tesseroid
+
+\begin{equation}
+    d = \left[
+        r^2 + r_t^2 - 2 r r_t \cos\psi_t
+        \right]^{\frac{1}{2}} ,
+    \label{eq:distance}
+\end{equation}
+
+\begin{equation}
+    \cos\psi_t =
+        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t) .
+\end{equation}
+
+\noindent
+\textit{Step 2}: If none of the dimensions of the tesseroid fail inequality
+\ref{eq:condition}, then compute the gravitational effect of the tesseroid using a
+second-order GLQ (Eq. \ref{eq:glq-var-dens})
 

### 2018-09-20 manuscript/manuscript.tex
Edit the adaptive discretization section

@@
 
-\noindent
 \textit{Step 1}: Check that the tesseroid satisfies the following inequality for each
@@
 
-\noindent
-\textit{Step 2}: If none of the dimensions of the tesseroid fail inequality
+\textit{Step 2}:
+If none of the dimensions of the tesseroid fail inequality
 \ref{eq:condition}, then compute the gravitational effect of the tesseroid using a
-second-order GLQ (Eq. \ref{eq:glq-var-dens})
-
-While keeping the GLQ fixed on second orders, the adaptive
-discretization algorithm starts by checking that the following
-inequality is satisfied:
-
-\begin{equation}
-    \frac{d}{L_i} \geq D,
-\label{eq:distance-size-ratio}
-\end{equation}
-
-\noindent where $i \in \{\lambda, \phi, r\}$, $d$ is the distance
-between the geometric centre of the tesseroid and the computation
-point, and $D$ is a predefined value called distance-size ratio.
-If the inequality \ref{eq:distance-size-ratio} is not held for at least
-one spherical coordinate, then the tesseroid must be divided by half in
-that direction.
-On the contrary, if the inequality is satisfied for every spherical
-coordinate, the algorithm computes the desired gravitational field using the
-GLQ approximation.
-A complete description of the modified adaptive discretization
-algorithm can be found in \citet{Uieda2016}.
-
-In summary, the distance-size ratio $D$ determines indirectly how many
-times the tesseroids will be divided, and therefore efficiently
-regulates both the accuracy of the algorithm and its computation time.
-Moreover, it limits the computation time by stopping the division of
-the tesseroids when their dimensions are lower than a certain value
-(1mm), assuming that such small tesseroids do not modifies
-significantly the desired gravitational field.
-
-On the other hand, the value assigned to the distance-size ratio $D$
-cannot be easily related to the error of the approximation, thus the
-choice of the value of $D$ to get an acceptable accuracy must be
-empirically determined beforehand.
-In order to overcome this, \citet{Uieda2016} compared the numerical
-computation with the analytical solution for a spherical shell with
-homogeneous density.
+second-order GLQ (Eq. \ref{eq:glq-var-dens}).
+Add the computed effect to a running total.
+
+\textit{Step 3}:
+If any dimension fails inequality \ref{eq:condition}, split the tesseroid in half along
+the offending dimensions.
+Repeat steps 1-3 for all smaller tesseroids until none are left.
+
+\textit{Final step}:
+By the end of the algorithm, the running total will be the gravitational effect of the
+tesseroid.
+
+Notice that the distance-size ratio $D$ determines how many
+times the tesseroids will be divided.
+Therefore, it effectively regulates both the accuracy of the algorithm and its
+computation time.
+An optimal value for $D$ cannot be directly calculated from the desired accuracy level.
+Instead, it is empirically determined by comparing the numerical results with the
+analytical solution for a spherical shell.
+\citet{Uieda2016} used a shell with homogeneous density to determine optimal values of
+$D$.
+Here, we will repeat the numerical experiment using analytical expressions for shells
+with density varying according to exponential and linear functions of $r$.
 
@@
     The computation point $Q$ is located on the $z$ axe at a distance $r$ from
-    the centre of the shell.
+    the center of the shell.
     For our purposes we will assume that $Q$ is outside the outer radii,
@@
 Due to the rotational symmetry along any axe that passes through the
-centre of the shell, the expression in equation \ref{eq:shell-pot}
+center of the shell, the expression in equation \ref{eq:shell-pot}
 reproduces the gravity potential field on any outside point at distance
-$r$ from its centre.
+$r$ from its center.
 
@@
 gravitational potential generated by a spherical shell with this density on
-any external point at distance $r$ from its centre:
+any external point at distance $r$ from its center:
 

### 2018-09-20 manuscript/manuscript.tex
In the middle of the density based discretization

@@
 
-Variations in the density function of an arbitrary tesseroid introduce
-a new type of error into the numerical approximation: only a few point
-masses would not reproduce the effect of the density variations.
-The modified version of the adaptive discretization algorithm may help
-in reducing this kind of error, but as it only takes into account the
-dimensions of the tesseroid and the distance to the computation point,
-it's not well suited to fully perform this task.
-
-In order to overcome this, we developed a complementary discretization
-algorithm that takes into account the variation of the density inside
-the tesseroid.
-Basically, it subdivides any tesseroid on the radial direction on the
-depth at which the \emph{maximum density variation} takes place, only
-if a certain discretization inequation holds.
-
-\begin{figure*}
-\centering
-\includegraphics[width=\linewidth]
-    {figures/density-based-discretization-algorithm.pdf}
-\caption{
-    Example intended to show how the density-based algorithm works.}
-\label{fig:density-discretization-algorithm}
-\end{figure*}
-
-In a general way, the algorithm can be summarised in the following steps:
-
-\begin{enumerate}
-\renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Normalise the density function to the [0, 1] interval,
-    \item Compute the absolute difference between the normalised
-          density and a linear function that assumes the same values on the
-          top and bottom radii that this normalised density.
-    \item If product of the maximum of the absolute difference and a
-          tesseroid size ratio is greater that a predefined delta ratio
-          ($\delta$), then the tesseroid will be subdivided on the
-          radius at which this maximum takes place. Otherwise, the
-          tesseroid won't be split.
-    \item If it's discretized, then apply items (2), (3) and (4) for
-          every subdividing tesseroid.
-\end{enumerate}
-
-Lets assume an arbitrary \emph{original} tesseroid with boundary radii
-$R_1$ and $R_2$ and a density function $\rho(r')$, whose radial
-dimension is $L_r = R_2 - R_1$.
-
-First we need to normalise the density function such as its minimum and
-maximum will be 0 and 1, respectively.
-We call this normalised density $\rho_n(r')$ and we calculate it as
-follows:
+The numerical integration of an arbitrary density function introduces
+a new type of problem:
+the integration error from using only a few nodes to discretize the density function.
+The adaptive discretization may help
+to reduce this kind of error.
+However, it does not take into account the density function
+and hence it is not well suited to fully perform this task.
+
+We have developed a complementary discretization
+algorithm that takes into account the variations of the density function.
+This density-based discretization happens prior to the adaptive discretization described
+in the previous section.
+In short, the algorithm divides the tesseroid along the radial dimension at the
+depths at which the \emph{maximum density variations} take place.
+
+Consider an \emph{original} tesseroid with density given by the function $\rho(r')$.
+Before the density-based discretization starts,
+we normalise the density function to the range $[0, 1]$ as follows
 
@@
 \end{equation}
-\noindent where
-\begin{equation}
-    \rho_\text{min} = \text{min}\{ \rho(r') \}, \quad
-    \rho_\text{max} = \text{max}\{ \rho(r') \} \quad
-    \forall \, r' \in [R_1, R_2].
-\end{equation}
-
-\noindent Because the algorithm will be iteratively applied for every
-subdividing tesseroid, we'd like to note that this normalised density
-function will be the same for every subdividing tesseroid.
-In case that the density function is constant, both the maximum and
-minimum densities are equal and the normalised function could not
-be defined, so the density-based discretization algorithm won't be
-applied.
-
-Secondly, we must define a linear function $\rho_l(r')$ that assumes
-the same values as the normalised density $\rho_n(r')$ on the boundary
-radii $R_1$ and $R_2$:
+
+\noindent in which $\rho_\text{min}$ and $\rho_\text{max}$ are the minimum and maximum
+density values inside the tesseroid boundaries.
+We emphasize that this normalised density function will not be modified throughout the
+algorithm.
+In case the density function is constant, both maximum and minimum densities will be
+equal and the density-based discretization algorithm will not be applied.
+
+The algorithm is comprised of the following steps:
+
+\textit{Step 1}:
+Define a linear function $\rho_l(r')$ that assumes
+the same values as the normalised density $\rho_n(r')$ on the boundaries of the
+tesseroid ($r_1$ and $r_2$):
 
@@
     \rho_l(r') =
-    \frac{ \rho_n(R_2) - \rho_n(R_1) }{ R_2 - R_1 } (r' - R_1) + \rho_n(R_1),
+    \frac{ \rho_n(r_2) - \rho_n(r_1) }{ r_2 - r_1 } (r' - r_1) + \rho_n(r_1),
     \label{eq:density-reference-line}
@@
 
-\noindent and compute the absolute difference between this linear
-function and the normalised density:
+\textit{Step 2}:
+Evaluate the normalized and linear density functions on a range of $N$ radii between
+$r_1$ and $r_2$.
+We have opted for $N = 101$ but the specific value of $N$ is not critical to the
+algorithm.
+
+\textit{Step 3}:
+Compute the absolute difference between the values of the linear and normalised density
+functions:
 
@@
 
-Finally, if the following inequation holds, the tesseroid will be divided on
+\textit{Step 4}:
+If the following inequality holds, the tesseroid will not be divided:
+
+\begin{equation}
+    \text{max}\{ \Delta \rho(r') \} \frac{L_r}{L_r^\text{orig}} \le \delta,
+    \label{eq:delta-density}
+\end{equation}
+
+\noindent
+in which $L_r^\text{orig}$ is the radial dimension of the original tesseroid and
+$\delta$ is a positive constant henceforth called the \textit{delta ratio}.
+
+on
 the radius $R_\text{max}$ at which the maximum absolute difference takes
 place:
-
-\begin{equation}
-    \text{max}\{ \Delta \rho(r') \} \frac{L_r^\text{sm}}{L_r} > \delta,
-    \label{eq:delta-density}
-\end{equation}
 
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]
+    {figures/density-based-discretization-algorithm.pdf}
+\caption{
+    Example intended to show how the density-based algorithm works.}
+\label{fig:density-discretization-algorithm}
+\end{figure*}
+
 Once the \emph{original} tesseroid is discretized, we proceed to
@@
 ensures an acceptable accuracy while minimising the computation time.
+
+In a general way, the algorithm can be summarised in the following steps:
+
+\begin{enumerate}
+\renewcommand{\theenumi}{(\arabic{enumi})}
+    \item Normalise the density function to the [0, 1] interval,
+    \item Compute the absolute difference between the normalised
+          density and a linear function that assumes the same values on the
+          top and bottom radii that this normalised density.
+    \item If product of the maximum of the absolute difference and a
+          tesseroid size ratio is greater that a predefined delta ratio
+          ($\delta$), then the tesseroid will be subdivided on the
+          radius at which this maximum takes place. Otherwise, the
+          tesseroid won't be split.
+    \item If it's discretized, then apply items (2), (3) and (4) for
+          every subdividing tesseroid.
+\end{enumerate}
 

### 2018-09-20 manuscript/manuscript.tex
Finished the density-based discretization section
Significantly reduced this part. Used a Step 1, ..., approach to the
algorithm description, like I did for the adaptive discretization. Moved
the description of figure 2 to the caption and removed some redundancy.
Now that the algorithm is in steps, the "Algorithm summary" isn't needed
anymore.

@@
 
-The algorithm is comprised of the following steps:
+The algorithm is comprised of the following steps
+(Fig. \ref{fig:density-discretization-algorithm}):
 
@@
 Define a linear function $\rho_l(r')$ that assumes
-the same values as the normalised density $\rho_n(r')$ on the boundaries of the
+the same values as the normalised density $\rho_n(r')$ at the boundaries of the
 tesseroid ($r_1$ and $r_2$):
@@
 \noindent
-in which $L_r^\text{orig}$ is the radial dimension of the original tesseroid and
+in which $L_r$ is the radial dimension of the tesseroid being considered for division,
+$L_r^\text{orig}$ is the radial dimension of the original tesseroid, and
 $\delta$ is a positive constant henceforth called the \textit{delta ratio}.
 
-on
-the radius $R_\text{max}$ at which the maximum absolute difference takes
-place:
-
-\noindent where the delta ratio $\delta$ is a predefined constant value and
-$L_r^\text{sm}/L_r$ is defined as the tesseroid size ratio, while
-$L_r^\text{sm}$ and $L_r$ are the radial dimensions of the tesseroid that is
-currently analysed and of the \emph{original} one, respectively.
-On the first iteration, the former tesseroid corresponds to the
-\emph{original} one, so $L_r^\text{sm}/L_r$ is equal to one, but for
-future iterations this radial dimension will be computed as the
-difference of the top and bottom radii of the subdividing tesseroid, making
-$L_r^\text{sm}/L_r$ lower than one.
-The tesseroid size ratio is intended to reduce the number of unneeded
-discretizations and focus on the ones that will significantly reduce the
-numerical error:
-a big subdividing tesseroid with small maximum absolute difference would
-cause more relative numerical error than a small subdividing tesseroid with
-higher maximum absolute difference.
+\textit{Step 5}:
+If inequality \ref{eq:delta-density} is not satisfied, then the tesseroid is split in
+two parts at the radius $r_\text{max}$ at which the maximum absolute difference (Eq.
+\ref{eq:density-abs-diff}) takes place.
+
+\textit{Step 6}:
+Repeat steps 1-6 for each smaller tesseroid produced in step 5.
 
@@
 \caption{
-    Example intended to show how the density-based algorithm works.}
+    Example application of the density-based discretization algorithm to a non-linear
+    density function.
+    (a) the normalised density function $\rho_n(r')$ (blue), current boundaries of the
+    tesseroid (orange dots), and the linear density function $\rho_l(r')$ (orange line).
+    The dashed red line represents the maximum density difference $\Delta \rho (r')$ at
+    which the tesseroid would be divided (assuming that the inequality
+    \ref{eq:delta-density} is not satisfied).
+    (b) second iteration of the algorithm with a new linear density function and maximum
+    density difference. The tesseroid would be divided at the depth indicated by the
+    dashed red line.
+    (c) third iteration of the algorithm.
+    (d) final output of the density-based discretization, assuming that all four new
+    tesseroids satisfy inequality \ref{eq:delta-density}.
+}
 \label{fig:density-discretization-algorithm}
@@
 
-Once the \emph{original} tesseroid is discretized, we proceed to
-subsequently apply the same algorithm on every subdividing tesseroid.
-We will start from equation \ref{eq:density-reference-line} defining a
-new linear function $\rho_l(r')$, an absolute difference with equation
-\ref{eq:density-abs-diff} and evaluate the discretization inequation
-\ref{eq:delta-density}, taking into account that now $R_1$ and $R_2$
-are the radial boundaries of this smaller tesseroid.
-If the inequation holds, the current tesseroid is discretized and the
-algorithm is subsequently applied to the resulting subdividing tesseroids.
-
-If on the contrary the inequation \ref{eq:delta-density} is not held,
-then the tesseroid will not be discretized.
-The same happens if the tesseroid radial dimension is bellow the
-threshold of 1mm.
-On the first case due to the fact that no more discretizations are
-needed, and on the second one because smaller tesseroids will not
-introduce meaningful increases in the accuracy of the computation.
-This two conditions guarantee that the algorithm is finite.
-Once any of them is met and no more discretizations are carried out,
-every subdividing tesseroid is subjected to the modified adaptive
-discretization algorithm and then a second order GLQ will be applied in
-order to compute the gravitational field generated by it.
-
-On Figure \ref{fig:density-discretization-algorithm}(a) we propose an
-arbitrary normalised density function as an example of how this
-algorithm works.
-We plotted its corresponding linear function $\rho_l(r')$ and a
-dashed line that represents the maximum absolute difference.
-We will assume that the discretization inequation
-\ref{eq:delta-density} holds in this case, so we must divide the
-tesseroid at the depth at which the maximum difference occurs.
-The dots on Figure \ref{fig:density-discretization-algorithm}
-represent the discretization points, i.e. the radial boundaries of the
-subdividing tesseroids produced by the algorithm.
-
-On Figure \ref{fig:density-discretization-algorithm}(b) we apply the
-algorithm on the shallower subdividing tesseroid.
-In the same way, we plot the linear function and compute the maximum
-absolute difference $\Delta \rho (r')$.
-In this case, the tesseroid size-ratio $L_r^\text{sm}/L_r$ is lower
-than one and will be taken into account in the discretization
-inequation, which will be assumed to hold also in this case, leading to
-the division of this tesseroid.
-
-On Figure \ref{fig:density-discretization-algorithm}(c) we perform the
-same analysis on the deeper subdividing tesseroid, and assuming that the
-inequality holds in this case we also subdivide it.
-
-Finally, if we suppose that the value of $\delta$ used in this example
-does not produce further discretizations, the resulting subdividing
-tesseroids can be seen on Figure
-\ref{fig:density-discretization-algorithm}(d).
-
-It's easy to see that the delta ratio indirectly controls how many
-times the tesseroids will be subdivided based on the variation of the
-density function.
-Therefore it determines the accuracy involving the numerical error that
-the density variation introduces and also the computation time: a lower
-value of $\delta$ will produce more discretizations leading to an
-increase in the accuracy of the approximation but also in the
-computation time.
+Once all smaller tesseroids satisfy Eq. \ref{eq:delta-density},
+each one is subjected to the adaptive
+discretization algorithm described earlier to calculate their gravitational effects.
+
+On the first iteration, the ratio $L_r/L_r^\text{orig} = 1$ because the tesseroid being
+divided is the original one.
+For future iterations, the ratio will be progressively smaller than one as the
+tesseroids get smaller.
+This is intended to limit the number of divisions to the ones that will
+significantly reduce the numerical error:
+dividing a large tesseroid with a small $\text{max}\{ \Delta \rho(r') \}$ would
+improve the integration accuracy more than dividing a small tesseroid with a
+higher $\text{max}\{ \Delta \rho(r') \}$.
+We also do not divide tesseroids with $L_r < 1\ mm$ because any integration errors would
+be negligible due to the small mass of the tesseroid..
+
+The higher $\delta$ is, the fewer divisions will be made, and vice-versa.
+Thus, it controls how many times the tesseroids will be divided based on the density
+function and, indirectly, determines the accuracy and computation time of
+numerical integration.
 This raises the need to determine a maximum value of $\delta$ that
 ensures an acceptable accuracy while minimising the computation time.
-
-In a general way, the algorithm can be summarised in the following steps:
-
-\begin{enumerate}
-\renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Normalise the density function to the [0, 1] interval,
-    \item Compute the absolute difference between the normalised
-          density and a linear function that assumes the same values on the
-          top and bottom radii that this normalised density.
-    \item If product of the maximum of the absolute difference and a
-          tesseroid size ratio is greater that a predefined delta ratio
-          ($\delta$), then the tesseroid will be subdivided on the
-          radius at which this maximum takes place. Otherwise, the
-          tesseroid won't be split.
-    \item If it's discretized, then apply items (2), (3) and (4) for
-          every subdividing tesseroid.
-\end{enumerate}
-
-\subsection{Algorithm summary}
-
-When computing any gravitational field generated by an arbitrary tesseroid on
-a specific computation point, we present the following methodology:
-
-\begin{enumerate}
-    \renewcommand{\theenumi}{(\arabic{enumi})}
-    \item Apply the density-based discretization algorithm in case of
-          variable density producing subdivisions of the original tesseroid,
-    \item Apply the modified adaptive discretization algorithm for
-          each subdivision, generating a set of smaller tesseroids,
-    \item Compute the desired gravitational field for every small tesseroid
-          through the GLQ approximation with fixed quadrature orders
-          (equation \ref{eq:glq-var-dens}).
-\end{enumerate}
-
-In most practical cases, the forward model will consist in a set of
-tesseroids whose gravitational effect will be computed on a set of
-computation points.
-These steps can be applied on every tesseroid and for each computation point,
-adding the contribution of each tesseroid to the resulting
-gravitational field.
 

### 2018-09-20 manuscript/manuscript.tex
Move the shell expression to an appendix
The derivation is not directly relevant to the error analysis because we
don't mention the expressions in the rest of the text. I moved this to
the appendix so it won't interrupt the flow of the text. Being in the
appendix doesn't make it less important or new. It just serves to make
the text for fluid.

@@
 
-\section{Determination of distance-size and delta ratios}
-
-The discretization algorithms described on the previous section are
-introduced to improve the accuracy of the numerical approximation of
-the gravitational fields.
-They involve two predefined ratios that determine how many
-times each tesseroid will be discretized and thus indirectly control
-the numerical error of the approximation.
-The modified version of the adaptive discretization algorithm involves
-the distance-size ratio $D$, while the density-based discretization
-algorithm makes use of the delta ratio $\delta$.
-These algorithms need predefined default values for $D$ and $\delta$
-in order to ensure both acceptable numerical accuracy and computation
-effectiveness.
-
-On the homogeneous density tesseroids case, \citet{Uieda2016} compared
-their numerical model with the analytical solution of a spherical
-shell in order to obtain default values for the distance-size ratio
-$D$.
-We could follow this idea, but for our needs the spherical shell must
-have the same density function as the numerical model.
-In contrast with the homogeneous density case, the analytical solutions
-for a variable density spherical shell are not present in the
-literature, so they must be obtained.
-
-We will then perform these comparisons between the analytical solutions
-and the numerical model for specific density functions: a linear and an
-exponential one.
-From these results we will finally obtain default values for $D$
-and $\delta$ that ensure a numerical error lower than 0.1\%.
-
-\subsection{Analytical Solutions for Spherical Shell}
-
-The integrals that define the gravity fields in equations
-\ref{eq:tesseroid-pot}-\ref{eq:tesseroid-tensor} have almost trivial
-analytical solutions in case of a spherical shell with homogeneous
-density \citep{Mikuska2006,Grombein2013}.
-But for a variable density spherical shell the problem haven't been
-addressed in the literature yet.
-
-Lets consider a spherical shell with inner and outer radii $R_1$ and
-$R_2$, respectively, whose density depends on the radial spherical
-coordinate (Figure \ref{fig:spherical-shell}).
-The gravity potential generated by the shell on an arbitrary external
-point $Q(0,0,r)$, located along the $z$ axe at a distance $r$ from the
-origin, can be written as follows:
-
-\begin{figure}
-\centering
-\includegraphics[width=0.7\linewidth]{figures/spherical-shell.pdf}
-\caption{
-    Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
-    The computation point $Q$ is located on the $z$ axe at a distance $r$ from
-    the center of the shell.
-    For our purposes we will assume that $Q$ is outside the outer radii,
-    i.e. $r > R_2$.
-    }
-\label{fig:spherical-shell}
-\end{figure}
-
-\begin{equation}
-    V_\text{sh}(r) = G
-    \int\limits_0^{2\pi}
-    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
-    \int\limits_{R_1}^{R_2}
-    \frac{\rho(r')}{\ell} {r'}^2 \cos\phi' \,
-    dr' d\phi' d\lambda',
-\end{equation}
-
-\noindent where $\ell$ is defined in equation \ref{eq:ell}. The computation
-point $Q$ is located at latitude $\phi=90^\circ$, so $\ell$ and $\cos\psi$
-(defined on equations \ref{eq:ell} and \ref{eq:cospsi}) get simplified:
-
-\begin{equation}
-    \cos\psi = \sin\phi', \quad
-    \ell = \sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}.
-\end{equation}
-
-Due to the rotational symmetry along the $z$ axe, the integration in
-$\lambda'$ is straightforward:
-
-\begin{equation}
-    V_\text{sh}(r) = 2\pi G
-    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
-    \int\limits_{R_1}^{R_2}
-    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}}
-    \, dr' d\phi',
-\end{equation}
-
-\noindent while the integration in $\phi'$ can be performed
-independently of the density function.
-Making use of SymPy \citep{sympy2017}, a Python library for symbolic
-mathematics, we obtained the following expression for the potential:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}(r) = 2\pi G
-        \int\limits_{R_1}^{R_2}
-        \Big[ & \sqrt{r^2 + r'^2 + 2rr'} - \\
-        & \sqrt{r^2 + r'^2 - 2rr'}
-        \Big] \frac{r'\rho(r')}{r} \, dr'.
-    \end{split}
-\label{eq:shell-pot-sqrts}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}(r) = 2\pi G
-    \int\limits_{R_1}^{R_2}
-    \Big[ \sqrt{r^2 + r'^2 + 2rr'}  -
-    \sqrt{r^2 + r'^2 - 2rr'}
-    \Big] \frac{r'\rho(r')}{r} \, dr'.
-\label{eq:shell-pot-sqrts}
-\end{equation}
-}
-
-For our purposes we can assume that the computation point $Q$ is
-farther from the origin than the outer radius, i.e. $r>R_2$.
-And taking into account that $r'$ is always lower than $R_2$, we can
-simplify the square roots in equation \ref{eq:shell-pot-sqrts}:
-
-\begin{equation}
-    \sqrt{r^2 + r'^2 + 2rr'} = |r + r'| = r + r',
-\end{equation}
-\begin{equation}
-    \sqrt{r^2 + r'^2 - 2rr'} = |r - r'| = r - r',
-\end{equation}
-
-\noindent what leads to the following expression of the potential:
-
-\begin{equation}
-    V_\text{sh}(r) = \frac{4\pi G}{r}
-    \int\limits_{R_1}^{R_2} {r'}^2 \rho(r') \, dr'.
-\label{eq:shell-pot}
-\end{equation}
-
-The equation \ref{eq:shell-pot} allows to easily obtain the exact
-gravity potential generated by a spherical shell with a variable
-density in depth on any outside point located on the $z$ axe.
-Due to the rotational symmetry along any axe that passes through the
-center of the shell, the expression in equation \ref{eq:shell-pot}
-reproduces the gravity potential field on any outside point at distance
-$r$ from its center.
-
-The gradient and the Marussi tensor derived from potentials that
-depends solely on $r$ have only a few non zero components: the vertical
-component of the gradient ($g_z$) and the diagonal components of the
-tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
-They can be easily obtained as \citet{Grombein2013} does, even for any
-arbitrary density function $\rho(r')$:
-
-\begin{equation}
-    g_z = \frac{V_\text{sh}(r)}{r},
-\end{equation}
-\begin{equation}
-    g_{xx} = g_{yy} = -\frac{V_\text{sh}(r)}{r^2}, \quad
-    g_{zz} = \frac{2V_\text{sh}(r)}{r^2}.
-\end{equation}
+\section{Determination of the distance-size and delta ratios}
+
+The distance-size ration $D$ of the adaptive discretization and the delta ratio $\delta$
+of the density-based discretization determine how many times each tesseroid will be
+divided and thus indirectly control the numerical error of the integration.
+Optimal values for $D$ and $\delta$ must be determined in order to ensure both
+acceptable numerical accuracy and computation efficiency for the algorithm.
+
+\citet{Uieda2016} compared the numerical integration of homogeneous density tesseroids
+with the analytical solution of a spherical shell \citep{Mikuska2006,Grombein2013} in
+order to obtain default values for the distance-size ratio $D$.
+We will follow this idea but for our needs the spherical shell must
+have the same density function of radius as our tesseroid model.
+Analytical solutions for a variable density spherical shell are not present in the
+literature and so they must be obtained first.
+We derive the expressions for the gravitational fields of spherical shells with linear
+and exponential density functions of radius in Appendix \ref{sec:shell}.
+
+We perform comparisons between the analytical solutions for the spherical shell  and the
+numerical integration results for linear and exponential density functions.
+From these results, we generalize default values for $D$ and $\delta$ that ensure a
+numerical error lower than 0.1\% of the spherical shell values.
 
@@
 \end{equation}
-
-First, by solving the integral on equation \ref{eq:shell-pot}, we obtained the
-gravitational potential generated by a spherical shell with this density on
-any external point at distance $r$ from its center:
-
-\begin{equation}
-    V_\text{sh}^\text{lin}(r) = \pi G \left[
-    a \frac{R_2^4 - R_1^4}{r} +
-    b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
-    \label{eq:shell-pot-linear}
-\end{equation}
-
-\noindent The first term on this equation reproduces the potential generated
-by a spherical shell with variable density $\rho(r') = ar'$, while the second
-one constitutes the potential generated by a spherical shell with homogeneous
-density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
 
@@
 
+\appendix
+
+\section{Analytical Solutions for Spherical Shell}
+\label{sec:shell}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{figures/spherical-shell.pdf}
+\caption{
+    Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
+    The computation point $Q$ is located on the $z$ axis at a distance $r$ from
+    the origin of the coordinates system.
+    For our purposes we will assume that $Q$ is outside of the shell,
+    i.e. $r > R_2$.
+}
+\label{fig:spherical-shell}
+\end{figure}
+
+Consider a spherical shell with inner radius $R_1$ and outer radius $R_2$,
+whose density is a function $\rho(r')$ of the radial coordinate
+(Fig. \ref{fig:spherical-shell}).
+The gravitational potential generated by the shell on an arbitrary external
+point $Q$ can be written as follows:
+
+\begin{equation}
+    V_\text{sh}(\phi, \lambda, r) = G
+    \int\limits_0^{2\pi}
+    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
+    \int\limits_{R_1}^{R_2}
+    \frac{\rho(r')}{\ell} {r'}^2 \cos\phi' \,
+    dr' d\phi' d\lambda',
+\end{equation}
+
+\noindent where $\ell$ is defined in equation \ref{eq:ell}.
+
+For $Q$ located along the $z$ axis (i.e., $\phi=90^\circ$) at a distance $r$ from the
+origin, equation \ref{eq:ell} simplifies to:
+
+\begin{equation}
+    \ell = \sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}.
+\end{equation}
+
+\noindent
+Because of the rotational symmetry along the $z$ axe, the integration in $\lambda'$ is
+straightforward:
+
+\begin{equation}
+    V_\text{sh}(r) = 2\pi G
+    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
+    \int\limits_{R_1}^{R_2}
+    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}}
+    \, dr' d\phi',
+\end{equation}
+
+\noindent
+while the integration in $\phi'$ can be performed independently of the density function.
+Making use of SymPy \citep{sympy2017}, a Python library for symbolic mathematics, we
+obtained the following expression for the potential:
+
+\iftwocol{
+\begin{equation}
+    \begin{split}
+        V_\text{sh}(r) = 2\pi G
+        \int\limits_{R_1}^{R_2}
+        \Big[ & \sqrt{r^2 + r'^2 + 2rr'} - \\
+        & \sqrt{r^2 + r'^2 - 2rr'}
+        \Big] \frac{r'\rho(r')}{r} \, dr'.
+    \end{split}
+\label{eq:shell-pot-sqrts}
+\end{equation}
+}{
+\begin{equation}
+    V_\text{sh}(r) = 2\pi G
+    \int\limits_{R_1}^{R_2}
+    \Big[ \sqrt{r^2 + r'^2 + 2rr'}  -
+    \sqrt{r^2 + r'^2 - 2rr'}
+    \Big] \frac{r'\rho(r')}{r} \, dr'.
+\label{eq:shell-pot-sqrts}
+\end{equation}
+}
+
+Because the computation point $Q$ is outside of the shell, $r > r'$ and the square roots
+in equation \ref{eq:shell-pot-sqrts} simplify to
+
+\begin{equation}
+    \sqrt{r^2 + r'^2 + 2rr'} = |r + r'| = r + r',
+\end{equation}
+
+\begin{equation}
+    \sqrt{r^2 + r'^2 - 2rr'} = |r - r'| = r - r',
+\end{equation}
+
+\noindent which leads to the following expression for the potential:
+
+\begin{equation}
+    V_\text{sh}(r) = \frac{4\pi G}{r}
+    \int\limits_{R_1}^{R_2} {r'}^2 \rho(r') \, dr'.
+\label{eq:shell-pot}
+\end{equation}
+
+The gradient and the Marussi tensor of potentials that
+depends solely on $r$ have only a few non zero components: the vertical
+component of the gradient ($g_z$) and the diagonal components of the
+tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
+Following \citet{Grombein2013}:
+
+\begin{equation}
+    g_z(r) = \frac{V_\text{sh}(r)}{r},
+\end{equation}
+
+\begin{equation}
+    g_{xx}(r) = g_{yy}(r) = -\frac{V_\text{sh}(r)}{r^2}
+\end{equation}
+
+\begin{equation}
+    g_{zz}(r) = \frac{2V_\text{sh}(r)}{r^2}.
+\end{equation}
+
+From Eq. \ref{eq:shell-pot} we can obtain expressions for gravitational potential for
+different density functions.
+For a linear density function
+
+\begin{equation}
+    \rho(r') = ar' + b\ ,
+\end{equation}
+
+\noindent
+the gravitational potential at any external point is
+
+\begin{equation}
+    V_\text{sh}^\text{lin}(r) = \pi G \left[
+    a \frac{R_2^4 - R_1^4}{r} +
+    b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
+    \label{eq:shell-pot-linear}
+\end{equation}
+
+\noindent The first term on this equation reproduces the potential generated
+by a spherical shell with variable density $\rho(r') = ar'$, while the second
+term constitutes the potential generated by a spherical shell with homogeneous
+density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
+
 \end{document}

### 2018-09-20 manuscript/manuscript.tex
Move exponential density shell to the appendix

@@
 increases the maximum slope of the density function.
-The analytical solution of the gravitational potential generated by a
-spherical shell with a density functions as the one in equation
-\ref{eq:density-exp} can be obtained by solving the equation
-\ref{eq:shell-pot} as follows:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
-        Ab
-        \Big[
-        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
-        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
-        \Big] + \\
-        & \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
-    \end{split}
-\label{eq:shell-pot-exp}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
-    Ab \, e^\frac{\Delta h}{b}
-    \Big[
-    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} -
-    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    \Big] +
-    \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
-\label{eq:shell-pot-exp}
-\end{equation}
-}
 
@@
 
+An exponential density function that assumes the values of $\rho_\text{out}$ and
+$\rho_\text{in}$ on the shell's outer and inner surfaces, respectively, can be defined
+as follows:
+
+\begin{equation}
+    \rho(r') = A e^{-(r' - R)/b} + C,
+\label{eq:density-exp}
+\end{equation}
+
+\noindent where
+
+\begin{equation}
+    A = (\rho_\text{in} - \rho_\text{out})
+        \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
+\end{equation}
+
+\begin{equation}
+    C = \rho_\text{out} - A,
+\end{equation}
+
+\noindent $R$ is the mean Earth radius, and $b$ is a constant
+that determines the variability of the function: a low value of $b$
+increases the maximum slope of the density function.
+
+The analytical solution of the gravitational potential generated by a
+spherical shell with a density function given by equation
+\ref{eq:density-exp} is
+
+\iftwocol{
+\begin{equation}
+    \begin{split}
+        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
+        Ab
+        \Big[
+        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - R}{b}} - \\
+        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - R}{b}}
+        \Big] + \\
+        & \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
+    \end{split}
+\label{eq:shell-pot-exp}
+\end{equation}
+}{
+\begin{equation}
+    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
+    Ab \, e^\frac{R}{b}
+    \Big[
+    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} -
+    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
+    \Big] +
+    \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
+\label{eq:shell-pot-exp}
+\end{equation}
+}
+
 \end{document}

### 2018-09-20 manuscript/manuscript.tex
Bit of editing in the linear density case

@@
 
-Lets perform a comparison between our numerical model and the
-analytical solution for a spherical shell with a linear density
-function like the following one:
-
-\begin{equation}
-    \rho(r') = ar' + b.
-\end{equation}
-
-Secondly we can analyse how the density-based discretization algorithm
-works in case of a linear density function.
-Because the absolute difference defined on equation
-\ref{eq:density-abs-diff} will always be zero, the discretization
-inequation \ref{eq:delta-density} will never be satisfied.
-Therefore, independently of the linear function proposed, the
-density-based algorithm will never be applied, leaving the modified
-adaptive discretization algorithm as the only mechanism to control the
-accuracy of the numerical approximation.
-That's why, for linear density scenarios, we only need to determine
-the minimum value of $D$ needed in order to guarantee an acceptable
-accuracy while ignoring the value of $\delta$.
+A spherical shell with a linear density function given by
+
+\begin{equation}
+    \rho(r') = ar' + b,
+\end{equation}
+
+\noindent
+has an analytical solution given by Eq. \ref{eq:shell-pot-linear}.
+
+The absolute density difference defined on equation
+\ref{eq:density-abs-diff} will always be zero for the linear density case.
+As a result, the inequality \ref{eq:delta-density} will always be satisfied and no
+divisions will ever be performed.
+Therefore, the distance-size ration $D$ of the adaptive discretization algorithm is the
+only mechanism that controls the accuracy of the numerical integration.
+For this reason, we will only determine the minimum value of $D$ needed in order to
+guarantee an acceptable accuracy while ignoring the value of $\delta$.
 
@@
 must build a spherical shell made of tesseroids.
-We propose a mesh of $30^\circ \times 30^\circ$ tesseroids whose outer and
+We use a mesh of $30^\circ \times 30^\circ$ tesseroids whose outer and
 inner radii are the same ones as the spherical shell in order to compute
 the gravitational potential, the vertical component of the gradient
-($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$,
+($g_z$), and the diagonal components of the Marussi tensor ($g_{xx}$,
 $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the

### 2018-09-21 manuscript/manuscript.tex
Finish the linear density analysis section

@@
     San Juan, Argentina\\
-    $^3$Department of Geology and Geophysics, SOEST, University of Hawai‘i at
+    $^3$Department of Earth Sciences, SOEST, University of Hawai‘i at
     M\={a}noa, Honolulu, Hawaii, USA
@@
 
+\begin{table}
+\caption{
+    Description of the tesseroid models of a spherical shell and computation grids
+    used to characterize the accuracy of the numerical integration.
+    Every grids consist of a set of $10 \times 10$ points.
+}
+\label{tab:grids}
+\begin{tabular}{lccc}
+    Name & Tesseroid size & Grid region (degrees) & Grid height (km)
+    \\ \hline
+    Pole & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 2 \\
+    Equator & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/0\text{N}/1\text{N}$ & 2 \\
+    Satellite & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 260 \\
+    Big Grid & $30^\circ \times 30^\circ$ & $0\text{E}/30\text{E}/60\text{N}/90\text{N}$ & 2 \\
+\end{tabular}
+\end{table}
+
 \subsection{Linear Density}
@@
     \rho(r') = ar' + b,
+    \label{eq:density-linear}
 \end{equation}
@@
 divisions will ever be performed.
-Therefore, the distance-size ration $D$ of the adaptive discretization algorithm is the
+Therefore, the distance-size ratio $D$ of the adaptive discretization algorithm is the
 only mechanism that controls the accuracy of the numerical integration.
@@
 
-In order to compare the numerical model with the analytical solution we
+In order to compare the numerical results with the analytical solution we
 must build a spherical shell made of tesseroids.
-We use a mesh of $30^\circ \times 30^\circ$ tesseroids whose outer and
-inner radii are the same ones as the spherical shell in order to compute
-the gravitational potential, the vertical component of the gradient
-($g_z$), and the diagonal components of the Marussi tensor ($g_{xx}$,
-$g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the
-pole, (2) another one on the Equator, (3) one located at satellite
-height and (4) a big grid of $30^\circ \times 30^\circ$.
-More information about this grids can be found on Table
+We use a combination of tesseroid models with different resolutions and compute their
+effects on $10 \times 10$ point grids at different heights and locations.
+All combinations of models and grids are shown in Table \ref{tab:grids}.
+The tesseroid model is composed of a single radial layer.
+The combinations in Table \ref{tab:grids} are repeated for two shells with thickness 1
+km and 35 km.
+The outer radius of both shells are equal to the mean Earth radius and
+the density varies linearly (Eq. \ref{eq:density-linear})
+with angular coefficient
+
+\begin{equation}
+    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R - R_1},
+\end{equation}
+
+\noindent and linear coefficient
+
+\begin{equation}
+    c = \frac{3300\text{kg/m$^3$} -
+        2670\text{kg/m$^3$}}{R - R_1} R +
+        2670\text{kg/m$^3$},
+\end{equation}
+
+\noindent
+in which $R$ = 6378.137 km is the mean Earth radius and $R_1$ is the inner radius of the
+shell (determined by the thickness).
+
+We compute the gravitational potential ($V$), the vertical component of the gradient
+($g_z$), and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$,
+$g_{zz}$) for each of the four tesseroid models and grids in Table
 \ref{tab:grids}.
-
-We will compute these gravitational fields on every point of the four
-grids for different values of $D$, ranging from 0.5 to 10 with a step
-of 0.5.
-Then we will calculate the maximum difference between these results and the
-value that the analytical solution assumes at the same height of the
-computation grid.
-Because the gravity fields generated by a tesseroid are approximated by
-the effect of point masses, the numerical results may vary between
-computation points at same height but at different latitude or
-longitude.
-That's why we will choose for comparison the maximum difference between the
-results on every point of the grid and the analytical solution.
-
-Finally, we will set the acceptable value of $D$ as the minimum value
-at which the corresponding error of the numerical approximation is
-lower than 0.1\%.
-
-\begin{table}
-\caption{
-    Description of the synthetic grids on which the comparison of the
-    numerical model against the analytical solutions for a spherical
-    shell was done.
-    This collection of grids tries to include every possible scenario
-    on which the numerical model can present a different behaviour
-    depending on the size of the grid, its location or its height over
-    the Earth surface. Every grid consists in a set of 10$\times$10
-    points.
-}
-\label{tab:grids}
-\begin{tabular}{lcccc}
-    Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
-    Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ &
-        $0^\circ - 1^\circ$ & 2km \\
-    Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ &
-        $0^\circ - 1^\circ$ & 2km \\
-    Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ &
-        $0^\circ - 1^\circ$ & 260km \\
-    Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ &
-        $0^\circ - 30^\circ$ & 2km \\
-\end{tabular}
-\end{table}
-
-In order to perform the comparison, we proposed a thin and a thick
-spherical shells with 1km and 35km of thickness, respectively, whose
-outer radii are equal to the mean Earth radius.
-On both cases we will suppose that the density assumes a value of
-2670~kg/m$^3$ on the outer surface and 3300~kg/m$^3$ at the inner
-radius:
-
-\begin{equation}
-    \rho(r') = ar' + c,
-    \label{eq:density-linear}
-\end{equation}
-\noindent with
-\begin{equation}
-    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R_2 - R_1},
-\end{equation}
-\begin{equation}
-    c = \frac{3300\text{kg/m$^3$} -
-        2670\text{kg/m$^3$}}{R_2 - R_1} R +
-        2670\text{kg/m$^3$},
-\end{equation}
-
-\noindent where $R$ = 6378.137 km is the mean Earth radius.
-
-We have computed the percentage difference between the numerical model
-with its corresponding analytical solution (equation
-\ref{eq:shell-pot-linear}) for the thick and the thin spherical shell
-on each of the four grids described in Table \ref{tab:grids}.
-We can see the results for the potential, the vertical component of the
-gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on
-Figure \ref{fig:D-linear}.
-We exclude the results corresponding to the other two diagonal
-components of the tensor because they are similar to the ones obtained
-for the $g_{zz}$, although they can be found on the repository.
-\todo[inline]{chequear}
-
-For the linear density case, we can observe that the calculation of the
-potential and the $g_z$ component of its gradient needs a value of
-$D=1$ and $D=2$, respectively, in order to ensure a numerical error
-bellow the 0.1\% threshold.
-On the other hand, the $g_{zz}$ computations show a noticeable
+Other components are equal to zero outside of the shell and are thus omitted from the
+analysis.
+The computations are repeated for values of $D$ ranging from 0.5 to 10 with a step of
+0.5.
+We then will calculate the maximum absolute difference between these results and the
+analytical solutions derived in Appendix \ref{sec:shell}.
+The differences are shown in Fig. \ref{fig:D-linear} as relative to the spherical shell
+values.
+We omit the differences for $g_{xx}$ and $g_{yy}$ because they are equivalent to the
+ones obtained for $g_{zz}$.
+Finally, we set the optimal value of $D$ as the minimum value at which the corresponding
+error of the numerical approximation is lower than 0.1\%.
+
+We observe from Fig. \ref{fig:D-linear} that the relative errors for the
+potential and the $g_z$ component fall below the 0.1\% threshold at
+$D=1$ and $D=2$, respectively, for both shell thicknesses.
+On the other hand, $g_{zz}$ displays a noticeable
 difference between the thin and the thick shell: the first one needs a
-value of $D$ equal to 8 while the later only a $D$ of 3.5.
-We will keep the more conservative result of $D$ equal to 8 in case of
-computing the Marussi tensor components.
+value of $D$ equal to 8 while the later only a $D$ of 3.
 
@@
 \caption{
-    Differences between the gravity fields generated by the numerical model
-    and the analytical solution for a thin and a thick spherical shell of 1km
-    and 35km of thickness, respectively.
-    Both models have the linear density defined on equation
-    \ref{eq:density-linear}.
-    The computations were performed on the four grids described in
-    Table \ref{tab:grids}, for different values of the distance-size
-    ratio $D$.
-    Due to the linearity of the density function, the density-based
-    discretization algorithm is not applied.
-    If the difference is less than 0.1\%, we consider that the model
-    has achieved an acceptable accuracy for the corresponding value of
-    $D$.
+    Differences between the gravitational fields generated by the tesseroid model
+    and the analytical solution for a thin (left) and a thick (right) spherical shell.
+    Both models have a linear density function of radius (Eq. \ref{eq:density-linear}).
+    The computations were performed on the four combinations described in
+    Table \ref{tab:grids}.
+    Due to the linearity of the density function, the density-based discretization
+    algorithm is not applied.
+    Differences are reported as a percentage of the shell values.
+    The horizontal dashed black line represents a target difference of 0.1\%.
 }

### 2018-09-21 manuscript/manuscript.tex
Edit the D-delta search section
Removed a lot of text from this section. As I was reading through it, I
didn't find the explanation about the different sections of the grid
necessary. In fact, it made the text more complicated than it needed to
be. The justification that we need was only:

1. We want the lowest D and highest delta
2. Anything out of the gray line is ruled out
3. D_linear is the best D for the exponential case so we can fix it to
   that.

This is pretty much what that whole explanation was about. Since we
don't need to explain the regions anymore, we don't really need the
extr

@@
 
-Lets perform a similar comparison between the numerical model and
-the analytical solution but now for a spherical shell with exponential
-density.
-On this scenario the problem is more complex due to the fact that the
-density-based discretization will be certainly applied together with
-the modified adaptive discretization algorithm.
-This means that both the distance-size ratio $D$ and the delta ratio
-$\delta$ must be simultaneously determined.
-
-\subsubsection{$D$, $\delta$ space exploration}
-
-From every combination of the $D$ and $\delta$ ratios we want to find
-the one that produces a numerical error lower than the 0.1\% threshold
-while keeping an effective computation, i.e. the ($D$, $\delta)$ pair
-that produces the minimum number of discretizations and ensures an
-acceptable numerical precision.
-We can propose an heuristic method to perform this task: compute the
-numerical error for every ($D$, $\delta$) points belonging to a grid on
-the $D$, $\delta$ space and choose the pair that combines an
-acceptable numerical error with an efficient computation.
-
-Firstly, lets consider a decreasing exponential density function that
-assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's
-outer and inner radii, respectively, that can be defined as follows:
+For an exponential density function, the density-based discretization will be applied
+before the adaptive discretization algorithm.
+This means that optimal values for both the distance-size ratio $D$ and the delta ratio
+$\delta$ must be determined.
+We perform an error analysis similar to what was done for the linear density case.
+However, we only consider the "Big Grid" scenario from Table \ref{tab:grids} for the
+sake of brevity.
+
+The spherical shell and tesseroid model will have an exponential density function given
+by
 
@@
     (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
-    \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
-\end{equation}
+    \left( e^{( R - R_1 )/b} - 1 \right)^{-1},
+\end{equation}
+
 \begin{equation}
@@
 
-\noindent $R$ is the mean Earth radius and $b$ is a constant
-that determines the variability of the function: a low value of $b$
-increases the maximum slope of the density function.
-
-Secondly, we propose a thick spherical shell with a thickness of 35km and
-modelled by a mesh of $30^\circ \times 30^\circ$ tesseroids in order to
-perform the exploration of the numerical error in the $D$, $\delta$
-space.
-We have assigned the density function proposed on equation
-\ref{eq:density-exp} to our spherical shell model with a $b$ value of
-1km, i.e. a highly varying density function.
-
-Finally, we have computed the gravitational potential ($V$), the vertical
-component of its gradient ($g_z$) and the $g_{zz}$ component of the
-Marussi tensor on the big grid (see Table \ref{tab:grids}) by exploring
-different values of $D$ and $\delta$.
-These results were used to obtain the percentage difference between
-them and their corresponding analytical solutions.
-
-On Figure \ref{fig:grid-search} we show
-the resulting percentage difference for each gravitational field.
-The points inside the dotted lines are the ones that present a
-numerical error lower than the 0.1\% threshold.
-Our intention is to choose the one that produces the most efficient computation
-and set it as the default values for $D$ and $\delta$.
+\noindent $R$ is the mean Earth radius, $R_1$ is the inner radius of the
+shell (determined by the thickness), and $b$ is a constant that determines the
+variability of the function: a low value of $b$ increases the maximum slope of the
+density function.
+
+\subsubsection{$D$-$\delta$ space exploration}
+
+We aim to find a combination of the $D$ and $\delta$ that produces a numerical error
+lower than the 0.1\% threshold while minimizing computation time.
+We use a grid search method and compute the numerical error for every ($D$, $\delta$)
+pair belonging to a grid on the $D$-$\delta$ space (Fig. \ref{fig:grid-search}).
+Because this is a time consuming computation, we limit the analysis to a shell of 35 km
+thickness and $b = 1\ \text{km}$ (Eq. \ref{eq:density-exp}), which yields a
+sharp density variation within the shell.
+For optimum algorithm performance, we search for the smallest possible value of $D$ and
+the highest possible value of $\delta$.
+
+We compute the relative difference between the numerical and analytical results for the
+gravitational potential ($V$), the vertical component of its gradient ($g_z$), and the
+$g_{zz}$ component of the Marussi tensor.
+The results are shown in Fig. \ref{fig:grid-search}, in which points inside the dotted
+lines are the ones that present a numerical error lower than the 0.1\% threshold.
+Also shown in Fig. \ref{fig:grid-search} are the $D$ values determined for the linear
+density function in the previous section ($D_\text{linear}$).
+The values of $D_\text{linear}$ are within the 0.1\% threshold and are the optimum
+values of $D$ for $V$ and $g_z$.
+The results for $g_{zz}$ are in agreement with those obtained for the linear density
+function in the case of a thick shell (Fig. \ref{fig:D-linear}).
+Thus, we also choose the conservative value of $D=8$ as the optimum value for $g_{zz}$.
+These results indicate that the values of $D_\text{linear}$ can be safely extrapolated
+to the exponential case.
+
+Having well controlled values for $D$, we are free to explore the integration error as a
+function of $\delta$ in more detail and how it varies for different values of $b$ (Eq.
+\ref{eq:density-exp}).
 
@@
 \caption{
-    Numerical error exploration in the $D$, $\delta$ space.
+    Numerical error exploration in the $D$-$\delta$ space.
     The percentage difference values were obtained from the comparison
@@
     the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a
-    spherical shell with an exponential density as the one defined on
-    equation \ref{eq:density-exp}.
-    These comparison have been carried out with a spherical shell of
-    35km of thickness, whose density function has a value of $b$ of 1km
-    and the gravitational fields were computed on the big grid
-    (defined in Table \ref{tab:grids}).
+    spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
+    These comparisons were carried out on the "Big Grid" configuration (Table
+    \ref{tab:grids}), with a spherical shell of
+    35km of thickness, and density function with $b$ of 1km.
     The points inside the dashed line are the ones that present an
@@
 \end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=\linewidth]
-    {figures/D-delta-grid-search.pdf}
-\caption{
-    $D$, $\delta$ space divided in three regions (A, B and C) based on
-    the definitions of $D_\text{linear}$ and $\delta_p$ (see text for more
-    information).
-    Light grey points are the non-eligible ($D$, $\delta$) pairs
-    belonging to the region A (with $D < D_\text{linear}$).
-    Green points have acceptable numerical error (lower than 0.1\%),
-    while the red ones represent those whose error is greater than the
-    threshold.
-    Dark grey points represent those from which we don't have a priori
-    information about their numerical error.}
-\label{fig:D-delta-grid-search}
-\end{figure}
-
-Because the algorithm is thought to work with default $D$ and $\delta$
-values that ensures these two conditions for every density function,
-the default value for $D$ cannot be lower than the one determined for
-the linear function, which will be called $D_\text{linear}$.
-That's why we can only choose the ($D$, $\delta$) points on which
-$D \geq D_\text{linear}$.
-Taking this fact into account we can divide the $D$, $\delta$ space in
-two subregions: one with $D < D_\text{linear}$ and other with $D \ge
-D_\text{linear}$.
-We can ignore points belonging to the former one (region A on Figure
-\ref{fig:D-delta-grid-search}) and focus on the latter (regions B and C on
-Figure \ref{fig:D-delta-grid-search}).
-
-From the $D \ge D_\text{linear}$ subregion we can explore the numerical
-error of points with $D = D_\text{linear}$ and different values of
-$\delta$.
-We will call $\delta_p$ to the maximum $\delta$ at which the numerical
-error is lower than the 0.1\% threshold while keeping
-$D = D_ \text{linear}$.
-Greater $\delta$ values than $\delta_p$ will produce computations with
-high error (red points in Figure \ref{fig:D-delta-grid-search}), while
-lower $\delta$ values than $\delta_p$ will produce more discretizations
-leading to non-efficient computation (green dots under the
-$(D_\text{linear}, \delta_p)$ point on Figure \ref{fig:D-delta-grid-search}).
-
-With the definition of $\delta_p$ in mind we can subdivide the
-subregion in two: one with points with $\delta \le \delta_p$ (region B
-on Figure \ref{fig:D-delta-grid-search}) and other with points with
-$\delta > \delta_p$ (region C on Figure \ref{fig:D-delta-grid-search}).
-
-Every point in the subregion with $\delta \le \delta_p$ (region B on
-Figure \ref{fig:D-delta-grid-search}) will have a numerical error lower
-than the 0.1\% threshold, but they will always produce more
-discretizations than the ($D_\text{linear}$, $\delta_p$) point leading
-to non-efficient computations.
-It's worth remembering that a lower $\delta$ or a greter $D$ will
-produce more discretizations, increasing the accuracy but also the
-computation time.
-
-On the other hand, we have no a priori information about the numerical
-error of points with $\delta > \delta_p$ (region C on
-Figure \ref{fig:D-delta-grid-search}).
-Nevertheless, the numerical error exploration shown on
-Figure \ref{fig:grid-search} doesn't present any
-point in this region with error lower than the threshold.
-Even though we are analysing a single case, it proves that no point in
-region C is eligible as default ($D$, $\delta$) pair because it should
-ensure numerical accuracy for every exponential density scenario.
-
-In conclusion, the ($D$, $\delta$) point that minimises the number of
-discretizations while keeping an acceptable numerical error under the
-0.1\% threshold is the one whose ratio $D$ is equal to
-$D_\text{linear}$ and its delta ratio is $\delta_p$.
 
@@
     \rho(r') = A e^{-(r' - R)/b} + C,
-\label{eq:density-exp}
 \end{equation}
@@
 The analytical solution of the gravitational potential generated by a
-spherical shell with a density function given by equation
-\ref{eq:density-exp} is
+spherical shell with a density function is
 

### 2018-09-21 manuscript/manuscript.tex
Edit the delta value determination section
This was an easy one as the text was already very good.
I removed a bit of redundancy and shortened the text a bit.

@@
 We perform an error analysis similar to what was done for the linear density case.
-However, we only consider the "Big Grid" scenario from Table \ref{tab:grids} for the
+However, we only consider the ``Big Grid'' scenario from Table \ref{tab:grids} for the
 sake of brevity.
@@
 
-Having well controlled values for $D$, we are free to explore the integration error as a
-function of $\delta$ in more detail and how it varies for different values of $b$ (Eq.
-\ref{eq:density-exp}).
-
 \begin{figure}
@@
     spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
-    These comparisons were carried out on the "Big Grid" configuration (Table
+    These comparisons were carried out on the ``Big Grid'' configuration (Table
     \ref{tab:grids}), with a spherical shell of
@@
 
-In order to determine a default value of $\delta$ that ensures an
-acceptable accuracy for every exponential density case we have
-performed comparisons between the numerical model and the analytical
-solutions, keeping the distance-size ratio $D$ equal to the values
-obtained for the linear density case while exploring different values of
-$\delta$.
-
-Firstly, we have considered the same spherical shell models we have
-used in the linear density case: a thin one (1km of thickness) and a
-thick one (35km of thickness) composed by $30^\circ \times 30^\circ$
-tesseroids.
-But in this case, we assigned an exponential density function like the
-one shown in equation \ref{eq:density-exp} to each shell, choosing
-a few values of $b$ in order to explore different density
-variations.
-
-Finally, we have computed the gravitational potential ($V$), the
-vertical component of its gradient ($g_z$) and the $g_{zz}$ component of
-the Marussi tensor on the big grid (see Table \ref{tab:grids}) and
-calculated the percentage difference between these numerical results and
-their corresponding analytical solutions.
-Our intention is to set the default $\delta$ as the highest value whose
-numerical error is bellow the $0.1\%$ threshold.
-
-On Figure \ref{fig:delta-exponential} we show the density functions
-used on each comparison and the resulting percentage difference for each
-gravitational field.
-The first column shows the results for the thin spherical shell, while the
-second one shows the ones corresponding to the thick shell.
-From these we can conclude that a $\delta = 0.2$ is needed in order to
-achieve an acceptable numerical error for the computation of the
-potential and the $g_z$ component.
-Because the same comparison using the other components of the gradient ($g_x$
-and $g_y$) could not be carried out, we will extrapolate this result
-setting a default $\delta = 0.2$ for their computation.
-
-On the other hand, the comparison with the $g_{zz}$ component of the
-Tensor shows a difference between the thin and the thick spherical
-shell: the former needs a $\delta = 0.2$, but the latter requires a
-$\delta = 0.01$.
-In order to guarantee a good accuracy for every case, we will keep the
-more conservative value of $\delta = 0.01$ for the computation of the
-Marussi tensor components.
+Having chosen values of $D$ equal to linear density case, we are free to explore the
+integration error as a function of $\delta$ in more detail and how it varies for
+different values of $b$ (Eq. \ref{eq:density-exp}).
+We perform the error calculations for the ``Big Grid'' configuration (Table
+\ref{tab:grids}) for two spherical shell models with 1km and 35km thickness.
+The calculations are repeated for different values of $b$ to examine the error variation
+with the sharpness of the density function.
+Because larger $\delta$ values result in fewer tesseroid divisions,
+our intention is to find the highest value of $\delta$ whose numerical error is bellow
+the $0.1\%$ threshold.
+
+Fig. \ref{fig:delta-exponential} shows the density functions and the resulting relative
+differences ($V$, $g_z$, and $g_{zz}$) for different values of $b$ and thickness of the
+shell and tesseroid model.
+The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold for
+$\delta = 0.2$ in both the thick and thin shell cases.
+On the other hand, optimal values of $\delta$ for $g_{zz}$
+are $\delta = 0.2$ for the thin shell and $\delta = 0.01$ for the thick shell.
+Once again, we opt for the conservative value of $\delta = 0.01$ and sacrifice
+performance for the sake of accuracy.
 
@@
 \caption{
-    Density functions used in the $\delta$ ratio determination and the
-    differences between the gravitational fields generated by the
-    numerical model and the analytical solution for a thin and a thick
-    spherical shell of 1km adn 35km of thickness, respectively, both
-    with the exponential density defined on equation
-    \ref{eq:density-exp}.
-    A few values of the constant $b$ were used in order to test the
-    numerical model for different density variations.
-    The computations were performed on the big grid described in
+    Numerical error for different exponential density functions and values of the delta
+    ratio.
+    The top panels show the exponential density functions for different values of $b$
+    (Eq. \ref{eq:density-exp}).
+    The error computations were performed on the ``Big Grid'' configuration described in
     Table \ref{tab:grids}, with fixed values of the distance-size ratio
-    $D$ obtained for in the linear density case ($D=1, 2, 8$ for the
-    potential, gradient components and tensor components, respectively)
-    and different values of the $\delta$ ratio.
-    If the difference is less than 0.1\%, we consider that the model
-    has achieved an acceptable accuracy for the corresponding value of
-    $\delta$.
+    $D$ obtained for the linear density case ($D=1, 2, 8$ for the
+    potential, gradient components, and tensor components, respectively).
+    The difference is reported as a percentage of the shell values.
     }

### 2018-09-21 manuscript/manuscript.tex
Remove the speed comparison section

@@
 
-\todo[inline]{I'm thinking to remove the speed comparison}
-
-\subsection{Speed Comparison}
-
-We have recorded the computation times for the gravity potential, the gradient component $g_z$ and the Marussi tensor component $g_{zz}$ generated by a single tesseroid model ($35$km of thickness) with homogeneous, linear and exponential density (with $b=50$km) at different heights.
-Then we calculated the ratio between these computation times:
-
-\begin{equation}
-    \text{Computation Times Ratio} =
-        \frac{\Delta t_\text{variable}}{\Delta t_\text{homogeneous}},
-    \label{eq:computation-times-ratio}
-\end{equation}
-
-\noindent where $\Delta t_\text{homogeneous}$ and $\Delta t_\text{variable}$ are the computation times corresponding to the homogeneous model and to the variable density ones (linear and exponential).
-
-On Figures \ref{fig:speed-comparison}a and \ref{fig:speed-comparison}b we can see the results of these comparison.
-In the linear density case the density-based discretization algorithm is not applied because the inequation \ref{eq:delta-density} is never satisfied.
-For this reason, the homogeneous density model is up to 5 times faster than the variable density one in this case.
-On the other hand, the density-based discretization algorithm makes the computation slower: up to 20 times slower than the homogeneous density code.
-Nevertheless, this increase in the computation time pays off with a guarantee\todo[inline]{guarantee is a noun?} on its accuracy.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\linewidth]{figures/speed-comparison.pdf}
-\caption{Computation time ratio between a single tesseroid model with variable and with homogeneous density at different heights (defined in equation \ref{eq:computation-times-ratio}) for (a) a linear and (b) an exponential density.}
-\label{fig:speed-comparison}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-09-21 manuscript/manuscript.tex
Move the implementation section to Methodology
The section is small and describes the basis for the method developed,
so it fits in with the methodology. Included a link to the Github repo
(Zenodo DOI will come later).
Removed the details about the Fatiando code using numba. That would be
good on a software paper but not here.

@@
 
+\subsection{Software implementation}
+
+We have implemented the algorithms described in the previous sections in the Python
+programming language.
+The software is based on the pre-existing code for the homogeneous density tesseroid
+\citep{Uieda2016}, more specifically the implementation in the Python library Fatiando a
+Terra v0.5 \citep{Uieda2013}.
+The more time consuming parts of the algorithm are written in the Cython language to
+achieve higher performance.
+This new code is freely available under the BSD 3-clause open-source license.
+It can be downloaded from the online repository
+\href{https://github.com/pinga-lab/tesseroid-variable-density}{github.com/pinga-lab/tesseroid-variable-density}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Implementation}
-
-In order to perform the determinations of the distance-size and the
-delta ratio, we have developed a Python code that implements the
-discretization algorithms and the GLQ to finally compute the
-gravitational fields.
-
-It is based on the pre-existing code developed by \citet{Uieda2016} (and
-included in Fatiando a Terra v.0.5, \citet{Uieda2013}) that computes
-the gravitational effects of tesseroids with homogeneous density.
-Our implementation keeps the the homogeneous tesseroid calculation and
-the modified adaptive discretization algorithm while incorporates the
-new variable density computations with the density-based discretization
-algorithm.
-On their software, \citet{Uieda2016} have optimised the more time
-consuming functions through the Numba compiler.
-In contrast, we have decided to develop our new library in Cython
-language due to difficulties when passing a Python function as argument
-to precompiled Numba methods.
-
-This new code is freely available under the BSD 3-clause open-source
-license and makes use of the pre-existing classes and functions of
-Fatiando a Terra v0.5 in order to define single tesseroid or meshes of
-them.
-It can be downloaded from ...
-\todo[inline]{include repo url}
 

### 2018-09-21 manuscript/manuscript.tex
Small edits to the Neuquen basin application

@@
 
-\section{Example with real data}
-
-Besides the determination of the distance-size and the delta ratio, we
-wanted to illustrate how the algorithm performs under a real data model.
-For this task we have chosen the Neuqu\'en Basin: a sedimentary basin
+\section{Application to the Neuqu\'en Basin}
+
+We applied the new algorithms and optimal values of $D$ and $\delta$ determined
+previously to calculate the gravitational effects of the Neuqu\'en Basin:
+a sedimentary basin
 located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude
-(see Figure \ref{fig:neuquen-basin}a), that includes continental and marine
-siliciclastics, carbonates and evaporites accumulated over the Jurassic and
-the Cretaceous constituting a stratigraphic record up to 5000m of depth
-\citep{Howell2005}.
+(Fig. \ref{fig:neuquen-basin}a).
+The basin includes continental and marine siliciclastics, carbonates, and evaporites
+accumulated over the Jurassic and the Cretaceous constituting a stratigraphic record up
+to 5000m of depth \citep{Howell2005}.
 
@@
 \caption{
-    Forward model example with real data: Nequ\'en sedimentary modelled
-    through tesseroids with exponential density.
-    (a) Topography of Neuqu\'en Basin (in km),
-    (b) thickness of the sedimentary basin (in meters),
-    (c) density function used to model the basin (depth in km and density in
-    kg/m$^3$),
-    (d)-(l) computed gravity fields: potential $V$ (in J/kg), gradient
-    components $g_x$, $g_y$ and $g_z$ (in mGal) and Marussi tensor components
-    (in Eötvös), calculated at 50km of height over the ellipsoid.}
+    Gravitational effects of the Nequ\'en sedimentary basin modelled
+    using tesseroids with an exponential density function of depth.
+    (a) Topography of the Neuqu\'en Basin (in km) and its location in South America,
+    (b) thickness of the sedimentary basin \citep[in meters;][]{Heine2007},
+    (c) the exponential density function used to model the basin
+        (depth in km and density in kg/m$^3$),
+    (d)-(l) computed gravitational fields: potential $V$ (in J/kg), gradient
+    components $g_x$, $g_y$ and $g_z$ (in mGal), and Marussi tensor components
+    (in Eötvös), calculated at 50km of height over the ellipsoid.
+}
 \label{fig:neuquen-basin}
@@
 
-The thickness of the sediment pack was obtained from \citet{Heine2007}.
-We managed to produce a regular grid with a resolution of 0.05$^\circ$ on both
-longitude and latitude directions with one value of thickness on each node
-(see Figure \ref{fig:neuquen-basin}(b)).
-
-In order to compute the gravity fields generated by this basin we created a
-mesh of tesseroids, each one located on one of the nodes of the grid, with
-dimensions of 0.05$^\circ$ $\times$ 0.05$^\circ$ and depth equal to the
-thickness of the basin on the corresponding node.
-
-We must also define a density function for the entire model.
+The thickness of the sediment pack was digitized from \citet{Heine2007} on a regular
+grid with a resolution of 0.05$^\circ$ on both longitude and latitude directions
+(Fig. \ref{fig:neuquen-basin}b).
+We created a tesseroid model of the sediment pack by placing a
+$0.05^\circ \times 0.05^\circ$ tesseroid on each node of the grid.
+The top of each tesseroid was fixed at 0 m depth and the bottom at corresponding
+thickness of the basin.
+
+We must also define a density function for the tesseroid model.
 \citet{Sigismondi2012} measured a minimum and maximum density contrast for
 the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
-So we have chosen, for this specific example, an exponential density (like the
-one in equation \ref{eq:density-exp}) that assumes the minimum value on the
-surface and the maximum at 5858m, i.e. the bottom of the basin, with a value
-of $b$ equal to 2km.
-This density function can be seen on Figure \ref{fig:neuquen-basin}(c).
-
-We have computed the gravity potential $V$, the gradient components $g_x$,
+We have chosen an exponential density variation (Eq. \ref{eq:density-exp}) that assumes
+the minimum value on the top surface and the maximum at 5858m depth (the thickest part
+of the basin), with a value of $b$ equal to 2km.
+This density function can be seen on Fig. \ref{fig:neuquen-basin}c.
+
+Finally, we computed the gravitational potential $V$, gradient components $g_x$,
 $g_y$ and $g_z$, and the Marussi tensor components $g_{xx}$, $g_{xy}$,
-$g_{xz}$, $g_{yy}$  and $g_{zz}$ on a computation grid of 159$\times$163 nodes
-with a 0.05$^\circ$ spacing on both longitude and latitude directions at a
-50km height over the reference ellipsoid.
-The resulting gravity fields can be seen in Figures
-\ref{fig:neuquen-basin}(d)-(l).
-We have excluded the other components of the Marussi tensor from the Figure,
-although they can be found in the repository.
+$g_{xz}$, $g_{yy}$, and $g_{zz}$ on a computation grid of $159\times163$ nodes
+($0.05^\circ$ spacing on both longitude and latitude) at a 50km height over the
+reference ellipsoid.
+The resulting fields can be seen in Fig.
+\ref{fig:neuquen-basin}d-l.
+We have excluded the other components of the Marussi tensor from the Fig.
+\ref{fig:neuquen-basin},
+although they can be found in the online repository.
 

### 2018-09-21 manuscript/manuscript.tex
Start editing the conclusion
Not much to change. This part is really good. Just removed some
unnecessary details.

@@
 
-We have developed a new methodology that computes the gravitational fields
-generated by a tesseroid with continuously variable density in depth.
+We have developed a new methodology to compute the gravitational fields
+generated by a tesseroid with a density given by a continuous function of depth.
 It numerically solves the integrals that define the gravitational potential,
-its gradient and the Marussi tensor components through the Gauss-Legendre
-Quadrature (GLQ), which essentially consists in approximating the
-gravitational fields of the tesseroids with the effect of point masses located
-on nodes of the Quadrature.
-
-Instead of enhancing the accuracy of the method by increasing the order of the
-GLQ on each integration, we make use of the pre-existing modified adaptive
+its gradient, and the Marussi tensor components through the Gauss-Legendre
+Quadrature (GLQ).
+The accuracy of the numerical integration is automatically controlled by an adaptive
 discretization algorithm and a new density-based discretization algorithm.
-
-The former one divides the tesseroid by half if the ratio of the distance to
-the computation point and its size is lower than a predefined distance-size
-ratio $D$.
-This algorithm is introduced in order to minimise the error when the
-computation point is close to the tesseroid, in which case the effect of the
-point masses becomes more noticeable and the approximation lose precision.
-Following this idea, we can summarise that the distance-size ratio $D$
-controls the accuracy of the method: a higher value of $D$ generates more
-divisions of the tesseroids, thus more point masses, what leads to a more
-precise approximation.
-Nevertheless, this algorithm is not sufficient to guarantee the accuracy of
-the method in case of tesseroid with variable density.
-
-For this reason we have developed a density-based discretization algorithm that
-divides the tesseroid on the radius at which the \emph{maximum variation} of
-the density function takes place, only if a certain inequation that relates
-this \emph{maximum variation} with a predefined delta ratio $\delta$ holds.
-The number of subdivisions is indirectly controlled by this delta
-ratio $\delta$, and thus the accuracy of the computation.
+The former divides the tesseroid in half if the ratio of the distance to the computation
+point and the size of the tesseroid is lower than a predefined distance-size ratio $D$.
+This algorithm minimises the integration error when the computation point is close to
+the tesseroid.
+Nevertheless, the adaptive discretization alone is not sufficient to guarantee the
+accuracy of the method in case of tesseroids with variable density.
+
+To overcome this challenge, we have developed a density-based discretization algorithm
+that divides the tesseroid on the points at which the maximum variations of the density
+function takes place.
+The density-based discretization is performed before adaptive discretization as a type
+of pre-processing step.
+The number of divisions performed, and thus the accuracy of the computation, is
+controlled by the delta ratio $\delta$.
 This new algorithm is intended to minimise the error due to the inability of
-the GLQ to produce precise approximations in case of highly varying density:
-the information about the density function is summed up with the values that
-it assumes at the node of the quadrature.
-Increasing the subdivisions of the tesseroid where the \emph{maximum
-variation} of the density takes place becomes an efficient way to reduce this
-kind of error.
+the GLQ to produce precise approximations of highly variable density functions.
 

### 2018-09-21 manuscript/manuscript.tex
Finished editing the conclusion
There was too much recap in the conclusion so I removed and summarized
some of it. I took out much of the algorithm and error analysis
description because it was repeated elsewhere. The conclusion should
give only a brief recap of the main results (in our case, the error
analysis). That's why I finished the conclusion with the results from
error analysis and a couple of sentences about why we can generalize
this to other density functions.

@@
 that divides the tesseroid on the points at which the maximum variations of the density
-function takes place.
+function take place.
 The density-based discretization is performed before adaptive discretization as a type
@@
 This new algorithm is intended to minimise the error due to the inability of
-the GLQ to produce precise approximations of highly variable density functions.
-
-Because there is no direct relation between the values assigned to $D$ and
-$\delta$ with the error of the computation, we had to empirically determine
-the minimum value of $D$ and the maximum value of $\delta$ that produce an
-acceptable accuracy for each gravitational field, while maintaining an
-efficient computation.
-The default values of $D$ and $\delta$ must produce accurate computations for
-most continuous and smooth densities, that's why we have chosen two density
-functions in order to establish them: a linear and an exponential one.
-
-In the linear density case, the densty-based discretization algorithm is not
-applied due to how it's build, so only a default value of $D$ had to be
-determined.
-By comparing the analytical solutions of the gravitational fields generated by
-a spherical shell with the ones produced by the numerical model for different
-values of $D$, we have obtained default values for the distance-size ratio of
-1, 2 and 8 for the potential, its gradient and the Marussi tensor components,
+the GLQ to produce precise approximations of density functions with sharp variations.
+
+Because there is no direct relation between the values assigned to $D$ and $\delta$ and
+the error of the computation, we empirically determined the optimum values for these
+parameters.
+The values minimize the computational load while maintaining the numerical error below
+0.1\% of an analytical solution.
+The density functions used to establish the optimum values were a linear and an
+exponential function.
+The linear function represents the smoothest variation of density and does not require
+density-based discretization at all.
+We analyzed the error for exponential functions ranging from smooth to sharp variations
+in depth to test the accuracy of the density-based discretization.
+Through this analysis, we have obtained optimal values for the distance-size ratio of
+1, 2 and 8 for the potential, its gradient, and the Marussi tensor components,
 respectively.
-From these comparisons we can guarantee a precision of 0.1\% or greater on the
-computations of the gravitational fields generated by any tesseroid with a
-linear density.
-
-The exponential density case is more complex, due to the fact that both $D$
-and $\delta$ must be determined simultaneously.
-We have explored regular gridded points in the $D$, $\delta$ space and
-computed the percentage difference between the analytical solutions for a
-spherical shell and the numerical model.
-We had to take into account that the $D$ values cannot be lower than the ones
-determined for the linear density case due to the fact that they must
-guarantee acceptable accuracy for every continuous density function, including
-the linear one.
-From these comparisons we have concluded that the more efficient computations
-take places when the $D$ values obtained in the linear density scenario are
-used, i.e. the minimum ones.
-This can be understood in terms of which kind of discretization produces more
-subdivisions: a higher $D$ may produce more discretizations than a lower
-$\delta$ because the first one divides the tesseroid in the radial, longitude
-and latitude directions, while the latter only divides it in the radial
-direction.
-In conclusion, we found out that keeping the $D$ values as the ones obtained
-for the linear density case, a $\delta = 0.2$ is needed for the computation of
-the potential and its gradient components, while a $\delta = 0.01$ must be
-used for the Marussi tensor components in order to ensure a 0.1\% precision.
-
-In order to implement this new methodology we have developed a Python code
-that carries out both discretization algorithms and the gravitational field
-computation through the GLQ. It's freely available under the BSD 3-clause
-open-source license and can be downloaded from
-\todo[inline]{add repo url}
-
-Finally, we have shown how this new method performs under a real data example:
-we have computed the gravitational fields generated by the Neuqu\'en
-sedimentary basin through a tesseroid model with an increasing exponential
-density in depth on a regular grid at a computation height of 50km over the
-reference ellipsoid.
+These values are in agreement with previous results for a homogeneous density tesseroid
+\citep{Uieda2016}.
+The optimum $\delta$ values are 0.2 for the potential and its gradient components, while
+a $\delta = 0.01$ must be used for the Marussi tensor components in order to ensure a
+0.1\% precision.
+A wide range of exponential functions were used to infer these results and they are
+conservative estimates, chosen to prioritize accuracy over speed of computations.
+Therefore, they can be safely generalized to other common depth-density relations, like
+quadratic functions, without loss of accuracy.
 
@@
 
-We are indebted to the developers and maintainers of the open-source
-software without which this work would not have been possible.
+We are indebted to the developers and maintainers of the open-source software without
+which this work would not have been possible.
 

### 2018-09-21 manuscript/manuscript.tex
Edit the abstract
Minor edits only. It was very good to begin. Just cleared up some
wording.

@@
 We present a new methodology to compute the gravitational fields generated by
-any tesseroid whose density varies continuously in depth, i.e. a forward
-gravity model in spherical coordinates for varying densities on the radial
-direction.
-It numerical approximates these gravitational fields through the
-Gauss-Legendre Quadrature along with two discretization algorithms that
-increase its accuracy by subdividing the tesseroid into smaller ones.
+tesseroids (spherical prisms) whose density varies continuously with depth according to
+an arbitrary function.
+It approximates the gravitational fields through the Gauss-Legendre Quadrature along
+with two discretization algorithms that automatically control its accuracy by adaptively
+dividing the tesseroid into smaller ones.
 The first one is a preexisting adaptive discretization algorithm that reduces
@@
 point.
-While the second one  is a new density-based discretization algorithm that
-decreases the errors introduced by the variation of the density function.
-The amount of discretizations made by each algorithm is indirectly controlled
-by two predefined scalar values: the distance-size ratio $D$ for the former
-and the delta ratio $\delta$ for the latter.
-We have also obtained analytical solutions for a spherical shell with variable
-density and compared them with the results of the numerical model in case of a
-linear and an exponential density function.
-These comparisons allowed to obtain default values for the distance-size and
-the delta ratio in order to guarantee the accuracy of the forward model.
-The resulting optimal values of $D$ for the gravity potential, the components
-of its gradient and the Marussi tensor components are 1, 2 and 8,
-respectively.
-While a $\delta=0.2$ is needed for the computation of the gravitational
-potential and its gradient components and a $\delta=0.01$ must be used for
-the Marussi tensor components.
-We have finally tested this new methodology by modelling the Neuqu\'en
-sedimentary basin, located to the east of the Andes between 32$^\circ$S and
-40$^\circ$S latitude, assigning an exponential density variation to it and
-computing all the gravitational fields it generates.
+The second is a new density-based discretization algorithm that
+decreases the errors introduced by the variation of the density function with depth.
+The amount of divisions made by each algorithm is indirectly controlled
+by two parameters: the distance-size ratio and the delta ratio.
+We have obtained analytical solutions for a spherical shell with radially variable
+density and compared them to the results of the numerical model for linear and
+exponential density functions.
+These comparisons allowed us to obtain optimum values for the distance-size and
+delta ratios that yield an accuracy of 0.1\% of the analytical solutions.
+The resulting optimal values of distance-size ratio for the gravitational potential, its
+gradient, and Marussi tensor are 1, 2 and 8, respectively.
+A delta ratio of 0.2 is needed for the computation of the gravitational potential and
+its gradient components, while a value of 0.01 must be used for the Marussi tensor
+components.
+Lastly, we apply this new methodology to model the Neuqu\'en Basin, a foreland basin in
+Argentina with a maximum depth of over 5000 m, using an exponential density function.
 \end{summary}

### 2018-09-21 manuscript/manuscript.tex
Fixed mm in math text
Fixed typo on distance-size ratio

@@
 higher $\text{max}\{ \Delta \rho(r') \}$.
-We also do not divide tesseroids with $L_r < 1\ mm$ because any integration errors would
+We also do not divide tesseroids with $L_r < 1$ mm because any integration errors would
 be negligible due to the small mass of the tesseroid..
@@
 
-The distance-size ration $D$ of the adaptive discretization and the delta ratio $\delta$
+The distance-size ratio $D$ of the adaptive discretization and the delta ratio $\delta$
 of the density-based discretization determine how many times each tesseroid will be

### 2018-09-21 manuscript/manuscript.tex
Changed modeling for modelling on the entire manuscript
Because gji requests British English, we must use "modelling"
instead of the American "modeling".

@@
 Furthermore, depth-variable densities have been used in forward and
-inverse gravity modeling, mostly applied to sedimentary basins
+inverse gravity modelling, mostly applied to sedimentary basins
 \citep{Cordell1973, Rao1986, Cowie1990, Rao1993, Rao1994, Zhang2001,
@@
 global scales.
-Hence, designing forward modeling methods that reproduce the gravity anomalies for
+Hence, designing forward modelling methods that reproduce the gravity anomalies for
 such scales is of high importance.
 
-To take into account the curvature of the Earth, many global forward modeling methods
+To take into account the curvature of the Earth, many global forward modelling methods
 are defined in geocentric spherical coordinates.

### 2018-09-21 manuscript/manuscript.tex
Fixed wrong Neuquen Basing typing
Started fixing grids issue

@@
 spherical shells.
-Finally, we applied the methodology to model the Nequen basin, Argentina, using
+Finally, we applied the methodology to model the Neuqu\'en basin, Argentina, using
 tesseroids with an exponentially increasing density with depth.
@@
 must build a spherical shell made of tesseroids.
-We use a combination of tesseroid models with different resolutions and compute their
-effects on $10 \times 10$ point grids at different heights and locations.
-All combinations of models and grids are shown in Table \ref{tab:grids}.
-The tesseroid model is composed of a single radial layer.
-The combinations in Table \ref{tab:grids} are repeated for two shells with thickness 1
-km and 35 km.
+We use a single radial layer of $30^\circ \times 30^\circ$ tesseroids to
+build the spherical shell.
+We compute their gravitational effects on $10 \times 10$ point grids at
+different heights and locations.
+These grids are defined in Table \ref{tab:grids}.
+The computation of the gravitational effects on each grid is repeated for
+two shells with thickness 1km and 35 km.
 The outer radius of both shells are equal to the mean Earth radius and

### 2018-09-21 manuscript/manuscript.tex
Edited grid table and its description

@@
 \caption{
-    Description of the tesseroid models of a spherical shell and computation grids
-    used to characterize the accuracy of the numerical integration.
-    Every grids consist of a set of $10 \times 10$ points.
+    Description of computation grids used to characterize the accuracy of
+    the numerical integration.
+    Every grid consist of a set of $10 \times 10$ points.
 }
@@
 \begin{tabular}{lccc}
-    Name & Tesseroid size & Grid region (degrees) & Grid height (km)
+    Name & Grid region (degrees) & Grid height (km)
     \\ \hline
-    Pole & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 2 \\
-    Equator & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/0\text{N}/1\text{N}$ & 2 \\
-    Satellite & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 260 \\
-    Big Grid & $30^\circ \times 30^\circ$ & $0\text{E}/30\text{E}/60\text{N}/90\text{N}$ & 2 \\
+    Pole & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 2 \\
+    Equator & $0\text{E}/1\text{E}/0\text{N}/1\text{N}$ & 2 \\
+    Satellite & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 260 \\
+    Big Grid & $0\text{E}/30\text{E}/60\text{N}/90\text{N}$ & 2 \\
 \end{tabular}

### 2018-09-21 manuscript/manuscript.tex
Changed arbitrary g function to f function in GLQ description
More fixes to grids issue

@@
     \begin{split}
-        \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda')
+        \iiint\limits_\Omega \rho(r') f(r', \phi', \lambda')
         d\Omega \approx& \\
@@
         \sum\limits_{k=1}^{N^\lambda}
-        & W_i^r W_j^\phi W_k^\lambda \rho(r_i) g(r_i, \phi_j, \lambda_k),
+        & W_i^r W_j^\phi W_k^\lambda \rho(r_i) f(r_i, \phi_j, \lambda_k),
     \end{split}
@@
 \begin{equation}
-    \iiint\limits_\Omega \rho(r') g(r', \phi', \lambda') d\Omega \approx
+    \iiint\limits_\Omega \rho(r') f(r', \phi', \lambda') d\Omega \approx
     A
@@
     \sum\limits_{k=1}^{N^\lambda}
-    W_i^r W_j^\phi W_k^\lambda \rho(r_i) g(r_i, \phi_j, \lambda_k),
+    W_i^r W_j^\phi W_k^\lambda \rho(r_i) f(r_i, \phi_j, \lambda_k),
 \label{eq:glq-var-dens}
@@
 ($g_z$), and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$,
-$g_{zz}$) for each of the four tesseroid models and grids in Table
-\ref{tab:grids}.
+$g_{zz}$) for each grid in Table \ref{tab:grids}.
 Other components are equal to zero outside of the shell and are thus omitted from the
@@
 We perform an error analysis similar to what was done for the linear density case.
-However, we only consider the ``Big Grid'' scenario from Table \ref{tab:grids} for the
+However, we only consider the ``Big Grid'' from Table \ref{tab:grids} for the
 sake of brevity.

### 2018-09-21 manuscript/manuscript.tex
Minor edits after suggestions by Santiago

@@
 tesseroid on any external point are given by volume
-integrals that are generally approximated by numerically.
+integrals that must be numerically approximated.
 The literature offers two main approaches: one involves Taylor series expansion
@@
 \caption{
-    Description of computation grids used to characterize the accuracy of
-    the numerical integration.
-    Every grid consist of a set of $10 \times 10$ points.
+    Description of the computation grids used to characterize the accuracy of the
+    numerical integration.
+    All grids consist of a set of $10 \times 10$ points.
 }
@@
 \begin{tabular}{lccc}
-    Name & Grid region (degrees) & Grid height (km)
+    Name & Grid size & Grid region (degrees) & Grid height (km)
     \\ \hline
-    Pole & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 2 \\
-    Equator & $0\text{E}/1\text{E}/0\text{N}/1\text{N}$ & 2 \\
-    Satellite & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 260 \\
-    Big Grid & $0\text{E}/30\text{E}/60\text{N}/90\text{N}$ & 2 \\
+    Pole & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 2 \\
+    Equator & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/0\text{N}/1\text{N}$ & 2 \\
+    Satellite & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 260 \\
+    Big Grid & $30^\circ \times 30^\circ$ & $0\text{E}/30\text{E}/60\text{N}/90\text{N}$ & 2 \\
 \end{tabular}
@@
 must build a spherical shell made of tesseroids.
-We use a single radial layer of $30^\circ \times 30^\circ$ tesseroids to
-build the spherical shell.
-We compute their gravitational effects on $10 \times 10$ point grids at
-different heights and locations.
+The outer radius of the shell is equal to the mean Earth radius and we repeat the
+calculations for shell thicknesses of 1km and 35km.
+We discretize the shell into a single layer of $30^\circ \times 30^\circ$ tesseroids and
+compute their gravitational effects on $10 \times 10$ point grids with different
+resolutions, heights, and locations on the sphere.
 These grids are defined in Table \ref{tab:grids}.
-The computation of the gravitational effects on each grid is repeated for
-two shells with thickness 1km and 35 km.
-The outer radius of both shells are equal to the mean Earth radius and
-the density varies linearly (Eq. \ref{eq:density-linear})
-with angular coefficient
+The density of the shell and tesseroid model varies linearly with $r$
+(Eq. \ref{eq:density-linear}) with angular coefficient
 
@@
 0.5.
-We then will calculate the maximum absolute difference between these results and the
+We will then calculate the maximum absolute difference between these results and the
 analytical solutions derived in Appendix \ref{sec:shell}.
@@
 parameters.
-The values minimize the computational load while maintaining the numerical error below
+These values minimize the computational load while maintaining the numerical error below
 0.1\% of an analytical solution.

### 2018-09-21 manuscript/manuscript.tex
Prepare for submission of the manuscript
Convert to single column layout and fix the size of some of the figures.
Fixed grid-search figure caption
Removed citation in conclusions

@@
-\documentclass[extra]{gji}
-%~ \documentclass[extra, referee]{gji}
+%\documentclass[extra]{gji}
+\documentclass[extra, referee]{gji}
 
@@
 \usepackage{graphicx}
-\usepackage{todonotes} % to make annotations on margins
-%~ \usepackage[disable]{todonotes} % to disable annotations on margins
 
@@
 \centering
-\includegraphics[width=0.9\linewidth]{figures/tesseroid-uieda.pdf}
+\includegraphics[width=0.6\linewidth]{figures/tesseroid-uieda.pdf}
 \caption{
@@
 
-The numerical integration of an arbitrary density function introduces
-a new type of problem:
-the integration error from using only a few nodes to discretize the density function.
-The adaptive discretization may help
-to reduce this kind of error.
-However, it does not take into account the density function
-and hence it is not well suited to fully perform this task.
-
-We have developed a complementary discretization
-algorithm that takes into account the variations of the density function.
-This density-based discretization happens prior to the adaptive discretization described
-in the previous section.
-In short, the algorithm divides the tesseroid along the radial dimension at the
-depths at which the \emph{maximum density variations} take place.
-
-Consider an \emph{original} tesseroid with density given by the function $\rho(r')$.
-Before the density-based discretization starts,
-we normalise the density function to the range $[0, 1]$ as follows
-
-\begin{equation}
-    \rho_n(r') =
-    \frac{\rho(r') - \rho_\text{min}}{\rho_\text{max} - \rho_\text{min}},
-\end{equation}
-
-\noindent in which $\rho_\text{min}$ and $\rho_\text{max}$ are the minimum and maximum
-density values inside the tesseroid boundaries.
-We emphasize that this normalised density function will not be modified throughout the
-algorithm.
-In case the density function is constant, both maximum and minimum densities will be
-equal and the density-based discretization algorithm will not be applied.
-
-The algorithm is comprised of the following steps
-(Fig. \ref{fig:density-discretization-algorithm}):
-
-\textit{Step 1}:
-Define a linear function $\rho_l(r')$ that assumes
-the same values as the normalised density $\rho_n(r')$ at the boundaries of the
-tesseroid ($r_1$ and $r_2$):
-
-\begin{equation}
-    \rho_l(r') =
-    \frac{ \rho_n(r_2) - \rho_n(r_1) }{ r_2 - r_1 } (r' - r_1) + \rho_n(r_1),
-    \label{eq:density-reference-line}
-\end{equation}
-
-\textit{Step 2}:
-Evaluate the normalized and linear density functions on a range of $N$ radii between
-$r_1$ and $r_2$.
-We have opted for $N = 101$ but the specific value of $N$ is not critical to the
-algorithm.
-
-\textit{Step 3}:
-Compute the absolute difference between the values of the linear and normalised density
-functions:
-
-\begin{equation}
-    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |.
-    \label{eq:density-abs-diff}
-\end{equation}
-
-\textit{Step 4}:
-If the following inequality holds, the tesseroid will not be divided:
-
-\begin{equation}
-    \text{max}\{ \Delta \rho(r') \} \frac{L_r}{L_r^\text{orig}} \le \delta,
-    \label{eq:delta-density}
-\end{equation}
-
-\noindent
-in which $L_r$ is the radial dimension of the tesseroid being considered for division,
-$L_r^\text{orig}$ is the radial dimension of the original tesseroid, and
-$\delta$ is a positive constant henceforth called the \textit{delta ratio}.
-
-\textit{Step 5}:
-If inequality \ref{eq:delta-density} is not satisfied, then the tesseroid is split in
-two parts at the radius $r_\text{max}$ at which the maximum absolute difference (Eq.
-\ref{eq:density-abs-diff}) takes place.
-
-\textit{Step 6}:
-Repeat steps 1-6 for each smaller tesseroid produced in step 5.
-
 \begin{figure*}
@@
 
+The numerical integration of an arbitrary density function introduces
+a new type of problem:
+the integration error from using only a few nodes to discretize the density function.
+The adaptive discretization may help
+to reduce this kind of error.
+However, it does not take into account the density function
+and hence it is not well suited to fully perform this task.
+
+We have developed a complementary discretization
+algorithm that takes into account the variations of the density function.
+This density-based discretization happens prior to the adaptive discretization described
+in the previous section.
+In short, the algorithm divides the tesseroid along the radial dimension at the
+depths at which the \emph{maximum density variations} take place.
+
+Consider an \emph{original} tesseroid with density given by the function $\rho(r')$.
+Before the density-based discretization starts,
+we normalise the density function to the range $[0, 1]$ as follows
+
+\begin{equation}
+    \rho_n(r') =
+    \frac{\rho(r') - \rho_\text{min}}{\rho_\text{max} - \rho_\text{min}},
+\end{equation}
+
+\noindent in which $\rho_\text{min}$ and $\rho_\text{max}$ are the minimum and maximum
+density values inside the tesseroid boundaries.
+We emphasize that this normalised density function will not be modified throughout the
+algorithm.
+In case the density function is constant, both maximum and minimum densities will be
+equal and the density-based discretization algorithm will not be applied.
+
+The algorithm is comprised of the following steps
+(Fig. \ref{fig:density-discretization-algorithm}):
+
+\textit{Step 1}:
+Define a linear function $\rho_l(r')$ that assumes
+the same values as the normalised density $\rho_n(r')$ at the boundaries of the
+tesseroid ($r_1$ and $r_2$):
+
+\begin{equation}
+    \rho_l(r') =
+    \frac{ \rho_n(r_2) - \rho_n(r_1) }{ r_2 - r_1 } (r' - r_1) + \rho_n(r_1),
+    \label{eq:density-reference-line}
+\end{equation}
+
+\textit{Step 2}:
+Evaluate the normalized and linear density functions on a range of $N$ radii between
+$r_1$ and $r_2$.
+We have opted for $N = 101$ but the specific value of $N$ is not critical to the
+algorithm.
+
+\textit{Step 3}:
+Compute the absolute difference between the values of the linear and normalised density
+functions:
+
+\begin{equation}
+    \Delta \rho (r') = | \rho_n(r') - \rho_l(r') |.
+    \label{eq:density-abs-diff}
+\end{equation}
+
+\textit{Step 4}:
+If the following inequality holds, the tesseroid will not be divided:
+
+\begin{equation}
+    \text{max}\{ \Delta \rho(r') \} \frac{L_r}{L_r^\text{orig}} \le \delta,
+    \label{eq:delta-density}
+\end{equation}
+
+\noindent
+in which $L_r$ is the radial dimension of the tesseroid being considered for division,
+$L_r^\text{orig}$ is the radial dimension of the original tesseroid, and
+$\delta$ is a positive constant henceforth called the \textit{delta ratio}.
+
+\textit{Step 5}:
+If inequality \ref{eq:delta-density} is not satisfied, then the tesseroid is split in
+two parts at the radius $r_\text{max}$ at which the maximum absolute difference (Eq.
+\ref{eq:density-abs-diff}) takes place.
+
+\textit{Step 6}:
+Repeat steps 1-6 for each smaller tesseroid produced in step 5.
+
 Once all smaller tesseroids satisfy Eq. \ref{eq:delta-density},
@@
 shell (determined by the thickness).
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{figures/linear-D.pdf}
+\caption{
+    Differences between the gravitational fields generated by the tesseroid model
+    and the analytical solution for a thin (left) and a thick (right) spherical shell.
+    Both models have a linear density function of radius (Eq. \ref{eq:density-linear}).
+    The computations were performed on the four combinations described in
+    Table \ref{tab:grids}.
+    Due to the linearity of the density function, the density-based discretization
+    algorithm is not applied.
+    Differences are reported as a percentage of the shell values.
+    The horizontal dashed black line represents a target difference of 0.1\%.
+}
+\label{fig:D-linear}
+\end{figure*}
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=\linewidth]{figures/linear-D.pdf}
-\caption{
-    Differences between the gravitational fields generated by the tesseroid model
-    and the analytical solution for a thin (left) and a thick (right) spherical shell.
-    Both models have a linear density function of radius (Eq. \ref{eq:density-linear}).
-    The computations were performed on the four combinations described in
-    Table \ref{tab:grids}.
-    Due to the linearity of the density function, the density-based discretization
-    algorithm is not applied.
-    Differences are reported as a percentage of the shell values.
-    The horizontal dashed black line represents a target difference of 0.1\%.
-}
-\label{fig:D-linear}
-\end{figure*}
-
 \subsection{Exponential Density}
@@
 \centering
-\includegraphics[width=\linewidth]
+\includegraphics[width=0.5\linewidth]
     {figures/grid-search.pdf}
@@
     spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
-    These comparisons were carried out on the ``Big Grid'' configuration (Table
+    These comparisons were carried out on the ``Big Grid'' (Table
     \ref{tab:grids}), with a spherical shell of
@@
 \subsubsection{Delta ratio determination}
-
-Having chosen values of $D$ equal to linear density case, we are free to explore the
-integration error as a function of $\delta$ in more detail and how it varies for
-different values of $b$ (Eq. \ref{eq:density-exp}).
-We perform the error calculations for the ``Big Grid'' configuration (Table
-\ref{tab:grids}) for two spherical shell models with 1km and 35km thickness.
-The calculations are repeated for different values of $b$ to examine the error variation
-with the sharpness of the density function.
-Because larger $\delta$ values result in fewer tesseroid divisions,
-our intention is to find the highest value of $\delta$ whose numerical error is bellow
-the $0.1\%$ threshold.
-
-Fig. \ref{fig:delta-exponential} shows the density functions and the resulting relative
-differences ($V$, $g_z$, and $g_{zz}$) for different values of $b$ and thickness of the
-shell and tesseroid model.
-The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold for
-$\delta = 0.2$ in both the thick and thin shell cases.
-On the other hand, optimal values of $\delta$ for $g_{zz}$
-are $\delta = 0.2$ for the thin shell and $\delta = 0.01$ for the thick shell.
-Once again, we opt for the conservative value of $\delta = 0.01$ and sacrifice
-performance for the sake of accuracy.
 
@@
 \end{figure*}
+
+Having chosen values of $D$ equal to linear density case, we are free to explore the
+integration error as a function of $\delta$ in more detail and how it varies for
+different values of $b$ (Eq. \ref{eq:density-exp}).
+We perform the error calculations for the ``Big Grid'' configuration (Table
+\ref{tab:grids}) for two spherical shell models with 1km and 35km thickness.
+The calculations are repeated for different values of $b$ to examine the error variation
+with the sharpness of the density function.
+Because larger $\delta$ values result in fewer tesseroid divisions,
+our intention is to find the highest value of $\delta$ whose numerical error is bellow
+the $0.1\%$ threshold.
+
+Fig. \ref{fig:delta-exponential} shows the density functions and the resulting relative
+differences ($V$, $g_z$, and $g_{zz}$) for different values of $b$ and thickness of the
+shell and tesseroid model.
+The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold for
+$\delta = 0.2$ in both the thick and thin shell cases.
+On the other hand, optimal values of $\delta$ for $g_{zz}$
+are $\delta = 0.2$ for the thin shell and $\delta = 0.01$ for the thick shell.
+Once again, we opt for the conservative value of $\delta = 0.01$ and sacrifice
+performance for the sake of accuracy.
 
@@
 respectively.
-These values are in agreement with previous results for a homogeneous density tesseroid
-\citep{Uieda2016}.
+These values are in agreement with previous results for a homogeneous density tesseroid.
 The optimum $\delta$ values are 0.2 for the potential and its gradient components, while
@@
 \centering
-\includegraphics[width=0.7\linewidth]{figures/spherical-shell.pdf}
+\includegraphics[width=0.5\linewidth]{figures/spherical-shell.pdf}
 \caption{
@@
 
-\iftwocol{
 \begin{equation}
@@
 \end{equation}
-}{
-\begin{equation}
-    V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
-    Ab \, e^\frac{R}{b}
-    \Big[
-    (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} -
-    (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
-    \Big] +
-    \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
-\label{eq:shell-pot-exp}
-\end{equation}
-}
 

### 2018-09-21 manuscript/manuscript.tex
Fixed some american to british word spelling

@@
 \begin{figure}
-\centering
+\centreing
 \includegraphics[width=0.6\linewidth]{figures/tesseroid-uieda.pdf}
@@
 The following algorithm computes the gravitational fields at a given point at $(r, \phi,
-\lambda)$ due to a tesseroid with its geometric center at $(r_t, \phi_t, \lambda_t)$ and
+\lambda)$ due to a tesseroid with its geometric centre at $(r_t, \phi_t, \lambda_t)$ and
 dimensions $L_r$, $L_\phi$, and $L_\lambda$ given by
@@
 in which $D$ is a positive scalar called the distance-size ratio and $d$ is the distance
-between the computation point and the geometric center of the tesseroid
+between the computation point and the geometric centre of the tesseroid
 
@@
 \begin{figure*}
-\centering
+\centreing
 \includegraphics[width=\linewidth]
@@
 \begin{figure*}
-\centering
+\centreing
 \includegraphics[width=\linewidth]{figures/linear-D.pdf}
@@
 \begin{figure}
-\centering
+\centreing
 \includegraphics[width=0.5\linewidth]
@@
 \begin{figure*}
-\centering
+\centreing
 \includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
@@
 \begin{figure*}
-\centering
+\centreing
 \includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
@@
 density-based discretization at all.
-We analyzed the error for exponential functions ranging from smooth to sharp variations
+We analysed the error for exponential functions ranging from smooth to sharp variations
 in depth to test the accuracy of the density-based discretization.
@@
 \begin{figure}
-\centering
+\centreing
 \includegraphics[width=0.5\linewidth]{figures/spherical-shell.pdf}

### 2018-11-12 manuscript/manuscript.tex
Fix wrong substitution of "\centreing"
Added iftwocol to grid-search figure in manuscript.tex
Fixed typo in author name in manuscript.tex
Fix editor's remark 5

@@
 \author[S.R. Soler, A. Pesce, L. Uieda and M.E. Gimenez]{
-    Santiado R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and
+    Santiago R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and
     Mario E. Gimenez$^{1,2}$ \\
@@
 \begin{figure}
-\centreing
+\centering
 \includegraphics[width=0.6\linewidth]{figures/tesseroid-uieda.pdf}
@@
 for the gravitational potential, acceleration, and gradient tensor components
-by comparing the numerical model with the fields generated by a spherical shell.
+by comparing the numerical model with the analytical solution of the fields generated by
+a spherical shell.
 
@@
 \begin{figure*}
-\centreing
+\centering
 \includegraphics[width=\linewidth]
@@
 \begin{figure*}
-\centreing
+\centering
 \includegraphics[width=\linewidth]{figures/linear-D.pdf}
@@
 \begin{figure}
-\centreing
+\centering
+\iftwocol{
+\includegraphics[width=0.9\linewidth]
+    {figures/grid-search.pdf}
+}{
 \includegraphics[width=0.5\linewidth]
     {figures/grid-search.pdf}
+}
 \caption{
@@
 \begin{figure*}
-\centreing
+\centering
 \includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
@@
 \begin{figure*}
-\centreing
+\centering
 \includegraphics[width=\linewidth]{figures/neuquen-basin.pdf}
@@
 \begin{figure}
-\centreing
+\centering
 \includegraphics[width=0.5\linewidth]{figures/spherical-shell.pdf}

### 2018-11-12 manuscript/manuscript.tex
Fix editor's remark 6

@@
 where the gravitational fields are going to be calculated.
-The first and second derivatives of the gravitational potential are taken with
+The first- and second-order derivatives of the gravitational potential are taken with
 respect to the local north-oriented Cartesian coordinate system of $P$
@@
 \citet{Grombein2013} provide efficient formulations for the volume integrals of the
-gravitational potential and its first and second derivatives of a tesseroid with
+gravitational potential and its first- and second-order derivatives of a tesseroid with
 homogeneous density.

### 2018-11-12 manuscript/manuscript.tex
Fix editor's remark 7

@@
 
-We define an external computation point $P(r, \phi, \lambda)$ in a geocentric spherical
-coordinate system at a radius $r$, geocentric latitude $\phi$, and longitude $\lambda$
-where the gravitational fields are going to be calculated.
+Consider an arbitrary tesseroid in a geocentric spherical coordinate system defined by
+pairs of geocentric latitudinal ($\phi_1$, $\phi_2$), longitudinal ($\lambda_1$,
+$\lambda_2$) and radial ($r_1$, $r_2$) boundaries.
+We define an external computation point $P(r, \phi, \lambda)$ located at radius $r$,
+geocentric latitude $\phi$, and longitude $\lambda$ where the gravitational fields the
+tesseroid generates are going to be calculated.
 The first- and second-order derivatives of the gravitational potential are taken with
@@
     \begin{split}
-        \iiint\limits_\Omega \rho(r') f(r', \phi', \lambda')
-        d\Omega \approx& \\
+        \int\limits_{\lambda_1}^{\lambda_2}
+        \int\limits_{\phi_1}^{\phi_2}
+        \int\limits_{r_1}^{r_2}
+        \rho(r') f(r', \phi', \lambda')
+        dr' d\phi' d\lambda' \approx& \\
         A
@@
         \sum\limits_{k=1}^{N^\lambda}
-        & W_i^r W_j^\phi W_k^\lambda \rho(r_i) f(r_i, \phi_j, \lambda_k),
+        W_i^r W_j^\phi W_k^\lambda
+        &\rho(r_i) f(r_i, \phi_j, \lambda_k),
     \end{split}
@@
 \begin{equation}
-    \iiint\limits_\Omega \rho(r') f(r', \phi', \lambda') d\Omega \approx
+    \int\limits_{\lambda_1}^{\lambda_2}
+    \int\limits_{\phi_1}^{\phi_2}
+    \int\limits_{r_1}^{r_2}
+    \rho(r') f(r', \phi', \lambda')
+    dr' d\phi' d\lambda' \approx
     A
@@
 
-\noindent $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$ are the
-quadrature weights.
+\noindent $f(r', \phi', \lambda')$ are the non-density dependent portion of the
+integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$
+are
+the quadrature weights.
 

### 2018-11-12 manuscript/manuscript.tex
Fix editor's remark 9

@@
 \noindent $f(r', \phi', \lambda')$ are the non-density dependent portion of the
-integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$
-are
-the quadrature weights.
+integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and
+$W_i^r$, $W_j^\phi$, $W_k^\lambda$ are the quadrature weights.
 

### 2018-11-12 manuscript/manuscript.tex
Fix minor spelling error

--- 2018-11-12 manuscript/manuscript.tex
@@
 
-\noindent $f(r', \phi', \lambda')$ are the non-density dependent portion of the
-integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$
-are
-the quadrature weights.
+\noindent $f(r', \phi', \lambda')$ is the non-density dependent portion of the
+integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and
+$W_i^r$, $W_j^\phi$, $W_k^\lambda$ are the quadrature weights.
 

### 2018-11-12 manuscript/manuscript.tex
Fix editor's remark 8
Fix editor's remark 11

@@
 $W_i^r$, $W_j^\phi$, $W_k^\lambda$ are the quadrature weights.
+The GLQ nodes and weights can be calculated as \citet[p.~391]{Hildebrand1987} (see
+\citet{Uieda2016} for nodes and weights definitions on similar notation as this paper).
 
@@
 divided is the original one.
-For future iterations, the ratio will be progressively smaller than one as the
+For further iterations, the ratio will be progressively smaller than one as the
 tesseroids get smaller.

### 2018-11-12 manuscript/manuscript.tex
Fix editor's remark 12
Fix editor's remark 14
Fix editor's remark 16

@@
 We also do not divide tesseroids with $L_r < 1$ mm because any integration errors would
-be negligible due to the small mass of the tesseroid..
+be negligible due to the small mass of the tesseroid.
 
@@
 \begin{equation}
-    c = \frac{3300\text{kg/m$^3$} -
+    b = \frac{3300\text{kg/m$^3$} -
         2670\text{kg/m$^3$}}{R - R_1} R +
@@
 
-Having chosen values of $D$ equal to linear density case, we are free to explore the
+Having chosen values of $D$ equal to the linear density case, we are free to explore the
 integration error as a function of $\delta$ in more detail and how it varies for

### 2018-11-15 manuscript/manuscript.tex
Fix editor's remark 21
Fix editor's remark 20
Specify smooth function in abstract

--- 2018-11-12 manuscript/manuscript.tex
@@
 tesseroids (spherical prisms) whose density varies continuously with depth according to
-an arbitrary function.
+an arbitrary smooth function.
 It approximates the gravitational fields through the Gauss-Legendre Quadrature along
@@
 for the gravitational potential, acceleration, and gradient tensor components
-by comparing the numerical model with the analytical solution of the fields generated by
-a spherical shell.
+by comparing the numerical model with the fields generated by a spherical shell.
 
@@
 
-Consider an arbitrary tesseroid in a geocentric spherical coordinate system defined by
-pairs of geocentric latitudinal ($\phi_1$, $\phi_2$), longitudinal ($\lambda_1$,
-$\lambda_2$) and radial ($r_1$, $r_2$) boundaries.
-We define an external computation point $P(r, \phi, \lambda)$ located at radius $r$,
-geocentric latitude $\phi$, and longitude $\lambda$ where the gravitational fields the
-tesseroid generates are going to be calculated.
-The first- and second-order derivatives of the gravitational potential are taken with
+We define an external computation point $P(r, \phi, \lambda)$ in a geocentric spherical
+coordinate system at a radius $r$, geocentric latitude $\phi$, and longitude $\lambda$
+where the gravitational fields are going to be calculated.
+The first and second derivatives of the gravitational potential are taken with
 respect to the local north-oriented Cartesian coordinate system of $P$
@@
 \citet{Grombein2013} provide efficient formulations for the volume integrals of the
-gravitational potential and its first- and second-order derivatives of a tesseroid with
+gravitational potential and its first and second derivatives of a tesseroid with
 homogeneous density.
@@
     \begin{split}
-        \int\limits_{\lambda_1}^{\lambda_2}
-        \int\limits_{\phi_1}^{\phi_2}
-        \int\limits_{r_1}^{r_2}
-        \rho(r') f(r', \phi', \lambda')
-        dr' d\phi' d\lambda' \approx& \\
+        \iiint\limits_\Omega \rho(r') f(r', \phi', \lambda')
+        d\Omega \approx& \\
         A
@@
         \sum\limits_{k=1}^{N^\lambda}
-        W_i^r W_j^\phi W_k^\lambda
-        &\rho(r_i) f(r_i, \phi_j, \lambda_k),
+        & W_i^r W_j^\phi W_k^\lambda \rho(r_i) f(r_i, \phi_j, \lambda_k),
     \end{split}
@@
 \begin{equation}
-    \int\limits_{\lambda_1}^{\lambda_2}
-    \int\limits_{\phi_1}^{\phi_2}
-    \int\limits_{r_1}^{r_2}
-    \rho(r') f(r', \phi', \lambda')
-    dr' d\phi' d\lambda' \approx
+    \iiint\limits_\Omega \rho(r') f(r', \phi', \lambda') d\Omega \approx
     A
@@
 
-\noindent $f(r', \phi', \lambda')$ is the non-density dependent portion of the
-integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and
-$W_i^r$, $W_j^\phi$, $W_k^\lambda$ are the quadrature weights.
+\noindent $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$ are the
+quadrature weights.
 

### 2018-11-15 manuscript/manuscript.tex
Added fukushima2018 and lin2018 review on the into

@@
 
+Two recent papers present other approaches for homogeneous tesseroids gravitational
+fields computation and incorporate methodologies for variable density tesseroid with
+specific density profiles.
+\citet{Fukushima2018} analytically integrates the volumetric integral for the
+gravitational potential in the radial direction, obtaining a surface integral, which is
+then numerically solved by applying a conditionally splitting and the double exponential
+quadrature rule. The gradient of the potential and the Marussi tensor components are
+then computed by finite differences. \citet{Fukushima2018} also generalizes his method
+to an arbitrary polynomial density profile on the radial coordinate.
+\citet{Lin2018} made a comparison of the different methodologies for homogeneous
+tesseroid gravitational fields computation by applying different discretizations
+algorithms to each one of them and developed a combined method.
+For computation points near the tesseroid they apply a GLQ numerical integration with
+a two dimensional adaptive discretization algorithm, i.e. only in the horizontal
+dimension. But, if the computation point is farther than a certain truncation distance,
+then a second order Taylor series approximation is applied along with the regular
+subdivision developed by \citet{Grombein2013}.
+They also introduce a variation of their combined method to compute the gravitational
+fields generated by tesseroids with linear density on the radial coordinate.
+
+In summary, the bibliography is offering different methodologies for computing
+gravitational fields of homogeneous tesseroids or with certain specific density
+functions in depth.
+Although any density profile could be approximated by one of these specific density
+functions, one should acknowledge the error that such approximation would propagate into
+the gravitational field in order to guarantee certain degree of accuracy.
+
 We present a new algorithm for computing the gravitational fields generated by any
-tesseroid with an arbitrary continuous density function on any external point.
+tesseroid with any arbitrary continuous density function on any external point.
 It is based on the GLQ approximation and the adaptive discretization of

### 2018-11-15 manuscript/manuscript.tex
Change smooth with continuous

@@
 tesseroids (spherical prisms) whose density varies continuously with depth according to
-an arbitrary smooth function.
+an arbitrary continuous function.
 It approximates the gravitational fields through the Gauss-Legendre Quadrature along

### 2018-11-15 manuscript/manuscript.tex
Remove continuously

--- 2018-11-15 manuscript/manuscript.tex
@@
 We present a new methodology to compute the gravitational fields generated by
-tesseroids (spherical prisms) whose density varies continuously with depth according to
-an arbitrary smooth function.
+tesseroids (spherical prisms) whose density varies with depth according to
+an arbitrary continuous function.
 It approximates the gravitational fields through the Gauss-Legendre Quadrature along

### 2018-11-16 manuscript/manuscript.tex
Specify arbitrary continuous density function

@@
 The Taylor series expansion is not well suited to develop an algorithm for
-a density varying with depth according to an arbitrary function.
+a density varying with depth according to an arbitrary continuous function.
 Different series expansion terms would have to obtained for each density function
 desired.
-Conversely, an arbitrary density function can be included in the GLQ without
+Conversely, an arbitrary continuous density function can be included in the GLQ without
 any change to the integration method.
@@
 functions in depth.
-Although any density profile could be approximated by one of these specific density
-functions, one should acknowledge the error that such approximation would propagate into
-the gravitational field in order to guarantee certain degree of accuracy.
+Although any continuous density profile could be approximated by one of these specific
+density functions, one should acknowledge the error that such approximation would
+propagate into the gravitational field in order to guarantee certain degree of accuracy.
 
@@
 Here, we will assume that the tesseroid has a density varying with $r$ according to an
-arbitrary function $\rho(r)$.
+arbitrary continuous function $\rho(r)$.
 Thus, the integrals for the gravitational fields are slightly modified to
@@
 
-The numerical integration of an arbitrary density function introduces
+The numerical integration of an arbitrary continuous density function introduces
 a new type of problem:

### 2018-11-16 manuscript/manuscript.tex
Normalise tensor and gradient words

@@
 \citet{Uieda2016} have also obtained standard values of $D$
-for the gravitational potential, acceleration, and gradient tensor components
+for the gravitational potential, its gradient and Marussi tensor components
 by comparing the numerical model with the fields generated by a spherical shell.
@@
     $D$ obtained for the linear density case ($D=1, 2, 8$ for the
-    potential, gradient components, and tensor components, respectively).
+    potential, gradient components, and Marussi tensor components, respectively).
     The difference is reported as a percentage of the shell values.

### 2018-11-16 code/tesseroid_density/_tesseroid.pyx
Removed tensor calculation in tesseroid_density

@@ -94,7 +94,7 @@ cdef rediscretizer(
                 # Put the nodes in the current range
                 scale = scale_nodes(w, e, s, n, top, bottom, lonc, sinlatc,
                                     coslatc, rc)
-                res += kernel(lon, sinlat, coslat, radius, scale, density, 
+                res += kernel(lon, sinlat, coslat, radius, scale, density,
                               lonc, sinlatc, coslatc, rc)
         result[l] += res
     return error_code
@@ -529,489 +529,4 @@ cdef inline double kernelz_variable(
     # Multiply by -1 so that z is pointing down for gz and the gravity anomaly
     # doesn't look inverted (ie, negative for positive density)
     result *= -1
-    return result*scale
-
-
-@cython.boundscheck(False)
-@cython.wraparound(False)
-@cython.cdivision(True)
-def gxx(
-    numpy.ndarray[double, ndim=1] bounds,
-    object density,
-    double ratio,
-    int STACK_SIZE,
-    numpy.ndarray[double, ndim=1] lons,
-    numpy.ndarray[double, ndim=1] sinlats,
-    numpy.ndarray[double, ndim=1] coslats,
-    numpy.ndarray[double, ndim=1] radii,
-    numpy.ndarray[double, ndim=1] result):
-    """
-    Calculate this gravity field of a tesseroid at given locations.
-    """
-    cdef int error
-    if callable(density):
-        error = rediscretizer(bounds, density, ratio, STACK_SIZE, lons,
-                              sinlats, coslats, radii, result,
-                              kernelxx_variable)
-    else:
-        error = rediscretizer(bounds, density, ratio, STACK_SIZE, lons,
-                              sinlats, coslats, radii, result, kernelxx)
-    return error
-
-
-@cython.boundscheck(False)
-@cython.wraparound(False)
-@cython.cdivision(True)
-cdef inline double kernelxx(
-    double lon, double sinlat, double coslat, double radius, double scale,
-    object density, double[::1] lonc, double[::1] sinlatc, double[::1] coslatc,
-    double[::1] rc):
-    cdef:
-        unsigned int i, j, k
-        double kappa, r_sqr, coslon, l_sqr, cospsi, kphi
-        double result, rc_sqr
-    r_sqr = radius**2
[…]

### 2018-11-16 code/tesseroid_density/tesseroid.py

@@ -99,11 +99,9 @@ from fatiando.constants import SI2MGAL, SI2EOTVOS, MEAN_EARTH_RADIUS, G
 
 RATIO_V = 1
 RATIO_G = 2
-RATIO_GG = 8
 STACK_SIZE = 100
 DELTA_V = 0.2
 DELTA_G = 0.2
-DELTA_GG = 0.01
 
 
 def _check_input(lon, lat, height, model, ratio, njobs, pool):
@@ -577,351 +575,4 @@ def gz(lon, lat, height, model, dens=None, ratio=RATIO_G,
                          ratio=ratio, njobs=njobs, pool=pool,
                          delta=delta)
     result *= SI2MGAL*G
-    return result
-
-
-def gxx(lon, lat, height, model, dens=None, ratio=RATIO_GG,
-        njobs=1, pool=None, delta=DELTA_GG):
-    """
-    Calculate the xx component of the gravity gradient tensor.
-
-    .. warning:: Tesseroids with dimensions < 10 cm will be ignored to avoid
-        numerical errors.
-
-    Implements the method of Uieda et al. (2016).
-
-    Parameters:
-
-    * lon, lat, height : arrays
-        Arrays with the longitude, latitude and height coordinates of the
-        computation points.
-    * model : list of :class:`~fatiando.mesher.Tesseroid`
-        The density model used to calculate the gravitational effect.
-        Tesseroids must have the property ``'density'``. Those that don't have
-        this property will be ignored in the computations. Elements that are
-        None will also be ignored.
-    * dens : float or None
-        If not None, will use this value instead of the ``'density'`` property
-        of the tesseroids. Use this, e.g., for sensitivity matrix building.
-    * ratio : float
-        Will divide each tesseroid until the distance between it and the
-        computation points is < ratio*size of tesseroid. Used to guarantee the
-        accuracy of the numerical integration.
-    * njobs : int
-        Split the computation into *njobs* parts and run it in parallel using
-        ``multiprocessing``. If ``njobs=1`` will run the computation in serial.
-    * pool : None or multiprocessing.Pool object
-        If not None, will use this pool to run the computation in parallel
-        instead of creating a new one. You must still specify *njobs* as the
-        number of processes in the pool. Use this to avoid spawning processes
-        on each call to this functions, which can have significant overhead.
-    * delta : float
-        Explain...
-
-    Returns:
-
-    * res : array
[…]

### 2018-11-16 code/tesseroid_density/_tesseroid.pyx
Fix wrong minimum tesseroid threshold

@@ -147,7 +147,7 @@ cdef inline int divisions(double distance, double Llon, double Llat, double Lr,
         else:
             nlat[0] = 2
     if distance <= ratio*Lr:
-        if Lr <= 1e3:
+        if Lr <= 1e-3:
             error = -1
         else:
             nr[0] = 2

### 2018-11-16 code/tesseroid_density/_tesseroid.pyx
Fix minimum tesseroid radial threshold

@@ -94,7 +94,7 @@ cdef rediscretizer(
                 # Put the nodes in the current range
                 scale = scale_nodes(w, e, s, n, top, bottom, lonc, sinlatc,
                                     coslatc, rc)
-                res += kernel(lon, sinlat, coslat, radius, scale, density, 
+                res += kernel(lon, sinlat, coslat, radius, scale, density,
                               lonc, sinlatc, coslatc, rc)
         result[l] += res
     return error_code
@@ -147,7 +147,7 @@ cdef inline int divisions(double distance, double Llon, double Llat, double Lr,
         else:
             nlat[0] = 2
     if distance <= ratio*Lr:
-        if Lr <= 1e3:
+        if Lr <= 1e-3:
             error = -1
         else:
             nr[0] = 2

### 2018-11-16 code/tesseroid_density/_tesseroid.pyx
Convert 3D adaptive discretization to 2D

@@ -53,7 +53,7 @@ cdef rediscretizer(
         double[::1] coslatc =  numpy.empty(2, numpy.float)
         double[::1] rc =  numpy.empty(2, numpy.float)
         double scale
-        int nlon, nlat, nr, new_cells
+        int nlon, nlat, new_cells
         int stktop, error, error_code
         double[:, ::1] stack =  numpy.empty((STACK_SIZE, 6), numpy.float)
         double w, e, s, n, top, bottom
@@ -82,14 +82,12 @@ cdef rediscretizer(
             distance = distance_n_size(w, e, s, n, top, bottom, lon, sinlat,
                                        coslat, radius, &Llon, &Llat, &Lr)
             # Check which dimensions I have to divide
-            error = divisions(distance, Llon, Llat, Lr, ratio, &nlon,
-                              &nlat, &nr, &new_cells)
+            error = divisions(distance, Llon, Llat, Lr, ratio, &nlon, &nlat, &new_cells)
             error_code += error
             if new_cells > 1:
-                if stktop + nlon*nlat*nr > STACK_SIZE:
+                if stktop + nlon*nlat > STACK_SIZE:
                     raise ValueError('Tesseroid stack overflow')
-                stktop = split(w, e, s, n, top, bottom, nlon, nlat, nr,
-                               stack, stktop)
+                stktop = split(w, e, s, n, top, bottom, nlon, nlat, stack, stktop)
             else:
                 # Put the nodes in the current range
                 scale = scale_nodes(w, e, s, n, top, bottom, lonc, sinlatc,
@@ -129,12 +127,10 @@ cdef inline double distance_n_size(
 @cython.wraparound(False)
 @cython.cdivision(True)
 cdef inline int divisions(double distance, double Llon, double Llat, double Lr,
-                          double ratio, int* nlon, int* nlat, int* nr,
-                          int* new_cells):
+                          double ratio, int* nlon, int* nlat, int* new_cells):
     "How many divisions should be made per dimension"
     nlon[0] = 1
     nlat[0] = 1
-    nr[0] = 1
     error = 0
     if distance <= ratio*Llon:
         if Llon <= 0.1:  # in meters. ~1e-6  degrees
@@ -146,12 +142,7 @@ cdef inline int divisions(double distance, double Llon, double Llat, double Lr,
             error = -1
         else:
             nlat[0] = 2
[…]

### 2018-11-16 code/tesseroid_density/tesseroid.py
Raise default stack size

@@ -99,7 +99,7 @@ from fatiando.constants import SI2MGAL, SI2EOTVOS, MEAN_EARTH_RADIUS, G
 
 RATIO_V = 1
 RATIO_G = 2
-STACK_SIZE = 100
+STACK_SIZE = 200
 DELTA_V = 0.2
 DELTA_G = 0.2
 

### 2018-11-27 code/notebooks/scripts/exponential_density.py
Change exponential density script computation
Now the computation is performed only if the resulting file does not
already exist.

@@ -77,7 +77,7 @@ grids = {"pole": gridder.regular((89, 90, 0, 1), (10, 10), z=0),
 # ---------------------
 fields = 'potential gz'.split()
 density_bottom, density_top = 3300, 2670
-b_factors = [1, 5, 10, 100]
+b_factors = [1, 2, 5, 10, 30, 100]
 delta_values = np.logspace(-3, 1, 9)
 
 
@@ -102,51 +102,50 @@ plt.show()
 
 # Compute differences
 # -------------------
-compute = True
-if compute:
-    for field in fields:
-        for model in models:
-            top, bottom = model.bounds[4], model.bounds[5]
-            thickness = top - bottom
+for field in fields:
+    for model in models:
+        top, bottom = model.bounds[4], model.bounds[5]
+        thickness = top - bottom
 
-            for b_factor in b_factors:
-                amplitude = (density_bottom - density_top) / \
-                    (np.exp(-bottom * b_factor / thickness) -
-                     np.exp(-top * b_factor / thickness))
-                constant_term = density_bottom - amplitude * \
-                    np.exp(-bottom * b_factor / thickness)
+        for b_factor in b_factors:
+            amplitude = (density_bottom - density_top) / \
+                (np.exp(-bottom * b_factor / thickness) -
+                 np.exp(-top * b_factor / thickness))
+            constant_term = density_bottom - amplitude * \
+                np.exp(-bottom * b_factor / thickness)
 
-                # Define density function
-                def density_exponential(height):
-                    return amplitude*np.exp(-height * b_factor / thickness) + \
-                        constant_term
+            # Define density function
+            def density_exponential(height):
+                return amplitude*np.exp(-height * b_factor / thickness) + constant_term
 
-                # Append density function to every tesseroid of the model
-                model.addprop(
-                    "density",
-                    [density_exponential for i in range(model.size)]
-                )
+            # Append density function to every tesseroid of the model
+            model.addprop(
+                "density",
+                [density_exponential for i in range(model.size)]
+            )
 
-                for grid_name, grid in grids.items():
-                    print("Thickness: {} Field: {} Grid: {} b: {}".format(
-                        int(thickness), field, grid_name, b_factor)
-                        )
-                    lats, lons, heights = grid
-                    analytical = shell_exponential_density(heights[0], top, bottom,
-                                                           amplitude, b_factor,
-                                                           constant_term)
-                    differences = []
-                    for delta in delta_values:
[…]

### 2018-11-27 code/notebooks/scripts/grid_search.py
Add grid_search script

@@ -1,0 +1,190 @@
+from __future__ import division, print_function
+import os
+import numpy as np
+from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
+from fatiando.mesher import TesseroidMesh
+from fatiando import gridder
+import matplotlib
+import matplotlib.pyplot as plt
+# This is our custom tesseroid code
+from tesseroid_density import tesseroid
+
+
+def shell_exponential_density(height, top, bottom, amplitude, b_factor, constant_term):
+    """
+    Analytical solution for a spherical shell with inner and outer radii `bottom
+    + MEAN_EARTH_RADIUS` and `top + MEAN_EARTH_RADIUS` at a computation point located at
+    the radial coordinate `height + MEAN_EARTH_RADIUS` with an exponential density as:
+
+    .. math :
+        \rho(r') = A e^{- b * (r' - R) / T} + C
+
+    Where $r$ is the radial coordinate where the density is going to be evaluated,
+    $A$ is the amplitude, $C$ the constant term, $T$ is the thickness of the tesseroid,
+    $b$ the b factor and $R$ the mean Earth radius.
+    """
+    r = height + MEAN_EARTH_RADIUS
+    r1 = bottom + MEAN_EARTH_RADIUS
+    r2 = top + MEAN_EARTH_RADIUS
+    thickness = r2 - r1
+    potential = 4 * np.pi * G * amplitude * thickness / b_factor / r * \
+        (
+         (r1**2 + 2 * r1 * thickness / b_factor + 2 * (thickness / b_factor)**2) *
+         np.exp(-(r1 - MEAN_EARTH_RADIUS) / thickness * b_factor) -
+         (r2**2 + 2 * r2 * thickness / b_factor + 2 * (thickness / b_factor)**2) *
+         np.exp(-(r2 - MEAN_EARTH_RADIUS) / thickness * b_factor)
+        ) + \
+        4/3 * np.pi * G * constant_term * (r2**3 - r1**3) / r
+    data = {'potential': potential,
+            'gx': 0,
+            'gy': 0,
+            'gz': SI2MGAL*(potential/r),
+            'gxx': SI2EOTVOS*(-potential/r**2),
+            'gxy': 0,
+            'gxz': 0,
+            'gyy': SI2EOTVOS*(-potential/r**2),
+            'gyz': 0,
+            'gzz': SI2EOTVOS*(2*potential/r**2)}
+    return data
+
+
[…]

### 2018-11-27 manuscript/manuscript.tex
Add missing citation of Uieda2016

--- 2018-11-15 manuscript/manuscript.tex
@@
 We present a new methodology to compute the gravitational fields generated by
-tesseroids (spherical prisms) whose density varies with depth according to
-an arbitrary continuous function.
+tesseroids (spherical prisms) whose density varies continuously with depth according to
+an arbitrary function.
 It approximates the gravitational fields through the Gauss-Legendre Quadrature along
@@
 for the gravitational potential, acceleration, and gradient tensor components
-by comparing the numerical model with the fields generated by a spherical shell.
-
-Two recent papers present other approaches for homogeneous tesseroids gravitational
-fields computation and incorporate methodologies for variable density tesseroid with
-specific density profiles.
-\citet{Fukushima2018} analytically integrates the volumetric integral for the
-gravitational potential in the radial direction, obtaining a surface integral, which is
-then numerically solved by applying a conditionally splitting and the double exponential
-quadrature rule. The gradient of the potential and the Marussi tensor components are
-then computed by finite differences. \citet{Fukushima2018} also generalizes his method
-to an arbitrary polynomial density profile on the radial coordinate.
-\citet{Lin2018} made a comparison of the different methodologies for homogeneous
-tesseroid gravitational fields computation by applying different discretizations
-algorithms to each one of them and developed a combined method.
-For computation points near the tesseroid they apply a GLQ numerical integration with
-a two dimensional adaptive discretization algorithm, i.e. only in the horizontal
-dimension. But, if the computation point is farther than a certain truncation distance,
-then a second order Taylor series approximation is applied along with the regular
-subdivision developed by \citet{Grombein2013}.
-They also introduce a variation of their combined method to compute the gravitational
-fields generated by tesseroids with linear density on the radial coordinate.
-
-In summary, the bibliography is offering different methodologies for computing
-gravitational fields of homogeneous tesseroids or with certain specific density
-functions in depth.
-Although any density profile could be approximated by one of these specific density
-functions, one should acknowledge the error that such approximation would propagate into
-the gravitational field in order to guarantee certain degree of accuracy.
+by comparing the numerical model with the analytical solution of the fields generated by
+a spherical shell.
 
 We present a new algorithm for computing the gravitational fields generated by any
-tesseroid with any arbitrary continuous density function on any external point.
+tesseroid with an arbitrary continuous density function on any external point.
 It is based on the GLQ approximation and the adaptive discretization of
@@
 
-We define an external computation point $P(r, \phi, \lambda)$ in a geocentric spherical
-coordinate system at a radius $r$, geocentric latitude $\phi$, and longitude $\lambda$
-where the gravitational fields are going to be calculated.
-The first and second derivatives of the gravitational potential are taken with
+Consider an arbitrary tesseroid in a geocentric spherical coordinate system defined by
+pairs of geocentric latitudinal ($\phi_1$, $\phi_2$), longitudinal ($\lambda_1$,
+$\lambda_2$) and radial ($r_1$, $r_2$) boundaries.
+We define an external computation point $P(r, \phi, \lambda)$ located at radius $r$,
+geocentric latitude $\phi$, and longitude $\lambda$ where the gravitational fields the
+tesseroid generates are going to be calculated.
+The first- and second-order derivatives of the gravitational potential are taken with
 respect to the local north-oriented Cartesian coordinate system of $P$
@@
 
-\noindent $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and $W_i^r$ are the
-quadrature weights.
+\noindent $f(r', \phi', \lambda')$ is the non-density dependent portion of the
+integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and
+$W_i^r$, $W_j^\phi$, $W_k^\lambda$ are the quadrature weights.
+The GLQ nodes and weights can be calculated as \citet[p.~391]{Hildebrand1987} (see
+\citet{Uieda2016} for nodes and weights definitions on similar notation as this paper).
 
@@
 \noindent
-Because of the rotational symmetry along the $z$ axe, the integration in $\lambda'$ is
+Because of the rotational symmetry along the $z$ axis, the integration in $\lambda'$ is
 straightforward:
@@
 \begin{equation}
-    g_{xx}(r) = g_{yy}(r) = -\frac{V_\text{sh}(r)}{r^2}
+    g_{xx}(r) = g_{yy}(r) = -\frac{V_\text{sh}(r)}{r^2},
 \end{equation}
@@
 
-From Eq. \ref{eq:shell-pot} we can obtain expressions for gravitational potential for
+From Eq. \ref{eq:shell-pot} we can obtain expressions of the gravitational potential for
 different density functions.

### 2018-11-27 manuscript/manuscript.tex

--- 2018-11-16 manuscript/manuscript.tex
@@
 respectively.
-These values are in agreement with previous results for a homogeneous density tesseroid.
+These values are in agreement with previous results for a homogeneous density tesseroid
+\citep{Uieda2016}.
 The optimum $\delta$ values are 0.2 for the potential and its gradient components, while

### 2018-11-27 manuscript/manuscript.tex
Change introduction and methodology in manuscript

@@
 dividing the tesseroid into smaller ones.
-The first one is a preexisting adaptive discretization algorithm that reduces
-the errors due to the distance between the tesseroid and the computation
-point.
+The first one is a preexisting two dimensional adaptive discretization algorithm that
+reduces the errors due to the distance between the tesseroid and the computation point.
 The second is a new density-based discretization algorithm that
@@
 gets closer to the tesseroid \citep{Ku1977}.
-\citet{Uieda2016} built on the adaptive discretization algorithm of \citet{Li2011}
-to automatically obtain integration results with 0.1\% accuracy.
+\citet{Uieda2016} built on the three dimensional adaptive discretization algorithm of
+\citet{Li2011} to automatically obtain integration results with 0.1\% accuracy.
 The algorithm consist in recursively splitting the tesseroid into smaller ones when a
@@
 tesseroid with any arbitrary continuous density function on any external point.
-It is based on the GLQ approximation and the adaptive discretization of
-\citet{Uieda2016}, which we extend to include a new density-based discretization step.
+It is based on the GLQ approximation and a two dimensional version of the adaptive
+discretization of \citet{Uieda2016}, which we extend to include a new density-based
+discretization algorithm.
 In order to ensure the accuracy of the numerical approximation we have
@@
 
-\subsection{Adaptive Discretization}
+\subsection{Two Dimensional Adaptive Discretization}
 
@@
 of point masses is produced where it is needed more.
-\citet{Uieda2016} developed a modified version of this algorithm, which we will use
-here.
+\citet{Uieda2016} developed a modified version of this algorithm that speeds the
+computation of the distances.
+Both algorithms (the original one from \citet{Li2011} and the modified
+version of \citet{Uieda2016}) performs tesseroid subdivisions in the latitudinal,
+longitudinal and radial directions, thus we can define them as three dimensional
+adaptive discretization algorithms.
+On the other hand, \citet{Lin2018} proposed a two dimensional discretization algorithm
+that subdivides the tesseroid only on the latitudinal and longitudinal directions.
+Removing a dimension from the splitting makes the computation more efficient, while
+keeping an acceptable accuracy on the gravitational fields \citep{Lin2018}.
+
+Here we will use a two dimensional version of the adaptive discretization algorithm
+developed by \citet{Uieda2016}.
 What follows is a summary of the algorithm and the reader is referred to
@@
 
-The following algorithm computes the gravitational fields at a given point at $(r, \phi,
-\lambda)$ due to a tesseroid with its geometric centre at $(r_t, \phi_t, \lambda_t)$ and
-dimensions $L_r$, $L_\phi$, and $L_\lambda$ given by
+Lets consider a computation point located at $(r, \phi, \lambda)$ and an arbitrary
+tesseroid with its geometric centre at $(r_t, \phi_t, \lambda_t)$.
+We can define its longitudinal ($L_\lambda$) and latitudinal ($L_\phi$) dimensions as
+follows:
 
@@
 \begin{equation}
-    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1),
-\end{equation}
-
-and
-
-\begin{equation}
-    L_r = r_2 - r_1.
-    \label{eq:sizer}
+    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1).
 \end{equation}
@@
 \textit{Step 1}: Check that the tesseroid satisfies the following inequality for each
-dimension $L_i$ of the tesseroid:
+dimension $L_i (i \in \{\lambda, \phi\})$ of the tesseroid:
 
@@
 with density varying according to exponential and linear functions of $r$.
+This numerical experiment will also show that the two dimensional adaptive
+discretization algorithm is as accurate as the original three dimensional version,
+though it reduces the number of tesseroid subdivisions and thus the computation time.
 
@@
 the integration error from using only a few nodes to discretize the density function.
-The adaptive discretization may help
-to reduce this kind of error.
+The three dimensional adaptive discretization may help
+to reduce this kind of error by adding more point masses in the radial direction.
 However, it does not take into account the density function
 and hence it is not well suited to fully perform this task.
+Moreover, because the two dimensional adaptive discretization algorithm does not
+subdivide the tesseroid on the radial dimension, it does not put any effort on lowering
+this new kind of error.
 
@@
 in which $L_r$ is the radial dimension of the tesseroid being considered for division,
-$L_r^\text{orig}$ is the radial dimension of the original tesseroid, and
+
+\begin{equation}
+    L_r = r_2 - r_1,
+\end{equation}
+
+\noindent $L_r^\text{orig}$ is the radial dimension of the original tesseroid, and
 $\delta$ is a positive constant henceforth called the \textit{delta ratio}.
@@
 ensures an acceptable accuracy while minimising the computation time.
+
+\subsection{Algorithm summary}
+
+In summary, given any tesseroid with variable density in depth and an arbitrary
+external computation point, we propose the following steps to numerically
+compute the gravitational fields:
+
+\textit{Step 1:}
+Apply the density-based discretization algorithm.
+It subdivides the tesseroid only in the radial dimension, thus it will return a set of
+tesseroids with the same longitudinal and latitudinal dimensions as the original one but
+with different radial boundaries.
+
+\textit{Step 2:}
+Apply the two dimensional adaptive discretization algorithm on each tesseroid obtained
+in the previous step.
+If needed, it will subdivide each tesseroid only in the latitudinal and longitudinal
+direction, generating a set of smaller tesseroids that resemble the original one.
+
+\textit{Step 3:}
+Apply a second-order GLQ to numerically compute the gravitational fields
+(Eq.\ref{eq:glq-var-dens}) generated by each tesseroid of the set obtained in the
+previous step.
+Then sum these results in order to obtain the gravitational field of the original
+tesseroid.
 

### 2018-11-28 manuscript/manuscript.tex
Edit in determination of D and delta

@@
 have the same density function of radius as our tesseroid model.
-Analytical solutions for a variable density spherical shell are not present in the
-literature and so they must be obtained first.
-We derive the expressions for the gravitational fields of spherical shells with linear
-and exponential density functions of radius in Appendix \ref{sec:shell}.
+\citet{Lin2018} show the analytical solution of the gravitational potential generated by
+a spherical shell with linear density in the radial coordinate.
+We derive an expression for such gravitational potential in case the spherical shell has
+an arbitrary density function in depth (Eq. \ref{eq:shell-pot} in Appendix
+\ref{sec:shell}), and obtained the expressions in case of linear and exponential
+density.
 

### 2018-11-28 manuscript/manuscript.tex
Fix spherical shell analytical solutions for exp density

@@
 different density functions.
+
+\subsection{Linear density}
+
 For a linear density function
@@
 
-An exponential density function that assumes the values of $\rho_\text{out}$ and
-$\rho_\text{in}$ on the shell's outer and inner surfaces, respectively, can be defined
-as follows:
-
-\begin{equation}
-    \rho(r') = A e^{-(r' - R)/b} + C,
+\subsection{Exponential density}
+
+For an exponential density function
+
+\begin{equation}
+    \rho(r') = A e^{- k (r' - R)},
+\end{equation}
+
+\noindent
+where $A$, $k$ and $R$ are constants, the gravitational potential on any external point
+is
+
+\begin{equation}
+    \begin{split}
+    V_\text{exp}(r) = \frac{4\pi G}{r} \frac{A}{k^3} \Big[
+    & \left( R_1^2 k^2 + 2 R_1 k + 2 \right) e^{- k (R_1 - R)} - \\
+    & \left( R_2^2 k^2 + 2 R_2 k + 2 \right) e^{- k (R_2 - R)}
+    \Big].
+    \end{split}
+\end{equation}
+
+Moreover, an exponential density function that assumes the values of $\rho_\text{out}$
+and $\rho_\text{in}$ on the shell's outer and inner surfaces, respectively, can be
+defined as follows:
+
+\begin{equation}
+    \rho(r') = A e^{- \frac{b}{T} (r' - R)} + C,
 \end{equation}
@@
 \begin{equation}
-    A = (\rho_\text{in} - \rho_\text{out})
-        \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
-\end{equation}
-
-\begin{equation}
-    C = \rho_\text{out} - A,
-\end{equation}
-
-\noindent $R$ is the mean Earth radius, and $b$ is a constant
-that determines the variability of the function: a low value of $b$
-increases the maximum slope of the density function.
+    A = \frac{\rho_\text{in} - \rho_\text{out}}{D},
+\end{equation}
+
+\begin{equation}
+    C = \frac{1}{D}
+    \left[\rho_\text{out} \, e^{- \frac{b}{T} (R_1 - R)} -
+    \rho_\text{in} \, e^{- \frac{b}{T} (R_2 - R)} \right],
+\end{equation}
+
+\begin{equation}
+    D = e^{- \frac{b}{T} (R_1 - R)} - e^{- \frac{b}{T} (R_2 - R)},
+\end{equation}
+
+\noindent $R$ is the mean Earth radius, $T$ is the thickness of the shell ($T = R_2
+- R_1$) and $b$ is a dimensionless constant that determines the variability of the
+function: a high value of $b$ increases the maximum slope of the density function.
 
 The analytical solution of the gravitational potential generated by a
-spherical shell with a density function is
+spherical shell with such density function is
 
@@
     \begin{split}
-        V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r}
-        Ab
-        \Big[
-        & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - R}{b}} - \\
-        & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - R}{b}}
+        V_\text{exp}(r) = & \frac{4\pi G}{r} \frac{A T^3}{b^3} \\ \Big[
+        & \left( \left( b R_1/T \right)^2 + 2 b R_1/T + 2 \right)
+        e^{- \frac{b}{T} (R_1 - R)} - \\
+        & \left( \left( b R_2/T \right)^2 + 2 b R_2/T + 2 \right)
+        e^{- \frac{b}{T} (R_2 - R)}
         \Big] + \\
-        & \frac{4 \pi G}{3 r} C (R_2^3 - R_1^3).
+        & \frac{4\pi G}{3 r} C \left( R_2^3 - R_1^3 \right).
     \end{split}

### 2018-11-28 manuscript/manuscript.tex
Fix line in determination of D and delta
Add tesseroid models table

@@
 We derive an expression for such gravitational potential in case the spherical shell has
-an arbitrary density function in depth (Eq. \ref{eq:shell-pot} in Appendix
-\ref{sec:shell}), and obtained the expressions in case of linear and exponential
-density.
+an arbitrary density function in depth (Eq. \ref{eq:shell-pot}), and obtained the
+expressions in case of linear and exponential density (see Appendix \ref{sec:shell}).
 
@@
 numerical error lower than 0.1\% of the spherical shell values.
+
+\begin{table}
+\caption{
+    Description of the tesseroid models used to build a spherical shell and characterize
+    the accuracy of the numerical integration.
+    The first three columns describe the shapes of the tesseroid models: each one
+    specifies the number of tesseroids on the corresponding dimension (radial,
+    latitudinal, longitudinal).
+    The last two columns specify the latitudinal and longitudinal size of each
+    tesseroid that compose the spherical shell model.
+    For each shape, we define several spherical shell models which the following
+    thicknesses: 100m, 1km, 10km, 100km and 1000km.
+    While the outer radius of every shell is equal to the mean Earth radius
+}
+\label{tab:tess-models}
+\begin{tabular}{cccccc}
+    Radial & Latitudinal & Longitudinal  & Total tesseroids & $L_\phi$ & $L_\lambda$
+    \\ \hline
+    1 & 1  & 2  & 2   & 180$^\circ$ & 180$^\circ$ \\
+    1 & 6  & 12 & 72  &  30$^\circ$ &  30$^\circ$ \\
+    1 & 12 & 24 & 288 &  15$^\circ$ &  15$^\circ$ \\
+    1 & 18 & 36 & 648 &  10$^\circ$ &  10$^\circ$ \\
+\end{tabular}
+\end{table}
 

### 2018-11-28 code/notebooks/scripts/grid_search.py
Edit grids shapes in scripts

@@ -67,8 +67,8 @@ models = [
 
 # Define computation grids
 # ------------------------
-grids = {"pole": gridder.regular((89, 90, 0, 1), (10, 10), z=0),
-         "equator": gridder.regular((0, 1, 0, 1), (10, 10), z=0),
+grids = {"pole": gridder.regular((89, 90, 0, 1), (11, 11), z=0),
+         "equator": gridder.regular((0, 1, 0, 1), (11, 11), z=0),
          "global": gridder.regular((-90, 90, 0, 360), (19, 13), z=0),
          "260km": gridder.regular((-90, 90, 0, 360), (19, 13), z=260e3),
          }

### 2018-11-28 code/notebooks/scripts/linear_density.py

@@ -50,8 +50,8 @@ models = [
 
 # Define computation grids
 # ------------------------
-grids = {"pole": gridder.regular((89, 90, 0, 1), (10, 10), z=0),
-         "equator": gridder.regular((0, 1, 0, 1), (10, 10), z=0),
+grids = {"pole": gridder.regular((89, 90, 0, 1), (11, 11), z=0),
+         "equator": gridder.regular((0, 1, 0, 1), (11, 11), z=0),
          "global": gridder.regular((-90, 90, 0, 360), (19, 13), z=0),
          "260km": gridder.regular((-90, 90, 0, 360), (19, 13), z=260e3),
          }

### 2018-11-29 code/notebooks/scripts/exponential_density.py
Fix exponential density analytical solution

@@ -19,22 +19,21 @@ def shell_exponential_density(height, top, bottom, amplitude, b_factor, constant
     .. math :
         \rho(r') = A e^{- b * (r' - R) / T} + C
 
-    Where $r$ is the radial coordinate where the density is going to be evaluated,
+    Where $r'$ is the radial coordinate where the density is going to be evaluated,
     $A$ is the amplitude, $C$ the constant term, $T$ is the thickness of the tesseroid,
     $b$ the b factor and $R$ the mean Earth radius.
     """
     r = height + MEAN_EARTH_RADIUS
     r1 = bottom + MEAN_EARTH_RADIUS
     r2 = top + MEAN_EARTH_RADIUS
-    thickness = r2 - r1
-    potential = 4 * np.pi * G * amplitude * thickness / b_factor / r * \
+    thickness = float(top - bottom)
+    k = b_factor / thickness
+    potential = 4 * np.pi * G * amplitude / k**3 / r * \
         (
-         (r1**2 + 2 * r1 * thickness / b_factor + 2 * (thickness / b_factor)**2) *
-         np.exp(-(r1 - MEAN_EARTH_RADIUS) / thickness * b_factor) -
-         (r2**2 + 2 * r2 * thickness / b_factor + 2 * (thickness / b_factor)**2) *
-         np.exp(-(r2 - MEAN_EARTH_RADIUS) / thickness * b_factor)
+         ((r1 * k)**2 + 2*r1*k + 2) * np.exp(-k * bottom) -
+         ((r2 * k)**2 + 2*r2*k + 2) * np.exp(-k * top)
         ) + \
-        4/3 * np.pi * G * constant_term * (r2**3 - r1**3) / r
+        4 / 3. * np.pi * G * constant_term * (r2**3 - r1**3) / r
     data = {'potential': potential,
             'gx': 0,
             'gy': 0,
@@ -86,9 +85,13 @@ delta_values = np.logspace(-3, 1, 9)
 bottom, top = 0, 1
 thickness = top - bottom
 for b_factor in b_factors:
-    amplitude = (density_bottom - density_top) / \
-        (np.exp(-bottom * b_factor / thickness) - np.exp(-top * b_factor / thickness))
-    constant_term = density_bottom - amplitude * np.exp(-bottom * b_factor / thickness)
+    denominator = np.exp(- bottom * b_factor / thickness) - \
+                  np.exp(- top * b_factor / thickness)
[…]

### 2018-11-29 code/notebooks/scripts/linear_density.py
Simplify linear density result dir name

@@ -32,7 +32,7 @@ def shell_linear_density(height, top, bottom, slope, constant_term):
 
 # Create results dir if it does not exist
 # ---------------------------------------
-result_dir = 'results/linear-D'
+result_dir = 'results/linear'
 if not os.path.isdir(result_dir):
     os.makedirs(result_dir)
 

### 2018-11-29 code/notebooks/scripts/linear_density.py
Ignore computation if outfile exists in linear density

@@ -60,41 +60,41 @@ grids = {"pole": gridder.regular((89, 90, 0, 1), (11, 11), z=0),
 # Compute differences
 # -------------------
 fields = 'potential gz'.split()
-compute = False
-if compute:
-    D_values = np.arange(0.5, 5.5, 0.5)
-    for field in fields:
-        for model in models:
-            top, bottom = model.bounds[4], model.bounds[5]
-            slope = -(3300 - 2670) / (top - bottom)
-            constant_term = (3300 - 2670) * MEAN_EARTH_RADIUS + 2670
+D_values = np.arange(0.5, 5.5, 0.5)
+for field in fields:
+    for model in models:
+        top, bottom = model.bounds[4], model.bounds[5]
+        slope = -(3300 - 2670) / (top - bottom)
+        constant_term = (3300 - 2670) * MEAN_EARTH_RADIUS + 2670
 
-            # Define density function
-            def density_linear(height):
-                r = height + MEAN_EARTH_RADIUS
-                return slope*r + constant_term
+        # Define density function
+        def density_linear(height):
+            r = height + MEAN_EARTH_RADIUS
+            return slope*r + constant_term
 
-            model.addprop("density", [density_linear for i in range(model.size)])
+        model.addprop("density", [density_linear for i in range(model.size)])
 
-            for grid_name, grid in grids.items():
-                print("Thickness: {} Model size: {} Field: {} Grid: {}".format(
-                    int(top - bottom), model.size, field, grid_name)
-                    )
-                lats, lons, heights = grid
-                analytical = shell_linear_density(heights[0], top, bottom,
-                                                  slope, constant_term)
-                differences = []
-                for D in D_values:
-                    result = getattr(tesseroid, field)(lons, lats, heights, model,
-                                                       ratio=D, delta=None)
-                    diff = np.abs((result - analytical[field]) / analytical[field])
-                    diff = 100 * np.max(diff)
-                    differences.append(diff)
-                differences = np.array(differences)
-                fname = "{}-{}-{}-{}".format(field, grid_name, int(top - bottom),
-                                             model.size)
-                np.savez(os.path.join(result_dir, fname),
-                         D_values=D_values, differences=differences)
+        for grid_name, grid in grids.items():
+            fname = "{}-{}-{}-{}".format(field, grid_name, int(top - bottom),
+                                         model.size)
+            if os.path.isfile(os.path.join(result_dir, fname)):
+                continue
+            print("Thickness: {} Model size: {} Field: {} Grid: {}".format(
[…]

### 2018-11-29 code/notebooks/scripts/grid_search.py
Fix analytical solution in grid_search

@@ -19,22 +19,21 @@ def shell_exponential_density(height, top, bottom, amplitude, b_factor, constant
     .. math :
         \rho(r') = A e^{- b * (r' - R) / T} + C
 
-    Where $r$ is the radial coordinate where the density is going to be evaluated,
+    Where $r'$ is the radial coordinate where the density is going to be evaluated,
     $A$ is the amplitude, $C$ the constant term, $T$ is the thickness of the tesseroid,
     $b$ the b factor and $R$ the mean Earth radius.
     """
     r = height + MEAN_EARTH_RADIUS
     r1 = bottom + MEAN_EARTH_RADIUS
     r2 = top + MEAN_EARTH_RADIUS
-    thickness = r2 - r1
-    potential = 4 * np.pi * G * amplitude * thickness / b_factor / r * \
+    thickness = float(top - bottom)
+    k = b_factor / thickness
+    potential = 4 * np.pi * G * amplitude / k**3 / r * \
         (
-         (r1**2 + 2 * r1 * thickness / b_factor + 2 * (thickness / b_factor)**2) *
-         np.exp(-(r1 - MEAN_EARTH_RADIUS) / thickness * b_factor) -
-         (r2**2 + 2 * r2 * thickness / b_factor + 2 * (thickness / b_factor)**2) *
-         np.exp(-(r2 - MEAN_EARTH_RADIUS) / thickness * b_factor)
+         ((r1 * k)**2 + 2*r1*k + 2) * np.exp(-k * bottom) -
+         ((r2 * k)**2 + 2*r2*k + 2) * np.exp(-k * top)
         ) + \
-        4/3 * np.pi * G * constant_term * (r2**3 - r1**3) / r
+        4 / 3. * np.pi * G * constant_term * (r2**3 - r1**3) / r
     data = {'potential': potential,
             'gx': 0,
             'gy': 0,
@@ -88,9 +87,13 @@ D_values = np.arange(0.5, 4.5, 0.5)
 bottom, top = 0, 1
 thickness = top - bottom
 for b_factor in b_factors:
-    amplitude = (density_bottom - density_top) / \
-        (np.exp(-bottom * b_factor / thickness) - np.exp(-top * b_factor / thickness))
-    constant_term = density_bottom - amplitude * np.exp(-bottom * b_factor / thickness)
+    denominator = np.exp(- bottom * b_factor / thickness) - \
+                  np.exp(- top * b_factor / thickness)
[…]

### 2018-12-10 code/scripts/exponential_density.py
Add gridspec for exponential density figure

@@ -5,6 +5,7 @@ from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
 from fatiando.mesher import TesseroidMesh
 from fatiando import gridder
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec, GridSpecFromSubplotSpec
 # This is our custom tesseroid code
 from tesseroid_density import tesseroid
 
@@ -166,13 +167,47 @@ grid_titles = {"pole": "Pole",
                "260km": "Satellite"}
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
+# Create outer grid
+fig = plt.figure(figsize=(7, 8))
+outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
+
+# Create grid specs for each grid
+grid_specs = dict(
+                  zip(
+                      grids.keys(),
+                      [GridSpecFromSubplotSpec(ncols=1, nrows=2,
+                                               subplot_spec=outer_grid[j, i],
+                                               hspace=0)
+                       for j in range(2)
+                       for i in range(2)]
+                     )
+                 )
+positions = dict(zip(grids.keys(), [[j, i] for j in range(2) for i in range(2)]))
+
+# Plot for each grid
 for grid_name in grids:
 
-    fig, axes = plt.subplots(nrows=len(fields), ncols=1, sharex=True)
-    fig.set_size_inches((5, 5))
-    fig.subplots_adjust(hspace=0)
+    # Choose gridspec and create axes
+    gs = grid_specs[grid_name]
+    position = positions[grid_name]
+    axes = [plt.subplot(gs[0, 0])]
+    axes.append(plt.subplot(gs[1, 0], sharex=axes[0]))
+    plt.setp(axes[0].get_xticklabels(), visible=False)
+    # Remove xticks if the axes in in the middle of the figure
+    if position[0] == 0:
+        plt.setp(axes[1].get_xticklabels(), visible=False)
+    else:
+        axes[1].set_xlabel(r"$\delta$")
+    # Move ticks and labels to the right if the axes are on the second column
+    if position[1] == 1:
+        for ax in axes:
+            ax.yaxis.tick_right()
+            ax.yaxis.set_label_position("right")
+    # Add title to axes
[…]

### 2018-12-10 code/scripts/exponential_density.py
Add exponential density figure

@@ -160,6 +160,8 @@ plt.show()
 
 # Plot Results
 # ------------
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/exponential-density.pdf")
 field_titles = dict(zip(fields, '$V$ $g_z$'.split()))
 grid_titles = {"pole": "Pole",
                "equator": "Equator",
@@ -251,5 +253,5 @@ for grid_name in grids:
         axes[0].legend(loc=0, prop={"size": 8})
 
 outer_grid.tight_layout(fig)
-plt.savefig("exponential_results.pdf", dpi=300)
+plt.savefig(figure_fname, dpi=300)
 plt.show()

### 2018-12-11 code/scripts/sine_density.py
Add gridspec to sine density script and figure

@@ -5,6 +5,7 @@ from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
 from fatiando.mesher import TesseroidMesh
 from fatiando import gridder
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec, GridSpecFromSubplotSpec
 # This is our custom tesseroid code
 from tesseroid_density import tesseroid
 
@@ -75,29 +76,6 @@ b_factors = [1, 2, 5, 10]
 delta_values = np.logspace(-4, 0, 5)
 
 
-# Plot Densities
-# --------------
-bottom, top = 0, 1
-thickness = top - bottom
-colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
-for b_factor in b_factors:
-
-    # Compute the k factor, i.e. the frequency for the sine function
-    k_factor = 2 * np.pi * b_factor / thickness
-
-    # Define density function
-    def density_sine(height):
-        return max_density * np.sin(k_factor * height) + max_density
-
-    heights = np.linspace(bottom, top, 101)
-    plt.plot(heights, density_sine(heights), color=colors[b_factor],
-             label="b={}".format(b_factor))
-plt.ylabel(r"Density [kg/m$^3$]")
-plt.xticks([bottom, top], ["Inner Radius", "Outer Radius"])
-plt.legend()
-plt.show()
-
-
 # Compute differences
 # -------------------
 for field in fields:
@@ -143,8 +121,33 @@ for field in fields:
                          delta_values=delta_values, differences=differences)
 
 
+# Plot Densities
+# --------------
+bottom, top = 0, 1
+thickness = top - bottom
+colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
+for b_factor in b_factors:
+
+    # Compute the k factor, i.e. the frequency for the sine function
+    k_factor = 2 * np.pi * b_factor / thickness
+
+    # Define density function
+    def density_sine(height):
+        return max_density * np.sin(k_factor * height) + max_density
+
+    heights = np.linspace(bottom, top, 101)
+    plt.plot(heights, density_sine(heights), color=colors[b_factor],
+             label="b={}".format(b_factor))
+plt.ylabel(r"Density [kg/m$^3$]")
+plt.xticks([bottom, top], ["Inner Radius", "Outer Radius"])
+plt.legend()
+plt.show()
+
+
[…]

### 2018-12-11 code/scripts/sine_density.py
Add sine densities figure

@@ -90,7 +90,7 @@ for field in fields:
 
             # Define density function
             def density_sine(height):
-                return max_density * np.sin(k_factor * height) + max_density
+                return max_density / 2 * np.sin(k_factor * height) + max_density / 2
 
             # Append density function to every tesseroid of the model
             model.addprop(
@@ -108,7 +108,7 @@ for field in fields:
                     )
                 lats, lons, heights = grid
                 analytical = shell_sine_density(heights[0], top, bottom,
-                                                max_density, k_factor, max_density)
+                                                max_density/2, k_factor, max_density/2)
                 differences = []
                 for delta in delta_values:
                     result = getattr(tesseroid, field)(lons, lats, heights, model,
@@ -126,28 +126,36 @@ for field in fields:
 bottom, top = 0, 1
 thickness = top - bottom
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
-for b_factor in b_factors:
+
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/sine-densities.pdf")
+fig, axes = plt.subplots(len(b_factors), 1, figsize=(6, 6), sharex=True)
+for b_factor, ax in zip(b_factors, axes):
 
     # Compute the k factor, i.e. the frequency for the sine function
     k_factor = 2 * np.pi * b_factor / thickness
 
     # Define density function
     def density_sine(height):
-        return max_density * np.sin(k_factor * height) + max_density
+        return max_density / 2 * np.sin(k_factor * height) + max_density / 2
 
-    heights = np.linspace(bottom, top, 101)
-    plt.plot(heights, density_sine(heights), color=colors[b_factor],
-             label="b={}".format(b_factor))
-plt.ylabel(r"Density [kg/m$^3$]")
-plt.xticks([bottom, top], ["Inner Radius", "Outer Radius"])
-plt.legend()
+    heights = np.linspace(bottom, top, 1001)
+    ax.plot(heights, density_sine(heights), color=colors[b_factor],
+            label="b={}".format(b_factor))
+    ax.set_xticks([bottom, top])
+    ax.set_xticklabels(["Inner Radius", "Outer Radius"])
+    ax.legend(loc=1)
+fig.text(0, 0.5, r"Density [kg/m$^3$]", va='center', rotation='vertical')
+plt.tight_layout(pad=1.8)
[…]

### 2018-12-11 code/scripts/exponential_density.py
Add exponential density figure

@@ -136,6 +136,10 @@ for field in fields:
 bottom, top = 0, 1
 thickness = top - bottom
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
+
+fig, ax = plt.subplots(figsize=(6, 4))
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/exponential-densities.pdf")
 for b_factor in b_factors:
     denominator = np.exp(- bottom * b_factor / thickness) - \
                   np.exp(- top * b_factor / thickness)
@@ -150,18 +154,21 @@ for b_factor in b_factors:
         return amplitude*np.exp(-height * b_factor / thickness) + constant_term
 
     heights = np.linspace(bottom, top, 101)
-    plt.plot(heights, density_exponential(heights), color=colors[b_factor],
-             label="b={}".format(b_factor))
-plt.ylabel(r"Density [kg/m$^3$]")
-plt.xticks([bottom, top], ["Inner Radius", "Outer Radius"])
-plt.legend()
+    ax.plot(heights, density_exponential(heights), color=colors[b_factor],
+            label="b={}".format(b_factor))
+ax.set_ylabel(r"Density [kg/m$^3$]")
+ax.set_xticks([bottom, top])
+ax.set_xticklabels(["Radius", "Outer Radius"])
+ax.legend()
+plt.tight_layout()
+plt.savefig(figure_fname, dpi=300)
 plt.show()
 
 
 # Plot Results
 # ------------
 figure_fname = os.path.join(script_path,
-                            "../../manuscript/figures/exponential-density.pdf")
+                            "../../manuscript/figures/exponential-density-diffs.pdf")
 field_titles = dict(zip(fields, '$V$ $g_z$'.split()))
 grid_titles = {"pole": "Pole",
                "equator": "Equator",

### 2018-12-11 code/scripts/linear_density.py
Change line density figure name

@@ -164,7 +164,8 @@ for grid_name in grids:
 
 # Plot single plot: one line per grid
 # -----------------------------------
-figure_fname = os.path.join(script_path, "../../manuscript/figures/linear-density.pdf")
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/linear-density-diffs.pdf")
 field_titles = dict(zip(fields, '$V$ $g_z$'.split()))
 grid_titles = {"pole": "Pole",
                "equator": "Equator",

### 2018-12-11 manuscript/manuscript.tex
Update grid table
Small improvements to introduction

@@
 tesseroid gravitational fields computation by applying different discretizations
-algorithms to each one of them and developed a combined method.
-For computation points near the tesseroid they apply a GLQ numerical integration with
+algorithms to each one of them.
+From this analysis they develop a combined method:
+for computation points near the tesseroid they apply a GLQ numerical integration with
 a two dimensional adaptive discretization algorithm, i.e. only in the horizontal
@@
 tesseroid with any arbitrary continuous density function on any external point.
-It is based on the GLQ approximation and a two dimensional version of the adaptive
-discretization of \citet{Uieda2016}, which we extend to include a new density-based
-discretization algorithm.
+It is based on the three dimensional GLQ approximation, a two dimensional version of the
+adaptive discretization of \citet{Uieda2016} and a new density-based discretization
+algorithm that we introduce in this work.
 In order to ensure the accuracy of the numerical approximation we have
@@
 Finally, we applied the methodology to model the Neuqu\'en basin, Argentina, using
-tesseroids with an exponentially increasing density with depth.
+tesseroids with linear and exponentially increasing density with depth.
 
@@
     numerical integration.
-    All grids consist of a set of $10 \times 10$ points.
+    Grid height is defined above the mean Earth radius.
 }
@@
     \\ \hline
-    Pole & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 2 \\
-    Equator & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/0\text{N}/1\text{N}$ & 2 \\
-    Satellite & $1^\circ \times 1^\circ$ & $0\text{E}/1\text{E}/89\text{N}/90\text{N}$ & 260 \\
-    Big Grid & $30^\circ \times 30^\circ$ & $0\text{E}/30\text{E}/60\text{N}/90\text{N}$ & 2 \\
+    Pole      & $0.1^\circ \times 0.1^\circ$ & 0E / 1E / 89N / 90N  & 0   \\
+    Equator   & $0.1^\circ \times 0.1^\circ$ & 0E / 1E / 0N  / 1N   & 0   \\
+    Global    & $ 10^\circ \times  10^\circ$ & 0E / 1E / 0N  / 1N   & 0   \\
+    Satellite & $ 10^\circ \times  10^\circ$ & 0E / 1E / 89N / 90N  & 260 \\
 \end{tabular}

### 2018-12-11 code/tesseroid_density/tesseroid.py
Updated default DELTA to 1e-2

@@ -100,8 +100,7 @@ from fatiando.constants import SI2MGAL, MEAN_EARTH_RADIUS, G
 RATIO_V = 1
 RATIO_G = 2.5
 STACK_SIZE = 200
-DELTA_V = 0.2
-DELTA_G = 0.2
+DELTA = 1e-2
 
 
 def _check_input(lon, lat, height, model, ratio, njobs, pool):
@@ -342,7 +341,7 @@ def _split_arrays(arrays, extra_args, nparts):
 
 
 def potential(lon, lat, height, model, dens=None, ratio=RATIO_V,
-              njobs=1, pool=None, delta=DELTA_V):
+              njobs=1, pool=None, delta=DELTA):
     """
     Calculate the gravitational potential due to a tesseroid model.
 
@@ -400,7 +399,7 @@ def potential(lon, lat, height, model, dens=None, ratio=RATIO_V,
 
 
 def gx(lon, lat, height, model, dens=None, ratio=RATIO_G,
-       njobs=1, pool=None, delta=DELTA_G):
+       njobs=1, pool=None, delta=DELTA):
     """
     Calculate the North component of the gravitational attraction.
 
@@ -458,7 +457,7 @@ def gx(lon, lat, height, model, dens=None, ratio=RATIO_G,
 
 
 def gy(lon, lat, height, model, dens=None, ratio=RATIO_G,
-       njobs=1, pool=None, delta=DELTA_G):
+       njobs=1, pool=None, delta=DELTA):
     """
     Calculate the East component of the gravitational attraction.
 
@@ -516,7 +515,7 @@ def gy(lon, lat, height, model, dens=None, ratio=RATIO_G,
 
 
 def gz(lon, lat, height, model, dens=None, ratio=RATIO_G,
-       njobs=1, pool=None, delta=DELTA_G):
+       njobs=1, pool=None, delta=DELTA):
     """
     Calculate the radial component of the gravitational attraction.
 

### 2018-12-11 manuscript/manuscript.tex
Minor improvements in methodology

@@
 
+\noindent and
+
 \begin{equation}
@@
 
-\noindent and
-
-\begin{equation}
-    g_{\alpha\beta}(r,\phi,\lambda) = G
-    \int\limits_{\lambda_1}^{\lambda_2}
-    \int\limits_{\phi_1}^{\phi_2}
-    \int\limits_{r_1}^{r_2}
-    \rho(r') I_{\alpha\beta} \, \kappa \, dr' d\phi' d\lambda' ,
-    \label{eq:tesseroid-tensor}
-\end{equation}
-
-\noindent in which
-
-\begin{equation}
-    I_{\alpha\beta} =
-    \left(
-        \frac{3\Delta_{\alpha} \Delta_{\beta}}{\ell^5} -
-        \frac{\delta_{\alpha\beta}}{\ell^3}
-    \right) ,
-    \label{eq:tesseroid-tensor-kernel}
-\end{equation}
-
-\noindent $\alpha, \beta \in \{x, y, z\}$, $\delta_{\alpha\beta}$ is Kronecker's delta,
-$G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant
-and
+\noindent in which $\alpha, \beta \in \{x, y, z\}$, $\delta_{\alpha\beta}$ is
+Kronecker's delta, $G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the
+gravitational constant and
 
@@
 Applying a $N$th order GLQ, we can approximate each integral in equations
-\ref{eq:tesseroid-pot}, \ref{eq:tesseroid-grav} and \ref{eq:tesseroid-tensor} by a
-weighted sum of the integration kernel evaluated on the roots of an $N$th order Legendre
-polynomial \citep[p.~390]{Hildebrand1987}.
+\ref{eq:tesseroid-pot} and \ref{eq:tesseroid-grav} by a weighted sum of the integration
+kernel evaluated on the roots of an $N$th order Legendre polynomial
+\citep[p.~390]{Hildebrand1987}.
 Unlike the homogeneous density case, the radial density function $\rho(r)$ must also be
@@
 The GLQ nodes and weights can be calculated as \citet[p.~391]{Hildebrand1987} (see
-\citet{Uieda2016} for nodes and weights definitions on similar notation as this paper).
+\citet{Uieda2016} for nodes and weights definitions on similar notation used on previous
+equations).
+
+It's worth noting that a GLQ is equivalent to approximate the tesseroid by $N^r \times
+N^\phi \times N^\lambda$ point masses located on the quadrature nodes
+\citep{Ku1977, Li2011, Uieda2016}.
 
@@
 Both algorithms (the original one from \citet{Li2011} and the modified
-version of \citet{Uieda2016}) performs tesseroid subdivisions in the latitudinal,
+version of \citet{Uieda2016}) perform tesseroid subdivisions in the latitudinal,
 longitudinal and radial directions, thus we can define them as three dimensional
@@
 Here we will use a two dimensional version of the adaptive discretization algorithm
-developed by \citet{Uieda2016}.
+developed by \citet{Uieda2016}: the distances from the tesseroid and the computation
+point are computed as \citet{Uieda2016} do, but the tesseroid will only be split on the
+longitudinal and latitudinal directions.
 What follows is a summary of the algorithm and the reader is referred to
@@
 Lets consider a computation point located at $(r, \phi, \lambda)$ and an arbitrary
-tesseroid with its geometric centre at $(r_t, \phi_t, \lambda_t)$.
-We can define its longitudinal ($L_\lambda$) and latitudinal ($L_\phi$) dimensions as
-follows:
+tesseroid defined by the boundaries $r_1$, $r_2$, $\phi_1$, $\phi_2$, $\lambda_1$ and
+$\lambda_2$, which geometric centre is located at $(r_t, \phi_t, \lambda_t)$, where
+
+\begin{equation}
+    r_t = \frac{r_2 + r_1}{2}, \quad
+    \phi_t = \frac{\phi_2 + \phi_1}{2}, \quad
+    \lambda_t = \frac{\lambda_2 + \lambda_1}{2}.
+\end{equation}
+
+We can define the longitudinal ($L_\lambda$) and latitudinal ($L_\phi$) dimensions of
+the tesseroid as follows:
 
@@
 \textit{Final step}:
-By the end of the algorithm, the running total will be the gravitational effect of the
-tesseroid.
+By the end of the algorithm, a second-order GLQ will be applied on each small tesseroid
+to compute the gravitational field they generate, and the sum of these results will be
+the gravitational effect of the original tesseroid.
 
@@
 Here, we will repeat the numerical experiment using analytical expressions for shells
-with density varying according to exponential and linear functions of $r$.
+with density varying according to certain functions of $r$.
 This numerical experiment will also show that the two dimensional adaptive
-discretization algorithm is as accurate as the original three dimensional version,
-though it reduces the number of tesseroid subdivisions and thus the computation time.
+discretization algorithm can achieve the same accuracy as the original three dimensional
+version, though it reduces the number of tesseroid subdivisions and thus the computation
+time.
 

### 2018-12-11 code/tesseroid_density/tesseroid.py
Remove tesseroid ignore and L_r hard threshold

@@ -154,13 +154,6 @@ def _check_tesseroid(tesseroid, dens):
     # Check if the dimensions given are valid
     assert w <= e and s <= n and top >= bottom, \
         "Invalid tesseroid dimensions {}".format(tesseroid.get_bounds())
-    # Check if the tesseroid has volume > 0
-    if (e - w <= 1e-6) or (n - s <= 1e-6) or (top - bottom <= 1e-3):
-        msg = ("Encountered tesseroid with dimensions smaller than the " +
-               "numerical threshold (1e-6 degrees or 1e-3 m). " +
-               "Ignoring this tesseroid.")
-        warnings.warn(msg, RuntimeWarning)
-        return None
     if dens is not None:
         density = dens
     else:
@@ -217,8 +210,8 @@ def _forward_model(args):
     lon, sinlat, coslat, radius = _convert_coords(lon, lat, height)
     func = getattr(_tesseroid, field)
     warning_msg = (
-        "Stopped dividing a tesseroid because it's dimensions would be " +
-        "below the minimum numerical threshold (1e-6 degrees or 1e-3 m). " +
+        "Stopped dividing a tesseroid because it's horizontal dimensions would be " +
+        "below the minimum numerical threshold (1e-6 degrees). " +
         "Will compute without division. Cannot guarantee the accuracy of " +
         "the solution.")
     for tesseroid in model:
@@ -279,15 +272,6 @@ def _density_based_discretization(bounds, density, delta):
 
         if divider is None:
             subset.append(np.array([w, e, s, n, top, bottom]))
-            continue
-
-        if abs(top - divider) < 1e-3 or abs(divider - bottom) < 1e-3:
-            subset.append(np.array([w, e, s, n, top, bottom]))
-            msg = ("Encountered tesseroid with dimensions smaller " +
-                   "than the numerical threshold (1e-3 m) on " +
-                   "density-based discretization. " +
-                   "Ignoring this tesseroid.")
-            warnings.warn(msg, RuntimeWarning)
             continue
 
         size_ratio = (top - bottom)/tesseroid_size

### 2018-12-11 manuscript/manuscript.tex
Improve density-based discretization algorithm

@@
 
-The numerical integration of an arbitrary continuous density function introduces
-a new type of problem:
-the integration error from using only a few nodes to discretize the density function.
+The numerical integration of an arbitrary continuous density function introduces a new
+type of problem: the integration error from using only a few nodes to account for the
+variation of the density function.
 The three dimensional adaptive discretization may help
 to reduce this kind of error by adding more point masses in the radial direction.
-However, it does not take into account the density function
-and hence it is not well suited to fully perform this task.
-Moreover, because the two dimensional adaptive discretization algorithm does not
-subdivide the tesseroid on the radial dimension, it does not put any effort on lowering
-this new kind of error.
+However, it does not take into account the density function when subdividing the
+tesseroid and hence it is not well suited to fully perform this task.
+The two dimensional adaptive discretization algorithm does not subdivide the tesseroid
+on the radial dimension, so it does not put any effort on lowering this new kind of
+error.
 
@@
 algorithm that takes into account the variations of the density function.
-This density-based discretization happens prior to the adaptive discretization described
-in the previous section.
+This density-based discretization is applied prior to the adaptive discretization
+described in the previous section.
 In short, the algorithm divides the tesseroid along the radial dimension at the
@@
 
-Consider an \emph{original} tesseroid with density given by the function $\rho(r')$.
+Consider an \emph{original} tesseroid with a continuous density given by the function
+$\rho(r')$.
 Before the density-based discretization starts,
@@
 Once all smaller tesseroids satisfy Eq. \ref{eq:delta-density},
-each one is subjected to the adaptive
+each one is subjected to the two dimensional adaptive
 discretization algorithm described earlier to calculate their gravitational effects.
@@
 higher $\text{max}\{ \Delta \rho(r') \}$.
-We also do not divide tesseroids with $L_r < 1$ mm because any integration errors would
-be negligible due to the small mass of the tesseroid.
 
@@
 Apply a second-order GLQ to numerically compute the gravitational fields
-(Eq.\ref{eq:glq-var-dens}) generated by each tesseroid of the set obtained in the
+(Eq.\ref{eq:glq-var-dens}) generated by each tesseroid obtained in the
 previous step.

### 2018-12-11 code/scripts/linear_density.py
Fix wrong constant term in linear density

@@ -64,13 +64,13 @@ D_values = np.arange(0.5, 5.5, 0.5)
 for field in fields:
     for model in models:
         top, bottom = model.bounds[4], model.bounds[5]
-        slope = -(3300 - 2670) / (top - bottom)
-        constant_term = (3300 - 2670) * MEAN_EARTH_RADIUS + 2670
+        slope = (2670 - 3300) / (top - bottom)
+        constant_term = 2670 - slope * (top + MEAN_EARTH_RADIUS)
 
         # Define density function
         def density_linear(height):
             r = height + MEAN_EARTH_RADIUS
-            return slope*r + constant_term
+            return slope * r + constant_term
 
         model.addprop("density", [density_linear for i in range(model.size)])
 

### 2018-12-11 code/scripts/linear_density.py
Add linear density fig with grid spec

@@ -5,6 +5,7 @@ from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
 from fatiando.mesher import TesseroidMesh
 from fatiando import gridder
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec, GridSpecFromSubplotSpec
 # This is our custom tesseroid code
 from tesseroid_density import tesseroid
 
@@ -99,26 +100,63 @@ for field in fields:
 
 # Plot Results
 # ------------
-# Plot one line per thickness using the maximum difference value for each case
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/linear-density-diffs.pdf")
 field_titles = dict(zip(fields, '$V$ $g_z$'.split()))
 grid_titles = {"pole": "Pole",
                "equator": "Equator",
                "global": "Global",
                "260km": "Satellite"}
 colors = dict(zip(thicknesses, plt.cm.viridis(np.linspace(0, 0.9, len(thicknesses)))))
+markers = dict(zip(thicknesses, ["o", "^", "s", "d", "x"]))
 
+# Create outer grid
+fig = plt.figure(figsize=(7, 8))
+outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
+
+# Create grid specs for each grid
+grid_specs = dict(
+                  zip(
+                      grids.keys(),
+                      [GridSpecFromSubplotSpec(ncols=1, nrows=2,
+                                               subplot_spec=outer_grid[j, i],
+                                               hspace=0)
+                       for j in range(2)
+                       for i in range(2)]
+                     )
+                 )
+positions = dict(zip(grids.keys(), [[j, i] for j in range(2) for i in range(2)]))
+
+# Plot for each grid
 for grid_name in grids:
 
-    fig, axes = plt.subplots(nrows=len(fields), ncols=1, sharex=True)
-    fig.set_size_inches((5, 5))
-    fig.subplots_adjust(hspace=0)
+    # Choose gridspec and create axes
+    gs = grid_specs[grid_name]
+    position = positions[grid_name]
+    axes = [plt.subplot(gs[0, 0])]
+    axes.append(plt.subplot(gs[1, 0], sharex=axes[0]))
+    plt.setp(axes[0].get_xticklabels(), visible=False)
+    # Remove xticks if the axes in in the middle of the figure
+    if position[0] == 0:
[…]

### 2018-12-11 code/scripts/linear_density.py
Fix location of legend in linear density

@@ -198,7 +198,7 @@ for grid_name in grids:
 
     # Add legend
     if position == [0, 0]:
-        axes[0].legend(loc=0, prop={"size": 8})
+        axes[0].legend(loc=1, prop={"size": 8})
 
 outer_grid.tight_layout(fig)
 plt.savefig(figure_fname, dpi=300)

### 2018-12-11 manuscript/manuscript.tex
Improve D and delta determination
Also update figures.

@@
 \citet{Uieda2016} compared the numerical integration of homogeneous density tesseroids
-with the analytical solution of a spherical shell \citep{Mikuska2006,Grombein2013} in
+with the analytical solution of a spherical shell \citep{Mikuska2006, Grombein2013} in
 order to obtain default values for the distance-size ratio $D$.
@@
     tesseroid that compose the spherical shell model.
-    For each shape, we define several spherical shell models which the following
+    The outer radius of the shell is equal to the mean Earth radius, while the inner
+    radius is defined based on its thickness.
+    For each shape, we define several spherical shell models with the following
     thicknesses: 100m, 1km, 10km, 100km and 1000km.
-    While the outer radius of every shell is equal to the mean Earth radius
 }
@@
 divisions will ever be performed.
-Therefore, the distance-size ratio $D$ of the adaptive discretization algorithm is the
+Therefore, the two dimensional adaptive discretization algorithm is the
 only mechanism that controls the accuracy of the numerical integration.
@@
 In order to compare the numerical results with the analytical solution we
-must build a spherical shell made of tesseroids.
-The outer radius of the shell is equal to the mean Earth radius and we repeat the
-calculations for shell thicknesses of 1km and 35km.
-We discretize the shell into a single layer of $30^\circ \times 30^\circ$ tesseroids and
-compute their gravitational effects on $10 \times 10$ point grids with different
-resolutions, heights, and locations on the sphere.
-These grids are defined in Table \ref{tab:grids}.
-The density of the shell and tesseroid model varies linearly with $r$
-(Eq. \ref{eq:density-linear}) with angular coefficient
-
-\begin{equation}
-    a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R - R_1},
-\end{equation}
-
-\noindent and linear coefficient
-
-\begin{equation}
-    b = \frac{3300\text{kg/m$^3$} -
-        2670\text{kg/m$^3$}}{R - R_1} R +
-        2670\text{kg/m$^3$},
+must build spherical shell models made of tesseroids.
+We propose the ones detailed on Table~\ref{tab:tess-models}: the outer radius of each
+shell is equal to the mean Earth radius (6378.137 km) while the inner radius is
+determined based on its thickness.
+We cover spherical shells composed by tesseroid of different sizes and shells with
+thicknesses of 100m, 1km, 10km, 100km and 1000km.
+
+The comparison between the numerical results and the analytical solution could be
+naively compared on a single computation point due to the rotation symmetry of the
+spherical shell.
+But the numerical results may adopt different values depending on the relative location
+between the computation point and the tesseroid, thus is not wise to carry out the
+comparison on a single point.
+Instead, we propose four computation grids detailed on Table~\ref{tab:grids}: a Pole and
+an Equator grids, and a Global and Satellite grids.
+The former ones are regional scale grids located on the outer surface of the spherical
+shell, while the latter are global scale grids located at zero height (Global) and at
+the mean orbital height of the gravimetric satellite missions (Satellite).
+We define these grids in the spirit of span a broad scenario of gravitational field
+computations and ensure an acceptable accuracy on each one of them.
+The comparison will be carried out by computing the maximum absolute difference between
+the analytical solution and the numerical result on each point of the computation grid.
+
+We assign a density that varies linearly with $r$ (Eq.~\ref{eq:density-linear}) to each
+shell, which assumes the values of $\rho_\text{in} = 3300$~kg/m$^3$ and
+$\rho_\text{out} = 2670$~kg/m$^3$ on the inner and outer radii, respectively.
+So, the constants $a$ and $b$ from Eq.~\ref{eq:density-linear} must be
+
+\begin{equation}
+    a = \frac{\rho_\text{out} - \rho_\text{in}}{R_\text{out} - R_\text{in}},
+\end{equation}
+
+\begin{equation}
+    b = \rho_\text{out} -
+    \frac{\rho_\text{out} - \rho_\text{in}}{R_\text{out} - R_\text{in}} R_\text{out},
 \end{equation}
@@
 \noindent
-in which $R$ = 6378.137 km is the mean Earth radius and $R_1$ is the inner radius of the
-shell (determined by the thickness).
+where $R_\text{out}$ and $R_text{in}$ are the outer and inner radii of the shell,
+respectively.
+For every shell defined in Table~\ref{tab:tess-models}, the outer radius will be equal
+to the mean Earth radius (6378.137 km).
 
@@
 \centering
-\includegraphics[width=\linewidth]{figures/linear-D.pdf}
+\includegraphics[width=\linewidth]{figures/linear-density-diffs.pdf}
 \caption{
-    Differences between the gravitational fields generated by the tesseroid model
-    and the analytical solution for a thin (left) and a thick (right) spherical shell.
-    Both models have a linear density function of radius (Eq. \ref{eq:density-linear}).
+    Differences between the gravitational fields generated by each tesseroid shell model
+    and the analytical solution.
+    Each model has a linear density function of radius (Eq. \ref{eq:density-linear}).
     The computations were performed on the four combinations described in
     Table \ref{tab:grids}.
+    Each line represents the maximum absolute difference between every shell model with
+    a certain thickness.
     Due to the linearity of the density function, the density-based discretization
@@
 
-We compute the gravitational potential ($V$), the vertical component of the gradient
-($g_z$), and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$,
-$g_{zz}$) for each grid in Table \ref{tab:grids}.
-Other components are equal to zero outside of the shell and are thus omitted from the
-analysis.
-The computations are repeated for values of $D$ ranging from 0.5 to 10 with a step of
+We compute the gravitational potential ($V$) and the vertical component of the gradient
+($g_z$) that each shell model detailed on Table~\ref{tab:tess-models} generates on each
+grid defined in Table~\ref{tab:grids}.
+Other components of the gradient are equal to zero outside of the shell due to the
+rotational symmetry and are thus omitted from the analysis.
+The computations are repeated for values of $D$ ranging from 0.5 to 5 with a step of
 0.5.
-We will then calculate the maximum absolute difference between these results and the
-analytical solutions derived in Appendix \ref{sec:shell}.
 The differences are shown in Fig. \ref{fig:D-linear} as relative to the spherical shell

### 2018-12-12 manuscript/manuscript.tex
Improve D determination text

@@
 We perform comparisons between the analytical solutions for the spherical shell and the
-numerical integration results for linear and exponential density functions.
+numerical integration results for linear, exponential and trigonometric density
+functions.
 From these results, we generalize default values for $D$ and $\delta$ that ensure a
@@
 \caption{
-    Description of the tesseroid models used to build a spherical shell and characterize
-    the accuracy of the numerical integration.
-    The first three columns describe the shapes of the tesseroid models: each one
+    Description of the tesseroid collections used to build a spherical shell and
+    characterize the accuracy of the numerical integration.
+    The first three columns describe the shapes of the tesseroid collections: each one
     specifies the number of tesseroids on the corresponding dimension (radial,
@@
     tesseroid that compose the spherical shell model.
-    The outer radius of the shell is equal to the mean Earth radius, while the inner
-    radius is defined based on its thickness.
-    For each shape, we define several spherical shell models with the following
-    thicknesses: 100m, 1km, 10km, 100km and 1000km.
+    For each collection, we define four spherical shell models with an outer radius
+    equal to the mean Earth radius (6378.137 km) and the inner radius determined by the
+    following thicknesses: 100m, 1km, 10km, 100km and 1000km.
 }
@@
 must build spherical shell models made of tesseroids.
-We propose the ones detailed on Table~\ref{tab:tess-models}: the outer radius of each
-shell is equal to the mean Earth radius (6378.137 km) while the inner radius is
-determined based on its thickness.
-We cover spherical shells composed by tesseroid of different sizes and shells with
-thicknesses of 100m, 1km, 10km, 100km and 1000km.
+These shell models are determined by their inner and outer radii and by the collection
+of tesseroids used to build the shell.
+For this analysis we have chosen to set the outer radii of every shell model equal to
+the mean Earth radius (6378.137 km), while the inner radius is determined by the desired
+thickness of the shell.
+We choose  different values of shell thickness (100m, 1km, 10km, 100km and 1000km) in
+order to account for different computation scenarios.
+Furthermore, we propose different tesseroids collections to build the shell model by
+regularly splitting it along the longitudinal and latitudinal directions.
+Each collection, described in Table~\ref{tab:tess-models}, is composed by a different
+number of tesseroids, i.e. with tesseroids of different sizes.
+Finally, for each collection we build a shell model with the thicknesses
+previously described, giving a total of 20 different shell models.
 

### 2018-12-12 code/scripts/linear_density.py
Improve linear density figure

@@ -175,7 +175,7 @@ for grid_name in grids:
                     label=label)
 
         # Add threshold line
-        ax.plot([0.5, 5], [1e-1, 1e-1], '--', color='k', linewidth=0.5)
+        ax.plot([0, 6], [1e-1, 1e-1], '--', color='k', linewidth=0.5)
 
         # Add field annotation on each axe
         ax.text(0.5, 0.87, field_title, fontsize=11,
@@ -187,14 +187,13 @@ for grid_name in grids:
                       'boxstyle': 'circle, pad=0.4'},
                 transform=ax.transAxes)
 
-
-
         # Configure axes
         ax.set_yscale('log')
         ax.set_yticks(ax.get_yticks()[2:-2])
         ax.set_ylabel('Difference (%)')
         ax.grid(True, linewidth=0.5, color='#aeaeae')
         ax.set_xticks(np.arange(0, 6, 1))
+        ax.set_xlim(0.3, 5.2)
 
     # Add legend
     if position == [0, 0]:

### 2018-12-12 manuscript/manuscript.tex
Improve D determination text

@@
 }
-\label{tab:tess-models}
+\label{tab:tess-collections}
 \begin{tabular}{cccccc}
@@
 regularly splitting it along the longitudinal and latitudinal directions.
-Each collection, described in Table~\ref{tab:tess-models}, is composed by a different
-number of tesseroids, i.e. with tesseroids of different sizes.
+Each collection, described in Table~\ref{tab:tess-collections}, is composed by
+a different number of tesseroids, i.e. with tesseroids of different sizes.
 Finally, for each collection we build a shell model with the thicknesses
@@
 We assign a density that varies linearly with $r$ (Eq.~\ref{eq:density-linear}) to each
-shell, which assumes the values of $\rho_\text{in} = 3300$~kg/m$^3$ and
+shell model, which assumes the values of $\rho_\text{in} = 3300$~kg/m$^3$ and
 $\rho_\text{out} = 2670$~kg/m$^3$ on the inner and outer radii, respectively.
@@
 \noindent
-where $R_\text{out}$ and $R_text{in}$ are the outer and inner radii of the shell,
+where $R_\text{out}$ and $R_\text{in}$ are the outer and inner radii of the shell,
 respectively.
-For every shell defined in Table~\ref{tab:tess-models}, the outer radius will be equal
-to the mean Earth radius (6378.137 km).
+Because the outer radius of every spherical shell model will be equal to the mean Earth
+radius, $R_\text{out} = 6378.137$km.
 
@@
 We compute the gravitational potential ($V$) and the vertical component of the gradient
-($g_z$) that each shell model detailed on Table~\ref{tab:tess-models} generates on each
-grid defined in Table~\ref{tab:grids}.
+($g_z$) that each shell model generates on each grid defined in Table~\ref{tab:grids}.
 Other components of the gradient are equal to zero outside of the shell due to the
@@
 0.5.
-The differences are shown in Fig. \ref{fig:D-linear} as relative to the spherical shell
-values.
-We omit the differences for $g_{xx}$ and $g_{yy}$ because they are equivalent to the
-ones obtained for $g_{zz}$.
+Then we calculate the absolute difference between these results and the analytical
+solution, taking only the maximum absolute difference on each computation grid.
+This differences are shown in Fig.~\ref{fig:D-linear}.
+In order to simplify the visualization of the results, only the maximum
+differences corresponding to tesseroids collections with equal thickness are shown
+in the Figure, obtaining one curve for every shell thickness.
 Finally, we set the optimal value of $D$ as the minimum value at which the corresponding
@@
 potential and the $g_z$ component fall below the 0.1\% threshold at
-$D=1$ and $D=2$, respectively, for both shell thicknesses.
-On the other hand, $g_{zz}$ displays a noticeable
-difference between the thin and the thick shell: the first one needs a
-value of $D$ equal to 8 while the later only a $D$ of 3.
+$D=1$ and $D=2.5$, respectively, for every shell model and computation grid.
+These results also show that there's no meaningful dependency between the accuracy of
+the method and the thickness of the shell, the collection of tesseroids used to build
+it or the height of the computation grid (the results for the Satellite grid and the
+Global grid are very similar).
 

### 2018-12-12 code/scripts/linear_density.py
Improve linear density figure

@@ -107,6 +107,7 @@ grid_titles = {"pole": "Pole",
                "equator": "Equator",
                "global": "Global",
                "260km": "Satellite"}
+grid_names = ["pole", "equator", "global", "260km"]
 colors = dict(zip(thicknesses, plt.cm.viridis(np.linspace(0, 0.9, len(thicknesses)))))
 markers = dict(zip(thicknesses, ["o", "^", "s", "d", "x"]))
 
@@ -117,7 +118,7 @@ outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 # Create grid specs for each grid
 grid_specs = dict(
                   zip(
-                      grids.keys(),
+                      grid_names,
                       [GridSpecFromSubplotSpec(ncols=1, nrows=2,
                                                subplot_spec=outer_grid[j, i],
                                                hspace=0)
@@ -125,10 +126,10 @@ grid_specs = dict(
                        for i in range(2)]
                      )
                  )
-positions = dict(zip(grids.keys(), [[j, i] for j in range(2) for i in range(2)]))
+positions = dict(zip(grid_names, [[j, i] for j in range(2) for i in range(2)]))
 
 # Plot for each grid
-for grid_name in grids:
+for grid_name in grid_names:
 
     # Choose gridspec and create axes
     gs = grid_specs[grid_name]
@@ -196,7 +197,7 @@ for grid_name in grids:
         ax.set_xlim(0.3, 5.2)
 
     # Add legend
-    if position == [0, 0]:
+    if position == [0, 1]:
         axes[0].legend(loc=1, prop={"size": 8})
 
 outer_grid.tight_layout(fig)

### 2018-12-12 manuscript/manuscript.tex
Improve exponential density text

@@
 We perform an error analysis similar to what was done for the linear density case.
-However, we only consider the ``Big Grid'' from Table \ref{tab:grids} for the
-sake of brevity.
-
-The spherical shell and tesseroid model will have an exponential density function given
+But, now the spherical shell will have an exponential density function that assumes the
+values of $\rho_\text{out} = 2670$~kg/m$^3$ and $\rho_\text{in} = 3300$~kg/m$^3$ on the
+outer and inner surfaces, respectively, defined as follows:
 by
@@
 \begin{equation}
-    \rho(r') = A e^{-(r' - R)/b} + C,
-\label{eq:density-exp}
+    \rho(r') = A e^{- \frac{b}{T} (r' - R)} + C,
 \end{equation}
@@
 \begin{equation}
-    A =
-    (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
-    \left( e^{( R - R_1 )/b} - 1 \right)^{-1},
-\end{equation}
-
-\begin{equation}
-    C =
-    2670 \text{kg/m}^3 - A,
-\end{equation}
-
-\noindent $R$ is the mean Earth radius, $R_1$ is the inner radius of the
-shell (determined by the thickness), and $b$ is a constant that determines the
-variability of the function: a low value of $b$ increases the maximum slope of the
-density function.
+    A = \frac{\rho_\text{in} - \rho_\text{out}}{D},
+\end{equation}
+
+\begin{equation}
+    C = \frac{1}{D}
+    \left[\rho_\text{out} \, e^{- \frac{b}{T} (R_1 - R)} -
+    \rho_\text{in} \, e^{- \frac{b}{T} (R_2 - R)} \right],
+\end{equation}
+
+\begin{equation}
+    D = e^{- \frac{b}{T} (R_1 - R)} - e^{- \frac{b}{T} (R_2 - R)},
+\end{equation}
+
+\noindent $R$ is the mean Earth radius, $T$ is the thickness of the shell ($T = R_2
+- R_1$) and $b$ is a dimensionless constant that determines the variability of the
+function: a high value of $b$ increases the maximum slope of the density function.
 
@@
 
-Moreover, an exponential density function that assumes the values of $\rho_\text{out}$
-and $\rho_\text{in}$ on the shell's outer and inner surfaces, respectively, can be
-defined as follows:
-
-\begin{equation}
-    \rho(r') = A e^{- \frac{b}{T} (r' - R)} + C,
-\end{equation}
-
-\noindent where
-
-\begin{equation}
-    A = \frac{\rho_\text{in} - \rho_\text{out}}{D},
-\end{equation}
-
-\begin{equation}
-    C = \frac{1}{D}
-    \left[\rho_\text{out} \, e^{- \frac{b}{T} (R_1 - R)} -
-    \rho_\text{in} \, e^{- \frac{b}{T} (R_2 - R)} \right],
-\end{equation}
-
-\begin{equation}
-    D = e^{- \frac{b}{T} (R_1 - R)} - e^{- \frac{b}{T} (R_2 - R)},
-\end{equation}
-
-\noindent $R$ is the mean Earth radius, $T$ is the thickness of the shell ($T = R_2
-- R_1$) and $b$ is a dimensionless constant that determines the variability of the
-function: a high value of $b$ increases the maximum slope of the density function.
-
-The analytical solution of the gravitational potential generated by a
-spherical shell with such density function is
-
-\begin{equation}
-    \begin{split}
-        V_\text{exp}(r) = & \frac{4\pi G}{r} \frac{A T^3}{b^3} \\ \Big[
-        & \left( \left( b R_1/T \right)^2 + 2 b R_1/T + 2 \right)
-        e^{- \frac{b}{T} (R_1 - R)} - \\
-        & \left( \left( b R_2/T \right)^2 + 2 b R_2/T + 2 \right)
-        e^{- \frac{b}{T} (R_2 - R)}
-        \Big] + \\
-        & \frac{4\pi G}{3 r} C \left( R_2^3 - R_1^3 \right).
-    \end{split}
-\label{eq:shell-pot-exp}
-\end{equation}
-
 \end{document}

### 2018-12-12 manuscript/manuscript.tex
Fix typo

@@
 outer and inner surfaces, respectively, defined as follows:
-by
 

### 2018-12-12 manuscript/manuscript.tex
Add values of b in exponential density

@@
     \rho(r') = A e^{- \frac{b}{T} (r' - R)} + C,
+    \label{eq:density-exp}
 \end{equation}
@@
 - R_1$) and $b$ is a dimensionless constant that determines the variability of the
-function: a high value of $b$ increases the maximum slope of the density function.
+function: a higher value of $b$ increases the maximum slope of the density function.
+On exponential density analysis we will use values of $b$ equal to 1, 2, 5, 10, 30
+and 100, in order to cover different kind of exponential density variations, from smooth
+to sharp.
 

### 2018-12-12 code/scripts/exponential_density.py
Fix missing word in exp densities xticklabel

@@ -158,7 +158,7 @@ for b_factor in b_factors:
             label="b={}".format(b_factor))
 ax.set_ylabel(r"Density [kg/m$^3$]")
 ax.set_xticks([bottom, top])
-ax.set_xticklabels(["Radius", "Outer Radius"])
+ax.set_xticklabels(["Inner Radius", "Outer Radius"])
 ax.legend()
 plt.tight_layout()
 plt.savefig(figure_fname, dpi=300)

### 2018-12-12 manuscript/manuscript.tex
Update grid search text

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{figures/exponential-densities.pdf}
+\caption{
+    Exponential density functions assigned to each spherical shell in the $D$ and
+    $\delta$ ratios determination.
+    Each density function corresponds to a different value of $b$ on
+    Eq.~\ref{eq:density-exp}.
+}
+\label{fig:exp-densities}
+\end{figure}
+
 \subsubsection{$D$-$\delta$ space exploration}
@@
 We use a grid search method and compute the numerical error for every ($D$, $\delta$)
-pair belonging to a grid on the $D$-$\delta$ space (Fig. \ref{fig:grid-search}).
-Because this is a time consuming computation, we limit the analysis to a shell of 35 km
-thickness and $b = 1\ \text{km}$ (Eq. \ref{eq:density-exp}), which yields a
-sharp density variation within the shell.
-For optimum algorithm performance, we search for the smallest possible value of $D$ and
-the highest possible value of $\delta$.
+pair belonging to a grid on the $D$-$\delta$ space (Fig.~\ref{fig:grid-search}).
+For optimum algorithm performance, we search for the ($D$, $\delta$) pair that minimize
+the number of splittings while keeping a numerical error under the 0.1\% threshold,
+i.e.~the smallest possible value of $D$ and the highest possible value of $\delta$.
 
 We compute the relative difference between the numerical and analytical results for the
-gravitational potential ($V$), the vertical component of its gradient ($g_z$), and the
-$g_{zz}$ component of the Marussi tensor.
-The results are shown in Fig. \ref{fig:grid-search}, in which points inside the dotted
+gravitational potential ($V$) and the vertical component of its gradient ($g_z$) on the
+computation grids detailed in Table~\ref{tab:grids} and using shell models with the same
+thicknesses as the linear density case.
+Because this is a time consuming computation, we limit the analysis only to the second
+collection of tesseroids in Table~\ref{tab:tess-collections}.
+For the sake of brevity, we compute the maximum differences between the results
+involving each computation grid, shell thickness and value of the $b$ factor.
+The results are shown in Fig.~\ref{fig:grid-search}, in which points inside the dotted
 lines are the ones that present a numerical error lower than the 0.1\% threshold.
@@
 values of $D$ for $V$ and $g_z$.
-The results for $g_{zz}$ are in agreement with those obtained for the linear density
-function in the case of a thick shell (Fig. \ref{fig:D-linear}).
-Thus, we also choose the conservative value of $D=8$ as the optimum value for $g_{zz}$.
 These results indicate that the values of $D_\text{linear}$ can be safely extrapolated
 to the exponential case.
+
+The two dimensional adaptive discretization and the density-based algorithms are
+independent of each other: the first one splits the tesseroid only in the latitudinal
+and longitudinal directions, while the latter only does it in the radial direction.
+This leads to the fact that the value of $\delta$ needed to limit the numerical error
+due to the density variation is independent of the value of $D$.
 

### 2018-12-12 code/scripts/linear_density.py
Add latex style in linear density and fix figsize

@@ -1,5 +1,6 @@
 from __future__ import division, print_function
 import os
+import warnings
 import numpy as np
 from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
 from fatiando.mesher import TesseroidMesh
@@ -98,8 +99,21 @@ for field in fields:
                      D_values=D_values, differences=differences)
 
 
+# Configure LaTeX style for plots
+# -------------------------------
+try:
+    plt.rcParams['text.usetex'] = True
+    plt.rcParams['font.family'] = 'serif'
+    plt.rcParams['font.serif'] = 'Computer Modern Roman'
+    plt.rcParams['xtick.major.size'] = 2
+    plt.rcParams['ytick.major.size'] = 2
+except Exception as e:
+    warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
+
+
 # Plot Results
 # ------------
+# Configure plot
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/linear-density-diffs.pdf")
 field_titles = dict(zip(fields, '$V$ $g_z$'.split()))
@@ -112,7 +126,7 @@ colors = dict(zip(thicknesses, plt.cm.viridis(np.linspace(0, 0.9, len(thicknesse
 markers = dict(zip(thicknesses, ["o", "^", "s", "d", "x"]))
 
 # Create outer grid
-fig = plt.figure(figsize=(7, 8))
+fig = plt.figure(figsize=(6.66, 8))
 outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 
 # Create grid specs for each grid
@@ -191,7 +205,7 @@ for grid_name in grid_names:
         # Configure axes
         ax.set_yscale('log')
         ax.set_yticks(ax.get_yticks()[2:-2])
-        ax.set_ylabel('Difference (%)')
+        ax.set_ylabel(r'Difference (\%)')
         ax.grid(True, linewidth=0.5, color='#aeaeae')
         ax.set_xticks(np.arange(0, 6, 1))
         ax.set_xlim(0.3, 5.2)

### 2018-12-12 code/scripts/exponential_density.py
Add latex style in exp density and fix figsize

@@ -1,5 +1,6 @@
 from __future__ import division, print_function
 import os
+import warnings
 import numpy as np
 from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
 from fatiando.mesher import TesseroidMesh
@@ -131,13 +132,25 @@ for field in fields:
                          delta_values=delta_values, differences=differences)
 
 
+# Configure LaTeX style for plots
+# -------------------------------
+try:
+    plt.rcParams['text.usetex'] = True
+    plt.rcParams['font.family'] = 'serif'
+    plt.rcParams['font.serif'] = 'Computer Modern Roman'
+    plt.rcParams['xtick.major.size'] = 2
+    plt.rcParams['ytick.major.size'] = 2
+except Exception as e:
+    warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
+
+
 # Plot Densities
 # --------------
 bottom, top = 0, 1
 thickness = top - bottom
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
-fig, ax = plt.subplots(figsize=(6, 4))
+fig, ax = plt.subplots(figsize=(3.33, 2.5))
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/exponential-densities.pdf")
 for b_factor in b_factors:
@@ -177,7 +190,7 @@ grid_titles = {"pole": "Pole",
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
 # Create outer grid
-fig = plt.figure(figsize=(7, 8))
+fig = plt.figure(figsize=(6.66, 8))
 outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 
 # Create grid specs for each grid
@@ -252,7 +265,7 @@ for grid_name in grids:
         ax.set_xscale('log')
         ax.set_yscale('log')
         ax.set_yticks(ax.get_yticks()[2:-2])
-        ax.set_ylabel('Difference (%)')
+        ax.set_ylabel(r'Difference (\%)')
         ax.grid(True, linewidth=0.5, color='#aeaeae')
 
     # Add legend

### 2018-12-12 manuscript/manuscript.tex
Fix exponential densities fig in manuscript

@@
 and 100, in order to cover different kind of exponential density variations, from smooth
-to sharp.
+to sharp (Fig.~\ref{fig:exp-densities}).
 
@@
 \centering
+\iftwocol{
 \includegraphics[width=\linewidth]{figures/exponential-densities.pdf}
+}{
+\includegraphics[width=0.5\linewidth]{figures/exponential-densities.pdf}
+}
 \caption{

### 2018-12-12 code/scripts/sine_density.py
Add latex style in sine density and fix figsize

@@ -1,5 +1,6 @@
 from __future__ import division, print_function
 import os
+import warnings
 import numpy as np
 from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
 from fatiando.mesher import TesseroidMesh
@@ -121,6 +122,18 @@ for field in fields:
                          delta_values=delta_values, differences=differences)
 
 
+# Configure LaTeX style for plots
+# -------------------------------
+try:
+    plt.rcParams['text.usetex'] = True
+    plt.rcParams['font.family'] = 'serif'
+    plt.rcParams['font.serif'] = 'Computer Modern Roman'
+    plt.rcParams['xtick.major.size'] = 2
+    plt.rcParams['ytick.major.size'] = 2
+except Exception as e:
+    warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
+
+
 # Plot Densities
 # --------------
 bottom, top = 0, 1
@@ -129,7 +142,7 @@ colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))
 
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/sine-densities.pdf")
-fig, axes = plt.subplots(len(b_factors), 1, figsize=(6, 6), sharex=True)
+fig, axes = plt.subplots(len(b_factors), 1, figsize=(3.33, 4), sharex=True)
 for b_factor, ax in zip(b_factors, axes):
 
     # Compute the k factor, i.e. the frequency for the sine function
@@ -164,7 +177,7 @@ grid_titles = {"pole": "Pole",
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
 # Create outer grid
-fig = plt.figure(figsize=(7, 8))
+fig = plt.figure(figsize=(6.66, 8))
 outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 
 # Create grid specs for each grid
@@ -239,7 +252,7 @@ for grid_name in grids:
         ax.set_xscale('log')
         ax.set_yscale('log')
         ax.set_yticks(ax.get_yticks()[2:-2])
-        ax.set_ylabel('Difference (%)')
+        ax.set_ylabel(r'Difference (\%)')
         ax.grid(True, linewidth=0.5, color='#aeaeae')
 
     # Add legend

### 2018-12-13 manuscript/manuscript.tex
Update linear density figure caption

@@
     and the analytical solution.
-    Each model has a linear density function of radius (Eq. \ref{eq:density-linear}).
-    The computations were performed on the four combinations described in
-    Table \ref{tab:grids}.
+    Each model has a linear density function (Eq. \ref{eq:density-linear}).
+    The computations were performed on the four grids described in
+    Table~\ref{tab:grids} and using 20 shell models built with the tesseroids
+    collections detailed in Table~\ref{tab:tess-collections} and five different
+    thicknesses.
     Each line represents the maximum absolute difference between every shell model with

### 2018-12-13 manuscript/manuscript.tex
Fix some LaTeX styling

@@
     Mario E. Gimenez$^{1,2}$ \\
-    $^1$CONICET, Argentina. e-mail: santiago.r.soler@gmail.com\\
+    $^1$CONICET, Argentina.~e-mail: santiago.r.soler@gmail.com\\
     $^2$Instituto Geofísico Sismológico Volponi, Universidad Nacional de
@@
 several density-depth relations have been proposed for different rock types
-\citep[e.g.,][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
+\citep[e.g.][]{Maxant1980, Rao1986, Rao1993, Rao1994}.
 Furthermore, depth-variable densities have been used in forward and
@@
 for computation points near the tesseroid they apply a GLQ numerical integration with
-a two dimensional adaptive discretization algorithm, i.e. only in the horizontal
+a two dimensional adaptive discretization algorithm, i.e.~only in the horizontal
 dimension. But, if the computation point is farther than a certain truncation distance,
@@
 
-Applying a $N$th order GLQ, we can approximate each integral in equations
-\ref{eq:tesseroid-pot} and \ref{eq:tesseroid-grav} by a weighted sum of the integration
-kernel evaluated on the roots of an $N$th order Legendre polynomial
+Applying a $N$th order GLQ, we can approximate each integral in
+equations~\ref{eq:tesseroid-pot} and~\ref{eq:tesseroid-grav} by a weighted sum of the
+integration kernel evaluated on the roots of an $N$th order Legendre polynomial
 \citep[p.~390]{Hildebrand1987}.
 Unlike the homogeneous density case, the radial density function $\rho(r)$ must also be
-included in the integration and evaluated on the Legendre polynomial roots (i.e.,
-quadrature nodes).
+included in the integration and evaluated on the Legendre polynomial roots
+(i.e.~quadrature nodes).
 
@@
     \cos\psi_t =
-        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t) .
+        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t).
 \end{equation}
@@
 \textit{Step 2}:
-If none of the dimensions of the tesseroid fail inequality
-\ref{eq:condition}, then compute the gravitational effect of the tesseroid using a
-second-order GLQ (Eq. \ref{eq:glq-var-dens}).
+If none of the dimensions of the tesseroid fail inequality~\ref{eq:condition}, then
+compute the gravitational effect of the tesseroid using a second-order GLQ
+(Eq.~\ref{eq:glq-var-dens}).
 Add the computed effect to a running total.
@@
 \textit{Step 3}:
-If any dimension fails inequality \ref{eq:condition}, split the tesseroid in half along
+If any dimension fails inequality~\ref{eq:condition}, split the tesseroid in half along
 the offending dimensions.
@@
     The dashed red line represents the maximum density difference $\Delta \rho (r')$ at
-    which the tesseroid would be divided (assuming that the inequality
-    \ref{eq:delta-density} is not satisfied).
+    which the tesseroid would be divided (assuming that the
+    inequality~\ref{eq:delta-density} is not satisfied).
     (b) second iteration of the algorithm with a new linear density function and maximum
@@
     (d) final output of the density-based discretization, assuming that all four new
-    tesseroids satisfy inequality \ref{eq:delta-density}.
+    tesseroids satisfy inequality~\ref{eq:delta-density}.
 }
@@
 The algorithm is comprised of the following steps
-(Fig. \ref{fig:density-discretization-algorithm}):
+(Fig.~\ref{fig:density-discretization-algorithm}):
 
@@
 \textit{Step 5}:
-If inequality \ref{eq:delta-density} is not satisfied, then the tesseroid is split in
-two parts at the radius $r_\text{max}$ at which the maximum absolute difference (Eq.
-\ref{eq:density-abs-diff}) takes place.
+If inequality~\ref{eq:delta-density} is not satisfied, then the tesseroid is split in
+two parts at the radius $r_\text{max}$ at which the maximum absolute difference
+(Eq.~\ref{eq:density-abs-diff}) takes place.
 
@@
 
-Once all smaller tesseroids satisfy Eq. \ref{eq:delta-density},
+Once all smaller tesseroids satisfy Eq.~\ref{eq:delta-density},
 each one is subjected to the two dimensional adaptive
@@
 We derive an expression for such gravitational potential in case the spherical shell has
-an arbitrary density function in depth (Eq. \ref{eq:shell-pot}), and obtained the
-expressions in case of linear and exponential density (see Appendix \ref{sec:shell}).
+an arbitrary density function in depth (Eq.~\ref{eq:shell-pot}), and obtained the
+expressions in case of linear and exponential density (see Appendix~\ref{sec:shell}).
 
@@
 \noindent
-has an analytical solution given by Eq. \ref{eq:shell-pot-linear}.
-
-The absolute density difference defined on equation
-\ref{eq:density-abs-diff} will always be zero for the linear density case.
-As a result, the inequality \ref{eq:delta-density} will always be satisfied and no
+has an analytical solution given by Eq.~\ref{eq:shell-pot-linear}.
+
+The absolute density difference defined on equation~\ref{eq:density-abs-diff} will
+always be zero for the linear density case.
+As a result, the inequality~\ref{eq:delta-density} will always be satisfied and no
 divisions will ever be performed.
@@
 Each collection, described in Table~\ref{tab:tess-collections}, is composed by
-a different number of tesseroids, i.e. with tesseroids of different sizes.
+a different number of tesseroids, i.e.~with tesseroids of different sizes.
 Finally, for each collection we build a shell model with the thicknesses
@@
     and the analytical solution.
-    Each model has a linear density function (Eq. \ref{eq:density-linear}).
+    Each model has a linear density function (Eq.~\ref{eq:density-linear}).
     The computations were performed on the four grids described in
@@
 
-We observe from Fig. \ref{fig:D-linear} that the relative errors for the
+We observe from Fig.~\ref{fig:D-linear} that the relative errors for the
 potential and the $g_z$ component fall below the 0.1\% threshold at
@@
 lines are the ones that present a numerical error lower than the 0.1\% threshold.
-Also shown in Fig. \ref{fig:grid-search} are the $D$ values determined for the linear
+Also shown in Fig.~\ref{fig:grid-search} are the $D$ values determined for the linear
 density function in the previous section ($D_\text{linear}$).
@@
     the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a
-    spherical shell with an exponential density function (Eq. \ref{eq:density-exp}).
-    These comparisons were carried out on the ``Big Grid'' (Table
-    \ref{tab:grids}), with a spherical shell of
-    35km of thickness, and density function with $b$ of 1km.
+    spherical shell with an exponential density function (Eq.~\ref{eq:density-exp}).
+    These comparisons were carried out on the ``Big Grid'' (Table~\ref{tab:grids}), with
+    a spherical shell of 35km of thickness, and density function with $b$ of 1km.
     The points inside the dashed line are the ones that present an
@@
     The top panels show the exponential density functions for different values of $b$
-    (Eq. \ref{eq:density-exp}).
+    (Eq.~\ref{eq:density-exp}).
     The error computations were performed on the ``Big Grid'' configuration described in
-    Table \ref{tab:grids}, with fixed values of the distance-size ratio
+    Table~\ref{tab:grids}, with fixed values of the distance-size ratio
     $D$ obtained for the linear density case ($D=1, 2, 8$ for the
@@
 integration error as a function of $\delta$ in more detail and how it varies for
-different values of $b$ (Eq. \ref{eq:density-exp}).
-We perform the error calculations for the ``Big Grid'' configuration (Table
-\ref{tab:grids}) for two spherical shell models with 1km and 35km thickness.
+different values of $b$ (Eq.~\ref{eq:density-exp}).
+We perform the error calculations for the ``Big Grid'' configuration
+(Table~\ref{tab:grids}) for two spherical shell models with 1km and 35km thickness.
 The calculations are repeated for different values of $b$ to examine the error variation
@@
 
-Fig. \ref{fig:delta-exponential} shows the density functions and the resulting relative
+Fig.~\ref{fig:delta-exponential} shows the density functions and the resulting relative
 differences ($V$, $g_z$, and $g_{zz}$) for different values of $b$ and thickness of the
@@
 located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude
-(Fig. \ref{fig:neuquen-basin}a).
+(Fig.~\ref{fig:neuquen-basin}a).
 The basin includes continental and marine siliciclastics, carbonates, and evaporites
@@
 grid with a resolution of 0.05$^\circ$ on both longitude and latitude directions
-(Fig. \ref{fig:neuquen-basin}b).
+(Fig.~\ref{fig:neuquen-basin}b).
 We created a tesseroid model of the sediment pack by placing a
@@
 the Neuqu\'en basin of -412kg/m$^3$ and -275kg/m$^3$, respectively.
-We have chosen an exponential density variation (Eq. \ref{eq:density-exp}) that assumes
+We have chosen an exponential density variation (Eq.~\ref{eq:density-exp}) that assumes
 the minimum value on the top surface and the maximum at 5858m depth (the thickest part
 of the basin), with a value of $b$ equal to 2km.
-This density function can be seen on Fig. \ref{fig:neuquen-basin}c.
+This density function can be seen on Fig.~\ref{fig:neuquen-basin}c.
 
@@
 reference ellipsoid.
-The resulting fields can be seen in Fig.
-\ref{fig:neuquen-basin}d-l.
-We have excluded the other components of the Marussi tensor from the Fig.
-\ref{fig:neuquen-basin},
-although they can be found in the online repository.
+The resulting fields can be seen in Fig.~\ref{fig:neuquen-basin}d-l.
+We have excluded the other components of the Marussi tensor from the
+Fig.~\ref{fig:neuquen-basin}, although they can be found in the online repository.
 
@@
     For our purposes we will assume that $Q$ is outside of the shell,
-    i.e. $r > R_2$.
+    i.e.~$r > R_2$.
 }
@@
 whose density is a function $\rho(r')$ of the radial coordinate
-(Fig. \ref{fig:spherical-shell}).
+(Fig.~\ref{fig:spherical-shell}).
 The gravitational potential generated by the shell on an arbitrary external
@@
 
-\noindent where $\ell$ is defined in equation \ref{eq:ell}.
+\noindent where $\ell$ is defined in equation~\ref{eq:ell}.
 
 For $Q$ located along the $z$ axis (i.e., $\phi=90^\circ$) at a distance $r$ from the
-origin, equation \ref{eq:ell} simplifies to:
+origin, equation~\ref{eq:ell} simplifies to:
 
@@
 Because the computation point $Q$ is outside of the shell, $r > r'$ and the square roots
-in equation \ref{eq:shell-pot-sqrts} simplify to
+in equation~\ref{eq:shell-pot-sqrts} simplify to
 
@@
 
-From Eq. \ref{eq:shell-pot} we can obtain expressions of the gravitational potential for
+From Eq.~\ref{eq:shell-pot} we can obtain expressions of the gravitational potential for
 different density functions.

### 2018-12-13 manuscript/manuscript.tex
Minor changes to linear density text

@@
     tesseroid that compose the spherical shell model.
-    For each collection, we define four spherical shell models with an outer radius
+    For each collection, we define five spherical shell models with an outer radius
     equal to the mean Earth radius (6378.137 km) and the inner radius determined by the
     following thicknesses: 100m, 1km, 10km, 100km and 1000km.
+    This makes a total of 20 shell models combining tesseroids collections and
+    thicknesses.
 }
@@
 a different number of tesseroids, i.e.~with tesseroids of different sizes.
-Finally, for each collection we build a shell model with the thicknesses
-previously described, giving a total of 20 different shell models.
+Finally, combining these collections and the thicknesses we obtain 20 different shell
+models.
 

### 2018-12-13 manuscript/manuscript.tex
Changes in grid search text

@@
 density function in the previous section ($D_\text{linear}$).
-The values of $D_\text{linear}$ are within the 0.1\% threshold and are the optimum
-values of $D$ for $V$ and $g_z$.
+
+The minimum values of $D$ that are within the 0.1\% threshold are the $D_\text{linear}$
+ones, i.e.~they are the optimal values of the distance-size ratios for $V$ and $g_z$.
 These results indicate that the values of $D_\text{linear}$ can be safely extrapolated
 to the exponential case.
-
-The two dimensional adaptive discretization and the density-based algorithms are
-independent of each other: the first one splits the tesseroid only in the latitudinal
-and longitudinal directions, while the latter only does it in the radial direction.
+This conclusion can be understood considering that the two dimensional adaptive
+discretization and the density-based algorithms are independent of each other: the first
+one splits the tesseroid only in the latitudinal and longitudinal directions, while the
+latter only does it in the radial direction.
 This leads to the fact that the value of $\delta$ needed to limit the numerical error
-due to the density variation is independent of the value of $D$.
+due to the density variation is independent of the value of $D$ and vice versa.
+That's why the optimal distance-size ratios for the exponential density case is the same
+as the ones for the linear density case.
+Because the grid search was limited to specific values of $b$ and used only one grid, we
+will determine the optimal value of $\delta$ for exponential density on the next
+subsection.
 
@@
     Numerical error exploration in the $D$-$\delta$ space.
-    The percentage difference values were obtained from the comparison
-    between the analytical solution and the numerical approximation of
-    the gravitational fields ($V$, $g_z$ and $g_{zz}$) generated by a
-    spherical shell with an exponential density function (Eq.~\ref{eq:density-exp}).
-    These comparisons were carried out on the ``Big Grid'' (Table~\ref{tab:grids}), with
-    a spherical shell of 35km of thickness, and density function with $b$ of 1km.
-    The points inside the dashed line are the ones that present an
-    error lower than 0.1\%.
+    The percentage difference values were obtained from the comparison between the
+    analytical solution and the numerical approximation of the gravitational fields ($V$
+    and $g_z$) generated by a spherical shell with an exponential density function
+    (Eq.~\ref{eq:density-exp}).
+    These comparisons were carried out on the ``Global'' grid (Table~\ref{tab:grids}),
+    with only five spherical shell models made out of the second tesseroid collection
+    from Table~\ref{tab:tess-colections} with the different thicknesses defined in the
+    same Table, and density function with $b = 30$.
+    The percentage difference values are obtained as the maximum difference
+    corresponding to each spherical shell model.
+    The points inside the dashed line are the ones that present an error lower than
+    0.1\%.
     }

### 2018-12-13 code/scripts/grid_search.py
Update grid search
Creates LaTeX style figure.
Only global grid and b=30 are taken into the computation.

@@ -1,5 +1,6 @@
 from __future__ import division, print_function
 import os
+import warnings
 import numpy as np
 from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
 from fatiando.mesher import TesseroidMesh
@@ -67,47 +68,17 @@ models = [
 
 # Define computation grids
 # ------------------------
-grids = {"pole": gridder.regular((89, 90, 0, 1), (11, 11), z=0),
-         "equator": gridder.regular((0, 1, 0, 1), (11, 11), z=0),
-         "global": gridder.regular((-90, 90, 0, 360), (19, 13), z=0),
-         "260km": gridder.regular((-90, 90, 0, 360), (19, 13), z=260e3),
-         }
+grid_name = "global"
+grid = gridder.regular((-90, 90, 0, 360), (19, 13), z=0)
 
 
 # Configure comparisons
 # ---------------------
 fields = 'potential gz'.split()
 density_bottom, density_top = 3300, 2670
-b_factors = [1, 2, 5, 10, 30, 100]
+b_factor = 30
 delta_values = np.logspace(-3, 1, 5)
 D_values = np.arange(0.5, 4.5, 0.5)
-
-
-# Plot Densities
-# --------------
-bottom, top = 0, 1
-thickness = top - bottom
-colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
-for b_factor in b_factors:
-    denominator = np.exp(- bottom * b_factor / thickness) - \
-                  np.exp(- top * b_factor / thickness)
-    amplitude = (density_bottom - density_top) / denominator
-    constant_term = (
-        density_top * np.exp(-bottom * b_factor / thickness) -
-        density_bottom * np.exp(-top * b_factor / thickness)
-        ) / denominator
-
-    # Define density function
-    def density_exponential(height):
-        return amplitude*np.exp(-height * b_factor / thickness) + constant_term
-
-    heights = np.linspace(bottom, top, 101)
-    plt.plot(heights, density_exponential(heights), color=colors[b_factor],
-             label="b={}".format(b_factor))
-plt.ylabel(r"Density [kg/m$^3$]")
-plt.xticks([bottom, top], ["Inner Radius", "Outer Radius"])
-plt.legend()
-plt.show()
 
 
 # Compute differences
@@ -117,85 +88,140 @@ for field in fields:
         top, bottom = model.bounds[4], model.bounds[5]
         thickness = top - bottom
 
[…]

### 2018-12-13 code/scripts/grid_search.py
Fix latex subscript in grid search figure

@@ -219,7 +219,7 @@ for field, ax in zip(fields, axes):
         label = labels[i]
         label = label.replace("$", "")
         if float(label) == D_linear[field]:
-            labels[i] = r"$D_\text{linear}$"
+            labels[i] = r"$D_\mathrm{linear}$"
     ax.set_xticklabels(labels)
 
 plt.tight_layout(pad=0.5, h_pad=0, w_pad=0)

### 2018-12-13 manuscript/manuscript.tex
Update grid search text

@@
 gravitational potential ($V$) and the vertical component of its gradient ($g_z$) on the
-computation grids detailed in Table~\ref{tab:grids} and using shell models with the same
-thicknesses as the linear density case.
-Because this is a time consuming computation, we limit the analysis only to the second
-collection of tesseroids in Table~\ref{tab:tess-collections}.
+computation grids detailed in Table~\ref{tab:grids}.
+Because this is a time consuming computation, we limit the analysis only to five
+spherical shell models made out of the second tesseroid collection from
+Table~\ref{tab:tess-collections} with the five different thicknesses, and only one value
+of $b$ equal to 30 (Fig.~\ref{fig:exp-densities}).
 For the sake of brevity, we compute the maximum differences between the results
-involving each computation grid, shell thickness and value of the $b$ factor.
+involving each shell thickness.
 The results are shown in Fig.~\ref{fig:grid-search}, in which points inside the dotted
@@
 as the ones for the linear density case.
+
 Because the grid search was limited to specific values of $b$ and used only one grid, we
-will determine the optimal value of $\delta$ for exponential density on the next
-subsection.
+will determine the optimal value of $\delta$ for exponential density tesseroids on the
+next subsection.
 
@@
 \iftwocol{
-\includegraphics[width=0.9\linewidth]
+\includegraphics[width=\linewidth]
     {figures/grid-search.pdf}

### 2018-12-13 manuscript/manuscript.tex
Update delta determination

@@
     with only five spherical shell models made out of the second tesseroid collection
-    from Table~\ref{tab:tess-colections} with the different thicknesses defined in the
+    from Table~\ref{tab:tess-collections} with the different thicknesses defined in the
     same Table, and density function with $b = 30$.
@@
     0.1\%.
+    The $D$ value obtained for the linear density case for each gravitational field is
+    also shown.
     }
@@
 
-Having chosen values of $D$ equal to the linear density case, we are free to explore the
-integration error as a function of $\delta$ in more detail and how it varies for
-different values of $b$ (Eq.~\ref{eq:density-exp}).
-We perform the error calculations for the ``Big Grid'' configuration
-(Table~\ref{tab:grids}) for two spherical shell models with 1km and 35km thickness.
+Having chosen values of $D$ equal to the ones obtained for the linear density case, we
+are free to explore the integration error as a function of $\delta$ in more detail and
+how it varies for different values of $b$ (Eq.~\ref{eq:density-exp}).
+Now we perform the error calculations for the every grid detailed in
+Table~\ref{tab:grids} for the 20 spherical shells models detailed in
+Table~\ref{tab:tess-collections}.
 The calculations are repeated for different values of $b$ to examine the error variation

### 2018-12-13 code/scripts/grid_search.py
Fix grid search dashed line contour

@@ -184,12 +184,12 @@ for field, ax in zip(fields, axes):
     plt.colorbar(cm, label=r"Differences (\%)", ax=ax)
 
     # Bellow error points contour
-    D_step = abs((D_grid.max() - D_grid.min())/(D_grid.shape[1] - 1))
-    delta_factor = abs(delta_grid.max()/delta_grid.min())**(1/(delta_grid.shape[0] - 1))
-    min_D = np.min(D_grid[differences <= 0.1]) - D_step/2
+    D_step = (D_values.max() - D_values.min()) / D_values.size
+    delta_factor = (delta_values.max()/delta_values.min())**(1/(delta_values.size - 1))
+    min_D = np.min(D_grid[total_differences <= 0.1]) - D_step/2
     max_D = D_grid.max() + D_step/8
     min_delta = delta_grid.min()*(delta_factor**(-0.125))
-    max_delta = np.max(delta_grid[differences <= 0.1])*(delta_factor**0.5)
+    max_delta = np.max(delta_grid[total_differences <= 0.1])*(delta_factor**0.5)
     ax.plot([min_D, min_D, max_D], [min_delta, max_delta, max_delta], '--', color='C7')
 
     # Add field title

### 2018-12-13 manuscript/manuscript.tex
Improve delta ratio determination for exp density

@@
 \centering
-\includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
+\includegraphics[width=\linewidth]{figures/exponential-density-diffs.pdf}
 \caption{
@@
     ratio.
-    The top panels show the exponential density functions for different values of $b$
-    (Eq.~\ref{eq:density-exp}).
-    The error computations were performed on the ``Big Grid'' configuration described in
-    Table~\ref{tab:grids}, with fixed values of the distance-size ratio
-    $D$ obtained for the linear density case ($D=1, 2, 8$ for the
-    potential, gradient components, and Marussi tensor components, respectively).
+    The percentage difference values were obtained from the comparison between the
+    analytical solution and the numerical approximation of the gravitational fields ($V$
+    and $g_z$) generated by a spherical shell with an exponential density function and
+    different values of $b$ (Eq.~\ref{eq:density-exp}).
+    These comparisons were carried out on every grid defined on Table~\ref{tab:grids},
+    making use of the 20 shell models detailed on Table~\ref{tab:tess-collections} and
+    fixing the distance-size ratio $D$ to the values obtained for the linear density
+    case ($D = 1, 2.5$ for the potential and $g_z$, respectively).
     The difference is reported as a percentage of the shell values.
+    To improve visualization of the results, each curve corresponds to the maximum
+    difference between every shell model.
     }
@@
 Now we perform the error calculations for the every grid detailed in
-Table~\ref{tab:grids} for the 20 spherical shells models detailed in
+Table~\ref{tab:grids} and for the 20 spherical shells models detailed in
 Table~\ref{tab:tess-collections}.
-The calculations are repeated for different values of $b$ to examine the error variation
-with the sharpness of the density function.
+The calculations are repeated for different values of $b$ (1, 2, 5, 10, 30, 100) to
+examine the error variation with the sharpness of the density function.
 Because larger $\delta$ values result in fewer tesseroid divisions,
@@
 
-Fig.~\ref{fig:delta-exponential} shows the density functions and the resulting relative
-differences ($V$, $g_z$, and $g_{zz}$) for different values of $b$ and thickness of the
-shell and tesseroid model.
-The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold for
-$\delta = 0.2$ in both the thick and thin shell cases.
-On the other hand, optimal values of $\delta$ for $g_{zz}$
-are $\delta = 0.2$ for the thin shell and $\delta = 0.01$ for the thick shell.
-Once again, we opt for the conservative value of $\delta = 0.01$ and sacrifice
-performance for the sake of accuracy.
+Fig.~\ref{fig:exp-densities} shows the density functions for each value of $b$ and
+Fig.~\ref{fig:delta-exponential} shows the resulting relative
+differences ($V$ and $g_z$).
+In order to improve the visualization of the results, each curve corresponds to the
+maximum difference between all shell models.
+The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold
+for each computation grid, shell model and $b$ value if $\delta = 0.1$.
+These results also show that there is no significant relation between the $b$
+value and the numerical error, proving that the density-based discretization algorithm
+is capable of limiting the error due to the density variation regardless the variation
+of the exponential density.
 

### 2018-12-13 code/scripts/exponential_density.py
Fix too big figures for linear and exp densities

@@ -190,7 +190,7 @@ grid_titles = {"pole": "Pole",
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
 # Create outer grid
-fig = plt.figure(figsize=(6.66, 8))
+fig = plt.figure(figsize=(6.66, 7))
 outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 
 # Create grid specs for each grid

### 2018-12-13 code/scripts/linear_density.py

@@ -126,7 +126,7 @@ colors = dict(zip(thicknesses, plt.cm.viridis(np.linspace(0, 0.9, len(thicknesse
 markers = dict(zip(thicknesses, ["o", "^", "s", "d", "x"]))
 
 # Create outer grid
-fig = plt.figure(figsize=(6.66, 8))
+fig = plt.figure(figsize=(6.66, 7))
 outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 
 # Create grid specs for each grid

### 2018-12-13 manuscript/manuscript.tex
Minor improvements in manuscript

@@
 Because the grid search was limited to specific values of $b$ and used only one grid, we
-will determine the optimal value of $\delta$ for exponential density tesseroids on the
-next subsection.
+will perform a more detailed analysis for $\delta$ determination in case of exponential
+density tesseroids.
 
@@
     fixing the distance-size ratio $D$ to the values obtained for the linear density
-    case ($D = 1, 2.5$ for the potential and $g_z$, respectively).
+    case.
     The difference is reported as a percentage of the shell values.
-    To improve visualization of the results, each curve corresponds to the maximum
-    difference between every shell model.
+    Each curve corresponds to the maximum difference between every shell model.
     }

### 2018-12-13 manuscript/manuscript.tex
Start writing sine density subsection

@@
 how it varies for different values of $b$ (Eq.~\ref{eq:density-exp}).
-Now we perform the error calculations for the every grid detailed in
-Table~\ref{tab:grids} and for the 20 spherical shells models detailed in
-Table~\ref{tab:tess-collections}.
+Now we perform the error calculations for values of delta ranging from $10^{-3}$ to
+$10^{1}$, for every grid detailed in Table~\ref{tab:grids} and for the 20 spherical
+shells models detailed in Table~\ref{tab:tess-collections}.
 The calculations are repeated for different values of $b$ (1, 2, 5, 10, 30, 100) to
@@
 of the exponential density.
+
+\subsection{Trigonometric Density}
+
+So far we have tested the density-based discretization algorithm against linear and
+exponential density functions.
+Nevertheless, this new algorithm is well suited to accurately reproduce the
+gravitational fields of tesseroids with more complex continuous functions, like
+non-monotonic or with one or more inflection points.
+Although such density functions are very rare (or maybe even non-existent) between
+geological structures, we want to subject the density-based algorithm to such cases in
+order to show that it can solve more complex situations other than linear and
+exponential functions.
+
+We will consider spherical shells with a trigonometric density defined as follows:
+
+\begin{equation}
+    \rho(r') = A \sin \left( 2 \pi b \frac{r' - R}{T} \right) + A,
+    \label{eq:sine-density}
+\end{equation}
+
+\noindent where $A$ is a constant, $T$ is the thickness of the shell, $R$ is the mean
+Earth radius and $b$ is a dimensionless constant that regulates how many periods of the
+trigonometric function are included inside the inner and outer radii.
+For this analysis we will use values of $b$ equal to 1, 2, 5 and 10
+(Fig.~\ref{fig:sine-densities}).
+We shift the sine function in order to prevent near zero values of the potential and the
+$g_z$.
+
+We compute the relative difference between the numerical and analytical results for the
+gravitational potential ($V$) and the $g_z$ of spherical shells with this kind of
+sinusoidal density function.
+The analytical solution for the potential $V$ can be found on Appendix~\ref{sec:shell}.
+The gravitational fields are computed on every grid detailed in Table~\ref{tab:grids}
+and making use of the spherical shells defined on Table~\ref{tab:tess-collections}.
+Each shell is assigned with different sinusoidal density functions by changing the
+values of $b$ in Eq.~\ref{eq:sine-density} (Fig.~\ref{fig:sine-densities}).
 
@@
 
+\subsection{Sinusoidal density}
+
+For an sinusoidal density function
+
+\begin{equation}
+    \rho(r') = A \sin ( k (r' - R)),
+\end{equation}
+
+\noindent
+where $A$, $k$ and $R$ are constants, the gravitational potential on any external point
+is
+
+\begin{equation}
+    \begin{split}
+    V_\text{sine}(r) = \frac{4\pi G}{r} \frac{A}{k^3} \Big[
+    & (2 - k^2 R_2^2) \cos(k(R_2 - R)) + 2 k R_2 \sin(k(R_2 - R)) -
+    & (2 - k^2 R_1^2) \cos(k(R_1 - R)) - 2 k R_1 \sin(k(R_1 - R))
+        \Big]
+    \end{split}
+\end{equation}
+
 \end{document}

### 2018-12-13 code/scripts/linear_density.py
Add comments in linear density script

@@ -65,17 +65,19 @@ fields = 'potential gz'.split()
 D_values = np.arange(0.5, 5.5, 0.5)
 for field in fields:
     for model in models:
+        # Define density function
         top, bottom = model.bounds[4], model.bounds[5]
         slope = (2670 - 3300) / (top - bottom)
         constant_term = 2670 - slope * (top + MEAN_EARTH_RADIUS)
 
-        # Define density function
         def density_linear(height):
             r = height + MEAN_EARTH_RADIUS
             return slope * r + constant_term
 
+        # Append density function to every tesseroid of the model
         model.addprop("density", [density_linear for i in range(model.size)])
 
+        # Compute differences
         for grid_name, grid in grids.items():
             fname = "{}-{}-{}-{}.npz".format(field, grid_name, int(top - bottom),
                                              model.size)
@@ -95,6 +97,8 @@ for field in fields:
                 diff = 100 * np.max(diff)
                 differences.append(diff)
             differences = np.array(differences)
+
+            # Save results to file
             np.savez(os.path.join(result_dir, fname),
                      D_values=D_values, differences=differences)
 

### 2018-12-14 code/scripts/linear_density.py
Remove multi shapes in linear density script
We keep only the (1, 6, 12) shape for the D and delta determination.
We've seen that the accuracy does not depend on the shape and it's very
time consuming to perform every test with several shapes.

@@ -42,11 +42,10 @@ if not os.path.isdir(result_dir):
 # Define Tesseroids models
 # ------------------------
 thicknesses = [100, 1e3, 1e4, 1e5, 1e6]
-shapes = [(1, 1, 2), (1, 6, 12), (1, 12, 24), (1, 18, 36)]
+shape = (1, 6, 12)
 models = [
     TesseroidMesh((0, 360, -90, 90, 0, -thickness), shape)
     for thickness in thicknesses
-    for shape in shapes
 ]
 
 
@@ -67,6 +66,7 @@ for field in fields:
     for model in models:
         # Define density function
         top, bottom = model.bounds[4], model.bounds[5]
+        thickness = top - bottom
         slope = (2670 - 3300) / (top - bottom)
         constant_term = 2670 - slope * (top + MEAN_EARTH_RADIUS)
 
@@ -79,8 +79,7 @@ for field in fields:
 
         # Compute differences
         for grid_name, grid in grids.items():
-            fname = "{}-{}-{}-{}.npz".format(field, grid_name, int(top - bottom),
-                                             model.size)
+            fname = "{}-{}-{}.npz".format(field, grid_name, int(thickness))
             if os.path.isfile(os.path.join(result_dir, fname)):
                 continue
             print("Thickness: {} Model size: {} Field: {} Grid: {}".format(
@@ -172,25 +171,19 @@ for grid_name in grid_names:
     # Plot
     for ax, field in zip(axes, fields):
         field_title = field_titles[field]
-        for thickness in thicknesses:
-            differences_per_thickness = []
+        for model in models:
+            thickness = model.bounds[4] - model.bounds[5]
             color = colors[thickness]
             marker = markers[thickness]
-            for model in models:
-                if model.bounds[4] - model.bounds[5] == thickness:
-                    fname = "{}-{}-{}-{}.npz".format(field, grid_name, int(thickness),
-                                                     model.size)
-                    diff_file = np.load(os.path.join(result_dir, fname))
-                    D_values = diff_file["D_values"]
-                    differences = diff_file["differences"]
-                    differences_per_thickness.append(differences)
-            differences_per_thickness = np.array(differences_per_thickness)
-            differences_per_thickness = np.max(differences_per_thickness, axis=0)
+            fname = "{}-{}-{}.npz".format(field, grid_name, int(thickness))
+            diff_file = np.load(os.path.join(result_dir, fname))
+            D_values = diff_file["D_values"]
[…]

### 2018-12-14 manuscript/manuscript.tex
Fix missing line break in equation split
Fix shell models to one single shape

@@
 \caption{
-    Description of the tesseroid collections used to build a spherical shell and
-    characterize the accuracy of the numerical integration.
-    The first three columns describe the shapes of the tesseroid collections: each one
-    specifies the number of tesseroids on the corresponding dimension (radial,
-    latitudinal, longitudinal).
+    Description of the tesseroid models used to build spherical shells and characterize
+    the accuracy of the numerical integration.
+    The first column indicates the thickness of the shell model and the next three
+    describe the shape of the tesseroid model: each one specifies the number of
+    tesseroids on the corresponding dimension (radial, latitudinal, longitudinal).
     The last two columns specify the latitudinal and longitudinal size of each
     tesseroid that compose the spherical shell model.
-    For each collection, we define five spherical shell models with an outer radius
-    equal to the mean Earth radius (6378.137 km) and the inner radius determined by the
-    following thicknesses: 100m, 1km, 10km, 100km and 1000km.
-    This makes a total of 20 shell models combining tesseroids collections and
-    thicknesses.
-}
-\label{tab:tess-collections}
+    The outer radius of every shell model is equal to the mean Earth radius (6378.137
+    km), while the inner radius is determined by its thickness.
+}
+\label{tab:shell-models}
 \begin{tabular}{cccccc}
-    Radial & Latitudinal & Longitudinal  & Total tesseroids & $L_\phi$ & $L_\lambda$
-    \\ \hline
-    1 & 1  & 2  & 2   & 180$^\circ$ & 180$^\circ$ \\
-    1 & 6  & 12 & 72  &  30$^\circ$ &  30$^\circ$ \\
-    1 & 12 & 24 & 288 &  15$^\circ$ &  15$^\circ$ \\
-    1 & 18 & 36 & 648 &  10$^\circ$ &  10$^\circ$ \\
+    Thickness & Radial & Latitudinal & Longitudinal & $L_\phi$ & $L_\lambda$ \\ \hline
+    100m   & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
+    1km    & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
+    10km   & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
+    100km  & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
+    1000km & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
 \end{tabular}
@@
 
-In order to compare the numerical results with the analytical solution we
-must build spherical shell models made of tesseroids.
-These shell models are determined by their inner and outer radii and by the collection
-of tesseroids used to build the shell.
-For this analysis we have chosen to set the outer radii of every shell model equal to
-the mean Earth radius (6378.137 km), while the inner radius is determined by the desired
-thickness of the shell.
-We choose  different values of shell thickness (100m, 1km, 10km, 100km and 1000km) in
-order to account for different computation scenarios.
-Furthermore, we propose different tesseroids collections to build the shell model by
-regularly splitting it along the longitudinal and latitudinal directions.
-Each collection, described in Table~\ref{tab:tess-collections}, is composed by
-a different number of tesseroids, i.e.~with tesseroids of different sizes.
-Finally, combining these collections and the thicknesses we obtain 20 different shell
-models.
+In order to compare the numerical results with the analytical solution we must build
+spherical shell models made of tesseroids: we choose to regularly split the spherical
+shell along the latitudinal and longitudinal directions obtaining a shell model made out
+of $30^\circ \times 30^\circ$ tesseroids.
+In order to cover different computation scenarios we use five spherical shells, each one
+with the outer radius equal to the mean Earth radius and the inner radius determined by
+the desired thickness (100m, 1km, 10km, 100km and 1000km).
+These shell models are detailed on Table~\ref{tab:shell-models}.
 
@@
 the mean orbital height of the gravimetric satellite missions (Satellite).
-We define these grids in the spirit of span a broad scenario of gravitational field
+We define these grids in the spirit of spanning a broad scenario of gravitational field
 computations and ensure an acceptable accuracy on each one of them.
@@
 Because the outer radius of every spherical shell model will be equal to the mean Earth
-radius, $R_\text{out} = 6378.137$km.
+radius, $R_\text{out} = 6378.137 \text{km}$.
 
@@
     Each model has a linear density function (Eq.~\ref{eq:density-linear}).
-    The computations were performed on the four grids described in
-    Table~\ref{tab:grids} and using 20 shell models built with the tesseroids
-    collections detailed in Table~\ref{tab:tess-collections} and five different
-    thicknesses.
-    Each line represents the maximum absolute difference between every shell model with
-    a certain thickness.
+    The computations were performed on the four grids described in Table~\ref{tab:grids}
+    and using the shell models detailed in Table~\ref{tab:shell-models}.
+    Each line represents the maximum absolute difference between every shell model.
     Due to the linearity of the density function, the density-based discretization
@@
 We compute the gravitational potential ($V$) and the vertical component of the gradient
-($g_z$) that each shell model generates on each grid defined in Table~\ref{tab:grids}.
+($g_z$) that each shell model generates on each computation grid.
 Other components of the gradient are equal to zero outside of the shell due to the
@@
 This differences are shown in Fig.~\ref{fig:D-linear}.
-In order to simplify the visualization of the results, only the maximum
-differences corresponding to tesseroids collections with equal thickness are shown
-in the Figure, obtaining one curve for every shell thickness.
 Finally, we set the optimal value of $D$ as the minimum value at which the corresponding
@@
 These results also show that there's no meaningful dependency between the accuracy of
-the method and the thickness of the shell, the collection of tesseroids used to build
-it or the height of the computation grid (the results for the Satellite grid and the
-Global grid are very similar).
+the method and the thickness of the shell or the height of the computation grid (the
+results for the Satellite grid and the Global grid are very similar).
 
@@
 computation grids detailed in Table~\ref{tab:grids}.
-Because this is a time consuming computation, we limit the analysis only to five
-spherical shell models made out of the second tesseroid collection from
-Table~\ref{tab:tess-collections} with the five different thicknesses, and only one value
-of $b$ equal to 30 (Fig.~\ref{fig:exp-densities}).
-For the sake of brevity, we compute the maximum differences between the results
-involving each shell thickness.
+Because this is a time consuming computation, we limit the analysis only to a single
+value of $b$ equal to 30 (Fig.~\ref{fig:exp-densities}).
+For the sake of brevity, we compute the maximum differences between the results for each
+shell model.
 The results are shown in Fig.~\ref{fig:grid-search}, in which points inside the dotted
@@
     These comparisons were carried out on the ``Global'' grid (Table~\ref{tab:grids}),
-    with only five spherical shell models made out of the second tesseroid collection
-    from Table~\ref{tab:tess-collections} with the different thicknesses defined in the
-    same Table, and density function with $b = 30$.
+    with the spherical shell models detailed in Table~\ref{tab:shell-models}, and an
+    exponential density function with $b = 30$.
     The percentage difference values are obtained as the maximum difference
-    corresponding to each spherical shell model.
+    between every shell model.
     The points inside the dashed line are the ones that present an error lower than
@@
     The $D$ value obtained for the linear density case for each gravitational field is
-    also shown.
+    also shown ($D_\text{linear}$).
     }
@@
     analytical solution and the numerical approximation of the gravitational fields ($V$
-    and $g_z$) generated by a spherical shell with an exponential density function and
+    and $g_z$) generated by the spherical shell with an exponential density function and
     different values of $b$ (Eq.~\ref{eq:density-exp}).
     These comparisons were carried out on every grid defined on Table~\ref{tab:grids},
-    making use of the 20 shell models detailed on Table~\ref{tab:tess-collections} and
+    making use of the shell models detailed on Table~\ref{tab:shell-models} and
     fixing the distance-size ratio $D$ to the values obtained for the linear density
     case.
-    The difference is reported as a percentage of the shell values.
-    Each curve corresponds to the maximum difference between every shell model.
+    Each curve corresponds to the maximum difference between every shell model with the
+    same value of $b$.
     }
@@
 Now we perform the error calculations for values of delta ranging from $10^{-3}$ to
-$10^{1}$, for every grid detailed in Table~\ref{tab:grids} and for the 20 spherical
-shells models detailed in Table~\ref{tab:tess-collections}.
-The calculations are repeated for different values of $b$ (1, 2, 5, 10, 30, 100) to
-examine the error variation with the sharpness of the density function.
+$10^{1}$, for every grid detailed in Table~\ref{tab:grids} and for the spherical shell
+models detailed in Table~\ref{tab:shell-models}.
+The calculations are repeated for distinct values of $b$ (1, 2, 5, 10, 30, 100) to
+examine the accuracy of the method for different sharpness of the density function.
 Because larger $\delta$ values result in fewer tesseroid divisions,
@@
 In order to improve the visualization of the results, each curve corresponds to the
-maximum difference between all shell models.
+maximum difference between all shell models for the same value of $b$.
 The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold
@@
 The gravitational fields are computed on every grid detailed in Table~\ref{tab:grids}
-and making use of the spherical shells defined on Table~\ref{tab:tess-collections}.
+and making use of the spherical shells defined on Table~\ref{tab:shell-models}.
 Each shell is assigned with different sinusoidal density functions by changing the
@@
     V_\text{sine}(r) = \frac{4\pi G}{r} \frac{A}{k^3} \Big[
-    & (2 - k^2 R_2^2) \cos(k(R_2 - R)) + 2 k R_2 \sin(k(R_2 - R)) -
+    & (2 - k^2 R_2^2) \cos(k(R_2 - R)) + 2 k R_2 \sin(k(R_2 - R)) - \\
     & (2 - k^2 R_1^2) \cos(k(R_1 - R)) - 2 k R_1 \sin(k(R_1 - R))

### 2018-12-14 code/scripts/sine_density.py
Fix sine density figsize

@@ -177,7 +177,7 @@ grid_titles = {"pole": "Pole",
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
 # Create outer grid
-fig = plt.figure(figsize=(6.66, 8))
+fig = plt.figure(figsize=(6.66, 7))
 outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 
 # Create grid specs for each grid

### 2018-12-14 manuscript/manuscript.tex
Fix grid table
Continue writing sinusoidal density subsection

@@
     Equator   & $0.1^\circ \times 0.1^\circ$ &   0E /   1E /  0N / 1N  & 0   \\
-    Global    & $ 10^\circ \times  10^\circ$ & 0E / 1E / 0N  / 1N   & 0   \\
-    Satellite & $ 10^\circ \times  10^\circ$ & 0E / 1E / 89N / 90N  & 260 \\
+    Global    & $ 10^\circ \times  10^\circ$ & 180W / 180E / 90S / 90N & 0   \\
+    Satellite & $ 10^\circ \times  10^\circ$ & 180W / 180E / 90S / 90N & 260 \\
 \end{tabular}
@@
 
-\subsection{Trigonometric Density}
+\subsection{Sinusoidal Density}
 
@@
 
-We will consider spherical shells with a trigonometric density defined as follows:
+We will consider spherical shells with a sinusoidal density defined as follows:
 
@@
     \rho(r') = A \sin \left( 2 \pi b \frac{r' - R}{T} \right) + A,
-    \label{eq:sine-density}
+    \label{eq:density-sine}
 \end{equation}
@@
 We compute the relative difference between the numerical and analytical results for the
-gravitational potential ($V$) and the $g_z$ of spherical shells with this kind of
-sinusoidal density function.
+gravitational potential ($V$) and the $g_z$ of the spherical shells detailed in
+Table~\ref{tab:shell-models} on each grid detailed in Table~\ref{tab:grids}.
+Each shell is assigned with different sinusoidal density functions by changing the
+values of $b$ in Eq.~\ref{eq:density-sine} (Fig.~\ref{fig:sine-densities}).
+We fixed the distance-size ratio $D$ to the ones obtained for the linear density case
+and explored values of $\delta$ ranging from $10^{-4}$ to $1$.
 The analytical solution for the potential $V$ can be found on Appendix~\ref{sec:shell}.
-The gravitational fields are computed on every grid detailed in Table~\ref{tab:grids}
-and making use of the spherical shells defined on Table~\ref{tab:shell-models}.
-Each shell is assigned with different sinusoidal density functions by changing the
-values of $b$ in Eq.~\ref{eq:sine-density} (Fig.~\ref{fig:sine-densities}).
+
+Fig.~\ref{fig:sine-densities} shows the sinusoidal density functions assigned to every
+spherical shell and Fig.~\ref{fig:delta-sine} shows the resulting relative differences
+between the analytical and the numerical solutions.
+In order to improve the visualization of the results, each curve corresponds to the
+maximum difference between all shell models for the same value of $b$.
+
+\begin{figure}
+\centering
+\iftwocol{
+\includegraphics[width=\linewidth]{figures/sine-densities.pdf}
+}{
+\includegraphics[width=0.5\linewidth]{figures/sine-densities.pdf}
+}
+\caption{
+    Sinusoidal density functions assigned to each spherical shell in the $\delta$ ratios
+    determination.
+    Each density function corresponds to a different value of $b$ on
+    Eq.~\ref{eq:density-sine}.
+}
+\label{fig:sine-densities}
+\end{figure}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{figures/sine-density-diffs.pdf}
+\caption{
+    Numerical error for different sinusoidal density functions and values of the delta
+    ratio.
+    The percentage difference values were obtained from the comparison between the
+    analytical solution and the numerical approximation of the gravitational fields ($V$
+    and $g_z$) generated by the spherical shell with an exponential density function and
+    different values of $b$ (Eq.~\ref{eq:density-sine}).
+    These comparisons were carried out on every grid defined on Table~\ref{tab:grids},
+    making use of the shell models detailed on Table~\ref{tab:shell-models} and
+    fixing the distance-size ratio $D$ to the values obtained for the linear density
+    case.
+    Each curve corresponds to the maximum difference between every shell model with the
+    same value of $b$.
+    }
+\label{fig:delta-exponential}
+\end{figure*}
 

### 2018-12-14 manuscript/manuscript.tex
Remove tensor components from Appendix

@@
 
-The gradient and the Marussi tensor of potentials that
-depends solely on $r$ have only a few non zero components: the vertical
-component of the gradient ($g_z$) and the diagonal components of the
-tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$).
+The gradient of potentials that
+depends solely on $r$ have only one non zero components: the vertical
+component of the gradient ($g_z$).
 Following \citet{Grombein2013}:
@@
     g_z(r) = \frac{V_\text{sh}(r)}{r},
-\end{equation}
-
-\begin{equation}
-    g_{xx}(r) = g_{yy}(r) = -\frac{V_\text{sh}(r)}{r^2},
-\end{equation}
-
-\begin{equation}
-    g_{zz}(r) = \frac{2V_\text{sh}(r)}{r^2}.
 \end{equation}

### 2018-12-14 manuscript/manuscript.tex
Minor modification in manuscript

@@
 for each computation grid, shell model and $b$ value if $\delta = 0.1$.
+This results imply that a value of $\delta=0.1$ is needed in order to
+achieve the 0.1\% accuracy threshold for both the potential and the $g_{z}$, every
+shell model, computation grid and value of $b$, in case of exponential density.
 These results also show that there is no significant relation between the $b$
@@
 The analytical solution for the potential $V$ can be found on Appendix~\ref{sec:shell}.
-
-Fig.~\ref{fig:sine-densities} shows the sinusoidal density functions assigned to every
-spherical shell and Fig.~\ref{fig:delta-sine} shows the resulting relative differences
-between the analytical and the numerical solutions.
-In order to improve the visualization of the results, each curve corresponds to the
-maximum difference between all shell models for the same value of $b$.
 
@@
 \end{figure*}
+
+Fig.~\ref{fig:sine-densities} shows the sinusoidal density functions assigned to every
+spherical shell and Fig.~\ref{fig:delta-sine} shows the resulting relative differences
+between the analytical and the numerical solutions.
+In order to improve the visualization of the results, each curve corresponds to the
+maximum difference between all shell models for the same value of $b$.
+For this kind of sinusoidal density, a value of $\delta=0.01$ is needed in order to
+achieve the 0.1\% accuracy threshold for both the potential and the $g_{z}$, every
+shell model, computation grid and value of $b$.
+This results also proof that even for a continuous density function with several local
+maximum and minimums and inflection points, the density-based discretization algorithm
+performs well on computing the gravitational fields that a tesseroid with such density
+generates.
 

### 2018-12-14 code/scripts/exponential_density.py
Improve exp and sine figures

@@ -249,7 +249,7 @@ for grid_name in grids:
             ax.plot(delta_values, differences_per_b, "-o", color=color, label=label)
 
         # Add threshold line
-        ax.plot([1e-3, 1e1], [1e-1, 1e-1], '--', color='k', linewidth=0.5)
+        ax.plot([1e-4, 1e2], [1e-1, 1e-1], '--', color='k', linewidth=0.5)
 
         # Add field annotation on each axe
         ax.text(0.5, 0.87, field_title, fontsize=11,
@@ -264,6 +264,7 @@ for grid_name in grids:
         # Configure axes
         ax.set_xscale('log')
         ax.set_yscale('log')
+        ax.set_xlim(6.31e-4, 1.58e1)
         ax.set_yticks(ax.get_yticks()[2:-2])
         ax.set_ylabel(r'Difference (\%)')
         ax.grid(True, linewidth=0.5, color='#aeaeae')

### 2018-12-14 code/scripts/sine_density.py

@@ -157,7 +157,19 @@ for b_factor, ax in zip(b_factors, axes):
             label="b={}".format(b_factor))
     ax.set_xticks([bottom, top])
     ax.set_xticklabels(["Inner Radius", "Outer Radius"])
-    ax.legend(loc=1)
+    # ax.legend(loc=1)
+
+    # Add field annotation on each axe
+    ax.text(0.85, 0.8, "b={}".format(b_factor),
+            fontsize=10,
+            horizontalalignment='center',
+            verticalalignment='center',
+            bbox={'facecolor': 'w',
+                  'edgecolor': '#9b9b9b',
+                  'alpha': 0.7,
+                  'linewidth': 0.5, 'pad': 5,
+                  'boxstyle': 'square, pad=0.4'},
+            transform=ax.transAxes)
 fig.text(0, 0.5, r"Density [kg/m$^3$]", va='center', rotation='vertical')
 plt.tight_layout(pad=1.8)
 fig.subplots_adjust(hspace=0)
@@ -236,7 +248,7 @@ for grid_name in grids:
             ax.plot(delta_values, differences_per_b, "-o", color=color, label=label)
 
         # Add threshold line
-        ax.plot([1e-4, 1e0], [1e-1, 1e-1], '--', color='k', linewidth=0.5)
+        ax.plot([1e-5, 1e1], [1e-1, 1e-1], '--', color='k', linewidth=0.5)
 
         # Add field annotation on each axe
         ax.text(0.5, 0.87, field_title, fontsize=11,
@@ -251,6 +263,7 @@ for grid_name in grids:
         # Configure axes
         ax.set_xscale('log')
         ax.set_yscale('log')
+        ax.set_xlim(6.31e-5, 1.58e0)
         ax.set_yticks(ax.get_yticks()[2:-2])
         ax.set_ylabel(r'Difference (\%)')
         ax.grid(True, linewidth=0.5, color='#aeaeae')

### 2018-12-14 code/scripts/computation_time.py
Improve computation time script
Add pandas DataFrame to store computation time data

@@ -1,8 +1,9 @@
 from __future__ import division, print_function
+import os
 import time
 import numpy as np
+import pandas as pd
 from fatiando.mesher import Tesseroid
-from fatiando import gridder
 import matplotlib.pyplot as plt
 # This is our custom tesseroid code
 from tesseroid_density import tesseroid
@@ -28,109 +29,128 @@ computation_points = [[np.array([0.]), np.array([0.]), np.array([height])]
                       for height in heights]
 
 
-# Compute gravitational fields for homogeneous tesseroid
-# ------------------------------------------------------
-for tess in model:
-    tess.addprop('density', 3000)
-
-times_homogeneous = {}
-for field in fields:
-    for point in computation_points:
-        times = []
-        lon, lat, height = point[:]
-        for run in range(nruns):
-            start_time = time.time()
-            getattr(tesseroid, field)(lon, lat, height, model)
-            end_time = time.time()
-            times.append(start_time - end_time)
-        times = np.mean(times)
-        if times_homogeneous[field]:
-            times_homogeneous[field].append(times)
-        else:
-            times_homogeneous[field] = [times]
+# Create results dir if it does not exist
+# -------------------------
+script_path = os.path.dirname(os.path.abspath(__file__))
+result_dir = os.path.join(script_path, 'results/computation_time')
+if not os.path.isdir(result_dir):
+    os.makedirs(result_dir)
 
 
-# Compute gravitational fields for linear density tesseroid
-# ---------------------------------------------------------
-density_top, density_bottom = 2670, 3300
-slope = (density_top - density_bottom) / (top - bottom)
-constant_term = density_top - slope * top
+# Check if DataFrame exists
+# -------------------------
+csv_fname = os.path.join(result_dir, "computation_time.csv")
+compute = True
+if os.path.isfile(csv_fname):
+    compute = False
 
 
-# Define density function
-def density_linear(height):
-    return slope * height + constant_term
+# Compute only if DataFrame does not exists
+# -----------------------------------------
+if compute:
+
+    # Create DataFrame
+    # ----------------
+    columns = ["{}_{}".format(density, field)
+               for density in ["constant", "linear", "exp"]
+               for field in fields]
+    df = pd.DataFrame(data=np.zeros((heights.size, len(columns))),
+                      index=heights,
+                      columns=columns)
+
+    # Compute gravitational fields for homogeneous tesseroid
+    # ------------------------------------------------------
[…]

### 2018-12-17 code/scripts/computation_time.py
Increase number of runs in computation time script

@@ -11,7 +11,7 @@ from tesseroid_density import tesseroid
 
 # Configure comparison
 # --------------------
-nruns = 1000
+nruns = 10000
 fields = "potential gz".split()
 
 

### 2018-12-17 code/scripts/neuquen_basin.py
Rerun neuquen basin

@@ -21,7 +21,6 @@ if not os.path.isdir(result_dir):
 
 # Load digital elevation map of Neuquen, Argentina
 # -----------------------------------------------
-script_path = os.path.dirname(os.path.abspath(__file__))
 data = np.load(os.path.join(script_path, "../../data/topography.npy"))
 lat, lon, topo = data[:, 0], data[:, 1], data[:, 2]
 shape = (571, 457)
@@ -52,11 +51,12 @@ sediments = {'lon': lon,
 # Create top and bottom of sediments layer
 # ----------------------------------------
 # Regrid dem to sediments regular grid
-topo = gridder.interp_at(topography["lat"], topography["lon"], topography["topo"],
-                         sediments["lat"], sediments["lon"],
-                         algorithm="linear")
-sediments["top"] = topo
-sediments["bottom"] = topo - sediments["thickness"]
+sediments_top = gridder.interp_at(topography["lat"], topography["lon"],
+                                  topography["topo"], sediments["lat"],
+                                  sediments["lon"], algorithm="linear")
+sediments_top[np.isnan(sediments["thickness"])] = np.nan
+sediments["top"] = sediments_top
+sediments["bottom"] = sediments_top - sediments["thickness"]
 
 
 # Create Tesseroids model of the sediments layer
@@ -110,8 +110,10 @@ for field in fields:
 slope = (density_top - density_bottom) / (max_height - min_depth)
 constant_term = density_top - slope * max_height
 
+
 def linear_density(h):
     return slope * h + constant_term
+
 
 basin.addprop("density", [linear_density for i in range(basin.size)])
 
@@ -131,7 +133,7 @@ for field in fields:
 
 # Compute gravitational effect of the basin with exponential density
 # ------------------------------------------------------------------
-b_factor = 10
+b_factor = 3
 thickness = max_height - min_depth
 denominator = np.exp(- min_depth * b_factor / thickness) - \
               np.exp(- max_height * b_factor / thickness)
@@ -141,11 +143,13 @@ constant_term = (
     density_bottom * np.exp(-max_height * b_factor / thickness)
     ) / denominator
 
+
 # Define density function
 def density_exponential(h):
     return amplitude * np.exp(-h * b_factor / thickness) + constant_term
 
-basin.addprop("density", [linear_density for i in range(basin.size)])
+
+basin.addprop("density", [density_exponential for i in range(basin.size)])
 
 fields = "potential gz".split()
 for field in fields:

### 2018-12-18 manuscript/manuscript.tex
Fix wrong fig label
Minor modification in manuscript
Simplify exp density definition on manuscript

@@
 \begin{equation}
-    \rho(r') = A e^{- \frac{b}{T} (r' - R)} + C,
+    \rho(r') = A e^{- \frac{b}{T} (r' - R_1)} + C,
     \label{eq:density-exp}
@@
 \begin{equation}
-    A = \frac{\rho_\text{in} - \rho_\text{out}}{D},
-\end{equation}
-
-\begin{equation}
-    C = \frac{1}{D}
-    \left[\rho_\text{out} \, e^{- \frac{b}{T} (R_1 - R)} -
-    \rho_\text{in} \, e^{- \frac{b}{T} (R_2 - R)} \right],
-\end{equation}
-
-\begin{equation}
-    D = e^{- \frac{b}{T} (R_1 - R)} - e^{- \frac{b}{T} (R_2 - R)},
-\end{equation}
-
-\noindent $R$ is the mean Earth radius, $T$ is the thickness of the shell ($T = R_2
-- R_1$) and $b$ is a dimensionless constant that determines the variability of the
-function: a higher value of $b$ increases the maximum slope of the density function.
+    A = \frac{\rho_\text{in} - \rho_\text{out}}{1 - e^{-b}},
+\end{equation}
+
+\begin{equation}
+    C = \rho_\text{in} - A,
+\end{equation}
+
+\noindent $T$ is the thickness of the shell ($T = R_2 - R_1$) and $b$ is a dimensionless
+constant that determines the variability of the function: a higher value of $b$
+increases the maximum slope of the density function.
 On exponential density analysis we will use values of $b$ equal to 1, 2, 5, 10, 30
@@
 values of $b$ in Eq.~\ref{eq:density-sine} (Fig.~\ref{fig:sine-densities}).
+The analytical solution for the potential $V$ can be found on Appendix~\ref{sec:shell}.
 We fixed the distance-size ratio $D$ to the ones obtained for the linear density case
 and explored values of $\delta$ ranging from $10^{-4}$ to $1$.
-The analytical solution for the potential $V$ can be found on Appendix~\ref{sec:shell}.
 
@@
 \caption{
-    Sinusoidal density functions assigned to each spherical shell in the $\delta$ ratios
+    Sinusoidal density functions assigned to each spherical shell in the $\delta$ ratio
     determination.
@@
     }
-\label{fig:delta-exponential}
+\label{fig:delta-sine}
 \end{figure*}

### 2018-12-18 code/scripts/exponential_density.py
Simplify exp density function in exp script

@@ -18,11 +18,11 @@ def shell_exponential_density(height, top, bottom, amplitude, b_factor, constant
     the radial coordinate `height + MEAN_EARTH_RADIUS` with an exponential density as:
 
     .. math :
-        \rho(r') = A e^{- b * (r' - R) / T} + C
+        \rho(r') = A e^{- b * (r' - R_1) / T} + C
 
     Where $r'$ is the radial coordinate where the density is going to be evaluated,
-    $A$ is the amplitude, $C$ the constant term, $T$ is the thickness of the tesseroid,
-    $b$ the b factor and $R$ the mean Earth radius.
+    $A$ is the amplitude, $C$ the constant term, $T$ is the thickness of the shell,
+    $b$ the b factor and $R_1$ the inner radius.
     """
     r = height + MEAN_EARTH_RADIUS
     r1 = bottom + MEAN_EARTH_RADIUS
@@ -31,8 +31,8 @@ def shell_exponential_density(height, top, bottom, amplitude, b_factor, constant
     k = b_factor / thickness
     potential = 4 * np.pi * G * amplitude / k**3 / r * \
         (
-         ((r1 * k)**2 + 2*r1*k + 2) * np.exp(-k * bottom) -
-         ((r2 * k)**2 + 2*r2*k + 2) * np.exp(-k * top)
+         ((r1 * k)**2 + 2*r1*k + 2) -
+         ((r2 * k)**2 + 2*r2*k + 2) * np.exp(-k * thickness)
         ) + \
         4 / 3. * np.pi * G * constant_term * (r2**3 - r1**3) / r
     data = {'potential': potential,
@@ -90,17 +90,12 @@ for field in fields:
         thickness = top - bottom
 
         for b_factor in b_factors:
-            denominator = np.exp(- bottom * b_factor / thickness) - \
-                          np.exp(- top * b_factor / thickness)
-            amplitude = (density_bottom - density_top) / denominator
-            constant_term = (
-                density_top * np.exp(-bottom * b_factor / thickness) -
-                density_bottom * np.exp(-top * b_factor / thickness)
-                ) / denominator
+            A = (density_bottom - density_top) / (1 - np.exp(-b_factor))
+            C = density_bottom - A
 
             # Define density function
             def density_exponential(height):
-                return amplitude*np.exp(-height * b_factor / thickness) + constant_term
+                return A * np.exp(-b_factor * (height - bottom) / thickness) + C
 
             # Append density function to every tesseroid of the model
[…]

### 2018-12-18 code/scripts/grid_search.py
Simplify exp density in grid_search

@@ -18,11 +18,11 @@ def shell_exponential_density(height, top, bottom, amplitude, b_factor, constant
     the radial coordinate `height + MEAN_EARTH_RADIUS` with an exponential density as:
 
     .. math :
-        \rho(r') = A e^{- b * (r' - R) / T} + C
+        \rho(r') = A e^{- b * (r' - R_1) / T} + C
 
     Where $r'$ is the radial coordinate where the density is going to be evaluated,
-    $A$ is the amplitude, $C$ the constant term, $T$ is the thickness of the tesseroid,
-    $b$ the b factor and $R$ the mean Earth radius.
+    $A$ is the amplitude, $C$ the constant term, $T$ is the thickness of the shell,
+    $b$ the b factor and $R_1$ the inner radius.
     """
     r = height + MEAN_EARTH_RADIUS
     r1 = bottom + MEAN_EARTH_RADIUS
@@ -31,8 +31,8 @@ def shell_exponential_density(height, top, bottom, amplitude, b_factor, constant
     k = b_factor / thickness
     potential = 4 * np.pi * G * amplitude / k**3 / r * \
         (
-         ((r1 * k)**2 + 2*r1*k + 2) * np.exp(-k * bottom) -
-         ((r2 * k)**2 + 2*r2*k + 2) * np.exp(-k * top)
+         ((r1 * k)**2 + 2*r1*k + 2) -
+         ((r2 * k)**2 + 2*r2*k + 2) * np.exp(-k * thickness)
         ) + \
         4 / 3. * np.pi * G * constant_term * (r2**3 - r1**3) / r
     data = {'potential': potential,
@@ -89,16 +89,12 @@ for field in fields:
         thickness = top - bottom
 
         # Define density function
-        denominator = np.exp(- bottom * b_factor / thickness) - \
-            np.exp(- top * b_factor / thickness)
-        amplitude = (density_bottom - density_top) / denominator
-        constant_term = (
-            density_top * np.exp(-bottom * b_factor / thickness) -
-            density_bottom * np.exp(-top * b_factor / thickness)
-            ) / denominator
+        A = (density_bottom - density_top) / (1 - np.exp(-b_factor))
+        C = density_bottom - A
 
+        # Define density function
         def density_exponential(height):
-            return amplitude*np.exp(-height * b_factor / thickness) + constant_term
+            return A * np.exp(-b_factor * (height - bottom) / thickness) + C
 
         # Append density function to every tesseroid of the model
[…]

### 2018-12-18 code/scripts/neuquen_basin.py
Simplify exp density in neuquen_basin

@@ -82,7 +82,7 @@ grid = {'lat': lat,
 # Common variables between computations
 # -------------------------------------
 density_top, density_bottom = -412, -275
-max_height, min_depth = basin.top.max(), basin.bottom.min()
+top, bottom = basin.top.max(), basin.bottom.min()
 
 
 # Compute gravitational effect of the basin with homogeneous density
@@ -107,12 +107,9 @@ for field in fields:
 
 # Compute gravitational effect of the basin with linear density
 # -------------------------------------------------------------
-slope = (density_top - density_bottom) / (max_height - min_depth)
-constant_term = density_top - slope * max_height
-
-
 def linear_density(h):
-    return slope * h + constant_term
+    slope = (density_top - density_bottom) / (top - bottom)
+    return slope * (h - bottom) + density_bottom
 
 
 basin.addprop("density", [linear_density for i in range(basin.size)])
@@ -133,20 +130,12 @@ for field in fields:
 
 # Compute gravitational effect of the basin with exponential density
 # ------------------------------------------------------------------
-b_factor = 3
-thickness = max_height - min_depth
-denominator = np.exp(- min_depth * b_factor / thickness) - \
-              np.exp(- max_height * b_factor / thickness)
-amplitude = (density_bottom - density_top) / denominator
-constant_term = (
-    density_top * np.exp(-min_depth * b_factor / thickness) -
-    density_bottom * np.exp(-max_height * b_factor / thickness)
-    ) / denominator
-
-
-# Define density function
-def density_exponential(h):
-    return amplitude * np.exp(-h * b_factor / thickness) + constant_term
+def density_exponential(height):
+    b_factor = 3
+    thickness = top - bottom
+    A = (density_bottom - density_top) / (1 - np.exp(-b_factor))
+    C = density_bottom - A
+    return A * np.exp(-b_factor * (height - bottom) / thickness) + C
 
 
 basin.addprop("density", [density_exponential for i in range(basin.size)])

### 2018-12-18 code/scripts/neuquen_basin.py
Start writing neuquen basin plots

@@ -1,11 +1,14 @@
 from __future__ import division
 import os
+import warnings
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib import ticker
 from matplotlib.patches import Rectangle
 from matplotlib.colors import LightSource
 from mpl_toolkits.basemap import Basemap
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+from matplotlib.gridspec import GridSpec
 from fatiando import gridder
 from tesseroid_density import tesseroid
 from tesseroid_model import TesseroidModel
@@ -154,121 +157,184 @@ for field in fields:
              height=grid["height"], shape=grid["shape"])
 
 
+# ------------
 # Plot Results
 # ------------
+
+# Configure LaTeX style for plots
+# -------------------------------
+try:
+    plt.rcParams['text.usetex'] = True
+    plt.rcParams['font.family'] = 'serif'
+    plt.rcParams['font.serif'] = 'Computer Modern Roman'
+    plt.rcParams['xtick.major.size'] = 2
+    plt.rcParams['ytick.major.size'] = 2
+except Exception as e:
+    warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
+
+# Create basemap and configuration
+# --------------------------------
+bm = Basemap(projection='merc',
+             llcrnrlon=topography["area"][2],
+             llcrnrlat=topography["area"][0],
+             urcrnrlon=topography["area"][3],
+             urcrnrlat=topography["area"][1],
+             resolution='i')
+
+config = {'countries': dict(linewidth=0.5, color='k'),
+          'states': dict(linewidth=0.4, linestyle='--', color='k'),
+          'coastlines': dict(linewidth=0.5, color='k'),
+          'meridians': dict(meridians=np.arange(-80, -50, 2),
+                            linewidth=0.5,
+                            labels=[False, False, True, False],
+                            labelstyle='+/-'),
+          'parallels': dict(circles=np.arange(-50, -30, 2),
+                            linewidth=0.5,
+                            labels=[True, False, False, False],
+                            labelstyle='+/-')}
+config['parallels-quiet'] = config['parallels'].copy()
[…]

### 2018-12-19 code/scripts/neuquen_basin.py
Improve top sediments layer definition as median

@@ -58,6 +58,11 @@ sediments_top = gridder.interp_at(topography["lat"], topography["lon"],
                                   topography["topo"], sediments["lat"],
                                   sediments["lon"], algorithm="linear")
 sediments_top[np.isnan(sediments["thickness"])] = np.nan
+# Compute the median of the topography on basin points
+nans = np.isnan(sediments_top)
+sediments_top[~nans] = np.median(sediments_top[~nans])
+print("Top of sediments: {}m".format(sediments_top[~nans][0]))
+# Add top and bottom arrays to sediments dictionary
 sediments["top"] = sediments_top
 sediments["bottom"] = sediments_top - sediments["thickness"]
 
@@ -65,8 +70,8 @@ sediments["bottom"] = sediments_top - sediments["thickness"]
 # Create Tesseroids model of the sediments layer
 # ----------------------------------------------
 bottom, top = sediments["bottom"].copy(), sediments["top"].copy()
-top[np.isnan(top)] = 0
-bottom[np.isnan(bottom)] = 0
+top[nans] = 0
+bottom[nans] = 0
 basin = TesseroidModel(sediments['area'], top, bottom, sediments['shape'])
 
 
@@ -84,8 +89,11 @@ grid = {'lat': lat,
 
 # Common variables between computations
 # -------------------------------------
+# Define density values for the top and the bottom of the sediment layer
 density_top, density_bottom = -412, -275
-top, bottom = basin.top.max(), basin.bottom.min()
+# Define top and bottom variables as the maximum and minimum sediments'
+# height and depth, respectively
+top, bottom = basin.top[~nans].max(), basin.bottom[~nans].min()
 
 
 # Compute gravitational effect of the basin with homogeneous density

### 2018-12-19 code/scripts/neuquen_basin.py
Simplify axes grid in neuquen basin

@@ -8,7 +8,6 @@ from matplotlib.patches import Rectangle
 from matplotlib.colors import LightSource
 from mpl_toolkits.basemap import Basemap
 from mpl_toolkits.axes_grid1.inset_locator import inset_axes
-from matplotlib.gridspec import GridSpec
 from fatiando import gridder
 from tesseroid_density import tesseroid
 from tesseroid_model import TesseroidModel
@@ -211,13 +210,11 @@ config['parallels-right']['labels'] = [False, True, False, False]
 
 
 # Create outer grid
-fig = plt.figure(figsize=(6.66, 8))
-grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
-
+fig, grid = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
 
 # Topography
 # ----------
-ax = plt.subplot(grid[0, 0])
+ax = grid[0, 0]
 bm.ax = ax
 ax.set_title("(a)", y=1.08, loc='left')
 ax.set_title("Neuquen Basin", y=1.08, loc="center")
@@ -277,15 +274,14 @@ axins.add_patch(rectangle)
 
 # Basin thickness
 # ---------------
-ax = plt.subplot(grid[0, 1])
+ax = grid[0, 1]
 bm.ax = ax
 ax.set_title("(b)", y=1.08, loc='left')
 ax.set_title("Basin Thickness", y=1.08, loc="center")
 x, y = bm(sediments['lon'], sediments['lat'])
-im = bm.contourf(x.reshape(sediments['shape']),
-                 y.reshape(sediments['shape']),
-                 sediments['thickness'].reshape(sediments['shape']),
-                 100)
+im = bm.pcolormesh(x.reshape(sediments['shape']),
+                   y.reshape(sediments['shape']),
+                   sediments['thickness'].reshape(sediments['shape']))
 bm.contour(x.reshape(sediments['shape']),
            y.reshape(sediments['shape']),
            sediments['thickness'].reshape(sediments['shape']),
@@ -313,7 +309,7 @@ grid_specs = {"potential": grid[1, 0], "gz": grid[1, 1]}
 
 density = "exponential"
 for field in fields:
-    ax = plt.subplot(grid_specs[field])
+    ax = grid_specs[field]
     bm.ax = ax
     ax.set_title(fig_titles[field], y=1.08, loc='left')
     ax.set_title(titles[field], y=1.08, loc='center')
@@ -322,8 +318,8 @@ for field in fields:
[…]

### 2018-12-19 code/scripts/neuquen_basin.py
Rename grid to axes in neuquen basin

@@ -209,12 +209,12 @@ config['parallels-right'] = config['parallels'].copy()
 config['parallels-right']['labels'] = [False, True, False, False]
 
 
-# Create outer grid
-fig, grid = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+# Initiate figure and subplots
+fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
 
 # Topography
 # ----------
-ax = grid[0, 0]
+ax = axes[0, 0]
 bm.ax = ax
 ax.set_title("(a)", y=1.08, loc='left')
 ax.set_title("Neuquen Basin", y=1.08, loc="center")
@@ -274,7 +274,7 @@ axins.add_patch(rectangle)
 
 # Basin thickness
 # ---------------
-ax = grid[0, 1]
+ax = axes[0, 1]
 bm.ax = ax
 ax.set_title("(b)", y=1.08, loc='left')
 ax.set_title("Basin Thickness", y=1.08, loc="center")
@@ -305,11 +305,11 @@ cbar.update_ticks()
 labels = {"potential": "J/kg", "gz": "mGal"}
 titles = {"potential": r"$V$", "gz": r"$g_{z}$"}
 fig_titles = {"potential": "(c)", "gz": "(d)"}
-grid_specs = {"potential": grid[1, 0], "gz": grid[1, 1]}
+axes_dict = {"potential": axes[1, 0], "gz": axes[1, 1]}
 
 density = "exponential"
 for field in fields:
-    ax = grid_specs[field]
+    ax = axes_dict[field]
     bm.ax = ax
     ax.set_title(fig_titles[field], y=1.08, loc='left')
     ax.set_title(titles[field], y=1.08, loc='center')

### 2018-12-19 code/scripts/neuquen_basin.py
Change cbar locations in neuquen basin

@@ -3,7 +3,6 @@ import os
 import warnings
 import numpy as np
 import matplotlib.pyplot as plt
-from matplotlib import ticker
 from matplotlib.patches import Rectangle
 from matplotlib.colors import LightSource
 from mpl_toolkits.basemap import Basemap
@@ -235,11 +234,7 @@ bm.imshow(rgb)
 # Proxy image for colorbar
 im = bm.imshow(1e-3*topography['topo'].reshape(shape), cmap=cmap)
 im.remove()
-cbar = bm.colorbar(im, location='bottom', pad="3%")
-tick_locator = ticker.MaxNLocator(nbins=5)
-cbar.locator = tick_locator
-cbar.update_ticks()
-cbar.ax.tick_params(labelsize=9)
+cbar = bm.colorbar(im)
 
 # Basemap configuration
 bm.drawcountries(**config['countries'])
@@ -293,11 +288,7 @@ bm.drawparallels(**config['parallels-quiet'])
 bm.drawmeridians(**config['meridians'])
 
 # Colorbar
-cbar = bm.colorbar(im, location='bottom', pad="3%")
-cbar.ax.tick_params(labelsize=9)
-tick_locator = ticker.MaxNLocator(nbins=5)
-cbar.locator = tick_locator
-cbar.update_ticks()
+cbar = bm.colorbar(im)
 
 
 # Results (exponential)
@@ -330,17 +321,11 @@ for field in fields:
     bm.drawcoastlines(**config['coastlines'])
     bm.drawparallels(**config['parallels'])
     bm.drawmeridians(**config['meridians'])
-    # bm.drawparallels(**config['parallels-quiet'])
-    # bm.drawmeridians(**config['meridians-quiet'])
 
     # Colorbar
-    cbar = bm.colorbar(im, location='bottom', pad="3%")
-    cbar.ax.tick_params(labelsize=9)
-    tick_locator = ticker.MaxNLocator(nbins=5)
-    cbar.locator = tick_locator
-    cbar.update_ticks()
+    cbar = bm.colorbar(im)
 
-fig.tight_layout()
+fig.tight_layout(h_pad=2)
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen_basin.pdf")
 plt.savefig(figure_fname, dpi=300)

### 2018-12-19 code/scripts/neuquen_basin.py
Add units labels to cbars in neuquen basin

@@ -234,7 +234,7 @@ bm.imshow(rgb)
 # Proxy image for colorbar
 im = bm.imshow(1e-3*topography['topo'].reshape(shape), cmap=cmap)
 im.remove()
-cbar = bm.colorbar(im)
+cbar = bm.colorbar(im, label="km")
 
 # Basemap configuration
 bm.drawcountries(**config['countries'])
@@ -288,12 +288,12 @@ bm.drawparallels(**config['parallels-quiet'])
 bm.drawmeridians(**config['meridians'])
 
 # Colorbar
-cbar = bm.colorbar(im)
+cbar = bm.colorbar(im, label="km")
 
 
 # Results (exponential)
 # ---------------------
-labels = {"potential": "J/kg", "gz": "mGal"}
+units = {"potential": "J/kg", "gz": "mGal"}
 titles = {"potential": r"$V$", "gz": r"$g_{z}$"}
 fig_titles = {"potential": "(c)", "gz": "(d)"}
 axes_dict = {"potential": axes[1, 0], "gz": axes[1, 1]}
@@ -323,7 +323,7 @@ for field in fields:
     bm.drawmeridians(**config['meridians'])
 
     # Colorbar
-    cbar = bm.colorbar(im)
+    cbar = bm.colorbar(im, label=units[field])
 
 fig.tight_layout(h_pad=2)
 figure_fname = os.path.join(script_path,

### 2018-12-19 code/scripts/neuquen_basin.py
Improve basin thickness subplot

@@ -276,7 +276,7 @@ ax.set_title("Basin Thickness", y=1.08, loc="center")
 x, y = bm(sediments['lon'], sediments['lat'])
 im = bm.pcolormesh(x.reshape(sediments['shape']),
                    y.reshape(sediments['shape']),
-                   sediments['thickness'].reshape(sediments['shape']))
+                   1e-3*sediments['thickness'].reshape(sediments['shape']))
 bm.contour(x.reshape(sediments['shape']),
            y.reshape(sediments['shape']),
            sediments['thickness'].reshape(sediments['shape']),
@@ -284,7 +284,7 @@ bm.contour(x.reshape(sediments['shape']),
 bm.drawcountries(**config['countries'])
 bm.drawstates(**config['states'])
 bm.drawcoastlines(**config['coastlines'])
-bm.drawparallels(**config['parallels-quiet'])
+bm.drawparallels(**config['parallels'])
 bm.drawmeridians(**config['meridians'])
 
 # Colorbar

### 2018-12-19 code/scripts/neuquen_basin.py
Add neuquen basin densities plot

@@ -162,10 +162,7 @@ for field in fields:
              result=result, lon=grid["lon"], lat=grid["lat"],
              height=grid["height"], shape=grid["shape"])
 
-
-# ------------
-# Plot Results
-# ------------
+# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 # Configure LaTeX style for plots
 # -------------------------------
@@ -177,6 +174,29 @@ try:
     plt.rcParams['ytick.major.size'] = 2
 except Exception as e:
     warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
+
+
+# --------------
+# Plot densities
+# --------------
+heights = np.linspace(bottom, top, 101)
+fig, ax = plt.subplots(figsize=(3.33, 3))
+ax.plot(heights, linear_density(heights), label="Linear")
+ax.plot(heights, density_exponential(heights), label="Exponential")
+ax.set_xlabel("Height [m]")
+ax.set_ylabel(r"Density [kg/m$^3$]")
+ax.legend()
+ax.grid()
+fig.tight_layout()
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/neuquen_basin_densities.pdf")
+plt.savefig(figure_fname, dpi=300)
+plt.show()
+
+
+# ------------
+# Plot Results
+# ------------
 
 # Create basemap and configuration
 # --------------------------------

### 2018-12-19 code/scripts/neuquen_basin.py
Fix wrong area in neuquen basin maps

@@ -201,10 +201,10 @@ plt.show()
 # Create basemap and configuration
 # --------------------------------
 bm = Basemap(projection='merc',
-             llcrnrlon=topography["area"][2],
-             llcrnrlat=topography["area"][0],
-             urcrnrlon=topography["area"][3],
-             urcrnrlat=topography["area"][1],
+             llcrnrlon=grid["area"][2],
+             llcrnrlat=grid["area"][0],
+             urcrnrlon=grid["area"][3],
+             urcrnrlat=grid["area"][1],
              resolution='i')
 
 config = {'countries': dict(linewidth=0.5, color='k'),

### 2018-12-19 code/scripts/neuquen_basin.py
Ignore warnings in neuquen basin script

@@ -175,6 +175,9 @@ try:
 except Exception as e:
     warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
 
+# Disable warnings
+warnings.filterwarnings("ignore")
+
 
 # --------------
 # Plot densities

### 2018-12-19 code/scripts/neuquen_basin.py
Minor change in tight_layout

@@ -348,7 +348,7 @@ for field in fields:
     # Colorbar
     cbar = bm.colorbar(im, label=units[field])
 
-fig.tight_layout(h_pad=2)
+fig.tight_layout(pad=1.3, h_pad=2)
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen_basin.pdf")
 plt.savefig(figure_fname, dpi=300)

### 2018-12-19 code/scripts/neuquen_basin.py
Add differences plot in neuquen basin

@@ -353,3 +353,58 @@ figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen_basin.pdf")
 plt.savefig(figure_fname, dpi=300)
 plt.show()
+
+
+# ----------------
+# Plot differences
+# ----------------
+fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+
+densities = ["homogeneous", "linear"]
+units = {"potential": "J/kg", "gz": "mGal"}
+titles = {"potential": r"$V$", "gz": r"$g_{z}$"}
+fig_titles = "(a) (b) (c) (d)".split()
+
+for i, density in enumerate(densities):
+    for j, field in enumerate(fields):
+        ax = axes[i, j]
+        bm.ax = ax
+        ax.set_title(fig_titles[2*i + j], y=1.08, loc='left')
+        ax.set_title(titles[field], y=1.08, loc='center')
+
+        # Plot result
+        exponential = np.load(
+            os.path.join(result_dir, "exponential-{}.npz".format(field))
+        )
+        other = np.load(
+            os.path.join(result_dir, "{}-{}.npz".format(density, field))
+        )
+        difference = exponential["result"] - other["result"]
+        vmax = np.abs(difference).max()
+
+        shape = exponential["shape"]
+        im = bm.pcolormesh(exponential["lon"].reshape(shape),
+                           exponential["lat"].reshape(shape),
+                           difference.reshape(shape),
+                           vmin=-vmax, vmax=vmax,
+                           cmap="RdBu_r", latlon=True)
+        bm.contour(exponential["lon"].reshape(shape),
+                   exponential["lat"].reshape(shape),
+                   difference.reshape(shape),
+                   5, colors='k', linewidths=0.7, latlon=True)
+
+        # Configure basemap
+        bm.drawcountries(**config['countries'])
+        bm.drawstates(**config['states'])
+        bm.drawcoastlines(**config['coastlines'])
+        bm.drawparallels(**config['parallels'])
+        bm.drawmeridians(**config['meridians'])
+
+        # Colorbar
+        cbar = bm.colorbar(im, label=units[field])
+
+fig.tight_layout(pad=1.3, h_pad=2)
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/neuquen_basin_diffs.pdf")
+plt.savefig(figure_fname, dpi=300)
+plt.show()

### 2018-12-19 code/scripts/neuquen_basin.py
Add try blocks in neuquen basin

@@ -184,6 +184,14 @@ warnings.filterwarnings("ignore")
 # --------------
 heights = np.linspace(bottom, top, 101)
 fig, ax = plt.subplots(figsize=(3.33, 3))
+
+# Initialize figure and subplots
+try:
+    fig, ax = plt.subplots(figsize=(3.33, 3))
+except Exception:
+    plt.switch_backend('agg')
+    fig, ax = plt.subplots(figsize=(3.33, 3))
+
 ax.plot(heights, linear_density(heights), label="Linear")
 ax.plot(heights, density_exponential(heights), label="Exponential")
 ax.set_xlabel("Height [m]")
@@ -194,7 +202,10 @@ fig.tight_layout()
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen_basin_densities.pdf")
 plt.savefig(figure_fname, dpi=300)
-plt.show()
+try:
+    plt.show()
+except Exception:
+    pass
 
 
 # ------------
@@ -232,7 +243,12 @@ config['parallels-right']['labels'] = [False, True, False, False]
 
 
 # Initiate figure and subplots
-fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+try:
+    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+except Exception:
+    plt.switch_backend('agg')
+    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+
 
 # Topography
 # ----------
@@ -352,7 +368,10 @@ fig.tight_layout(pad=1.3, h_pad=2)
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen_basin.pdf")
 plt.savefig(figure_fname, dpi=300)
-plt.show()
+try:
+    plt.show()
+except Exception:
+    pass
 
 
 # ----------------
@@ -407,4 +426,7 @@ fig.tight_layout(pad=1.3, h_pad=2)
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen_basin_diffs.pdf")
 plt.savefig(figure_fname, dpi=300)
-plt.show()
+try:
+    plt.show()
+except Exception:
+    pass

### 2018-12-19 code/scripts/neuquen_basin.py
Remove unwanted line in neuquen basin

@@ -183,7 +183,6 @@ warnings.filterwarnings("ignore")
 # Plot densities
 # --------------
 heights = np.linspace(bottom, top, 101)
-fig, ax = plt.subplots(figsize=(3.33, 3))
 
 # Initialize figure and subplots
 try:

### 2018-12-19 code/scripts/neuquen_basin.py
Print information of top and bottom variables

@@ -92,6 +92,7 @@ density_top, density_bottom = -412, -275
 # Define top and bottom variables as the maximum and minimum sediments'
 # height and depth, respectively
 top, bottom = basin.top[~nans].max(), basin.bottom[~nans].min()
+print("Top: {} Bottom: {}".format(top, bottom))
 
 
 # Compute gravitational effect of the basin with homogeneous density

### 2018-12-19 code/scripts/neuquen_basin.py
Change figures' names in neuquen basin

@@ -200,7 +200,7 @@ ax.legend()
 ax.grid()
 fig.tight_layout()
 figure_fname = os.path.join(script_path,
-                            "../../manuscript/figures/neuquen_basin_densities.pdf")
+                            "../../manuscript/figures/neuquen-basin-densities.pdf")
 plt.savefig(figure_fname, dpi=300)
 try:
     plt.show()
@@ -366,7 +366,7 @@ for field in fields:
 
 fig.tight_layout(pad=1.3, h_pad=2)
 figure_fname = os.path.join(script_path,
-                            "../../manuscript/figures/neuquen_basin.pdf")
+                            "../../manuscript/figures/neuquen-basin.pdf")
 plt.savefig(figure_fname, dpi=300)
 try:
     plt.show()
@@ -424,7 +424,7 @@ for i, density in enumerate(densities):
 
 fig.tight_layout(pad=1.3, h_pad=2)
 figure_fname = os.path.join(script_path,
-                            "../../manuscript/figures/neuquen_basin_diffs.pdf")
+                            "../../manuscript/figures/neuquen-basin-diffs.pdf")
 plt.savefig(figure_fname, dpi=300)
 try:
     plt.show()

### 2018-12-19 code/scripts/neuquen_basin.py
Fix wrong printing of information

@@ -59,7 +59,6 @@ sediments_top[np.isnan(sediments["thickness"])] = np.nan
 # Compute the median of the topography on basin points
 nans = np.isnan(sediments_top)
 sediments_top[~nans] = np.median(sediments_top[~nans])
-print("Top of sediments: {}m".format(sediments_top[~nans][0]))
 # Add top and bottom arrays to sediments dictionary
 sediments["top"] = sediments_top
 sediments["bottom"] = sediments_top - sediments["thickness"]
@@ -92,7 +91,7 @@ density_top, density_bottom = -412, -275
 # Define top and bottom variables as the maximum and minimum sediments'
 # height and depth, respectively
 top, bottom = basin.top[~nans].max(), basin.bottom[~nans].min()
-print("Top: {} Bottom: {}".format(top, bottom))
+print("Top of sediments: {} \nBottom of sediments: {}".format(top, bottom))
 
 
 # Compute gravitational effect of the basin with homogeneous density

### 2018-12-19 code/scripts/neuquen_basin.py
Improve image size and tight layout in neuquen

@@ -243,10 +243,10 @@ config['parallels-right']['labels'] = [False, True, False, False]
 
 # Initiate figure and subplots
 try:
-    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 7))
 except Exception:
     plt.switch_backend('agg')
-    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 7))
 
 
 # Topography
@@ -363,7 +363,7 @@ for field in fields:
     # Colorbar
     cbar = bm.colorbar(im, label=units[field])
 
-fig.tight_layout(pad=1.3, h_pad=2)
+fig.tight_layout()
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen-basin.pdf")
 plt.savefig(figure_fname, dpi=300)
@@ -376,7 +376,12 @@ except Exception:
 # ----------------
 # Plot differences
 # ----------------
-fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 8))
+# Initiate figure and subplots
+try:
+    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 7))
+except Exception:
+    plt.switch_backend('agg')
+    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 7))
 
 densities = ["homogeneous", "linear"]
 units = {"potential": "J/kg", "gz": "mGal"}
@@ -421,7 +426,7 @@ for i, density in enumerate(densities):
         # Colorbar
         cbar = bm.colorbar(im, label=units[field])
 
-fig.tight_layout(pad=1.3, h_pad=2)
+fig.tight_layout()
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen-basin-diffs.pdf")
 plt.savefig(figure_fname, dpi=300)

### 2018-12-19 manuscript/manuscript.tex
Update neuquen basin text in manuscript

@@
 
+\begin{figure}
+\centering
+\iftwocol{
+\includegraphics[width=\linewidth]{figures/neuquen-basin-densities.pdf}
+}{
+\includegraphics[width=0.5\linewidth]{figures/neuquen-basin-densities.pdf}
+}
+\caption{
+    Linear and exponential densities used to compute the gravitational fields generated
+    by a tesseroid model of the Neuquen sedimentary basin.
+    The height is defined above the mean Earth radius, and its axis is spanned between
+    the deepest and the highest point of the basin.
+}
+\label{fig:neuquen-basin-densities}
+\end{figure}
+
 \begin{figure*}
@@
     (b) thickness of the sedimentary basin \citep[in meters;][]{Heine2007},
-    (c) the exponential density function used to model the basin
-        (depth in km and density in kg/m$^3$),
-    (d)-(l) computed gravitational fields: potential $V$ (in J/kg), gradient
-    components $g_x$, $g_y$ and $g_z$ (in mGal), and Marussi tensor components
-    (in Eötvös), calculated at 50km of height over the ellipsoid.
+    (c) resulting gravitational potential $V$,
+    (d) resulting vertical component of the gradient ($g_z$),
+    calculated at 10km of height over the mean Earth radius.
 }
 \label{fig:neuquen-basin}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=\linewidth]{figures/neuquen-basin-diffs.pdf}
+\caption{
+    Differences between the gravitational fields generated by the tesseroid model of the
+    Neuquen basin with an exponential density contrast and with a homogenenous and
+    linear density variation.
+    (a)-(b) Differences on $V$ and $g_z$ between the exponential density model and the
+    homogenenous density one,
+    (c)-(f) differences on $V$ and $g_z$ between the exponential density model and the
+    linear density one,
+    calculated at 10km of height over the mean Earth radius.
+}
+\label{fig:neuquen-basin-diffs}
 \end{figure*}
@@
 $0.05^\circ \times 0.05^\circ$ tesseroid on each node of the grid.
-The top of each tesseroid was fixed at 0 m depth and the bottom at corresponding
-thickness of the basin.
+The top of each tesseroid was fixed at the median of the topography of the basin
+(845~m above mean Earth radius) and the bottom at corresponding thickness of the basin.
 
@@
 We have chosen an exponential density variation (Eq.~\ref{eq:density-exp}) that assumes
-the minimum value on the top surface and the maximum at 5858m depth (the thickest part
-of the basin), with a value of $b$ equal to 2km.
-This density function can be seen on Fig.~\ref{fig:neuquen-basin}c.
-
-Finally, we computed the gravitational potential $V$, gradient components $g_x$,
-$g_y$ and $g_z$, and the Marussi tensor components $g_{xx}$, $g_{xy}$,
-$g_{xz}$, $g_{yy}$, and $g_{zz}$ on a computation grid of $159\times163$ nodes
-($0.05^\circ$ spacing on both longitude and latitude) at a 50km height over the
-reference ellipsoid.
-The resulting fields can be seen in Fig.~\ref{fig:neuquen-basin}d-l.
-We have excluded the other components of the Marussi tensor from the
-Fig.~\ref{fig:neuquen-basin}, although they can be found in the online repository.
+the minimum value on the top surface and the maximum at 5014m depth (the thickest part
+of the basin), with a value of $b$ equal to 3.
+This density variation is in agreement with the ones used by
+This density function can be seen on Fig.~\ref{fig:neuquen-basin-densities}.
+
+Finally, we computed the gravitational potential $V$ and the vertical component of the
+gradient ($g_z$) on a computation grid of $159\times163$ nodes ($0.05^\circ$ spacing on
+both longitude and latitude) at a 10km height over the mean Earth radius.
+The resulting fields can be seen in Fig.~\ref{fig:neuquen-basin}c-d.
+Moreover, we computed the differences between these fields and the ones generated by the
+same model but now with a constant density as the mean value between -412kg/m$^3$ and
+-275kg/m$^3$, and a linear density that assumes these two values on the highest and the
+deepest point of the basin, respectively (Fig.~\ref{fig:neuquen-basin-densities}).
+Fig.~\ref{fig:neuquen-basin-diffs}a-b and Fig.~\ref{fig:neuquen-basin-diffs}c-d show the
+differences with the constant and the linear density, respectively
 

### 2018-12-19 code/scripts/computation_time.py
Improve computation time figure

@@ -1,6 +1,7 @@
 from __future__ import division, print_function
 import os
 import time
+import warnings
 import numpy as np
 import pandas as pd
 from fatiando.mesher import Tesseroid
@@ -143,14 +144,59 @@ if compute:
 
 # Plot relative computation times
 # -------------------------------
+# Load computation time differences
 df = pd.read_csv(csv_fname, index_col=0)
-fig, axes = plt.subplots(nrows=2, ncols=1)
+
+# Configure LaTeX style for plots
+try:
+    plt.rcParams['text.usetex'] = True
+    plt.rcParams['font.family'] = 'serif'
+    plt.rcParams['font.serif'] = 'Computer Modern Roman'
+    plt.rcParams['xtick.major.size'] = 2
+    plt.rcParams['ytick.major.size'] = 2
+except Exception as e:
+    warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
+
+# Initialize figure and subplots
+try:
+    fig, axes = plt.subplots(nrows=2, ncols=1, figsize=(3.33, 4), sharex=True)
+except Exception:
+    plt.switch_backend('agg')
+    fig, axes = plt.subplots(nrows=2, ncols=1, figsize=(3.33, 4), sharex=True)
+
+# Plot results
 ax = axes[0]
-ax.plot(df["linear_potential"] / df["constant_potential"], '-o')
-ax.plot(df["linear_gz"] / df["constant_gz"], '-o')
+ax.plot(df["linear_potential"] / df["constant_potential"], '-o', label=r"$V$")
+ax.plot(df["linear_gz"] / df["constant_gz"], '-o', label=r"$g_z$")
 ax = axes[1]
-ax.plot(df["exp_gz"] / df["constant_gz"], '-o')
-ax.plot(df["exp_potential"] / df["constant_potential"], '-o')
-for ax in axes:
+ax.plot(df["exp_gz"] / df["constant_gz"], '-o', label=r"$V$")
+ax.plot(df["exp_potential"] / df["constant_potential"], '-o', label=r"$g_z$")
+
+# Configure axes
+densities = 'Linear Exponential'.split()
+ax.set_xlabel("Computation height [m]")
+fig.text(0, 0.5, "Computation time ratio", va='center', rotation='vertical')
+fig.subplots_adjust(hspace=0.1)
+axes[0].legend()
+for density, ax in zip(densities, axes):
+    ax.grid()
     ax.set_xscale("log")
-plt.show()
+
+    # Add field annotation on each axe
[…]

### 2018-12-19 code/scripts/computation_time.py
Reduced value of b on computation time script

@@ -105,7 +105,7 @@ if compute:
 
     # Compute gravitational fields for exp density tesseroid
     # ------------------------------------------------------
-    b_factor = 10
+    b_factor = 3
     density_top, density_bottom = 2670, 3300
     thickness = top - bottom
 

### 2018-12-19 code/scripts/computation_time.py
Save fig for computation time script

@@ -196,6 +196,10 @@ for density, ax in zip(densities, axes):
                   'boxstyle': 'square, pad=0.4'},
             transform=ax.transAxes)
 
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/computation-times.pdf")
+plt.savefig(figure_fname, dpi=300)
+
 try:
     plt.show()
 except Exception:

### 2018-12-19 code/scripts/computation_time.py
Fix wrong time computation

@@ -133,7 +133,7 @@ if compute:
                 start_time = time.time()
                 getattr(tesseroid, field)(lon, lat, height, model)
                 end_time = time.time()
-                times[run] = start_time - end_time
+                times[run] = end_time - start_time
             times = np.mean(times)
             df.at[height, "{}_{}".format(density, field)] = times
 

### 2018-12-19 manuscript/manuscript.tex
Add missing citations in manuscript
Add computation time section in manuscript

@@
 
+\section{Computation time comparison}
+
+Because the density-based algorithm produces discretizations along the radial direction,
+besides the latitudinal and longitudinal ones generated by the two dimensional adaptive
+one, the computation time of the gravitational fields of variable density tesseroids
+will be increased in comparison to the homogeneous density case.
+In order to quantize this increase we perform a computation time comparison between the
+homogeneous density case with the linear and the exponential ones.
+
+We define a $20^\circ \times 20^\circ$ tesseroid of 1km of thickness and a single
+computation point located at different heights (0m, 1km, 10km, 100km and 1000km) above
+the top surface of the tesseroid.
+Then we compute the gravitational potential $V$ and the $g_z$ that the tesseroid
+generates on the computation point in case of an homogeneous density, a linear density
+and an exponential density with $b = 3$ (Eq.~\ref{eq:density-exp}).
+We perform 10000 computations for each field, density and computation height while
+tracking the computation time for each one of them.
+The final computation time is obtained as the mean value of these runs.
+
+Fig.~\ref{fig:computation-times} shows the computation time ratio between the
+computations times of the linear and exponential density with the one for an homogeneous
+density, defined as:
+
+\begin{equation}
+    \text{Computation time ratio} = \frac{\Delta t}{\Delta t_\text{homog}},
+\end{equation}
+
+\noindent where $\Delta t_\text{homog}$ is the computation time for the homogeneous
+density and $\Delta t$ is the computation time for the linear or the exponential
+density.
+
+\begin{figure}
+\centering
+\iftwocol{
+\includegraphics[width=\linewidth]{figures/computation-times.pdf}
+}{
+\includegraphics[width=0.5\linewidth]{figures/computation-times.pdf}
+}
+\caption{
+    Computation times ratio between linear and exponential density tesseroids with an
+    homogeneous one.
+}
+\label{fig:computation-times}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Application to the Neuqu\'en Basin}
@@
 of the basin), with a value of $b$ equal to 3.
-This density variation is in agreement with the ones used by
+This density variation is in the order of magnitude of the ones used by
+\citet{Cowie1990} and \citet{Cordell1973}.
 This density function can be seen on Fig.~\ref{fig:neuquen-basin-densities}.

### 2018-12-19 code/scripts/computation_time.py
Fix wrong computation time calculation

@@ -73,7 +73,7 @@ if compute:
                 start_time = time.time()
                 getattr(tesseroid, field)(lon, lat, height, model)
                 end_time = time.time()
-                times[run] = start_time - end_time
+                times[run] = end_time - start_time
             times = np.mean(times)
             df.at[height, "{}_{}".format(density, field)] = times
 
@@ -99,7 +99,7 @@ if compute:
                 start_time = time.time()
                 getattr(tesseroid, field)(lon, lat, height, model)
                 end_time = time.time()
-                times[run] = start_time - end_time
+                times[run] = end_time - start_time
             times = np.mean(times)
             df.at[height, "{}_{}".format(density, field)] = times
 

### 2018-12-19 code/scripts/computation_time.py
Change computation heights on computation times

@@ -25,7 +25,7 @@ model = [Tesseroid(w, e, s, n, top, bottom)]
 
 # Define computation points
 # -------------------------
-heights = np.array([0., 1e3, 1e4, 1e5, 1e6])
+heights = np.array([1e0, 1e1, 1e2, 1e3, 1e4, 1e5, 1e6])
 computation_points = [[np.array([0.]), np.array([0.]), np.array([height])]
                       for height in heights]
 

### 2018-12-19 code/scripts/computation_time.py
Reduced nruns

@@ -12,7 +12,7 @@ from tesseroid_density import tesseroid
 
 # Configure comparison
 # --------------------
-nruns = 10000
+nruns = 1000
 fields = "potential gz".split()
 
 

### 2018-12-19 code/scripts/computation_time.py
Improve computation times figure

@@ -177,16 +177,17 @@ densities = 'Linear Exponential'.split()
 ax.set_xlabel("Computation height [m]")
 fig.text(0, 0.5, "Computation time ratio", va='center', rotation='vertical')
 fig.subplots_adjust(hspace=0.1)
-axes[0].legend()
+axes[0].legend(loc=6)
 for density, ax in zip(densities, axes):
     ax.grid()
     ax.set_xscale("log")
+    ax.set_xticks(np.logspace(0, 6, 7))
 
     # Add field annotation on each axe
     if density == "Linear":
-        yloc = 0.8
+        yloc = 0.7
     elif density == "Exponential":
-        yloc = 0.89
+        yloc = 0.2
     ax.text(0.5, yloc, density, fontsize=11,
             horizontalalignment='center',
             verticalalignment='center',

### 2018-12-19 manuscript/manuscript.tex
Improve computation times text in manuscript

@@
 We define a $20^\circ \times 20^\circ$ tesseroid of 1km of thickness and a single
-computation point located at different heights (0m, 1km, 10km, 100km and 1000km) above
-the top surface of the tesseroid.
+computation point located at different heights (1m, 10m, 100m, 1km, 10km, 100km and
+1000km) above the top surface of the tesseroid.
 Then we compute the gravitational potential $V$ and the $g_z$ that the tesseroid
@@
 and an exponential density with $b = 3$ (Eq.~\ref{eq:density-exp}).
-We perform 10000 computations for each field, density and computation height while
+We perform 1000 computations for each field, density and computation height while
 tracking the computation time for each one of them.

### 2018-12-21 code/scripts/density-based-discretization.py
Create density-based-discretization example script

@@ -1,0 +1,169 @@
+from __future__ import division
+import os
+import warnings
+import numpy as np
+import matplotlib.pyplot as plt
+
+# This is our custom tesseroid code
+from tesseroid_density import tesseroid
+
+
+# Define functions
+def density_function(height):
+    return np.exp(b_factor * height / thickness)
+
+
+def normalized_density(height):
+    return (density_function(height) - rho_min)/(rho_max - rho_min)
+
+
+def line(height, top, bottom):
+    a = (normalized_density(top) - normalized_density(bottom))/(top - bottom)
+    b = normalized_density(bottom)
+    return a*(height - bottom) + b
+
+
+# Define tesseroid boundaries
+w, e, s, n, top, bottom = -10, 10, -10, 10, 0, -1
+thickness = top - bottom
+bounds = [w, e, s, n, top, bottom]
+
+# Set b factor for exp density variation
+b_factor = 6
+
+# Define heights array and boundary density values
+heights = np.linspace(bottom, top, 101)
+density = density_function(heights)
+rho_min, rho_max = density.min(), density.max()
+
+
+# Calculate discretizations
+# -------------------------
+delta = 1
+delta_step = 0.001
+max_divisions = 4
+subsets = [[np.array(bounds)]]
+deltas = [1]
+
+while True:
+    subset = tesseroid._density_based_discretization(bounds,
+                                                     density_function,
+                                                     delta)
+    divisions = len(subset)
+    if divisions == len(subsets[-1]) + 1:
+        subsets.append(subset)
+        deltas.append(delta)
+    elif divisions > len(subsets[-1]) + 1:
+        print("More discretizations than 1 in delta {}".format(delta))
+    if divisions >= max_divisions:
+        break
+    else:
+        delta -= delta_step
+
+
+# Plot discretization steps
+# -------------------------
+
+# Configure LaTeX style for plots
+try:
+    plt.rcParams['text.usetex'] = True
+    plt.rcParams['font.family'] = 'serif'
+    plt.rcParams['font.serif'] = 'Computer Modern Roman'
+    plt.rcParams['xtick.major.size'] = 2
+    plt.rcParams['ytick.major.size'] = 2
+except Exception as e:
+    warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
+
+# Initialize figure and subplots
+fig, axes = plt.subplots(1, len(subsets),
+                         figsize=(6.66, 2),
[…]

### 2018-12-21 manuscript/manuscript.tex
Fix unwanted lowercase in density-based fig caption

@@
     density function.
-    (a) the normalised density function $\rho_n(r')$ (blue), current boundaries of the
+    (a) The normalised density function $\rho_n(r')$ (blue), current boundaries of the
     tesseroid (orange dots), and the linear density function $\rho_l(r')$ (orange line).
@@
     inequality~\ref{eq:delta-density} is not satisfied).
-    (b) second iteration of the algorithm with a new linear density function and maximum
+    (b) Second iteration of the algorithm with a new linear density function and maximum
     density difference. The tesseroid would be divided at the depth indicated by the
     dashed red line.
-    (c) third iteration of the algorithm.
-    (d) final output of the density-based discretization, assuming that all four new
+    (c) Third iteration of the algorithm.
+    (d) Final output of the density-based discretization, assuming that all four new
     tesseroids satisfy inequality~\ref{eq:delta-density}.

### 2018-12-21 manuscript/manuscript.tex
Add missing dot

@@
 Fig.~\ref{fig:neuquen-basin-diffs}a-b and Fig.~\ref{fig:neuquen-basin-diffs}c-d show the
-differences with the constant and the linear density, respectively
+differences with the constant and the linear density, respectively.
 

### 2018-12-21 manuscript/manuscript.tex
Minor change in manuscript

@@
 -275kg/m$^3$, and a linear density that assumes these two values on the highest and the
-deepest point of the basin, respectively (Fig.~\ref{fig:neuquen-basin-densities}).
+deepest point of the basin (Fig.~\ref{fig:neuquen-basin-densities}), respectively.
 Fig.~\ref{fig:neuquen-basin-diffs}a-b and Fig.~\ref{fig:neuquen-basin-diffs}c-d show the

### 2018-12-21 code/scripts/neuquen_basin.py
Remove white lines in pcolormeshes

@@ -314,7 +314,8 @@ ax.set_title("Basin Thickness", y=1.08, loc="center")
 x, y = bm(sediments['lon'], sediments['lat'])
 im = bm.pcolormesh(x.reshape(sediments['shape']),
                    y.reshape(sediments['shape']),
-                   1e-3*sediments['thickness'].reshape(sediments['shape']))
+                   1e-3*sediments['thickness'].reshape(sediments['shape']),
+                   rasterized=True)
 bm.contour(x.reshape(sediments['shape']),
            y.reshape(sediments['shape']),
            sediments['thickness'].reshape(sediments['shape']),
@@ -348,7 +349,7 @@ for field in fields:
 
     shape = result["shape"]
     im = bm.pcolormesh(result["lon"].reshape(shape), result["lat"].reshape(shape),
-                       result["result"].reshape(shape), latlon=True)
+                       result["result"].reshape(shape), rasterized=True, latlon=True)
     bm.contour(result["lon"].reshape(shape), result["lat"].reshape(shape),
                result["result"].reshape(shape), 5, colors='k', linewidths=0.7,
                latlon=True)
@@ -410,7 +411,7 @@ for i, density in enumerate(densities):
                            exponential["lat"].reshape(shape),
                            difference.reshape(shape),
                            vmin=-vmax, vmax=vmax,
-                           cmap="RdBu_r", latlon=True)
+                           cmap="RdBu_r", rasterized=True, latlon=True)
         bm.contour(exponential["lon"].reshape(shape),
                    exponential["lat"].reshape(shape),
                    difference.reshape(shape),

### 2018-12-21 manuscript/manuscript.tex
Add first approach to results and discussion

@@
 
+\section{Results and Discussion}
+
+The $D$ and $\delta$ ratios determinations proved that for a linear density case only
+the two dimensional adaptive discretization algorithm and a second-order GLQ
+approximation is needed to achieve a 0.1\% accuracy of the gravitational potential and
+the vertical component of the gradient ($g_z$) by using a $D=$ 1 and 2.5 for each field,
+respectively.
+Although for non-linear functions the adaptive discretization is not sufficient and the
+density-based discretization algorithm must be applied.
+The $D$-$\delta$ space exploration showed that the more efficient computation takes
+place when the value of $D$ equals to the ones determined in the linear density
+scenario, while the value of $\delta$ must be determined for each specific density
+variation.
+In case of an exponential density, a $\delta$ equal to 0.1 is sufficient to achieve the
+0.1\% accuracy threshold, while a $\delta = 0.01$ is needed for a sinusoidal density
+function.
+
+This result could be extrapolated: a default value of $\delta = 0.01$ would guarantee an
+accuracy of 0.1\% in case of an arbitrary continuous density function.
+Nevertheless, we encourage to perform similar accuracy tests in order to proof that this
+extrapolation works for the desired density function, specially for more complex
+variations than a polynomial or exponential one.
+
+\citet{Lin2018} proposed a combined method to efficiently compute gravitational fields
+generated by tesseroids: it uses a GLQ with a two dimensional adaptive discretization
+algorithm for near computation points and a Taylor series expansion method for far
+computation points.
+Because our goal is to build a method that could be applied to any continuous density
+function, the Taylor series wouldn't be well suited for this task and therefore we
+focused only on a pure GLQ approximation method.
+Nevertheless, our method could be easily combined with a Taylor series expansion for
+specific density function, i.e.~a linear variation.
+
+Moreover, \citet{Lin2018} show that the combined method can be applied to a linear
+density variation by splitting the gravitational field integral in two: one for each
+term of the linear function.
+Our results show that in case of a linear density variation a single GLQ computation is
+sufficient to accurately compute the fields, thus reducing the computation time by half.
+
+\citet{Fukushima2018} propose a generalization of his method to polynomial density
+functions.
+On the other hand, our method is not bounded to a specific density function.
+Nevertheless, the density-based discretization algorithm could be used before applying
+the numerical integration developed by \citet{Fukushima2018} assigning a low polynomial
+order density to each discretizing tesseroid.
+Although this method would involve a computational error due to the approximation of the
+density function as a polynomial one, it's feasibly if we account for this error and
+a method to limit it.
+
+We have also carried out $D$ and $\delta$ determinations for the vertical component of
+the Marussi tensor ($g_{zz}$) for linear and exponential densities.
+We didn't include these results because our methodology wouldn't achieve the 0.1\%
+accuracy threshold for the general case of such density functions.
+Moreover, the values of $D$ and $\delta$ needed for the cases at which the accuracy was
+below 0.1\% implied a high number of discretizations, thus greatly increasing the
+computation time.
+In summary, we found the computation of the Marussi tensor components in case of
+variable density tesseroids a subject for future research.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conclusions}

### 2018-12-26 manuscript/manuscript.tex
Update abstract

@@
 We have obtained analytical solutions for a spherical shell with radially variable
-density and compared them to the results of the numerical model for linear and
-exponential density functions.
+density and compared them to the results of the numerical model for linear,
+exponential and sinusoidal density functions.
 These comparisons allowed us to obtain optimum values for the distance-size and
 delta ratios that yield an accuracy of 0.1\% of the analytical solutions.
-The resulting optimal values of distance-size ratio for the gravitational potential, its
-gradient, and Marussi tensor are 1, 2 and 8, respectively.
-A delta ratio of 0.2 is needed for the computation of the gravitational potential and
-its gradient components, while a value of 0.01 must be used for the Marussi tensor
-components.
+The resulting optimal values of distance-size ratio for the gravitational potential and
+its gradient are 1 and 2.5, respectively.
+The density-based discretization algorithm produces no discretizations in the linear
+density case, but a delta ratio of 0.1 and 0.01 is needed for the exponential and the
+sinusoidal density function, respectively.
 Lastly, we apply this new methodology to model the Neuqu\'en Basin, a foreland basin in

### 2018-12-26 manuscript/manuscript.tex
Minor text modification

@@
 the mean orbital height of the gravimetric satellite missions (Satellite).
-We define these grids in the spirit of spanning a broad scenario of gravitational field
-computations and ensure an acceptable accuracy on each one of them.
+These grids are defined in the spirit of spanning a broad scenario of gravitational
+field computations and ensure an acceptable accuracy on each one of them.
 The comparison will be carried out by computing the maximum absolute difference between

### 2018-12-26 manuscript/manuscript.tex
Update conclusions

@@
 generated by a tesseroid with a density given by a continuous function of depth.
-It numerically solves the integrals that define the gravitational potential,
-its gradient, and the Marussi tensor components through the Gauss-Legendre
-Quadrature (GLQ).
-The accuracy of the numerical integration is automatically controlled by an adaptive
-discretization algorithm and a new density-based discretization algorithm.
+It numerically solves the integrals that define the gravitational potential and
+its gradient through the Gauss-Legendre Quadrature (GLQ).
+The accuracy of the numerical integration is automatically controlled by a two
+dimensional adaptive discretization algorithm and a new density-based discretization
+algorithm.
 The former divides the tesseroid in half if the ratio of the distance to the computation
-point and the size of the tesseroid is lower than a predefined distance-size ratio $D$.
+point and the latitudinal or the longitudinal size of the tesseroid is lower than
+a predefined distance-size ratio $D$.
 This algorithm minimises the integration error when the computation point is close to
@@
 0.1\% of an analytical solution.
-The density functions used to establish the optimum values were a linear and an
-exponential function.
+The density functions used to establish the optimum values were a linear, an exponential
+and a sinusoidal function.
 The linear function represents the smoothest variation of density and does not require
@@
 in depth to test the accuracy of the density-based discretization.
+And the error for sinusoidal functions with different wavelengths.
 Through this analysis, we have obtained optimal values for the distance-size ratio of
-1, 2 and 8 for the potential, its gradient, and the Marussi tensor components,
-respectively.
-These values are in agreement with previous results for a homogeneous density tesseroid
-\citep{Uieda2016}.
-The optimum $\delta$ values are 0.2 for the potential and its gradient components, while
-a $\delta = 0.01$ must be used for the Marussi tensor components in order to ensure a
-0.1\% precision.
-A wide range of exponential functions were used to infer these results and they are
-conservative estimates, chosen to prioritize accuracy over speed of computations.
-Therefore, they can be safely generalized to other common depth-density relations, like
-quadratic functions, without loss of accuracy.
+1 and 2.5 for the potential an its gradient, respectively.
+While values of $\delta$ equal to 0.1 and 0.01 are needed in order to guarantee a 0.1\%
+accuracy for an exponential and a sinusoidal density, respectively.
+These results could be extrapolated: a $\delta$ value equal to 0.01 is needed in order
+to achieve the accuracy threshold in case of an arbitrary continuous density function.
+Although we encourage to perform similar accuracy tests for functions other than the
+ones we studied, this results could be safely generalized to other common depth-density
+relations among geophysical applications, like quadratic functions, without loss of
+accuracy.
 

### 2018-12-26 manuscript/manuscript.tex
Improve enumerated figure captions

@@
     density function.
-    (a) The normalised density function $\rho_n(r')$ (blue), current boundaries of the
+    (a)~The normalised density function $\rho_n(r')$ (blue), current boundaries of the
     tesseroid (orange dots), and the linear density function $\rho_l(r')$ (orange line).
@@
     inequality~\ref{eq:delta-density} is not satisfied).
-    (b) Second iteration of the algorithm with a new linear density function and maximum
+    (b)~Second iteration of the algorithm with a new linear density function and maximum
     density difference. The tesseroid would be divided at the depth indicated by the
     dashed red line.
-    (c) Third iteration of the algorithm.
-    (d) Final output of the density-based discretization, assuming that all four new
+    (c)~Third iteration of the algorithm.
+    (d)~Final output of the density-based discretization, assuming that all four new
     tesseroids satisfy inequality~\ref{eq:delta-density}.
@@
     using tesseroids with an exponential density function of depth.
-    (a) Topography of the Neuqu\'en Basin (in km) and its location in South America,
-    (b) thickness of the sedimentary basin \citep[in meters;][]{Heine2007},
-    (c) resulting gravitational potential $V$,
-    (d) resulting vertical component of the gradient ($g_z$),
+    (a)~Topography of the Neuqu\'en Basin (in km) and its location in South America,
+    (b)~thickness of the sedimentary basin \citep[in meters;][]{Heine2007},
+    (c)~resulting gravitational potential $V$,
+    (d)~resulting vertical component of the gradient ($g_z$),
     calculated at 10km of height over the mean Earth radius.
@@
     linear density variation.
-    (a)-(b) Differences on $V$ and $g_z$ between the exponential density model and the
-    homogenenous density one,
-    (c)-(f) differences on $V$ and $g_z$ between the exponential density model and the
-    linear density one,
-    calculated at 10km of height over the mean Earth radius.
+    \mbox{(a)-(b)}~Differences on $V$ and $g_z$ between the exponential density model
+    and the homogenenous density one,
+    \mbox{(c)-(f)}~differences on $V$ and $g_z$ between the exponential density model
+    and the linear density one, calculated at 10km of height over the mean Earth radius.
 }

### 2018-12-26 manuscript/manuscript.tex
Add Anderson citation
By request of second reviewer

@@
 are defined in geocentric spherical coordinates.
-A common approach is to discretize the Earth into tesseroids (spherical prisms),
-which are defined by pairs of latitude, longitude, and
-radial boundaries (see Fig.~\ref{fig:tesseroid}).
+A common approach is to discretize the Earth into tesseroids \citep{Anderson1976},
+i.e.~spherical prisms, which are defined by pairs of latitude, longitude, and radial
+boundaries (see Fig.~\ref{fig:tesseroid}).
 The gravitational fields generated by an arbitrary

### 2018-12-26 code/scripts/computation_time.py
Fix DISPLAY variable related error in scripts

@@ -147,6 +147,12 @@ if compute:
 # Load computation time differences
 df = pd.read_csv(csv_fname, index_col=0)
 
+# Check for DISPLAY variable for matplotlib
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
 # Configure LaTeX style for plots
 try:
     plt.rcParams['text.usetex'] = True
@@ -158,11 +164,7 @@ except Exception as e:
     warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
 
 # Initialize figure and subplots
-try:
-    fig, axes = plt.subplots(nrows=2, ncols=1, figsize=(3.33, 4), sharex=True)
-except Exception:
-    plt.switch_backend('agg')
-    fig, axes = plt.subplots(nrows=2, ncols=1, figsize=(3.33, 4), sharex=True)
+fig, axes = plt.subplots(nrows=2, ncols=1, figsize=(3.33, 4), sharex=True)
 
 # Plot results
 ax = axes[0]
@@ -200,8 +202,4 @@ for density, ax in zip(densities, axes):
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/computation-times.pdf")
 plt.savefig(figure_fname, dpi=300)
-
-try:
-    plt.show()
-except Exception:
-    pass
+plt.show()

### 2018-12-26 code/scripts/density-based-discretization.py

@@ -64,6 +64,12 @@ while True:
 # Plot discretization steps
 # -------------------------
 
+# Check for DISPLAY variable for matplotlib
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
 # Configure LaTeX style for plots
 try:
     plt.rcParams['text.usetex'] = True

### 2018-12-26 code/scripts/exponential_density.py

@@ -126,6 +126,14 @@ for field in fields:
                          delta_values=delta_values, differences=differences)
 
 
+# Check for DISPLAY variable for matplotlib
+# -----------------------------------------
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
+
 # Configure LaTeX style for plots
 # -------------------------------
 try:

### 2018-12-26 code/scripts/grid_search.py

@@ -131,6 +131,14 @@ for field in fields:
                  differences=differences)
 
 
+# Check for DISPLAY variable for matplotlib
+# -----------------------------------------
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
+
 # Configure LaTeX style for plots
 # -------------------------------
 try:

### 2018-12-26 code/scripts/linear_density.py

@@ -102,6 +102,14 @@ for field in fields:
                      D_values=D_values, differences=differences)
 
 
+# Check for DISPLAY variable for matplotlib
+# -----------------------------------------
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
+
 # Configure LaTeX style for plots
 # -------------------------------
 try:

### 2018-12-26 code/scripts/neuquen_basin.py

@@ -164,6 +164,14 @@ for field in fields:
 
 # %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+# Check for DISPLAY variable for matplotlib
+# -----------------------------------------
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
+
 # Configure LaTeX style for plots
 # -------------------------------
 try:
@@ -185,11 +193,7 @@ warnings.filterwarnings("ignore")
 heights = np.linspace(bottom, top, 101)
 
 # Initialize figure and subplots
-try:
-    fig, ax = plt.subplots(figsize=(3.33, 3))
-except Exception:
-    plt.switch_backend('agg')
-    fig, ax = plt.subplots(figsize=(3.33, 3))
+fig, ax = plt.subplots(figsize=(3.33, 3))
 
 ax.plot(heights, linear_density(heights), label="Linear")
 ax.plot(heights, density_exponential(heights), label="Exponential")
@@ -201,10 +205,7 @@ fig.tight_layout()
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen-basin-densities.pdf")
 plt.savefig(figure_fname, dpi=300)
-try:
-    plt.show()
-except Exception:
-    pass
+plt.show()
 
 
 # ------------
@@ -242,11 +243,7 @@ config['parallels-right']['labels'] = [False, True, False, False]
 
 
 # Initiate figure and subplots
-try:
-    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 7))
-except Exception:
-    plt.switch_backend('agg')
-    fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 7))
+fig, axes = plt.subplots(nrows=2, ncols=2, figsize=(6.66, 7))
 
 
 # Topography
@@ -368,21 +365,14 @@ fig.tight_layout()
 figure_fname = os.path.join(script_path,
                             "../../manuscript/figures/neuquen-basin.pdf")
 plt.savefig(figure_fname, dpi=300)
-try:
-    plt.show()
-except Exception:
-    pass
+plt.show()
 
 
 # ----------------
 # Plot differences
 # ----------------
 # Initiate figure and subplots
-try:
[…]

### 2018-12-26 code/scripts/sine_density.py
Fix DISPLAY variable related error in sine script

@@ -122,6 +122,14 @@ for field in fields:
                          delta_values=delta_values, differences=differences)
 
 
+# Check for DISPLAY variable for matplotlib
+# -----------------------------------------
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
+
 # Configure LaTeX style for plots
 # -------------------------------
 try:

### 2018-12-28 manuscript/manuscript.tex
Put myself as last author
In this case, I supervised most of the work so it makes more sense for
me to be last author and Santiago first.

@@
 }
-\author[S.R. Soler, A. Pesce, L. Uieda and M.E. Gimenez]{
-    Santiago R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and
-    Mario E. Gimenez$^{1,2}$ \\
+\author[S.R. Soler, A. Pesce, M.E. Gimenez, and L. Uieda]{
+    Santiago R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$,
+    Mario E. Gimenez$^{1,2}$, and Leonardo Uieda$^3$ \\
     $^1$CONICET, Argentina.~e-mail: santiago.r.soler@gmail.com\\
@@
 Lastly, we apply this new methodology to model the Neuqu\'en Basin, a foreland basin in
-Argentina with a maximum depth of over 5000 m, using an exponential density function.
+Argentina with a maximum depth of over 5000~m, using an exponential density function.
 \end{summary}
@@
 These results indicate that the values of $D_\text{linear}$ can be safely extrapolated
-to the exponential case.
+to non linear density cases.
 This conclusion can be understood considering that the two dimensional adaptive
@@
 The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold
-for each computation grid, shell model and $b$ value if $\delta = 0.1$.
-This results imply that a value of $\delta=0.1$ is needed in order to
-achieve the 0.1\% accuracy threshold for both the potential and the $g_{z}$, every
-shell model, computation grid and value of $b$, in case of exponential density.
-These results also show that there is no significant relation between the $b$
-value and the numerical error, proving that the density-based discretization algorithm
-is capable of limiting the error due to the density variation regardless the variation
-of the exponential density.
+for each computation grid, shell model and $b$ value if $\delta$ is lower than 0.1.
+This results imply that the optimal value of $\delta$ is 0.1 in case of exponential
+density.
+They also show that there is no significant relation between the $b$ value and the
+numerical error, proving that the density-based discretization algorithm is capable of
+limiting the error due to the density variation regardless the variation of the
+exponential density.
 
@@
 shell model, computation grid and value of $b$.
-This results also proof that even for a continuous density function with several local
-maximum and minimums and inflection points, the density-based discretization algorithm
-performs well on computing the gravitational fields that a tesseroid with such density
-generates.
+Even in case of a continuous density function with several local maximums and minimums
+and inflection points, the GLQ along with the two discretization algorithms reproduce
+the gravitational fields with an acceptable degree of accuracy.
 
@@
     Linear and exponential densities used to compute the gravitational fields generated
-    by a tesseroid model of the Neuquen sedimentary basin.
+    by a tesseroid model of the Neuqu\'en sedimentary basin.
     The height is defined above the mean Earth radius, and its axis is spanned between
@@
 \caption{
-    Gravitational effects of the Nequ\'en sedimentary basin modelled
+    Gravitational effects of the Neuqu\'en sedimentary basin modelled
     using tesseroids with an exponential density function of depth.
@@
     Differences between the gravitational fields generated by the tesseroid model of the
-    Neuquen basin with an exponential density contrast and with a homogenenous and
+    Neuqu\'en basin with an exponential density contrast and with a homogeneous and
     linear density variation.
     \mbox{(a)-(b)}~Differences on $V$ and $g_z$ between the exponential density model
-    and the homogenenous density one,
+    and the homogeneous density one,
     \mbox{(c)-(f)}~differences on $V$ and $g_z$ between the exponential density model
@@
 density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
+Eq~\ref{eq:shell-pot-linear} is in agreement with the one obtained by \citet{Lin2018}.
 

### 2018-12-28 manuscript/manuscript.tex
Edits to the introduction
I re-worded some sections to make things clearer and included more
citations to Lin and Denker. I re-wrote the part justifying why we need
a new method. The accuracy of the integration is probably not an issue
if you did a piecewise linear function (as our own results suggest).
Even using a high-degree polynomial wouldn't be too bad if the fit to
the density data is good. The main problem is that piecewise linear
raises the question of "what is the discretization interval" and "how
can we do this automatically"? We answer these with our algorithm. The
high-degree poly

@@
 i.e.~spherical prisms, which are defined by pairs of latitude, longitude, and radial
-boundaries (see Fig.~\ref{fig:tesseroid}).
-The gravitational fields generated by an arbitrary
+boundaries (Fig.~\ref{fig:tesseroid}).
+The gravitational fields generated by a
 tesseroid on any external point are given by volume
@@
 Quadrature (GLQ)
-\citep{Asgharzadeh2007, Wild-Pfeiffer2008, Li2011, Uieda2016}.
+\citep{Asgharzadeh2007, Wild-Pfeiffer2008, Li2011, Uieda2016, Lin2018}.
 The Taylor series expansion is not well suited to develop an algorithm for
@@
 any change to the integration method.
+Thus, we will focus only on GLQ-based methods henceforth.
 
@@
 
-The main challenge of the GLQ integration is loss of accuracy when the computation point
-gets closer to the tesseroid \citep{Ku1977}.
+The main challenge of the GLQ integration is the loss of accuracy that occurs when the
+computation point approaches the tesseroid \citep{Ku1977}.
 \citet{Uieda2016} built on the three dimensional adaptive discretization algorithm of
 \citet{Li2011} to automatically obtain integration results with 0.1\% accuracy.
-The algorithm consist in recursively splitting the tesseroid into smaller ones when a
+The algorithm recursively divides the tesseroid into smaller ones when a
 threshold is exceeded,
 namely when the normalized distance to the computation point is greater than a
-``distance-size ratio'' ($D$).
+``distance-size ratio'' parameter ($D$).
 \citet{Uieda2016} have also obtained standard values of $D$
 for the gravitational potential, its gradient and Marussi tensor components
-by comparing the numerical model with the fields generated by a spherical shell.
-
-Two recent papers present other approaches for homogeneous tesseroids gravitational
-fields computation and incorporate methodologies for variable density tesseroid with
-specific density profiles.
+by comparing the numerical integration results with the fields generated by a spherical
+shell.
+
+Two recent publications present alternative approaches for calculating the gravitational
+fields of homogeneous tesseroids and incorporate methodologies for tesseroids with
+variable densities in depth.
 \citet{Fukushima2018} analytically integrates the volumetric integral for the
 gravitational potential in the radial direction, obtaining a surface integral, which is
-then numerically solved by applying a conditionally splitting and the double exponential
-quadrature rule. The gradient of the potential and the Marussi tensor components are
-then computed by finite differences. \citet{Fukushima2018} also generalizes his method
-to an arbitrary polynomial density profile on the radial coordinate.
-\citet{Lin2018} made a comparison of the different methodologies for homogeneous
-tesseroid gravitational fields computation by applying different discretizations
-algorithms to each one of them.
-From this analysis they develop a combined method:
-for computation points near the tesseroid they apply a GLQ numerical integration with
-a two dimensional adaptive discretization algorithm, i.e.~only in the horizontal
-dimension. But, if the computation point is farther than a certain truncation distance,
-then a second order Taylor series approximation is applied along with the regular
+then numerically solved by conditionally splitting the tesseroid and applying the double
+exponential quadrature rule.
+The gradient of the potential and the Marussi tensor components are computed by finite
+differences.
+\citet{Fukushima2018} also generalized their method to tesseroids with a radial
+polynomial density function of arbitrary degree.
+\citet{Lin2018} compared the different integration and discretization methodologies for
+homogeneous tesseroids.
+From this analysis they developed a combined method:
+for computation points near the tesseroid, they use a GLQ integration with an adaptive
+discretization based on \citet{Uieda2016} but only applied to the horizontal dimensions.
+If the computation point is farther than a certain truncation distance,
+a second order Taylor series approximation is applied instead along with the regular
 subdivision developed by \citet{Grombein2013}.
-They also introduce a variation of their combined method to compute the gravitational
-fields generated by tesseroids with linear density on the radial coordinate.
-
-In summary, the bibliography is offering different methodologies for computing
-gravitational fields of homogeneous tesseroids or with certain specific density
-functions in depth.
-Although any continuous density profile could be approximated by one of these specific
-density functions, one should acknowledge the error that such approximation would
-propagate into the gravitational field in order to guarantee certain degree of accuracy.
-
-We present a new algorithm for computing the gravitational fields generated by any
-tesseroid with any arbitrary continuous density function on any external point.
-It is based on the three dimensional GLQ approximation, a two dimensional version of the
-adaptive discretization of \citet{Uieda2016} and a new density-based discretization
-algorithm that we introduce in this work.
-In order to ensure the accuracy of the numerical approximation we have
-determined optimal values for the controlling parameters by
-comparing the numerical approximation with analytical solutions for
-spherical shells.
-Finally, we applied the methodology to model the Neuqu\'en basin, Argentina, using
+\citet{Lin2018} also introduced a variation of their combined method to compute the
+gravitational fields generated by tesseroids with a linearly varying density in the
+radial dimension.
+
+Both the \citet{Lin2018} and the \citet{Fukushima2018} studies limit the radial density
+variation to polynomial functions.
+While most continuous and smooth functions can be approximated by piecewise linear
+functions, the choice of a discretization interval is neither straight forward nor
+automatic.
+Furthermore, it is well known that the use of high-degree polynomials to approximate a
+highly variable function produces unstable results when extrapolating beyond the data
+domain.
+These shortcomings could make piecewise linear or high-degree polynomial density
+functions unwieldy for non-linear gravity inversions if the density function in depth
+is highly variable.
+
+We present a new algorithm for computing the gravitational fields generated by a
+tesseroid with an arbitrary continuous density function on an external point.
+It is based on the three dimensional GLQ integration, a two dimensional version of the
+adaptive discretization of \citet{Uieda2016} (following \citet{Lin2018}),
+and a new density-based radial discretization algorithm.
+To ensure the accuracy of the numerical approximation, we empirically determine
+optimal values for the controlling parameters by comparing the numerical results with
+analytical solutions for spherical shells.
+Finally, we apply the methodology to model the Neuqu\'en basin, Argentina, using
 tesseroids with linear and exponentially increasing density with depth.

### 2018-12-28 manuscript/manuscript.tex
Edits to start of methodology and GLQ
Minor changes to the text.

@@
 
-We present a new algorithm for computing the gravitational fields generated by a
-tesseroid with an arbitrary continuous density function on an external point.
+We present a new algorithm for computing the gravitational potential and its gradient
+generated by a tesseroid with an arbitrary continuous density function on an external
+point.
 It is based on the three dimensional GLQ integration, a two dimensional version of the
@@
 
-Consider an arbitrary tesseroid in a geocentric spherical coordinate system defined by
+Consider a tesseroid in a geocentric spherical coordinate system defined by
 pairs of geocentric latitudinal ($\phi_1$, $\phi_2$), longitudinal ($\lambda_1$,
-$\lambda_2$) and radial ($r_1$, $r_2$) boundaries.
+$\lambda_2$), and radial ($r_1$, $r_2$) boundaries.
 We define an external computation point $P(r, \phi, \lambda)$ located at radius $r$,
-geocentric latitude $\phi$, and longitude $\lambda$ where the gravitational fields the
-tesseroid generates are going to be calculated.
-The first- and second-order derivatives of the gravitational potential are taken with
-respect to the local north-oriented Cartesian coordinate system of $P$
+geocentric latitude $\phi$, and longitude $\lambda$.
+The derivatives of the gravitational potential are taken with
+respect to the local north-oriented Cartesian coordinate system with origin at $P$
 (Fig.~\ref{fig:tesseroid}).
@@
 arbitrary continuous function $\rho(r)$.
-Thus, the integrals for the gravitational fields are slightly modified to
+Thus, the integrals for the gravitational potential and its first-order derivatives are
+slightly modified to
 
@@
 \end{equation}
-
-\noindent in which $\alpha, \beta \in \{x, y, z\}$, $\delta_{\alpha\beta}$ is
-Kronecker's delta, $G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the
-gravitational constant and
+\\
+\noindent in which $\alpha \in \{x, y, z\}$,
+$G = 6.674\times10^{-11}\, \text{m$^3$kg$^{-1}$s$^{-1}$}$ is the gravitational constant
+and
 
@@
 }
-
+\\
 \noindent where
@@
 \begin{equation}
-    A =
-    \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8},
-\end{equation}
-
-\noindent $f(r', \phi', \lambda')$ is the non-density dependent portion of the
-integration kernels, $(r_i, \phi_j, \lambda_k)$ are the quadrature nodes, and
-$W_i^r$, $W_j^\phi$, $W_k^\lambda$ are the quadrature weights.
-The GLQ nodes and weights can be calculated as \citet[p.~391]{Hildebrand1987} (see
-\citet{Uieda2016} for nodes and weights definitions on similar notation used on previous
-equations).
-
-It's worth noting that a GLQ is equivalent to approximate the tesseroid by $N^r \times
-N^\phi \times N^\lambda$ point masses located on the quadrature nodes
-\citep{Ku1977, Li2011, Uieda2016}.
+    A = \frac{(\lambda_2 - \lambda_1)(\phi_2 - \phi_1)(r_2 - r_1)}{8},
+\end{equation}
+\\
+\noindent $f(r', \phi', \lambda')$ is an integral kernel for a homogeneous tesseroid
+\citep{Grombein2013},
+$(r_i, \phi_j, \lambda_k)$ are the coordinates of the quadrature nodes,
+$N^r$, $N^\phi$, $M^\lambda$ are the quadrature orders and $W_i^r$, $W_j^\phi$,
+$W_k^\lambda$ are the quadrature weights in the radial, latitudinal, and longitudinal
+directions, respectively.
+\citet[p.~391]{Hildebrand1987} provides the formulas for calculating the GLQ weights and
+pre-computed values of the Legendre polynomial roots for low orders.
+See \citet{Uieda2016} for a more detailed description using a similar notation to the
+one used here.
+It is worth noting that a GLQ is equivalent to approximating the tesseroid by $N^r
+\times N^\phi \times N^\lambda$ point masses located on the quadrature nodes
+\citep{Ku1977, Asgharzadeh2007}.
 
@@
 Although this method would involve a computational error due to the approximation of the
-density function as a polynomial one, it's feasibly if we account for this error and
+density function as a polynomial one, its feasibly if we account for this error and
 a method to limit it.

### 2018-12-28 manuscript/manuscript.tex
Minor changes in exponential density text

--- 2018-12-28 manuscript/manuscript.tex
@@
 }
-\author[S.R. Soler, A. Pesce, M.E. Gimenez, and L. Uieda]{
-    Santiago R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$,
-    Mario E. Gimenez$^{1,2}$, and Leonardo Uieda$^3$ \\
+\author[S.R. Soler, A. Pesce, L. Uieda and M.E. Gimenez]{
+    Santiago R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and
+    Mario E. Gimenez$^{1,2}$ \\
     $^1$CONICET, Argentina.~e-mail: santiago.r.soler@gmail.com\\
@@
 Lastly, we apply this new methodology to model the Neuqu\'en Basin, a foreland basin in
-Argentina with a maximum depth of over 5000~m, using an exponential density function.
+Argentina with a maximum depth of over 5000 m, using an exponential density function.
 \end{summary}
@@
 shell model, computation grid and value of $b$.
-Even in case of a continuous density function with several local maximums and minimums
-and inflection points, the GLQ along with the two discretization algorithms reproduce
-the gravitational fields with an acceptable degree of accuracy.
+This results also proof that even for a continuous density function with several local
+maximum and minimums and inflection points, the density-based discretization algorithm
+performs well on computing the gravitational fields that a tesseroid with such density
+generates.
 
@@
     Linear and exponential densities used to compute the gravitational fields generated
-    by a tesseroid model of the Neuqu\'en sedimentary basin.
+    by a tesseroid model of the Neuquen sedimentary basin.
     The height is defined above the mean Earth radius, and its axis is spanned between
@@
 \caption{
-    Gravitational effects of the Neuqu\'en sedimentary basin modelled
+    Gravitational effects of the Nequ\'en sedimentary basin modelled
     using tesseroids with an exponential density function of depth.
@@
     Differences between the gravitational fields generated by the tesseroid model of the
-    Neuqu\'en basin with an exponential density contrast and with a homogeneous and
+    Neuquen basin with an exponential density contrast and with a homogenenous and
     linear density variation.
     \mbox{(a)-(b)}~Differences on $V$ and $g_z$ between the exponential density model
-    and the homogeneous density one,
+    and the homogenenous density one,
     \mbox{(c)-(f)}~differences on $V$ and $g_z$ between the exponential density model
@@
 density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
-Eq~\ref{eq:shell-pot-linear} is in agreement with the one obtained by \citet{Lin2018}.
 

### 2018-12-28 manuscript/manuscript.tex
Minor fixes to the adaptive discretization text
Mostly moved the definition of the tesseroid size to Step 1 so that it
follows the definition of the inequality. It was breaking the flow of
the text before that.

--- 2018-12-28 manuscript/manuscript.tex
@@
 of point masses is produced where it is needed more.
-\citet{Uieda2016} developed a modified version of this algorithm that speeds the
-computation of the distances.
-Both algorithms (the original one from \citet{Li2011} and the modified
-version of \citet{Uieda2016}) perform tesseroid subdivisions in the latitudinal,
-longitudinal and radial directions, thus we can define them as three dimensional
+\citet{Uieda2016} developed a modified version of this algorithm along with an efficient
+computational implementation.
+Both algorithms perform tesseroid subdivisions in the latitudinal,
+longitudinal and radial directions, thus we can define them as three-dimensional
 adaptive discretization algorithms.
@@
 that subdivides the tesseroid only on the latitudinal and longitudinal directions.
-Removing a dimension from the splitting makes the computation more efficient, while
-keeping an acceptable accuracy on the gravitational fields \citep{Lin2018}.
-
-Here we will use a two dimensional version of the adaptive discretization algorithm
-developed by \citet{Uieda2016}: the distances from the tesseroid and the computation
-point are computed as \citet{Uieda2016} do, but the tesseroid will only be split on the
-longitudinal and latitudinal directions.
+Removing a dimension from the discretization makes the computation more efficient by
+reducing the number of tesseroids in the model, while
+retaining an acceptable accuracy \citep{Lin2018}.
+
+Here we will follow \citet{Lin2018} and use a two dimensional version of the adaptive
+discretization of \citet{Uieda2016}.
 What follows is a summary of the algorithm and the reader is referred to
@@
 
-Lets consider a computation point located at $(r, \phi, \lambda)$ and an arbitrary
-tesseroid defined by the boundaries $r_1$, $r_2$, $\phi_1$, $\phi_2$, $\lambda_1$ and
-$\lambda_2$, which geometric centre is located at $(r_t, \phi_t, \lambda_t)$, where
-
-\begin{equation}
-    r_t = \frac{r_2 + r_1}{2}, \quad
-    \phi_t = \frac{\phi_2 + \phi_1}{2}, \quad
-    \lambda_t = \frac{\lambda_2 + \lambda_1}{2}.
-\end{equation}
-
-We can define the longitudinal ($L_\lambda$) and latitudinal ($L_\phi$) dimensions of
-the tesseroid as follows:
-
-\begin{equation}
-    L_\lambda = r_2 \arccos(\sin^2\phi_t +
-        \cos^2\phi_t\cos(\lambda_2 - \lambda_1)),
-    \label{eq:sizelon}
-\end{equation}
-
-\begin{equation}
-    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1).
-\end{equation}
-
 \textit{Step 1}: Check that the tesseroid satisfies the following inequality for each
-dimension $L_i (i \in \{\lambda, \phi\})$ of the tesseroid:
+dimension $L_i\ (i \in \{\lambda, \phi\})$ of the tesseroid:
 
@@
 \noindent
-in which $D$ is a positive scalar called the distance-size ratio and $d$ is the distance
+in which $D$ is a positive scalar called the distance-size ratio, $d$ is the distance
 between the computation point and the geometric centre of the tesseroid
@@
         \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t).
+\end{equation}
+
+\begin{equation}
+    r_t = \frac{r_2 + r_1}{2}, \quad
+    \phi_t = \frac{\phi_2 + \phi_1}{2}, \quad
+    \lambda_t = \frac{\lambda_2 + \lambda_1}{2}.
+\end{equation}
+\\
+\noindent
+and the dimensions of the tesseroid are defined as
+
+\begin{equation}
+    L_\lambda = r_2 \arccos(\sin^2\phi_t +
+        \cos^2\phi_t\cos(\lambda_2 - \lambda_1)),
+    \label{eq:sizelon}
+\end{equation}
+\\
+\noindent and
+
+\begin{equation}
+    L_\phi = r_2 \arccos(\sin\phi_2\sin\phi_1 + \cos\phi_2\cos\phi_1).
 \end{equation}
@@
 \textit{Step 3}:
-If any dimension fails inequality~\ref{eq:condition}, split the tesseroid in half along
-the offending dimensions.
-Repeat steps 1-3 for all smaller tesseroids until none are left.
+If inequality~\ref{eq:condition} does not hold for one of the dimensions
+(longitudinal or latitudinal), split the tesseroid in half along that dimension.
+Repeat steps 1-3 for all smaller tesseroids until none are left that violate
+inequality~\ref{eq:condition}.
 
 \textit{Final step}:
-By the end of the algorithm, a second-order GLQ will be applied on each small tesseroid
-to compute the gravitational field they generate, and the sum of these results will be
-the gravitational effect of the original tesseroid.
-
-Notice that the distance-size ratio $D$ determines how many
-times the tesseroids will be divided.
+By the end of the algorithm, a second-order GLQ will have been applied to each smaller
+tesseroid and the running total gathered in step 2 will be the gravitational effect of
+the original tesseroid.
+
+The distance-size ratio $D$ determines how many times the tesseroids will be divided.
 Therefore, it effectively regulates both the accuracy of the algorithm and its
@@
 Here, we will repeat the numerical experiment using analytical expressions for shells
-with density varying according to certain functions of $r$.
-This numerical experiment will also show that the two dimensional adaptive
-discretization algorithm can achieve the same accuracy as the original three dimensional
-version, though it reduces the number of tesseroid subdivisions and thus the computation
-time.
+with density varying according to functions of $r$.
+This experiment will also test whether the same values of $D$ determined by
+\citet{Uieda2016} can be used for the two dimensional adaptive discretization.
 

### 2018-12-28 manuscript/manuscript.tex
Add missing accents to Neuquen

--- 2018-12-28 manuscript/manuscript.tex
@@
     Differences between the gravitational fields generated by the tesseroid model of the
-    Neuquen basin with an exponential density contrast and with a homogenenous and
+    Neuqu\'en basin with an exponential density contrast and with a homogenenous and
     linear density variation.

### 2018-12-28 manuscript/manuscript.tex
Fix typos

@@
     Santiago R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$, Leonardo Uieda$^3$ and
-    Mario E. Gimenez$^{1,2}$ \\
+    Mario E. Gim\'enez$^{1,2}$ \\
     $^1$CONICET, Argentina.~e-mail: santiago.r.soler@gmail.com\\
@@
     Differences between the gravitational fields generated by the tesseroid model of the
-    Neuqu\'en basin with an exponential density contrast and with a homogenenous and
+    Neuqu\'en basin with an exponential density contrast and with a homogeneous and
     linear density variation.
     \mbox{(a)-(b)}~Differences on $V$ and $g_z$ between the exponential density model
-    and the homogenenous density one,
+    and the homogeneous density one,
     \mbox{(c)-(f)}~differences on $V$ and $g_z$ between the exponential density model

### 2018-12-28 code/scripts/exponential_density.py
Minor change in plots ordering

@@ -184,6 +184,7 @@ grid_titles = {"pole": "Pole",
                "equator": "Equator",
                "global": "Global",
                "260km": "Satellite"}
+grid_names = ["pole", "equator", "global", "260km"]
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
 # Create outer grid
@@ -193,7 +194,7 @@ outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 # Create grid specs for each grid
 grid_specs = dict(
                   zip(
-                      grids.keys(),
+                      grid_names,
                       [GridSpecFromSubplotSpec(ncols=1, nrows=2,
                                                subplot_spec=outer_grid[j, i],
                                                hspace=0)
@@ -201,10 +202,10 @@ grid_specs = dict(
                        for i in range(2)]
                      )
                  )
-positions = dict(zip(grids.keys(), [[j, i] for j in range(2) for i in range(2)]))
+positions = dict(zip(grid_names, [[j, i] for j in range(2) for i in range(2)]))
 
 # Plot for each grid
-for grid_name in grids:
+for grid_name in grid_names:
 
     # Choose gridspec and create axes
     gs = grid_specs[grid_name]

### 2018-12-28 code/scripts/sine_density.py

@@ -194,6 +194,7 @@ grid_titles = {"pole": "Pole",
                "equator": "Equator",
                "global": "Global",
                "260km": "Satellite"}
+grid_names = ["pole", "equator", "global", "260km"]
 colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 
 # Create outer grid
@@ -203,7 +204,7 @@ outer_grid = GridSpec(ncols=2, nrows=2, wspace=0.001, hspace=0.1)
 # Create grid specs for each grid
 grid_specs = dict(
                   zip(
-                      grids.keys(),
+                      grid_names,
                       [GridSpecFromSubplotSpec(ncols=1, nrows=2,
                                                subplot_spec=outer_grid[j, i],
                                                hspace=0)
@@ -211,10 +212,10 @@ grid_specs = dict(
                        for i in range(2)]
                      )
                  )
-positions = dict(zip(grids.keys(), [[j, i] for j in range(2) for i in range(2)]))
+positions = dict(zip(grid_names, [[j, i] for j in range(2) for i in range(2)]))
 
 # Plot for each grid
-for grid_name in grids:
+for grid_name in grid_names:
 
     # Choose gridspec and create axes
     gs = grid_specs[grid_name]

### 2018-12-28 code/scripts/exponential_density.py
Change location of legend in exponential fig

@@ -268,7 +268,7 @@ for grid_name in grid_names:
         ax.grid(True, linewidth=0.5, color='#aeaeae')
 
     # Add legend
-    if position == [0, 0]:
+    if position == [0, 1]:
         axes[0].legend(loc=0, prop={"size": 8})
 
 outer_grid.tight_layout(fig)

### 2018-12-28 manuscript/manuscript.tex
Add citation to Lin2018 in Appendix
Edits to the density discretization and software
Minor edits to the algorithm to make it more like the adaptive
discretization description.
Added some more comments about how we can use any function because the
GLQ is the same and our software can handle user defined functions.

--- 2018-12-28 manuscript/manuscript.tex
@@
 The numerical integration of an arbitrary continuous density function introduces a new
-type of problem: the integration error from using only a few nodes to account for the
-variation of the density function.
-The three dimensional adaptive discretization may help
-to reduce this kind of error by adding more point masses in the radial direction.
-However, it does not take into account the density function when subdividing the
+type of problem: the integration error from using only a few quadrature nodes to account
+for the variation of the density function.
+The three dimensional adaptive discretization may help to reduce this kind of error by
+adding more point masses in the radial direction.
+However, it does not take into account the density function itself when dividing the
 tesseroid and hence it is not well suited to fully perform this task.
-The two dimensional adaptive discretization algorithm does not subdivide the tesseroid
-on the radial dimension, so it does not put any effort on lowering this new kind of
-error.
-
-We have developed a complementary discretization
-algorithm that takes into account the variations of the density function.
-This density-based discretization is applied prior to the adaptive discretization
-described in the previous section.
+
+We have developed a complementary discretization algorithm in the redial direction that
+takes into account the variations of the density function.
+This density-based discretization is applied prior to the two dimensional adaptive
+discretization described in the previous section.
 In short, the algorithm divides the tesseroid along the radial dimension at the
@@
 
-Consider an \emph{original} tesseroid with a continuous density given by the function
+Consider an \emph{original} tesseroid with a density given by the continuous function
 $\rho(r')$.
@@
 \textit{Step 1}:
-Define a linear function $\rho_l(r')$ that assumes
-the same values as the normalised density $\rho_n(r')$ at the boundaries of the
-tesseroid ($r_1$ and $r_2$):
+Define a linear function $\rho_l(r')$ that assumes the same values as the normalised
+density function $\rho_n(r')$ at the boundaries of the tesseroid ($r_1$ and $r_2$):
 
@@
 (Eq.~\ref{eq:density-abs-diff}) takes place.
-
-\textit{Step 6}:
-Repeat steps 1-6 for each smaller tesseroid produced in step 5.
-
-Once all smaller tesseroids satisfy Eq.~\ref{eq:delta-density},
-each one is subjected to the two dimensional adaptive
-discretization algorithm described earlier to calculate their gravitational effects.
+Repeat steps 1-5 for each smaller tesseroid produced in this step.
+
+\textit{Final step}:
+Once all smaller tesseroids satisfy inequality.~\ref{eq:delta-density}, each one is
+subjected to the two dimensional adaptive discretization algorithm described earlier to
+calculate their gravitational effects.
 
@@
 
-In summary, given any tesseroid with variable density in depth and an arbitrary
-external computation point, we propose the following steps to numerically
-compute the gravitational fields:
+In summary, given a tesseroid with density varying in depth according to an arbitrary
+continuous function, we propose the following steps to numerically compute its
+gravitational fields on an external point:
 
 \textit{Step 1:}
-Apply the density-based discretization algorithm.
-It subdivides the tesseroid only in the radial dimension, thus it will return a set of
-tesseroids with the same longitudinal and latitudinal dimensions as the original one but
-with different radial boundaries.
+Apply the density-based discretization algorithm to discretize the tesseroid in the
+radial dimension, producing a set of tesseroids with the same longitudinal and
+latitudinal dimensions as the original one but with different radial boundaries.
 
 \textit{Step 2:}
-Apply the two dimensional adaptive discretization algorithm on each tesseroid obtained
+Apply the two dimensional adaptive discretization algorithm for each tesseroid obtained
 in the previous step.
-If needed, it will subdivide each tesseroid only in the latitudinal and longitudinal
-direction, generating a set of smaller tesseroids that resemble the original one.
+If needed, the algorithm will divide each tesseroid in the latitudinal and longitudinal
+direction, generating a set of smaller tesseroids.
 
@@
 previous step.
-Then sum these results in order to obtain the gravitational field of the original
-tesseroid.
+The numerical integration includes the density function and can be applied
+without modification to any continuous function.
+The sum of these results is the gravitational field of the original tesseroid.
 
@@
 programming language.
-The software is based on the pre-existing code for the homogeneous density tesseroid
-\citep{Uieda2016}, more specifically the implementation in the Python library Fatiando a
-Terra v0.5 \citep{Uieda2013}.
+The software is based on the Python implementation of \citet{Uieda2016} that is included
+in the Fatiando a Terra library v0.5 \citep{Uieda2013}.
 The more time consuming parts of the algorithm are written in the Cython language to
 achieve higher performance.
+We leverage the dynamic nature of the Python language to allow user-defined radial
+density functions as inputs to the software.
+Thus, our code can evaluate linear, exponential, polynomial, sinusoidal, cubic splines,
+or any other continuous density function without modification.
 This new code is freely available under the BSD 3-clause open-source license.
-It can be downloaded from the online repository
+It can be freely downloaded from the online repository
 \href{https://github.com/pinga-lab/tesseroid-variable-density}{github.com/pinga-lab/tesseroid-variable-density}.

### 2018-12-28 manuscript/manuscript.tex
Move the description of grids and shells to the top
The description of each grid and shell model was in the linear density
section but we them for the other functions as well. I moved them up to
the beginning of the section and added some more justification about
why we chose these values (as requested by the reviewers).
I also simplified the formatting of the shell models table. I think this
is simpler and the caption can be a bit shorter because of it.

@@
 a spherical shell with linear density in the radial coordinate.
-We derive an expression for such gravitational potential in case the spherical shell has
-an arbitrary density function in depth (Eq.~\ref{eq:shell-pot}), and obtained the
-expressions in case of linear and exponential density (see Appendix~\ref{sec:shell}).
+We derive an expression for the gravitational potential of a spherical shell with
+an arbitrary density function in depth (Eq.~\ref{eq:shell-pot}),
+from which we obtained
+expressions for linear, exponential, and sinusoidal density functions
+(see Appendix~\ref{sec:shell}).
+
+In order to compare the numerical results with the analytical solution we must build
+spherical shell models out of tesseroids.
+We divide the spherical shell along the latitudinal and longitudinal directions to
+obtain a shell model made out of $6 \times 12 = 72$ tesseroids of size
+$30^\circ \times 30^\circ$.
+In order to asses the density-based discretization in the radial dimension,
+we use spherical shell models with different thicknesses (Table~\ref{tab:shell-models}).
+The thickness values were chosen to represent a range of applications, from topographic
+to lithospheric scale models.
+Because the amount of tesseroid divisions in the adaptive discretization will be
+proportional to the tesseroid size (Eq.~\ref{eq:condition}),
+some of these configurations represent worse-case scenarios.
+Most practical applications will use tesseroids smaller than
+$30^\circ \times 30^\circ \times 1000\ \text{km}$.
+
+The differences between the analytical and numerical solutions could be calculated on a
+single computation point due to the rotational symmetry of the spherical shell.
+However, the numerical results depend on the relative location between the computation
+point and the tesseroid \citep{Ku1977, Asgharzadeh2007, Uieda2016}.
+We account for this effect by calculating the differences on regular grids and keeping
+only the maximum absolute difference.
+All calculations are repeated for four different computation grids
+(Table~\ref{tab:grids}):
+a local grid at the pole, a local grid at the equator, a global grid at zero height
+above the shell, and a global grid at a height of 260 km (representing the nominal
+height of the GOCE satellite).
+These grids span a broad scenario of use cases and ensure an acceptable accuracy on each
+one of them.
 
 We perform comparisons between the analytical solutions for the spherical shell and the
-numerical integration results for linear, exponential and trigonometric density
+numerical integration results for linear, exponential, and sinusoidal density
 functions.
-From these results, we generalize default values for $D$ and $\delta$ that ensure a
-numerical error lower than 0.1\% of the spherical shell values.
+The comparisons are repeated for all combinations of tesseroid models in
+Table~\ref{tab:shell-models} and computation grids in Table~\ref{tab:grids}.
+From these results, we generalize optimal values for $D$ and $\delta$ that ensure a
+numerical error lower than 0.1\% of the spherical shell values for most use cases.
 
@@
     the accuracy of the numerical integration.
-    The first column indicates the thickness of the shell model and the next three
-    describe the shape of the tesseroid model: each one specifies the number of
-    tesseroids on the corresponding dimension (radial, latitudinal, longitudinal).
-    The last two columns specify the latitudinal and longitudinal size of each
-    tesseroid that compose the spherical shell model.
-    The outer radius of every shell model is equal to the mean Earth radius (6378.137
-    km), while the inner radius is determined by its thickness.
+    The outer radius ($R_2$) of every shell model is equal to the mean Earth radius
+    (6378.137 km), while the inner radius ($R_1$) is determined by its thickness.
+    The horizontal dimensions of the tesseroids and the total number of
+    tesseroids in the shell model are given in the latitudinal and longitudinal
+    dimensions, respectively.
 }
 \label{tab:shell-models}
-\begin{tabular}{cccccc}
-    Thickness & Radial & Latitudinal & Longitudinal & $L_\phi$ & $L_\lambda$ \\ \hline
-    100m   & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
-    1km    & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
-    10km   & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
-    100km  & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
-    1000km & 1 & 6 & 12 & 30$^\circ$ & 30$^\circ$ \\
+\begin{tabular}{rccccc}
+    Thickness & Tesseroid size & Number of tesseroids \\ \hline
+    0.1 km  & $30^\circ \times 30^\circ$ & $6 \times 12 = 72$ \\
+    1 km    & $30^\circ \times 30^\circ$ & $6 \times 12 = 72$ \\
+    10 km   & $30^\circ \times 30^\circ$ & $6 \times 12 = 72$ \\
+    100 km  & $30^\circ \times 30^\circ$ & $6 \times 12 = 72$ \\
+    1000 km & $30^\circ \times 30^\circ$ & $6 \times 12 = 72$ \\
 \end{tabular}
@@
 \begin{tabular}{lccc}
-    Name & Grid size & Grid region (degrees) & Grid height (km)
+    Name & Grid spacing & Grid region (degrees) & Grid height (km)
     \\ \hline
-    Pole      & $0.1^\circ \times 0.1^\circ$ &   0E /   1E / 89N / 90N & 0   \\
-    Equator   & $0.1^\circ \times 0.1^\circ$ &   0E /   1E /  0N / 1N  & 0   \\
-    Global    & $ 10^\circ \times  10^\circ$ & 180W / 180E / 90S / 90N & 0   \\
-    Satellite & $ 10^\circ \times  10^\circ$ & 180W / 180E / 90S / 90N & 260 \\
+    Pole      & $0.1^\circ$ &   0E /   1E / 89N / 90N & 0   \\
+    Equator   & $0.1^\circ$ &   0E /   1E /  0N / 1N  & 0   \\
+    Global    & $ 10^\circ$ & 180W / 180E / 90S / 90N & 0   \\
+    Satellite & $ 10^\circ$ & 180W / 180E / 90S / 90N & 260 \\
 \end{tabular}

### 2018-12-28 manuscript/manuscript.tex
Edits for the linear density case
Moved some things down to the discussion and deleted the parts that I
moved up to the previous section.

@@
 \noindent
-has an analytical solution given by Eq.~\ref{eq:shell-pot-linear}.
-
-The absolute density difference defined on equation~\ref{eq:density-abs-diff} will
-always be zero for the linear density case.
-As a result, the inequality~\ref{eq:delta-density} will always be satisfied and no
-divisions will ever be performed.
-Therefore, the two dimensional adaptive discretization algorithm is the
-only mechanism that controls the accuracy of the numerical integration.
-For this reason, we will only determine the minimum value of $D$ needed in order to
-guarantee an acceptable accuracy while ignoring the value of $\delta$.
-
-In order to compare the numerical results with the analytical solution we must build
-spherical shell models made of tesseroids: we choose to regularly split the spherical
-shell along the latitudinal and longitudinal directions obtaining a shell model made out
-of $30^\circ \times 30^\circ$ tesseroids.
-In order to cover different computation scenarios we use five spherical shells, each one
-with the outer radius equal to the mean Earth radius and the inner radius determined by
-the desired thickness (100m, 1km, 10km, 100km and 1000km).
-These shell models are detailed on Table~\ref{tab:shell-models}.
-
-The comparison between the numerical results and the analytical solution could be
-naively compared on a single computation point due to the rotation symmetry of the
-spherical shell.
-But the numerical results may adopt different values depending on the relative location
-between the computation point and the tesseroid, thus is not wise to carry out the
-comparison on a single point.
-Instead, we propose four computation grids detailed on Table~\ref{tab:grids}: a Pole and
-an Equator grids, and a Global and Satellite grids.
-The former ones are regional scale grids located on the outer surface of the spherical
-shell, while the latter are global scale grids located at zero height (Global) and at
-the mean orbital height of the gravimetric satellite missions (Satellite).
-These grids are defined in the spirit of spanning a broad scenario of gravitational
-field computations and ensure an acceptable accuracy on each one of them.
-The comparison will be carried out by computing the maximum absolute difference between
-the analytical solution and the numerical result on each point of the computation grid.
-
-We assign a density that varies linearly with $r$ (Eq.~\ref{eq:density-linear}) to each
-shell model, which assumes the values of $\rho_\text{in} = 3300$~kg/m$^3$ and
-$\rho_\text{out} = 2670$~kg/m$^3$ on the inner and outer radii, respectively.
-So, the constants $a$ and $b$ from Eq.~\ref{eq:density-linear} must be
+has an analytical solutions for the gravitational potential and its vertical derivative
+given by Eqs.~\ref{eq:shell-pot-linear} and \ref{eq:shell-gz}.
+The values of the angular and linear coefficients ($a$ and $b$)
+can be chosen so that the density assumes the values of $\rho_\text{in} = 3300$~kg/m$^3$
+and $\rho_\text{out} = 2670$~kg/m$^3$ on the inner ($R_\text{in}$) and outer
+($R_\text{out}$) radii of the shell, respectively,
 
@@
     b = \rho_\text{out} -
-    \frac{\rho_\text{out} - \rho_\text{in}}{R_\text{out} - R_\text{in}} R_\text{out},
-\end{equation}
-
-\noindent
-where $R_\text{out}$ and $R_\text{in}$ are the outer and inner radii of the shell,
-respectively.
-Because the outer radius of every spherical shell model will be equal to the mean Earth
-radius, $R_\text{out} = 6378.137 \text{km}$.
+    \frac{\rho_\text{out} - \rho_\text{in}}{R_\text{out} - R_\text{in}} R_\text{out}.
+\end{equation}
 
@@
 
-We compute the gravitational potential ($V$) and the vertical component of the gradient
-($g_z$) that each shell model generates on each computation grid.
-Other components of the gradient are equal to zero outside of the shell due to the
-rotational symmetry and are thus omitted from the analysis.
-The computations are repeated for values of $D$ ranging from 0.5 to 5 with a step of
-0.5.
-Then we calculate the absolute difference between these results and the analytical
-solution, taking only the maximum absolute difference on each computation grid.
-This differences are shown in Fig.~\ref{fig:D-linear}.
-Finally, we set the optimal value of $D$ as the minimum value at which the corresponding
-error of the numerical approximation is lower than 0.1\%.
+The absolute density difference defined on equation~\ref{eq:density-abs-diff} will
+always be zero for the linear density case.
+As a result, the inequality~\ref{eq:delta-density} will always be satisfied and no
+divisions will ever be performed during the density-based discretization.
+Therefore, the two dimensional adaptive discretization algorithm is the
+only mechanism that controls the accuracy of the numerical integration for the linear
+density case.
+For this reason, we will ignore the values of $\delta$ and only determine the minimum
+value of $D$ needed in order to guarantee an acceptable accuracy.
+
+We compute the gravitational potential ($V$) and its vertical derivative ($g_z$) for
+each shell model in Table~\ref{tab:shell-models} on each computation grid in
+Table~\ref{tab:grids}.
+The horizontal derivatives of the potential are equal to zero outside of the shell due
+to the rotational symmetry and are thus omitted from the analysis.
+The computations are repeated for values of the distance-size ratio $D$ ranging from 0.5
+to 5 in increments of 0.5.
+We then calculate the absolute difference between the numerical results and the
+analytical solution for the shell.
+Fig.~\ref{fig:D-linear} shows the maximum absolute difference for each shell model and
+computation grid as a function of $D$.
+The differences are relative to the shell value.
+Finally, we set the optimal value of $D$ as the smallest value for which the difference
+is below 0.1\%.
 
 We observe from Fig.~\ref{fig:D-linear} that the relative errors for the
-potential and the $g_z$ component fall below the 0.1\% threshold at
-$D=1$ and $D=2.5$, respectively, for every shell model and computation grid.
-These results also show that there's no meaningful dependency between the accuracy of
-the method and the thickness of the shell or the height of the computation grid (the
-results for the Satellite grid and the Global grid are very similar).
+potential and $g_z$ fall below the 0.1\% threshold at
+$D=1$ and $D=2.5$, respectively.
+These values are consistent for all shell models and computation grids.
 
@@
 \section{Results and Discussion}
+
+% Brought this down from the results. It fits better in the discussion. I'll find a
+% place for it afterwards.
+These results also show that there's no meaningful dependency between the accuracy of
+the method and the thickness of the shell or the height of the computation grid (the
+results for the Satellite grid and the Global grid are very similar).
 
@@
     g_z(r) = \frac{V_\text{sh}(r)}{r},
+\label{eq:shell-gz}
 \end{equation}

### 2018-12-29 manuscript/manuscript.tex
Remark that D=2 would be ok for gz at 260km
Edits to exponential density case
Fixes some inconsistencies between the text and figure caption. Move
some descriptions of the different functions for varying b further down
to where they are needed.

@@
 exponential and sinusoidal density functions.
-These comparisons allowed us to obtain optimum values for the distance-size and
+These comparisons allowed us to obtain optimal values for the distance-size and
 delta ratios that yield an accuracy of 0.1\% of the analytical solutions.
@@
 
-We observe from Fig.~\ref{fig:D-linear} that the relative errors for the
-potential and $g_z$ fall below the 0.1\% threshold at
-$D=1$ and $D=2.5$, respectively.
-These values are consistent for all shell models and computation grids.
+We observe from Fig.~\ref{fig:D-linear} that the relative errors for the potential and
+$g_z$ fall below the 0.1\% threshold at $D=1$ and $D=2.5$, respectively.
+Notably, a value of $D=2$ would suffice for $g_z$ in the case of the satellite height
+grid.
+For all other configurations, these values are consistent and independent of the shell
+model thickness or geographic location.
 
@@
 This means that optimal values for both the distance-size ratio $D$ and the delta ratio
-$\delta$ must be determined.
+$\delta$ (Eq.~\ref{eq:delta-density}) must be determined.
 We perform an error analysis similar to what was done for the linear density case.
-But, now the spherical shell will have an exponential density function that assumes the
+Now the spherical shell will have an exponential density function that assumes the
 values of $\rho_\text{out} = 2670$~kg/m$^3$ and $\rho_\text{in} = 3300$~kg/m$^3$ on the
@@
 \begin{equation}
-    \rho(r') = A e^{- \frac{b}{T} (r' - R_1)} + C,
+    \rho(r') = A e^{- \frac{b}{R_2 - R_1} (r' - R_1)} + C,
     \label{eq:density-exp}
@@
 
-\noindent $T$ is the thickness of the shell ($T = R_2 - R_1$) and $b$ is a dimensionless
-constant that determines the variability of the function: a higher value of $b$
-increases the maximum slope of the density function.
-On exponential density analysis we will use values of $b$ equal to 1, 2, 5, 10, 30
-and 100, in order to cover different kind of exponential density variations, from smooth
-to sharp (Fig.~\ref{fig:exp-densities}).
+\noindent and $b$ is a dimensionless
+constant that determines the variability of the function.
+A higher value of $b$ increases the maximum slope of the density function
+(Fig.~\ref{fig:exp-densities}).
 
@@
 pair belonging to a grid on the $D$-$\delta$ space (Fig.~\ref{fig:grid-search}).
-For optimum algorithm performance, we search for the ($D$, $\delta$) pair that minimize
-the number of splittings while keeping a numerical error under the 0.1\% threshold,
-i.e.~the smallest possible value of $D$ and the highest possible value of $\delta$.
-
-We compute the relative difference between the numerical and analytical results for the
-gravitational potential ($V$) and the vertical component of its gradient ($g_z$) on the
-computation grids detailed in Table~\ref{tab:grids}.
-Because this is a time consuming computation, we limit the analysis only to a single
-value of $b$ equal to 30 (Fig.~\ref{fig:exp-densities}).
-For the sake of brevity, we compute the maximum differences between the results for each
-shell model.
-The results are shown in Fig.~\ref{fig:grid-search}, in which points inside the dotted
-lines are the ones that present a numerical error lower than the 0.1\% threshold.
+For optimum algorithm performance, we search for the ($D$, $\delta$) pair that minimizes
+the number of tesseroid division while keeping a numerical error under the 0.1\%
+threshold.
+This requirement translates into the smallest possible value of $D$ and the highest
+possible value of $\delta$.
+
+Because this is a time consuming computation, we limit the analysis to a high value of
+$b=30$ (Fig.\ref{fig:exp-densities}) and the global grid defined in
+Table~\ref{tab:grids}.
+We compute the relative absolute difference between the numerical and analytical results
+of the gravitational potential ($V$) and its vertical derivative ($g_z$) for all shell
+models defined in Table~\ref{tab:shell-models}.
+For the sake of brevity, Fig.~\ref{fig:grid-search} shows the maximum difference values
+obtained from all shell models.
+The dotted lines in Fig.~\ref{fig:grid-search} represent a contour of 0.1\% relative
+error.
+Points inside the dotted are within the acceptable threshold.
 Also shown in Fig.~\ref{fig:grid-search} are the $D$ values determined for the linear
@@
 
-The minimum values of $D$ that are within the 0.1\% threshold are the $D_\text{linear}$
-ones, i.e.~they are the optimal values of the distance-size ratios for $V$ and $g_z$.
-These results indicate that the values of $D_\text{linear}$ can be safely extrapolated
-to non linear density cases.
-This conclusion can be understood considering that the two dimensional adaptive
-discretization and the density-based algorithms are independent of each other: the first
-one splits the tesseroid only in the latitudinal and longitudinal directions, while the
-latter only does it in the radial direction.
-This leads to the fact that the value of $\delta$ needed to limit the numerical error
-due to the density variation is independent of the value of $D$ and vice versa.
-That's why the optimal distance-size ratios for the exponential density case is the same
-as the ones for the linear density case.
-
-Because the grid search was limited to specific values of $b$ and used only one grid, we
-will perform a more detailed analysis for $\delta$ determination in case of exponential
-density tesseroids.
+The smallest values of $D$ that are within the 0.1\% threshold coincide with
+$D_\text{linear}$ for both $V$ and $g_z$.
+These results indicate that $D_\text{linear}$ can be safely extrapolated to non-linear
+density cases.
+This is not surprising considering that the two dimensional adaptive
+discretization and the density-based discretization are independent of each other.
+The former divides the tesseroid in the horizontal dimensions, while the latter only
+divides in the radial dimension.
+Hence, the optimal value of $\delta$ is likely to be independent of the optimal value of
+$D$.
+Because the grid search was limited to a specific computation grid and value of $b$,
+we perform a more detailed analysis in the following section to determine an optimal
+value of $\delta$ for the exponential density case.
 
@@
 Having chosen values of $D$ equal to the ones obtained for the linear density case, we
-are free to explore the integration error as a function of $\delta$ in more detail and
-how it varies for different values of $b$ (Eq.~\ref{eq:density-exp}).
-Now we perform the error calculations for values of delta ranging from $10^{-3}$ to
-$10^{1}$, for every grid detailed in Table~\ref{tab:grids} and for the spherical shell
-models detailed in Table~\ref{tab:shell-models}.
-The calculations are repeated for distinct values of $b$ (1, 2, 5, 10, 30, 100) to
-examine the accuracy of the method for different sharpness of the density function.
+are free to explore the integration error as a function of $\delta$ in more detail.
+We will compute the difference between the numerical and analytical results for all
+combinations of computation grid (Table~\ref{tab:grids}) and spherical shell model
+(Table~\ref{tab:shell-models}), varying $\delta$ from $10^{-3}$ to $10^{1}$.
+The calculations are repeated for each $b \in \{1, 2, 5, 10, 30, 100\}$
+(Fig.~\ref{fig:exp-densities}) to examine the accuracy of the method for density
+functions of different sharpness.
 Because larger $\delta$ values result in fewer tesseroid divisions,
-our intention is to find the highest value of $\delta$ whose numerical error is bellow
-the $0.1\%$ threshold.
-
-Fig.~\ref{fig:exp-densities} shows the density functions for each value of $b$ and
+our intention is to find the highest value of $\delta$ that produces a relative
+difference bellow the $0.1\%$ threshold.
+
 Fig.~\ref{fig:delta-exponential} shows the resulting relative
-differences ($V$ and $g_z$).
-In order to improve the visualization of the results, each curve corresponds to the
-maximum difference between all shell models for the same value of $b$.
-The relative difference for $V$ and $g_z$ falls below the 0.1\% threshold
-for each computation grid, shell model and $b$ value if $\delta$ is lower than 0.1.
-This results imply that the optimal value of $\delta$ is 0.1 in case of exponential
-density.
-They also show that there is no significant relation between the $b$ value and the
-numerical error, proving that the density-based discretization algorithm is capable of
-limiting the error due to the density variation regardless the variation of the
-exponential density.
+absolute differences for $V$ and $g_z$ as a function of $\delta$.
+For the sake of brevity, each curve corresponds to the maximum difference between all
+shell models.
+The curves for $b=1$ and $b=2$ are below the 0.1\% threshold for all values of $\delta$
+and do not change for $\delta > 0.8$, indicating that the density
+functions are sufficiently smooth and no density discretizations are necessary.
+For all other values of $b$, the difference falls below the 0.1\% threshold
+if $\delta = 0.1$.
+These results indicate that there is no significant relationship between the
+sharpness of the exponential density function and the numerical error.
 
@@
 exponential density functions.
-Nevertheless, this new algorithm is well suited to accurately reproduce the
-gravitational fields of tesseroids with more complex continuous functions, like
-non-monotonic or with one or more inflection points.
-Although such density functions are very rare (or maybe even non-existent) between
-geological structures, we want to subject the density-based algorithm to such cases in
-order to show that it can solve more complex situations other than linear and
-exponential functions.
+Nevertheless, this new algorithm is well suited to more complex continuous functions,
+for example non-monotonic functions or those with multiple inflection points.
+Although such density functions are rarely, if ever, seen in geological structures,
+we want to subject our algorithm to such cases in order to show that it can solve more
+complex situations than linear and exponential functions.
 
@@
 Because there is no direct relation between the values assigned to $D$ and $\delta$ and
-the error of the computation, we empirically determined the optimum values for these
+the error of the computation, we empirically determined the optimal values for these
 parameters.
@@
 0.1\% of an analytical solution.
-The density functions used to establish the optimum values were a linear, an exponential
+The density functions used to establish the optimal values were a linear, an exponential
 and a sinusoidal function.

### 2018-12-29 manuscript/manuscript.tex
Edits to the sinusoidal case
Added some more justification of why we're doing this (to show that we
can). Added missing definition of A and its value. delta=0.1 is good for
all but the most extreme case, emphasize that. Point out that even if
delta=0.1, the error is below 1% for b=10.

@@
 
-We will consider spherical shells with a sinusoidal density defined as follows:
-
-\begin{equation}
-    \rho(r') = A \sin \left( 2 \pi b \frac{r' - R}{T} \right) + A,
+We will consider spherical shells with a sinusoidal density function defined as follows:
+
+\begin{equation}
+    \rho(r') = A \sin \left( 2 \pi b \frac{r' - R}{R_2 - R_1} \right) + A,
     \label{eq:density-sine}
@@
 
-\noindent where $A$ is a constant, $T$ is the thickness of the shell, $R$ is the mean
-Earth radius and $b$ is a dimensionless constant that regulates how many periods of the
-trigonometric function are included inside the inner and outer radii.
-For this analysis we will use values of $b$ equal to 1, 2, 5 and 10
-(Fig.~\ref{fig:sine-densities}).
-We shift the sine function in order to prevent near zero values of the potential and the
-$g_z$.
-
-We compute the relative difference between the numerical and analytical results for the
-gravitational potential ($V$) and the $g_z$ of the spherical shells detailed in
-Table~\ref{tab:shell-models} on each grid detailed in Table~\ref{tab:grids}.
-Each shell is assigned with different sinusoidal density functions by changing the
-values of $b$ in Eq.~\ref{eq:density-sine} (Fig.~\ref{fig:sine-densities}).
-The analytical solution for the potential $V$ can be found on Appendix~\ref{sec:shell}.
+\noindent in which $A$ is a constant that controls the amplitude and vertical shift of
+the sine function, $R$ is the mean Earth radius, and $b$ is a dimensionless constant
+that regulates how many periods of the trigonometric function are included inside the
+inner and outer radii.
+The analytical solutions for $V$ and $g_z$ of a spherical shell with a sinusoidal
+density function can be found on Appendix~\ref{sec:shell}.
+
+We compute the relative difference between the numerical and analytical results for $V$
+and $g_z$ for all combinations of spherical shell models (Table~\ref{tab:shell-models})
+and computation grids (Table~\ref{tab:grids}).
 We fixed the distance-size ratio $D$ to the ones obtained for the linear density case
 and explored values of $\delta$ ranging from $10^{-4}$ to $1$.
+The calculations are repeated for values of $b$ equal to 1, 2, 5 and 10
+(Fig.~\ref{fig:sine-densities}).
+For all computations, the value of $A$ is fixed at 1650 kg/m$^3$.
 
@@
 \caption{
-    Sinusoidal density functions assigned to each spherical shell in the $\delta$ ratio
+    Sinusoidal density functions assigned to the spherical shells in the $\delta$ ratio
     determination.
@@
 
-Fig.~\ref{fig:sine-densities} shows the sinusoidal density functions assigned to every
-spherical shell and Fig.~\ref{fig:delta-sine} shows the resulting relative differences
-between the analytical and the numerical solutions.
-In order to improve the visualization of the results, each curve corresponds to the
-maximum difference between all shell models for the same value of $b$.
-For this kind of sinusoidal density, a value of $\delta=0.01$ is needed in order to
-achieve the 0.1\% accuracy threshold for both the potential and the $g_{z}$, every
-shell model, computation grid and value of $b$.
-Even in case of a continuous density function with several local maximums and minimums
-and inflection points, the GLQ along with the two discretization algorithms reproduce
-the gravitational fields with an acceptable degree of accuracy.
+Fig.~\ref{fig:delta-sine} shows the relative absolute differences between the analytical
+and the numerical solutions for sinusoidal density case.
+Once again, each curve corresponds to the maximum difference between all shell models.
+For all values of $b$ except for $b=10$,
+the differences fall below the 0.1\% threshold for $\delta = 0.1$.
+In the case of $b=10$, a lower value of $\delta = 0.01$ is required to achieve 0.1\%
+difference.
+We note that, even for the case of $b=10$, the differences for $\delta = 0.1$ are below
+1\%.
 

### 2018-12-29 manuscript/manuscript.tex
Edit the computational performance section
Rename it to Algorithm Performance. Not much is changed here, just some
re-wording.

@@
 
-\section{Computation time comparison}
-
-Because the density-based algorithm produces discretizations along the radial direction,
-besides the latitudinal and longitudinal ones generated by the two dimensional adaptive
-one, the computation time of the gravitational fields of variable density tesseroids
-will be increased in comparison to the homogeneous density case.
-In order to quantize this increase we perform a computation time comparison between the
-homogeneous density case with the linear and the exponential ones.
-
-We define a $20^\circ \times 20^\circ$ tesseroid of 1km of thickness and a single
-computation point located at different heights (1m, 10m, 100m, 1km, 10km, 100km and
-1000km) above the top surface of the tesseroid.
-Then we compute the gravitational potential $V$ and the $g_z$ that the tesseroid
-generates on the computation point in case of an homogeneous density, a linear density
-and an exponential density with $b = 3$ (Eq.~\ref{eq:density-exp}).
-We perform 1000 computations for each field, density and computation height while
-tracking the computation time for each one of them.
-The final computation time is obtained as the mean value of these runs.
-
-Fig.~\ref{fig:computation-times} shows the computation time ratio between the
-computations times of the linear and exponential density with the one for an homogeneous
-density, defined as:
-
-\begin{equation}
-    \text{Computation time ratio} = \frac{\Delta t}{\Delta t_\text{homog}},
-\end{equation}
-
-\noindent where $\Delta t_\text{homog}$ is the computation time for the homogeneous
-density and $\Delta t$ is the computation time for the linear or the exponential
-density.
+\section{Algorithm performance}
+
+Because the density-based discretization algorithm introduces more division along the
+radial dimension, it is reasonable that the computation time for variable densities will
+be increased in comparison to the homogeneous density case.
+In addition to the extra division, there is also the overhead imposed by the extra
+function calls to evaluate the density function.
+In order to quantify these effects, compare the computation times between the
+homogeneous density case and the linear and the exponential density ones.
+We note that these results are highly dependent on our particular implementation of the
+algorithm and choice of programming language.
+
+We define a $20^\circ \times 20^\circ$ tesseroid of 1 km of thickness and a single
+computation point located at different heights above the top surface of the tesseroid (1
+m, 10 m, 100 m, 1 km, 10 km, 100 km and 1000 km).
+We then compute $V$ and $g_z$ for the homogeneous, linear, and exponential (with $b =
+3$) density cases.
+We perform 1000 computations for each field, density function, and height while tracking
+their computation time.
+The final computation time is the mean of these 1000 executions.
+
+Fig.~\ref{fig:computation-times} shows the computation times relative to the homogeneous
+case (i.e., ${\Delta t}/{\Delta t_\text{homog}}$).
 

### 2018-12-29 manuscript/manuscript.tex
Edit Neuquen basin application
Not much to add here. Only commented about the magnitude of the
differences and how they are larger than the accuracy of most gravity
data.

@@
 We applied the new algorithms and optimal values of $D$ and $\delta$ determined
-previously to calculate the gravitational effects of the Neuqu\'en Basin:
-a sedimentary basin
-located to the east of the Andes, between 32$^\circ$S and 40$^\circ$S latitude
-(Fig.~\ref{fig:neuquen-basin}a).
+previously to calculate the gravitational effects of the Neuqu\'en Basin,
+a sedimentary basin located to the east of the Andes, between 32$^\circ$S and
+40$^\circ$S latitude (Fig.~\ref{fig:neuquen-basin}a).
 The basin includes continental and marine siliciclastics, carbonates, and evaporites
@@
 to 5000m of depth \citep{Howell2005}.
-
-\begin{figure}
-\centering
-\iftwocol{
-\includegraphics[width=\linewidth]{figures/neuquen-basin-densities.pdf}
-}{
-\includegraphics[width=0.5\linewidth]{figures/neuquen-basin-densities.pdf}
-}
-\caption{
-    Linear and exponential densities used to compute the gravitational fields generated
-    by a tesseroid model of the Neuqu\'en sedimentary basin.
-    The height is defined above the mean Earth radius, and its axis is spanned between
-    the deepest and the highest point of the basin.
-}
-\label{fig:neuquen-basin-densities}
-\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\iftwocol{
+\includegraphics[width=\linewidth]{figures/neuquen-basin-densities.pdf}
+}{
+\includegraphics[width=0.5\linewidth]{figures/neuquen-basin-densities.pdf}
+}
+\caption{
+    Linear and exponential densities used to compute the gravitational fields generated
+    by a tesseroid model of the Neuqu\'en sedimentary basin.
+    The height is defined above the mean Earth radius, and its axis is spanned between
+    the deepest and the highest point of the basin.
+}
+\label{fig:neuquen-basin-densities}
+\end{figure}
+
 \begin{figure*}
@@
 gradient ($g_z$) on a computation grid of $159\times163$ nodes ($0.05^\circ$ spacing on
-both longitude and latitude) at a 10km height over the mean Earth radius.
+both longitude and latitude) at a 10 km height over the mean Earth radius.
 The resulting fields can be seen in Fig.~\ref{fig:neuquen-basin}c-d.
-Moreover, we computed the differences between these fields and the ones generated by the
-same model but now with a constant density as the mean value between -412kg/m$^3$ and
--275kg/m$^3$, and a linear density that assumes these two values on the highest and the
-deepest point of the basin (Fig.~\ref{fig:neuquen-basin-densities}), respectively.
+
+We computed the differences between the results for the exponential density function and
+those generated by the same model but now with a constant density (the mean of
+-412kg/m$^3$ and -275kg/m$^3$).
+We also computed the differences between the exponential density and a linear density
+that assumes these two values on the highest and the deepest point of the basin
+(Fig.~\ref{fig:neuquen-basin-densities}).
 Fig.~\ref{fig:neuquen-basin-diffs}a-b and Fig.~\ref{fig:neuquen-basin-diffs}c-d show the
 differences with the constant and the linear density, respectively.
+The maximum absolute difference in the computed $g_z$ is approximately 8 mGal for the
+homogeneous case and approximately 6 mGal for the linear density case.
+Both are well above the accuracy of data derived from spherical harmonic models, which
+is often used for regional interpretations.
 

### 2018-12-29 manuscript/manuscript.tex
Edit the discussion
Rename to just Discussion. Remove some of the things repeated in the
intro about Lin2018 and Fukushima2018. I kept the part that mentions
that we could use the density discretization combined with piecewise
linear or polynomial functions (citing both papers).
Added some parts about why our method is independent of the density
function. Will reiterate this in the conclusions as well.
Argue why we can probably extrapolate the results here to any realistic
density function.

@@
 homogeneous case and approximately 6 mGal for the linear density case.
-Both are well above the accuracy of data derived from spherical harmonic models, which
-is often used for regional interpretations.
+Both are well above the accuracy of most available data products.
 
@@
 
-\section{Results and Discussion}
-
-% Brought this down from the results. It fits better in the discussion. I'll find a
-% place for it afterwards.
-These results also show that there's no meaningful dependency between the accuracy of
-the method and the thickness of the shell or the height of the computation grid (the
-results for the Satellite grid and the Global grid are very similar).
-
-The $D$ and $\delta$ ratios determinations proved that for a linear density case only
-the two dimensional adaptive discretization algorithm and a second-order GLQ
-approximation is needed to achieve a 0.1\% accuracy of the gravitational potential and
-the vertical component of the gradient ($g_z$) by using a $D=$ 1 and 2.5 for each field,
-respectively.
-Although for non-linear functions the adaptive discretization is not sufficient and the
-density-based discretization algorithm must be applied.
-The $D$-$\delta$ space exploration showed that the more efficient computation takes
-place when the value of $D$ equals to the ones determined in the linear density
-scenario, while the value of $\delta$ must be determined for each specific density
-variation.
-In case of an exponential density, a $\delta$ equal to 0.1 is sufficient to achieve the
-0.1\% accuracy threshold, while a $\delta = 0.01$ is needed for a sinusoidal density
-function.
-
-This result could be extrapolated: a default value of $\delta = 0.01$ would guarantee an
-accuracy of 0.1\% in case of an arbitrary continuous density function.
-Nevertheless, we encourage to perform similar accuracy tests in order to proof that this
-extrapolation works for the desired density function, specially for more complex
-variations than a polynomial or exponential one.
-
-\citet{Lin2018} proposed a combined method to efficiently compute gravitational fields
-generated by tesseroids: it uses a GLQ with a two dimensional adaptive discretization
-algorithm for near computation points and a Taylor series expansion method for far
-computation points.
-Because our goal is to build a method that could be applied to any continuous density
-function, the Taylor series wouldn't be well suited for this task and therefore we
-focused only on a pure GLQ approximation method.
-Nevertheless, our method could be easily combined with a Taylor series expansion for
-specific density function, i.e.~a linear variation.
-
-Moreover, \citet{Lin2018} show that the combined method can be applied to a linear
-density variation by splitting the gravitational field integral in two: one for each
-term of the linear function.
-Our results show that in case of a linear density variation a single GLQ computation is
-sufficient to accurately compute the fields, thus reducing the computation time by half.
-
-\citet{Fukushima2018} propose a generalization of his method to polynomial density
-functions.
-On the other hand, our method is not bounded to a specific density function.
-Nevertheless, the density-based discretization algorithm could be used before applying
-the numerical integration developed by \citet{Fukushima2018} assigning a low polynomial
-order density to each discretizing tesseroid.
-Although this method would involve a computational error due to the approximation of the
-density function as a polynomial one, its feasibly if we account for this error and
-a method to limit it.
-
-We have also carried out $D$ and $\delta$ determinations for the vertical component of
-the Marussi tensor ($g_{zz}$) for linear and exponential densities.
-We didn't include these results because our methodology wouldn't achieve the 0.1\%
-accuracy threshold for the general case of such density functions.
-Moreover, the values of $D$ and $\delta$ needed for the cases at which the accuracy was
-below 0.1\% implied a high number of discretizations, thus greatly increasing the
-computation time.
-In summary, we found the computation of the Marussi tensor components in case of
-variable density tesseroids a subject for future research.
+\section{Discussion}
+
+By including the density function into the GLQ integration, the method described here
+can be applied without modification to tesseroids with density following any continuous
+function of radius.
+A density-based discretization algorithm divides the tesseroid in the radial dimension
+to ensure accurate integration of the density function.
+This algorithm is independent of the GLQ integration and could potentially be used to
+determine an optimal discretization when approximating a density function by piecewise
+linear \citep{Lin2018} or piecewise polynomial \citep{Fukushima2018} functions.
+
+Our numerical experiments show that the two dimensional adaptive discretization is
+enough to achieve 0.1\% accuracy with a second-order GLQ in the case of a linear density
+function (Fig.~\ref{fig:D-linear}).
+Values of the distance-size ratio determined here for the gravitational potential
+($D=1$) and its vertical derivative ($D=2.5$) are compatible with \citet{Uieda2016}.
+These results also show that there is no significant relationship between the accuracy
+of the method and the thickness of the tesseroid model.
+
+The $D$-$\delta$ space exploration for the exponential density function
+(Fig.~\ref{fig:grid-search}) showed that values of $D$ determined for the linear density
+case are equally applicable to the exponential density case.
+Furthermore, the difference with respect to the analytical solution only falls below
+0.1\% for values of $\delta$ below 0.1.
+It follows that the density-based discretization is required to achieve the desired
+accuracy level for non-linear density functions.
+
+More detailed error analyses showed that $\delta = 0.1$ is sufficient to guarantee an
+accuracy of 0.1\% for all exponential density functions tested
+(Fig.~\ref{fig:delta-exponential})
+and most of the sinusoidal functions tested (Fig.~\ref{fig:delta-sine}).
+The exception is the case of a sinusoidal function with $b = 10$
+(Eq.~\ref{eq:density-sine}), for which $\delta = 0.01$ is required to achieve 0.1\%
+accuracy.
+Nevertheless, using $\delta = 0.1$ in this case would still produce results with an
+error of less than 1\%.
+
+While the algorithms proposed here can be applied to any continuous density function,
+the optimal values for $D$ and $\delta$ are empirically determined only for linear,
+exponential, and sinusoidal density functions.
+Thus, we can only say with certainty that these values will produce results with 0.1\%
+accuracy for these functions.
+However, all of our numerical tests include worse-case scenarios (zero computation
+heights, large tesseroids, highly variable density functions, etc).
+For this reason, these optimal values of $D$ and $\delta$ can plausibly be
+extrapolated to any realistic continuous density function.
+Notwithstanding, we encourage users of the algorithms and software to perform similar
+tests in order evaluate the accuracy when using density functions more complex than the
+ones tested here.
 
@@
 
+% Moved here from the discussion. The conclusions are the place to talk about future
+% work and speculations.
+We have also carried out $D$ and $\delta$ determinations for the vertical component of
+the Marussi tensor ($g_{zz}$) for linear and exponential densities.
+We didn't include these results because our methodology wouldn't achieve the 0.1\%
+accuracy threshold for the general case of such density functions.
+Moreover, the values of $D$ and $\delta$ needed for the cases at which the accuracy was
+below 0.1\% implied a high number of discretizations, thus greatly increasing the
+computation time.
+In summary, we found the computation of the Marussi tensor components in case of
+variable density tesseroids a subject for future research.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-12-29 manuscript/manuscript.tex
Edit the conclusions
Emphasize that the method is general purpose. We sacrifice speed for
usability and accuracy.
Added some more speculation about using the density-based discretization
with other methods.
Talk a bit about the Neuquen application and how the errors we see there
would translate into bad gravity inversions.

I removed the Gzz part because it felt very out of place. We don't talk
about it anywhere and have nothing to show. The only people who would
ask are the reviewers so we should include that in our response instead.

@@
 These shortcomings could make piecewise linear or high-degree polynomial density
-functions unwieldy for non-linear gravity inversions if the density function in depth
-is highly variable.
+functions unwieldy for non-linear gravity inversions \citep[e.g.][]{Uieda2017} if the
+density function in depth is highly variable.
 
@@
 
-We have developed a new methodology to compute the gravitational fields
-generated by a tesseroid with a density given by a continuous function of depth.
-It numerically solves the integrals that define the gravitational potential and
-its gradient through the Gauss-Legendre Quadrature (GLQ).
+We have developed a new methodology to compute the gravitational potential and its
+gradient for a tesseroid with a density given by a continuous function of radius.
+It numerically solves the volume integrals through Gauss-Legendre Quadrature (GLQ) by
+including the density function in the numerical integration.
+By implementing the algorithm in the dynamic Python programming language, users can
+define their own density function to supply to the software.
+This allows the use of any arbitrary continuous function without modification to the
+method or software.
 The accuracy of the numerical integration is automatically controlled by a two
@@
 algorithm.
-The former divides the tesseroid in half if the ratio of the distance to the computation
-point and the latitudinal or the longitudinal size of the tesseroid is lower than
-a predefined distance-size ratio $D$.
+The former divides the tesseroid in half if the ratios of the distance to the
+computation point and the latitudinal or the longitudinal size of the tesseroid are
+smaller than a predefined distance-size ratio $D$.
 This algorithm minimises the integration error when the computation point is close to
@@
 Nevertheless, the adaptive discretization alone is not sufficient to guarantee the
-accuracy of the method in case of tesseroids with variable density.
-
-To overcome this challenge, we have developed a density-based discretization algorithm
-that divides the tesseroid on the points at which the maximum variations of the density
-function take place.
-The density-based discretization is performed before adaptive discretization as a type
-of pre-processing step.
-The number of divisions performed, and thus the accuracy of the computation, is
-controlled by the delta ratio $\delta$.
+accuracy of the method in case of tesseroids with non-linear density functions.
+To overcome this challenge, the density-based discretization algorithm
+divides the tesseroid along the radial dimension on the points at which the maximum
+variations of the density function take place.
+The number of radial divisions performed, and thus the accuracy of the computation, is
+controlled by the $\delta$ parameter.
 This new algorithm is intended to minimise the error due to the inability of
@@
 
-Because there is no direct relation between the values assigned to $D$ and $\delta$ and
-the error of the computation, we empirically determined the optimal values for these
-parameters.
+We empirically determined optimal values for $D$ and $\delta$ using common model and
+computation grid configurations and density functions.
 These values minimize the computational load while maintaining the numerical error below
 0.1\% of an analytical solution.
-The density functions used to establish the optimal values were a linear, an exponential
-and a sinusoidal function.
+The density functions used to establish the optimal values were a linear, an
+exponential, and a sinusoidal function.
 The linear function represents the smoothest variation of density and does not require
 density-based discretization at all.
-We analysed the error for exponential functions ranging from smooth to sharp variations
-in depth to test the accuracy of the density-based discretization.
-And the error for sinusoidal functions with different wavelengths.
-Through this analysis, we have obtained optimal values for the distance-size ratio of
-1 and 2.5 for the potential an its gradient, respectively.
-While values of $\delta$ equal to 0.1 and 0.01 are needed in order to guarantee a 0.1\%
-accuracy for an exponential and a sinusoidal density, respectively.
-These results could be extrapolated: a $\delta$ value equal to 0.01 is needed in order
-to achieve the accuracy threshold in case of an arbitrary continuous density function.
-Although we encourage to perform similar accuracy tests for functions other than the
-ones we studied, this results could be safely generalized to other common depth-density
-relations among geophysical applications, like quadratic functions, without loss of
-accuracy.
-
-% Moved here from the discussion. The conclusions are the place to talk about future
-% work and speculations.
-We have also carried out $D$ and $\delta$ determinations for the vertical component of
-the Marussi tensor ($g_{zz}$) for linear and exponential densities.
-We didn't include these results because our methodology wouldn't achieve the 0.1\%
-accuracy threshold for the general case of such density functions.
-Moreover, the values of $D$ and $\delta$ needed for the cases at which the accuracy was
-below 0.1\% implied a high number of discretizations, thus greatly increasing the
-computation time.
-In summary, we found the computation of the Marussi tensor components in case of
-variable density tesseroids a subject for future research.
+Through this analysis, we have obtained optimal values for the distance-size ratio $D$
+of 1 and 2.5 for the potential an its gradient, respectively.
+We analysed the error for exponential functions ranging from smooth to sharp and
+sinusoidal function of different wavelengths to test the accuracy of the density-based
+discretization.
+Values of $\delta$ equal to 0.1 are sufficient to guarantee a 0.1\% accuracy for most
+realist cases tested.
+These results could be extrapolated to other continuous density functions that are
+sufficiently smooth, as is the case for most common geophysical applications.
+Smaller values of $\delta$ can used for more highly variable density functions to
+increase the accuracy of the integration accordingly.
+
+Some computational performance is sacrificed in order to obtain an accurate and general
+purpose method.
+The density-based discretization adds to the computation time by increasing
+the number of GLQ integrations.
+Likewise, allowing users to supply custom density functions incurs an overhead cost
+because of the added function evaluations.
+It is also impossible to optimize the source code and formulas without knowledge of the
+specific density function.
+Nonetheless, the density-based discretization is independent of the adaptive
+discretization and GLQ integration.
+It can be viewed as a type of pre-processing step that can be combined with other more
+specialized integration methods.
+
+An application to modeling the Neuqu\'en Basin, Argentina, demonstrates that the effects
+of sediment compaction cannot be ignored.
+When compared to an exponential density function, a homogeneous density with depth would
+result in an error of up to -8 mGal, while a linear density function would result in up
+to -6 mGal error.
+Accurate and robust forward modeling is a key component to any gravity inversion method.
+Errors of this magnitude could result in significant overestimation of sedimentary
+basin thickness, for example.
 

### 2018-12-29 manuscript/manuscript.tex
Move a sentence in the methods
Add missing ~ before reference to equation

@@
 geocentric latitude $\phi$, and longitude $\lambda$.
+\citet{Grombein2013} provide efficient formulations for the volume integrals of the
+gravitational potential and its first- and second-order derivatives of a tesseroid with
+homogeneous density.
 The derivatives of the gravitational potential are taken with
@@
 (Fig.~\ref{fig:tesseroid}).
-\citet{Grombein2013} provide efficient formulations for the volume integrals of the
-gravitational potential and its first- and second-order derivatives of a tesseroid with
-homogeneous density.
 Here, we will assume that the tesseroid has a density varying with $r$ according to an
@@
 has an analytical solutions for the gravitational potential and its vertical derivative
-given by Eqs.~\ref{eq:shell-pot-linear} and \ref{eq:shell-gz}.
+given by Eqs.~\ref{eq:shell-pot-linear} and~\ref{eq:shell-gz}.
 The values of the angular and linear coefficients ($a$ and $b$)

### 2018-12-29 code/scripts/number-of-tesseroids.py
Add number-of-tesseroids.py script

@@ -1,0 +1,95 @@
+"""
+We will compute the number of discretization that the density-based algorithm performs
+in case of an exponential density and a delta value of 0.1.
+We will explore different values of the constant b to see how it affects the
+computational load.
+"""
+from __future__ import division
+import os
+import warnings
+import numpy as np
+import matplotlib.pyplot as plt
+
+# This is our custom tesseroid code
+from tesseroid_density import tesseroid
+
+
+# Define Tesseroid boundaries
+# ---------------------------
+w, e, s, n = -10, 10, -10, 10
+top, bottom = 0, -1e3
+
+
+# Compute number of discretized tesseroids (exp)
+# ----------------------------------------------
+delta = 0.1
+density_top, density_bottom = 0, 1
+b_factors = np.array([1, 2, 5, 10, 30, 100])
+n_tess = np.zeros_like(b_factors, dtype=np.int64)
+
+fig, ax = plt.subplots()
+heights = np.linspace(bottom, top, 101)
+for i, b_factor in enumerate(b_factors):
+    thickness = top - bottom
+    A = (density_bottom - density_top) / (1 - np.exp(-b_factor))
+    C = density_bottom - A
+
+    # Define density function
+    def density_exponential(height):
+        return A * np.exp(-b_factor * (height - bottom) / thickness) + C
+
+    # Calculate number of discretizations
+    subset = tesseroid._density_based_discretization([w, e, s, n, top, bottom],
+                                                     density_exponential,
+                                                     delta)
+    n_tess[i] = len(subset)
+
+    # Plot density profile and boundaries of discretized tesseroids
+    bounds = []
+    for tess in subset:
+        bounds.append(tess[-1])
+        bounds.append(tess[-2])
+    bounds = np.array(bounds)
+    ax.plot(heights, density_exponential(heights), '-', color="C0")
+    ax.plot(bounds, density_exponential(bounds), 'o', color="C1")
+
+plt.show()
+
+
+# Compute number of discretized tesseroids (sine)
+# -----------------------------------------------
+delta = 0.1
+b_factors = [1, 2, 5, 10]
+n_tess = np.zeros_like(b_factors, dtype=np.int64)
+
+fig, axes = plt.subplots(nrows=len(b_factors), sharex=True)
+heights = np.linspace(bottom, top, 101)
+for i, b_factor in enumerate(b_factors):
+    max_density = density_bottom
+    thickness = top - bottom
+    k_factor = 2 * np.pi * b_factor / thickness
+
[…]

### 2018-12-29 manuscript/manuscript.tex
Rewrite exponential density equation
Replace "below" for "lower than"
Add missing space before equation reference

@@
 Apply a second-order GLQ to numerically compute the gravitational fields
-(Eq.\ref{eq:glq-var-dens}) generated by each tesseroid obtained in the
+(Eq.~\ref{eq:glq-var-dens}) generated by each tesseroid obtained in the
 previous step.
@@
 \begin{equation}
-    \rho(r') = A e^{- \frac{b}{R_2 - R_1} (r' - R_1)} + C,
+    \rho(r') = A e^{- b \frac{r' - R_1}{R_2 - R_1}} + C,
     \label{eq:density-exp}
@@
 Furthermore, the difference with respect to the analytical solution only falls below
-0.1\% for values of $\delta$ below 0.1.
+0.1\% for values of $\delta$ lower than 0.1.
 It follows that the density-based discretization is required to achieve the desired

### 2018-12-29 manuscript/manuscript.tex
Remove "relative absolute differences"

@@
 Table~\ref{tab:grids}.
-We compute the relative absolute difference between the numerical and analytical results
+We compute the relative difference between the numerical and analytical results
 of the gravitational potential ($V$) and its vertical derivative ($g_z$) for all shell
@@
 Fig.~\ref{fig:delta-exponential} shows the resulting relative
-absolute differences for $V$ and $g_z$ as a function of $\delta$.
+differences for $V$ and $g_z$ as a function of $\delta$.
 For the sake of brevity, each curve corresponds to the maximum difference between all
@@
 
-Fig.~\ref{fig:delta-sine} shows the relative absolute differences between the analytical
+Fig.~\ref{fig:delta-sine} shows the relative differences between the analytical
 and the numerical solutions for sinusoidal density case.

### 2018-12-29 manuscript/manuscript.tex
Less "ands"

@@
 
-We empirically determined optimal values for $D$ and $\delta$ using common model and
-computation grid configurations and density functions.
+We empirically determined optimal values for $D$ and $\delta$ by comparing the numerical
+results with analytical solutions.
+Our analysis included a range of tesseroid models and computation grids as well as
+common density functions.
 These values minimize the computational load while maintaining the numerical error below

### 2018-12-29 manuscript/manuscript.tex
Edits to the abstract

@@
 density and compared them to the results of the numerical model for linear,
-exponential and sinusoidal density functions.
+exponential, and sinusoidal density functions.
 These comparisons allowed us to obtain optimal values for the distance-size and
@@
 The density-based discretization algorithm produces no discretizations in the linear
-density case, but a delta ratio of 0.1 and 0.01 is needed for the exponential and the
-sinusoidal density function, respectively.
+density case, but a delta ratio of 0.1 is needed for the exponential and the
+sinusoidal density functions.
+These values can be extrapolated to cover most common use cases.
+However, the distance-size and delta ratios can be configured by the user to increase
+the accuracy of the results at the expense of computational speed.
 Lastly, we apply this new methodology to model the Neuqu\'en Basin, a foreland basin in

### 2018-12-29 manuscript/manuscript.tex

@@
     Differences between the gravitational fields generated by each tesseroid shell model
-    and the analytical solution.
+    and the analytical solution as a function of the distance-size ratio $D$.
     Each model has a linear density function (Eq.~\ref{eq:density-linear}).
@@
     and using the shell models detailed in Table~\ref{tab:shell-models}.
-    Each line represents the maximum absolute difference between every shell model.
+    Each line represents the maximum absolute difference between the numerical results
+    and the analytical solution for a given shell model.
     Due to the linearity of the density function, the density-based discretization
     algorithm is not applied.
-    Differences are reported as a percentage of the shell values.
+    Differences are reported as a percentage of the analytical solutions.
     The horizontal dashed black line represents a target difference of 0.1\%.
@@
 \caption{
-    Exponential density functions assigned to each spherical shell in the $D$ and
-    $\delta$ ratios determination.
+    Exponential density functions assigned to the spherical shell models for
+    $\delta$ ratio determination.
     Each density function corresponds to a different value of $b$ on
@@
 \caption{
-    Numerical error for different exponential density functions and values of the delta
-    ratio.
-    The percentage difference values were obtained from the comparison between the
-    analytical solution and the numerical approximation of the gravitational fields ($V$
-    and $g_z$) generated by the spherical shell with an exponential density function and
-    different values of $b$ (Eq.~\ref{eq:density-exp}).
-    These comparisons were carried out on every grid defined on Table~\ref{tab:grids},
-    making use of the shell models detailed on Table~\ref{tab:shell-models} and
-    fixing the distance-size ratio $D$ to the values obtained for the linear density
-    case.
-    Each curve corresponds to the maximum difference between every shell model with the
-    same value of $b$.
+    Difference between the numerical and analytical solutions as a function of $\delta$
+    ratio for different exponential density functions.
+    These comparisons were carried out for every shell model
+    (Table~\ref{tab:shell-models}) and computation grid (Table~\ref{tab:grids})
+    using a fixed value of $D$.
+    Each curve corresponds to the maximum difference out of all shell models
+    for a particular value of $b$ (Eq.~\ref{eq:density-exp}).
+    Differences are reported as a percentage of the analytical solutions.
+    The horizontal dashed black line represents a target difference of 0.1\%.
     }
@@
 \caption{
-    Numerical error for different sinusoidal density functions and values of the delta
-    ratio.
-    The percentage difference values were obtained from the comparison between the
-    analytical solution and the numerical approximation of the gravitational fields ($V$
-    and $g_z$) generated by the spherical shell with an exponential density function and
-    different values of $b$ (Eq.~\ref{eq:density-sine}).
-    These comparisons were carried out on every grid defined on Table~\ref{tab:grids},
-    making use of the shell models detailed on Table~\ref{tab:shell-models} and
-    fixing the distance-size ratio $D$ to the values obtained for the linear density
-    case.
-    Each curve corresponds to the maximum difference between every shell model with the
-    same value of $b$.
+    Difference between the numerical and analytical solutions as a function of $\delta$
+    ratio for different sinusoidal density functions.
+    These comparisons were carried out for every shell model
+    (Table~\ref{tab:shell-models}) and computation grid (Table~\ref{tab:grids})
+    using a fixed value of $D$.
+    Each curve corresponds to the maximum difference out of all shell models
+    for a particular value of $b$ (Eq.~\ref{eq:density-sine}).
+    Differences are reported as a percentage of the analytical solutions.
+    The horizontal dashed black line represents a target difference of 0.1\%.
     }

### 2018-12-29 code/scripts/number-of-tesseroids.py
Add number-of-tesseroids figure
The plots were included into one single figure and formated
for publication.

@@ -9,9 +9,30 @@ import os
 import warnings
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec, GridSpecFromSubplotSpec
 
 # This is our custom tesseroid code
 from tesseroid_density import tesseroid
+
+
+# Check for DISPLAY variable for matplotlib
+# -----------------------------------------
+try:
+    os.environ["DISPLAY"]
+except Exception:
+    plt.switch_backend('agg')
+
+
+# Configure LaTeX style for plots
+# -------------------------------
+try:
+    plt.rcParams['text.usetex'] = True
+    plt.rcParams['font.family'] = 'serif'
+    plt.rcParams['font.serif'] = 'Computer Modern Roman'
+    plt.rcParams['xtick.major.size'] = 2
+    plt.rcParams['ytick.major.size'] = 2
+except Exception as e:
+    warnings.warn("Couldn't configure LaTeX style for plots:" + str(e))
 
 
 # Define Tesseroid boundaries
@@ -20,15 +41,29 @@ w, e, s, n = -10, 10, -10, 10
 top, bottom = 0, -1e3
 
 
+# Initialize figure and GridSpec
+# ------------------------------
+fig = plt.figure(figsize=(6.66, 2.5))
+outer_grid = GridSpec(ncols=3, nrows=1)
+script_path = os.path.dirname(os.path.abspath(__file__))
+figure_fname = os.path.join(script_path,
+                            "../../manuscript/figures/number-of-tesseroids.pdf")
+
+
 # Compute number of discretized tesseroids (exp)
 # ----------------------------------------------
+# Define denstiy related parameters
 delta = 0.1
 density_top, density_bottom = 0, 1
 b_factors = np.array([1, 2, 5, 10, 30, 100])
 n_tess = np.zeros_like(b_factors, dtype=np.int64)
 
-fig, ax = plt.subplots()
+# Define axe for this plot
+ax = plt.subplot(outer_grid[0])
+
+# Compute number of discretizations
 heights = np.linspace(bottom, top, 101)
+colors = dict(zip(b_factors, plt.cm.viridis(np.linspace(0, 0.9, len(b_factors)))))
 for i, b_factor in enumerate(b_factors):
     thickness = top - bottom
     A = (density_bottom - density_top) / (1 - np.exp(-b_factor))
[…]

### 2018-12-29 code/scripts/number-of-tesseroids.py
Add enumeration to number-of-tesseroid axes

@@ -43,7 +43,7 @@ top, bottom = 0, -1e3
 
 # Initialize figure and GridSpec
 # ------------------------------
-fig = plt.figure(figsize=(6.66, 2.5))
+fig = plt.figure(figsize=(6.66, 2.8))
 outer_grid = GridSpec(ncols=3, nrows=1)
 script_path = os.path.dirname(os.path.abspath(__file__))
 figure_fname = os.path.join(script_path,
@@ -94,6 +94,7 @@ for i, b_factor in enumerate(b_factors):
     ax.set_xticklabels([r"$r_1$", r"$r_2$"])
     ax.set_ylabel("Density")
     ax.set_xlabel("Depth")
+    ax.set_title("(a)")
 
 
 # Compute number of discretized tesseroids (sine)
@@ -143,6 +144,8 @@ for i, b_factor in enumerate(b_factors):
     axes[i].set_xticks([bottom, top])
     axes[i].set_xticklabels([r"$r_1$", r"$r_2$"])
     axes[i].set_xlabel("Depth")
+    if i == 0:
+        axes[i].set_title("(b)")
 
 
 # Plot number of tesseroids (sine)
@@ -157,6 +160,7 @@ ax.set_yticks(np.arange(3, 21, 2))
 ax.set_xticks(np.arange(1, 11, 2))
 ax.set_xlabel(r"$b$")
 ax.set_ylabel("Number of tesseroids")
+ax.set_title("(c)")
 
 outer_grid.tight_layout(fig)
 plt.savefig(figure_fname, dpi=300)

### 2018-12-29 manuscript/manuscript.tex
Add number-of-tesseroid figure and modify text

@@
 be increased in comparison to the homogeneous density case.
-In addition to the extra division, there is also the overhead imposed by the extra
-function calls to evaluate the density function.
-In order to quantify these effects, compare the computation times between the
-homogeneous density case and the linear and the exponential density ones.
-We note that these results are highly dependent on our particular implementation of the
-algorithm and choice of programming language.
-
-We define a $20^\circ \times 20^\circ$ tesseroid of 1 km of thickness and a single
-computation point located at different heights above the top surface of the tesseroid (1
-m, 10 m, 100 m, 1 km, 10 km, 100 km and 1000 km).
-We then compute $V$ and $g_z$ for the homogeneous, linear, and exponential (with $b =
-3$) density cases.
-We perform 1000 computations for each field, density function, and height while tracking
-their computation time.
-The final computation time is the mean of these 1000 executions.
-
-Fig.~\ref{fig:computation-times} shows the computation times relative to the homogeneous
-case (i.e., ${\Delta t}/{\Delta t_\text{homog}}$).
-
-\begin{figure}
+In order to get a qualitative indicator of this increase in the computation time we
+gathered the number of subdivisions the density-based discretization performs under
+exponential and sinusoidal density cases.
+
+We assign an exponential density (Eq.~\ref{eq:density-exp}) with values of $b$ equal to
+1, 2, 5, 10, 30 and 100, and a sinusoidal density (Eq.~\ref{eq:density-sine}) with
+values of $b$ equal to 1, 2, 5, and 10 to an arbitrary tesseroid.
+Then we applied the density-based discretization algorithm on each case recording how
+many subdivisions it carried out and where the boundaries of the smaller tesseroids are
+located.
+On both exponential and sinusoidal density cases a value of $\delta$ equal to 0.1 has
+been used.
+
+Figs.~\ref{fig:number-of-tesseroids}(a) and~\ref{fig:number-of-tesseroids}(b) show the
+exponential and sinusoidal densities used and the resulting boundaries of the
+discretized tesseroids as orange dots.
+The number of discretizations are independent of the value of factor $b$ in the
+exponential density case: the algorithm only performs one split independently of $b$.
+On the other hand, the sinusoidal density results show a significant relation between
+the number of discretizations and the constant $b$.
+Fig.~\ref{fig:number-of-tesseroids} shows an almost linear relation between these two
+magnitudes.
+
+Therefore, the computation time for the exponential density would be increased in
+comparison to the homogeneous density case because the
+single split the density-based algorithm performs, although it will be independent of
+the value of $b$.
+On the other hand, the computation time for the sinusoidal density is bounded to the
+value of $b$ used, greatly increasing it for a large $b$ as 100.
+
+\begin{figure*}
 \centering
-\iftwocol{
-\includegraphics[width=\linewidth]{figures/computation-times.pdf}
-}{
-\includegraphics[width=0.5\linewidth]{figures/computation-times.pdf}
-}
+\includegraphics[width=\linewidth]{figures/number-of-tesseroids.pdf}
 \caption{
-    Computation times ratio between linear and exponential density tesseroids with an
-    homogeneous one.
-}
-\label{fig:computation-times}
-\end{figure}
+    Number of discretizations performed by the density-based algorithm in case of
+    (a)~exponential densities with the same values of $b$ shown in
+    Fig.~\ref{fig:exp-densities},
+    (b)~sinusoidal densities with the same values of $b$ shown in
+    Fig.~\ref{fig:sine-densities}.
+    On both figures the boundaries of the discretized tesseroids are marked with orange
+    dots.
+    Subfigure (c) shows the number of discretized tesseroids for the sinusoidal
+    densities as a function of the factor $b$.
+}
+\label{fig:number-of-tesseroids}
+\end{figure*}
 

### 2018-12-29 manuscript/manuscript.tex
Edit algorithm performance text

@@
 located.
-On both exponential and sinusoidal density cases a value of $\delta$ equal to 0.1 has
-been used.
-
 Figs.~\ref{fig:number-of-tesseroids}(a) and~\ref{fig:number-of-tesseroids}(b) show the
@@
 discretized tesseroids as orange dots.
-The number of discretizations are independent of the value of factor $b$ in the
-exponential density case: the algorithm only performs one split independently of $b$.
-On the other hand, the sinusoidal density results show a significant relation between
-the number of discretizations and the constant $b$.
+On both exponential and sinusoidal density cases a value of $\delta$ equal to 0.1 has
+been used.
+
+On the exponential density scenario, the algorithm performs a single split independently
+of the value of $b$, i.e.~the number of discretizations is independent of the
+variability of the exponential function.
+Therefore, the computation time would be increased in comparison to the homogeneous
+density case due to this single split, but would be independent of the value of $b$.
+
+On the other hand, the sinusoidal density case shows a strong dependence of the
+number of discretizations on the value of $b$.
 Fig.~\ref{fig:number-of-tesseroids} shows an almost linear relation between these two
 magnitudes.
-
-Therefore, the computation time for the exponential density would be increased in
-comparison to the homogeneous density case because the
-single split the density-based algorithm performs, although it will be independent of
-the value of $b$.
-On the other hand, the computation time for the sinusoidal density is bounded to the
-value of $b$ used, greatly increasing it for a large $b$ as 100.
+This means that the computation time is heavily bounded to the value of $b$ used.
+It can be greatly increased for large values of $b$ such as 100.
+
+From this results we can conclude that the computation time would be greatly increased
+in case of multiple inflection points on the density function.
+Nevertheless, for most geophysical applications on which the density profiles may have
+low number of inflection points, the computation time would be in the same order of
+magnitude as the one for the homogeneous density case.
 

### 2018-12-29 manuscript/manuscript.tex
Minor edit in number of tesseroids caption

@@
 \caption{
-    Number of discretizations performed by the density-based algorithm in case of
+    Number of discretizations performed by the density-based algorithm
+    (with a $\delta = 0.1$) in case of
     (a)~exponential densities with the same values of $b$ shown in

### 2018-12-29 manuscript/manuscript.tex
Edit algorithm performance

@@
 On the exponential density scenario, the algorithm performs a single split independently
-of the value of $b$, i.e.~the number of discretizations is independent of the
-variability of the exponential function.
+of the value of $b$.
 Therefore, the computation time would be increased in comparison to the homogeneous
-density case due to this single split, but would be independent of the value of $b$.
+density case due to this single split, but would be independent of the variation of the
+function.
 
@@
 magnitude as the one for the homogeneous density case.
+
+There are some elements related to the implementation of the algorithm and the
+programming language used that may increase the computation time, e.g.~we found that
+density function callbacks on a Python implementation plays an important role on the
+computation time increase.
+For simplicity, these elements were not taken into account, and because they are
+dependent on the specific implementation of the algorithm, they cannot be generalized.
 

### 2018-12-29 manuscript/manuscript.tex
Minor edit on Algorithm performance
Review the new performance section

--- 2018-12-29 manuscript/manuscript.tex
@@
 
-Because the density-based discretization algorithm introduces more division along the
+Because the density-based discretization algorithm introduces more divisions along the
 radial dimension, it is reasonable that the computation time for variable densities will
-be increased in comparison to the homogeneous density case.
-In order to get a qualitative indicator of this increase in the computation time we
-gathered the number of subdivisions the density-based discretization performs under
-exponential and sinusoidal density cases.
-
-We assign an exponential density (Eq.~\ref{eq:density-exp}) with values of $b$ equal to
-1, 2, 5, 10, 30 and 100, and a sinusoidal density (Eq.~\ref{eq:density-sine}) with
-values of $b$ equal to 1, 2, 5, and 10 to an arbitrary tesseroid.
-Then we applied the density-based discretization algorithm on each case recording how
-many subdivisions it carried out and where the boundaries of the smaller tesseroids are
-located.
+be higher than the homogeneous density case.
+Directly comparing the computation time of both cases cannot be done in a meaningful
+way because it is highly dependent on the implementation of the algorithm,
+the choice of programming language, and the particular density function used.
+In order to obtain an indicator of the increase in the computation time,
+we chose the number of tesseroid divisions in the density-based discretization as a
+proxy measure.
+
+We analyse exponential density functions (Eq.~\ref{eq:density-exp}) with values of $b$ equal to
+1, 2, 5, 10, 30 and 100, and sinusoidal density functions (Eq.~\ref{eq:density-sine}) with
+values of $b$ equal to 1, 2, 5, and 10.
+We then apply the density-based discretization algorithm on each function and record the
+number divisions carried out.
 Figs.~\ref{fig:number-of-tesseroids}(a) and~\ref{fig:number-of-tesseroids}(b) show the
-exponential and sinusoidal densities used and the resulting boundaries of the
-discretized tesseroids as orange dots.
-On both exponential and sinusoidal density cases a value of $\delta$ equal to 0.1 has
-been used.
-
-On the exponential density scenario, the algorithm performs a single split independently
-of the value of $b$, i.e.~the number of discretizations is independent of the
-variability of the exponential function.
-Therefore, the computation time would be increased in comparison to the homogeneous
-density case due to this single split, but would be independent of the value of $b$.
-
-On the other hand, the sinusoidal density case shows a strong dependence of the
-number of discretizations on the value of $b$.
-Fig.~\ref{fig:number-of-tesseroids} shows an almost linear relation between these two
-magnitudes.
-This means that the computation time is heavily bounded to the value of $b$ used.
-It can be greatly increased for large values of $b$ such as 100.
-
-From this results we can conclude that the computation time would be greatly increased
-in case of multiple inflection points on the density function.
-Nevertheless, for most geophysical applications on which the density profiles may have
-low number of inflection points, the computation time would be in the same order of
-magnitude as the one for the homogeneous density case.
+exponential and sinusoidal density functions and the resulting discretization points as
+orange dots.
+
+For the exponential density functions, the algorithm performs a single division on the
+inflection point independently of the value of $b$.
+Therefore, the computation time would likely be at least twice that of the homogeneous
+density case (assuming equal implementations).
+On the other hand, Fig.~\ref{fig:number-of-tesseroids}c shows an almost linear relation
+between the number of divisions in the sinusoidal density case and the value of $b$.
+Thus, the computation time is likely to be dependent on the number of wavelengths
+of the sinusoidal function that are contained within the tesseroid.
 
@@
 \caption{
-    Number of discretizations performed by the density-based algorithm in case of
-    (a)~exponential densities with the same values of $b$ shown in
+    Number of divisions performed by the density-based discretization algorithm
+    (with $\delta = 0.1$) in case of
+    (a)~exponential density functions with the same values of $b$ shown in
     Fig.~\ref{fig:exp-densities},
-    (b)~sinusoidal densities with the same values of $b$ shown in
+    (b)~sinusoidal density functions with the same values of $b$ shown in
     Fig.~\ref{fig:sine-densities}.
-    On both figures the boundaries of the discretized tesseroids are marked with orange
+    On both figures, the locations of the tesseroid divisions are are marked with orange
     dots.
-    Subfigure (c) shows the number of discretized tesseroids for the sinusoidal
-    densities as a function of the factor $b$.
+    (c) shows the number of discretized tesseroids for the sinusoidal density functions
+    as a function of $b$.
 }
@@
 ones tested here.
+
+The algorithm performance analysis shows that the computation time using variable
+densities is likely to be at least two times larger than the homogeneous density case.
+It is also likely to increase proportionally to the number of inflection points of the
+density function.
+As is the case for most numerical methods, there is a trade-off between computation time
+and accuracy.
+Nevertheless, density profiles will have few inflection points for most geophysical
+applications.
+Therefore, the computation time would be in the same order of
+magnitude as the one for the homogeneous density in most real world applications.
 

### 2018-12-29 manuscript/manuscript.tex
No inflection point for exponential

@@
 
-We analyse exponential density functions (Eq.~\ref{eq:density-exp}) with values of $b$ equal to
-1, 2, 5, 10, 30 and 100, and sinusoidal density functions (Eq.~\ref{eq:density-sine}) with
-values of $b$ equal to 1, 2, 5, and 10.
+We analyse exponential density functions (Eq.~\ref{eq:density-exp}) with values of $b$
+equal to 1, 2, 5, 10, 30 and 100, and sinusoidal density functions
+(Eq.~\ref{eq:density-sine}) with values of $b$ equal to 1, 2, 5, and 10.
 We then apply the density-based discretization algorithm on each function and record the
 number divisions carried out.
-Figs.~\ref{fig:number-of-tesseroids}(a) and~\ref{fig:number-of-tesseroids}(b) show the
+Figs.~\ref{fig:number-of-tesseroids}a and~\ref{fig:number-of-tesseroids}b show the
 exponential and sinusoidal density functions and the resulting discretization points as
@@
 
-For the exponential density functions, the algorithm performs a single division on the
-inflection point independently of the value of $b$.
+For the exponential density functions, the algorithm performs a single division
+independently of the value of $b$.
 Therefore, the computation time would likely be at least twice that of the homogeneous

### 2018-12-29 manuscript/manuscript.tex
Fix British spelling for modelling

@@
 
-An application to modeling the Neuqu\'en Basin, Argentina, demonstrates that the effects
-of sediment compaction cannot be ignored.
+An application to modelling the Neuqu\'en Basin, Argentina, demonstrates that the
+effects of sediment compaction cannot be ignored.
 When compared to an exponential density function, a homogeneous density with depth would
@@
 to -6 mGal error.
-Accurate and robust forward modeling is a key component to any gravity inversion method.
+Accurate and robust forward modelling is a key component to any gravity inversion
+method.
 Errors of this magnitude could result in significant overestimation of sedimentary

### 2019-02-12 manuscript/manuscript.tex
Add a version with differences highlighted
Used git latexdiff to create it and include here in case we don't have
it installed.
Fix wrong or missing punctuation on equations
Fix typos highlighted by reviewer

@@
 \end{equation}
-\\
+
 \noindent in which $\alpha \in \{x, y, z\}$,
@@
 }
-\\
+
 \noindent where
@@
 \end{equation}
-\\
+
 \noindent $f(r', \phi', \lambda')$ is an integral kernel for a homogeneous tesseroid
@@
     \cos\psi_t =
-        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t).
+        \sin\phi\sin\phi_t + \cos\phi\cos\phi_t\cos(\lambda - \lambda_t),
 \end{equation}
@@
     \phi_t = \frac{\phi_2 + \phi_1}{2}, \quad
-    \lambda_t = \frac{\lambda_2 + \lambda_1}{2}.
-\end{equation}
-\\
+    \lambda_t = \frac{\lambda_2 + \lambda_1}{2},
+\end{equation}
+
 \noindent
@@
 \end{equation}
-\\
+
 \noindent and
@@
     \rho_l(r') =
-    \frac{ \rho_n(r_2) - \rho_n(r_1) }{ r_2 - r_1 } (r' - r_1) + \rho_n(r_1),
+    \frac{ \rho_n(r_2) - \rho_n(r_1) }{ r_2 - r_1 } (r' - r_1) + \rho_n(r_1).
     \label{eq:density-reference-line}
@@
 
-On the first iteration, the ratio $L_r/L_r^\text{orig} = 1$ because the tesseroid being
-divided is the original one.
+On the first iteration, the ratio $L_r/L_r^\text{orig}$ equals one because the tesseroid
+being divided is the original one.
 For further iterations, the ratio will be progressively smaller than one as the
@@
 Through this analysis, we have obtained optimal values for the distance-size ratio $D$
-of 1 and 2.5 for the potential an its gradient, respectively.
+of 1 and 2.5 for the potential and its gradient, respectively.
 We analysed the error for exponential functions ranging from smooth to sharp and
@@
 sufficiently smooth, as is the case for most common geophysical applications.
-Smaller values of $\delta$ can used for more highly variable density functions to
+Smaller values of $\delta$ can be used for more highly variable density functions to
 increase the accuracy of the integration accordingly.
@@
 \begin{equation}
-    g_z(r) = \frac{V_\text{sh}(r)}{r},
+    g_z(r) = \frac{V_\text{sh}(r)}{r}.
 \label{eq:shell-gz}
@@
     & (2 - k^2 R_1^2) \cos(k(R_1 - R)) - 2 k R_1 \sin(k(R_1 - R))
-        \Big]
+        \Big].
     \end{split}

### 2019-02-12 code/scripts/neuquen_basin.py
Set same colormap for Neuquen basin diffs figures

@@ -379,6 +379,29 @@ units = {"potential": "J/kg", "gz": "mGal"}
 titles = {"potential": r"$V$", "gz": r"$g_{z}$"}
 fig_titles = "(a) (b) (c) (d)".split()
 
+# Compute differences between exponential density results with linear and homogeneous
+differences = dict(zip(fields, [{}, {}]))
+for field in fields:
+    exponential = np.load(
+        os.path.join(result_dir, "exponential-{}.npz".format(field))
+    )
+    shape = exponential["shape"]
+    longitude = exponential["lon"].reshape(shape)
+    latitude = exponential["lat"].reshape(shape)
+    for density in densities:
+        other = np.load(
+            os.path.join(result_dir, "{}-{}.npz".format(density, field))
+        )
+        difference = exponential["result"] - other["result"]
+        differences[field][density] = difference.reshape(shape)
+
+min_value = {}
+for field in fields:
+    min_value[field] = min(
+        [differences[field][density].min() for density in densities]
+    )
+
+# Plot result
 for i, density in enumerate(densities):
     for j, field in enumerate(fields):
         ax = axes[i, j]
@@ -386,22 +409,11 @@ for i, density in enumerate(densities):
         ax.set_title(fig_titles[2*i + j], y=1.08, loc='left')
         ax.set_title(titles[field], y=1.08, loc='center')
 
-        # Plot result
-        exponential = np.load(
-            os.path.join(result_dir, "exponential-{}.npz".format(field))
-        )
-        other = np.load(
-            os.path.join(result_dir, "{}-{}.npz".format(density, field))
-        )
-        difference = exponential["result"] - other["result"]
-        vmax = np.abs(difference).max()
-
-        shape = exponential["shape"]
-        im = bm.pcolormesh(exponential["lon"].reshape(shape),
-                           exponential["lat"].reshape(shape),
-                           difference.reshape(shape),
-                           vmin=-vmax, vmax=vmax,
-                           cmap="RdBu_r", rasterized=True, latlon=True)
+        vmin = min_value[field]
+        # Because max value of differences is bellow zero, we set vmax=0
+        im = bm.pcolormesh(longitude, latitude, differences[field][density],
+                           vmin=vmin, vmax=0,
+                           cmap="Blues_r", rasterized=True, latlon=True)
         bm.contour(exponential["lon"].reshape(shape),
                    exponential["lat"].reshape(shape),
                    difference.reshape(shape),

### 2019-03-08 manuscript/manuscript.tex
Replace "linear density" for "straight line"
On the density-based discretization algorithm we use a "straight line"
for computing the new discretization depth (if it's needed).
Naming this "straight line" as "linear density" could cause confusions
with the "linear density" used in the ratii determinations and may
induce readers to think that we are applying a piecewise linear
approximation of the density function.

@@
     (a)~The normalised density function $\rho_n(r')$ (blue), current boundaries of the
-    tesseroid (orange dots), and the linear density function $\rho_l(r')$ (orange line).
+    tesseroid (orange dots), and the straight line $\rho_l(r')$ (orange line).
     The dashed red line represents the maximum density difference $\Delta \rho (r')$ at
@@
     inequality~\ref{eq:delta-density} is not satisfied).
-    (b)~Second iteration of the algorithm with a new linear density function and maximum
+    (b)~Second iteration of the algorithm with a new straight line and maximum
     density difference. The tesseroid would be divided at the depth indicated by the
@@
 \textit{Step 1}:
-Define a linear function $\rho_l(r')$ that assumes the same values as the normalised
+Define a straight line $\rho_l(r')$ that assumes the same values as the normalised
 density function $\rho_n(r')$ at the boundaries of the tesseroid ($r_1$ and $r_2$):
@@
 \textit{Step 2}:
-Evaluate the normalized and linear density functions on a range of $N$ radii between
-$r_1$ and $r_2$.
+Evaluate the normalized density function and straight line on a range of $N$ radii
+between $r_1$ and $r_2$.
 We have opted for $N = 101$ but the specific value of $N$ is not critical to the
@@
 \textit{Step 3}:
-Compute the absolute difference between the values of the linear and normalised density
-functions:
+Compute the absolute difference between the values of the straight line and the
+normalised density function:
 

### 2019-03-08 manuscript/manuscript.tex
Replace "normalize" for "normalize"

@@
 threshold is exceeded,
-namely when the normalized distance to the computation point is greater than a
+namely when the normalised distance to the computation point is greater than a
 ``distance-size ratio'' parameter ($D$).
@@
 
-We have developed a complementary discretization algorithm in the redial direction that
+We have developed a complementary discretization algorithm in the radial direction that
 takes into account the variations of the density function.
@@
 \textit{Step 2}:
-Evaluate the normalized density function and straight line on a range of $N$ radii
+Evaluate the normalised density function and straight line on a range of $N$ radii
 between $r_1$ and $r_2$.

### 2019-03-09 manuscript/manuscript.tex
Add missing the before straight line
Add PWL sentences to manuscript

@@
 functions, the choice of a discretization interval is neither straight forward nor
-automatic.
+automatic for the general case.
+Although many piecewise linear approximation algorithms that automatize this process
+exist \citep{Ketkov1969, Vandewalle1975, Imamoto2008, Ahmadi2013}, they either require
+a fixed number of discretization intervals or are meant to be used only on convex
+functions.
+They would limit the range of density functions that can be assigned to the tesseroids
+or wouldn't guarantee a fully automatic process.
 Furthermore, it is well known that the use of high-degree polynomials to approximate a
@@
 \textit{Step 2}:
-Evaluate the normalised density function and straight line on a range of $N$ radii
+Evaluate the normalised density function and the straight line on a range of $N$ radii
 between $r_1$ and $r_2$.

### 2019-03-09 manuscript/manuscript.tex
Add sinusoidal disclaimer on abstract

--- 2019-03-08 manuscript/manuscript.tex
@@
 exponential, and sinusoidal density functions.
+The heavily oscillating density functions are intended only to test the algorithm to its
+limits and not to emulate a real world case.
 These comparisons allowed us to obtain optimal values for the distance-size and

### 2019-03-09 manuscript/manuscript.tex
Add more sinusoidal disclaimer in the abstract

@@
 The density-based discretization algorithm produces no discretizations in the linear
-density case, but a delta ratio of 0.1 is needed for the exponential and the
-sinusoidal density functions.
-These values can be extrapolated to cover most common use cases.
+density case, but a delta ratio of 0.1 is needed for the exponential and most sinusoidal
+density functions.
+These values can be extrapolated to cover most common use cases, which are simpler than
+oscillating density profiles.
 However, the distance-size and delta ratios can be configured by the user to increase

### 2019-03-09 manuscript/manuscript.tex
Add more sinusoidal disclaimer in text
Add more sinusoidal disclaimer

@@
 functions.
+The sinusoidal density case is included in order to test the numerical approximation to
+its limits.
 The comparisons are repeated for all combinations of tesseroid models in
@@
 complex situations than linear and exponential functions.
+Furthermore, acquiring an optimal value for the $\delta$ ratio in case of unrealistic
+density functions allow to extrapolate it to simpler realistic cases.
+Thus this tests are only intended to submit the algorithm to the extreme and not to
+emulate a real case scenario.
 

### 2019-03-09 manuscript/manuscript.tex
Add sinusoidal disclaimer in discussion
Move disclaimer from discussion to conclusions
Change Appendix A: shell solutions from Newton

--- 2019-03-09 manuscript/manuscript.tex
@@
 exponential, and sinusoidal density functions.
-The heavily oscillating density functions are intended only to test the algorithm to its
-limits and not to emulate a real world case.
 These comparisons allowed us to obtain optimal values for the distance-size and
@@
 a spherical shell with linear density in the radial coordinate.
-We derive an expression for the gravitational potential of a spherical shell with
-an arbitrary density function in depth (Eq.~\ref{eq:shell-pot}),
-from which we obtained
-expressions for linear, exponential, and sinusoidal density functions
+Applying the Newton's Shell Theorem \citep{Chandrasekhar1995, Binney2008},
+we derive expressions for the gravitational potential of a spherical shell with
+linear, exponential, and sinusoidal density functions
 (see Appendix~\ref{sec:shell}).
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.5\linewidth]{figures/spherical-shell.pdf}
-\caption{
-    Spherical shell with inner and outer radii $R_1$ and $R_2$, respectively.
-    The computation point $Q$ is located on the $z$ axis at a distance $r$ from
-    the origin of the coordinates system.
-    For our purposes we will assume that $Q$ is outside of the shell,
-    i.e.~$r > R_2$.
-}
-\label{fig:spherical-shell}
-\end{figure}
-
 Consider a spherical shell with inner radius $R_1$ and outer radius $R_2$,
-whose density is a function $\rho(r')$ of the radial coordinate
-(Fig.~\ref{fig:spherical-shell}).
-The gravitational potential generated by the shell on an arbitrary external
-point $Q$ can be written as follows:
-
-\begin{equation}
-    V_\text{sh}(\phi, \lambda, r) = G
-    \int\limits_0^{2\pi}
-    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
-    \int\limits_{R_1}^{R_2}
-    \frac{\rho(r')}{\ell} {r'}^2 \cos\phi' \,
-    dr' d\phi' d\lambda',
-\end{equation}
-
-\noindent where $\ell$ is defined in equation~\ref{eq:ell}.
-
-For $Q$ located along the $z$ axis (i.e., $\phi=90^\circ$) at a distance $r$ from the
-origin, equation~\ref{eq:ell} simplifies to:
-
-\begin{equation}
-    \ell = \sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}.
-\end{equation}
-
-\noindent
-Because of the rotational symmetry along the $z$ axis, the integration in $\lambda'$ is
-straightforward:
-
-\begin{equation}
-    V_\text{sh}(r) = 2\pi G
-    \int\limits_{-\frac{\pi}{2}}^\frac{\pi}{2}
-    \int\limits_{R_1}^{R_2}
-    \frac{\rho(r') {r'}^2 \cos\phi'}{\sqrt{r'^2 + r^2 - 2 r r' \sin\phi'}}
-    \, dr' d\phi',
-\end{equation}
-
-\noindent
-while the integration in $\phi'$ can be performed independently of the density function.
-Making use of SymPy \citep{sympy2017}, a Python library for symbolic mathematics, we
-obtained the following expression for the potential:
-
-\iftwocol{
-\begin{equation}
-    \begin{split}
-        V_\text{sh}(r) = 2\pi G
-        \int\limits_{R_1}^{R_2}
-        \Big[ & \sqrt{r^2 + r'^2 + 2rr'} - \\
-        & \sqrt{r^2 + r'^2 - 2rr'}
-        \Big] \frac{r'\rho(r')}{r} \, dr'.
-    \end{split}
-\label{eq:shell-pot-sqrts}
-\end{equation}
-}{
-\begin{equation}
-    V_\text{sh}(r) = 2\pi G
-    \int\limits_{R_1}^{R_2}
-    \Big[ \sqrt{r^2 + r'^2 + 2rr'}  -
-    \sqrt{r^2 + r'^2 - 2rr'}
-    \Big] \frac{r'\rho(r')}{r} \, dr'.
-\label{eq:shell-pot-sqrts}
-\end{equation}
-}
-
-Because the computation point $Q$ is outside of the shell, $r > r'$ and the square roots
-in equation~\ref{eq:shell-pot-sqrts} simplify to
-
-\begin{equation}
-    \sqrt{r^2 + r'^2 + 2rr'} = |r + r'| = r + r',
-\end{equation}
-
-\begin{equation}
-    \sqrt{r^2 + r'^2 - 2rr'} = |r - r'| = r - r',
-\end{equation}
-
-\noindent which leads to the following expression for the potential:
+whose density is a function $\rho(r')$ of the radial coordinate.
+We want to get an analytical expression for the gravitational fields the shell generates
+on any external point located at a distance $r$ from the centre of the shell
+($r > R_2$).
+
+According to Newton's Shell Theorem (Theorem XXXI)
+\citep{Chandrasekhar1995, Binney2008}, the gravitational potential generated by the
+shell on any external point is the same as it would be if its mass were concentrated on
+a point located at its centre:
+
+\begin{equation}
+    V_\text{sh}(\phi, \lambda, r) = \frac{GM}{r},
+\end{equation}
+
+\noindent where $M$ is the total mass of the shell, which can be easily computed as
+follows:
+
+\begin{equation}
+    M =
+    \iiint\limits_{\Omega} \rho(r') dV =
+    4\pi \int\limits_{R_1}^{R_2} \rho(r') {r'}^2 dr',
+\end{equation}
+
+\noindent where $\Omega$ symbolizes the volume of the shell.
+
+Combining the two previous equations, we get the following expression for the potential:
 
@@
     V_\text{sh}(r) = \frac{4\pi G}{r}
-    \int\limits_{R_1}^{R_2} {r'}^2 \rho(r') \, dr'.
+    \int\limits_{R_1}^{R_2} {r'}^2 \rho(r') \, dr',
 \label{eq:shell-pot}
@@
 
-The gradient of potentials that
-depends solely on $r$ have only one non zero components: the vertical
-component of the gradient ($g_z$).
+\noindent which is in agreement with the one obtained by \citet[p.62]{Binney2008}.
+
+The gradient of potentials that depends solely on $r$ have only one non zero components:
+the vertical component of the gradient ($g_z$).
 Following \citet{Grombein2013}:
@@
 different density functions.
+The integration of the following density functions have been carried out by SymPy
+\citep{sympy2017}, a Python library for symbolic mathematics.
 

### 2019-04-04 manuscript/manuscript.tex
Specify horizontal dims in adaptive discretization (#56)
Fixes #55

--- 2019-03-09 manuscript/manuscript.tex
@@
 The density-based discretization algorithm produces no discretizations in the linear
-density case, but a delta ratio of 0.1 is needed for the exponential and most sinusoidal
-density functions.
-These values can be extrapolated to cover most common use cases, which are simpler than
-oscillating density profiles.
+density case, but a delta ratio of 0.1 is needed for the exponential and the
+sinusoidal density functions.
+These values can be extrapolated to cover most common use cases.
 However, the distance-size and delta ratios can be configured by the user to increase
@@
 
-\textit{Step 1}: Check that the tesseroid satisfies the following inequality for each
-dimension $L_i\ (i \in \{\lambda, \phi\})$ of the tesseroid:
+\textit{Step 1}: Check that the tesseroid satisfies the following inequality for its
+latitudinal and longitudinal dimensions $L_i\ (i \in \{\lambda, \phi\})$:
 

### 2019-04-04 manuscript/manuscript.tex
Add sinusoidal density function disclaimers (#53)
Fixes #46

@@
 discretization.
+It worth noting that the sinusoidal density tests were carried out in order to test the
+algorithm to the extreme and not to emulate a real world scenario.
 Values of $\delta$ equal to 0.1 are sufficient to guarantee a 0.1\% accuracy for most

### 2019-04-04 manuscript/manuscript.tex
Minor typos

--- 2019-04-04 manuscript/manuscript.tex
@@
 functions, the choice of a discretization interval is neither straight forward nor
-automatic.
+automatic for the general case.
+Although many piecewise linear approximation algorithms that automate this process
+exist \citep{Ketkov1969, Vandewalle1975, Imamoto2008, Ahmadi2013}, they either require
+a fixed number of discretization intervals or are meant to be used only on convex
+functions.
+Using such algorithms would limit the range of density functions that can be assigned 
+to the tesseroids and would not guarantee a fully automatic process.
 Furthermore, it is well known that the use of high-degree polynomials to approximate a
@@
 
-\textit{Step 1}: Check that the tesseroid satisfies the following inequality for its
-latitudinal and longitudinal dimensions $L_i\ (i \in \{\lambda, \phi\})$:
+\textit{Step 1}: Check that the tesseroid satisfies the following inequality for each
+dimension $L_i\ (i \in \{\lambda, \phi\})$ of the tesseroid:
 

### 2019-06-06 manuscript/manuscript.tex
Format a version for the EarthArxiv
Use the Latex article class instead of GJI and insert the self-archiving
text from the publisher.

@@
-%\documentclass[extra]{gji}
-\documentclass[extra, referee]{gji}
-
+\documentclass[twocolumn]{article}
+
+\newcommand{\Title}{
+    Gravitational field calculation in spherical coordinates using variable
+    densities in depth
+}
+\newcommand{\Author}{S.R. Soler, A. Pesce, M.E. Gimenez, L. Uieda}
+\newcommand{\AuthorAffil}{
+    {\large
+        Santiago R. Soler$^{1,2,*}$, Agustina Pesce$^{1,2}$,
+        Mario E. Gimenez$^{1,2}$, and Leonardo Uieda$^3$
+    }
+    \\[0.4cm]
+    {\small $^1$ CONICET, Argentina}
+    \\
+    {\small $^2$ Instituto Geofísico Sismológico Volponi, Universidad Nacional de San Juan, Argentina}
+    \\
+    {\small $^3$ Department of Earth Sciences, SOEST, University of Hawai'i at M\={a}noa, USA}
+    \\
+    {\small $^*$ e-mail: santiago.r.soler@gmail.com}
+}
+
+\usepackage[left=0.7in,right=0.7in,top=1in,bottom=1in]{geometry}
+\setlength{\columnsep}{2\columnsep}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
-
+\usepackage[round]{natbib}
+\usepackage{fixltx2e}
 \usepackage{url}
@@
     allcolors=blue,
-}
+    pdftitle={\Title},
+    pdfauthor={\Author},
+}
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{
+    \fontsize{9pt}{12pt}\selectfont
+    \Author (2019).
+    Accepted for publication in \textit{Geophysical Journal International}.
+}
+\rhead{\fontsize{9pt}{12pt}\selectfont \thepage}
+\renewcommand{\headrulewidth}{0pt}
 
@@
 
-\title[Variable Density Tesseroids]{
-    Gravitational field calculation in spherical coordinates using variable
-    densities in depth
-}
-\author[S.R. Soler, A. Pesce, M.E. Gimenez, and L. Uieda]{
-    Santiago R. Soler$^{1,2}$, Agustina Pesce$^{1,2}$,
-    Mario E. Gimenez$^{1,2}$, and Leonardo Uieda$^3$ \\
-    $^1$CONICET, Argentina.~e-mail: santiago.r.soler@gmail.com\\
-    $^2$Instituto Geofísico Sismológico Volponi, Universidad Nacional de
-    San Juan, Argentina\\
-    $^3$Department of Earth Sciences, SOEST, University of Hawai‘i at
-    M\={a}noa, Honolulu, Hawaii, USA
-}
-
+\title{\Title}
+\author{\AuthorAffil}
+\date{
+    \normalsize
+    Accepted 2019 June 05. Received 2019 May 10; in original form 2018 December 29
+    \\[0.4cm]
+    This is a pre-copyedited, author-produced PDF of an article accepted for
+    publication in \textit{Geophysical Journal International} following peer review.
+    The version of record
+    ``\textit{INSERT COMPLETE CITATION INFORMATION HERE}''
+    is available online at: \href{https://doi.org/XXXXXXX}{doi.org/XXXXXX}
+}
 \maketitle
 
-\begin{summary}
+\begin{abstract}
 We present a new methodology to compute the gravitational fields generated by
@@
 Argentina with a maximum depth of over 5000~m, using an exponential density function.
-\end{summary}
-
-\begin{keywords}
+\\[0.5cm]
+\textbf{Keywords:}
 Numerical modelling, Numerical approximations and analysis, Gravity anomalies
 and Earth structure, Satellite gravity
-\end{keywords}
+\end{abstract}
 
@@
 \centering
-\includegraphics[width=0.6\linewidth]{figures/tesseroid-uieda.pdf}
+\includegraphics[width=\linewidth]{figures/tesseroid-uieda.pdf}
 \caption{
@@
 
-\iftwocol{
+%\iftwocol{
 \begin{equation}
@@
 \end{equation}
-}{
-\begin{equation}
-    \int\limits_{\lambda_1}^{\lambda_2}
-    \int\limits_{\phi_1}^{\phi_2}
-    \int\limits_{r_1}^{r_2}
-    \rho(r') f(r', \phi', \lambda')
-    dr' d\phi' d\lambda' \approx
-    A
-    \sum\limits_{i=1}^{N^r}
-    \sum\limits_{j=1}^{N^\phi}
-    \sum\limits_{k=1}^{N^\lambda}
-    W_i^r W_j^\phi W_k^\lambda \rho(r_i) f(r_i, \phi_j, \lambda_k),
-\label{eq:glq-var-dens}
-\end{equation}
-}
+%}{
+%\begin{equation}
+    %\int\limits_{\lambda_1}^{\lambda_2}
+    %\int\limits_{\phi_1}^{\phi_2}
+    %\int\limits_{r_1}^{r_2}
+    %\rho(r') f(r', \phi', \lambda')
+    %dr' d\phi' d\lambda' \approx
+    %A
+    %\sum\limits_{i=1}^{N^r}
+    %\sum\limits_{j=1}^{N^\phi}
+    %\sum\limits_{k=1}^{N^\lambda}
+    %W_i^r W_j^\phi W_k^\lambda \rho(r_i) f(r_i, \phi_j, \lambda_k),
+%\label{eq:glq-var-dens}
+%\end{equation}
+%}
 
@@
     dimensions, respectively.
+    \newline
 }
 \label{tab:shell-models}
+\centering
 \begin{tabular}{rccccc}
@@
 
-\begin{table}
+\begin{table*}
 \caption{
@@
     Grid height is defined above the mean Earth radius.
+    \newline
 }
 \label{tab:grids}
+\centering
 \begin{tabular}{lccc}
@@
 \end{tabular}
-\end{table}
+\end{table*}
 
@@
 \centering
-\iftwocol{
+%\iftwocol{
 \includegraphics[width=\linewidth]{figures/exponential-densities.pdf}
-}{
-\includegraphics[width=0.5\linewidth]{figures/exponential-densities.pdf}
-}
+%}{
+%\includegraphics[width=0.5\linewidth]{figures/exponential-densities.pdf}
+%}
 \caption{
@@
 \centering
-\iftwocol{
+%\iftwocol{
 \includegraphics[width=\linewidth]
     {figures/grid-search.pdf}
-}{
-\includegraphics[width=0.5\linewidth]
-    {figures/grid-search.pdf}
-}
+%}{
+%\includegraphics[width=0.5\linewidth]
+    %{figures/grid-search.pdf}
+%}
 \caption{
@@
 \centering
-\iftwocol{
+%\iftwocol{
 \includegraphics[width=\linewidth]{figures/sine-densities.pdf}
-}{
-\includegraphics[width=0.5\linewidth]{figures/sine-densities.pdf}
-}
+%}{
+%\includegraphics[width=0.5\linewidth]{figures/sine-densities.pdf}
+%}
 \caption{
@@
 \centering
-\iftwocol{
+%\iftwocol{
 \includegraphics[width=\linewidth]{figures/neuquen-basin-densities.pdf}
-}{
-\includegraphics[width=0.5\linewidth]{figures/neuquen-basin-densities.pdf}
-}
+%}{
+%\includegraphics[width=0.5\linewidth]{figures/neuquen-basin-densities.pdf}
+%}
 \caption{
@@
         V_\text{sine}(r) = \frac{4\pi G}{r} \frac{A}{k^3} \Big[
-    & (2 - k^2 R_2^2) \cos(k(R_2 - R)) + 2 k R_2 \sin(k(R_2 - R)) - \\
-    & (2 - k^2 R_1^2) \cos(k(R_1 - R)) - 2 k R_1 \sin(k(R_1 - R))
+    & (2 - k^2 R_2^2) \cos(k(R_2 - R)) + \\
+    & 2 k R_2 \sin(k(R_2 - R)) - \\
+    & (2 - k^2 R_1^2) \cos(k(R_1 - R)) - \\
+    & 2 k R_1 \sin(k(R_1 - R))
     \Big].

### 2019-06-06 manuscript/manuscript.tex
Add the SOEST number and thank the editor/reviewers (#62)
Fixes #60

--- 2019-04-04 manuscript/manuscript.tex
@@
 This new code is freely available under the BSD 3-clause open-source license.
-It can be freely downloaded from the online repository
+All source code, Python scripts, data, and model results are made available through
+an online repository
+\href{https://doi.org/10.6084/m9.figshare.8239622}{doi.org/10.6084/m9.figshare.8239622}
+\citep{Soler2019} or
 \href{https://github.com/pinga-lab/tesseroid-variable-density}{github.com/pinga-lab/tesseroid-variable-density}.
+The repository also contains instructions for replicating all results
+presented here.
 
@@
 which this work would not have been possible.
+We would also like to thank editor Kosuke Heki and two anonymous reviewers for their
+comments and contributions.
+This is SOEST contribution 10717.
 

### 2019-06-10 manuscript/manuscript.tex
Add publication DOI reserved by Oxford

@@
     ``\textit{INSERT COMPLETE CITATION INFORMATION HERE}''
-    is available online at: \href{https://doi.org/XXXXXXX}{doi.org/XXXXXX}
+    is available online at: \href{https://doi.org/10.1093/gji/ggz277}{doi.org/10.1093/gji/ggz277}
 }

### 2019-06-11 manuscript/manuscript.tex
Format a version for the EarthArxiv (#63)
Use the Latex article class instead of GJI and insert the self-archiving
text from the publisher.
Add citation information for version of record and its DOI.
Add DOI link and citation to README.
Better formatting for citation in README.

--- 2019-06-06 manuscript/manuscript.tex
@@
 }
+\newcommand{\DOI}{doi:\href{https://doi.org/10.1093/gji/ggz277}{10.1093/gji/ggz277}}
+\newcommand{\DOILink}{\href{https://doi.org/10.1093/gji/ggz277}{doi.org/10.1093/gji/ggz277}}
 
@@
     \fontsize{9pt}{12pt}\selectfont
-    \Author (2019).
-    Accepted for publication in \textit{Geophysical Journal International}.
+    \Author{}, 2019. \DOI{}
 }
@@
     The version of record
-    ``\textit{INSERT COMPLETE CITATION INFORMATION HERE}''
-    is available online at: \href{https://doi.org/XXXXXXX}{doi.org/XXXXXX}
+    ``\textit{Soler, S. R., Pesce, A., Gimenez, M. E., \& Uieda, L., 2019.
+    \Title{}, Geophysical Journal International, \DOI{}}\ ''
+    is available online at: \DOILink{}
 }

### 2019-06-21 manuscript/manuscript.tex
Update reference to Lin2019

@@
 Quadrature (GLQ)
-\citep{Asgharzadeh2007, Wild-Pfeiffer2008, Li2011, Uieda2016, Lin2018}.
+\citep{Asgharzadeh2007, Wild-Pfeiffer2008, Li2011, Uieda2016, Lin2019}.
 The Taylor series expansion is not well suited to develop an algorithm for
@@
 polynomial density function of arbitrary degree.
-\citet{Lin2018} compared the different integration and discretization methodologies for
+\citet{Lin2019} compared the different integration and discretization methodologies for
 homogeneous tesseroids.
@@
 subdivision developed by \citet{Grombein2013}.
-\citet{Lin2018} also introduced a variation of their combined method to compute the
+\citet{Lin2019} also introduced a variation of their combined method to compute the
 gravitational fields generated by tesseroids with a linearly varying density in the
@@
 
-Both the \citet{Lin2018} and the \citet{Fukushima2018} studies limit the radial density
+Both the \citet{Lin2019} and the \citet{Fukushima2018} studies limit the radial density
 variation to polynomial functions.
@@
 It is based on the three dimensional GLQ integration, a two dimensional version of the
-adaptive discretization of \citet{Uieda2016} (following \citet{Lin2018}),
+adaptive discretization of \citet{Uieda2016} (following \citet{Lin2019}),
 and a new density-based radial discretization algorithm.
@@
 adaptive discretization algorithms.
-On the other hand, \citet{Lin2018} proposed a two dimensional discretization algorithm
+On the other hand, \citet{Lin2019} proposed a two dimensional discretization algorithm
 that subdivides the tesseroid only on the latitudinal and longitudinal directions.
@@
 reducing the number of tesseroids in the model, while
-retaining an acceptable accuracy \citep{Lin2018}.
-
-Here we will follow \citet{Lin2018} and use a two dimensional version of the adaptive
+retaining an acceptable accuracy \citep{Lin2019}.
+
+Here we will follow \citet{Lin2019} and use a two dimensional version of the adaptive
 discretization of \citet{Uieda2016}.
@@
 have the same density function of radius as our tesseroid model.
-\citet{Lin2018} show the analytical solution of the gravitational potential generated by
+\citet{Lin2019} show the analytical solution of the gravitational potential generated by
 a spherical shell with linear density in the radial coordinate.
@@
 determine an optimal discretization when approximating a density function by piecewise
-linear \citep{Lin2018} or piecewise polynomial \citep{Fukushima2018} functions.
+linear \citep{Lin2019} or piecewise polynomial \citep{Fukushima2018} functions.
 
@@
 density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
-Eq~\ref{eq:shell-pot-linear} is in agreement with the one obtained by \citet{Lin2018}.
+Eq~\ref{eq:shell-pot-linear} is in agreement with the one obtained by \citet{Lin2019}.
 

### 2019-06-27 code/tests/test_constant_density.py
Update code by removing old notebooks and scripts (#70)
Remove old notebooks, unused scripts.
Update README.md and Makefile.
Remove unimplemented tests

@@ -1,290 +1,0 @@
-from __future__ import division, absolute_import
-import numpy as np
-from numpy.testing import assert_array_almost_equal, assert_allclose
-from pytest import raises
-import multiprocessing
-import warnings
-from fatiando.mesher import Tesseroid, TesseroidMesh
-from fatiando import gridder
-from fatiando.constants import SI2MGAL, SI2EOTVOS, G, MEAN_EARTH_RADIUS
-
-from tesseroid_density import tesseroid
-
-
-def test_warn_if_division_makes_too_small():
-    "gravmag.tesseroid warn if not dividing further bc tesseroid got too small"
-    # When tesseroids get below a threshold, they should not divide further and
-    # compute as is instead. Otherwise results in ZeroDivisionError involving
-    # some trigonometric functions.
-    ds = 1e-6
-    models = [
-        [Tesseroid(-ds, ds, -ds, ds, 0, -1000, {'density': 100})],
-        [Tesseroid(-1e-3, 1e-3, -1e-3, 1e-3, 0, -1e-2, {'density': 100})]]
-    lat, lon = np.zeros((2, 1))
-    h = np.array([0.1])
-    warning_msg = (
-        "Stopped dividing a tesseroid because it's dimensions would be " +
-        "below the minimum numerical threshold (1e-6 degrees or 1e-3 m). " +
-        "Will compute without division. Cannot guarantee the accuracy of " +
-        "the solution.")
-    for i, model in enumerate(models):
-        with warnings.catch_warnings(record=True) as w:
-            # Cause all warnings to always be triggered.
-            warnings.simplefilter("always")
-            tesseroid.gz(lon, lat, h, model)
-            msg = ("Failed model {}. Got {} warnings.\n\n".format(i, len(w)) +
-                   "\n\n".join([str(j.message) for j in w]))
-            assert len(w) >= 1, msg
-            assert any(issubclass(j.category, RuntimeWarning) for j in w), \
-                "No RuntimeWarning found. " + msg
-            assert any(warning_msg in str(j.message) for j in w), \
-                "Message mismatch. " + msg
-
-
-def test_warn_if_too_small():
-    "gravmag.tesseroid warns if ignoring tesseroid that is too small"
[…]

### 2019-06-27 code/tests/test_linear_density.py

@@ -1,125 +1,0 @@
-from __future__ import division, absolute_import
-import numpy as np
-from fatiando.constants import G, MEAN_EARTH_RADIUS, SI2MGAL, SI2EOTVOS
-from fatiando.mesher import TesseroidMesh
-from fatiando import gridder
-
-from tesseroid_density import tesseroid
-
-
-def shell_linear_density(height, top, bottom, a, b):
-    """
-    Returns the analytical solution of the gravity fields generated by a
-    spherical shell with linear density function given by:
-
-    .. math::
-        \rho(r) = a*r + b
-
-    where :math:`r` is the radial coordinate, while :math:`a` and :math:`b`
-    are constants.
-
-    The solution corresponds with the value of the gravity fields in a
-    computation point at a height given by `height` above the Earth surface.
-    """
-    r = height + MEAN_EARTH_RADIUS
-    r1 = bottom + MEAN_EARTH_RADIUS
-    r2 = top + MEAN_EARTH_RADIUS
-    constant = np.pi*G*a*(r2**4 - r1**4) + 4/3.*np.pi*G*b*(r2**3 - r1**3)
-    potential = constant/r
-    data = {'potential': potential,
-            'gx': 0,
-            'gy': 0,
-            'gz': SI2MGAL*(potential/r),
-            'gxx': SI2EOTVOS*(-potential/r**2),
-            'gxy': 0,
-            'gxz': 0,
-            'gyy': SI2EOTVOS*(-potential/r**2),
-            'gyz': 0,
-            'gzz': SI2EOTVOS*(2*potential/r**2)}
-    return data
-
-
-def test_thin_shell():
-
-    def density_linear(height):
-        r = height + MEAN_EARTH_RADIUS
-        return a*r + b
-
-    # Spherical shell model with a Tesseroid Mesh
-    top, bottom = 0, -1000
-    density_1, density_2 = 2670, 3300
-    a = -(density_2 - density_1)/abs(bottom)
-    b = (density_2 - density_1)/abs(bottom)*MEAN_EARTH_RADIUS + 2670
-    model = TesseroidMesh((0, 360, -90, 90, top, bottom), (1, 6, 12))
-    model.addprop("density", [density_linear for i in range(model.size)])
-
-    # Computation grids
-    shape = (10, 10)
-    grids = {"pole": gridder.regular((89, 90, 0, 1), shape, z=2e3),
[…]
